### 2009-07-01 doc/user_guide.tex

\documentclass{article}
\usepackage{hyperlatex}

%\topmargin=-0.5in
\texonly{\oddsidemargin=0in}
\texonly{\evensidemargin=0in}
\texonly{\textwidth=5.6in}
%\textheight=8.5in

%TODO: specify use of BDD's under efficiency in section 1 (overview)

\htmlmathitalic{0}
\htmltitle{User Guide}
\htmladdress{chord@cs.stanford.edu, \today}

\providecommand\Mdom{{\bf M}}
\providecommand\Idom{{\bf I}}
\providecommand\Vdom{{\bf V}}
\providecommand\Zdom{{\bf Z}}
\providecommand\Hdom{{\bf H}}
\providecommand\Edom{{\bf E}}
\providecommand\Fdom{{\bf F}}
\providecommand\Cdom{{\bf C}}
\providecommand\Tdom{{\bf T}}
\providecommand\Pdom{{\bf P}}
\providecommand\Sdom{{\bf S}}

\providecommand\instM{{\bf instM}}
\providecommand\statM{{\bf statM}}
\providecommand\MH{{\bf MH}}
\providecommand\MI{{\bf MI}}
\providecommand\ME{{\bf ME}}

\providecommand\cha{{\bf cha}}
\providecommand\sub{{\bf sub}}
\providecommand\HT{{\bf HT}}
\providecommand\VT{{\bf VT}}
\providecommand\EF{{\bf EF}}
\providecommand\writeE{{\bf writeE}}

\providecommand\virtIM{{\bf virtIM}}
\providecommand\statIM{{\bf statIM}}
\providecommand\specIM{{\bf specIM}}

\providecommand\MmethArg{{\bf MmethArg}}
\providecommand\MmethRet{{\bf MmethRet}}
\providecommand\IinvkArg{{\bf IinvkArg}}
\providecommand\IinvkRet{{\bf IinvkRet}}
\providecommand\MordMethArg{{\bf MordMethArg}}
\providecommand\MspcMethArg{{\bf MspcMethArg}}

\providecommand\MobjVarAsgnInst{{\bf MobjVarAsgnInst}}
\providecommand\MclsVarAsgnInst{{\bf MclsVarAsgnInst}}
\providecommand\MobjValAsgnInst{{\bf MobjValAsgnInst}}
\providecommand\MgetInstFldInst{{\bf MgetInstFldInst}}
\providecommand\MputInstFldInst{{\bf MputInstFldInst}}
\providecommand\MgetStatFldInst{{\bf MgetStatFldInst}}
\providecommand\MputStatFldInst{{\bf MputStatFldInst}}

\providecommand\syncMV{{\bf syncMV}}
\providecommand\nestedLocks{{\bf nestedLocks}}

\providecommand\VH{{\bf VH}}
\providecommand\FH{{\bf FH}}
\providecommand\HFH{{\bf HFH}}
\providecommand\IM{{\bf IM}}

\providecommand\forkC{{\bf forkC}}

\providecommand\CVC{{\bf CVC}}
\providecommand\FC{{\bf FC}}
\providecommand\CFC{{\bf CFC}}
\providecommand\CICM{{\bf CICM}}
\providecommand\CHC{{\bf CHC}}

\providecommand\clinitM{{\bf clinitM}}

\providecommand\Chord{{Chord}}

\providecommand\NameToTrgtMap{{\tt NameToTrgtMap}}
\providecommand\NameToTaskMap{{\tt NameToTaskMap}}
\providecommand\ConsumedNameToTasksMap{{\tt ConsumedNameToTasksMap}}
\providecommand\TaskGraph{{\tt TaskGraph}}
\providecommand\TaskMapEntries{{\tt TaskMapEntries}}
\providecommand\TrgtMapEntries{{\tt TrgtMapEntries}}
\providecommand\DlogMapEntries{{\tt DlogMapEntries}}
\providecommand\Tasks{{\tt Tasks}}
\providecommand\Trgts{{\tt Trgts}}
\providecommand\ValidTasks{{\tt ValidTasks}}
\providecommand\TrgtNames{{\tt TrgtNames}}

\title{Chord: A Static Analysis Framework for Java}
\author{Mayur Naik and Alex Aiken}
\date{\today}
\begin{document}
\maketitle
\texonly{\tableofcontents}

\texonly{\newpage}
\section{Overview}

\Chord\ is a static analysis framework for Java with an emphasis on the analysis of objects and concurrency.
It has the following characteristics:
\begin{enumerate}
\item
It is {\it stand-alone} in that it provides various off-the-shelf static analyses (e.g., may alias analysis, thread escape analysis, datarace analysis, deadlock analysis, etc.).
\item
It is {\it extensible} in that it allows users to build their own static analyses on top of the provided static analyses.
\item
It is {\it compositional} in that it enables different static analyses to coexist and interact with each other.
\item
It is {\it efficient} in that it
caches results computed by each static analysis so that they can be reused by other static analyses without re-computation.
\item
It is {\it flexible} in that it allows each static analysis to be expressed either declaratively as a Datalog program or imperatively as a Java program.
\end{enumerate}

\texonly{\newpage}
\section{Download and Installation}

Before installing \Chord\ ensure that the following software is installed on your machine:

\begin{itemize}
\item
JDK 5 or higher (e.g., from
Sun\footnote{\xlink{{\tt http://java.sun.com/javase/downloads/index.jsp}}{http://java.sun.com/javase/downloads/index.jsp}}
or IBM\footnote{\xlink{{\tt http://www.ibm.com/developerworks/java/jdk/}}{http://www.ibm.com/developerworks/java/jdk/}}).

\item
Apache Ant (a Java build tool)\footnote{\xlink{{\tt http://ant.apache.org/}}{http://ant.apache.org/}}
\end{itemize}

Download the latest version of \Chord\ from \xlink{{\tt http://code.google.com/p/jchord/}}{http://code.google.com/p/jchord/}
(say {\tt chord-x.y.tar.gz}) and uncompress it in a suitable directory by running the following command:

\begin{verbatim}
tar xvzf chord-x.y.tar.gz
\end{verbatim}

This will create a directory structure with root directory {\tt chord-x.y/}.

\texonly{\newpage}
\section{Building Chord}

The root directory {\tt chord-x.y/} contains a {\tt build.xml} file which is interpreted by
Apache Ant (a Java build tool).
To see the various targets available run the following command in the root directory:

\begin{verbatim}
ant -p
\end{verbatim}

The default target {\tt compile} compiles \Chord.  So simply run the following command in the
root directory:

\begin{verbatim}
ant
\end{verbatim}

This will compile the \Chord\ {\tt .java} source files and store the {\tt .class} files
under the {\tt classes/} sub-directory of the root directory.

\texonly{\newpage}
\section{Running Chord}

Suppose the program to be analyzed has the following directory structure:

\begin{verbatim}
example/
    src/
        foo/
            Main.java
            ...
    classes/
        foo/
            Main.class
            ...
    lib/
        src/
            taz/
                ...
        jars/
            taz.jar
    chord/
        chord.properties
\end{verbatim}

The above structure is typical: the program's Java source files are under {\tt example/src/}, its class files
are under {\tt example/classes/}, the source and jar files of the libraries used by the program are under
{\tt example/lib/src/} and {\tt example/lib/jars/}, respectively, and finally {\tt example/chord/} is the directory
to which Chord will output its intermediate and final results.

Execute Chord on the above program by running the
following command in Chord's root directory {\tt chord-x.y/}:

\begin{verbatim}
ant -Dchord.work.dir=.../example/chord run
\end{verbatim}

where ``{\tt ...}'' denotes either the absolute path, or the path relative to Chord's root directory,
of the parent of directory {\tt example/}.
Chord will read the system properties specified in the file named {\tt chord.properties}
that it presumes exists in the directory specified
by system property {\tt chord.work.dir}
which in the above case is directory {\tt example/chord/}.
A sample such file for the above program is as follows:

\begin{verbatim}
chord.main.class=foo.Main
chord.class.path=../classes:../lib/classes/taz.jar
chord.src.path=../src:../lib/src
chord.out=log.txt
chord.err=log.txt
chord.serial.file=program.ser
chord.java.analysis.path=${chord.home.dir}/classes/main:
    ${chord.home.dir}/classes/downcast:${chord.home.dir}/src/dlog/monosite
chord.dlog.analysis.path=${chord.home.dir}/src/dlog/main:
    ${chord.home.dir}/src/dlog/downcast:${chord.home.dir}/src/dlog/monosite
chord.analyses=downcast-java,monosite-java
\end{verbatim}

System property {\tt chord.home.dir} specifies Chord's root directory
{\tt chord-x.y/}.
Every non-absolute location used in the value of any system property defined in the file
is taken to be relative to the directory specified by system
property {\tt chord.work.dir}.

Section \ref{sec-sys-prop} presents an exhaustive
listing of system properties that are recognized by \Chord.
Briefly:
\begin{itemize}
\item
{\tt chord.main.class} specifies the main class of the program
to be analyzed.
\item
{\tt chord.class.path} specifies the application classpath of the program to be analyzed (as opposed to the
JDK library classpath which consists of jar files such as {\tt rt.jar}
which are implicitly appended to the program's application classpath to yield the
complete program classpath).
\item
{\tt chord.out} and {\tt chord.err} specify the files (typically one and the same) to which the standard output and
standard error streams, respectively, must be redirected while running Chord.
\item
{\tt chord.serial.file} specifies the location of a file from (resp. to) which
the serialized form of the intermediate representation of the program to be analyzed must be loaded (resp. stored).
This option is provided because loading the serialized form is much faster than
constructing the intermediate representation of the program each time.
\item
{\tt chord.java.analysis.path} and {\tt chord.dlog.analysis.path} 
specify all program domains, program relations, and program analyses
to be included in the project and the
dependencies between them (see Section \ref{sec-tg}).
In the above case, they specify certain
 program domains, program relations, and program analyses
pre-defined in \Chord, in particular, all those in (1) Chord's standard library, (2) Chord's downcast safety
analysis, and (3) Chord's monomorphic call site analysis.
Users can also define their own
program domains, program relations, and program analyses and include them in the project.
\item
{\tt chord.analyses} is a space- or comma-separated list of the names of program domains,
program relations, and program analyses that must be run in that order.  In the above case,
these are downcast safety analysis, denoted {\tt downcast-java}, and monomorphic call site analysis,
denoted {\tt monosite-java}.
Both are program analyses pre-defined in \Chord\ 
but users can also include the names
of program domains, program relations, and program analyses they have defined and wish to run.
\end{itemize}

\texonly{\newpage}
\section{Specifying Program Scope}

\Chord\ takes as input a complete (i.e., closed) program with a main method.
The scope of the program must be specified in one of the
following ways:

\begin{itemize}
\item
Defining system property {\tt chord.serial.file} as
the location of a file containing in serialized form the intermediate representation
of the program generated by a previous run of \Chord.
\item
Defining system property {\tt chord.main.class} as the fully-qualified name
of the program's main class (e.g., {\tt foo.Main}) and system property
{\tt chord.class.path} as the program's application classpath.
\end{itemize}

The former option is provided because loading the serialized form
is much faster than constructing the intermediate representation
of the program each time.

\Chord\ first checks whether system property {\tt chord.serial.file} is defined.
If it is defined, then it further checks whether the file specified by it
exists.
If the file exists, then \Chord\ treats the file as a serialized form of
the intermediate representation of the program to be analyzed, and loads it.

If system property {\tt chord.serial.file} is not defined or if the file
specified by the property does not exist, then \Chord\ requires system properties
{\tt chord.main.class} and {\tt chord.class.path} 
to be defined, and uses them to
build an intermediate representation of the program (see Section \ref{sec-prog-build})
and perform source-to-source transformations on that representation
(see Section \ref{sec-prog-transform}) that simplify several program analyses.
Finally, if system property {\tt chord.serial.file} was defined but
specified a non-existent file, then \Chord\ creates the file and stores the
transformed intermediate representation of the program into the file in
serialized form so that it can be loaded in subsequent runs.

Defining system property {\tt chord.src.path} is optional.
It specifies the program's Java source path which provides the
locations of the program's {\tt .java} files so that, for instance, the
results of an analysis (which operates on the intermediate representation of
the program extracted from {\tt .class} files) can be mapped to the
program's {\tt .java} files and presented to the user for
manual inspection.

\texonly{\newpage}
\section{Program Representation}
\label{sec-prog-build}

Given the program's program's main class via system property {\tt chord.main.class} and the
program's application classpath via system property {\tt chord.class.path},
\Chord\ uses system property {\tt chord.program.builder.class} to locate the
builder to be used for building an intermediate representation of the program,
namely, an object of class {\tt chord.program.Program}.
By default, the value of this property is class {\tt chord.program.builders.SootProgramBuilder}
which uses the Soot Java compiler framework \cite{Soot}.
Any builder, however, must implement interface
{\tt chord.program.builders.IProgramBuilder}.

The intermediate representation
elides values of primitive type (e.g., int, boolean, etc.) and operations
on them (e.g., arithmetic and relational operations) in the original program.
The current emphasis of \Chord\ is on the analysis of objects
but the analysis of primitives may be supported in future
releases.

The intermediate representation represents the program as a set of types.
Each type has a set of (instance or static) fields and a set of
(instance or static) methods.
Each method has an optional control-flow graph.
The control-flow graph has a unique entry point and
a unique exit point\footnote{Even the control-flow graph of a method with an infinite loop
has an exit point because loops are transformed into tail-recursive methods
(see Section \ref{sec-prog-transform}).}.
The entry and exit points are always distinct.
Each point in the control-flow graph has one or more immediate predecessors
(except the entry point which has none) and, likewise,
one or more immediate successors (except the exit point which has none).
Associated with each point is a simple statement, which may be of one of
the following kinds:

\begin{itemize}
\item
{\tt HeadInst}

The statement associated with the unique entry point of a control-flow graph.

It serves as a placeholder for:

\begin{itemize}
\item
the (possibly empty) list of the method's {\it argument variables} of reference type;
in the case of an instance method, this list is non-empty,
with the $0^{th}$ element being the distinguished implicit {\tt this}
argument variable.
\item
the (possibly empty) list of the method's {\it temporary variables} of reference type.
\end{itemize}

Argument variables and temporary variables of primitive type are
elided due to \Chord's emphasis on the analysis of objects.

Argument variables and temporary variables are distinct, namely,
the set of a method's {\it local variables} (of reference type)
is the disjoint union of the sets of the method's argument variables (of reference type) and
temporary variables (of reference type).

\item
{\tt TailInst}

The statement associated with the unique exit point of a control-flow graph.

It serves as a placeholder for 
the (possibly empty) list of the method's {\it return variables} of reference type.

Return variables of primitive type are
elided due to \Chord's emphasis on the analysis of objects.

The set of a method's return variables (of reference type) is a subset of the set
of the method's local variables (of reference type).

A Java method has at most one return variable but
\Chord's intermediate representation allows multiple return variables
to enable compiling away exceptions, loops, synchronized blocks, etc.

\item
{\tt AcqLockInst}

A statement of the form {\tt monitorenter v} where
{\tt v} is a local variable of reference type.

\item
{\tt RelLockInst}

A statement of the form {\tt monitorexit v} where {\tt v}
is a local variable of reference type.

\item
{\tt StrValAsgnInst}

A statement of the form {\tt v = "..."} where {\tt v} is
a local variable of any reference type that can store a string
constant.  It is typically (but not necessarily) of type
{\tt java.lang.String}.

\item
{\tt NilValAsgnInst}

A statement of the form {\tt v = null} where {\tt v}
is a local variable of reference type.

\item
{\tt StatFldRefInst}

A statement of one of the following forms:

{\tt v = f} \\
{\tt f = v} \\
{\tt * = f} \\
{\tt f = *}

where {\tt v} is a local variable of reference type and
{\tt f} is a static field.

The {\tt *} denotes an expression of primitive type in the
original program which is elided due to \Chord's emphasis on the
analysis of objects.

\item
{\tt InstFldRefInst}

A statement of one of the following forms:

{\tt v = b.f} \\
{\tt b.f = v} \\
{\tt * = b.f} \\
{\tt b.f = *}

where {\tt v} and {\tt b} are local variables of reference
type and {\tt f} is an instance field.

The {\tt *} denotes an expression of primitive type in the
original program which is elided due to \Chord's emphasis on the
analysis of objects.

\item
{\tt SkipInst}

A no-op statement.

Due to \Chord's emphasis on the analysis of objects, each statement
in the original program that does not access
any value of reference type (i.e., only accesses values of primitive
type) is represented by this statement.

\item
{\tt InvkInst}

A method invocation statement of the form
{\tt r1,...,rn = m(a1,...,ak)} where
{\tt r1}, ..., {\tt rn} (return variables) and
{\tt a1}, ..., {\tt ak} (argument variables)
are local variables of reference type.

Argument variables and return variables of primitive type are
elided due to \Chord's emphasis on the analysis of objects.

A method invocation statement in Java has at most
one return variable but
\Chord's intermediate representation allows multiple return variables
to enable compiling away exceptions, loops, synchronized blocks, etc.

\item
{\tt ObjVarAsgnInst}

A statement of the form {\tt v1 = v2} where {\tt v1}
and {\tt v2} are local variables of (possibly different) reference types.

\item
{\tt ObjValAsgnInst}

A statement of the form {\tt v = new t} where {\tt t} is a
concrete reference type and
{\tt v} is a local variable of a (possibly different) reference type.

\item
{\tt ClsVarAsgnInst}

A statement of the form {\tt v = b.getClass()} where
{\tt v} is a local variable of type {\tt java.lang.Class}
and {\tt b} is a local variable of reference type.

The reason this kind of statement is modeled is because
\Chord\ has an emphasis on concurrency and many Java programs
hold locks on the runtime class of an object
(which is obtained by calling the native {\tt getClass}
instance method of the object).

\item
{\tt AryElemRefInst}

A statement of one of the following forms:

{\tt v = b[*]}  \\
{\tt b[*] = v} \\
{\tt * = b[*]} \\
{\tt b[*] = *}

where {\tt v} and {\tt b} are local variables of reference
type.

The {\tt *} denotes an expression of primitive type in the
original program which is elided due to \Chord's emphasis on the
analysis of objects.
\end{itemize}

\texonly{\newpage}
\section{Program Transformations}
\label{sec-prog-transform}

After building the intermediate representation of the
program, \Chord\ performs
the following source-to-source transformations 
that simplify several program analyses:

\begin{enumerate}
\item
{\bf Sync Removal}

Removes all synchronized blocks from each method M in each class
C in the program by generating a fresh static method (in class C)
per synchronized block in the body of M.

\item
{\bf Loop Removal}

Removes all loops from each method M in each class C in the program
by generating a fresh static method (in class C)
per loop in the body of M.

\item
{\bf Skip Removal}

Removes all skip (i.e., no-op) statements from each method in each class
in the program.
Such statements arise primarily because \Chord, due to its emphasis
on the analysis of objects, replaces each
statement in the original program that does not access any
value of reference type (i.e., only accesses values of primitive
type) by a skip statement.

\end{enumerate}

\texonly{\newpage}
\section{Task Graph Construction}
\label{sec-tg}

\Chord\ requires the following three system properties to construct the task graph:

\begin{enumerate}
\item
{\tt chord.dlogmap.path}

The path specifying the mapping from the name of each Datalog-based program
analysis that must be included in the project to the location
of the file defining that program analysis.

Each path element must be the location of a file containing a
list of entries (one per line) of the form:

{\tt <name> <file-name>}

The path elements must be separated by the system-specific path separator
character (e.g., `:' on UNIX and `;' on Windows).

Let \DlogMapEntries\ denote the set of all entries of the
above form.

\item
{\tt chord.trgtmap.path}

The path partially specifying the mapping from the name of each program domain,
program relation, and Java-based program analysis that must be
included in the project to the fully-qualified name of a subclass, either
pre-defined in \Chord\ or defined by the user, of {\tt chord.Dom},
{\tt chord.Ary*Rel}, and
{\tt chord.Entity}, respectively, the instance fields of which define the
state of the corresponding domain, relation, or analysis.
The `{\tt *}' in {\tt chord.Ary*Rel} is 1, 2, ..., 6, or N for
program relations of arity 1, 2, ..., 6, or $>$ 6, respectively.

Each path element must be the location of a file containing a
list of entries (one per line) of the form:

{\tt <name> <class-name>}

The path elements must be separated by the system-specific path separator
character (e.g., `:` on UNIX and ';' on Windows).

Let \TaskMapEntries\ denote the set of all entries of the
above form.

\item
{\tt chord.taskmap.path}

The path partially specifying the mapping from the name of each program domain,
program relation, and Java-based program analysis that must be included in
the project to the fully-qualified name of a
subclass, either pre-defined in \Chord\ or defined by the user, of
{\tt chord.Dom}, {\tt chord.Ary*Rel}, and {\tt chord.Entity}, respectively,
which define the following instance methods:
\begin{itemize}
\item
the {\tt execute} method specifying
the computation that updates the state of the corresponding domain, relation,
or analysis,
\item
the {\tt invalidate} method specifying the action to be taken when the
state of the corresponding domain, relation, or analysis is invalidated.
\item
the {\tt getConsumedNames} method specifying names of any domains, relations,
and Datalog-based or Java-based program analyses upon whose state the
computation depends, and
\item
the {\tt getProducedNames} method specifying names of any domains, relations,
and Datalog-based or Java-based program analyses whose state depends upon
the computation.
\end{itemize}

Each path element must be the location of a file containing a
list of entries (one per line) of the form:

{\tt <name> <class-name>}

The path elements must be separated by the system-specific path separator
character (e.g., `:` on UNIX and ';' on Windows).

Let \TrgtMapEntries\ denote the set of all entries of the above form.
\end{enumerate}

Then, \Chord\ uses the values of \DlogMapEntries, \TaskMapEntries, and \TrgtMapEntries\ 
to do the following:

\begin{enumerate}
\item
Compute {\tt Tasks\ :\ Set<chord.Entity>}, the set of all tasks,
each of which is an object of (a subclass of) {\tt chord.Entity},
as follows:

\begin{enumerate}
\item
For each entry in \DlogMapEntries\ (i.e., a Datalog-based program analysis,
there is a unique task, namely, a unique object of class {\tt chord.DatalogAnalysis}.

\item
For each entry in \TaskMapEntries\ (i.e., of the form {\tt <name> <class-name>} specifying
a program domain, a program relation, or a Java-based program analysis), there is a
unique task, namely, a unique object
of the class named {\tt <class-name>} which must be (a subclass of) {\tt chord.Dom} in the
case of a program domain, (a subclass of) {\tt chord.Ary*Rel} in the case of a
program relation, and (a subclass of) {\tt chord.Entity} in the case of a Java-based
program analysis.
\end{enumerate}

\item
Compute the following maps as shown below:

{\tt ConsumedNameToTasksMap\ :\ String -> Set<Task>},
mapping each name to the set of (possibly empty) tasks that consume that name.

{\tt NameToTaskMap\ :\ String -> Task},
mapping each name to the unique task that produces that name.

\Chord\ reports an error if a name is produced by no task or by more than one task.
{\small 
\begin{verbatim}
for each (<task> in Tasks) {
    case <task> came from item 1(a) above: {
        // <task> corresponds to a Datalog-based program analysis
        for each (<name> declared as input relation) {
            add <task> to ConsumedNameToTasksMap(<name>)
        }
        for each (<name> declared as output relation) {
            if (<name> is in domain of map NameToTaskMap)
                error ``multiple tasks produce name <name>''
            else
                NameToTaskMap(<name>) = <task>
        }
    }
    case <task> came from item 1(b) above: {
        // <task> corresponds to a program domain, a program relation, or
        // a Java-based program analysis
        // Suppose <task> came from entry <name> <class-name> in
        // TaskMapEntries
        // i.e. <task> is an object of class named <class-name>
        for each (<name2> in <task>.getConsumedNames())
            add <task> to ConsumedNameToTaskMaps(<name2>)
        for each (<name2> in <task>.getProducedNames() union { <name> })
            if (<name2> is in domain of map NameToTaskMap)
                error ``multiple tasks produce name <name>''
            else
                NameToTaskMap(<name2>) = <task>
    }
}
\end{verbatim}
}

{\small 
\begin{verbatim}
for each (<name> in domain of map ConsumedNameToTasksMap) {
    if (<name> is not in domain of map NameToTaskMap)
        error ``no task produces name <name>''
}
\end{verbatim}
}

\item
Compute \NameToTrgtMap : {\tt String -> Entity},
mapping each name to a unique target, as below.
\Chord\ reports an error if a name is mapped to multiple targets
or to no target.

{\small 
\begin{verbatim}
for each (entry <name> <class-name> in TaskMapEntries) {
    if (<name> is in domain of map NameToTrgtMap)
        error ``name <name> mapped to multiple targets''
    NameToTrgtMap(<name>) = NameToTaskMap(<name>)
}
for each (entry <name> <class-name> in TrgtMapEntries) {
    if (<name> is in domain of map NameToTrgtMap)
        error ``name <name> mapped to multiple targets''
    NameToTrgtMap(<name>) = a fresh object of class named <class-name>
}
\end{verbatim}
}

To reduce user burden in providing \TaskMapEntries\ and
\TrgtMapEntries, \Chord\ implicitly extends \NameToTrgtMap\ using
\DlogMapEntries\ as follows:

{\small 
\begin{verbatim}
for each (<name> in domain of map NameToTaskMap) {
    if (<name> is not in domain of map NameToTrgtMap) {
        if (<name> is declared as a domain in some datalog file in DlogMapEntries)
            NameToTrgtMap(<name>) = a fresh object of class chord.Dom
        else if (<name> is declared as a (input or output) relation
                in some datalog file in DlogMapEntries)
            NameToTrgtMap(<name>) = a fresh object of class chord.Rel
        else
            error ``<name> not mapped to any target''
    } else {
        chord.Entity <trgt> = NameToTrgtMap(<name>)
        if (<name> is declared as a domain in some datalog file in DlogMapEntries
                and <trgt> is not of type (subclass of) chord.Dom)
            error ``<name> has conflicting types''
        else if (<name> is declared as a (input or output) relation
                in some datalog file in DlogMapEntries
                and <trgt> is not of type (subclass of) chord.Rel)
            error ``<name> has conflicting types''
    }
}
\end{verbatim}
}

\item
Compute graph \TaskGraph\ as a graph whose:

\begin{itemize}
\item
set of nodes is \Tasks\ and
\item
set of edges containing an edge from task {\tt <task1>} to task {\tt <task2>} 
whenever there exists a name {\tt <name>} such that
\NameToTaskMap({\tt <name>}) = {\tt <task1>} and
\ConsumedNameToTasksMap({\tt <name>}) contains {\tt <task2>}.
\end{itemize}

The task graph in XML and HTML formats is dumped to files {\tt taskgraph.xml}
and {\tt taskgraph.html} in the project's working directory specified using
system property {\tt chord.work.dir}.

\end{enumerate}

\texonly{\newpage}
\section{Project Organization}

\Chord\ interprets system property {\tt chord.trgts}, if defined, as
a space-separated list of names of program domains, program relations, and
Datalog-based or Java-based program analyses, denoted \TrgtNames,
to be computed in that order.

To avoid redundant computation, \Chord\ maintains during its execution a set of
up-to-date tasks, denoted \ValidTasks,
which is a subset of the set of all tasks, denoted \Tasks.
Set \ValidTasks\ is empty at startup.
A future release of \Chord\ may provide the ability to retrieve a set of up-to-date tasks from
a previous run of \Chord\ to prevent re-computing those tasks in subsequent runs.
The following procedure is executed upon startup:

{\small
\begin{verbatim}
for each (<name> in TrgtNames in order) {
    if (<name> is not in domain of map NameToTaskMap) {
        error ``name <name> is an invalid target name''
    else {
        chord.Entity <task> = NameToTaskMap(<name>)
        executeTask(<task>)
    }
}
\end{verbatim}
}

\noindent where procedure {\tt executeTask} is defined as follows:

{\small
\begin{verbatim}
procedure executeTask(chord.Entity <task>) {
    if (<task> is not in ValidTasks) {
        for each (immed. pred. <task2> of <task> in TaskGraph)
            executeTask(<task2>)
        <task>.execute()
        for each (immed. succ. <task2> of <task> in TaskGraph)
            invalidateTask(<task2>)
        add <task> to ValidTasks
    }
}
\end{verbatim}
}

\noindent where procedure {\tt invalidateTask} is defined as follows:

{\small
\begin{verbatim}
procedure invalidateTask(chord.Entity t) {
    if (<task> is in ValidTasks) {
        remove <task> from ValidTasks
        <task>.invalidate()
        for each (immed. succ. <task2> of <task> in TaskGraph)
            invalidateTask(<task2>)
    }
}
\end{verbatim}
}

\texonly{\newpage}
\section{Chord System Properties}
\label{sec-sys-prop}

\Chord\ recognizes the following system properties:

\begin{itemize}
\item
{\tt chord.out}

The location of a file to which the standard output stream
of \Chord's execution is redirected.

It may be the same file to which the standard error stream is redirected (see
the description of system property {\tt chord.err}).

The file is created if it does not exist and its contents are overwritten if it exists.

\item
{\tt chord.err}

The location of a file to which the standard error stream
of \Chord's execution is redirected.

It may be the same file to which the standard output stream is redirected (see
the description of system property {\tt chord.out}).

The file is created if it does not exist and its contents are overwritten if it exists.

\item
{\tt chord.program.builder.class}

The fully-qualified name of the class to be used for building the
intermediate representation of the program to be analyzed.
It must be a concrete class with a nullary constructor that
implements interface {\tt chord.program.builders.IProgramBuilder}.

The default value of this property is {\tt chord.program.builders.SootProgramBuilder}
which uses the Soot Java compiler framework \cite{Soot}.

\item
{\tt chord.serial.file}

The location of a file from (resp.\ to) which
the serialized form of the intermediate representation of the
program to be analyzed is loaded (resp.\ stored).

\item
{\tt chord.main.class}

The fully-qualified name of the main class (e.g., {\tt foo.Main}) of the
program to be analyzed.

This system property (and {\tt chord.class.path})
must be defined if system property {\tt chord.serial.file}
does not specify the file containing the serialized form of the intermediate representation
of the program to be analyzed.

\item
{\tt chord.class.path}

The application classpath of the program to be analyzed (as opposed to the
JDK library classpath which consists of jar files such as {\tt rt.jar}
which are implicitly appended to the program's application classpath to
yield the complete program classpath).

Each path element must be the location of a directory or a jar/zip file.

The path elements must be separated by the system-specific path separator character
(e.g., `:' on UNIX and `;' on Windows).

This system property (and {\tt chord.main.class}) must
be defined if system property {\tt chord.serial.file}
does not specify the file containing the serialized form of the intermediate representation
of the program to be analyzed.

\item
{\tt chord.src.path}

The Java source path of the program to be analyzed.

Each path element must be the location of a directory.

The purpose of this path is to provide the location of the
{\tt .java} files of the program being analyzed so that, for instance, the
results of an analysis (which operates on an intermediate representation of the program
extracted from {\tt .class} files) can be mapped to the corresponding {\tt .java}
files and presented to the user for manual inspection.

Each file having suffix {\tt .java} in a directory (or a sub-directory of a directory)
specified in this path is regarded as a source file of the program.

\item
{\tt chord.dlog.analysis.path}

The path specifying the mapping from the name of each Datalog-based program
analysis that must be included in the project to the location
of the file defining that program analysis.

Each path element must be the location of a file containing a
list of entries (one per line) of the form:

{\tt <name> <file-name>}

The path elements must be separated by the system-specific path separator character
(e.g., `:' on UNIX and `;' on Windows).

\item
{\tt chord.java.analysis.path}

The path specifying the mapping from the name of each program domain, program relation,
and Java-based program analysis that must be included in the project to the
fully-qualified name of a subclass, either pre-defined in \Chord\ or defined by the user, of
{\tt chord.Dom}, {\tt chord.Ary*Rel}, and {\tt chord.Entity}, respectively, which
define the following instance methods:
\begin{itemize}
\item
the {\tt execute} method specifying
the computation that updates the state of the corresponding domain, relation,
or analysis, 
\item
the {\tt invalidate} method specifying the action to be taken when the
state of the corresponding domain, relation, or analysis is invalidated.
\item
the {\tt getConsumedNames} method specifying names of any domains, relations,
and analyses upon whose state the computation depends, and
\item
the {\tt getProducedNames} method specifying names of any domains, relations,
and analyses whose state depends upon the computation.
\end{itemize}

Each path element must be the location of a file containing a
list of entries (one per line) of the form:

{\tt <name> <class-name>}

The path elements must be separated by the system-specific path separator character
(e.g., `:' on UNIX and `;' on Windows).

\item
{\tt chord.trgtmap.path}

The path specifying the mapping from the name of each program domain,
program relation, and Java-based program analysis that must be
included in the project to the fully-qualified name of a subclass,
either pre-defined in \Chord\ or defined by the user, of {\tt chord.Dom},
{\tt chord.Ary*Rel}, and
{\tt chord.Entity}, respectively, the instance fields of which define the
state of the corresponding domain, relation, or analysis.

Each path element in the list must be the location of a file containing a
list of entries (one per line) of the form:

{\tt <name> <class-name>}

The path elements must be separated by the system-specific path separator character
(e.g., `:` on UNIX and ';' on Windows).

\item
{\tt chord.analyses}

A space-separated list of the names of program domains, program relations, and
Java-based or Datalog-based program analyses that must be computed in that order.

\item
{\tt chord.ext.class.path}

The classpath of user-defined program domains, user-defined program relations,
and user-defined Java-based program analyses.
It is implicitly appended to the classpath of pre-defined program domains,
pre-defined program relations, and pre-defined Java-based program analyses
to yield the complete runtime classpath of \Chord.

Each path element must be the location of a directory or a jar/zip file.

The path elements must be separated by the system-specific path separator character
(e.g., `:' on UNIX and `;' on Windows).
\end{itemize}

\texonly{\newpage}
\section{Pre-defined Domains}
\label{sec-predef-domains}

The following domains are pre-defined in \Chord:

\begin{itemize}
\item
$\Cdom$

The domain of all k-object-sensitive contexts.

The $0^{th}$ element in this domain denotes a distinguished hypothetical
context (denoted \texorhtml{$\epsilon$}{{\it epsilon}} in the literature) that is regarded as the
lone abstract context in which certain static methods are analyzed.
The main static method with signature {\tt void main(java.lang.String[])}
has no callers and is therefore analyzed in this context.
For simplicity, the k-object-sensitive may alias analysis provided in
\Chord\ also analyzes each class initializer method in this context.

\item
$\Edom$

The domain of all statements that access (read or write) an instance field,
a static field, or an array element, namely, all statements
represented by instances of (subclasses of) {\tt chord.program.insts.HeapInst}.

\item
$\Fdom$

The domain of  all instance and static fields.

The $0^{th}$ element in this domain denotes a distinguished hypothetical
field that is regarded as accessed whenever an array element is
accessed.

\item
$\Hdom$

The domain of all object allocation statements, namely, all statements
represented by instances of {\tt chord.program.insts.ObjValAsgnInst}.

\item
$\Idom$

The domain of all method invocation statements, namely, all statements
represented by instances of {\tt chord.program.insts.InvkInst}.

\item
$\Mdom$

The domain of all instance and static methods.

The $0^{th}$ element in this domain is the main method of the program.

The $1^{st}$ element in this domain is the method with signature
{\tt void start()} of class {\tt java.lang.Thread}, if this
method is deemed reachable from the main method of the program.

The above two methods are the root methods of the implicitly spawned
main thread and each explicitly spawned child thread, respectively.
Due to \Chord's emphasis on concurrency, these methods are
referenced frequently by various pre-defined program analyses
expressed in Datalog, and giving them special indices makes it
convenient to reference them in those analyses.

\item
$\Pdom$

The domain of all simple statements, namely, all statements
represented by instances of (subclasses of) {\tt chord.program.insts.Inst}.

The $0^{th}$ element in this domain is the statement at the unique
entry point of the main method of the program.

The statements of each method in the program are assigned
contiguous indices in this domain, with the statements at the
unique entry and exit points of each method being assigned
the smallest and largest indices, respectively, of all
indices assigned to statements in that method.

\item
$\Tdom$

The domain of all types, including the types of classes, interfaces,
arrays, and primitives.

\item
$\Vdom$

The domain of all local variables.

Each local variable declared in each block of each method is
represented by a unique element in this domain, in particular,
local variables that have the same name but are declared in
different methods or in different blocks of the same method
are represented by different elements in this domain.

All local variables of the same method are assigned contiguous
indices in this domain, in particular, the set of local
variables of a method is the disjoint union of its argument
variables and its temporary variables, and the argument
variables are assigned contiguous indices in order followed by
the temporary variables.

\item
$\Zdom$

The domain of all argument and return variable positions.
Let $N$ be the largest number or arguments or return variables of any
method.  Then, this domain contains the $N$ elements $0, 1, ..., N-1$.
\end{itemize}

\texonly{\newpage}
\section{Pre-defined Relations}
\label{sec-predef-relations}

The following relations are pre-defined in \Chord:

\begin{itemize}

\item
$\cha : \Mdom \times \Tdom \times \Mdom$

Relation containing each tuple $(m_1,t,m_2)$ such that $m_2$ is
the target method of a method invocation statement whose {\it resolved method}
is $m_1$ and whose receiver has run-time type $t$.
It is named $\cha$ because it represents the relation
computed by Class Hierarchy Analysis.

\item
$\sub : \Tdom \times \Tdom$

Relation containing each tuple $(s,t)$ such that type $s$ is a
subtype of type $t$.

\item
$\HT : \Hdom \times \Tdom$

Relation containing each tuple $(h,t)$ such that object allocation
statement $h$ allocates objects of type $t$.

\item
$\VT : \Vdom \times \Tdom$

Relation containing each tuple $(v,t)$ such that local variable $v$
has type $t$.

\item
$\MH : \Mdom \times \Hdom$

Relation containing each tuple $(m,h)$ such that method $m$ contains
object allocation statement $h$.

\item
$\MI : \Mdom \times \Idom$

Relation containing each tuple $(m,i)$ such that method $m$ contains
method invocation statement $i$.

\item
$\ME : \Mdom \times \Edom$

Relation containing each tuple $(m,e)$ such that method $m$ contains
statement $e$ that accesses (reads or writes) an instance field, a
static field, or an array element.

\item
$\instM : \Mdom$

Relation containing all instance (as opposed static) methods.

\item
$\statM : \Mdom$

Relation containing all static (as opposed instance) methods.

\item
$\clinitM : \Mdom$

Relation containing all class initializer methods, namely,
methods having signature {\tt void <clinit>()}.

\item
$\IinvkArg : \Idom \times \Zdom \times \Vdom$

Relation containing each tuple $(i,z,v)$ such that local variable $v$
is the $z^{th}$ argument variable of method invocation statement $i$.

\item
$\IinvkRet : \Idom \times \Zdom \times \Vdom$

Relation containing each tuple $(i,z,v)$ such that local variable $v$
is the $z^{th}$ return variable of method invocation statement $i$.

\item
$\MmethArg : \Mdom \times \Zdom \times \Vdom$

Relation containing each tuple $(m,z,v)$ such that local variable
$v$ is the $z^{th}$ argument variable of method $m$.

\item
$\MmethRet : \Mdom \times \Zdom \times \Vdom$

Relation containing each tuple $(m,z,v)$ such that local variable
$v$ is the $z^{th}$ return variable of method $m$.

\item
$\MspcMethArg : \Mdom \times \Vdom$

Relation containing each tuple $(m,v)$ such that local variable $v$ is
the implicit {\tt this} argument variable of instance method $m$.

\item

$\MordMethArg : \Mdom \times \Zdom \times \Vdom$

Relation containing each tuple $(m,z,v)$ such that local variable $v$
is the $z^{th}$ argument variable of method $m$ other than the implicit
{\tt this} argument variable (in the case in which $m$ is an instance
method).

\item
$\MobjVarAsgnInst : \Mdom \times \Vdom \times \Vdom$

Relation containing each tuple $(m,l,r)$ such that method $m$
contains a copy statement of the form $l = r$ where $l$ and $r$ are
local variables.

\item
$\MobjValAsgnInst : \Mdom \times \Vdom \times \Vdom$

Relation containing each tuple $(m,l,h)$ such that method $m$ contains
object allocation statement $h$ which assigns to local variable $l$.

\item
$\MclsVarAsgnInst : \Mdom \times \Vdom \times \Vdom$

Relation containing each tuple $(m,l,b)$ such that method $m$
contains a statement of the form $l = b.{\tt getClass()}$.

\item
$\MgetInstFldInst : \Mdom \times \Vdom \times \Fdom \times \Vdom$

Relation containing each tuple $(m,b,f,r)$ such that method $m$
contains a statement of the form $b.f = r$ which assigns to
instance field or array element $f$ of the object denoted by local
variable $b$ the value denoted by local variable $r$.

\item
$\MputInstFldInst : \Mdom \times \Vdom \times \Fdom \times \Vdom$

Relation containing each tuple $(m,b,f,r)$ such that method $m$
contains a statement of the form $b.f = r$ which writes the value
denoted by local variable $r$ to instance field or array element
$f$ of the object denoted by local variable $b$.

\item
$\MgetStatFldInst : \Mdom \times \Vdom \times \Fdom$

Relation containing each tuple $(m,l,f)$ such that method $m$ contains
a statement of the form $l = f$ that assigns to local variable $l$ the
value denoted by static field $f$.

\item
$\MputStatFldInst : \Mdom \times \Fdom \times \Vdom$

Relation containing each tuple $(m,f,r)$ such that method $m$ contains
a statement of the form $f = r$ which assigns to static field $f$ the
value denoted by local variable $r$.

\item
$\virtIM : \Idom \times \Mdom$

Relation containing each tuple $(i,m)$ such that $m$ is the
{\it resolved method} of {\tt invokevirtual} or {\tt invokeinterface} method
invocation statement $i$.

\item
$\statIM : \Idom \times \Mdom$

Relation containing each tuple $(i,m)$ such that $m$ is the
{\it resolved method} of {\tt invokestatic} method invocation statement $i$.

\item
$\specIM : \Idom \times \Mdom$

Relation containing each tuple $(i,m)$ such that $m$ is the
{\it resolved method} of {\tt invokespecial} method invocation statement $i$.

\item
$\syncMV : \Mdom \times \Vdom$

Relation containing each tuple $(m,v)$ such that method $m$ is
synchronized on argument variable $v$.

\item
$\writeE : \Edom$

Relation containing all statements that write (as opposed to read)
an instance field, static field, or array element.

\item
$\EF : \Edom \times \Fdom$

Relation containing each tuple $(e,f)$ such that statement $e$
accesses (reads or writes) instance field, static field, or
array element $f$.

\item
$\VH : \Vdom \times \Hdom$

Relation containing each tuple $(v,h)$ such that local variable $v$
may point to abstract object $h$, as determined by 0CFA-based may
alias analysis.

\item
$\FH : \Fdom \times \Hdom$

Relation containing each tuple $(f,h)$ such that static field $f$ may
point to abstract object $h$, as determined by 0CFA-based may alias
analysis.

\item

$\HFH : \Hdom \times \Fdom \times \Hdom$

Relation containing each tuple $(h_1,f,h_2)$ such that instance field
or array element $f$ of abstract object $h_1$ may point to abstract
object $h_2$, as determined by 0CFA-based may alias analysis.

\item
$\IM : \Idom \times \Mdom$

Relation containing each tuple $(i,m)$ such that method invocation
statement $i$ may call method $m$ as determined by 0CFA-based may alias
analysis.

\item
$\nestedLocks : \Cdom \times \Cdom \times \Mdom \times \Cdom \times \Mdom$

Relation containing each tuple $(c,c_1,m_1,c_2,m_2)$ such that in the
context-sensitive call graph computed by k-object-sensitive may
alias analysis, abstract thread $c$ may first acquire the lock at
the entry of method $m_1$ in abstract context $c_1$ and without
releasing it proceed to acquire the lock at the entry of method
$m_2$ in abstract context $c_2$.

\item
$\forkC : \Cdom$

Relation containing all abstract contexts in which the
{\tt start()} method of class {\tt java.lang.Thread} may
be reachable from the main method, as determined by
k-object-sensitive may alias analysis.

\item
$\CVC : \Cdom \times \Vdom \times \Cdom$

Relation containing each tuple $(c_1,v,c_2)$ such that local variable
$v$ may point to abstract object $c_2$ in abstract context $c_1$ of its
containing method, as determined by k-object-sensitive may alias
analysis.

\item
$\FC : \Fdom \times \Cdom$

Relation containing each tuple $(f,c)$ such that static field $f$
may point to abstract object $c$, as determined by k-object-sensitive
may alias analysis.

\item
$\CFC : \Cdom \times \Fdom \times \Cdom$

Relation containing each tuple $(c_1,f,c_2)$ such that instance field
or array element $f$ of abstract object $c_1$ may point to abstract
object $c_2$, as determined by k-object-sensitive may alias analysis.

\item
$\CICM : \Cdom \times \Idom \times \Cdom \times \Mdom$

Relation containing each tuple $(c_1,i,c_2,m)$ such that method
invocation statement $i$ in abstract context $c_1$ of its containing method
may call method $m$ in abstract context $c_2$, as determined by
k-object-sensitive may alias analysis.

\item
$\CHC : \Cdom \times \Hdom \times \Cdom$

Relation containing each tuple $(c_1,h,c_2)$ such that the method
containing object allocation statement $h$ may be reachable in
abstract context $c_1$, and $c_2 = (h::c_1) + k$, where $k$ is the k
value of statement $h$, provided as an argument to k-object-sensitive
may alias analysis.

\end{itemize}

\texonly{\newpage}
\section{Pre-defined Datalog-based Analyses}
\label{sec-predef-datalog-analyses}

\texonly{\newpage}
\section{Pre-defined Java-based Analyses}
\label{sec-predef-java-analyses}

\texonly{\newpage}
\bibliographystyle{plain}
\bibliography{citations}

\end{document}

\newpage
\section{Defining Tasks and Targets}

\Chord\ uses the notion of {\it tasks} to express computations
and the notion of {\it targets} to express the structure of
inputs/outputs of computations.

Tasks and targets are both implemented uniformly as {\it entities}.
A generic entity is represented by class {\tt chord.Entity}
depicted in Figure \ref{fig-entity}.

\begin{figure}
\begin{verbatim}
public class Entity {
    protected static final String[] emptyStringArray = new String[0];
    // subclasses may override
    public void execute() {
        // do nothing
    }
    // subclasses may override
    public String[] getConsumedTargetNames() {
        return emptyStringArray;
    }
    // subclasses may override
    public String[] getProducedTargetNames() {
        return emptyStringArray;
    }
    ...
}
\end{verbatim}
\caption{Snippet of class {\tt chord.Entity}}
\label{fig-entity}
\end{figure}

Whenever an enntity is regarded as a task:

\begin{itemize}
\item
The {\tt execute} method defines the computation that must be performed by the task.

\item
The {\tt getConsumedTargetNames} method specifies the names of targets
that are inputs to (the computation performed by) the task.

\item
The {\tt getProducedTargetNames} method specifies the names of targets
that are outputs of (the computation performed by) the task.
\end{itemize}

There are four kinds of entities:
\begin{itemize}
\item
A program domain

\item
A program relation

\item
A program analysis expressed in Datalog

\item
A program analysis expressed in Java

\end{itemize}

We next describe each of these kinds of entities in detail.

\subsection{Program Domains}

A program domain is a set of related elements in a given program (e.g.,
the set of all methods, the set of all object allocation statements, etc.).
The structure of a generic program domain is represented by
class {\tt chord.Dom} depicted in Figure \ref{fig-dom}.

\begin{figure}
\begin{verbatim}
public class Dom<T> extends Entity {
    public final IndexMap<T> map;
    public Dom() {
        this.map = new IndexMap<T>();
    }
    public void execute() {
        populate();
        saveToDomFile();
    }
    // subclasses cannot override
    public final String[] getProducedTargetNames() {
        return emptyStringArray;
    }
    // subclasses may override
    protected void populate() {
        // do nothing
    }
    ...
\end{verbatim}
\caption{Snippet of class {\tt chord.Dom}}
\label{fig-dom}
\end{figure}

A program domain provides a map from each element in the domain
to a unique integer in the range $[0..n-1]$ where $n$ is the size of
the domain.
This map is required by {\tt bddbddb} (a BDD-based Datalog solver).
The {\tt execute} method is specialized to construct the map and then
save the map to disk for use by {\tt bddbddb}.
Subclasses must override the {\tt populate} method to construct the
map as desired.

It is not possible to override the {\tt getProducedTargetNames}
method (note that it is declared {\tt final}) because a program domain
task is a specialized task whose output is exactly one target, namely,
the program domain it constructs.
It is, however, possible to override the {\tt getConsumedTargetNames}
method (inherited from superclass {\tt chord.Entity}) since constructing
a program domain may require other targets (i.e., other program
domains, program relations, or program analyses).

Figure \ref{fig-domI} defines the program domain
named $\Idom$ as the domain of all method invocation statements.
Note from method {\tt getConsumedTargetNames} that it takes as input
the domain named $\Mdom$ which is the domain of all methods.
This is because the {\tt populate} method refers to that domain.
Intuitively, the {\tt populate} method gives adjacent indices in
domain $\Idom$ to method invocation statements contained in
methods that have adjacent indices in domain $\Mdom$ are given
adjacent indices in domain $\Idom$.

\begin{figure}
\begin{verbatim}
public class DomI extends Dom<InvkInst> {
    private IndexMap<Method> mapM;
    private final String[] consumedTaskNames = new String[] { "M" };
    public String[] getConsumedTargetNames() {
        return consumedTaskNames;
    }
    protected void populate() {
        DomM domM = (DomM) project.getEntity("M");
        mapM = domM.map;
        for (int m = 0; m < mapM.size(); m++) {
            Method meth = mapM.get(m);
            Body body = meth.body;
            if (body == null)
                continue;
            Set<Inst> insts = body.cfg.getNodes();
            for (Inst inst : insts) {
                if (inst instanceof InvkInst)
                    map.set((InvkInst) inst);
            }
        }
    }
    ...
}
\end{verbatim}
\caption{Snippet of class {\tt chord.DomI}}
\label{fig-domI}
\end{figure}

A list of pre-defined program domains is presented in Section \ref{sec-predef-domains}.

\subsection{Program Relations}

A program relation is a relation over one or more program domains.
The structure of a generic program relation is represented by class
{\tt chord.Rel} depicted in Figure \ref{fig-rel}.

\begin{figure}
\begin{verbatim}
\end{verbatim}
\caption{Snippet of class {\tt chord.Rel}}
\label{fig-rel}
\end{figure}

It is not possible to override the {\tt getProducedTargetNames}
method (note that it is declared {\tt final}) because a program relation
task is a specialized task whose output is exactly one target, namely,
the program relation it constructs.
It is, however, possible to override the {\tt getConsumedTargetNames}
method (inherited from superclass {\tt chord.Entity}) since constructing
a program relation may require other targets (i.e., program
domains, other program relations, or program analyses).

A list of pre-defined program relations is presented in Section \ref{sec-predef-relations}.

\subsection{Program Analyses in Datalog}

A list of pre-defined program analyses in Datalog is presented in
Section \ref{sec-predef-datalog-analyses}.

\subsection{Program Analyses in Java}

A list of pre-defined program analyses in Java is presented in
Section \ref{sec-predef-java-analyses}.

### 2009-07-01 agent/jvmti_agent.cpp

@@ -1,0 +1,243 @@
+#include <jvmti.h>
+#include <assert.h>
+#include <iostream>
+#include <fstream>
+#include <set>
+#include <map>
+#include <vector>
+#include <string>
+using namespace std;
+
+static jvmtiEnv* jvmti_env;
+
+static jlong num_otags = 1;
+
+static fstream fp;
+
+static jrawMonitorID lock;
+
+static void enterAgentMonitor()
+{
+    jvmtiError retval = jvmti_env->RawMonitorEnter(lock);
+    assert(retval == JVMTI_ERROR_NONE);
+}
+
+static void exitAgentMonitor()
+{
+    jvmtiError retval = jvmti_env->RawMonitorExit(lock);
+    assert(retval == JVMTI_ERROR_NONE);
+}
+
+static jlong get_or_set_otag(jobject o) {
+    jlong otag;
+    jvmti_env->GetTag(o, &otag);
+    if (otag == 0) {
+        otag = num_otags;
+        jvmtiError result = jvmti_env->SetTag(o, otag);
+        assert(result == JVMTI_ERROR_NONE);
+        num_otags++;
+    }
+    return otag;
+}
+
+char* get_token(char *str, char *seps, char *buf, int max)
+{
+    int len;
+    buf[0] = 0;
+    if (str == NULL || str[0] == 0)
+        return NULL;
+    str += strspn(str, seps);
+    if (str[0] == 0)
+        return NULL;
+    len = (int) strcspn(str, seps);
+    if (len >= max) {
+		cerr << "ERROR: get_token failed" << endl;
+		exit(1);
+    }
+    strncpy(buf, str, len);
+    buf[len] = 0;
+    return str + len;
+}
+
+static char* getMethod(jmethodID method)
+{
+	jclass decl_class;
+	jvmti_env->GetMethodDeclaringClass(method, &decl_class);
+
+    char* class_name_ptr;
+	jvmti_env->GetClassSignature(decl_class, &class_name_ptr, NULL);
+	int n1 = strlen(class_name_ptr);
+
+	char* meth_name_ptr;
+	char* meth_sign_ptr;
+	jvmti_env->GetMethodName(method, &meth_name_ptr, &meth_sign_ptr, NULL);
+	int n2 = strlen(meth_name_ptr);
+	int n3 = strlen(meth_sign_ptr);
+
[…]

### 2009-07-01 main/java/chord/analyses/alias/CIAliasAnalysis.java

@@ -1,0 +1,124 @@
+/*
+ * Copyright (c) 2008-2009, Intel Corporation.
+ * Copyright (c) 2006-2007, The Trustees of Stanford University.
+ * All rights reserved.
+ */
+package chord.analyses.alias;
+
+import java.util.HashSet;
+import java.util.Set;
+
+import joeq.Class.jq_Field;
+import joeq.Compiler.Quad.Quad;
+import joeq.Compiler.Quad.RegisterFactory.Register;
+
+import chord.bddbddb.Rel.RelView;
+import chord.project.Chord;
+import chord.project.JavaTask;
+import chord.project.ProgramRel;
+import chord.project.Project;
+
+import chord.util.SetUtils;
+
+/**
+ * Context-insensitive may alias analysis.
+ * 
+ * @author Mayur Naik (mhn@cs.stanford.edu)
+ */
+@Chord(
+	name = "ci-alias-java",
+	consumedNames = { "VH", "FH", "HFH" }
+)
+public class CIAliasAnalysis extends JavaTask {
+	private ProgramRel relVH;
+	private ProgramRel relFH;
+	private ProgramRel relHFH;
+	public void run() {
+		relVH  = (ProgramRel) Project.getTrgt("VH");
+		relFH  = (ProgramRel) Project.getTrgt("FH");
+		relHFH = (ProgramRel) Project.getTrgt("HFH");
+	}
+	/**
+	 * Provides the abstract object to which a given local variable
+	 * may point.
+	 * 
+	 * @param	var	A local variable.
+	 * 
+	 * @return	The abstract object to which the given local variable
+	 * 			may point.
+	 */
+	public Obj pointsTo(Register var) {
+		if (!relVH.isOpen())
+			relVH.load();
+		RelView view = relVH.getView();
+		view.selectAndDelete(0, var);
+		Iterable<Quad> res = view.getAry1ValTuples();
+		Set<Ctxt> pts = SetUtils.newSet(view.size());
+		for (Quad inst : res)
+			pts.add(new Ctxt(new Quad[] { inst }));
+		view.free();
+		return new Obj(pts);
+	}
+	/**
+	 * Provides the abstract object to which a given static field
+	 * may point.
+	 * 
+	 * @param	field	A static field.
+	 * 
+	 * @return	The abstract object to which the given static field
+	 * 			may point.
+	 */
+	public Obj pointsTo(jq_Field field) {
+		if (!relFH.isOpen())
+			relFH.load();
[…]

### 2009-07-01 main/java/chord/analyses/alias/CICG.java

@@ -1,0 +1,139 @@
+/*
+ * Copyright (c) 2008-2009, Intel Corporation.
+ * Copyright (c) 2006-2007, The Trustees of Stanford University.
+ * All rights reserved.
+ */
+package chord.analyses.alias;
+
+import java.util.Set;
+
+import joeq.Class.jq_Method;
+import joeq.Compiler.Quad.Quad;
+
+import chord.bddbddb.Rel.RelView;
+import chord.doms.DomM;
+import chord.project.ProgramRel;
+import chord.util.SetUtils;
+import chord.util.graph.AbstractGraph;
+import chord.util.ArraySet;
+
+/**
+ * Implementation of a context-insensitive call graph.
+ * 
+ * @author Mayur Naik (mhn@cs.stanford.edu)
+ */
+public class CICG extends AbstractGraph<jq_Method> implements ICICG {
+	private DomM domM;
+	private ProgramRel relIM;
+	private ProgramRel relMM;
+	private ProgramRel relReachableM;
+	public CICG(DomM domM, ProgramRel relIM, ProgramRel relMM,
+			ProgramRel relReachableM) {
+		this.domM = domM;
+		this.relIM = relIM;
+		this.relMM = relMM;
+		this.relReachableM = relReachableM;
+	}
+	public Set<Quad> getCallers(jq_Method meth) {
+		if (!relIM.isOpen())
+			relIM.load();
+		RelView view = relIM.getView();
+		view.selectAndDelete(1, meth);
+		Iterable<Quad> res = view.getAry1ValTuples();
+		Set<Quad> invks = SetUtils.newSet(view.size());
+		for (Quad invk : res)
+			invks.add(invk);
+		return invks;
+	}
+	public Set<jq_Method> getTargets(Quad invk) {
+		if (!relIM.isOpen())
+			relIM.load();
+		RelView view = relIM.getView();
+		view.selectAndDelete(0, invk);
+		Iterable<jq_Method> res = view.getAry1ValTuples();
+		Set<jq_Method> meths = SetUtils.newSet(view.size());
+		for (jq_Method meth : res)
+			meths.add(meth);
+		return meths;
+	}
+	public int numRoots() {
+		return 1;
+	}
+	public int numNodes() {
+		if (!relReachableM.isOpen())
+			relReachableM.load();
+		return relReachableM.size();
+	}
+	public int numPreds(jq_Method node) {
+		throw new UnsupportedOperationException();
+	}
[…]

### 2009-07-01 main/java/chord/analyses/alias/CSAliasAnalysis.java

@@ -1,0 +1,120 @@
+/*
+ * Copyright (c) 2008-2009, Intel Corporation.
+ * Copyright (c) 2006-2007, The Trustees of Stanford University.
+ * All rights reserved.
+ */
+package chord.analyses.alias;
+
+import java.util.Set;
+import java.util.HashSet;
+
+import joeq.Class.jq_Field;
+import joeq.Compiler.Quad.RegisterFactory.Register;
+
+import chord.bddbddb.Rel.RelView;
+import chord.util.SetUtils;
+import chord.project.Chord;
+import chord.project.JavaTask;
+import chord.project.ProgramRel;
+import chord.project.Project;
+
+/**
+ * Context-sensitive may alias analysis.
+ * 
+ * @author Mayur Naik (mhn@cs.stanford.edu)
+ */
+@Chord(
+	name = "cs-alias-java"
+)
+public class CSAliasAnalysis extends JavaTask {
+	private ProgramRel relCVC;
+	private ProgramRel relFC;
+	private ProgramRel relCFC;
+	public void run() {
+		relCVC = (ProgramRel) Project.getTrgt("CVC");
+		relFC  = (ProgramRel) Project.getTrgt("FC");
+		relCFC = (ProgramRel) Project.getTrgt("CFC");
+	}
+	/**
+	 * Provides the abstract object to which a given local variable
+	 * may point in a given abstract context of its declaring method.
+	 * 
+	 * @param	ctxt	An abstract context of a method.
+	 * @param	var		A local variable declared in the method.
+	 * 
+	 * @return	The abstract object to which the given local variable
+	 *  		may point in the given abstract context.
+	 */
+	public Obj pointsTo(Ctxt ctxt, Register var) {
+		if (!relCVC.isOpen())
+			relCVC.load();
+		RelView view = relCVC.getView();
+		view.selectAndDelete(0, ctxt);
+		view.selectAndDelete(1, var);
+		Iterable<Ctxt> res = view.getAry1ValTuples();
+		Set<Ctxt> pts = SetUtils.iterableToSet(
+				res, view.size());
+		view.free();
+		return new Obj(pts);
+	}
+	/**
+	 * Provides the abstract object to which a given static field
+	 * may point.
+	 * 
+	 * @param	field	A static field.
+
+	 * @return	The abstract object to which the given static field
+	 * 			may point.
+	 */
+	public Obj pointsTo(jq_Field field) {
[…]

### 2009-07-01 main/java/chord/analyses/alias/ICSCG.java

@@ -1,0 +1,79 @@
+/*
+ * Copyright (c) 2008-2009, Intel Corporation.
+ * Copyright (c) 2006-2007, The Trustees of Stanford University.
+ * All rights reserved.
+ */
+package chord.analyses.alias;
+
+import java.util.Set;
+
+import joeq.Class.jq_Method;
+import joeq.Compiler.Quad.Quad;
+
+import chord.util.graph.ILabeledGraph;
+import chord.util.tuple.object.Pair;
+
+/**
+ * Specification of a context-sensitive call graph.
+ * 
+ * @author Mayur Naik (mhn@cs.stanford.edu)
+ */
+public interface ICSCG extends ILabeledGraph<Pair<Ctxt, jq_Method>, Quad> {
+	/**
+	 * Provides all abstract contexts in which a given jq_Method
+	 * may be reachable.
+	 * 
+	 * @param	meth	A jq_Method.
+	 * 
+	 * @return	All abstract contexts in which the jq_Method may
+	 * 			be reachable.
+	 */
+	public Set<Ctxt> getContexts(jq_Method meth);
+    /**
+     * Provides each jq_Method along with each abstract context in
+     * which it may be called by a given jq_Method invocation site
+     * in a given abstract context.
+     * 
+     * @param   ctxt    An abstract context.
+     * @param   invk    A jq_Method invocation site.
+     * 
+     * @return  All (<tt>ctxt2</tt>, <tt>meth</tt>) pairs such that
+     * 			jq_Method <tt>meth</tt> may be called in abstract
+     *			context <tt>ctxt2</tt> from jq_Method invocation site
+     * 			<tt>invk</tt> in abstract context <tt>ctxt</tt>.
+     */
+	public Set<Pair<Ctxt, jq_Method>>
+		getTargets(Ctxt ctxt, Quad invk);
+	/**
+	 * Provides each jq_Method invocation site along with each abstract
+	 * context from which it may call a given jq_Method in a given
+	 * abstract context.
+	 * 
+	 * @param   ctxt    An abstract context.
+	 * @param   meth    A jq_Method.
+	 * 
+	 * @return  All (<tt>ctxt2</tt>, <tt>invk</tt>) pairs such that
+	 * 			jq_Method invocation site <tt>invk</tt> in abstract
+	 * 			context <tt>ctxt2</tt> may call jq_Method <tt>meth</tt>
[…]

### 2009-07-01 main/java/chord/analyses/thread/FldObj.java

@@ -1,0 +1,7 @@
+package chord.analyses.thread;
+
+public class FldObj {
+    public int f;
+    public int o;
+    public FldObj(int f, int o) { this.f = f; this.o = o; }
+}

### 2009-07-01 main/java/chord/analyses/thread/FlowSenDynamicThreadEscapeAnalysis.java

@@ -1,0 +1,177 @@
+package chord.analyses.thread;
+
+import java.util.List;
+import java.util.ArrayList;
+
+import chord.project.Chord;
+import chord.project.ProgramRel;
+import chord.project.DynamicAnalysis;
+import chord.project.Project;
+import chord.util.Assertions;
+
+import gnu.trove.TIntHashSet;
+import gnu.trove.TIntArrayList;
+import gnu.trove.TIntObjectHashMap;
+
+@Chord(
+    name = "flowsen-dyn-thresc-java",
+    producedNames = { "escE", "visitedE" },
+    namesOfSigns = { "escE", "visitedE" },
+    signs = { "E0:E0", "E0:E0" }
+)
+public class FlowSenDynamicThreadEscapeAnalysis extends DynamicAnalysis {
+	// map from each object to a list containing each non-null-valued
+	// instance field of reference type along with that value
+	protected TIntObjectHashMap/*<List>*/ objToFldObjs;
+    // set of all currently escaping objects
+    protected TIntHashSet escObjs;
+
+    // isEidxEsc[e] == true iff instance field/array deref
+	// site having index e in domain E is deemed thread-escaping
+	protected boolean[] isEidxEsc;
+	// isEidxVisited[e] == true iff instance field/array deref
+	// site having index e in domain E is visited at least once
+	protected boolean[] isEidxVisited;
+
+    public boolean handlesObjValAsgnInst() { return true; }
+	public boolean handlesInstFldRdInst() { return true; }
+	public boolean handlesInstFldWrInst() { return true; }
+	public boolean handlesAryElemRdInst() { return true; }
+	public boolean handlesAryElemWrInst() { return true; }
+	public boolean handlesStatFldWrInst() { return true; }
+	public boolean handlesForkHeadInst() { return true; }
+
+	protected boolean isFirst = true;
+	protected int numE;
+
+	protected ProgramRel relEscE;
+	protected ProgramRel relVisitedE;
+
+	public void init() {
+		if (isFirst) {
+			isFirst = false;
+ 			escObjs = new TIntHashSet();
+			numE = Emap.size();
+			isEidxEsc = new boolean[numE];
+			isEidxVisited = new boolean[numE];
+			relEscE = (ProgramRel) Project.getTrgt("escE");
+			relVisitedE = (ProgramRel) Project.getTrgt("visitedE");
+		} else {
+			escObjs.clear();
+			printNumEsc();
+		}
+		objToFldObjs = new TIntObjectHashMap();
+	}
[…]

### 2009-07-01 main/java/chord/doms/DomL.java

@@ -1,0 +1,63 @@
+/*
+ * Copyright (c) 2008-2009, Intel Corporation.
+ * Copyright (c) 2006-2007, The Trustees of Stanford University.
+ * All rights reserved.
+ */
+package chord.doms;
+
+import joeq.Class.jq_Class;
+import joeq.Class.jq_Method;
+import joeq.Compiler.Quad.BasicBlock;
+import joeq.Compiler.Quad.ControlFlowGraph;
+import joeq.Compiler.Quad.Inst;
+import joeq.Compiler.Quad.Quad;
+import joeq.Compiler.Quad.Operator.Monitor;
+
+import chord.project.Chord;
+import chord.project.Program;
+import chord.project.ProgramDom;
+import chord.project.Project;
+import chord.visitors.ILockInstVisitor;
+
+/**
+ * Domain of all monitorenter statements.
+ * 
+ * @author Mayur Naik (mhn@cs.stanford.edu)
+ */
+@Chord(
+	name = "L",
+	consumedNames = { "M" }
+)
+public class DomL extends ProgramDom<Inst> implements ILockInstVisitor {
+	protected DomM domM;
+	protected jq_Method ctnrMethod;
+	public void init() {
+		domM = (DomM) Project.getTrgt("M");
+	}
+	public void visit(jq_Class c) { }
+	public void visit(jq_Method m) {
+		ctnrMethod = m;
+		if (m.isSynchronized()) {
+			ControlFlowGraph cfg = Program.getCFG(m);
+			if (cfg != null) {
+				BasicBlock head = cfg.entry();
+				Program.mapInstToMethod(head, m);
+				set(head);
+			}
+		}
+	}
+	public void visitLockInst(Quad q) {
+		Program.mapInstToMethod(q, ctnrMethod);
+		set(q);
+	}
+	public String toXMLAttrsString(Inst i) {
+		jq_Method m = Program.getMethod(i);
+		String fileName = Program.getSourceFileName(m.getDeclaringClass());
+		int lineNumber = Program.getLineNumber(i, m);
+		return "file=\"" + fileName + "\" " + "line=\"" + lineNumber + "\" " +
+			"Mid=\"M" + domM.get(m) + "\"";
+	}
+	public String toString(Inst q) {
+		return Program.toStringLockInst(q);
+	}
+}

### 2009-07-01 main/java/chord/doms/QuadDom.java

@@ -1,0 +1,50 @@
+/*
+ * Copyright (c) 2008-2009, Intel Corporation.
+ * Copyright (c) 2006-2007, The Trustees of Stanford University.
+ * All rights reserved.
+ */
+package chord.doms;
+
+import chord.project.Program;
+import chord.project.ProgramDom;
+import chord.project.Project;
+
+import joeq.Class.jq_Class;
+import joeq.Class.jq_Method;
+import joeq.Compiler.Quad.Quad;
+
+public abstract class QuadDom extends ProgramDom<Quad> {
+	protected DomM domM;
+	protected jq_Method ctnrMethod;
+	public void init() {
+		domM = (DomM) Project.getTrgt("M");
+	}
+	public void visit(jq_Class c) { }
+	public void visit(jq_Method m) {
+		ctnrMethod = m;
+	}
+	public int set(Quad q) {
+		// XXX
+		if (q != null)
+			Program.mapInstToMethod(q, ctnrMethod);
+		return super.set(q);
+	}
+	public String toUniqueIdString(Quad inst) {
+		/*
+		Method method = inst.getOutermostCtnrMethod();
+		Type type = method.getCtnrType();
+		return inst.getBCI() + "@" + method.getSign() + "@" + type.getName();
+		*/
+		return null; // TODO
+	}
+	public String toXMLAttrsString(Quad q) {
+		jq_Method m = Program.getMethod(q);
+		String fileName = Program.getSourceFileName(m.getDeclaringClass());
+		int lineNumber = Program.getLineNumber(q, m);
+		return "file=\"" + fileName + "\" " + "line=\"" + lineNumber + "\" " +
+			"Mid=\"M" + domM.get(m) + "\"";
+	}
+	public String toString(Quad q) {
+		return Program.toString(q);
+	}
+}

### 2009-07-01 main/java/chord/project/ChordAnnotParser.java

@@ -1,0 +1,225 @@
+/*
+ * Copyright (c) 2008-2009, Intel Corporation.
+ * Copyright (c) 2006-2007, The Trustees of Stanford University.
+ * All rights reserved.
+ */
+package chord.project;
+
+import java.util.Map;
+import java.util.Set;
+import java.util.HashSet;
+import java.util.HashMap;
+
+import chord.bddbddb.RelSign;
+import chord.util.Assertions;
+import chord.util.ClassUtils;
+
+/**
+ * Parser of a Chord annotation on a class defining a Java task.
+ *
+ * The annotation specifies aspects of the task such as its name,
+ * its consumed and produced targets, etc.
+ * 
+ * @author Mayur Naik (mhn@cs.stanford.edu)
+ */
+public class ChordAnnotParser {
+	private final Class type;
+	private String name;
+	private Map<String, RelSign> nameToSignMap;
+	private Map<String, Class> nameToTypeMap;
+	private boolean hasNoErrors;
+	private Set<String> consumedNames;
+	private Set<String> producedNames;
+	/**
+	 * Constructor.
+	 * 
+	 * @param	type	A class annotated with a Chord annotation.
+	 */
+    public ChordAnnotParser(Class type) {
+		this.type = type;
+	}
+    /**
+     * Parses this Chord annotation.
+     * 
+     * @return	true iff the Chord annotation parses successfully.
+     */
+	public boolean parse() {
+		Chord chord = (Chord) type.getAnnotation(Chord.class);
+		Assertions.Assert(chord != null);
+		hasNoErrors = true;
+		this.name = chord.name();
+
+		String[] names;
+
+		names = chord.consumedNames();
+		consumedNames = new HashSet<String>(names.length);
+		for (String name2 : names) {
+			consumedNames.add(name2);
+		}
+		names = chord.producedNames();
+		producedNames = new HashSet<String>(names.length);
+		for (String name2 : names) {
+			producedNames.add(name2);
+		}
+		producedNames.add(name);
+
+		nameToTypeMap = new HashMap<String, Class>();
+		nameToSignMap = new HashMap<String, RelSign>();
+
+		nameToTypeMap.put(name, type);
+
+		String sign = chord.sign();
+		if (ClassUtils.isSubclass(type, ProgramRel.class)) {
+			if (sign.equals("")) {
+				error("Method sign() cannot return empty string " +
+					"for Java analysis '" + type + "'");
+			} else {
[…]

### 2009-07-01 main/java/chord/project/ITask.java

@@ -1,0 +1,38 @@
+/*
+ * Copyright (c) 2008-2009, Intel Corporation.
+ * Copyright (c) 2006-2007, The Trustees of Stanford University.
+ * All rights reserved.
+ */
+package chord.project;
+
+/**
+ * Specification of a program analysis.
+ * 
+ * @author Mayur Naik (mhn@cs.stanford.edu)
+ */
+public interface ITask {
+	/**
+	 * Sets the name of this program analysis.
+	 * 
+	 * @param	name	A name unique across all program analyses
+	 *			included in a Chord project.
+	 */
+    public void setName(String name);
+    /**
+     * Provides the name of this program analysis.
+     * 
+     * @return	The name of this program analysis.
+     */
+    public String getName();
+	/**
+	 * Executes this program analysis.
+	 * 
+	 * This method must usually not be called directly from another
+	 * program analysis that wishes to execute this program analysis.
+	 * The correct way to achieve this is to call
+	 * {@link chord.project.Project#runTask(String)} or
+	 * {@link chord.project.Project#runTask(ITask)}, providing this
+	 * program analysis either by its name or its object.
+	 */
+	public void run();
+}

### 2009-07-01 main/java/chord/project/Instrumentor.java

@@ -1,0 +1,264 @@
+/*
+ * Copyright (c) 2008-2009, Intel Corporation.
+ * Copyright (c) 2006-2007, The Trustees of Stanford University.
+ * All rights reserved.
+ */
+package chord.project;
+
+import javassist.*;
+import javassist.bytecode.Descriptor;
+import javassist.expr.*;
+import java.util.*;
+import java.io.*;
+
+import chord.util.ClasspathUtils;
+import chord.util.FileUtils;
+import chord.util.PropertyUtils;
+import chord.util.IndexMap;
+import chord.util.Assertions;
+import java.net.URL;
+import java.net.MalformedURLException;
+import org.scannotation.AnnotationDB;
+
+/**
+ *
+ * @author Mayur Naik (mhn@cs.stanford.edu)
+ */
+public class Instrumentor {
+	private static IndexMap<String> Hmap;
+	private static IndexMap<String> Emap;
+	private static IndexMap<String> Lmap;
+	private static IndexMap<String> Fmap;
+	private static int numH, numE, numL;
+
+	private static IndexMap<String> readFileToMap(String dirName, String fileName) {
+		File file = new File(dirName, fileName);
+		if (!file.exists())
+			return new IndexMap<String>();
+		return FileUtils.readFileToMap(file);
+	}
+
+	private static List<String> readFileToList(String dirName, String fileName) {
+		File file = new File(dirName, fileName);
+		if (!file.exists())
+			return new ArrayList<String>();
+		return FileUtils.readFileToList(file);
+	}
+
+	public static void main(String[] args) throws Exception {
+		String classPath = System.getProperty("chord.class.path", "");
+		String outDirName = System.getProperty("chord.out.dir", ".");
+		String classesDirName = System.getProperty("chord.classes.dir", ".");
+		String jdkOutDirName = System.getProperty("chord.jdk.out.dir");
+		if (jdkOutDirName == null) {
+			Hmap = new IndexMap<String>();
+			Emap = new IndexMap<String>();
+			Lmap = new IndexMap<String>();
+			Fmap = new IndexMap<String>();
+		} else {
+			Hmap = readFileToMap(jdkOutDirName, "H.dynamic.txt");
+			Emap = readFileToMap(jdkOutDirName, "E.dynamic.txt");
+			Lmap = readFileToMap(jdkOutDirName, "L.dynamic.txt");
[…]

### 2009-07-01 main/java/chord/project/Runtime.java

@@ -1,0 +1,168 @@
+/*
+ * Copyright (c) 2008-2009, Intel Corporation.
+ * Copyright (c) 2006-2007, The Trustees of Stanford University.
+ * All rights reserved.
+ * Licensed under the terms of the BSD License; see COPYING for details.
+ */
+package chord.project;
+
+import chord.util.IntBuffer;
+import java.io.IOException;
+
+/**
+ * 
+ * @author Mayur Naik (mhn@cs.stanford.edu)
+ */
+public class Runtime {
+	private static IntBuffer buffer;
+	private static boolean trace = false;
+
+	public synchronized static void befNewInst(int hIdx) {
+		if (trace) {
+			trace = false;
+			Thread t = Thread.currentThread();
+			int tIdx = System.identityHashCode(t);
+			try {
+				buffer.put(InstKind.BEF_NEW_INST);
+				buffer.put(tIdx);
+				buffer.put(hIdx);
+			} catch (IOException ex) { throw new RuntimeException(ex); }
+			trace = true;
+		}
+	}
+	public synchronized static void aftNewInst(int hIdx, Object o) {
+		if (trace) {
+			trace = false;
+			Thread t = Thread.currentThread();
+			int tIdx = System.identityHashCode(t);
+			int oIdx = System.identityHashCode(o);
+			try {
+				buffer.put(InstKind.AFT_NEW_INST);
+				buffer.put(tIdx);
+				buffer.put(hIdx);
+				buffer.put(oIdx);
+			} catch (IOException ex) { throw new RuntimeException(ex); }
+			trace = true;
+		}
+	}
+	public synchronized static void newArrayInst(int hIdx, Object o) {
+		if (trace) {
+			trace = false;
+			int oIdx = System.identityHashCode(o);
+			try {
+				buffer.put(InstKind.NEW_ARRAY_INST);
+				buffer.put(hIdx);
+				buffer.put(oIdx);
+			} catch (IOException ex) { throw new RuntimeException(ex); }
+			trace = true;
+		}
+	}
+	public synchronized static void instFldRdInst(int eIdx, Object b,
+			int fIdx) {
+		if (trace) {
+			trace = false;
+			int bIdx = System.identityHashCode(b);
+			try {
+				buffer.put(InstKind.INST_FLD_RD_INST);
+				buffer.put(eIdx);
+				buffer.put(bIdx);
+				buffer.put(fIdx);
[…]

### 2009-07-01 main/java/chord/rels/RelClinitM.java

@@ -1,0 +1,32 @@
+/*
+ * Copyright (c) 2008-2009, Intel Corporation.
+ * Copyright (c) 2006-2007, The Trustees of Stanford University.
+ * All rights reserved.
+ */
+package chord.rels;
+
+import joeq.Class.jq_Class;
+import joeq.Class.jq_ClassInitializer;
+import joeq.Class.jq_Method;
+import chord.project.Chord;
+import chord.project.ProgramRel;
+import chord.visitors.IMethodVisitor;
+
+/**
+ * Relation containing all class initializer methods, namely, methods
+ * having signature <tt>&lt;clinit&gt;()</tt>.
+ * 
+ * @author Mayur Naik (mhn@cs.stanford.edu)
+ */
+@Chord(
+	name = "clinitM",
+	sign = "M0"
+)
+public class RelClinitM extends ProgramRel
+		implements IMethodVisitor {
+	public void visit(jq_Class c) { }
+	public void visit(jq_Method m) {
+		if (m instanceof jq_ClassInitializer)
+			add(m);
+	}
+}

### 2009-07-01 main/java/chord/rels/RelHT.java

@@ -1,0 +1,46 @@
+/*
+ * Copyright (c) 2008-2009, Intel Corporation.
+ * Copyright (c) 2006-2007, The Trustees of Stanford University.
+ * All rights reserved.
+ */
+package chord.rels;
+
+import joeq.Class.jq_Type;
+import joeq.Compiler.Quad.Operator;
+import joeq.Compiler.Quad.Quad;
+import joeq.Compiler.Quad.Operator.New;
+import joeq.Compiler.Quad.Operator.NewArray;
+import chord.doms.DomH;
+import chord.doms.DomT;
+import chord.project.Chord;
+import chord.project.ProgramRel;
+
+/**
+ * Relation containing each tuple (h,t) such that object allocation
+ * statement h allocates objects of type t.
+ *
+ * @author Mayur Naik (mhn@cs.stanford.edu)
+ */
+@Chord(
+	name = "HT",
+	sign = "H0,T1:T1_H0"
+)
+public class RelHT extends ProgramRel {
+	public void fill() {
+		DomH domH = (DomH) doms[0];
+		DomT domT = (DomT) doms[1];
+		int numH = domH.size();
+		for (int hIdx = 0; hIdx < numH; hIdx++) {
+			Quad h = domH.get(hIdx);
+			Operator op = h.getOperator();
+			jq_Type t;
+			if (op instanceof New)
+				t = New.getType(h).getType();
+			else
+				t = NewArray.getType(h).getType();
+			int tIdx = domT.get(t);
+			if (tIdx != -1)
+				add(hIdx, tIdx);
+		}
+	}
+}

### 2009-07-01 main/java/chord/rels/RelImplsTT.java

@@ -1,0 +1,39 @@
+/*
+ * Copyright (c) 2008-2009, Intel Corporation.
+ * Copyright (c) 2006-2007, The Trustees of Stanford University.
+ * All rights reserved.
+ */
+package chord.rels;
+
+import joeq.Class.jq_Class;
+
+import chord.doms.DomT;
+import chord.project.Chord;
+import chord.project.Program;
+import chord.project.ProgramRel;
+
+/**
+ * Relation containing each tuple (s,t) such that class type s
+ * implements interface type t.
+ *
+ * @author Mayur Naik (mhn@cs.stanford.edu)
+ */
+@Chord(
+	name = "implsTT",
+	sign = "T1,T0:T0_T1"
+)
+public class RelImplsTT extends ProgramRel {
+	public void fill() {
+		DomT domT = (DomT) doms[0];
+		for (jq_Class c : Program.getClasses()) {
+			if (!c.isInterface()) {
+				int t1Idx = domT.get(c);
+				for (jq_Class d : c.getInterfaces()) {
+					int t2Idx = domT.get(d);
+					if (t2Idx != -1)
+						add(t1Idx, t2Idx);
+				}
+			}
+		}
+	}
+}

### 2009-07-01 main/java/chord/rels/RelInitM.java

@@ -1,0 +1,31 @@
+/*
+ * Copyright (c) 2008-2009, Intel Corporation.
+ * Copyright (c) 2006-2007, The Trustees of Stanford University.
+ * All rights reserved.
+ */
+package chord.rels;
+
+import joeq.Class.jq_Class;
+import joeq.Class.jq_Initializer;
+import joeq.Class.jq_Method;
+import chord.project.Chord;
+import chord.project.ProgramRel;
+import chord.visitors.IMethodVisitor;
+
+/**
+ * Relation containing all constructor methods, that is, methods
+ * having name <tt>&lt;init&gt;</tt>.
+ * 
+ * @author Mayur Naik (mhn@cs.stanford.edu)
+ */
+@Chord(
+	name = "initM",
+	sign = "M0"
+)
+public class RelInitM extends ProgramRel implements IMethodVisitor {
+	public void visit(jq_Class c) { }
+	public void visit(jq_Method m) {
+		if (m instanceof jq_Initializer)
+			add(m);
+	}
+}

### 2009-07-01 main/java/chord/rels/RelInstM.java

@@ -1,0 +1,29 @@
+/*
+ * Copyright (c) 2008-2009, Intel Corporation.
+ * Copyright (c) 2006-2007, The Trustees of Stanford University.
+ * All rights reserved.
+ */
+package chord.rels;
+
+import joeq.Class.jq_Class;
+import joeq.Class.jq_Method;
+import chord.project.Chord;
+import chord.project.ProgramRel;
+import chord.visitors.IMethodVisitor;
+
+/**
+ * Relation containing all instance (as opposed to static) methods.
+ *
+ * @author Mayur Naik (mhn@cs.stanford.edu)
+ */
+@Chord(
+	name = "instM",
+	sign = "M0"
+)
+public class RelInstM extends ProgramRel implements IMethodVisitor {
+	public void visit(jq_Class c) { }
+	public void visit(jq_Method m) {
+		if (!m.isStatic())
+			add(m);
+	}
+}

### 2009-07-01 main/java/chord/rels/RelME.java

@@ -1,0 +1,39 @@
+/*
+ * Copyright (c) 2008-2009, Intel Corporation.
+ * Copyright (c) 2006-2007, The Trustees of Stanford University.
+ * All rights reserved.
+ */
+package chord.rels;
+
+import joeq.Class.jq_Method;
+import joeq.Compiler.Quad.Quad;
+import chord.doms.DomE;
+import chord.doms.DomM;
+import chord.project.Chord;
+import chord.project.Program;
+import chord.project.ProgramRel;
+
+/**
+ * Relation containing each tuple (m,e) such that method m contains
+ * statement e that accesses (reads or writes) an instance field, a
+ * static field, or an array element.
+ *
+ * @author Mayur Naik (mhn@cs.stanford.edu)
+ */
+@Chord(
+	name = "ME",
+	sign = "M0,E0:M0_E0"
+)
+public class RelME extends ProgramRel {
+	public void fill() {
+		DomM domM = (DomM) doms[0];
+		DomE domE = (DomE) doms[1];
+		int numE = domE.size();
+		for (int hIdx = 0; hIdx < numE; hIdx++) {
+			Quad q = domE.get(hIdx);
+			jq_Method m = Program.getMethod(q);
+			int mIdx = domM.get(m);
+			add(mIdx, hIdx);
+		}
+	}
+}

### 2009-07-01 main/java/chord/rels/RelMS.java

@@ -1,0 +1,31 @@
+/*
+ * Copyright (c) 2008-2009, Intel Corporation.
+ * Copyright (c) 2006-2007, The Trustees of Stanford University.
+ * All rights reserved.
+ */
+package chord.rels;
+
+import joeq.Class.jq_Class;
+import joeq.Class.jq_Method;
+import chord.project.Chord;
+import chord.project.Program;
+import chord.project.ProgramRel;
+import chord.visitors.IMethodVisitor;
+
+/**
+ * Relation containing each tuple (m,s) such that method m has
+ * signature s.
+ *
+ * @author Mayur Naik (mhn@cs.stanford.edu)
+ */
+@Chord(
+	name = "MS",
+	sign = "M0,S0:S0_M0"
+)
+public class RelMS extends ProgramRel implements IMethodVisitor {
+	public void visit(jq_Class c) { }
+	public void visit(jq_Method m) {
+		String s = Program.getSign(m);
+		add(m, s);
+	}
+}

### 2009-07-01 main/java/chord/rels/RelMgetStatFldInst.java

@@ -1,0 +1,52 @@
+/*
+ * Copyright (c) 2008-2009, Intel Corporation.
+ * Copyright (c) 2006-2007, The Trustees of Stanford University.
+ * All rights reserved.
+ */
+package chord.rels;
+
+import joeq.Class.jq_Class;
+import joeq.Class.jq_Field;
+import joeq.Class.jq_Method;
+import joeq.Compiler.Quad.Operator;
+import joeq.Compiler.Quad.Quad;
+import joeq.Compiler.Quad.Operand.RegisterOperand;
+import joeq.Compiler.Quad.Operator.Getstatic;
+import joeq.Compiler.Quad.RegisterFactory.Register;
+import chord.project.Chord;
+import chord.project.ProgramRel;
+import chord.visitors.IHeapInstVisitor;
+
+/**
+ * Relation containing each tuple (m,v,f) such that method m contains
+ * a statement of the form <tt>v = f</tt>.
+ *
+ * @author Mayur Naik (mhn@cs.stanford.edu)
+ */
+@Chord(
+	name = "MgetStatFldInst",
+	sign = "M0,V0,F0:F0_M0_V0"
+)
+public class RelMgetStatFldInst extends ProgramRel
+		implements IHeapInstVisitor {
+	private jq_Method ctnrMethod;
+	public void visit(jq_Class c) { }
+	public void visit(jq_Method m) {
+		ctnrMethod = m;
+	}
+	public void visitHeapInst(Quad q) {
+		Operator op = q.getOperator();
+		if (op instanceof Getstatic) {
+			jq_Field f = Getstatic.getField(q).getField();
+			if (f.getType().isReferenceType()) {
+				RegisterOperand lo = Getstatic.getDest(q);
+				Register l = lo.getRegister();
+				try {
+					add(ctnrMethod, l, f);
+				} catch (RuntimeException ex) {
+					// TODO
+				}
+			}
+		}
+	}
+}

### 2009-07-01 main/java/chord/rels/RelMmethRet.java

@@ -1,0 +1,49 @@
+/*
+ * Copyright (c) 2008-2009, Intel Corporation.
+ * Copyright (c) 2006-2007, The Trustees of Stanford University.
+ * All rights reserved.
+ */
+package chord.rels;
+
+import joeq.Class.jq_Class;
+import joeq.Class.jq_Method;
+import joeq.Compiler.Quad.Operand;
+import joeq.Compiler.Quad.Quad;
+import joeq.Compiler.Quad.Operand.RegisterOperand;
+import joeq.Compiler.Quad.Operator.Return;
+import joeq.Compiler.Quad.RegisterFactory.Register;
+
+import chord.project.Chord;
+import chord.project.ProgramRel;
+import chord.visitors.IReturnInstVisitor;
+
+/**
+ * Relation containing each tuple (m,z,v) such that local variable
+ * v is the zth return variable of method m.
+ *
+ * @author Mayur Naik (mhn@cs.stanford.edu)
+ */
+@Chord(
+	name = "MmethRet",
+	sign = "M0,Z0,V1:M0_V1_Z0"
+)
+public class RelMmethRet extends ProgramRel
+		implements IReturnInstVisitor {
+	private static Integer ZERO = new Integer(0);
+	private jq_Method ctnrMethod;
+	public void visit(jq_Class c) { }
+	public void visit(jq_Method m) {
+		ctnrMethod = m;
+	}
+	public void visitReturnInst(Quad q) {
+		Operand rx = Return.getSrc(q);
+		// note: rx is null if this method returns void
+		if (rx instanceof RegisterOperand) {
+			RegisterOperand ro = (RegisterOperand) rx;
+			if (ro.getType().isReferenceType()) {
+				Register v = ro.getRegister();
+				add(ctnrMethod, ZERO, v);
+			}
+		}
+	}
+}

### 2009-07-01 main/java/chord/rels/RelMputStatFldInst.java

@@ -1,0 +1,56 @@
+/*
+ * Copyright (c) 2008-2009, Intel Corporation.
+ * Copyright (c) 2006-2007, The Trustees of Stanford University.
+ * All rights reserved.
+ */
+package chord.rels;
+
+import joeq.Class.jq_Class;
+import joeq.Class.jq_Field;
+import joeq.Class.jq_Method;
+import joeq.Compiler.Quad.Operand;
+import joeq.Compiler.Quad.Operator;
+import joeq.Compiler.Quad.Quad;
+import joeq.Compiler.Quad.Operand.RegisterOperand;
+import joeq.Compiler.Quad.Operator.Putstatic;
+import joeq.Compiler.Quad.RegisterFactory.Register;
+import chord.project.Chord;
+import chord.project.ProgramRel;
+import chord.visitors.IHeapInstVisitor;
+
+/**
+ * Relation containing each tuple (m,f,v) such that method m contains
+ * a statement of the form <tt>f = v</tt>.
+ *
+ * @author Mayur Naik (mhn@cs.stanford.edu)
+ */
+@Chord(
+	name = "MputStatFldInst",
+	sign = "M0,F0,V0:F0_M0_V0"
+)
+public class RelMputStatFldInst extends ProgramRel
+		implements IHeapInstVisitor {
+	private jq_Method ctnrMethod;
+	public void visit(jq_Class c) { }
+	public void visit(jq_Method m) {
+		ctnrMethod = m;
+	}
+	public void visitHeapInst(Quad q) {
+		Operator op = q.getOperator();
+		if (op instanceof Putstatic) {
+			jq_Field f = Putstatic.getField(q).getField();
+			if (f.getType().isReferenceType()) {
+				Operand rx = Putstatic.getSrc(q);
+				if (rx instanceof RegisterOperand) {
+					RegisterOperand ro = (RegisterOperand) rx;
+					Register r = ro.getRegister();
+					try {
+						add(ctnrMethod, f, r);
+					} catch (RuntimeException ex) {
+						// TODO
+					}
+				}
+			}
+		}
+	}
+}

### 2009-07-01 main/java/chord/rels/RelPrivateM.java

@@ -1,0 +1,31 @@
+/*
+ * Copyright (c) 2008-2009, Intel Corporation.
+ * Copyright (c) 2006-2007, The Trustees of Stanford University.
+ * All rights reserved.
+ */
+package chord.rels;
+
+import joeq.Class.jq_Class;
+import joeq.Class.jq_Method;
+import chord.project.Chord;
+import chord.project.ProgramRel;
+import chord.visitors.IMethodVisitor;
+
+/**
+ * Relation containing all private (as opposed to protected or
+ * public) methods.
+ *
+ * @author Mayur Naik (mhn@cs.stanford.edu)
+ */
+@Chord(
+	name = "privateM",
+	sign = "M0"
+)
+public class RelPrivateM extends ProgramRel
+		implements IMethodVisitor {
+	public void visit(jq_Class c) { }
+	public void visit(jq_Method m) {
+		if (m.isPrivate())
+			add(m);
+	}
+}

### 2009-07-01 main/java/chord/rels/RelSpecIM.java

@@ -1,0 +1,45 @@
+/*
+ * Copyright (c) 2008-2009, Intel Corporation.
+ * Copyright (c) 2006-2007, The Trustees of Stanford University.
+ * All rights reserved.
+ */
+package chord.rels;
+
+import joeq.Class.jq_Method;
+import joeq.Compiler.Quad.Operator;
+import joeq.Compiler.Quad.Quad;
+import joeq.Compiler.Quad.Operator.Invoke.InvokeStatic;
+import chord.doms.DomI;
+import chord.doms.DomM;
+import chord.project.Chord;
+import chord.project.ProgramRel;
+
+/**
+ * Relation containing each tuple (i,m) such that m is the resolved
+ * method of method invocation statement i of kind <tt>INVK_SPECIAL</tt>.
+ *
+ * @author Mayur Naik (mhn@cs.stanford.edu)
+ */
+@Chord(
+	name = "specIM",
+	sign = "I0,M0:I0xM0"
+)
+public class RelSpecIM extends ProgramRel {
+	public void fill() {
+		DomI domI = (DomI) doms[0];
+		DomM domM = (DomM) doms[1];
+		int numI = domI.size();
+		for (int iIdx = 0; iIdx < numI; iIdx++) {
+			Quad i = domI.get(iIdx);
+			Operator op = i.getOperator();
+			if (op instanceof InvokeStatic) {
+				jq_Method m = InvokeStatic.getMethod(i).getMethod();
+				if (!m.isStatic()) {
+					int mIdx = domM.get(m);
+					if (mIdx != -1)
+						add(iIdx, mIdx);
+				}
+			}
+		}
+	}
+}

### 2009-07-01 main/java/chord/rels/RelStatFldE.java

@@ -1,0 +1,38 @@
+/*
+ * Copyright (c) 2008-2009, Intel Corporation.
+ * Copyright (c) 2006-2007, The Trustees of Stanford University.
+ * All rights reserved.
+ */
+package chord.rels;
+
+import joeq.Compiler.Quad.Operator;
+import joeq.Compiler.Quad.Quad;
+import joeq.Compiler.Quad.Operator.Getstatic;
+import joeq.Compiler.Quad.Operator.Putstatic;
+import chord.doms.DomE;
+import chord.project.Chord;
+import chord.project.ProgramRel;
+
+/**
+ * Relation containing each statement that accesses (reads or writes)
+ * a static field (as opposed to an instance field or array element).
+ *
+ * @author Mayur Naik (mhn@cs.stanford.edu)
+ */
+@Chord(
+	name = "statFldE",
+	sign = "E0"
+)
+public class RelStatFldE extends ProgramRel {
+	public void fill() {
+		DomE domE = (DomE) doms[0];
+		int numE = domE.size();
+		for (int eIdx = 0; eIdx < numE; eIdx++) {
+			Quad e = domE.get(eIdx);
+			Operator op = e.getOperator();
+			if (op instanceof Getstatic || op instanceof Putstatic) {
+				add(eIdx);
+			}
+		}
+	}
+}

### 2009-07-01 main/java/chord/rels/RelStaticTM.java

@@ -1,0 +1,33 @@
+/*
+ * Copyright (c) 2008-2009, Intel Corporation.
+ * Copyright (c) 2006-2007, The Trustees of Stanford University.
+ * All rights reserved.
+ */
+package chord.rels;
+
+import joeq.Class.jq_Class;
+import joeq.Class.jq_Method;
+import chord.project.Chord;
+import chord.project.ProgramRel;
+import chord.visitors.IMethodVisitor;
+
+/**
+ * Relation containing each tuple (t,m) such that m is a
+ * static method defined in type t.
+ * 
+ * @author Mayur Naik (mhn@cs.stanford.edu)
+ */
+@Chord(
+	name = "staticTM",
+	sign = "T0,M0:M0_T0"
+)
+public class RelStaticTM extends ProgramRel
+		implements IMethodVisitor {
+	public void visit(jq_Class c) { }
+	public void visit(jq_Method m) {
+		if (m.isStatic()) {
+			jq_Class t = m.getDeclaringClass();
+			add(t, m);
+		}
+	}
+}

### 2009-07-01 main/java/chord/rels/RelSyncLM.java

@@ -1,0 +1,41 @@
+/*
+ * Copyright (c) 2008-2009, Intel Corporation.
+ * Copyright (c) 2006-2007, The Trustees of Stanford University.
+ * All rights reserved.
+ */
+package chord.rels;
+
+import joeq.Class.jq_Method;
+import joeq.Compiler.Quad.Inst;
+import joeq.Compiler.Quad.Quad;
+import chord.doms.DomL;
+import chord.doms.DomM;
+import chord.project.Chord;
+import chord.project.Program;
+import chord.project.ProgramRel;
+
+/**
+ * Relation containing each tuple (l,v) such that monitorenter
+ * statement l is synchronized on variable v.
+ *
+ * @author Mayur Naik (mhn@cs.stanford.edu)
+ */
+@Chord(
+	name = "syncLM",
+	sign = "L0,M0:L0_M0"
+)
+public class RelSyncLM extends ProgramRel {
+	public void fill() {
+		DomL domL = (DomL) doms[0];
+		DomM domM = (DomM) doms[1];
+		int numL = domL.size();
+		for (int lIdx = 0; lIdx < numL; lIdx++) {
+			Inst i = domL.get(lIdx);
+			if (!(i instanceof Quad)) {
+				jq_Method m = Program.getMethod(i);
+				int mIdx = domM.get(m);
+				add(lIdx, mIdx);
+			}
+		}
+	}
+}

### 2009-07-01 main/java/chord/rels/RelSyncM.java

@@ -1,0 +1,29 @@
+/*
+ * Copyright (c) 2008-2009, Intel Corporation.
+ * Copyright (c) 2006-2007, The Trustees of Stanford University.
+ * All rights reserved.
+ */
+package chord.rels;
+
+import joeq.Class.jq_Class;
+import joeq.Class.jq_Method;
+import chord.project.Chord;
+import chord.project.ProgramRel;
+import chord.visitors.IMethodVisitor;
+
+/**
+ * Relation containing all synchronized methods.
+ *
+ * @author Mayur Naik (mhn@cs.stanford.edu)
+ */
+@Chord(
+	name = "syncM",
+	sign = "M0"
+)
+public class RelSyncM extends ProgramRel implements IMethodVisitor {
+	public void visit(jq_Class c) { }
+	public void visit(jq_Method m) {
+		if (m.isSynchronized())
+			add(m);
+	}
+}

### 2009-07-01 main/java/chord/rels/RelTM.java

@@ -1,0 +1,32 @@
+/*
+ * Copyright (c) 2008-2009, Intel Corporation.
+ * Copyright (c) 2006-2007, The Trustees of Stanford University.
+ * All rights reserved.
+ */
+package chord.rels;
+
+import joeq.Class.jq_Class;
+import joeq.Class.jq_Method;
+import chord.project.Chord;
+import chord.project.ProgramRel;
+import chord.visitors.IMethodVisitor;
+
+/**
+ * Relation containing each tuple (t,m) such that method m is
+ * declared in type t.
+ *
+ * @author Mayur Naik (mhn@cs.stanford.edu)
+ */
+@Chord(
+	name = "TM",
+	sign = "T0,M0:M0_T0"
+)
+public class RelTM extends ProgramRel implements IMethodVisitor {
+	private jq_Class ctnrClass;
+	public void visit(jq_Class c) {
+		ctnrClass = c;
+	}
+	public void visit(jq_Method m) {
+		add(ctnrClass, m);
+	}
+}

### 2009-07-01 main/java/chord/rels/RelThisMV.java

@@ -1,0 +1,47 @@
+/*
+ * Copyright (c) 2008-2009, Intel Corporation.
+ * Copyright (c) 2006-2007, The Trustees of Stanford University.
+ * All rights reserved.
+ */
+package chord.rels;
+
+import joeq.Class.jq_Method;
+import joeq.Compiler.Quad.ControlFlowGraph;
+import joeq.Compiler.Quad.RegisterFactory;
+import joeq.Compiler.Quad.RegisterFactory.Register;
+
+import chord.doms.DomM;
+import chord.doms.DomV;
+import chord.project.Program;
+import chord.project.Chord;
+import chord.project.ProgramRel;
+
+/**
+ * Relation containing each tuple (m,z,v) such that local variable
+ * v is the zth argument variable of method m.
+ *
+ * @author Mayur Naik (mhn@cs.stanford.edu)
+ */
+@Chord(
+	name = "thisMV",
+	sign = "M0,V0:M0_V0"
+)
+public class RelThisMV extends ProgramRel {
+	public void fill() {
+		DomM domM = (DomM) doms[0];
+		DomV domV = (DomV) doms[1];
+		int numM = domM.size();
+		for (int mIdx = 0; mIdx < numM; mIdx++) {
+			jq_Method m = domM.get(mIdx);
+			if (!m.isStatic()) {
+				ControlFlowGraph cfg = Program.getCFG(m);
+				if (cfg != null) {
+					RegisterFactory rf = cfg.getRegisterFactory();
+					Register v = rf.get(0);
+					int vIdx = domV.get(v);
+					add(mIdx, vIdx);
+				}
+			}
+		}
+	}
+}

### 2009-07-01 main/java/chord/visitors/IPhiInstVisitor.java

@@ -1,0 +1,22 @@
+/*
+ * Copyright (c) 2008-2009, Intel Corporation.
+ * Copyright (c) 2006-2007, The Trustees of Stanford University.
+ * All rights reserved.
+ */
+package chord.visitors;
+
+import joeq.Compiler.Quad.Quad;
+
+/**
+ * Visitor over all phi statements in all methods in the program.
+ * 
+ * @author Mayur Naik (mayur.naik@intel.com)
+ */
+public interface IPhiInstVisitor extends IMethodVisitor {
+	/**
+	 * Visits all phi statements in all methods in the program.
+	 * 
+	 * @param	q	A copy assignment statement.
+	 */
+	public void visitPhiInst(Quad q);
+}

### 2009-07-01 main/java/chord/visitors/IVarVisitor.java

@@ -1,0 +1,22 @@
+/*
+ * Copyright (c) 2008-2009, Intel Corporation.
+ * Copyright (c) 2006-2007, The Trustees of Stanford University.
+ * All rights reserved.
+ */
+package chord.visitors;
+
+import joeq.Compiler.Quad.RegisterFactory.Register;
+
+/**
+ * Visitor over all local variables of all methods in the program.
+ * 
+ * @author Mayur Naik (mayur.naik@intel.com)
+ */
+public interface IVarVisitor extends IMethodVisitor {
+	/**
+	 * Visits all local variables of all methods in the program.
+	 * 
+	 * @param	v	A local variable.
+	 */
+	public void visit(Register v);
+}

### 2009-07-01 util/java/chord/util/ArraySet.java

@@ -1,0 +1,81 @@
+/*
+ * Copyright (c) 2008-2009, Intel Corporation.
+ * Copyright (c) 2006-2007, The Trustees of Stanford University.
+ * All rights reserved.
+ */
+package chord.util;
+
+import java.util.ArrayList;
+import java.util.Collection;
+import java.util.Set;
+import java.util.Iterator;
+
+/**
+ * Array-based implementation of a set.
+ * 
+ * @author Mayur Naik (mhn@cs.stanford.edu)
+ */
+public class ArraySet<T> extends ArrayList<T> implements Set<T> {
+	/**
+	 * 
+	 */
+	private static final long serialVersionUID = -2510316860750967764L;
+	public ArraySet(Collection<T> s) {
+		super(s);
+	}
+	public ArraySet() { }
+	public ArraySet(int initialCapacity) {
+		super(initialCapacity);
+	}
+	public boolean add(T e) {
+		if (contains(e)) {
+			return false;
+		}
+		return super.add(e);
+	}
+	/**
+	 * Adds a given value to the set without checking if it already
+	 * exists in the set.
+	 * 
+	 * @param	e	A value to be added to the set.
+	 */
+	public void addForcibly(T e) {
+		super.add(e);
+	}
+	public boolean addAll(Collection<? extends T> c) {
+		boolean modified = false;
+		for (T e : c) {
+			if (add(e)) {
+				modified = true;
+			}
+		}
+		return modified;
+	}
+    public boolean equals(Object o) {
+   		if (o == this)
+        	return true;
+    	if (!(o instanceof Set)) {
+        	return false;
+		}
+    	Collection c = (Collection) o;
+    	if (c.size() != size()) {
+        	return false;
+		}
+		for (Iterator it = c.iterator(); it.hasNext();) {
+			Object e = it.next();
+			if (!contains(e)) {
+				return false;
+			}
+		}
+		return true;
+    }
+	public int hashCode() {
+		int h = 0;
+		for (T e : this) {
+			if (e != null)
+				h += e.hashCode();
+		}
+		return h;
+	}
+}
+

### 2009-07-01 util/java/chord/util/CollectionUtils.java

@@ -1,0 +1,52 @@
+/*
+ * Copyright (c) 2008-2009, Intel Corporation.
+ * Copyright (c) 2006-2007, The Trustees of Stanford University.
+ * All rights reserved.
+ */
+package chord.util;
+
+import java.util.Collection;
+import java.util.Iterator;
+import java.util.List;
+
+/**
+ * Collection related utilities.
+ * 
+ * @author Mayur Naik (mhn@cs.stanford.edu)
+ */
+public class CollectionUtils {
+	/**
+	 * Determines whether a given collection contains duplicate
+	 * values.
+	 * 
+	 * @param	<T>	The type of the collection elements.
+	 * @param	elems	A collection.
+	 * 
+	 * @return	true iff the given collection contains duplicate
+	 * 			values.
+	 */
+	public static <T> boolean hasDuplicates(List<T> elems) {
+		for (int i = 0; i < elems.size() - 1; i++) {
+			T elem = elems.get(i);
+			for (int j = i + 1; j < elems.size(); j++) {
+				if (elems.get(j).equals(elem))
+					return true;
+			}
+		}
+		return false;
+	}
+	public static <T> String toString(Collection<T> a,
+			String begStr, String sepStr, String endStr) {
+		int n = a.size();
+		if (n == 0)
+			return begStr + endStr;
+		Iterator<T> it = a.iterator();
+		String s = begStr + it.next();
+		while (it.hasNext())
+			s += sepStr + it.next();
+		return s + endStr;
+	}
+	public static <T> String toString(Collection<T> a) {
+		return toString(a, "", ",", "");
+	}
+}

### 2009-07-01 util/java/chord/util/CompareUtils.java

@@ -1,0 +1,17 @@
+/*
+ * Copyright (c) 2008-2009, Intel Corporation.
+ * Copyright (c) 2006-2007, The Trustees of Stanford University.
+ * All rights reserved.
+ */
+package chord.util;
+
+/**
+ * Object comparison utilities.
+ * 
+ * @author Mayur Naik (mhn@cs.stanford.edu)
+ */
+public class CompareUtils {
+	public static boolean areEqual(Object a, Object b) {
+		return a == null ? b == null : a.equals(b);
+	}
+}

### 2009-07-01 util/java/chord/util/FileUtils.java

@@ -1,0 +1,155 @@
+/*
+ * Copyright (c) 2008-2009, Intel Corporation.
+ * Copyright (c) 2006-2007, The Trustees of Stanford University.
+ * All rights reserved.
+ */
+package chord.util;
+
+import java.io.File;
+import java.io.FileInputStream;
+import java.io.FileNotFoundException;
+import java.io.FileOutputStream;
+import java.io.ObjectOutputStream;
+import java.io.ObjectInputStream;
+import java.io.PrintWriter;
+import java.io.BufferedReader;
+import java.io.FileReader;
+import java.util.List;
+import java.util.ArrayList;
+
+/**
+ * File related utilities.
+ *  
+ * @author Mayur Naik (mhn@cs.stanford.edu)
+ */
+public class FileUtils {
+	public static String getAbsolutePath(String fileName, String baseName) {
+		File file = new File(fileName);
+		if (file.isAbsolute())
+			return fileName;
+		file = new File(baseName, fileName);
+		return file.getAbsolutePath();
+	}
+	public static void copy(String fromFileName, String toFileName) {
+		try {
+			FileInputStream fis = new FileInputStream(fromFileName);
+			FileOutputStream fos = new FileOutputStream(toFileName);
+			byte[] buf = new byte[1024];
+			int i = 0;
+			while((i = fis.read(buf)) != -1) {
+				fos.write(buf, 0, i);
+			}
+			fis.close();
+			fos.close();
+		} catch (Exception e) {
+			throw new RuntimeException(e.getMessage(), e);
+		}
+	}
+	public static boolean mkdir(String dirName) {
+		return mkdir(new File(dirName));
+	}
+	public static boolean mkdir(String parentName, String childName) {
+		return mkdir(new File(parentName, childName));
+	}
+	public static boolean mkdir(File file) {
+		if (file.exists()) {
+			if (!file.isDirectory()) {
+				throw new RuntimeException(
+					"File '" + file + "' is not a directory.");
+			}
+			return false;
+		}
+		if (file.mkdirs())
+			return true;
+		throw new RuntimeException("Failed to create directory '" +
+			file + "'"); 
+	}
+	public static PrintWriter newPrintWriter(String fileName) {
+		PrintWriter out;
+		try {
+			out = new PrintWriter(fileName);
+		} catch (FileNotFoundException ex) {
+			throw new RuntimeException(ex);
+		}
+		return out;
+	}
[…]

### 2009-07-01 util/java/chord/util/ProcessExecutor.java

@@ -1,0 +1,67 @@
+/*
+ * Copyright (c) 2008-2009, Intel Corporation.
+ * Copyright (c) 2006-2007, The Trustees of Stanford University.
+ * All rights reserved.
+ */
+package chord.util;
+
+import java.io.BufferedReader;
+import java.io.IOException;
+import java.io.InputStream;
+import java.io.InputStreamReader;
+
+/**
+ * Utility to execute a system command specified as a string in a
+ * separate process.
+ * 
+ * @author Mayur Naik (mhn@cs.stanford.edu)
+ */
+public final class ProcessExecutor {
+	/**
+	 * Executes a given system command specified as a string in a
+	 * separate process.
+	 * <p>
+	 * The invoking process waits till the invoked process finishes.
+	 * 
+	 * @param	A system command to be executed.
+	 * 
+	 * @return	The exit value of the invoked process.
+	 * 			By convention, 0 indicates normal termination.
+	 */
+    public static final int execute(String cmd) {
+		System.out.println("EXEC " + cmd);
+        try {
+	    	Process proc = Runtime.getRuntime().exec(cmd);
+	    	StreamGobbler err =
+	    		new StreamGobbler(proc.getErrorStream(), "ERR");
+	        StreamGobbler out = 
+	        	new StreamGobbler(proc.getInputStream(), "OUT");
+	        err.start();
+	        out.start();
+	        return proc.waitFor();
+        } catch (Exception e) {
+        	throw new RuntimeException(e.getMessage());
+        }
+        
+    }
+    private static class StreamGobbler extends Thread {
+        private final InputStream s;
+        private final String n;
+        private StreamGobbler(InputStream s, String n) {
+            this.s = s;
+            this.n = n;
+        }
+        public void run() {
+            try {
+                BufferedReader r =
+                	new BufferedReader(new InputStreamReader(s));
+                String l;
+                while ((l = r.readLine()) != null)
+                    System.out.println(n + "> " + l);
+            } catch (IOException e) {
+            	throw new RuntimeException(e.getMessage());
+            }
+        }
+    }
+}
+

### 2009-07-01 util/java/chord/util/SetUtils.java

@@ -1,0 +1,42 @@
+/*
+ * Copyright (c) 2008-2009, Intel Corporation.
+ * Copyright (c) 2006-2007, The Trustees of Stanford University.
+ * All rights reserved.
+ */
+package chord.util;
+
+import java.util.HashSet;
+import java.util.Set;
+
+/**
+ * Set related utilities.
+ * 
+ * @author Mayur Naik (mhn@cs.stanford.edu)
+ */
+public class SetUtils {
+	public static final int DEFAULT_INIT_SIZE = 10;
+	public static final int THRESHOLD = 10;
+	public static <E> Set<E> newSet() {
+		return newSet(DEFAULT_INIT_SIZE);
+	}
+	public static <E> Set<E> newSet(int initSize) {
+		Set<E> set;
+		if (initSize < THRESHOLD) {
+			set = new ArraySet<E>(initSize);
+		} else {
+			set = new HashSet<E>(initSize);
+		}
+		return set;
+	}
+	public static <E> Set<E> newSet(Set<E> c) {
+		Set<E> set = newSet(c.size());
+		set.addAll(c);
+		return set;
+	}
+    public static <E> Set<E> iterableToSet(Iterable<E> c, int size) {
+		Set<E> set = newSet(size);
+        for (E e : c)
+            set.add(e);
+        return set;
+    }
+}

### 2009-07-01 util/java/chord/util/graph/GraphEntityVisitorException.java

@@ -1,0 +1,17 @@
+/*
+ * Copyright (c) 2008-2009, Intel Corporation.
+ * Copyright (c) 2006-2007, The Trustees of Stanford University.
+ * All rights reserved.
+ */
+package chord.util.graph;
+
+/**
+ * Exception that may be thrown by a client of
+ * {@link IGraphEntityVisitor} to terminate the visitor's traversal.
+ * 
+ * @author Mayur Naik (mhn@cs.stanford.edu)
+ */
+
+public class GraphEntityVisitorException extends RuntimeException {
+
+}

### 2009-07-01 util/java/chord/util/graph/IGraph.java

@@ -1,0 +1,198 @@
+/*
+ * Copyright (c) 2008-2009, Intel Corporation.
+ * Copyright (c) 2006-2007, The Trustees of Stanford University.
+ * All rights reserved.
+ */
+package chord.util.graph;
+
+import java.util.List;
+import java.util.Set;
+
+import chord.util.IndexMap;
+import chord.util.tuple.object.Pair;
+
+/**
+ * Specification of an immutable, unlabeled, directed graph along
+ * with useful operations on it.
+ * <p>
+ * This interface must be used when it is desirable to provide an
+ * immutable view of an unlabeled, directed graph to clients even
+ * though the underlying implementation may be that of a mutable
+ * graph.
+ * <p>
+ * Classes implementing this interface are:
+ * <ul>
+ * <li>{@link MutableGraph}, a complete implementation that provides
+ *     both, the useful operations and a representation of the
+ *     graph.</li>
+ * <li>{@link AbstractGraph}, a partial implementation that provides
+ *     the useful operations but leaves the representation of the
+ *     graph unspecified.</li>
+ * </ul>
+ * 
+ * @author Mayur Naik (mhn@cs.stanford.edu)
+ *
+ * @param	<Node>	The type of the graph's nodes.
+ */
+public interface IGraph<Node> extends java.io.Serializable {
+	/**
+	 * Determines whether this graph contains a given node as a
+	 * root node.
+	 * 
+	 * @param	node	A node.
+	 * 
+	 * @return	true if this graph contains the given node as a
+	 * 			root node.
+	 */
+	public boolean hasRoot(Node node);
+	/**
+	 * Determines whether this graph contains a given node.
+	 *
+	 * @param	node	A node.
+	 *
+	 * @return	true if this graph contains the given node.
+	 */
+	public boolean hasNode(Node node);
+	/**
+	 * Determines whether this graph contains a given directed edge.
+	 * 
+	 * @param	node1	A node.
+	 * @param	node2	A node.
+	 * 
+	 * @return	true if this graph contains a directed edge from
+	 * 			<tt>node1</tt> to <tt>node2</tt>.
+	 */
+	public boolean hasEdge(Node node1, Node node2);
+	/**
+	 * Provides the total number of roots of this graph.
+	 *
+	 * @return	The total number of roots of this graph.
+	 */
+	public int numRoots();
+	/**
+	 * Provides the total number of nodes in this graph.
[…]

### 2009-07-01 util/java/chord/util/graph/SimpleCyclesFinder.java

@@ -1,0 +1,113 @@
+/*
+ * Copyright (c) 2008-2009, Intel Corporation.
+ * Copyright (c) 2006-2007, The Trustees of Stanford University.
+ * All rights reserved.
+ */
+package chord.util.graph;
+
+import java.util.HashMap;
+import java.util.Iterator;
+import java.util.Map;
+import java.util.Set;
+
+import gnu.trove.TIntStack;
+import chord.util.ArraySet;
+import chord.util.IndexMap;
+
+/**
+ * Algorithm for finding all simple cycles in a directed graph.
+ *  
+ * @author Mayur Naik (mhn@cs.stanford.edu)
+ *
+ * @param	<Node>	The type of the graph's nodes.
+ */
+public class SimpleCyclesFinder<Node> {
+	private final IGraph<Node> graph;
+	private final IGraphEntityVisitor<Node> visitor;
+	private Map<Node, Set<Node>> nodeToSuccsMap; 
+	private IndexMap<Node> nodeIdxMap;
+	private int currNodeIdx;
+	private TIntStack markedStack;
+	private TIntStack pointStack;
+	private boolean[] mark;
+	/**
+	 * Finds all simple cycles in a given directed graph by
+     * invoking a given visitor on each simple cycle.
+	 * 
+	 * @param	<Node>	The type of the graph's nodes.
+	 * 
+	 * @param	graph	A directed graph.
+	 * @param	visitor	A visitor over the graph's simple cycles.
+	 */
+	public static <Node> void run(IGraph<Node> graph,
+			IGraphEntityVisitor<Node> visitor) {
+		(new SimpleCyclesFinder<Node>(graph, visitor)).run();
+	}
+	private SimpleCyclesFinder(IGraph<Node> graph,
+			IGraphEntityVisitor<Node> visitor) {
+		this.graph = graph;
+		this.visitor = visitor;
+	}
+	private void run() {
+		markedStack = new TIntStack();
+		pointStack = new TIntStack();
+		nodeToSuccsMap = new HashMap<Node, Set<Node>>();
+		nodeIdxMap = new IndexMap<Node>();
+		int numNodes = 0;
+		for (Node node : graph.getNodes()) {
+			numNodes++;
+			Set<Node> succs = new ArraySet<Node>();
+			nodeToSuccsMap.put(node, succs);
+			for (Node node2 : graph.getSuccs(node)) {
[…]

### 2009-07-01 util/java/chord/util/tuple/integer/IntHext.java

@@ -1,0 +1,67 @@
+/*
+ * Copyright (c) 2008-2009, Intel Corporation.
+ * Copyright (c) 2006-2007, The Trustees of Stanford University.
+ * All rights reserved.
+ */
+package chord.util.tuple.integer;
+
+/**
+ * An ordered 6-tuple of integers.
+ * 
+ * @author Mayur Naik (mhn@cs.stanford.edu)
+ */
+public class IntHext implements java.io.Serializable {
+	/**
+	 * 
+	 */
+	private static final long serialVersionUID = -7833966053817524146L;
+	/**
+	 * The 1st integer in the ordered 6-tuple.
+	 */
+	public int idx0;
+	/**
+	 * The 2nd integer in the ordered 6-tuple.
+	 */
+	public int idx1;
+	/**
+	 * The 3rd integer in the ordered 6-tuple.
+	 */
+	public int idx2;
+	/**
+	 * The 4th integer in the ordered 6-tuple.
+	 */
+	public int idx3;
+	/**
+	 * The 5th integer in the ordered 6-tuple.
+	 */
+	public int idx4;
+	/**
+	 * The 6th integer in the ordered 6-tuple.
+	 */
+	public int idx5;
+	public IntHext(int idx0, int idx1, int idx2,
+				   int idx3, int idx4, int idx5) {
+		this.idx0 = idx0;
+		this.idx1 = idx1;
+		this.idx2 = idx2;
+		this.idx3 = idx3;
+		this.idx4 = idx4;
+		this.idx5 = idx5;
+	}
+	public boolean equals(Object o) {
+		if (o instanceof IntHext) {
+			IntHext that = (IntHext) o;
+			return that.idx0 == idx0 && that.idx1 == idx1 &&
+				   that.idx2 == idx2 && that.idx3 == idx3 &&
+				   that.idx4 == idx4 && that.idx5 == idx5;
+		}
+		return false;
+	}
+	public int hashCode() {
+		return idx0 + idx1 + idx2 + idx3 + idx4 + idx5;
+	}
+	public String toString() {
+        return "<" + idx0 + ", " + idx1 + ", " + idx2 + ", " +
+					 idx3 + ", " + idx4 + ", " + idx5 + ">";
+	}
+}

### 2009-07-01 util/java/chord/util/tuple/integer/IntPair.java

@@ -1,0 +1,43 @@
+/*
+ * Copyright (c) 2008-2009, Intel Corporation.
+ * Copyright (c) 2006-2007, The Trustees of Stanford University.
+ * All rights reserved.
+ */
+package chord.util.tuple.integer;
+
+/**
+ * An ordered 2-tuple of integers.
+ * 
+ * @author Mayur Naik (mhn@cs.stanford.edu)
+ */
+public class IntPair implements java.io.Serializable {
+	/**
+	 * 
+	 */
+	private static final long serialVersionUID = 6950035759159189325L;
+	/**
+	 * The 0th integer in the ordered 2-tuple.
+	 */
+	public int idx0;
+	/**
+	 * The 1st integer in the ordered 2-tuple.
+	 */
+	public int idx1;
+	public IntPair(int idx0, int idx1) {
+		this.idx0 = idx0;
+		this.idx1 = idx1;
+	}
+	public boolean equals(Object o) {
+		if (o instanceof IntPair) {
+			IntPair that = (IntPair) o;
+			return this.idx0 == that.idx0 && this.idx1 == that.idx1;
+		}
+		return false;
+	}
+	public int hashCode() {
+		return idx0 + idx1;
+	}
+	public String toString() {
+		return "fst: " + idx0 + " snd: " + idx1;
+	}
+}

### 2009-07-01 util/java/chord/util/tuple/integer/IntQuad.java

@@ -1,0 +1,55 @@
+/*
+ * Copyright (c) 2008-2009, Intel Corporation.
+ * Copyright (c) 2006-2007, The Trustees of Stanford University.
+ * All rights reserved.
+ */
+package chord.util.tuple.integer;
+
+/**
+ * An ordered 4-tuple of integers.
+ * 
+ * @author Mayur Naik (mhn@cs.stanford.edu)
+ */
+public class IntQuad implements java.io.Serializable {
+	/**
+	 * 
+	 */
+	private static final long serialVersionUID = 8505832814282826196L;
+	/**
+	 * The 0th integer in the ordered 4-tuple.
+	 */
+	public int idx0;
+	/**
+	 * The 1st integer in the ordered 4-tuple.
+	 */
+	public int idx1;
+	/**
+	 * The 2nd integer in the ordered 4-tuple.
+	 */
+	public int idx2;
+	/**
+	 * The 3rd integer in the ordered 4-tuple.
+	 */
+	public int idx3;
+	public IntQuad(int idx0, int idx1, int idx2, int idx3) {
+		this.idx0 = idx0;
+		this.idx1 = idx1;
+		this.idx2 = idx2;
+		this.idx3 = idx3;
+	}
+	public boolean equals(Object o) {
+		if (o instanceof IntQuad) {
+			IntQuad that = (IntQuad) o;
+			return that.idx0 == idx0 && that.idx1 == idx1 &&
+				   that.idx2 == idx2 && that.idx3 == idx3;
+		}
+		return false;
+	}
+	public int hashCode() {
+		return idx0 + idx1 + idx2 + idx3;
+	}
+	public String toString() {
+        return "<" + idx0 + ", " + idx1 + ", " +
+        			 idx2 + ", " + idx3 + ">";
+	}
+}

### 2009-07-01 util/java/chord/util/tuple/integer/IntTrio.java

@@ -1,0 +1,49 @@
+/*
+ * Copyright (c) 2008-2009, Intel Corporation.
+ * Copyright (c) 2006-2007, The Trustees of Stanford University.
+ * All rights reserved.
+ */
+package chord.util.tuple.integer;
+
+/**
+ * An ordered 3-tuple of integers.
+ * 
+ * @author Mayur Naik (mhn@cs.stanford.edu)
+ */
+public class IntTrio implements java.io.Serializable {
+	/**
+	 * 
+	 */
+	private static final long serialVersionUID = 1728579358137591136L;
+	/**
+	 * The 0th integer in the ordered 3-tuple.
+	 */
+	public int idx0;
+	/**
+	 * The 1st integer in the ordered 3-tuple.
+	 */
+	public int idx1;
+	/**
+	 * The 2nd integer in the ordered 3-tuple.
+	 */
+	public int idx2;
+	public IntTrio(int idx0, int idx1, int idx2) {
+		this.idx0 = idx0;
+		this.idx1 = idx1;
+		this.idx2 = idx2;
+	}
+	public boolean equals(Object o) {
+		if (o instanceof IntTrio) {
+			IntTrio that = (IntTrio) o;
+			return that.idx0 == idx0 && that.idx1 == idx1 &&
+				that.idx2 == idx2;
+		}
+		return false;
+	}
+	public int hashCode() {
+		return idx0 + idx1 + idx2;
+	}
+	public String toString() {
+        return "<" + idx0 + ", " + idx1 + ", " + idx2 + ">";
+	}
+}

### 2010-02-16 main/doc/user_guide.tex
added user guide, removed redundant property file loading in pathconvert.xml, and other minor changes

--- 2009-07-01 doc/user_guide.tex
@@
 
-%TODO: specify use of BDD's under efficiency in section 1 (overview)
-
 \htmlmathitalic{0}
 \htmltitle{User Guide}
-\htmladdress{chord@cs.stanford.edu, \today}
+\htmladdress{mayur.naik@gmail.com, \today}
 
@@
 \providecommand\Pdom{{\bf P}}
-\providecommand\Sdom{{\bf S}}
 
@@
 
-\providecommand\Chord{{Chord}}
+\providecommand\Chord{Chord}
 
@@
 
-\title{Chord: A Static Analysis Framework for Java}
-\author{Mayur Naik and Alex Aiken}
+\title{Chord: A Static and Dynamic Program Analysis Framework for Java}
+\author{Mayur Naik}
 \date{\today}
@@
 \texonly{\newpage}
-\section{Overview}
-
-\Chord\ is a static analysis framework for Java with an emphasis on the analysis of objects and concurrency.
-It has the following characteristics:
-\begin{enumerate}
-\item
-It is {\it stand-alone} in that it provides various off-the-shelf static analyses (e.g., may alias analysis, thread escape analysis, datarace analysis, deadlock analysis, etc.).
-\item
-It is {\it extensible} in that it allows users to build their own static analyses on top of the provided static analyses.
-\item
-It is {\it compositional} in that it enables different static analyses to coexist and interact with each other.
-\item
-It is {\it efficient} in that it
-caches results computed by each static analysis so that they can be reused by other static analyses without re-computation.
-\item
-It is {\it flexible} in that it allows each static analysis to be expressed either declaratively as a Datalog program or imperatively as a Java program.
-\end{enumerate}
+\section{Introduction}
+
+\Chord\ is a static and dynamic program analysis framework for Java.
+It has the following key characteristics:
+
+\begin{itemize}
+\item
+Stand-alone: various off-the-shelf program analyses are provided (e.g., may-alias, thread-escape, datarace, deadlock, etc.).
+\item
+Extensible: users can build their own program analyses on top of the provided ones.
+\item
+Compositional: each program analysis can be written independently and yet made to interact in complex ways with other program analyses.
+\item
+Efficient: results computed by each program analysis are cached for reuse by other program analyses without re-computation.
+\item
+Flexible: a broad range of program analyses can be expressed, including those
+written imperatively in Java or declaratively in Datalog, summary-based as well as cloning-based context-sensitive analyses,
+iterative refinement-based analyses, client-driven analyses, and combined static and dynamic analyses.
+\end{itemize}
+
+\Chord\ is intended for use by a broad range of users:
+
+\begin{itemize}
+\item
+Software engineers: Programmers wanting to use program analyses in Chord to assist with software development and testing.
+\item
+Program analysis appliers: Researchers with possibly a limited program analysis background wanting to build applications on top of program analyses in Chord used as black boxes.
+\item
+Program analysis writers: Program analysis researchers wanting to implement and evaluate new program analysis algorithms.
+\end{itemize}
+
+\Chord\ is intended to work on a variety of platforms, including Linux, Windows, and MacOS.
+It has been tested most extensively on Linux.
+
+\Chord\ is open source software distributed under the \xlink{New BSD License}{http://www.opensource.org/licenses/bsd-license.php}.
+Contributions from users are welcome and encouraged.
 
@@
 
-Before installing \Chord\ ensure that the following software is installed on your machine:
+Ensure that the following software is installed on your machine:
 
@@
 \item
-JDK 5 or higher (e.g., from
-Sun\footnote{\xlink{{\tt http://java.sun.com/javase/downloads/index.jsp}}{http://java.sun.com/javase/downloads/index.jsp}}
-or IBM\footnote{\xlink{{\tt http://www.ibm.com/developerworks/java/jdk/}}{http://www.ibm.com/developerworks/java/jdk/}}).
-
-\item
-Apache Ant (a Java build tool)\footnote{\xlink{{\tt http://ant.apache.org/}}{http://ant.apache.org/}}
+JDK 5 or higher, e.g. from \xlink{IBM}{http://www.ibm.com/developerworks/java/jdk/} or \xlink{Sun}{http://java.sun.com/javase/downloads/index.jsp}
+\item
+\xlink{Apache Ant}{http://ant.apache.org/}, a Java-based build tool
+\item
+a C++ compiler, e.g. \xlink{GCC}{http://gcc.gnu.org/}
+\item
+a Make utility, e.g. \xlink{GNU Make}{http://www.gnu.org/software/make/}
+\item
+\xlink{Cygwin}{http://www.cygwin.com/}, if it is Windows machine
 \end{itemize}
 
-Download the latest version of \Chord\ from \xlink{{\tt http://code.google.com/p/jchord/}}{http://code.google.com/p/jchord/}
-(say {\tt chord-x.y.tar.gz}) and uncompress it in a suitable directory by running the following command:
-
+Download the latest \Chord\ version from \xlink{here}{http://code.google.com/p/jchord/downloads/list} and uncompress it.
+
+Directory {\tt chord/main} (henceforth called the main directory) contains a {\tt build.xml} file which is interpreted by Apache Ant.
+To see the various targets available, run the following command in that directory:
+
+\begin{quote}
 \begin{verbatim}
-tar xvzf chord-x.y.tar.gz
+prompt> ant
 \end{verbatim}
-
-This will create a directory structure with root directory {\tt chord-x.y/}.
+\end{quote}
+
+To compile \Chord, run the following command in the same directory:
+
+\begin{quote}
+\begin{verbatim}
+prompt> ant compile
+\end{verbatim}
+\end{quote}
+
+This will compile the following:
+\begin{itemize}
+\item
+the Java source code of \Chord\ from {\tt chord/main/src/java/} to Java bytecode in {\tt chord/main/classes/}
+\item
+the C source code of BDD library \xlink{BuDDy}{http://buddy.sourceforge.net/} from {\tt chord/main/src/bdd/}
+to a shared library in {\tt chord/main/lib/} ({\tt libbuddy.so} on Linux, {\tt buddy.dll} on Windows, and {\tt libbuddy.dylib} on MacOS);
+this library is needed for executing program analyses written in Datalog using
+\xlink{bddbddd}{http://bddbddb.sourceforge.net/} (a BDD-based Datalog solver).
+\item
+the C++ source code of the Chord instrumentation agent from {\tt chord/main/src/agent/} to a shared
+library in {\tt chord/main/lib} ({\tt libchord\_instr\_agent.so} on Linux, {\tt chord\_instr\_agent.dll} on Windows,
+and {\tt libchord\_instr\_agent.dylib} on MacOS); this agent is needed for building program scope dynamically
+and for executing dynamic program analyses.
+\end{itemize}
 
 \texonly{\newpage}
-\section{Building Chord}
-
-The root directory {\tt chord-x.y/} contains a {\tt build.xml} file which is interpreted by
-Apache Ant (a Java build tool).
-To see the various targets available run the following command in the root directory:
-
-\begin{verbatim}
-ant -p
-\end{verbatim}
-
-The default target {\tt compile} compiles \Chord.  So simply run the following command in the
-root directory:
-
-\begin{verbatim}
-ant
-\end{verbatim}
-
-This will compile the \Chord\ {\tt .java} source files and store the {\tt .class} files
-under the {\tt classes/} sub-directory of the root directory.
-
-\texonly{\newpage}
-\section{Running Chord}
+\section{A Quick Tour}
 
@@
 
+\begin{quote}
 \begin{verbatim}
@@
                 ...
-        jars/
+        jar/
             taz.jar
-    chord/
     chord.properties
 \end{verbatim}
-
-The above structure is typical: the program's Java source files are under {\tt example/src/}, its class files
-are under {\tt example/classes/}, the source and jar files of the libraries used by the program are under
-{\tt example/lib/src/} and {\tt example/lib/jars/}, respectively, and finally {\tt example/chord/} is the directory
-to which Chord will output its intermediate and final results.
-
-Execute Chord on the above program by running the
-following command in Chord's root directory {\tt chord-x.y/}:
-
-\begin{verbatim}
-ant -Dchord.work.dir=.../example/chord run
-\end{verbatim}
-
-where ``{\tt ...}'' denotes either the absolute path, or the path relative to Chord's root directory,
-of the parent of directory {\tt example/}.
-Chord will read the system properties specified in the file named {\tt chord.properties}
-that it presumes exists in the directory specified
-by system property {\tt chord.work.dir}
-which in the above case is directory {\tt example/chord/}.
+\end{quote}
+
+The above structure is typical: the program's Java source files are under \verb+example/src/+,
+its class files are under \verb+example/classes/+, the source and jar files of the libraries
+used by the program are under \verb+example/lib/src/+ and \verb+example/lib/jar/+, respectively.
+
+File \verb+chord.properties+ specifies properties to be passed to Chord (alternatively,
+these properties may be passed on the command-line, in format \verb+-D<name>=<value>+).
 A sample such file for the above program is as follows:
 
+\begin{quote}
 \begin{verbatim}
 chord.main.class=foo.Main
-chord.class.path=../classes:../lib/classes/taz.jar
-chord.src.path=../src:../lib/src
-chord.out=log.txt
-chord.err=log.txt
-chord.serial.file=program.ser
-chord.java.analysis.path=${chord.home.dir}/classes/main:
-    ${chord.home.dir}/classes/downcast:${chord.home.dir}/src/dlog/monosite
-chord.dlog.analysis.path=${chord.home.dir}/src/dlog/main:
-    ${chord.home.dir}/src/dlog/downcast:${chord.home.dir}/src/dlog/monosite
-chord.analyses=downcast-java,monosite-java
+chord.class.path=classes:lib/jar/taz.jar
+chord.src.path=src:../lib/src
+chord.run.ids=0,1
+chord.args.0="-thread 1 -n 10"
+chord.args.1="-thread 2 -n 50"
 \end{verbatim}
-
-System property {\tt chord.home.dir} specifies Chord's root directory
-{\tt chord-x.y/}.
-Every non-absolute location used in the value of any system property defined in the file
-is taken to be relative to the directory specified by system
-property {\tt chord.work.dir}.
-
-Section \ref{sec-sys-prop} presents an exhaustive
-listing of system properties that are recognized by \Chord.
-Briefly:
+\end{quote}
+
+Each relative (instead of absolute) path element in the value of any property
+named \verb+chord.<...>.path+ set in this file is converted to an absolute path element with respect
+to the directory containing this file, which in the above case is \verb+example/+.
+
+Section \ref{sec-sys-props} presents an exhaustive
+listing of properties recognized by \Chord.
+Here, we only describe those set in the above
+sample \verb+chord.properties+ file:
+
 \begin{itemize}
 \item
-{\tt chord.main.class} specifies the main class of the program
-to be analyzed.
-\item
-{\tt chord.class.path} specifies the application classpath of the program to be analyzed (as opposed to the
-JDK library classpath which consists of jar files such as {\tt rt.jar}
-which are implicitly appended to the program's application classpath to yield the
-complete program classpath).
-\item
-{\tt chord.out} and {\tt chord.err} specify the files (typically one and the same) to which the standard output and
-standard error streams, respectively, must be redirected while running Chord.
-\item
-{\tt chord.serial.file} specifies the location of a file from (resp. to) which
-the serialized form of the intermediate representation of the program to be analyzed must be loaded (resp. stored).
-This option is provided because loading the serialized form is much faster than
-constructing the intermediate representation of the program each time.
-\item
-{\tt chord.java.analysis.path} and {\tt chord.dlog.analysis.path} 
-specify all program domains, program relations, and program analyses
-to be included in the project and the
-dependencies between them (see Section \ref{sec-tg}).
-In the above case, they specify certain
- program domains, program relations, and program analyses
-pre-defined in \Chord, in particular, all those in (1) Chord's standard library, (2) Chord's downcast safety
-analysis, and (3) Chord's monomorphic call site analysis.
-Users can also define their own
-program domains, program relations, and program analyses and include them in the project.
-\item
-{\tt chord.analyses} is a space- or comma-separated list of the names of program domains,
-program relations, and program analyses that must be run in that order.  In the above case,
-these are downcast safety analysis, denoted {\tt downcast-java}, and monomorphic call site analysis,
-denoted {\tt monosite-java}.
-Both are program analyses pre-defined in \Chord\ 
-but users can also include the names
-of program domains, program relations, and program analyses they have defined and wish to run.
+\verb+chord.main.class+ specifies the fully-qualified name of the main class of the
+program.
+\item
+\verb+chord.class.path+ specifies the classpath of the program.
+The value of pre-defined property \verb+${sun.boot.class.path}+ is implicitly appended
+to it.
+\item
+\verb+chord.src.path+ specifies the Java source path of the program.
+All program analyses in Chord operate on Java bytecode (specified by \verb+chord.class.path+).
+The only use of this property is to HTMLize the Java source files of the program so that the
+results of program analyses can be reported at the Java source code level.
+\item
+\verb+chord.run.ids+ specifies a list of IDs to identify runs of the program.
+It is used by dynamic program analyses to determine how many times the program must be run.
+An additional use of this property is to allow specifying the command-line arguments to use
+in the run having ID \verb+XXX+ via property \verb+chord.args.XXX+, as illustrated by
+properties \verb+chord.args.0+ and \verb+chord.args.1+ in the above example.
 \end{itemize}
 
+Execute Chord on the above program by running the following command in Chord's main directory:
+
+\begin{quote}
+\begin{verbatim}
+prompt> ant -Dchord.work.dir=<...>/example run
+\end{verbatim}
+\end{quote}
+
+where \verb+<...>+ denotes either the absolute path, or the path relative to Chord's main directory,
+of the parent of directory {\tt example/}.
+System property \verb+chord.work.dir+ specifies the working directory during Chord's execution.
+
+The above command causes Chord to load the \verb+chord.properties+ file if present.
+The location of this file may be specified explicitly by property
+\verb+chord.props.file+ on the above command line.
+By default, it is \verb+${chord.work.dir}/chord.properties+.
+
+In the above case, the above command does not do much beyond loading the sample \verb+chord.properties+ file.
+For Chord to do something interesting, you need to set additional properties, either
+on the above command line or in the \verb+chord.properties+ file, specifying the task(s)
+Chord must perform.
+All tasks are summarized in the ``Chord task properties" section at \ref{sec-sys-props}.
+The two most common tasks, described next, are building program scope and
+running program analyses.
+
+\paragraph{Building Program Scope}
+
+Chord computes program scope (i.e. reachable classes and methods of a given Java program) if property
+\verb+chord.build.scope+ is set to \verb+true+.
+The algorithm used to compute program scope is specified by 
+properties \verb+chord.scope.kind+ and \verb+chord.reuse.scope+.
+
+\begin{itemize}
+\item
+If property \verb+chord.reuse.scope+ is set to \verb+true+ and the files specified by
+properties \verb+chord.classes.file+ and \verb+chord.methods.file+ exist,
+then Chord regards those files as specifying the classes and methods, respectively,
+to be regarded as reachable.  The format
+of the classes file is a class name per line (e.g. \verb+foo.Main+).  The format
+of the methods file is an entry of the form \verb+<method name>:<method descriptor>@<class name>+ per line,
+specifying the method's name, the method's descriptor, and the method's declaring class
+(e.g. \verb+main:([Ljava/lang/String;)V@foo.Main+).
+Chord throws a runtime exception if the declaring class of a method in the methods file is not
+specified in the classes file.
+\item
+If property \verb+chord.reuse.scope+ is set to \verb+false+ or the file specified by
+property \verb+chord.classes.file+ or \verb+chord.methods.file+ does not exist,
+then Chord computes program scope
+using the algorithm specified by property \verb+chord.scope.kind+ and then
+writes the classes and methods deemed reachable by that algorithm to those files.
+
+The possible legal values of property \verb+chord.scope.kind+ are \verb+rta+ and \verb+dynamic+.
+In both cases, Chord expects properties \verb+chord.main.class+ and \verb+chord.class.path+
+to be set.
+\begin{itemize}
+\item
+The \verb+rta+ value instructs Chord to compute the program scope statically using Rapid Type Analysis.
+\item
+The \verb+dynamic+ value instructs Chord to compute the program scope dynamically, by running the program
+and observing using JVMTI the classes that are loaded at run-time.
+The number of times the program is run and the command-line arguments to be supplied to
+the program in each run is specified by properties \verb+chord.run.ids+ and
+\verb+chord.args.XXX+ for each run ID \verb+XXX+.  By default, the program is run only once, using run ID \verb+0+,
+and without any command-line arguments.
+Only classes loaded in some run are regarded as reachable but {\it all} methods of each loaded class are regarded
+as reachable regardless of whether they were invoked in the run.
+The rationale behind this decision is to both reduce run-time overhead of JVMTI and to increase the
+predictive power of program analyses performed using the computed program scope.
+\end{itemize}
+\end{itemize}
+
+The default value of \verb+chord.reuse.scope+ is \verb+false+ and the default value of \verb+chord.scope.kind+
+is \verb+rta+.
+Finally, the default values of \verb+chord.classes.file+ and \verb+chord.methods.file+ are
+\verb+${chord.out.dir}/classes.txt+ and \verb+${chord.out.dir}/methods.txt+, respectively,
+where property \verb+chord.out.dir+ defaults to \verb+${chord.work.dir}/chord_output/+,
+and property \verb+chord.work.dir+ defaults to the current working directory.
+
+\paragraph{Running Program Analyses}
+
+TODO
+
+%chord.run.analyses
+%List of names of program analyses to be run in order.
+%chord.java.analysis.path	 path	 Classpath containing program analyses in Java (i.e. @Chord-annotated classes).	${chord.main.dir}/classes/
+%chord.dlog.analysis.path	 path	 Path of directories containing program analyses in Datalog (i.e. *.datalog and *.dlog files).	${chord.main.dir}/src/main/dlog/
+%chord.reuse.rels	 bool	 Construct program relations from BDDs stored on disk (from a previous run of Chord) whenever possible instead of re-computing them.	false
+%chord.publish.results	 bool	 Publish the results of program analyses in HTML.	true
+
+%\texonly{\newpage}
+%\section{Program Representation (todo)}
+
+%TODO
+
+%\texonly{\newpage}
+%\section{Implementing Program Analyses (todo)}
+
+%TODO
+
 \texonly{\newpage}
-\section{Specifying Program Scope}
-
-\Chord\ takes as input a complete (i.e., closed) program with a main method.
-The scope of the program must be specified in one of the
-following ways:
+\section{Chord System Properties}
+\label{sec-sys-props}
+
+The following properties are recognized by Chord.
+The separator for list-valued properties can be either a blank space, a comma, a colon, or a semi-colon.
+
+\begin{table}[htp]
+%\caption{}
+\xmlattributes*{table}{border="1"}
+\begin{center}
+\begin{tabular}{|l|lll|}
+\hline
+{\bf Property name} & {\bf Type} & {\bf Description} & {\bf Default value (if any)} \\
+\hline
+\multicolumn{4}{l}{{\bf Chord resource properties}} \\
+\verb+chord.main.dir+ & location & \verb+main+ directory in Chord's installation. & \\
+\hline
+\verb+chord.lib.dir+ & location & Directory containing libraries needed by Chord. & \verb+${chord.main.dir}/lib/+ \\
+\hline
+\verb+chord.main.class.path+ & path & Classpath of Chord.
+It includes the path specified by property \verb+chord.java.analysis.path+ to allow running user-defined program analyses.
+& See \verb+${chord.main.dir}/build.xml+ \\
+\hline
+\verb+chord.bddbddb.class.path+ & path & Classpath of bddbddb. & See \verb+${chord.main.dir}/build.xml+ \\
+\hline
+\verb+chord.instr.agent.file+ & location & Shared library implementing Chord instrumentation agent.
+& \verb+${chord.main.dir}/lib/libchord_instr_agent.so+ (on Linux; \verb+.dll+ on Windows and \verb+.dylib+ on MacOS)\\
+\hline
+\verb+chord.javadoc.url+ & string & URL of the Javadoc location of program analyses.
+It is used when publishing targets (i.e.
+when property \verb+chord.publish.targets+ is set to true).
+& \verb+http://chord.stanford.edu/javadoc/+ \\
+\hline
+\multicolumn{4}{l}{{\bf Chord boot properties}} \\
+\hline
+\verb+chord.work.dir+ & location & Working directory during Chord's execution. & current working directory \\
+\hline
+\verb+chord.props.file+ & location & Properties file loaded by \verb+${chord.main.dir}/build.xml+.
+Any of the below properties, as well as other user-defined properties to be passed to Chord (e.g.
+for user-defined analyses), may be set in this file.
+Each relative (instead of absolute) path element in the value of any property named \verb+chord.<...>.path+ set in this file is converted to an absolute path element with respect to the directory containing this file.
+& \verb+${chord.work.dir}/chord.properties+ \\
+\hline
+\verb+chord.max.heap+ & string & Maximum memory size of JVM running Chord. & \verb+1024m+ \\
+\hline 
+\verb+chord.max.stack+ & string & Maximum thread stack size of JVM running Chord. & \verb+32m+ \\
+\hline
+\verb+chord.jvmargs+ & string & Arguments passed to JVM running Chord. & \verb+"-showversion -ea -Xmx${chord.max.heap} -Xss${chord.max.stack}"+ \\
+\hline
+\verb+chord.bddbddb.max.heap+ & string & Maximum memory size of JVM running bddbddb. & \verb+1024m+ \\
+\hline
+\multicolumn{4}{l}{{\bf Chord task properties}} \\
+\hline
+\verb+chord.build.scope+ & bool & Compute program scope (i.e.
+reachable classes and methods) using algorithm specified by properties \verb+chord.scope.kind+ and \verb+chord.reuse.scope+.
+Program scope is computed regardless of the value of this property if another task (e.g.
+a program analysis to be run) needs it.
+& \verb+false+ \\
+\hline
+\verb+chord.run.analyses+ & string list & List of names of program analyses to be run in order. & \verb+""+ \\
+\hline
+\verb+chord.print.rels+ & string list & List of names of program relations whose contents must be printed to files \verb+${chord.out.dir}/<...>.txt+ where \verb+<...>+ denotes the relation name.
+& \verb+""+ \\
+\hline
+\verb+chord.publish.targets+ & bool & Publish all targets defined by program analyses in paths specified by properties \verb+chord.java.analysis.path+ and \verb+chord.dlog.analysis.path+.
+& \verb+false+ \\
+\hline
+\multicolumn{4}{l}{{\bf Basic program properties}} \\
+\hline
+\verb+chord.main.class+ & class & Fully-qualified name of main class of given program (e.g. \verb+com.example.Main+). &  \\
+\hline
+\verb+chord.class.path+ & path & Classpath of given program. \verb+${sun.boot.class.path}+ is implicitly appended to it. & \verb+${sun.boot.class.path}+ \\
+\hline
+\verb+chord.src.path+ & path & Java source path of given program. & \verb+""+ \\
+\hline
+\verb+chord.run.ids+ & string list & List of IDs to identify runs of given program. & \verb+0+ \\
+\hline
+\verb+chord.args.XXX+ & string & Command-line argument string to be used for given program in run having ID \verb+XXX+. & \verb+""+ \\
+\hline
+\verb+chord.runtime.jvmargs+ & string & Arguments to JVM when running given program.
+It is used when the program is executed to compute its scope (i.e.
+when property \verb+chord.scope.kind+ is set to \verb+dynamic+) as well as when the instrumented program is executed for dynamic analyses.
+& \verb+"-ea -Xmx1024m"+ \\
+\hline
+\multicolumn{4}{l}{{\bf Program scope properties}} \\ 
+\hline
+\verb+chord.scope.kind+ & [\verb+rta+|\verb+dynamic+] & Algorithm to compute program scope (i.e.
+reachable classes and methods).
+Current choices include \verb+rta+ (static Rapid Type Analysis) and \verb+dynamic+ (dynamic analysis).
+& \verb+rta+ \\
+\hline
+\verb+chord.reuse.scope+ & bool & Treat program scope as the classes and methods listed in files specified by properties \verb+chord.classes.file+ and \verb+chord.methods.file+, respectively.
+Property \verb+chord.scope.kind+ is ignored if this property is set to \verb+true+ and the two files exist.
+& \verb+false+ \\
+\hline
+\verb+chord.check.exclude+ & string list & List of prefixes of names of classes and packages to be excluded by program analyses.
+Interpretation of this property is analysis-specific.
+& \verb+"java.,sun.,com.sun.,com.ibm.,org.apache.harmony.,joeq.,jwutil."+ \\
+\hline
+\verb+chord.instr.exclude+ & string list & List of prefixes of names of classes and packages to be excluded from being instrumented.
+Classes in package \verb+java.lang.ref+ are not instrumented regardless of the value of this property.
+& \verb+"java.,sun.,com.sun.,com.ibm.,org.apache.harmony.,joeq.,jwutil."+ \\
+\hline
+\multicolumn{4}{l}{{\bf Program analysis properties}} \\
+\hline
+\verb+chord.java.analysis.path+ & path & Classpath containing program analyses written in Java (i.e. \verb+@Chord+-annotated classes). & \verb+${chord.main.dir}/classes/+ \\
+\hline
+\verb+chord.dlog.analysis.path+ & path & Path of directories containing program analyses written in Datalog (i.e. \verb+*.datalog+ and \verb+*.dlog+ files). & \verb+${chord.main.dir}/src/dlog/+ \\
+\hline
+\verb+chord.reuse.rels+ & bool & Construct program relations from BDDs stored on disk (from a previous run of Chord) whenever possible instead of re-computing them. & \verb+false+ \\
+\hline
+\verb+chord.publish.results+ & bool & Publish the results of program analyses in HTML. Interpretation of this property is analysis-specific. & \verb+true+ \\
+\hline
+\multicolumn{4}{l}{{\bf Program transformation properties}} \\
+\hline
+\verb+chord.ssa+ & bool & Do SSA transformation for all methods deemed reachable by the algorithm used to compute program scope. & \verb+true+ \\
+\hline
+\multicolumn{4}{l}{{\bf Chord debug properties}} \\
+\hline
+\verb+chord.bddbddb.noisy+ & bool & Make bddbddb output debug info when solving Datalog programs. & \verb+false+ \\
+\hline
+\verb+chord.save.maps+ & bool & Write to file \verb+${chord.bddbddb.work.dir}/<...>.map+ when saving program domain named \verb+<...>+ (useful for debugging Datalog programs using \verb+debug+ target of \verb+${chord.main.dir}/build.xml+).
+& \verb+true+ \\
+\hline
+\multicolumn{4}{l}{{\bf Chord instrumentation properties}} \\
+\hline
+\verb+chord.trace.pipe+ & bool & Implement the dynamic trace file as a POSIX pipe instead of a regular file. & \verb+true+ \\
+\hline
+\verb+chord.trace.block.size+ & int & Number of bytes to read/write in a single operation from/to the dynamic trace file. & \verb+4096+ \\
+\hline
+\verb+chord.runtime.class+ & class & Subclass of \verb+chord.project.Runtime+ used by dynamic program analyses to handle events generated during an instrumented program's execution.
+& \verb+chord.project.BufferedRuntime+ \\
+\hline
+\hline
+\multicolumn{4}{l}{{\bf Chord output properties}} \\
+\hline
+\verb+chord.out.dir+ & location & Directory to which Chord dumps all files. & \verb+${chord.work.dir}/chord_output/+ \\
+\hline
+\verb+chord.out.file+ & location & File to redirect standard output stream during Chord's execution. & \verb+${chord.out.dir}/log.txt+ \\
+\hline
+\verb+chord.err.file+ & location & File to redirect standard error stream during Chord's execution. & \verb+${chord.out.dir}/log.txt+ \\
+\hline
+\verb+chord.classes.file+ & location & File from/to which list of classes deemed reachable is read/written. & \verb+${chord.out.dir}/classes.txt+ \\
+\hline
+\verb+chord.methods.file+ & location & File from/to which list of methods deemed reachable is read/written. & \verb+${chord.out.dir}/methods.txt+ \\
+\hline
+\verb+chord.bddbddb.work.dir+ & location & Directory used by bddbddb as its input/output directory. & \verb+${chord.out.dir}/bddbddb/+ \\
+\hline
+\verb+chord.boot.classes.dir+ & location & Directory containing instrumented JDK classes used by given program. & \verb+${chord.out.dir}/boot_classes/+ \\
+\hline
+\verb+chord.user.classes.dir+ & location & Directory containing instrumented non-JDK classes of given program. & \verb+${chord.out.dir}/user_classes/+  \\
+\hline
+\verb+chord.instr.scheme.file+ & location & File specifying kind and format of events in dynamic trace file. & \verb+${chord.out.dir}/scheme.ser+ \\
+\hline
+\verb+chord.crude.trace.file+ & location & Crude dynamic trace file. & \verb+${chord.out.dir}/crude_trace.txt+ \\
+\hline
+\verb+chord.final.trace.file+ & location & Final dynamic trace file. & \verb+${chord.out.dir}/final_trace.txt+
+\end{tabular}
+\end{center}
+\end{table}
+
+\texonly{\newpage}
+\section{Acknowledgments}
+
+Chord would not be possible without the following open-source software:
 
@@
 \item
-Defining system property {\tt chord.serial.file} as
-the location of a file containing in serialized form the intermediate representation
-of the program generated by a previous run of \Chord.
-\item
-Defining system property {\tt chord.main.class} as the fully-qualified name
-of the program's main class (e.g., {\tt foo.Main}) and system property
-{\tt chord.class.path} as the program's application classpath.
+\xlink{Joeq}{http://joeq.sourceforge.net/}, a Java compiler framework
+\item
+\xlink{Javassist}{http://www.csg.is.titech.ac.jp/~chiba/javassist/}, a Java bytecode manipulation tool
+\item
+\xlink{bddbddb}{http://bddbddb.sourceforge.net/}, a BDD-based Datalog solver
 \end{itemize}
 
-The former option is provided because loading the serialized form
-is much faster than constructing the intermediate representation
-of the program each time.
-
-\Chord\ first checks whether system property {\tt chord.serial.file} is defined.
-If it is defined, then it further checks whether the file specified by it
-exists.
-If the file exists, then \Chord\ treats the file as a serialized form of
-the intermediate representation of the program to be analyzed, and loads it.
-
-If system property {\tt chord.serial.file} is not defined or if the file
-specified by the property does not exist, then \Chord\ requires system properties
-{\tt chord.main.class} and {\tt chord.class.path} 
-to be defined, and uses them to
-build an intermediate representation of the program (see Section \ref{sec-prog-build})
-and perform source-to-source transformations on that representation
-(see Section \ref{sec-prog-transform}) that simplify several program analyses.
-Finally, if system property {\tt chord.serial.file} was defined but
-specified a non-existent file, then \Chord\ creates the file and stores the
-transformed intermediate representation of the program into the file in
-serialized form so that it can be loaded in subsequent runs.
-
-Defining system property {\tt chord.src.path} is optional.
-It specifies the program's Java source path which provides the
-locations of the program's {\tt .java} files so that, for instance, the
-results of an analysis (which operates on the intermediate representation of
-the program extracted from {\tt .class} files) can be mapped to the
-program's {\tt .java} files and presented to the user for
-manual inspection.
+Chord additionally uses the following open-source tools and libraries:
+\begin{itemize}
+\item
+\xlink{Ant-Contrib}{http://ant-contrib.sourceforge.net/}, a collection of useful Ant tasks  
+\item
+\xlink{BuDDy}{http://buddy.sourceforge.net/}, a BDD library
+\item
+\xlink{GNU Trove}{http://trove4j.sourceforge.net/}, a primitive collections library for Java
+\item
+\xlink{Java2HTML}{http://www.java2html.com/} and \xlink{Java2Html}{http://www.java2html.de/}, Java to HTML tools
+\item
+\xlink{Saxon}{http://saxon.sourceforge.net/}, an XSLT processor
+\end{itemize}
+
+Chord was supported in part by grants from the National Science Foundation, an equipment grant from Intel, and a Microsoft fellowship during 2005-2007.
+
+\end{document}
 
@@
 
-Given the program's program's main class via system property {\tt chord.main.class} and the
-program's application classpath via system property {\tt chord.class.path},
+Given the program's application classpath via system property {\tt chord.app.class.class}
+and either the program's main class via system property {\tt chord.main.class} or
+the program's root methods via system property {\tt chord.root.classes}
+(and, optionally, system properties {\tt chord.include.root.methods}
+and {\tt chord.exclude.root.methods}),
 \Chord\ uses system property {\tt chord.program.builder.class} to locate the
@@
 \end{itemize}
-
-\texonly{\newpage}
-\section{Program Transformations}
-\label{sec-prog-transform}
-
-After building the intermediate representation of the
-program, \Chord\ performs
-the following source-to-source transformations 
-that simplify several program analyses:
-
-\begin{enumerate}
-\item
-{\bf Sync Removal}
-
-Removes all synchronized blocks from each method M in each class
-C in the program by generating a fresh static method (in class C)
-per synchronized block in the body of M.
-
-\item
-{\bf Loop Removal}
-
-Removes all loops from each method M in each class C in the program
-by generating a fresh static method (in class C)
-per loop in the body of M.
-
-\item
-{\bf Skip Removal}
-
-Removes all skip (i.e., no-op) statements from each method in each class
-in the program.
-Such statements arise primarily because \Chord, due to its emphasis
-on the analysis of objects, replaces each
-statement in the original program that does not access any
-value of reference type (i.e., only accesses values of primitive
-type) by a skip statement.
-
-\end{enumerate}
 
@@
 \texonly{\newpage}
-\section{Project Organization}
+\section{Executing Targets}
 
@@
 \section{Chord System Properties}
-\label{sec-sys-prop}
+\label{sec-sys-props}
 
@@
 \item
-{\tt chord.serial.file}
+{\tt chord.program.file}
 
@@
 \item
-{\tt chord.main.class}
-
-The fully-qualified name of the main class (e.g., {\tt foo.Main}) of the
-program to be analyzed.
-
-This system property (and {\tt chord.class.path})
-must be defined if system property {\tt chord.serial.file}
+{\tt chord.app.class.path}
+
+The application classpath of the program to be analyzed (as opposed to the
+JDK library classpath which consists of jar files {\tt rt.jar}, {\tt jsse.jar}, and
+{\tt jce.jar} and is implicitly appended to the program's application classpath to
+yield the complete program classpath).
+
+Each path element must be the location of a directory or a jar/zip file.
+
+The path elements must be separated by the system-specific path separator character
+(e.g., `:' on UNIX and `;' on Windows).
+
+This system property (and either {\tt chord.main.class} or {\tt chord.root.classes}) must
+be defined if system property {\tt chord.program.file}
 does not specify the file containing the serialized form of the intermediate representation
@@
 \item
-{\tt chord.class.path}
-
-The application classpath of the program to be analyzed (as opposed to the
-JDK library classpath which consists of jar files such as {\tt rt.jar}
-which are implicitly appended to the program's application classpath to
-yield the complete program classpath).
-
-Each path element must be the location of a directory or a jar/zip file.
-
-The path elements must be separated by the system-specific path separator character
-(e.g., `:' on UNIX and `;' on Windows).
-
-This system property (and {\tt chord.main.class}) must
-be defined if system property {\tt chord.serial.file}
+{\tt chord.main.class}
+
+The fully-qualified name of the main class (e.g., {\tt com.example.Main}) of the
+program to be analyzed if the program is a complete one.
+
+This system property (and {\tt chord.app.class.path})
+must be defined if system property {\tt chord.program.file}
 does not specify the file containing the serialized form of the intermediate representation
@@
 \item
-{\tt chord.src.path}
+{\tt chord.root.classes}
+
+A space-separated list of fully-qualified names of classes (e.g., {\tt com.example.Main})
+whose public methods must be regarded as root methods (i.e., entry points) if the program
+is an incomplete one.
+
+This system property (and {\tt chord.app.class.path} plus, optionally,
+system properties {\tt chord.include.root.methods} and {\tt chord.exclude.root.methods})
+must be defined if system property {\tt chord.program.file}
+does not specify the file containing the serialized form of the intermediate representation
+of the program to be analyzed.
+
+\item
+{\tt chord.include.root.methods}
+
+A space-separated list of fully-qualified signatures of
+methods (e.g., {\tt com.example.Foo.bar(int,com.example.Baz[])}) that must be included
+into the set of root methods (i.e., entry points)
+if the program is an incomplete one.
+
+\item
+{\tt chord.exclude.root.methods}
+
+A space-separated list of fully-qualified signatures of
+methods (e.g., {\tt com.example.Foo.bar(int,com.example.Baz[])}) that must be excluded
+from the set of root methods (i.e., entry points)
+if the program is an incomplete one.
+
+\item
+{\tt chord.java.src.path}
 
@@
 \item
-{\tt chord.dlog.analysis.path}
+{\tt chord.dlogmap.path}
 
@@
 \item
-{\tt chord.java.analysis.path}
+{\tt chord.taskmap.path}
 
@@
 \item
-{\tt chord.analyses}
+{\tt chord.trgts}
 
@@
 Section \ref{sec-predef-java-analyses}.
+
+\texonly{\newpage}
+\section{Program Transformations}
+\label{sec-prog-transform}
+
+After building the intermediate representation of the
+program, \Chord\ performs
+the following source-to-source transformations 
+that simplify several program analyses:
+
+\begin{enumerate}
+\item
+{\bf Sync Removal}
+
+Removes all synchronized blocks from each method M in each class
+C in the program by generating a fresh static method (in class C)
+per synchronized block in the body of M.
+
+\item
+{\bf Loop Removal}
+
+Removes all loops from each method M in each class C in the program
+by generating a fresh static method (in class C)
+per loop in the body of M.
+
+\item
+{\bf Skip Removal}
+
+Removes all skip (i.e., no-op) statements from each method in each class
+in the program.
+Such statements arise primarily because \Chord, due to its emphasis
+on the analysis of objects, replaces each
+statement in the original program that does not access any
+value of reference type (i.e., only accesses values of primitive
+type) by a skip statement.
+
+\end{enumerate}

### 2010-02-16 main/src/java/chord/analyses/thread/DomA.java

@@ -34,7 +34,7 @@ import chord.doms.DomM;
  * <tt>java.lang.Thread</tt> and <tt>c</tt> is an abstract context
  * of that method.
  * 
- * @see chord.thread.ForkCtxtsAnalysis
+ * @see chord.analyses.thread.ForkCtxtsAnalysis
  * 
  * @author Mayur Naik (mhn@cs.stanford.edu)
  */

### 2010-04-01 main/doc/acknowledgments.tex

\section{Acknowledgments}
\label{sec:acknowledgments}

Chord would not be possible without the following open-source software:

\begin{itemize}
\item
\href{http://joeq.sourceforge.net/}{Joeq}, a Java compiler framework
\item
\href{http://www.csg.is.titech.ac.jp/~chiba/javassist/}{Javassist},
a Java bytecode manipulation tool
\item
\href{ttp://bddbddb.sourceforge.net/}{bddbddb}, a BDD-based Datalog solver
\end{itemize}

Chord additionally relies on the following open-source tools and libraries:
\begin{itemize}
\item
\href{http://ant-contrib.sourceforge.net/}{Ant-Contrib},
a collection of useful Ant tasks  
\item
\href{http://buddy.sourceforge.net/}{BuDDy}, a BDD library
\item
\href{http://trove4j.sourceforge.net/}{GNU Trove},
a primitive collections library for Java
\item
\href{http://www.java2html.com/}{Java2HTML} and
\href{http://www.java2html.de/}{Java2Html}, Java to HTML tools
\item
\href{http://saxon.sourceforge.net/}{Saxon}, an XSLT processor
\end{itemize}

Chord was supported in part by grants from the National Science Foundation,
an equipment grant from Intel, and a Microsoft fellowship during 2005-2007.

### 2010-04-01 main/doc/chord_sysprops.tex

\section{Chord System Properties}
\label{sec:chord-sysprops}

The following properties are recognized by Chord.
The separator for list-valued properties can be either a blank space, a comma, a colon, or a semi-colon.

\begin{longtable}{|l|p{0.4in}|p{4.0in}|}
%\caption{TODO} \\
\hline
{\bf Property name} & {\bf Type} & {\bf Description and default value if any} \\
\endfirsthead
\hline
{\bf Property name} & {\bf Type} & {\bf Description and default value if any} \\
\endhead
\hline
\hline
\multicolumn{3}{|l|}{{\bf Chord resource properties}} \\
\verb+chord.main.dir+ & location & \verb+main+ directory in Chord's installation. \\
\hline
\verb+chord.lib.dir+ & location & Directory containing libraries needed by Chord.
\newline Default: \verb+${chord.main.dir}/lib/+ \\
\hline
\verb+chord.main.class.path+ & path & Classpath of Chord.
It includes the path specified by property \verb+chord.java.analysis.path+
to allow running user-defined program analyses.
\newline Default: See \verb+${chord.main.dir}/build.xml+ \\
\hline
\verb+chord.bddbddb.class.path+ & path & Classpath of \verb+bddbddb+.
\newline Default: See \verb+${chord.main.dir}/build.xml+ \\
\hline
\verb+chord.instr.agent.file+ & location &
Shared library implementing Chord instrumentation agent. \\
%\newline Default: \verb+${chord.main.dir}/lib/libchord_instr_agent.so+ \\
\hline
\verb+chord.javadoc.url+ & string & URL of the Javadoc location of program analyses.
It is used when publishing targets (i.e., when property \verb+chord.publish.targets+
is set to true).
\newline Default: \verb+http://chord.stanford.edu/javadoc/+ \\
\hline
\hline
\multicolumn{3}{|l|}{{\bf Chord boot properties}} \\
\hline
\verb+chord.work.dir+ & location & Working directory during Chord's execution.
\newline Default: current working directory \\
\hline
\verb+chord.props.file+ & location & Properties file loaded by \verb+${chord.main.dir}/build.xml+.
Any of the below properties, as well as other user-defined properties to be passed to Chord
(e.g., for user-defined analyses), may be set in this file.  Each relative (instead of absolute)
path element in the value of any property named \verb+chord.<...>.path+ set in this file is
converted to an absolute path element with respect to the directory containing this file.
\newline Default: \verb+${chord.work.dir}/chord.properties+ \\
\hline
\verb+chord.max.heap+ & string & Maximum memory size of JVM running Chord.
\newline Default: \verb+1024m+ \\
\hline 
\verb+chord.max.stack+ & string & Maximum thread stack size of JVM running Chord.
\newline Default: \verb+32m+ \\
\hline
\verb+chord.jvmargs+ & string & Arguments passed to JVM running Chord. \\
%\newline Default: \verb+"-showversion -ea -Xmx${chord.max.heap} -Xss${chord.max.stack}"+ \\
\hline
\verb+chord.bddbddb.max.heap+ & string & Maximum memory size of JVM running bddbddb.
\newline Default: \verb+1024m+ \\
\hline
\hline
\multicolumn{3}{|l|}{{\bf Chord task properties}} \\
\hline
\verb+chord.build.scope+ & bool & Compute analysis scope (i.e., reachable classes and methods)
using algorithm specified by properties \verb+chord.scope.kind+ and \verb+chord.reuse.scope+.
Program scope is computed regardless of the value of this property if another task (e.g., a
program analysis to be run) needs it.
\newline Default: \verb+false+ \\
\hline
\verb+chord.run.analyses+ & string list & List of names of program analyses to be run
in order.
\newline Default: \verb+""+ \\
\hline
\verb+chord.print.rels+ & string list & List of names of program relations whose contents
must be printed to files \verb+${chord.out.dir}/<...>.txt+ where \verb+<...>+ denotes the
relation name. 
\newline Default: \verb+""+ \\
\hline
\verb+chord.publish.targets+ & bool & Publish all targets defined by program analyses in
paths specified by properties \verb+chord.java.analysis.path+ and
\verb+chord.dlog.analysis.path+.
\newline Default: \verb+false+ \\
\hline
\hline
\multicolumn{3}{|l|}{{\bf Basic program properties}} \\
\hline
\verb+chord.main.class+ & class & Fully-qualified name of main class of given program
(e.g., \verb+com.example.Main+). \\
\hline
\verb+chord.class.path+ & path & Classpath of given program. \verb+${sun.boot.class.path}+
is implicitly appended to it.
\newline Default: \verb+${sun.boot.class.path}+ \\
\hline
\verb+chord.src.path+ & path & Java source path of given program.
\newline Default: \verb+""+ \\
\hline
\verb+chord.run.ids+ & string list & List of IDs to identify runs of given program.
\newline Default: \verb+0+ \\
\hline
\verb+chord.args.XXX+ & string & Command-line argument string to be used for given
program in run having ID \verb+XXX+.
\newline Default: \verb+""+ \\
\hline
\verb+chord.runtime.jvmargs+ & string & Arguments to JVM when running given program.
It is used when the program is executed to compute its scope (i.e., when property
\verb+chord.scope.kind+ is set to \verb+dynamic+) as well as when the instrumented
program is executed for dynamic analyses.
\newline Default: \verb+"-ea -Xmx1024m"+ \\ 
\hline
\hline
\multicolumn{3}{|l|}{{\bf Program scope properties}} \\ 
\hline
\verb+chord.scope.kind+ & [\verb+rta+ $|$ \verb+dynamic+] & Algorithm to compute
analysis scope (i.e., reachable classes and methods).  Current choices include
\verb+rta+ (static Rapid Type Analysis) and \verb+dynamic+ (dynamic analysis).
\newline Default: \verb+rta+ \\
\hline
\verb+chord.reuse.scope+ & bool & Treat analysis scope as the classes and methods listed
in files specified by properties \verb+chord.classes.file+ and \verb+chord.methods.file+,
respectively. Property \verb+chord.scope.kind+ is ignored if this property is set to
\verb+true+ and the two files exist.
\newline Default: \verb+false+ \\
\hline
\verb+chord.check.exclude+ & string list & List of prefixes of names of classes and
packages to be excluded by program analyses.
Interpretation of this property is analysis-specific. \\
%\newline Default: \verb+"java.,sun.,com.sun.,com.ibm.,org.apache.harmony.,joeq.,jwutil."+ \\
\hline
\verb+chord.instr.exclude+ & string list & List of prefixes of names of classes and
packages to be excluded from being instrumented.  Classes in package \verb+java.lang.ref+
are not instrumented regardless of the value of this property. \\
%\newline Default: \verb+"java.,sun.,com.sun.,com.ibm.,org.apache.harmony.,joeq.,jwutil."+ \\
\hline
\hline
\multicolumn{3}{|l|}{{\bf Program analysis properties}} \\
\hline
\verb+chord.java.analysis.path+ & path & Classpath containing program analyses written
in Java (i.e., \verb+@Chord+-annotated classes).
\newline Default: \verb+${chord.main.dir}/classes/+ \\
\hline
\verb+chord.dlog.analysis.path+ & path & Path of directories containing program analyses
written in Datalog (i.e., \verb+*.datalog+ and \verb+*.dlog+ files).
\newline Default: \verb+${chord.main.dir}/src/dlog/+ \\
\hline
\verb+chord.reuse.rels+ & bool & Construct program relations from BDDs stored on disk
(from a previous run of Chord) whenever possible instead of re-computing them.
\newline Default: \verb+false+ \\
\hline
\verb+chord.publish.results+ & bool & Publish the results of program analyses in HTML.
Interpretation of this property is analysis-specific.
\newline Default: \verb+true+ \\
\hline
\hline
\multicolumn{3}{|l|}{{\bf Program transformation properties}} \\
\hline
\verb+chord.ssa+ & bool & Do SSA transformation for all methods deemed reachable by the
algorithm used to compute analysis scope.
\newline Default: \verb+true+ \\
\hline
\hline
\multicolumn{3}{|l|}{{\bf Chord debug properties}} \\
\hline
\verb+chord.bddbddb.noisy+ & bool & Make bddbddb output debug info when solving Datalog programs.
\newline Default: \verb+false+ \\
\hline
\verb+chord.save.maps+ & bool & Write to file \verb+${chord.bddbddb.work.dir}/<...>.map+
when saving program domain named \verb+<...>+ (useful for debugging Datalog programs using
\verb+debug+ target of \verb+${chord.main.dir}/build.xml+).
\newline Default: \verb+true+ \\
\hline
\hline
\multicolumn{3}{|l|}{{\bf Chord instrumentation properties}} \\
\hline
\verb+chord.trace.pipe+ & bool & Implement the dynamic trace file as a POSIX pipe instead of a regular file.
\newline Default: \verb+true+ \\
\hline
\verb+chord.trace.block.size+ & int & Number of bytes to read/write in a single operation from/to the dynamic trace file.
\newline Default: \verb+4096+ \\
\hline
\verb+chord.runtime.class+ & class & Subclass of \verb+chord.project.Runtime+ used by dynamic program
analyses to handle events generated during an instrumented program's execution.
\newline Default: \verb+chord.project.BufferedRuntime+ \\
\hline
\hline
\multicolumn{3}{|l|}{{\bf Chord output properties}} \\
\hline
\verb+chord.out.dir+ & location & Directory to which Chord dumps all files.
\newline Default: \verb+${chord.work.dir}/chord_output/+ \\
\hline
\verb+chord.out.file+ & loc. & File to redirect standard output stream during Chord's execution.
\newline Default: \verb+${chord.out.dir}/log.txt+ \\
\hline
\verb+chord.err.file+ & loc. & File to redirect standard error stream during Chord's execution.
\newline Default: \verb+${chord.out.dir}/log.txt+ \\
\hline
\verb+chord.classes.file+ & loc. & File from/to which list of classes deemed reachable is read/written.
\newline Default: \verb+${chord.out.dir}/classes.txt+ \\
\hline
\verb+chord.methods.file+ & loc. & File from/to which list of methods deemed reachable is read/written.
\newline Default: \verb+${chord.out.dir}/methods.txt+ \\
\hline
\verb+chord.bddbddb.work.dir+ & loc. & Directory used by bddbddb as its input/output directory.
\newline Default: \verb+${chord.out.dir}/bddbddb/+ \\
\hline
\verb+chord.boot.classes.dir+ & loc. & Directory containing instrumented JDK classes used by given program.
\newline Default: \verb+${chord.out.dir}/boot_classes/+ \\
\hline
\verb+chord.user.classes.dir+ & loc. & Directory containing instrumented non-JDK classes of given program.
\newline Default: \verb+${chord.out.dir}/user_classes/+  \\
\hline
\verb+chord.instr.scheme.file+ & loc. & File specifying kind and format of events in dynamic trace file.
\newline Default: \verb+${chord.out.dir}/scheme.ser+ \\
\hline
\verb+chord.crude.trace.file+ & loc. & Crude dynamic trace file.
\newline Default: \verb+${chord.out.dir}/crude_trace.txt+ \\
\hline
\verb+chord.final.trace.file+ & loc. & Final dynamic trace file.
\newline Default: \verb+${chord.out.dir}/final_trace.txt+ \\
\hline
\end{longtable}

### 2010-04-01 main/doc/deadlock.tex

\section{Deadlock Analysis}
\label{sec:deadlock}

Run the deadlock analysis provided in Chord by running the following command in Chord's \verb+main/+ directory:

\begin{quote}
\begin{verbatim}
prompt> ant -Dchord.work.dir=<...> -Dchord.run.analyses=deadlock-java run
\end{verbatim}
\end{quote}

where directory \verb+<...>+ contains a file named \verb+chord.properties+ which 
defines properties \verb+chord.main.class+, \verb+chord.class.path+, and \verb+chord.src.path+.
See Section \ref{sec:chord-sysprops} for the meaning of these properties.

Directory \verb+main/examples/deadlock_test/+ provides a toy Java program on which you can run the deadlock
analysis by first running \verb+ant+ in that directory (in order to compile the program's \verb+.java+ files to
\verb+.class+ files) and then running the following command in Chord's \verb+main/+ directory:

\begin{quote}
\begin{verbatim}
prompt> ant -Dchord.work.dir=examples/deadlock_test/ -Dchord.run.analyses=deadlock-java run
\end{verbatim}
\end{quote}

Upon successful completion, the file \verb+deadlocks.html+ should be produced in directory
\verb+main/examples/deadlock_test/chord_output/+.

### 2010-04-01 main/doc/program_representation.tex

--- 2009-07-01 doc/user_guide.tex
@@
-\documentclass{article}
-\usepackage{hyperlatex}
+\section{Program Representation}
+\label{sec:program-representation}
 
-%\topmargin=-0.5in
-\texonly{\oddsidemargin=0in}
-\texonly{\evensidemargin=0in}
-\texonly{\textwidth=5.6in}
-%\textheight=8.5in
-
-%TODO: specify use of BDD's under efficiency in section 1 (overview)
-
-\htmlmathitalic{0}
-\htmltitle{User Guide}
-\htmladdress{chord@cs.stanford.edu, \today}
-
-\providecommand\Mdom{{\bf M}}
-\providecommand\Idom{{\bf I}}
-\providecommand\Vdom{{\bf V}}
-\providecommand\Zdom{{\bf Z}}
-\providecommand\Hdom{{\bf H}}
-\providecommand\Edom{{\bf E}}
-\providecommand\Fdom{{\bf F}}
-\providecommand\Cdom{{\bf C}}
-\providecommand\Tdom{{\bf T}}
-\providecommand\Pdom{{\bf P}}
-\providecommand\Sdom{{\bf S}}
-
-\providecommand\instM{{\bf instM}}
-\providecommand\statM{{\bf statM}}
-\providecommand\MH{{\bf MH}}
-\providecommand\MI{{\bf MI}}
-\providecommand\ME{{\bf ME}}
-
-\providecommand\cha{{\bf cha}}
-\providecommand\sub{{\bf sub}}
-\providecommand\HT{{\bf HT}}
-\providecommand\VT{{\bf VT}}
-\providecommand\EF{{\bf EF}}
-\providecommand\writeE{{\bf writeE}}
-
-\providecommand\virtIM{{\bf virtIM}}
-\providecommand\statIM{{\bf statIM}}
-\providecommand\specIM{{\bf specIM}}
-
-\providecommand\MmethArg{{\bf MmethArg}}
-\providecommand\MmethRet{{\bf MmethRet}}
-\providecommand\IinvkArg{{\bf IinvkArg}}
-\providecommand\IinvkRet{{\bf IinvkRet}}
-\providecommand\MordMethArg{{\bf MordMethArg}}
-\providecommand\MspcMethArg{{\bf MspcMethArg}}
-
-\providecommand\MobjVarAsgnInst{{\bf MobjVarAsgnInst}}
-\providecommand\MclsVarAsgnInst{{\bf MclsVarAsgnInst}}
-\providecommand\MobjValAsgnInst{{\bf MobjValAsgnInst}}
-\providecommand\MgetInstFldInst{{\bf MgetInstFldInst}}
-\providecommand\MputInstFldInst{{\bf MputInstFldInst}}
-\providecommand\MgetStatFldInst{{\bf MgetStatFldInst}}
-\providecommand\MputStatFldInst{{\bf MputStatFldInst}}
-
-\providecommand\syncMV{{\bf syncMV}}
-\providecommand\nestedLocks{{\bf nestedLocks}}
-
-\providecommand\VH{{\bf VH}}
-\providecommand\FH{{\bf FH}}
-\providecommand\HFH{{\bf HFH}}
-\providecommand\IM{{\bf IM}}
-
-\providecommand\forkC{{\bf forkC}}
-
-\providecommand\CVC{{\bf CVC}}
-\providecommand\FC{{\bf FC}}
-\providecommand\CFC{{\bf CFC}}
-\providecommand\CICM{{\bf CICM}}
-\providecommand\CHC{{\bf CHC}}
-
-\providecommand\clinitM{{\bf clinitM}}
-
-\providecommand\Chord{{Chord}}
-
-\providecommand\NameToTrgtMap{{\tt NameToTrgtMap}}
-\providecommand\NameToTaskMap{{\tt NameToTaskMap}}
-\providecommand\ConsumedNameToTasksMap{{\tt ConsumedNameToTasksMap}}
-\providecommand\TaskGraph{{\tt TaskGraph}}
-\providecommand\TaskMapEntries{{\tt TaskMapEntries}}
-\providecommand\TrgtMapEntries{{\tt TrgtMapEntries}}
-\providecommand\DlogMapEntries{{\tt DlogMapEntries}}
-\providecommand\Tasks{{\tt Tasks}}
-\providecommand\Trgts{{\tt Trgts}}
-\providecommand\ValidTasks{{\tt ValidTasks}}
-\providecommand\TrgtNames{{\tt TrgtNames}}
-
-\title{Chord: A Static Analysis Framework for Java}
-\author{Mayur Naik and Alex Aiken}
-\date{\today}
-\begin{document}
-\maketitle
-\texonly{\tableofcontents}
-
-\texonly{\newpage}
-\section{Overview}
-
-\Chord\ is a static analysis framework for Java with an emphasis on the analysis of objects and concurrency.
-It has the following characteristics:
-\begin{enumerate}
-\item
-It is {\it stand-alone} in that it provides various off-the-shelf static analyses (e.g., may alias analysis, thread escape analysis, datarace analysis, deadlock analysis, etc.).
-\item
-It is {\it extensible} in that it allows users to build their own static analyses on top of the provided static analyses.
-\item
-It is {\it compositional} in that it enables different static analyses to coexist and interact with each other.
-\item
-It is {\it efficient} in that it
-caches results computed by each static analysis so that they can be reused by other static analyses without re-computation.
-\item
-It is {\it flexible} in that it allows each static analysis to be expressed either declaratively as a Datalog program or imperatively as a Java program.
-\end{enumerate}
-
-\texonly{\newpage}
-\section{Download and Installation}
-
-Before installing \Chord\ ensure that the following software is installed on your machine:
-
-\begin{itemize}
-\item
-JDK 5 or higher (e.g., from
-Sun\footnote{\xlink{{\tt http://java.sun.com/javase/downloads/index.jsp}}{http://java.sun.com/javase/downloads/index.jsp}}
-or IBM\footnote{\xlink{{\tt http://www.ibm.com/developerworks/java/jdk/}}{http://www.ibm.com/developerworks/java/jdk/}}).
-
-\item
-Apache Ant (a Java build tool)\footnote{\xlink{{\tt http://ant.apache.org/}}{http://ant.apache.org/}}
-\end{itemize}
-
-Download the latest version of \Chord\ from \xlink{{\tt http://code.google.com/p/jchord/}}{http://code.google.com/p/jchord/}
-(say {\tt chord-x.y.tar.gz}) and uncompress it in a suitable directory by running the following command:
-
-\begin{verbatim}
-tar xvzf chord-x.y.tar.gz
-\end{verbatim}
-
-This will create a directory structure with root directory {\tt chord-x.y/}.
-
-\texonly{\newpage}
-\section{Building Chord}
-
-The root directory {\tt chord-x.y/} contains a {\tt build.xml} file which is interpreted by
-Apache Ant (a Java build tool).
-To see the various targets available run the following command in the root directory:
-
-\begin{verbatim}
-ant -p
-\end{verbatim}
-
-The default target {\tt compile} compiles \Chord.  So simply run the following command in the
-root directory:
-
-\begin{verbatim}
-ant
-\end{verbatim}
-
-This will compile the \Chord\ {\tt .java} source files and store the {\tt .class} files
-under the {\tt classes/} sub-directory of the root directory.
-
-\texonly{\newpage}
-\section{Running Chord}
-
-Suppose the program to be analyzed has the following directory structure:
-
-\begin{verbatim}
-example/
-    src/
-        foo/
-            Main.java
-            ...
-    classes/
-        foo/
-            Main.class
-            ...
-    lib/
-        src/
-            taz/
-                ...
-        jars/
-            taz.jar
-    chord/
-        chord.properties
-\end{verbatim}
-
-The above structure is typical: the program's Java source files are under {\tt example/src/}, its class files
-are under {\tt example/classes/}, the source and jar files of the libraries used by the program are under
-{\tt example/lib/src/} and {\tt example/lib/jars/}, respectively, and finally {\tt example/chord/} is the directory
-to which Chord will output its intermediate and final results.
-
-Execute Chord on the above program by running the
-following command in Chord's root directory {\tt chord-x.y/}:
-
-\begin{verbatim}
-ant -Dchord.work.dir=.../example/chord run
-\end{verbatim}
-
-where ``{\tt ...}'' denotes either the absolute path, or the path relative to Chord's root directory,
-of the parent of directory {\tt example/}.
-Chord will read the system properties specified in the file named {\tt chord.properties}
-that it presumes exists in the directory specified
-by system property {\tt chord.work.dir}
-which in the above case is directory {\tt example/chord/}.
-A sample such file for the above program is as follows:
-
-\begin{verbatim}
-chord.main.class=foo.Main
-chord.class.path=../classes:../lib/classes/taz.jar
-chord.src.path=../src:../lib/src
-chord.out=log.txt
-chord.err=log.txt
-chord.serial.file=program.ser
-chord.java.analysis.path=${chord.home.dir}/classes/main:
-    ${chord.home.dir}/classes/downcast:${chord.home.dir}/src/dlog/monosite
-chord.dlog.analysis.path=${chord.home.dir}/src/dlog/main:
-    ${chord.home.dir}/src/dlog/downcast:${chord.home.dir}/src/dlog/monosite
-chord.analyses=downcast-java,monosite-java
-\end{verbatim}
-
-System property {\tt chord.home.dir} specifies Chord's root directory
-{\tt chord-x.y/}.
-Every non-absolute location used in the value of any system property defined in the file
-is taken to be relative to the directory specified by system
-property {\tt chord.work.dir}.
-
-Section \ref{sec-sys-prop} presents an exhaustive
-listing of system properties that are recognized by \Chord.
-Briefly:
-\begin{itemize}
-\item
-{\tt chord.main.class} specifies the main class of the program
-to be analyzed.
-\item
-{\tt chord.class.path} specifies the application classpath of the program to be analyzed (as opposed to the
-JDK library classpath which consists of jar files such as {\tt rt.jar}
-which are implicitly appended to the program's application classpath to yield the
-complete program classpath).
-\item
-{\tt chord.out} and {\tt chord.err} specify the files (typically one and the same) to which the standard output and
-standard error streams, respectively, must be redirected while running Chord.
-\item
-{\tt chord.serial.file} specifies the location of a file from (resp. to) which
-the serialized form of the intermediate representation of the program to be analyzed must be loaded (resp. stored).
-This option is provided because loading the serialized form is much faster than
-constructing the intermediate representation of the program each time.
-\item
-{\tt chord.java.analysis.path} and {\tt chord.dlog.analysis.path} 
-specify all program domains, program relations, and program analyses
-to be included in the project and the
-dependencies between them (see Section \ref{sec-tg}).
-In the above case, they specify certain
- program domains, program relations, and program analyses
-pre-defined in \Chord, in particular, all those in (1) Chord's standard library, (2) Chord's downcast safety
-analysis, and (3) Chord's monomorphic call site analysis.
-Users can also define their own
-program domains, program relations, and program analyses and include them in the project.
-\item
-{\tt chord.analyses} is a space- or comma-separated list of the names of program domains,
-program relations, and program analyses that must be run in that order.  In the above case,
-these are downcast safety analysis, denoted {\tt downcast-java}, and monomorphic call site analysis,
-denoted {\tt monosite-java}.
-Both are program analyses pre-defined in \Chord\ 
-but users can also include the names
-of program domains, program relations, and program analyses they have defined and wish to run.
-\end{itemize}
-
-\texonly{\newpage}
-\section{Specifying Program Scope}
-
-\Chord\ takes as input a complete (i.e., closed) program with a main method.
-The scope of the program must be specified in one of the
-following ways:
-
-\begin{itemize}
-\item
-Defining system property {\tt chord.serial.file} as
-the location of a file containing in serialized form the intermediate representation
-of the program generated by a previous run of \Chord.
-\item
-Defining system property {\tt chord.main.class} as the fully-qualified name
-of the program's main class (e.g., {\tt foo.Main}) and system property
-{\tt chord.class.path} as the program's application classpath.
-\end{itemize}
-
-The former option is provided because loading the serialized form
-is much faster than constructing the intermediate representation
-of the program each time.
-
-\Chord\ first checks whether system property {\tt chord.serial.file} is defined.
-If it is defined, then it further checks whether the file specified by it
-exists.
-If the file exists, then \Chord\ treats the file as a serialized form of
-the intermediate representation of the program to be analyzed, and loads it.
-
-If system property {\tt chord.serial.file} is not defined or if the file
-specified by the property does not exist, then \Chord\ requires system properties
-{\tt chord.main.class} and {\tt chord.class.path} 
-to be defined, and uses them to
-build an intermediate representation of the program (see Section \ref{sec-prog-build})
-and perform source-to-source transformations on that representation
-(see Section \ref{sec-prog-transform}) that simplify several program analyses.
-Finally, if system property {\tt chord.serial.file} was defined but
-specified a non-existent file, then \Chord\ creates the file and stores the
-transformed intermediate representation of the program into the file in
-serialized form so that it can be loaded in subsequent runs.
-
-Defining system property {\tt chord.src.path} is optional.
-It specifies the program's Java source path which provides the
-locations of the program's {\tt .java} files so that, for instance, the
-results of an analysis (which operates on the intermediate representation of
-the program extracted from {\tt .class} files) can be mapped to the
-program's {\tt .java} files and presented to the user for
-manual inspection.
-
-\texonly{\newpage}
-\section{Program Representation}
-\label{sec-prog-build}
-
-Given the program's program's main class via system property {\tt chord.main.class} and the
-program's application classpath via system property {\tt chord.class.path},
-\Chord\ uses system property {\tt chord.program.builder.class} to locate the
-builder to be used for building an intermediate representation of the program,
-namely, an object of class {\tt chord.program.Program}.
-By default, the value of this property is class {\tt chord.program.builders.SootProgramBuilder}
-which uses the Soot Java compiler framework \cite{Soot}.
-Any builder, however, must implement interface
-{\tt chord.program.builders.IProgramBuilder}.
-
-The intermediate representation
-elides values of primitive type (e.g., int, boolean, etc.) and operations
-on them (e.g., arithmetic and relational operations) in the original program.
-The current emphasis of \Chord\ is on the analysis of objects
-but the analysis of primitives may be supported in future
-releases.
-
-The intermediate representation represents the program as a set of types.
-Each type has a set of (instance or static) fields and a set of
-(instance or static) methods.
-Each method has an optional control-flow graph.
-The control-flow graph has a unique entry point and
-a unique exit point\footnote{Even the control-flow graph of a method with an infinite loop
-has an exit point because loops are transformed into tail-recursive methods
-(see Section \ref{sec-prog-transform}).}.
-The entry and exit points are always distinct.
-Each point in the control-flow graph has one or more immediate predecessors
-(except the entry point which has none) and, likewise,
-one or more immediate successors (except the exit point which has none).
-Associated with each point is a simple statement, which may be of one of
-the following kinds:
-
-\begin{itemize}
-\item
-{\tt HeadInst}
-
-The statement associated with the unique entry point of a control-flow graph.
-
-It serves as a placeholder for:
-
-\begin{itemize}
-\item
-the (possibly empty) list of the method's {\it argument variables} of reference type;
-in the case of an instance method, this list is non-empty,
-with the $0^{th}$ element being the distinguished implicit {\tt this}
-argument variable.
-\item
-the (possibly empty) list of the method's {\it temporary variables} of reference type.
-\end{itemize}
-
-Argument variables and temporary variables of primitive type are
-elided due to \Chord's emphasis on the analysis of objects.
-
-Argument variables and temporary variables are distinct, namely,
-the set of a method's {\it local variables} (of reference type)
-is the disjoint union of the sets of the method's argument variables (of reference type) and
-temporary variables (of reference type).
-
-\item
-{\tt TailInst}
-
-The statement associated with the unique exit point of a control-flow graph.
-
-It serves as a placeholder for 
-the (possibly empty) list of the method's {\it return variables} of reference type.
-
-Return variables of primitive type are
-elided due to \Chord's emphasis on the analysis of objects.
-
-The set of a method's return variables (of reference type) is a subset of the set
-of the method's local variables (of reference type).
-
-A Java method has at most one return variable but
-\Chord's intermediate representation allows multiple return variables
-to enable compiling away exceptions, loops, synchronized blocks, etc.
-
-\item
-{\tt AcqLockInst}
-
-A statement of the form {\tt monitorenter v} where
-{\tt v} is a local variable of reference type.
-
-\item
-{\tt RelLockInst}
-
-A statement of the form {\tt monitorexit v} where {\tt v}
-is a local variable of reference type.
-
-\item
-{\tt StrValAsgnInst}
-
-A statement of the form {\tt v = "..."} where {\tt v} is
-a local variable of any reference type that can store a string
-constant.  It is typically (but not necessarily) of type
-{\tt java.lang.String}.
-
-\item
-{\tt NilValAsgnInst}
-
-A statement of the form {\tt v = null} where {\tt v}
-is a local variable of reference type.
-
-\item
-{\tt StatFldRefInst}
-
-A statement of one of the following forms:
-
-{\tt v = f} \\
-{\tt f = v} \\
-{\tt * = f} \\
-{\tt f = *}
-
-where {\tt v} is a local variable of reference type and
-{\tt f} is a static field.
-
-The {\tt *} denotes an expression of primitive type in the
-original program which is elided due to \Chord's emphasis on the
-analysis of objects.
-
-\item
-{\tt InstFldRefInst}
-
-A statement of one of the following forms:
-
-{\tt v = b.f} \\
-{\tt b.f = v} \\
-{\tt * = b.f} \\
-{\tt b.f = *}
-
-where {\tt v} and {\tt b} are local variables of reference
-type and {\tt f} is an instance field.
-
-The {\tt *} denotes an expression of primitive type in the
-original program which is elided due to \Chord's emphasis on the
-analysis of objects.
-
-\item
-{\tt SkipInst}
-
-A no-op statement.
-
-Due to \Chord's emphasis on the analysis of objects, each statement
-in the original program that does not access
-any value of reference type (i.e., only accesses values of primitive
-type) is represented by this statement.
-
-\item
-{\tt InvkInst}
-
-A method invocation statement of the form
-{\tt r1,...,rn = m(a1,...,ak)} where
-{\tt r1}, ..., {\tt rn} (return variables) and
-{\tt a1}, ..., {\tt ak} (argument variables)
-are local variables of reference type.
-
-Argument variables and return variables of primitive type are
-elided due to \Chord's emphasis on the analysis of objects.
-
-A method invocation statement in Java has at most
-one return variable but
-\Chord's intermediate representation allows multiple return variables
-to enable compiling away exceptions, loops, synchronized blocks, etc.
-
-\item
-{\tt ObjVarAsgnInst}
-
-A statement of the form {\tt v1 = v2} where {\tt v1}
-and {\tt v2} are local variables of (possibly different) reference types.
-
-\item
-{\tt ObjValAsgnInst}
-
-A statement of the form {\tt v = new t} where {\tt t} is a
-concrete reference type and
-{\tt v} is a local variable of a (possibly different) reference type.
-
-\item
-{\tt ClsVarAsgnInst}
-
-A statement of the form {\tt v = b.getClass()} where
-{\tt v} is a local variable of type {\tt java.lang.Class}
-and {\tt b} is a local variable of reference type.
-
-The reason this kind of statement is modeled is because
-\Chord\ has an emphasis on concurrency and many Java programs
-hold locks on the runtime class of an object
-(which is obtained by calling the native {\tt getClass}
-instance method of the object).
-
-\item
-{\tt AryElemRefInst}
-
-A statement of one of the following forms:
-
-{\tt v = b[*]}  \\
-{\tt b[*] = v} \\
-{\tt * = b[*]} \\
-{\tt b[*] = *}
-
-where {\tt v} and {\tt b} are local variables of reference
-type.
-
-The {\tt *} denotes an expression of primitive type in the
-original program which is elided due to \Chord's emphasis on the
-analysis of objects.
-\end{itemize}
-
-\texonly{\newpage}
-\section{Program Transformations}
-\label{sec-prog-transform}
-
-After building the intermediate representation of the
-program, \Chord\ performs
-the following source-to-source transformations 
-that simplify several program analyses:
-
-\begin{enumerate}
-\item
-{\bf Sync Removal}
-
-Removes all synchronized blocks from each method M in each class
-C in the program by generating a fresh static method (in class C)
-per synchronized block in the body of M.
-
-\item
-{\bf Loop Removal}
-
-Removes all loops from each method M in each class C in the program
-by generating a fresh static method (in class C)
-per loop in the body of M.
-
-\item
-{\bf Skip Removal}
-
-Removes all skip (i.e., no-op) statements from each method in each class
-in the program.
-Such statements arise primarily because \Chord, due to its emphasis
-on the analysis of objects, replaces each
-statement in the original program that does not access any
-value of reference type (i.e., only accesses values of primitive
-type) by a skip statement.
-
-\end{enumerate}
-
-\texonly{\newpage}
-\section{Task Graph Construction}
-\label{sec-tg}
-
-\Chord\ requires the following three system properties to construct the task graph:
-
-\begin{enumerate}
-\item
-{\tt chord.dlogmap.path}
-
-The path specifying the mapping from the name of each Datalog-based program
-analysis that must be included in the project to the location
-of the file defining that program analysis.
-
-Each path element must be the location of a file containing a
-list of entries (one per line) of the form:
-
-{\tt <name> <file-name>}
-
-The path elements must be separated by the system-specific path separator
-character (e.g., `:' on UNIX and `;' on Windows).
-
-Let \DlogMapEntries\ denote the set of all entries of the
-above form.
-
-\item
-{\tt chord.trgtmap.path}
-
-The path partially specifying the mapping from the name of each program domain,
-program relation, and Java-based program analysis that must be
-included in the project to the fully-qualified name of a subclass, either
-pre-defined in \Chord\ or defined by the user, of {\tt chord.Dom},
-{\tt chord.Ary*Rel}, and
-{\tt chord.Entity}, respectively, the instance fields of which define the
-state of the corresponding domain, relation, or analysis.
-The `{\tt *}' in {\tt chord.Ary*Rel} is 1, 2, ..., 6, or N for
-program relations of arity 1, 2, ..., 6, or $>$ 6, respectively.
-
-Each path element must be the location of a file containing a
-list of entries (one per line) of the form:
-
-{\tt <name> <class-name>}
-
-The path elements must be separated by the system-specific path separator
-character (e.g., `:` on UNIX and ';' on Windows).
-
-Let \TaskMapEntries\ denote the set of all entries of the
-above form.
-
-\item
-{\tt chord.taskmap.path}
-
-The path partially specifying the mapping from the name of each program domain,
-program relation, and Java-based program analysis that must be included in
-the project to the fully-qualified name of a
-subclass, either pre-defined in \Chord\ or defined by the user, of
-{\tt chord.Dom}, {\tt chord.Ary*Rel}, and {\tt chord.Entity}, respectively,
-which define the following instance methods:
-\begin{itemize}
-\item
-the {\tt execute} method specifying
-the computation that updates the state of the corresponding domain, relation,
-or analysis,
-\item
-the {\tt invalidate} method specifying the action to be taken when the
-state of the corresponding domain, relation, or analysis is invalidated.
-\item
-the {\tt getConsumedNames} method specifying names of any domains, relations,
-and Datalog-based or Java-based program analyses upon whose state the
-computation depends, and
-\item
-the {\tt getProducedNames} method specifying names of any domains, relations,
-and Datalog-based or Java-based program analyses whose state depends upon
-the computation.
-\end{itemize}
-
-Each path element must be the location of a file containing a
-list of entries (one per line) of the form:
-
-{\tt <name> <class-name>}
-
-The path elements must be separated by the system-specific path separator
-character (e.g., `:` on UNIX and ';' on Windows).
-
-Let \TrgtMapEntries\ denote the set of all entries of the above form.
-\end{enumerate}
-
-Then, \Chord\ uses the values of \DlogMapEntries, \TaskMapEntries, and \TrgtMapEntries\ 
-to do the following:
-
-\begin{enumerate}
-\item
-Compute {\tt Tasks\ :\ Set<chord.Entity>}, the set of all tasks,
-each of which is an object of (a subclass of) {\tt chord.Entity},
-as follows:
-
-\begin{enumerate}
-\item
-For each entry in \DlogMapEntries\ (i.e., a Datalog-based program analysis,
-there is a unique task, namely, a unique object of class {\tt chord.DatalogAnalysis}.
-
-\item
-For each entry in \TaskMapEntries\ (i.e., of the form {\tt <name> <class-name>} specifying
-a program domain, a program relation, or a Java-based program analysis), there is a
-unique task, namely, a unique object
-of the class named {\tt <class-name>} which must be (a subclass of) {\tt chord.Dom} in the
-case of a program domain, (a subclass of) {\tt chord.Ary*Rel} in the case of a
-program relation, and (a subclass of) {\tt chord.Entity} in the case of a Java-based
-program analysis.
-\end{enumerate}
-
-\item
-Compute the following maps as shown below:
-
-{\tt ConsumedNameToTasksMap\ :\ String -> Set<Task>},
-mapping each name to the set of (possibly empty) tasks that consume that name.
-
-{\tt NameToTaskMap\ :\ String -> Task},
-mapping each name to the unique task that produces that name.
-
-\Chord\ reports an error if a name is produced by no task or by more than one task.
-{\small 
-\begin{verbatim}
-for each (<task> in Tasks) {
-    case <task> came from item 1(a) above: {
-        // <task> corresponds to a Datalog-based program analysis
-        for each (<name> declared as input relation) {
-            add <task> to ConsumedNameToTasksMap(<name>)
-        }
-        for each (<name> declared as output relation) {
-            if (<name> is in domain of map NameToTaskMap)
-                error ``multiple tasks produce name <name>''
-            else
-                NameToTaskMap(<name>) = <task>
-        }
-    }
-    case <task> came from item 1(b) above: {
-        // <task> corresponds to a program domain, a program relation, or
-        // a Java-based program analysis
-        // Suppose <task> came from entry <name> <class-name> in
-        // TaskMapEntries
-        // i.e. <task> is an object of class named <class-name>
-        for each (<name2> in <task>.getConsumedNames())
-            add <task> to ConsumedNameToTaskMaps(<name2>)
-        for each (<name2> in <task>.getProducedNames() union { <name> })
-            if (<name2> is in domain of map NameToTaskMap)
-                error ``multiple tasks produce name <name>''
-            else
-                NameToTaskMap(<name2>) = <task>
-    }
-}
-\end{verbatim}
-}
-
-{\small 
-\begin{verbatim}
-for each (<name> in domain of map ConsumedNameToTasksMap) {
-    if (<name> is not in domain of map NameToTaskMap)
-        error ``no task produces name <name>''
-}
-\end{verbatim}
-}
-
-\item
-Compute \NameToTrgtMap : {\tt String -> Entity},
-mapping each name to a unique target, as below.
-\Chord\ reports an error if a name is mapped to multiple targets
-or to no target.
-
-{\small 
-\begin{verbatim}
-for each (entry <name> <class-name> in TaskMapEntries) {
-    if (<name> is in domain of map NameToTrgtMap)
-        error ``name <name> mapped to multiple targets''
-    NameToTrgtMap(<name>) = NameToTaskMap(<name>)
-}
-for each (entry <name> <class-name> in TrgtMapEntries) {
-    if (<name> is in domain of map NameToTrgtMap)
-        error ``name <name> mapped to multiple targets''
-    NameToTrgtMap(<name>) = a fresh object of class named <class-name>
-}
-\end{verbatim}
-}
-
-To reduce user burden in providing \TaskMapEntries\ and
-\TrgtMapEntries, \Chord\ implicitly extends \NameToTrgtMap\ using
-\DlogMapEntries\ as follows:
-
-{\small 
-\begin{verbatim}
-for each (<name> in domain of map NameToTaskMap) {
-    if (<name> is not in domain of map NameToTrgtMap) {
-        if (<name> is declared as a domain in some datalog file in DlogMapEntries)
-            NameToTrgtMap(<name>) = a fresh object of class chord.Dom
-        else if (<name> is declared as a (input or output) relation
-                in some datalog file in DlogMapEntries)
-            NameToTrgtMap(<name>) = a fresh object of class chord.Rel
-        else
-            error ``<name> not mapped to any target''
-    } else {
-        chord.Entity <trgt> = NameToTrgtMap(<name>)
-        if (<name> is declared as a domain in some datalog file in DlogMapEntries
-                and <trgt> is not of type (subclass of) chord.Dom)
-            error ``<name> has conflicting types''
-        else if (<name> is declared as a (input or output) relation
-                in some datalog file in DlogMapEntries
-                and <trgt> is not of type (subclass of) chord.Rel)
-            error ``<name> has conflicting types''
-    }
-}
-\end{verbatim}
-}
-
-\item
-Compute graph \TaskGraph\ as a graph whose:
-
-\begin{itemize}
-\item
-set of nodes is \Tasks\ and
-\item
-set of edges containing an edge from task {\tt <task1>} to task {\tt <task2>} 
-whenever there exists a name {\tt <name>} such that
-\NameToTaskMap({\tt <name>}) = {\tt <task1>} and
-\ConsumedNameToTasksMap({\tt <name>}) contains {\tt <task2>}.
-\end{itemize}
-
-The task graph in XML and HTML formats is dumped to files {\tt taskgraph.xml}
-and {\tt taskgraph.html} in the project's working directory specified using
-system property {\tt chord.work.dir}.
-
-\end{enumerate}
-
-\texonly{\newpage}
-\section{Project Organization}
-
-\Chord\ interprets system property {\tt chord.trgts}, if defined, as
-a space-separated list of names of program domains, program relations, and
-Datalog-based or Java-based program analyses, denoted \TrgtNames,
-to be computed in that order.
-
-To avoid redundant computation, \Chord\ maintains during its execution a set of
-up-to-date tasks, denoted \ValidTasks,
-which is a subset of the set of all tasks, denoted \Tasks.
-Set \ValidTasks\ is empty at startup.
-A future release of \Chord\ may provide the ability to retrieve a set of up-to-date tasks from
-a previous run of \Chord\ to prevent re-computing those tasks in subsequent runs.
-The following procedure is executed upon startup:
-
-{\small
-\begin{verbatim}
-for each (<name> in TrgtNames in order) {
-    if (<name> is not in domain of map NameToTaskMap) {
-        error ``name <name> is an invalid target name''
-    else {
-        chord.Entity <task> = NameToTaskMap(<name>)
-        executeTask(<task>)
-    }
-}
-\end{verbatim}
-}
-
-\noindent where procedure {\tt executeTask} is defined as follows:
-
-{\small
-\begin{verbatim}
-procedure executeTask(chord.Entity <task>) {
-    if (<task> is not in ValidTasks) {
-        for each (immed. pred. <task2> of <task> in TaskGraph)
-            executeTask(<task2>)
-        <task>.execute()
-        for each (immed. succ. <task2> of <task> in TaskGraph)
-            invalidateTask(<task2>)
-        add <task> to ValidTasks
-    }
-}
-\end{verbatim}
-}
-
-\noindent where procedure {\tt invalidateTask} is defined as follows:
-
-{\small
-\begin{verbatim}
-procedure invalidateTask(chord.Entity t) {
-    if (<task> is in ValidTasks) {
-        remove <task> from ValidTasks
-        <task>.invalidate()
-        for each (immed. succ. <task2> of <task> in TaskGraph)
-            invalidateTask(<task2>)
-    }
-}
-\end{verbatim}
-}
-
-\texonly{\newpage}
-\section{Chord System Properties}
-\label{sec-sys-prop}
-
-\Chord\ recognizes the following system properties:
-
-\begin{itemize}
-\item
-{\tt chord.out}
-
-The location of a file to which the standard output stream
-of \Chord's execution is redirected.
-
-It may be the same file to which the standard error stream is redirected (see
-the description of system property {\tt chord.err}).
-
-The file is created if it does not exist and its contents are overwritten if it exists.
-
-\item
-{\tt chord.err}
-
-The location of a file to which the standard error stream
-of \Chord's execution is redirected.
-
-It may be the same file to which the standard output stream is redirected (see
-the description of system property {\tt chord.out}).
-
-The file is created if it does not exist and its contents are overwritten if it exists.
-
-\item
-{\tt chord.program.builder.class}
-
-The fully-qualified name of the class to be used for building the
-intermediate representation of the program to be analyzed.
-It must be a concrete class with a nullary constructor that
-implements interface {\tt chord.program.builders.IProgramBuilder}.
-
-The default value of this property is {\tt chord.program.builders.SootProgramBuilder}
-which uses the Soot Java compiler framework \cite{Soot}.
-
-\item
-{\tt chord.serial.file}
-
-The location of a file from (resp.\ to) which
-the serialized form of the intermediate representation of the
-program to be analyzed is loaded (resp.\ stored).
-
-\item
-{\tt chord.main.class}
-
-The fully-qualified name of the main class (e.g., {\tt foo.Main}) of the
-program to be analyzed.
-
-This system property (and {\tt chord.class.path})
-must be defined if system property {\tt chord.serial.file}
-does not specify the file containing the serialized form of the intermediate representation
-of the program to be analyzed.
-
-\item
-{\tt chord.class.path}
-
-The application classpath of the program to be analyzed (as opposed to the
-JDK library classpath which consists of jar files such as {\tt rt.jar}
-which are implicitly appended to the program's application classpath to
-yield the complete program classpath).
-
-Each path element must be the location of a directory or a jar/zip file.
-
-The path elements must be separated by the system-specific path separator character
-(e.g., `:' on UNIX and `;' on Windows).
-
-This system property (and {\tt chord.main.class}) must
-be defined if system property {\tt chord.serial.file}
-does not specify the file containing the serialized form of the intermediate representation
-of the program to be analyzed.
-
-\item
-{\tt chord.src.path}
-
-The Java source path of the program to be analyzed.
-
-Each path element must be the location of a directory.
-
-The purpose of this path is to provide the location of the
-{\tt .java} files of the program being analyzed so that, for instance, the
-results of an analysis (which operates on an intermediate representation of the program
-extracted from {\tt .class} files) can be mapped to the corresponding {\tt .java}
-files and presented to the user for manual inspection.
-
-Each file having suffix {\tt .java} in a directory (or a sub-directory of a directory)
-specified in this path is regarded as a source file of the program.
-
-\item
-{\tt chord.dlog.analysis.path}
-
-The path specifying the mapping from the name of each Datalog-based program
-analysis that must be included in the project to the location
-of the file defining that program analysis.
-
-Each path element must be the location of a file containing a
-list of entries (one per line) of the form:
-
-{\tt <name> <file-name>}
-
-The path elements must be separated by the system-specific path separator character
-(e.g., `:' on UNIX and `;' on Windows).
-
-\item
-{\tt chord.java.analysis.path}
-
-The path specifying the mapping from the name of each program domain, program relation,
-and Java-based program analysis that must be included in the project to the
-fully-qualified name of a subclass, either pre-defined in \Chord\ or defined by the user, of
-{\tt chord.Dom}, {\tt chord.Ary*Rel}, and {\tt chord.Entity}, respectively, which
-define the following instance methods:
-\begin{itemize}
-\item
-the {\tt execute} method specifying
-the computation that updates the state of the corresponding domain, relation,
-or analysis, 
-\item
-the {\tt invalidate} method specifying the action to be taken when the
-state of the corresponding domain, relation, or analysis is invalidated.
-\item
-the {\tt getConsumedNames} method specifying names of any domains, relations,
-and analyses upon whose state the computation depends, and
-\item
-the {\tt getProducedNames} method specifying names of any domains, relations,
-and analyses whose state depends upon the computation.
-\end{itemize}
-
-Each path element must be the location of a file containing a
-list of entries (one per line) of the form:
-
-{\tt <name> <class-name>}
-
-The path elements must be separated by the system-specific path separator character
-(e.g., `:' on UNIX and `;' on Windows).
-
-\item
-{\tt chord.trgtmap.path}
-
-The path specifying the mapping from the name of each program domain,
-program relation, and Java-based program analysis that must be
-included in the project to the fully-qualified name of a subclass,
-either pre-defined in \Chord\ or defined by the user, of {\tt chord.Dom},
-{\tt chord.Ary*Rel}, and
-{\tt chord.Entity}, respectively, the instance fields of which define the
-state of the corresponding domain, relation, or analysis.
-
-Each path element in the list must be the location of a file containing a
-list of entries (one per line) of the form:
-
-{\tt <name> <class-name>}
-
-The path elements must be separated by the system-specific path separator character
-(e.g., `:` on UNIX and ';' on Windows).
-
-\item
-{\tt chord.analyses}
-
-A space-separated list of the names of program domains, program relations, and
-Java-based or Datalog-based program analyses that must be computed in that order.
-
-\item
-{\tt chord.ext.class.path}
-
-The classpath of user-defined program domains, user-defined program relations,
-and user-defined Java-based program analyses.
-It is implicitly appended to the classpath of pre-defined program domains,
-pre-defined program relations, and pre-defined Java-based program analyses
-to yield the complete runtime classpath of \Chord.
-
-Each path element must be the location of a directory or a jar/zip file.
-
-The path elements must be separated by the system-specific path separator character
-(e.g., `:' on UNIX and `;' on Windows).
-\end{itemize}
-
-\texonly{\newpage}
-\section{Pre-defined Domains}
-\label{sec-predef-domains}
-
-The following domains are pre-defined in \Chord:
-
-\begin{itemize}
-\item
-$\Cdom$
-
-The domain of all k-object-sensitive contexts.
-
-The $0^{th}$ element in this domain denotes a distinguished hypothetical
-context (denoted \texorhtml{$\epsilon$}{{\it epsilon}} in the literature) that is regarded as the
-lone abstract context in which certain static methods are analyzed.
-The main static method with signature {\tt void main(java.lang.String[])}
-has no callers and is therefore analyzed in this context.
-For simplicity, the k-object-sensitive may alias analysis provided in
-\Chord\ also analyzes each class initializer method in this context.
-
-\item
-$\Edom$
-
-The domain of all statements that access (read or write) an instance field,
-a static field, or an array element, namely, all statements
-represented by instances of (subclasses of) {\tt chord.program.insts.HeapInst}.
-
-\item
-$\Fdom$
-
-The domain of  all instance and static fields.
-
-The $0^{th}$ element in this domain denotes a distinguished hypothetical
-field that is regarded as accessed whenever an array element is
-accessed.
-
-\item
-$\Hdom$
-
-The domain of all object allocation statements, namely, all statements
-represented by instances of {\tt chord.program.insts.ObjValAsgnInst}.
-
-\item
-$\Idom$
-
-The domain of all method invocation statements, namely, all statements
-represented by instances of {\tt chord.program.insts.InvkInst}.
-
-\item
-$\Mdom$
-
-The domain of all instance and static methods.
-
-The $0^{th}$ element in this domain is the main method of the program.
-
-The $1^{st}$ element in this domain is the method with signature
-{\tt void start()} of class {\tt java.lang.Thread}, if this
-method is deemed reachable from the main method of the program.
-
-The above two methods are the root methods of the implicitly spawned
-main thread and each explicitly spawned child thread, respectively.
-Due to \Chord's emphasis on concurrency, these methods are
-referenced frequently by various pre-defined program analyses
-expressed in Datalog, and giving them special indices makes it
-convenient to reference them in those analyses.
-
-\item
-$\Pdom$
-
-The domain of all simple statements, namely, all statements
-represented by instances of (subclasses of) {\tt chord.program.insts.Inst}.
-
-The $0^{th}$ element in this domain is the statement at the unique
-entry point of the main method of the program.
-
-The statements of each method in the program are assigned
-contiguous indices in this domain, with the statements at the
-unique entry and exit points of each method being assigned
-the smallest and largest indices, respectively, of all
-indices assigned to statements in that method.
-
-\item
-$\Tdom$
-
-The domain of all types, including the types of classes, interfaces,
-arrays, and primitives.
-
-\item
-$\Vdom$
-
-The domain of all local variables.
-
-Each local variable declared in each block of each method is
-represented by a unique element in this domain, in particular,
-local variables that have the same name but are declared in
-different methods or in different blocks of the same method
-are represented by different elements in this domain.
-
-All local variables of the same method are assigned contiguous
-indices in this domain, in particular, the set of local
-variables of a method is the disjoint union of its argument
-variables and its temporary variables, and the argument
-variables are assigned contiguous indices in order followed by
-the temporary variables.
-
-\item
-$\Zdom$
-
-The domain of all argument and return variable positions.
-Let $N$ be the largest number or arguments or return variables of any
-method.  Then, this domain contains the $N$ elements $0, 1, ..., N-1$.
-\end{itemize}
-
-\texonly{\newpage}
-\section{Pre-defined Relations}
-\label{sec-predef-relations}
-
-The following relations are pre-defined in \Chord:
-
-\begin{itemize}
-
-\item
-$\cha : \Mdom \times \Tdom \times \Mdom$
-
-Relation containing each tuple $(m_1,t,m_2)$ such that $m_2$ is
-the target method of a method invocation statement whose {\it resolved method}
-is $m_1$ and whose receiver has run-time type $t$.
-It is named $\cha$ because it represents the relation
-computed by Class Hierarchy Analysis.
-
-\item
-$\sub : \Tdom \times \Tdom$
-
-Relation containing each tuple $(s,t)$ such that type $s$ is a
-subtype of type $t$.
-
-\item
-$\HT : \Hdom \times \Tdom$
-
-Relation containing each tuple $(h,t)$ such that object allocation
-statement $h$ allocates objects of type $t$.
-
-\item
-$\VT : \Vdom \times \Tdom$
-
-Relation containing each tuple $(v,t)$ such that local variable $v$
-has type $t$.
-
-\item
-$\MH : \Mdom \times \Hdom$
-
-Relation containing each tuple $(m,h)$ such that method $m$ contains
-object allocation statement $h$.
-
-\item
-$\MI : \Mdom \times \Idom$
-
-Relation containing each tuple $(m,i)$ such that method $m$ contains
-method invocation statement $i$.
-
-\item
-$\ME : \Mdom \times \Edom$
-
-Relation containing each tuple $(m,e)$ such that method $m$ contains
-statement $e$ that accesses (reads or writes) an instance field, a
-static field, or an array element.
-
-\item
-$\instM : \Mdom$
-
-Relation containing all instance (as opposed static) methods.
-
-\item
-$\statM : \Mdom$
-
-Relation containing all static (as opposed instance) methods.
-
-\item
-$\clinitM : \Mdom$
-
-Relation containing all class initializer methods, namely,
-methods having signature {\tt void <clinit>()}.
-
-\item
-$\IinvkArg : \Idom \times \Zdom \times \Vdom$
-
-Relation containing each tuple $(i,z,v)$ such that local variable $v$
-is the $z^{th}$ argument variable of method invocation statement $i$.
-
-\item
-$\IinvkRet : \Idom \times \Zdom \times \Vdom$
-
-Relation containing each tuple $(i,z,v)$ such that local variable $v$
-is the $z^{th}$ return variable of method invocation statement $i$.
-
-\item
-$\MmethArg : \Mdom \times \Zdom \times \Vdom$
-
-Relation containing each tuple $(m,z,v)$ such that local variable
-$v$ is the $z^{th}$ argument variable of method $m$.
-
-\item
-$\MmethRet : \Mdom \times \Zdom \times \Vdom$
-
-Relation containing each tuple $(m,z,v)$ such that local variable
-$v$ is the $z^{th}$ return variable of method $m$.
-
-\item
-$\MspcMethArg : \Mdom \times \Vdom$
-
-Relation containing each tuple $(m,v)$ such that local variable $v$ is
-the implicit {\tt this} argument variable of instance method $m$.
-
-\item
-
-$\MordMethArg : \Mdom \times \Zdom \times \Vdom$
-
-Relation containing each tuple $(m,z,v)$ such that local variable $v$
-is the $z^{th}$ argument variable of method $m$ other than the implicit
-{\tt this} argument variable (in the case in which $m$ is an instance
-method).
-
-\item
-$\MobjVarAsgnInst : \Mdom \times \Vdom \times \Vdom$
-
-Relation containing each tuple $(m,l,r)$ such that method $m$
-contains a copy statement of the form $l = r$ where $l$ and $r$ are
-local variables.
-
-\item
-$\MobjValAsgnInst : \Mdom \times \Vdom \times \Vdom$
-
-Relation containing each tuple $(m,l,h)$ such that method $m$ contains
-object allocation statement $h$ which assigns to local variable $l$.
-
-\item
-$\MclsVarAsgnInst : \Mdom \times \Vdom \times \Vdom$
-
-Relation containing each tuple $(m,l,b)$ such that method $m$
-contains a statement of the form $l = b.{\tt getClass()}$.
-
-\item
-$\MgetInstFldInst : \Mdom \times \Vdom \times \Fdom \times \Vdom$
-
-Relation containing each tuple $(m,b,f,r)$ such that method $m$
-contains a statement of the form $b.f = r$ which assigns to
-instance field or array element $f$ of the object denoted by local
-variable $b$ the value denoted by local variable $r$.
-
-\item
-$\MputInstFldInst : \Mdom \times \Vdom \times \Fdom \times \Vdom$
-
-Relation containing each tuple $(m,b,f,r)$ such that method $m$
-contains a statement of the form $b.f = r$ which writes the value
-denoted by local variable $r$ to instance field or array element
-$f$ of the object denoted by local variable $b$.
-
-\item
-$\MgetStatFldInst : \Mdom \times \Vdom \times \Fdom$
-
-Relation containing each tuple $(m,l,f)$ such that method $m$ contains
-a statement of the form $l = f$ that assigns to local variable $l$ the
-value denoted by static field $f$.
-
-\item
-$\MputStatFldInst : \Mdom \times \Fdom \times \Vdom$
-
-Relation containing each tuple $(m,f,r)$ such that method $m$ contains
-a statement of the form $f = r$ which assigns to static field $f$ the
-value denoted by local variable $r$.
-
-\item
-$\virtIM : \Idom \times \Mdom$
-
-Relation containing each tuple $(i,m)$ such that $m$ is the
-{\it resolved method} of {\tt invokevirtual} or {\tt invokeinterface} method
-invocation statement $i$.
-
-\item
-$\statIM : \Idom \times \Mdom$
-
-Relation containing each tuple $(i,m)$ such that $m$ is the
-{\it resolved method} of {\tt invokestatic} method invocation statement $i$.
-
-\item
-$\specIM : \Idom \times \Mdom$
-
-Relation containing each tuple $(i,m)$ such that $m$ is the
-{\it resolved method} of {\tt invokespecial} method invocation statement $i$.
-
-\item
-$\syncMV : \Mdom \times \Vdom$
-
-Relation containing each tuple $(m,v)$ such that method $m$ is
-synchronized on argument variable $v$.
-
-\item
-$\writeE : \Edom$
-
-Relation containing all statements that write (as opposed to read)
-an instance field, static field, or array element.
-
-\item
-$\EF : \Edom \times \Fdom$
-
-Relation containing each tuple $(e,f)$ such that statement $e$
-accesses (reads or writes) instance field, static field, or
-array element $f$.
-
-\item
-$\VH : \Vdom \times \Hdom$
-
-Relation containing each tuple $(v,h)$ such that local variable $v$
-may point to abstract object $h$, as determined by 0CFA-based may
-alias analysis.
-
-\item
-$\FH : \Fdom \times \Hdom$
-
-Relation containing each tuple $(f,h)$ such that static field $f$ may
-point to abstract object $h$, as determined by 0CFA-based may alias
-analysis.
-
-\item
-
-$\HFH : \Hdom \times \Fdom \times \Hdom$
-
-Relation containing each tuple $(h_1,f,h_2)$ such that instance field
-or array element $f$ of abstract object $h_1$ may point to abstract
-object $h_2$, as determined by 0CFA-based may alias analysis.
-
-\item
-$\IM : \Idom \times \Mdom$
-
-Relation containing each tuple $(i,m)$ such that method invocation
-statement $i$ may call method $m$ as determined by 0CFA-based may alias
-analysis.
-
-\item
-$\nestedLocks : \Cdom \times \Cdom \times \Mdom \times \Cdom \times \Mdom$
-
-Relation containing each tuple $(c,c_1,m_1,c_2,m_2)$ such that in the
-context-sensitive call graph computed by k-object-sensitive may
-alias analysis, abstract thread $c$ may first acquire the lock at
-the entry of method $m_1$ in abstract context $c_1$ and without
-releasing it proceed to acquire the lock at the entry of method
-$m_2$ in abstract context $c_2$.
-
-\item
-$\forkC : \Cdom$
-
-Relation containing all abstract contexts in which the
-{\tt start()} method of class {\tt java.lang.Thread} may
-be reachable from the main method, as determined by
-k-object-sensitive may alias analysis.
-
-\item
-$\CVC : \Cdom \times \Vdom \times \Cdom$
-
-Relation containing each tuple $(c_1,v,c_2)$ such that local variable
-$v$ may point to abstract object $c_2$ in abstract context $c_1$ of its
-containing method, as determined by k-object-sensitive may alias
-analysis.
-
-\item
-$\FC : \Fdom \times \Cdom$
-
-Relation containing each tuple $(f,c)$ such that static field $f$
-may point to abstract object $c$, as determined by k-object-sensitive
-may alias analysis.
-
-\item
-$\CFC : \Cdom \times \Fdom \times \Cdom$
-
-Relation containing each tuple $(c_1,f,c_2)$ such that instance field
-or array element $f$ of abstract object $c_1$ may point to abstract
-object $c_2$, as determined by k-object-sensitive may alias analysis.
-
-\item
-$\CICM : \Cdom \times \Idom \times \Cdom \times \Mdom$
-
-Relation containing each tuple $(c_1,i,c_2,m)$ such that method
-invocation statement $i$ in abstract context $c_1$ of its containing method
-may call method $m$ in abstract context $c_2$, as determined by
-k-object-sensitive may alias analysis.
-
-\item
-$\CHC : \Cdom \times \Hdom \times \Cdom$
-
-Relation containing each tuple $(c_1,h,c_2)$ such that the method
-containing object allocation statement $h$ may be reachable in
-abstract context $c_1$, and $c_2 = (h::c_1) + k$, where $k$ is the k
-value of statement $h$, provided as an argument to k-object-sensitive
-may alias analysis.
-
-\end{itemize}
-
-\texonly{\newpage}
-\section{Pre-defined Datalog-based Analyses}
-\label{sec-predef-datalog-analyses}
-
-\texonly{\newpage}
-\section{Pre-defined Java-based Analyses}
-\label{sec-predef-java-analyses}
-
-\texonly{\newpage}
-\bibliographystyle{plain}
-\bibliography{citations}
-
-\end{document}
-
-\newpage
-\section{Defining Tasks and Targets}
-
-\Chord\ uses the notion of {\it tasks} to express computations
-and the notion of {\it targets} to express the structure of
-inputs/outputs of computations.
-
-Tasks and targets are both implemented uniformly as {\it entities}.
-A generic entity is represented by class {\tt chord.Entity}
-depicted in Figure \ref{fig-entity}.
-
-\begin{figure}
-\begin{verbatim}
-public class Entity {
-    protected static final String[] emptyStringArray = new String[0];
-    // subclasses may override
-    public void execute() {
-        // do nothing
-    }
-    // subclasses may override
-    public String[] getConsumedTargetNames() {
-        return emptyStringArray;
-    }
-    // subclasses may override
-    public String[] getProducedTargetNames() {
-        return emptyStringArray;
-    }
-    ...
-}
-\end{verbatim}
-\caption{Snippet of class {\tt chord.Entity}}
-\label{fig-entity}
-\end{figure}
-
-Whenever an enntity is regarded as a task:
-
-\begin{itemize}
-\item
-The {\tt execute} method defines the computation that must be performed by the task.
-
-\item
-The {\tt getConsumedTargetNames} method specifies the names of targets
-that are inputs to (the computation performed by) the task.
-
-\item
-The {\tt getProducedTargetNames} method specifies the names of targets
-that are outputs of (the computation performed by) the task.
-\end{itemize}
-
-There are four kinds of entities:
-\begin{itemize}
-\item
-A program domain
-
-\item
-A program relation
-
-\item
-A program analysis expressed in Datalog
-
-\item
-A program analysis expressed in Java
-
-\end{itemize}
-
-We next describe each of these kinds of entities in detail.
-
-\subsection{Program Domains}
-
-A program domain is a set of related elements in a given program (e.g.,
-the set of all methods, the set of all object allocation statements, etc.).
-The structure of a generic program domain is represented by
-class {\tt chord.Dom} depicted in Figure \ref{fig-dom}.
-
-\begin{figure}
-\begin{verbatim}
-public class Dom<T> extends Entity {
-    public final IndexMap<T> map;
-    public Dom() {
-        this.map = new IndexMap<T>();
-    }
-    public void execute() {
-        populate();
-        saveToDomFile();
-    }
-    // subclasses cannot override
-    public final String[] getProducedTargetNames() {
-        return emptyStringArray;
-    }
-    // subclasses may override
-    protected void populate() {
-        // do nothing
-    }
-    ...
-\end{verbatim}
-\caption{Snippet of class {\tt chord.Dom}}
-\label{fig-dom}
-\end{figure}
-
-A program domain provides a map from each element in the domain
-to a unique integer in the range $[0..n-1]$ where $n$ is the size of
-the domain.
-This map is required by {\tt bddbddb} (a BDD-based Datalog solver).
-The {\tt execute} method is specialized to construct the map and then
-save the map to disk for use by {\tt bddbddb}.
-Subclasses must override the {\tt populate} method to construct the
-map as desired.
-
-It is not possible to override the {\tt getProducedTargetNames}
-method (note that it is declared {\tt final}) because a program domain
-task is a specialized task whose output is exactly one target, namely,
-the program domain it constructs.
-It is, however, possible to override the {\tt getConsumedTargetNames}
-method (inherited from superclass {\tt chord.Entity}) since constructing
-a program domain may require other targets (i.e., other program
-domains, program relations, or program analyses).
-
-Figure \ref{fig-domI} defines the program domain
-named $\Idom$ as the domain of all method invocation statements.
-Note from method {\tt getConsumedTargetNames} that it takes as input
-the domain named $\Mdom$ which is the domain of all methods.
-This is because the {\tt populate} method refers to that domain.
-Intuitively, the {\tt populate} method gives adjacent indices in
-domain $\Idom$ to method invocation statements contained in
-methods that have adjacent indices in domain $\Mdom$ are given
-adjacent indices in domain $\Idom$.
-
-\begin{figure}
-\begin{verbatim}
-public class DomI extends Dom<InvkInst> {
-    private IndexMap<Method> mapM;
-    private final String[] consumedTaskNames = new String[] { "M" };
-    public String[] getConsumedTargetNames() {
-        return consumedTaskNames;
-    }
-    protected void populate() {
-        DomM domM = (DomM) project.getEntity("M");
-        mapM = domM.map;
-        for (int m = 0; m < mapM.size(); m++) {
-            Method meth = mapM.get(m);
-            Body body = meth.body;
-            if (body == null)
-                continue;
-            Set<Inst> insts = body.cfg.getNodes();
-            for (Inst inst : insts) {
-                if (inst instanceof InvkInst)
-                    map.set((InvkInst) inst);
-            }
-        }
-    }
-    ...
-}
-\end{verbatim}
-\caption{Snippet of class {\tt chord.DomI}}
-\label{fig-domI}
-\end{figure}
-
-A list of pre-defined program domains is presented in Section \ref{sec-predef-domains}.
-
-\subsection{Program Relations}
-
-A program relation is a relation over one or more program domains.
-The structure of a generic program relation is represented by class
-{\tt chord.Rel} depicted in Figure \ref{fig-rel}.
-
-\begin{figure}
-\begin{verbatim}
-\end{verbatim}
-\caption{Snippet of class {\tt chord.Rel}}
-\label{fig-rel}
-\end{figure}
-
-It is not possible to override the {\tt getProducedTargetNames}
-method (note that it is declared {\tt final}) because a program relation
-task is a specialized task whose output is exactly one target, namely,
-the program relation it constructs.
-It is, however, possible to override the {\tt getConsumedTargetNames}
-method (inherited from superclass {\tt chord.Entity}) since constructing
-a program relation may require other targets (i.e., program
-domains, other program relations, or program analyses).
-
-A list of pre-defined program relations is presented in Section \ref{sec-predef-relations}.
-
-\subsection{Program Analyses in Datalog}
-
-A list of pre-defined program analyses in Datalog is presented in
-Section \ref{sec-predef-datalog-analyses}.
-
-\subsection{Program Analyses in Java}
-
-A list of pre-defined program analyses in Java is presented in
-Section \ref{sec-predef-java-analyses}.
+TODO

### 2010-04-01 main/doc/program_setup.tex

--- 2009-07-01 doc/user_guide.tex
@@
-\documentclass{article}
-\usepackage{hyperlatex}
-
-%\topmargin=-0.5in
-\texonly{\oddsidemargin=0in}
-\texonly{\evensidemargin=0in}
-\texonly{\textwidth=5.6in}
-%\textheight=8.5in
-
-%TODO: specify use of BDD's under efficiency in section 1 (overview)
-
-\htmlmathitalic{0}
-\htmltitle{User Guide}
-\htmladdress{chord@cs.stanford.edu, \today}
-
-\providecommand\Mdom{{\bf M}}
-\providecommand\Idom{{\bf I}}
-\providecommand\Vdom{{\bf V}}
-\providecommand\Zdom{{\bf Z}}
-\providecommand\Hdom{{\bf H}}
-\providecommand\Edom{{\bf E}}
-\providecommand\Fdom{{\bf F}}
-\providecommand\Cdom{{\bf C}}
-\providecommand\Tdom{{\bf T}}
-\providecommand\Pdom{{\bf P}}
-\providecommand\Sdom{{\bf S}}
-
-\providecommand\instM{{\bf instM}}
-\providecommand\statM{{\bf statM}}
-\providecommand\MH{{\bf MH}}
-\providecommand\MI{{\bf MI}}
-\providecommand\ME{{\bf ME}}
-
-\providecommand\cha{{\bf cha}}
-\providecommand\sub{{\bf sub}}
-\providecommand\HT{{\bf HT}}
-\providecommand\VT{{\bf VT}}
-\providecommand\EF{{\bf EF}}
-\providecommand\writeE{{\bf writeE}}
-
-\providecommand\virtIM{{\bf virtIM}}
-\providecommand\statIM{{\bf statIM}}
-\providecommand\specIM{{\bf specIM}}
-
-\providecommand\MmethArg{{\bf MmethArg}}
-\providecommand\MmethRet{{\bf MmethRet}}
-\providecommand\IinvkArg{{\bf IinvkArg}}
-\providecommand\IinvkRet{{\bf IinvkRet}}
-\providecommand\MordMethArg{{\bf MordMethArg}}
-\providecommand\MspcMethArg{{\bf MspcMethArg}}
-
-\providecommand\MobjVarAsgnInst{{\bf MobjVarAsgnInst}}
-\providecommand\MclsVarAsgnInst{{\bf MclsVarAsgnInst}}
-\providecommand\MobjValAsgnInst{{\bf MobjValAsgnInst}}
-\providecommand\MgetInstFldInst{{\bf MgetInstFldInst}}
-\providecommand\MputInstFldInst{{\bf MputInstFldInst}}
-\providecommand\MgetStatFldInst{{\bf MgetStatFldInst}}
-\providecommand\MputStatFldInst{{\bf MputStatFldInst}}
-
-\providecommand\syncMV{{\bf syncMV}}
-\providecommand\nestedLocks{{\bf nestedLocks}}
-
-\providecommand\VH{{\bf VH}}
-\providecommand\FH{{\bf FH}}
-\providecommand\HFH{{\bf HFH}}
-\providecommand\IM{{\bf IM}}
-
-\providecommand\forkC{{\bf forkC}}
-
-\providecommand\CVC{{\bf CVC}}
-\providecommand\FC{{\bf FC}}
-\providecommand\CFC{{\bf CFC}}
-\providecommand\CICM{{\bf CICM}}
-\providecommand\CHC{{\bf CHC}}
-
-\providecommand\clinitM{{\bf clinitM}}
-
-\providecommand\Chord{{Chord}}
-
-\providecommand\NameToTrgtMap{{\tt NameToTrgtMap}}
-\providecommand\NameToTaskMap{{\tt NameToTaskMap}}
-\providecommand\ConsumedNameToTasksMap{{\tt ConsumedNameToTasksMap}}
-\providecommand\TaskGraph{{\tt TaskGraph}}
-\providecommand\TaskMapEntries{{\tt TaskMapEntries}}
-\providecommand\TrgtMapEntries{{\tt TrgtMapEntries}}
-\providecommand\DlogMapEntries{{\tt DlogMapEntries}}
-\providecommand\Tasks{{\tt Tasks}}
-\providecommand\Trgts{{\tt Trgts}}
-\providecommand\ValidTasks{{\tt ValidTasks}}
-\providecommand\TrgtNames{{\tt TrgtNames}}
-
-\title{Chord: A Static Analysis Framework for Java}
-\author{Mayur Naik and Alex Aiken}
-\date{\today}
-\begin{document}
-\maketitle
-\texonly{\tableofcontents}
-
-\texonly{\newpage}
-\section{Overview}
-
-\Chord\ is a static analysis framework for Java with an emphasis on the analysis of objects and concurrency.
-It has the following characteristics:
-\begin{enumerate}
-\item
-It is {\it stand-alone} in that it provides various off-the-shelf static analyses (e.g., may alias analysis, thread escape analysis, datarace analysis, deadlock analysis, etc.).
-\item
-It is {\it extensible} in that it allows users to build their own static analyses on top of the provided static analyses.
-\item
-It is {\it compositional} in that it enables different static analyses to coexist and interact with each other.
-\item
-It is {\it efficient} in that it
-caches results computed by each static analysis so that they can be reused by other static analyses without re-computation.
-\item
-It is {\it flexible} in that it allows each static analysis to be expressed either declaratively as a Datalog program or imperatively as a Java program.
-\end{enumerate}
-
-\texonly{\newpage}
-\section{Download and Installation}
-
-Before installing \Chord\ ensure that the following software is installed on your machine:
-
-\begin{itemize}
-\item
-JDK 5 or higher (e.g., from
-Sun\footnote{\xlink{{\tt http://java.sun.com/javase/downloads/index.jsp}}{http://java.sun.com/javase/downloads/index.jsp}}
-or IBM\footnote{\xlink{{\tt http://www.ibm.com/developerworks/java/jdk/}}{http://www.ibm.com/developerworks/java/jdk/}}).
-
-\item
-Apache Ant (a Java build tool)\footnote{\xlink{{\tt http://ant.apache.org/}}{http://ant.apache.org/}}
-\end{itemize}
-
-Download the latest version of \Chord\ from \xlink{{\tt http://code.google.com/p/jchord/}}{http://code.google.com/p/jchord/}
-(say {\tt chord-x.y.tar.gz}) and uncompress it in a suitable directory by running the following command:
-
-\begin{verbatim}
-tar xvzf chord-x.y.tar.gz
-\end{verbatim}
-
-This will create a directory structure with root directory {\tt chord-x.y/}.
-
-\texonly{\newpage}
-\section{Building Chord}
-
-The root directory {\tt chord-x.y/} contains a {\tt build.xml} file which is interpreted by
-Apache Ant (a Java build tool).
-To see the various targets available run the following command in the root directory:
-
-\begin{verbatim}
-ant -p
-\end{verbatim}
-
-The default target {\tt compile} compiles \Chord.  So simply run the following command in the
-root directory:
-
-\begin{verbatim}
-ant
-\end{verbatim}
-
-This will compile the \Chord\ {\tt .java} source files and store the {\tt .class} files
-under the {\tt classes/} sub-directory of the root directory.
-
-\texonly{\newpage}
-\section{Running Chord}
+\section{Setting up a Java Program for Analysis}
+\label{sec:program-setup}
 
@@
                 ...
-        jars/
+        jar/
             taz.jar
-    chord/
     chord.properties
 \end{verbatim}
+\end{quote}
 
-The above structure is typical: the program's Java source files are under {\tt example/src/}, its class files
-are under {\tt example/classes/}, the source and jar files of the libraries used by the program are under
-{\tt example/lib/src/} and {\tt example/lib/jars/}, respectively, and finally {\tt example/chord/} is the directory
-to which Chord will output its intermediate and final results.
+The above structure is typical: the program's Java source files are under \verb+src/+,
+its class files are under \verb+classes/+, the source and jar files of the libraries
+used by the program are under \verb+lib/src/+ and \verb+lib/jar/+, respectively.
 
-Execute Chord on the above program by running the
-following command in Chord's root directory {\tt chord-x.y/}:
-
-\begin{verbatim}
-ant -Dchord.work.dir=.../example/chord run
-\end{verbatim}
-
-where ``{\tt ...}'' denotes either the absolute path, or the path relative to Chord's root directory,
-of the parent of directory {\tt example/}.
-Chord will read the system properties specified in the file named {\tt chord.properties}
-that it presumes exists in the directory specified
-by system property {\tt chord.work.dir}
-which in the above case is directory {\tt example/chord/}.
+File \verb+chord.properties+ specifies properties to be passed to Chord (alternatively,
+these properties may be passed on the command-line, in format \verb+-D<name>=<value>+).
 A sample such file for the above program is as follows:
 
+\begin{quote}
 \begin{verbatim}
 chord.main.class=foo.Main
-chord.class.path=../classes:../lib/classes/taz.jar
-chord.src.path=../src:../lib/src
-chord.out=log.txt
-chord.err=log.txt
-chord.serial.file=program.ser
-chord.java.analysis.path=${chord.home.dir}/classes/main:
-    ${chord.home.dir}/classes/downcast:${chord.home.dir}/src/dlog/monosite
-chord.dlog.analysis.path=${chord.home.dir}/src/dlog/main:
-    ${chord.home.dir}/src/dlog/downcast:${chord.home.dir}/src/dlog/monosite
-chord.analyses=downcast-java,monosite-java
+chord.class.path=classes:lib/jar/taz.jar
+chord.src.path=src:../lib/src
+chord.run.ids=0,1
+chord.args.0="-thread 1 -n 10"
+chord.args.1="-thread 2 -n 50"
 \end{verbatim}
+\end{quote}
 
-System property {\tt chord.home.dir} specifies Chord's root directory
-{\tt chord-x.y/}.
-Every non-absolute location used in the value of any system property defined in the file
-is taken to be relative to the directory specified by system
-property {\tt chord.work.dir}.
+Each relative (instead of absolute) path element in the value of any property
+named \verb+chord.<...>.path+ defined in this file is converted to an absolute path element with respect
+to the directory containing this file, which in the above case is \verb+example/+.
 
-Section \ref{sec-sys-prop} presents an exhaustive
-listing of system properties that are recognized by \Chord.
-Briefly:
-\begin{itemize}
-\item
-{\tt chord.main.class} specifies the main class of the program
-to be analyzed.
-\item
-{\tt chord.class.path} specifies the application classpath of the program to be analyzed (as opposed to the
-JDK library classpath which consists of jar files such as {\tt rt.jar}
-which are implicitly appended to the program's application classpath to yield the
-complete program classpath).
-\item
-{\tt chord.out} and {\tt chord.err} specify the files (typically one and the same) to which the standard output and
-standard error streams, respectively, must be redirected while running Chord.
-\item
-{\tt chord.serial.file} specifies the location of a file from (resp. to) which
-the serialized form of the intermediate representation of the program to be analyzed must be loaded (resp. stored).
-This option is provided because loading the serialized form is much faster than
-constructing the intermediate representation of the program each time.
-\item
-{\tt chord.java.analysis.path} and {\tt chord.dlog.analysis.path} 
-specify all program domains, program relations, and program analyses
-to be included in the project and the
-dependencies between them (see Section \ref{sec-tg}).
-In the above case, they specify certain
- program domains, program relations, and program analyses
-pre-defined in \Chord, in particular, all those in (1) Chord's standard library, (2) Chord's downcast safety
-analysis, and (3) Chord's monomorphic call site analysis.
-Users can also define their own
-program domains, program relations, and program analyses and include them in the project.
-\item
-{\tt chord.analyses} is a space- or comma-separated list of the names of program domains,
-program relations, and program analyses that must be run in that order.  In the above case,
-these are downcast safety analysis, denoted {\tt downcast-java}, and monomorphic call site analysis,
-denoted {\tt monosite-java}.
-Both are program analyses pre-defined in \Chord\ 
-but users can also include the names
-of program domains, program relations, and program analyses they have defined and wish to run.
-\end{itemize}
-
-\texonly{\newpage}
-\section{Specifying Program Scope}
-
-\Chord\ takes as input a complete (i.e., closed) program with a main method.
-The scope of the program must be specified in one of the
-following ways:
+Section \ref{sec:chord-sysprops} presents an exhaustive
+listing of properties recognized by \Chord.
+Here, we only describe those defined in the above
+sample \verb+chord.properties+ file:
 
@@
 \item
-Defining system property {\tt chord.serial.file} as
-the location of a file containing in serialized form the intermediate representation
-of the program generated by a previous run of \Chord.
+\verb+chord.main.class+ specifies the fully-qualified name of the main class of the
+program.
 \item
-Defining system property {\tt chord.main.class} as the fully-qualified name
-of the program's main class (e.g., {\tt foo.Main}) and system property
-{\tt chord.class.path} as the program's application classpath.
+\verb+chord.class.path+ specifies the classpath of the program.
+The value of pre-defined property \verb+${sun.boot.class.path}+ is implicitly appended
+to it.
+\item
+\verb+chord.src.path+ specifies the Java source path of the program.
+All program analyses in Chord operate on Java bytecode (specified by \verb+chord.class.path+).
+The only use of this property is to HTMLize the Java source files of the program so that the
+results of program analyses can be reported at the Java source code level.
+\item
+\verb+chord.run.ids+ specifies a list of IDs to identify runs of the program.
+It is used by dynamic program analyses to determine how many times the program must be run.
+An additional use of this property is to allow specifying the command-line arguments to use
+in the run having ID \verb+XXX+ via property \verb+chord.args.XXX+, as illustrated by
+properties \verb+chord.args.0+ and \verb+chord.args.1+ in the above example.
 \end{itemize}
 
-The former option is provided because loading the serialized form
-is much faster than constructing the intermediate representation
-of the program each time.
+To run Chord on the above program, run the following command in Chord's \verb+main/+ directory:
 
-\Chord\ first checks whether system property {\tt chord.serial.file} is defined.
-If it is defined, then it further checks whether the file specified by it
-exists.
-If the file exists, then \Chord\ treats the file as a serialized form of
-the intermediate representation of the program to be analyzed, and loads it.
+\begin{quote}
+\begin{verbatim}
+prompt> ant -Dchord.work.dir=<...>/example run
+\end{verbatim}
+\end{quote}
 
-If system property {\tt chord.serial.file} is not defined or if the file
-specified by the property does not exist, then \Chord\ requires system properties
-{\tt chord.main.class} and {\tt chord.class.path} 
-to be defined, and uses them to
-build an intermediate representation of the program (see Section \ref{sec-prog-build})
-and perform source-to-source transformations on that representation
-(see Section \ref{sec-prog-transform}) that simplify several program analyses.
-Finally, if system property {\tt chord.serial.file} was defined but
-specified a non-existent file, then \Chord\ creates the file and stores the
-transformed intermediate representation of the program into the file in
-serialized form so that it can be loaded in subsequent runs.
+where \verb+<...>+ denotes either the absolute path, or the path relative to Chord's main directory,
+of the parent of directory {\tt example/}.
+System property \verb+chord.work.dir+ specifies the working directory during Chord's execution.
 
-Defining system property {\tt chord.src.path} is optional.
-It specifies the program's Java source path which provides the
-locations of the program's {\tt .java} files so that, for instance, the
-results of an analysis (which operates on the intermediate representation of
-the program extracted from {\tt .class} files) can be mapped to the
-program's {\tt .java} files and presented to the user for
-manual inspection.
+The above command causes Chord to load the \verb+chord.properties+ file if present.
+The location of this file may be specified explicitly via property
+\verb+chord.props.file+ on the above command line.
+By default, it is \verb+${chord.work.dir}/chord.properties+.
 
-\texonly{\newpage}
-\section{Program Representation}
-\label{sec-prog-build}
-
-Given the program's program's main class via system property {\tt chord.main.class} and the
-program's application classpath via system property {\tt chord.class.path},
-\Chord\ uses system property {\tt chord.program.builder.class} to locate the
-builder to be used for building an intermediate representation of the program,
-namely, an object of class {\tt chord.program.Program}.
-By default, the value of this property is class {\tt chord.program.builders.SootProgramBuilder}
-which uses the Soot Java compiler framework \cite{Soot}.
-Any builder, however, must implement interface
-{\tt chord.program.builders.IProgramBuilder}.
-
-The intermediate representation
-elides values of primitive type (e.g., int, boolean, etc.) and operations
-on them (e.g., arithmetic and relational operations) in the original program.
-The current emphasis of \Chord\ is on the analysis of objects
-but the analysis of primitives may be supported in future
-releases.
-
-The intermediate representation represents the program as a set of types.
-Each type has a set of (instance or static) fields and a set of
-(instance or static) methods.
-Each method has an optional control-flow graph.
-The control-flow graph has a unique entry point and
-a unique exit point\footnote{Even the control-flow graph of a method with an infinite loop
-has an exit point because loops are transformed into tail-recursive methods
-(see Section \ref{sec-prog-transform}).}.
-The entry and exit points are always distinct.
-Each point in the control-flow graph has one or more immediate predecessors
-(except the entry point which has none) and, likewise,
-one or more immediate successors (except the exit point which has none).
-Associated with each point is a simple statement, which may be of one of
-the following kinds:
-
-\begin{itemize}
-\item
-{\tt HeadInst}
-
-The statement associated with the unique entry point of a control-flow graph.
-
-It serves as a placeholder for:
-
-\begin{itemize}
-\item
-the (possibly empty) list of the method's {\it argument variables} of reference type;
-in the case of an instance method, this list is non-empty,
-with the $0^{th}$ element being the distinguished implicit {\tt this}
-argument variable.
-\item
-the (possibly empty) list of the method's {\it temporary variables} of reference type.
-\end{itemize}
-
-Argument variables and temporary variables of primitive type are
-elided due to \Chord's emphasis on the analysis of objects.
-
-Argument variables and temporary variables are distinct, namely,
-the set of a method's {\it local variables} (of reference type)
-is the disjoint union of the sets of the method's argument variables (of reference type) and
-temporary variables (of reference type).
-
-\item
-{\tt TailInst}
-
-The statement associated with the unique exit point of a control-flow graph.
-
-It serves as a placeholder for 
-the (possibly empty) list of the method's {\it return variables} of reference type.
-
-Return variables of primitive type are
-elided due to \Chord's emphasis on the analysis of objects.
-
-The set of a method's return variables (of reference type) is a subset of the set
-of the method's local variables (of reference type).
-
-A Java method has at most one return variable but
-\Chord's intermediate representation allows multiple return variables
-to enable compiling away exceptions, loops, synchronized blocks, etc.
-
-\item
-{\tt AcqLockInst}
-
-A statement of the form {\tt monitorenter v} where
-{\tt v} is a local variable of reference type.
-
-\item
-{\tt RelLockInst}
-
-A statement of the form {\tt monitorexit v} where {\tt v}
-is a local variable of reference type.
-
-\item
-{\tt StrValAsgnInst}
-
-A statement of the form {\tt v = "..."} where {\tt v} is
-a local variable of any reference type that can store a string
-constant.  It is typically (but not necessarily) of type
-{\tt java.lang.String}.
-
-\item
-{\tt NilValAsgnInst}
-
-A statement of the form {\tt v = null} where {\tt v}
-is a local variable of reference type.
-
-\item
-{\tt StatFldRefInst}
-
-A statement of one of the following forms:
-
-{\tt v = f} \\
-{\tt f = v} \\
-{\tt * = f} \\
-{\tt f = *}
-
-where {\tt v} is a local variable of reference type and
-{\tt f} is a static field.
-
-The {\tt *} denotes an expression of primitive type in the
-original program which is elided due to \Chord's emphasis on the
-analysis of objects.
-
-\item
-{\tt InstFldRefInst}
-
-A statement of one of the following forms:
-
-{\tt v = b.f} \\
-{\tt b.f = v} \\
-{\tt * = b.f} \\
-{\tt b.f = *}
-
-where {\tt v} and {\tt b} are local variables of reference
-type and {\tt f} is an instance field.
-
-The {\tt *} denotes an expression of primitive type in the
-original program which is elided due to \Chord's emphasis on the
-analysis of objects.
-
-\item
-{\tt SkipInst}
-
-A no-op statement.
-
-Due to \Chord's emphasis on the analysis of objects, each statement
-in the original program that does not access
-any value of reference type (i.e., only accesses values of primitive
-type) is represented by this statement.
-
-\item
-{\tt InvkInst}
-
-A method invocation statement of the form
-{\tt r1,...,rn = m(a1,...,ak)} where
-{\tt r1}, ..., {\tt rn} (return variables) and
-{\tt a1}, ..., {\tt ak} (argument variables)
-are local variables of reference type.
-
-Argument variables and return variables of primitive type are
-elided due to \Chord's emphasis on the analysis of objects.
-
-A method invocation statement in Java has at most
-one return variable but
-\Chord's intermediate representation allows multiple return variables
-to enable compiling away exceptions, loops, synchronized blocks, etc.
-
-\item
-{\tt ObjVarAsgnInst}
-
-A statement of the form {\tt v1 = v2} where {\tt v1}
-and {\tt v2} are local variables of (possibly different) reference types.
-
-\item
-{\tt ObjValAsgnInst}
-
-A statement of the form {\tt v = new t} where {\tt t} is a
-concrete reference type and
-{\tt v} is a local variable of a (possibly different) reference type.
-
-\item
-{\tt ClsVarAsgnInst}
-
-A statement of the form {\tt v = b.getClass()} where
-{\tt v} is a local variable of type {\tt java.lang.Class}
-and {\tt b} is a local variable of reference type.
-
-The reason this kind of statement is modeled is because
-\Chord\ has an emphasis on concurrency and many Java programs
-hold locks on the runtime class of an object
-(which is obtained by calling the native {\tt getClass}
-instance method of the object).
-
-\item
-{\tt AryElemRefInst}
-
-A statement of one of the following forms:
-
-{\tt v = b[*]}  \\
-{\tt b[*] = v} \\
-{\tt * = b[*]} \\
-{\tt b[*] = *}
-
-where {\tt v} and {\tt b} are local variables of reference
-type.
-
-The {\tt *} denotes an expression of primitive type in the
-original program which is elided due to \Chord's emphasis on the
-analysis of objects.
-\end{itemize}
-
-\texonly{\newpage}
-\section{Program Transformations}
-\label{sec-prog-transform}
-
-After building the intermediate representation of the
-program, \Chord\ performs
-the following source-to-source transformations 
-that simplify several program analyses:
-
-\begin{enumerate}
-\item
-{\bf Sync Removal}
-
-Removes all synchronized blocks from each method M in each class
-C in the program by generating a fresh static method (in class C)
-per synchronized block in the body of M.
-
-\item
-{\bf Loop Removal}
-
-Removes all loops from each method M in each class C in the program
-by generating a fresh static method (in class C)
-per loop in the body of M.
-
-\item
-{\bf Skip Removal}
-
-Removes all skip (i.e., no-op) statements from each method in each class
-in the program.
-Such statements arise primarily because \Chord, due to its emphasis
-on the analysis of objects, replaces each
-statement in the original program that does not access any
-value of reference type (i.e., only accesses values of primitive
-type) by a skip statement.
-
-\end{enumerate}
-
-\texonly{\newpage}
-\section{Task Graph Construction}
-\label{sec-tg}
-
-\Chord\ requires the following three system properties to construct the task graph:
-
-\begin{enumerate}
-\item
-{\tt chord.dlogmap.path}
-
-The path specifying the mapping from the name of each Datalog-based program
-analysis that must be included in the project to the location
-of the file defining that program analysis.
-
-Each path element must be the location of a file containing a
-list of entries (one per line) of the form:
-
-{\tt <name> <file-name>}
-
-The path elements must be separated by the system-specific path separator
-character (e.g., `:' on UNIX and `;' on Windows).
-
-Let \DlogMapEntries\ denote the set of all entries of the
-above form.
-
-\item
-{\tt chord.trgtmap.path}
-
-The path partially specifying the mapping from the name of each program domain,
-program relation, and Java-based program analysis that must be
-included in the project to the fully-qualified name of a subclass, either
-pre-defined in \Chord\ or defined by the user, of {\tt chord.Dom},
-{\tt chord.Ary*Rel}, and
-{\tt chord.Entity}, respectively, the instance fields of which define the
-state of the corresponding domain, relation, or analysis.
-The `{\tt *}' in {\tt chord.Ary*Rel} is 1, 2, ..., 6, or N for
-program relations of arity 1, 2, ..., 6, or $>$ 6, respectively.
-
-Each path element must be the location of a file containing a
-list of entries (one per line) of the form:
-
-{\tt <name> <class-name>}
-
-The path elements must be separated by the system-specific path separator
-character (e.g., `:` on UNIX and ';' on Windows).
-
-Let \TaskMapEntries\ denote the set of all entries of the
-above form.
-
-\item
-{\tt chord.taskmap.path}
-
-The path partially specifying the mapping from the name of each program domain,
-program relation, and Java-based program analysis that must be included in
-the project to the fully-qualified name of a
-subclass, either pre-defined in \Chord\ or defined by the user, of
-{\tt chord.Dom}, {\tt chord.Ary*Rel}, and {\tt chord.Entity}, respectively,
-which define the following instance methods:
-\begin{itemize}
-\item
-the {\tt execute} method specifying
-the computation that updates the state of the corresponding domain, relation,
-or analysis,
-\item
-the {\tt invalidate} method specifying the action to be taken when the
-state of the corresponding domain, relation, or analysis is invalidated.
-\item
-the {\tt getConsumedNames} method specifying names of any domains, relations,
-and Datalog-based or Java-based program analyses upon whose state the
-computation depends, and
-\item
-the {\tt getProducedNames} method specifying names of any domains, relations,
-and Datalog-based or Java-based program analyses whose state depends upon
-the computation.
-\end{itemize}
-
-Each path element must be the location of a file containing a
-list of entries (one per line) of the form:
-
-{\tt <name> <class-name>}
-
-The path elements must be separated by the system-specific path separator
-character (e.g., `:` on UNIX and ';' on Windows).
-
-Let \TrgtMapEntries\ denote the set of all entries of the above form.
-\end{enumerate}
-
-Then, \Chord\ uses the values of \DlogMapEntries, \TaskMapEntries, and \TrgtMapEntries\ 
-to do the following:
-
-\begin{enumerate}
-\item
-Compute {\tt Tasks\ :\ Set<chord.Entity>}, the set of all tasks,
-each of which is an object of (a subclass of) {\tt chord.Entity},
-as follows:
-
-\begin{enumerate}
-\item
-For each entry in \DlogMapEntries\ (i.e., a Datalog-based program analysis,
-there is a unique task, namely, a unique object of class {\tt chord.DatalogAnalysis}.
-
-\item
-For each entry in \TaskMapEntries\ (i.e., of the form {\tt <name> <class-name>} specifying
-a program domain, a program relation, or a Java-based program analysis), there is a
-unique task, namely, a unique object
-of the class named {\tt <class-name>} which must be (a subclass of) {\tt chord.Dom} in the
-case of a program domain, (a subclass of) {\tt chord.Ary*Rel} in the case of a
-program relation, and (a subclass of) {\tt chord.Entity} in the case of a Java-based
-program analysis.
-\end{enumerate}
-
-\item
-Compute the following maps as shown below:
-
-{\tt ConsumedNameToTasksMap\ :\ String -> Set<Task>},
-mapping each name to the set of (possibly empty) tasks that consume that name.
-
-{\tt NameToTaskMap\ :\ String -> Task},
-mapping each name to the unique task that produces that name.
-
-\Chord\ reports an error if a name is produced by no task or by more than one task.
-{\small 
-\begin{verbatim}
-for each (<task> in Tasks) {
-    case <task> came from item 1(a) above: {
-        // <task> corresponds to a Datalog-based program analysis
-        for each (<name> declared as input relation) {
-            add <task> to ConsumedNameToTasksMap(<name>)
-        }
-        for each (<name> declared as output relation) {
-            if (<name> is in domain of map NameToTaskMap)
-                error ``multiple tasks produce name <name>''
-            else
-                NameToTaskMap(<name>) = <task>
-        }
-    }
-    case <task> came from item 1(b) above: {
-        // <task> corresponds to a program domain, a program relation, or
-        // a Java-based program analysis
-        // Suppose <task> came from entry <name> <class-name> in
-        // TaskMapEntries
-        // i.e. <task> is an object of class named <class-name>
-        for each (<name2> in <task>.getConsumedNames())
-            add <task> to ConsumedNameToTaskMaps(<name2>)
-        for each (<name2> in <task>.getProducedNames() union { <name> })
-            if (<name2> is in domain of map NameToTaskMap)
-                error ``multiple tasks produce name <name>''
-            else
-                NameToTaskMap(<name2>) = <task>
-    }
-}
-\end{verbatim}
-}
-
-{\small 
-\begin{verbatim}
-for each (<name> in domain of map ConsumedNameToTasksMap) {
-    if (<name> is not in domain of map NameToTaskMap)
-        error ``no task produces name <name>''
-}
-\end{verbatim}
-}
-
-\item
-Compute \NameToTrgtMap : {\tt String -> Entity},
-mapping each name to a unique target, as below.
-\Chord\ reports an error if a name is mapped to multiple targets
-or to no target.
-
-{\small 
-\begin{verbatim}
-for each (entry <name> <class-name> in TaskMapEntries) {
-    if (<name> is in domain of map NameToTrgtMap)
-        error ``name <name> mapped to multiple targets''
-    NameToTrgtMap(<name>) = NameToTaskMap(<name>)
-}
-for each (entry <name> <class-name> in TrgtMapEntries) {
-    if (<name> is in domain of map NameToTrgtMap)
-        error ``name <name> mapped to multiple targets''
-    NameToTrgtMap(<name>) = a fresh object of class named <class-name>
-}
-\end{verbatim}
-}
-
-To reduce user burden in providing \TaskMapEntries\ and
-\TrgtMapEntries, \Chord\ implicitly extends \NameToTrgtMap\ using
-\DlogMapEntries\ as follows:
-
-{\small 
-\begin{verbatim}
-for each (<name> in domain of map NameToTaskMap) {
-    if (<name> is not in domain of map NameToTrgtMap) {
-        if (<name> is declared as a domain in some datalog file in DlogMapEntries)
-            NameToTrgtMap(<name>) = a fresh object of class chord.Dom
-        else if (<name> is declared as a (input or output) relation
-                in some datalog file in DlogMapEntries)
-            NameToTrgtMap(<name>) = a fresh object of class chord.Rel
-        else
-            error ``<name> not mapped to any target''
-    } else {
-        chord.Entity <trgt> = NameToTrgtMap(<name>)
-        if (<name> is declared as a domain in some datalog file in DlogMapEntries
-                and <trgt> is not of type (subclass of) chord.Dom)
-            error ``<name> has conflicting types''
-        else if (<name> is declared as a (input or output) relation
-                in some datalog file in DlogMapEntries
-                and <trgt> is not of type (subclass of) chord.Rel)
-            error ``<name> has conflicting types''
-    }
-}
-\end{verbatim}
-}
-
-\item
-Compute graph \TaskGraph\ as a graph whose:
-
-\begin{itemize}
-\item
-set of nodes is \Tasks\ and
-\item
-set of edges containing an edge from task {\tt <task1>} to task {\tt <task2>} 
-whenever there exists a name {\tt <name>} such that
-\NameToTaskMap({\tt <name>}) = {\tt <task1>} and
-\ConsumedNameToTasksMap({\tt <name>}) contains {\tt <task2>}.
-\end{itemize}
-
-The task graph in XML and HTML formats is dumped to files {\tt taskgraph.xml}
-and {\tt taskgraph.html} in the project's working directory specified using
-system property {\tt chord.work.dir}.
-
-\end{enumerate}
-
-\texonly{\newpage}
-\section{Project Organization}
-
-\Chord\ interprets system property {\tt chord.trgts}, if defined, as
-a space-separated list of names of program domains, program relations, and
-Datalog-based or Java-based program analyses, denoted \TrgtNames,
-to be computed in that order.
-
-To avoid redundant computation, \Chord\ maintains during its execution a set of
-up-to-date tasks, denoted \ValidTasks,
-which is a subset of the set of all tasks, denoted \Tasks.
-Set \ValidTasks\ is empty at startup.
-A future release of \Chord\ may provide the ability to retrieve a set of up-to-date tasks from
-a previous run of \Chord\ to prevent re-computing those tasks in subsequent runs.
-The following procedure is executed upon startup:
-
-{\small
-\begin{verbatim}
-for each (<name> in TrgtNames in order) {
-    if (<name> is not in domain of map NameToTaskMap) {
-        error ``name <name> is an invalid target name''
-    else {
-        chord.Entity <task> = NameToTaskMap(<name>)
-        executeTask(<task>)
-    }
-}
-\end{verbatim}
-}
-
-\noindent where procedure {\tt executeTask} is defined as follows:
-
-{\small
-\begin{verbatim}
-procedure executeTask(chord.Entity <task>) {
-    if (<task> is not in ValidTasks) {
-        for each (immed. pred. <task2> of <task> in TaskGraph)
-            executeTask(<task2>)
-        <task>.execute()
-        for each (immed. succ. <task2> of <task> in TaskGraph)
-            invalidateTask(<task2>)
-        add <task> to ValidTasks
-    }
-}
-\end{verbatim}
-}
-
-\noindent where procedure {\tt invalidateTask} is defined as follows:
-
-{\small
-\begin{verbatim}
-procedure invalidateTask(chord.Entity t) {
-    if (<task> is in ValidTasks) {
-        remove <task> from ValidTasks
-        <task>.invalidate()
-        for each (immed. succ. <task2> of <task> in TaskGraph)
-            invalidateTask(<task2>)
-    }
-}
-\end{verbatim}
-}
-
-\texonly{\newpage}
-\section{Chord System Properties}
-\label{sec-sys-prop}
-
-\Chord\ recognizes the following system properties:
-
-\begin{itemize}
-\item
-{\tt chord.out}
-
-The location of a file to which the standard output stream
-of \Chord's execution is redirected.
-
-It may be the same file to which the standard error stream is redirected (see
-the description of system property {\tt chord.err}).
-
-The file is created if it does not exist and its contents are overwritten if it exists.
-
-\item
-{\tt chord.err}
-
-The location of a file to which the standard error stream
-of \Chord's execution is redirected.
-
-It may be the same file to which the standard output stream is redirected (see
-the description of system property {\tt chord.out}).
-
-The file is created if it does not exist and its contents are overwritten if it exists.
-
-\item
-{\tt chord.program.builder.class}
-
-The fully-qualified name of the class to be used for building the
-intermediate representation of the program to be analyzed.
-It must be a concrete class with a nullary constructor that
-implements interface {\tt chord.program.builders.IProgramBuilder}.
-
-The default value of this property is {\tt chord.program.builders.SootProgramBuilder}
-which uses the Soot Java compiler framework \cite{Soot}.
-
-\item
-{\tt chord.serial.file}
-
-The location of a file from (resp.\ to) which
-the serialized form of the intermediate representation of the
-program to be analyzed is loaded (resp.\ stored).
-
-\item
-{\tt chord.main.class}
-
-The fully-qualified name of the main class (e.g., {\tt foo.Main}) of the
-program to be analyzed.
-
-This system property (and {\tt chord.class.path})
-must be defined if system property {\tt chord.serial.file}
-does not specify the file containing the serialized form of the intermediate representation
-of the program to be analyzed.
-
-\item
-{\tt chord.class.path}
-
-The application classpath of the program to be analyzed (as opposed to the
-JDK library classpath which consists of jar files such as {\tt rt.jar}
-which are implicitly appended to the program's application classpath to
-yield the complete program classpath).
-
-Each path element must be the location of a directory or a jar/zip file.
-
-The path elements must be separated by the system-specific path separator character
-(e.g., `:' on UNIX and `;' on Windows).
-
-This system property (and {\tt chord.main.class}) must
-be defined if system property {\tt chord.serial.file}
-does not specify the file containing the serialized form of the intermediate representation
-of the program to be analyzed.
-
-\item
-{\tt chord.src.path}
-
-The Java source path of the program to be analyzed.
-
-Each path element must be the location of a directory.
-
-The purpose of this path is to provide the location of the
-{\tt .java} files of the program being analyzed so that, for instance, the
-results of an analysis (which operates on an intermediate representation of the program
-extracted from {\tt .class} files) can be mapped to the corresponding {\tt .java}
-files and presented to the user for manual inspection.
-
-Each file having suffix {\tt .java} in a directory (or a sub-directory of a directory)
-specified in this path is regarded as a source file of the program.
-
-\item
-{\tt chord.dlog.analysis.path}
-
-The path specifying the mapping from the name of each Datalog-based program
-analysis that must be included in the project to the location
-of the file defining that program analysis.
-
-Each path element must be the location of a file containing a
-list of entries (one per line) of the form:
-
-{\tt <name> <file-name>}
-
-The path elements must be separated by the system-specific path separator character
-(e.g., `:' on UNIX and `;' on Windows).
-
-\item
-{\tt chord.java.analysis.path}
-
-The path specifying the mapping from the name of each program domain, program relation,
-and Java-based program analysis that must be included in the project to the
-fully-qualified name of a subclass, either pre-defined in \Chord\ or defined by the user, of
-{\tt chord.Dom}, {\tt chord.Ary*Rel}, and {\tt chord.Entity}, respectively, which
-define the following instance methods:
-\begin{itemize}
-\item
-the {\tt execute} method specifying
-the computation that updates the state of the corresponding domain, relation,
-or analysis, 
-\item
-the {\tt invalidate} method specifying the action to be taken when the
-state of the corresponding domain, relation, or analysis is invalidated.
-\item
-the {\tt getConsumedNames} method specifying names of any domains, relations,
-and analyses upon whose state the computation depends, and
-\item
-the {\tt getProducedNames} method specifying names of any domains, relations,
-and analyses whose state depends upon the computation.
-\end{itemize}
-
-Each path element must be the location of a file containing a
-list of entries (one per line) of the form:
-
-{\tt <name> <class-name>}
-
-The path elements must be separated by the system-specific path separator character
-(e.g., `:' on UNIX and `;' on Windows).
-
-\item
-{\tt chord.trgtmap.path}
-
-The path specifying the mapping from the name of each program domain,
-program relation, and Java-based program analysis that must be
-included in the project to the fully-qualified name of a subclass,
-either pre-defined in \Chord\ or defined by the user, of {\tt chord.Dom},
-{\tt chord.Ary*Rel}, and
-{\tt chord.Entity}, respectively, the instance fields of which define the
-state of the corresponding domain, relation, or analysis.
-
-Each path element in the list must be the location of a file containing a
-list of entries (one per line) of the form:
-
-{\tt <name> <class-name>}
-
-The path elements must be separated by the system-specific path separator character
-(e.g., `:` on UNIX and ';' on Windows).
-
-\item
-{\tt chord.analyses}
-
-A space-separated list of the names of program domains, program relations, and
-Java-based or Datalog-based program analyses that must be computed in that order.
-
-\item
-{\tt chord.ext.class.path}
-
-The classpath of user-defined program domains, user-defined program relations,
-and user-defined Java-based program analyses.
-It is implicitly appended to the classpath of pre-defined program domains,
-pre-defined program relations, and pre-defined Java-based program analyses
-to yield the complete runtime classpath of \Chord.
-
-Each path element must be the location of a directory or a jar/zip file.
-
-The path elements must be separated by the system-specific path separator character
-(e.g., `:' on UNIX and `;' on Windows).
-\end{itemize}
-
-\texonly{\newpage}
-\section{Pre-defined Domains}
-\label{sec-predef-domains}
-
-The following domains are pre-defined in \Chord:
-
-\begin{itemize}
-\item
-$\Cdom$
-
-The domain of all k-object-sensitive contexts.
-
-The $0^{th}$ element in this domain denotes a distinguished hypothetical
-context (denoted \texorhtml{$\epsilon$}{{\it epsilon}} in the literature) that is regarded as the
-lone abstract context in which certain static methods are analyzed.
-The main static method with signature {\tt void main(java.lang.String[])}
-has no callers and is therefore analyzed in this context.
-For simplicity, the k-object-sensitive may alias analysis provided in
-\Chord\ also analyzes each class initializer method in this context.
-
-\item
-$\Edom$
-
-The domain of all statements that access (read or write) an instance field,
-a static field, or an array element, namely, all statements
-represented by instances of (subclasses of) {\tt chord.program.insts.HeapInst}.
-
-\item
-$\Fdom$
-
-The domain of  all instance and static fields.
-
-The $0^{th}$ element in this domain denotes a distinguished hypothetical
-field that is regarded as accessed whenever an array element is
-accessed.
-
-\item
-$\Hdom$
-
-The domain of all object allocation statements, namely, all statements
-represented by instances of {\tt chord.program.insts.ObjValAsgnInst}.
-
-\item
-$\Idom$
-
-The domain of all method invocation statements, namely, all statements
-represented by instances of {\tt chord.program.insts.InvkInst}.
-
-\item
-$\Mdom$
-
-The domain of all instance and static methods.
-
-The $0^{th}$ element in this domain is the main method of the program.
-
-The $1^{st}$ element in this domain is the method with signature
-{\tt void start()} of class {\tt java.lang.Thread}, if this
-method is deemed reachable from the main method of the program.
-
-The above two methods are the root methods of the implicitly spawned
-main thread and each explicitly spawned child thread, respectively.
-Due to \Chord's emphasis on concurrency, these methods are
-referenced frequently by various pre-defined program analyses
-expressed in Datalog, and giving them special indices makes it
-convenient to reference them in those analyses.
-
-\item
-$\Pdom$
-
-The domain of all simple statements, namely, all statements
-represented by instances of (subclasses of) {\tt chord.program.insts.Inst}.
-
-The $0^{th}$ element in this domain is the statement at the unique
-entry point of the main method of the program.
-
-The statements of each method in the program are assigned
-contiguous indices in this domain, with the statements at the
-unique entry and exit points of each method being assigned
-the smallest and largest indices, respectively, of all
-indices assigned to statements in that method.
-
-\item
-$\Tdom$
-
-The domain of all types, including the types of classes, interfaces,
-arrays, and primitives.
-
-\item
-$\Vdom$
-
-The domain of all local variables.
-
-Each local variable declared in each block of each method is
-represented by a unique element in this domain, in particular,
-local variables that have the same name but are declared in
-different methods or in different blocks of the same method
-are represented by different elements in this domain.
-
-All local variables of the same method are assigned contiguous
-indices in this domain, in particular, the set of local
-variables of a method is the disjoint union of its argument
-variables and its temporary variables, and the argument
-variables are assigned contiguous indices in order followed by
-the temporary variables.
-
-\item
-$\Zdom$
-
-The domain of all argument and return variable positions.
-Let $N$ be the largest number or arguments or return variables of any
-method.  Then, this domain contains the $N$ elements $0, 1, ..., N-1$.
-\end{itemize}
-
-\texonly{\newpage}
-\section{Pre-defined Relations}
-\label{sec-predef-relations}
-
-The following relations are pre-defined in \Chord:
-
-\begin{itemize}
-
-\item
-$\cha : \Mdom \times \Tdom \times \Mdom$
-
-Relation containing each tuple $(m_1,t,m_2)$ such that $m_2$ is
-the target method of a method invocation statement whose {\it resolved method}
-is $m_1$ and whose receiver has run-time type $t$.
-It is named $\cha$ because it represents the relation
-computed by Class Hierarchy Analysis.
-
-\item
-$\sub : \Tdom \times \Tdom$
-
-Relation containing each tuple $(s,t)$ such that type $s$ is a
-subtype of type $t$.
-
-\item
-$\HT : \Hdom \times \Tdom$
-
-Relation containing each tuple $(h,t)$ such that object allocation
-statement $h$ allocates objects of type $t$.
-
-\item
-$\VT : \Vdom \times \Tdom$
-
-Relation containing each tuple $(v,t)$ such that local variable $v$
-has type $t$.
-
-\item
-$\MH : \Mdom \times \Hdom$
-
-Relation containing each tuple $(m,h)$ such that method $m$ contains
-object allocation statement $h$.
-
-\item
-$\MI : \Mdom \times \Idom$
-
-Relation containing each tuple $(m,i)$ such that method $m$ contains
-method invocation statement $i$.
-
-\item
-$\ME : \Mdom \times \Edom$
-
-Relation containing each tuple $(m,e)$ such that method $m$ contains
-statement $e$ that accesses (reads or writes) an instance field, a
-static field, or an array element.
-
-\item
-$\instM : \Mdom$
-
-Relation containing all instance (as opposed static) methods.
-
-\item
-$\statM : \Mdom$
-
-Relation containing all static (as opposed instance) methods.
-
-\item
-$\clinitM : \Mdom$
-
-Relation containing all class initializer methods, namely,
-methods having signature {\tt void <clinit>()}.
-
-\item
-$\IinvkArg : \Idom \times \Zdom \times \Vdom$
-
-Relation containing each tuple $(i,z,v)$ such that local variable $v$
-is the $z^{th}$ argument variable of method invocation statement $i$.
-
-\item
-$\IinvkRet : \Idom \times \Zdom \times \Vdom$
-
-Relation containing each tuple $(i,z,v)$ such that local variable $v$
-is the $z^{th}$ return variable of method invocation statement $i$.
-
-\item
-$\MmethArg : \Mdom \times \Zdom \times \Vdom$
-
-Relation containing each tuple $(m,z,v)$ such that local variable
-$v$ is the $z^{th}$ argument variable of method $m$.
-
-\item
-$\MmethRet : \Mdom \times \Zdom \times \Vdom$
-
-Relation containing each tuple $(m,z,v)$ such that local variable
-$v$ is the $z^{th}$ return variable of method $m$.
-
-\item
-$\MspcMethArg : \Mdom \times \Vdom$
-
-Relation containing each tuple $(m,v)$ such that local variable $v$ is
-the implicit {\tt this} argument variable of instance method $m$.
-
-\item
-
-$\MordMethArg : \Mdom \times \Zdom \times \Vdom$
-
-Relation containing each tuple $(m,z,v)$ such that local variable $v$
-is the $z^{th}$ argument variable of method $m$ other than the implicit
-{\tt this} argument variable (in the case in which $m$ is an instance
-method).
-
-\item
-$\MobjVarAsgnInst : \Mdom \times \Vdom \times \Vdom$
-
-Relation containing each tuple $(m,l,r)$ such that method $m$
-contains a copy statement of the form $l = r$ where $l$ and $r$ are
-local variables.
-
-\item
-$\MobjValAsgnInst : \Mdom \times \Vdom \times \Vdom$
-
-Relation containing each tuple $(m,l,h)$ such that method $m$ contains
-object allocation statement $h$ which assigns to local variable $l$.
-
-\item
-$\MclsVarAsgnInst : \Mdom \times \Vdom \times \Vdom$
-
-Relation containing each tuple $(m,l,b)$ such that method $m$
-contains a statement of the form $l = b.{\tt getClass()}$.
-
-\item
-$\MgetInstFldInst : \Mdom \times \Vdom \times \Fdom \times \Vdom$
-
-Relation containing each tuple $(m,b,f,r)$ such that method $m$
-contains a statement of the form $b.f = r$ which assigns to
-instance field or array element $f$ of the object denoted by local
-variable $b$ the value denoted by local variable $r$.
-
-\item
-$\MputInstFldInst : \Mdom \times \Vdom \times \Fdom \times \Vdom$
-
-Relation containing each tuple $(m,b,f,r)$ such that method $m$
-contains a statement of the form $b.f = r$ which writes the value
-denoted by local variable $r$ to instance field or array element
-$f$ of the object denoted by local variable $b$.
-
-\item
-$\MgetStatFldInst : \Mdom \times \Vdom \times \Fdom$
-
-Relation containing each tuple $(m,l,f)$ such that method $m$ contains
-a statement of the form $l = f$ that assigns to local variable $l$ the
-value denoted by static field $f$.
-
-\item
-$\MputStatFldInst : \Mdom \times \Fdom \times \Vdom$
-
-Relation containing each tuple $(m,f,r)$ such that method $m$ contains
-a statement of the form $f = r$ which assigns to static field $f$ the
-value denoted by local variable $r$.
-
-\item
-$\virtIM : \Idom \times \Mdom$
-
-Relation containing each tuple $(i,m)$ such that $m$ is the
-{\it resolved method} of {\tt invokevirtual} or {\tt invokeinterface} method
-invocation statement $i$.
-
-\item
-$\statIM : \Idom \times \Mdom$
-
-Relation containing each tuple $(i,m)$ such that $m$ is the
-{\it resolved method} of {\tt invokestatic} method invocation statement $i$.
-
-\item
-$\specIM : \Idom \times \Mdom$
-
-Relation containing each tuple $(i,m)$ such that $m$ is the
-{\it resolved method} of {\tt invokespecial} method invocation statement $i$.
-
-\item
-$\syncMV : \Mdom \times \Vdom$
-
-Relation containing each tuple $(m,v)$ such that method $m$ is
-synchronized on argument variable $v$.
-
-\item
-$\writeE : \Edom$
-
-Relation containing all statements that write (as opposed to read)
-an instance field, static field, or array element.
-
-\item
-$\EF : \Edom \times \Fdom$
-
-Relation containing each tuple $(e,f)$ such that statement $e$
-accesses (reads or writes) instance field, static field, or
-array element $f$.
-
-\item
-$\VH : \Vdom \times \Hdom$
-
-Relation containing each tuple $(v,h)$ such that local variable $v$
-may point to abstract object $h$, as determined by 0CFA-based may
-alias analysis.
-
-\item
-$\FH : \Fdom \times \Hdom$
-
-Relation containing each tuple $(f,h)$ such that static field $f$ may
-point to abstract object $h$, as determined by 0CFA-based may alias
-analysis.
-
-\item
-
-$\HFH : \Hdom \times \Fdom \times \Hdom$
-
-Relation containing each tuple $(h_1,f,h_2)$ such that instance field
-or array element $f$ of abstract object $h_1$ may point to abstract
-object $h_2$, as determined by 0CFA-based may alias analysis.
-
-\item
-$\IM : \Idom \times \Mdom$
-
-Relation containing each tuple $(i,m)$ such that method invocation
-statement $i$ may call method $m$ as determined by 0CFA-based may alias
-analysis.
-
-\item
-$\nestedLocks : \Cdom \times \Cdom \times \Mdom \times \Cdom \times \Mdom$
-
-Relation containing each tuple $(c,c_1,m_1,c_2,m_2)$ such that in the
-context-sensitive call graph computed by k-object-sensitive may
-alias analysis, abstract thread $c$ may first acquire the lock at
-the entry of method $m_1$ in abstract context $c_1$ and without
-releasing it proceed to acquire the lock at the entry of method
-$m_2$ in abstract context $c_2$.
-
-\item
-$\forkC : \Cdom$
-
-Relation containing all abstract contexts in which the
-{\tt start()} method of class {\tt java.lang.Thread} may
-be reachable from the main method, as determined by
-k-object-sensitive may alias analysis.
-
-\item
-$\CVC : \Cdom \times \Vdom \times \Cdom$
-
-Relation containing each tuple $(c_1,v,c_2)$ such that local variable
-$v$ may point to abstract object $c_2$ in abstract context $c_1$ of its
-containing method, as determined by k-object-sensitive may alias
-analysis.
-
-\item
-$\FC : \Fdom \times \Cdom$
-
-Relation containing each tuple $(f,c)$ such that static field $f$
-may point to abstract object $c$, as determined by k-object-sensitive
-may alias analysis.
-
-\item
-$\CFC : \Cdom \times \Fdom \times \Cdom$
-
-Relation containing each tuple $(c_1,f,c_2)$ such that instance field
-or array element $f$ of abstract object $c_1$ may point to abstract
-object $c_2$, as determined by k-object-sensitive may alias analysis.
-
-\item
-$\CICM : \Cdom \times \Idom \times \Cdom \times \Mdom$
-
-Relation containing each tuple $(c_1,i,c_2,m)$ such that method
-invocation statement $i$ in abstract context $c_1$ of its containing method
-may call method $m$ in abstract context $c_2$, as determined by
-k-object-sensitive may alias analysis.
-
-\item
-$\CHC : \Cdom \times \Hdom \times \Cdom$
-
-Relation containing each tuple $(c_1,h,c_2)$ such that the method
-containing object allocation statement $h$ may be reachable in
-abstract context $c_1$, and $c_2 = (h::c_1) + k$, where $k$ is the k
-value of statement $h$, provided as an argument to k-object-sensitive
-may alias analysis.
-
-\end{itemize}
-
-\texonly{\newpage}
-\section{Pre-defined Datalog-based Analyses}
-\label{sec-predef-datalog-analyses}
-
-\texonly{\newpage}
-\section{Pre-defined Java-based Analyses}
-\label{sec-predef-java-analyses}
-
-\texonly{\newpage}
-\bibliographystyle{plain}
-\bibliography{citations}
-
-\end{document}
-
-\newpage
-\section{Defining Tasks and Targets}
-
-\Chord\ uses the notion of {\it tasks} to express computations
-and the notion of {\it targets} to express the structure of
-inputs/outputs of computations.
-
-Tasks and targets are both implemented uniformly as {\it entities}.
-A generic entity is represented by class {\tt chord.Entity}
-depicted in Figure \ref{fig-entity}.
-
-\begin{figure}
-\begin{verbatim}
-public class Entity {
-    protected static final String[] emptyStringArray = new String[0];
-    // subclasses may override
-    public void execute() {
-        // do nothing
-    }
-    // subclasses may override
-    public String[] getConsumedTargetNames() {
-        return emptyStringArray;
-    }
-    // subclasses may override
-    public String[] getProducedTargetNames() {
-        return emptyStringArray;
-    }
-    ...
-}
-\end{verbatim}
-\caption{Snippet of class {\tt chord.Entity}}
-\label{fig-entity}
-\end{figure}
-
-Whenever an enntity is regarded as a task:
-
-\begin{itemize}
-\item
-The {\tt execute} method defines the computation that must be performed by the task.
-
-\item
-The {\tt getConsumedTargetNames} method specifies the names of targets
-that are inputs to (the computation performed by) the task.
-
-\item
-The {\tt getProducedTargetNames} method specifies the names of targets
-that are outputs of (the computation performed by) the task.
-\end{itemize}
-
-There are four kinds of entities:
-\begin{itemize}
-\item
-A program domain
-
-\item
-A program relation
-
-\item
-A program analysis expressed in Datalog
-
-\item
-A program analysis expressed in Java
-
-\end{itemize}
-
-We next describe each of these kinds of entities in detail.
-
-\subsection{Program Domains}
-
-A program domain is a set of related elements in a given program (e.g.,
-the set of all methods, the set of all object allocation statements, etc.).
-The structure of a generic program domain is represented by
-class {\tt chord.Dom} depicted in Figure \ref{fig-dom}.
-
-\begin{figure}
-\begin{verbatim}
-public class Dom<T> extends Entity {
-    public final IndexMap<T> map;
-    public Dom() {
-        this.map = new IndexMap<T>();
-    }
-    public void execute() {
-        populate();
-        saveToDomFile();
-    }
-    // subclasses cannot override
-    public final String[] getProducedTargetNames() {
-        return emptyStringArray;
-    }
-    // subclasses may override
-    protected void populate() {
-        // do nothing
-    }
-    ...
-\end{verbatim}
-\caption{Snippet of class {\tt chord.Dom}}
-\label{fig-dom}
-\end{figure}
-
-A program domain provides a map from each element in the domain
-to a unique integer in the range $[0..n-1]$ where $n$ is the size of
-the domain.
-This map is required by {\tt bddbddb} (a BDD-based Datalog solver).
-The {\tt execute} method is specialized to construct the map and then
-save the map to disk for use by {\tt bddbddb}.
-Subclasses must override the {\tt populate} method to construct the
-map as desired.
-
-It is not possible to override the {\tt getProducedTargetNames}
-method (note that it is declared {\tt final}) because a program domain
-task is a specialized task whose output is exactly one target, namely,
-the program domain it constructs.
-It is, however, possible to override the {\tt getConsumedTargetNames}
-method (inherited from superclass {\tt chord.Entity}) since constructing
-a program domain may require other targets (i.e., other program
-domains, program relations, or program analyses).
-
-Figure \ref{fig-domI} defines the program domain
-named $\Idom$ as the domain of all method invocation statements.
-Note from method {\tt getConsumedTargetNames} that it takes as input
-the domain named $\Mdom$ which is the domain of all methods.
-This is because the {\tt populate} method refers to that domain.
-Intuitively, the {\tt populate} method gives adjacent indices in
-domain $\Idom$ to method invocation statements contained in
-methods that have adjacent indices in domain $\Mdom$ are given
-adjacent indices in domain $\Idom$.
-
-\begin{figure}
-\begin{verbatim}
-public class DomI extends Dom<InvkInst> {
-    private IndexMap<Method> mapM;
-    private final String[] consumedTaskNames = new String[] { "M" };
-    public String[] getConsumedTargetNames() {
-        return consumedTaskNames;
-    }
-    protected void populate() {
-        DomM domM = (DomM) project.getEntity("M");
-        mapM = domM.map;
-        for (int m = 0; m < mapM.size(); m++) {
-            Method meth = mapM.get(m);
-            Body body = meth.body;
-            if (body == null)
-                continue;
-            Set<Inst> insts = body.cfg.getNodes();
-            for (Inst inst : insts) {
-                if (inst instanceof InvkInst)
-                    map.set((InvkInst) inst);
-            }
-        }
-    }
-    ...
-}
-\end{verbatim}
-\caption{Snippet of class {\tt chord.DomI}}
-\label{fig-domI}
-\end{figure}
-
-A list of pre-defined program domains is presented in Section \ref{sec-predef-domains}.
-
-\subsection{Program Relations}
-
-A program relation is a relation over one or more program domains.
-The structure of a generic program relation is represented by class
-{\tt chord.Rel} depicted in Figure \ref{fig-rel}.
-
-\begin{figure}
-\begin{verbatim}
-\end{verbatim}
-\caption{Snippet of class {\tt chord.Rel}}
-\label{fig-rel}
-\end{figure}
-
-It is not possible to override the {\tt getProducedTargetNames}
-method (note that it is declared {\tt final}) because a program relation
-task is a specialized task whose output is exactly one target, namely,
-the program relation it constructs.
-It is, however, possible to override the {\tt getConsumedTargetNames}
-method (inherited from superclass {\tt chord.Entity}) since constructing
-a program relation may require other targets (i.e., program
-domains, other program relations, or program analyses).
-
-A list of pre-defined program relations is presented in Section \ref{sec-predef-relations}.
-
-\subsection{Program Analyses in Datalog}
-
-A list of pre-defined program analyses in Datalog is presented in
-Section \ref{sec-predef-datalog-analyses}.
-
-\subsection{Program Analyses in Java}
-
-A list of pre-defined program analyses in Java is presented in
-Section \ref{sec-predef-java-analyses}.
+In the above case, the above command does not do much beyond loading the sample \verb+chord.properties+ file.
+For Chord to do something interesting, you need to set additional properties, either
+on the above command line or in the \verb+chord.properties+ file, specifying the task(s)
+Chord must perform.
+All tasks are summarized in the ``Chord task properties" portion of the table in Section \ref{sec:chord-sysprops}.
+The two most common tasks, described next, are building program scope (Section \ref{sec:building-scope})
+and running program analyses (Section \ref{sec:running-analysis}).

### 2010-04-01 main/doc/running_analysis.tex

--- 2009-07-01 doc/user_guide.tex
@@
-\documentclass{article}
-\usepackage{hyperlatex}
+\section{Running a Program Analysis}
+\label{sec:running-analysis}
 
-%\topmargin=-0.5in
-\texonly{\oddsidemargin=0in}
-\texonly{\evensidemargin=0in}
-\texonly{\textwidth=5.6in}
-%\textheight=8.5in
+TODO
 
-%TODO: specify use of BDD's under efficiency in section 1 (overview)
-
-\htmlmathitalic{0}
-\htmltitle{User Guide}
-\htmladdress{chord@cs.stanford.edu, \today}
-
-\providecommand\Mdom{{\bf M}}
-\providecommand\Idom{{\bf I}}
-\providecommand\Vdom{{\bf V}}
-\providecommand\Zdom{{\bf Z}}
-\providecommand\Hdom{{\bf H}}
-\providecommand\Edom{{\bf E}}
-\providecommand\Fdom{{\bf F}}
-\providecommand\Cdom{{\bf C}}
-\providecommand\Tdom{{\bf T}}
-\providecommand\Pdom{{\bf P}}
-\providecommand\Sdom{{\bf S}}
-
-\providecommand\instM{{\bf instM}}
-\providecommand\statM{{\bf statM}}
-\providecommand\MH{{\bf MH}}
-\providecommand\MI{{\bf MI}}
-\providecommand\ME{{\bf ME}}
-
-\providecommand\cha{{\bf cha}}
-\providecommand\sub{{\bf sub}}
-\providecommand\HT{{\bf HT}}
-\providecommand\VT{{\bf VT}}
-\providecommand\EF{{\bf EF}}
-\providecommand\writeE{{\bf writeE}}
-
-\providecommand\virtIM{{\bf virtIM}}
-\providecommand\statIM{{\bf statIM}}
-\providecommand\specIM{{\bf specIM}}
-
-\providecommand\MmethArg{{\bf MmethArg}}
-\providecommand\MmethRet{{\bf MmethRet}}
-\providecommand\IinvkArg{{\bf IinvkArg}}
-\providecommand\IinvkRet{{\bf IinvkRet}}
-\providecommand\MordMethArg{{\bf MordMethArg}}
-\providecommand\MspcMethArg{{\bf MspcMethArg}}
-
-\providecommand\MobjVarAsgnInst{{\bf MobjVarAsgnInst}}
-\providecommand\MclsVarAsgnInst{{\bf MclsVarAsgnInst}}
-\providecommand\MobjValAsgnInst{{\bf MobjValAsgnInst}}
-\providecommand\MgetInstFldInst{{\bf MgetInstFldInst}}
-\providecommand\MputInstFldInst{{\bf MputInstFldInst}}
-\providecommand\MgetStatFldInst{{\bf MgetStatFldInst}}
-\providecommand\MputStatFldInst{{\bf MputStatFldInst}}
-
-\providecommand\syncMV{{\bf syncMV}}
-\providecommand\nestedLocks{{\bf nestedLocks}}
-
-\providecommand\VH{{\bf VH}}
-\providecommand\FH{{\bf FH}}
-\providecommand\HFH{{\bf HFH}}
-\providecommand\IM{{\bf IM}}
-
-\providecommand\forkC{{\bf forkC}}
-
-\providecommand\CVC{{\bf CVC}}
-\providecommand\FC{{\bf FC}}
-\providecommand\CFC{{\bf CFC}}
-\providecommand\CICM{{\bf CICM}}
-\providecommand\CHC{{\bf CHC}}
-
-\providecommand\clinitM{{\bf clinitM}}
-
-\providecommand\Chord{{Chord}}
-
-\providecommand\NameToTrgtMap{{\tt NameToTrgtMap}}
-\providecommand\NameToTaskMap{{\tt NameToTaskMap}}
-\providecommand\ConsumedNameToTasksMap{{\tt ConsumedNameToTasksMap}}
-\providecommand\TaskGraph{{\tt TaskGraph}}
-\providecommand\TaskMapEntries{{\tt TaskMapEntries}}
-\providecommand\TrgtMapEntries{{\tt TrgtMapEntries}}
-\providecommand\DlogMapEntries{{\tt DlogMapEntries}}
-\providecommand\Tasks{{\tt Tasks}}
-\providecommand\Trgts{{\tt Trgts}}
-\providecommand\ValidTasks{{\tt ValidTasks}}
-\providecommand\TrgtNames{{\tt TrgtNames}}
-
-\title{Chord: A Static Analysis Framework for Java}
-\author{Mayur Naik and Alex Aiken}
-\date{\today}
-\begin{document}
-\maketitle
-\texonly{\tableofcontents}
-
-\texonly{\newpage}
-\section{Overview}
-
-\Chord\ is a static analysis framework for Java with an emphasis on the analysis of objects and concurrency.
-It has the following characteristics:
-\begin{enumerate}
-\item
-It is {\it stand-alone} in that it provides various off-the-shelf static analyses (e.g., may alias analysis, thread escape analysis, datarace analysis, deadlock analysis, etc.).
-\item
-It is {\it extensible} in that it allows users to build their own static analyses on top of the provided static analyses.
-\item
-It is {\it compositional} in that it enables different static analyses to coexist and interact with each other.
-\item
-It is {\it efficient} in that it
-caches results computed by each static analysis so that they can be reused by other static analyses without re-computation.
-\item
-It is {\it flexible} in that it allows each static analysis to be expressed either declaratively as a Datalog program or imperatively as a Java program.
-\end{enumerate}
-
-\texonly{\newpage}
-\section{Download and Installation}
-
-Before installing \Chord\ ensure that the following software is installed on your machine:
-
-\begin{itemize}
-\item
-JDK 5 or higher (e.g., from
-Sun\footnote{\xlink{{\tt http://java.sun.com/javase/downloads/index.jsp}}{http://java.sun.com/javase/downloads/index.jsp}}
-or IBM\footnote{\xlink{{\tt http://www.ibm.com/developerworks/java/jdk/}}{http://www.ibm.com/developerworks/java/jdk/}}).
-
-\item
-Apache Ant (a Java build tool)\footnote{\xlink{{\tt http://ant.apache.org/}}{http://ant.apache.org/}}
-\end{itemize}
-
-Download the latest version of \Chord\ from \xlink{{\tt http://code.google.com/p/jchord/}}{http://code.google.com/p/jchord/}
-(say {\tt chord-x.y.tar.gz}) and uncompress it in a suitable directory by running the following command:
-
-\begin{verbatim}
-tar xvzf chord-x.y.tar.gz
-\end{verbatim}
-
-This will create a directory structure with root directory {\tt chord-x.y/}.
-
-\texonly{\newpage}
-\section{Building Chord}
-
-The root directory {\tt chord-x.y/} contains a {\tt build.xml} file which is interpreted by
-Apache Ant (a Java build tool).
-To see the various targets available run the following command in the root directory:
-
-\begin{verbatim}
-ant -p
-\end{verbatim}
-
-The default target {\tt compile} compiles \Chord.  So simply run the following command in the
-root directory:
-
-\begin{verbatim}
-ant
-\end{verbatim}
-
-This will compile the \Chord\ {\tt .java} source files and store the {\tt .class} files
-under the {\tt classes/} sub-directory of the root directory.
-
-\texonly{\newpage}
-\section{Running Chord}
-
-Suppose the program to be analyzed has the following directory structure:
-
-\begin{verbatim}
-example/
-    src/
-        foo/
-            Main.java
-            ...
-    classes/
-        foo/
-            Main.class
-            ...
-    lib/
-        src/
-            taz/
-                ...
-        jars/
-            taz.jar
-    chord/
-        chord.properties
-\end{verbatim}
-
-The above structure is typical: the program's Java source files are under {\tt example/src/}, its class files
-are under {\tt example/classes/}, the source and jar files of the libraries used by the program are under
-{\tt example/lib/src/} and {\tt example/lib/jars/}, respectively, and finally {\tt example/chord/} is the directory
-to which Chord will output its intermediate and final results.
-
-Execute Chord on the above program by running the
-following command in Chord's root directory {\tt chord-x.y/}:
-
-\begin{verbatim}
-ant -Dchord.work.dir=.../example/chord run
-\end{verbatim}
-
-where ``{\tt ...}'' denotes either the absolute path, or the path relative to Chord's root directory,
-of the parent of directory {\tt example/}.
-Chord will read the system properties specified in the file named {\tt chord.properties}
-that it presumes exists in the directory specified
-by system property {\tt chord.work.dir}
-which in the above case is directory {\tt example/chord/}.
-A sample such file for the above program is as follows:
-
-\begin{verbatim}
-chord.main.class=foo.Main
-chord.class.path=../classes:../lib/classes/taz.jar
-chord.src.path=../src:../lib/src
-chord.out=log.txt
-chord.err=log.txt
-chord.serial.file=program.ser
-chord.java.analysis.path=${chord.home.dir}/classes/main:
-    ${chord.home.dir}/classes/downcast:${chord.home.dir}/src/dlog/monosite
-chord.dlog.analysis.path=${chord.home.dir}/src/dlog/main:
-    ${chord.home.dir}/src/dlog/downcast:${chord.home.dir}/src/dlog/monosite
-chord.analyses=downcast-java,monosite-java
-\end{verbatim}
-
-System property {\tt chord.home.dir} specifies Chord's root directory
-{\tt chord-x.y/}.
-Every non-absolute location used in the value of any system property defined in the file
-is taken to be relative to the directory specified by system
-property {\tt chord.work.dir}.
-
-Section \ref{sec-sys-prop} presents an exhaustive
-listing of system properties that are recognized by \Chord.
-Briefly:
-\begin{itemize}
-\item
-{\tt chord.main.class} specifies the main class of the program
-to be analyzed.
-\item
-{\tt chord.class.path} specifies the application classpath of the program to be analyzed (as opposed to the
-JDK library classpath which consists of jar files such as {\tt rt.jar}
-which are implicitly appended to the program's application classpath to yield the
-complete program classpath).
-\item
-{\tt chord.out} and {\tt chord.err} specify the files (typically one and the same) to which the standard output and
-standard error streams, respectively, must be redirected while running Chord.
-\item
-{\tt chord.serial.file} specifies the location of a file from (resp. to) which
-the serialized form of the intermediate representation of the program to be analyzed must be loaded (resp. stored).
-This option is provided because loading the serialized form is much faster than
-constructing the intermediate representation of the program each time.
-\item
-{\tt chord.java.analysis.path} and {\tt chord.dlog.analysis.path} 
-specify all program domains, program relations, and program analyses
-to be included in the project and the
-dependencies between them (see Section \ref{sec-tg}).
-In the above case, they specify certain
- program domains, program relations, and program analyses
-pre-defined in \Chord, in particular, all those in (1) Chord's standard library, (2) Chord's downcast safety
-analysis, and (3) Chord's monomorphic call site analysis.
-Users can also define their own
-program domains, program relations, and program analyses and include them in the project.
-\item
-{\tt chord.analyses} is a space- or comma-separated list of the names of program domains,
-program relations, and program analyses that must be run in that order.  In the above case,
-these are downcast safety analysis, denoted {\tt downcast-java}, and monomorphic call site analysis,
-denoted {\tt monosite-java}.
-Both are program analyses pre-defined in \Chord\ 
-but users can also include the names
-of program domains, program relations, and program analyses they have defined and wish to run.
-\end{itemize}
-
-\texonly{\newpage}
-\section{Specifying Program Scope}
-
-\Chord\ takes as input a complete (i.e., closed) program with a main method.
-The scope of the program must be specified in one of the
-following ways:
-
-\begin{itemize}
-\item
-Defining system property {\tt chord.serial.file} as
-the location of a file containing in serialized form the intermediate representation
-of the program generated by a previous run of \Chord.
-\item
-Defining system property {\tt chord.main.class} as the fully-qualified name
-of the program's main class (e.g., {\tt foo.Main}) and system property
-{\tt chord.class.path} as the program's application classpath.
-\end{itemize}
-
-The former option is provided because loading the serialized form
-is much faster than constructing the intermediate representation
-of the program each time.
-
-\Chord\ first checks whether system property {\tt chord.serial.file} is defined.
-If it is defined, then it further checks whether the file specified by it
-exists.
-If the file exists, then \Chord\ treats the file as a serialized form of
-the intermediate representation of the program to be analyzed, and loads it.
-
-If system property {\tt chord.serial.file} is not defined or if the file
-specified by the property does not exist, then \Chord\ requires system properties
-{\tt chord.main.class} and {\tt chord.class.path} 
-to be defined, and uses them to
-build an intermediate representation of the program (see Section \ref{sec-prog-build})
-and perform source-to-source transformations on that representation
-(see Section \ref{sec-prog-transform}) that simplify several program analyses.
-Finally, if system property {\tt chord.serial.file} was defined but
-specified a non-existent file, then \Chord\ creates the file and stores the
-transformed intermediate representation of the program into the file in
-serialized form so that it can be loaded in subsequent runs.
-
-Defining system property {\tt chord.src.path} is optional.
-It specifies the program's Java source path which provides the
-locations of the program's {\tt .java} files so that, for instance, the
-results of an analysis (which operates on the intermediate representation of
-the program extracted from {\tt .class} files) can be mapped to the
-program's {\tt .java} files and presented to the user for
-manual inspection.
-
-\texonly{\newpage}
-\section{Program Representation}
-\label{sec-prog-build}
-
-Given the program's program's main class via system property {\tt chord.main.class} and the
-program's application classpath via system property {\tt chord.class.path},
-\Chord\ uses system property {\tt chord.program.builder.class} to locate the
-builder to be used for building an intermediate representation of the program,
-namely, an object of class {\tt chord.program.Program}.
-By default, the value of this property is class {\tt chord.program.builders.SootProgramBuilder}
-which uses the Soot Java compiler framework \cite{Soot}.
-Any builder, however, must implement interface
-{\tt chord.program.builders.IProgramBuilder}.
-
-The intermediate representation
-elides values of primitive type (e.g., int, boolean, etc.) and operations
-on them (e.g., arithmetic and relational operations) in the original program.
-The current emphasis of \Chord\ is on the analysis of objects
-but the analysis of primitives may be supported in future
-releases.
-
-The intermediate representation represents the program as a set of types.
-Each type has a set of (instance or static) fields and a set of
-(instance or static) methods.
-Each method has an optional control-flow graph.
-The control-flow graph has a unique entry point and
-a unique exit point\footnote{Even the control-flow graph of a method with an infinite loop
-has an exit point because loops are transformed into tail-recursive methods
-(see Section \ref{sec-prog-transform}).}.
-The entry and exit points are always distinct.
-Each point in the control-flow graph has one or more immediate predecessors
-(except the entry point which has none) and, likewise,
-one or more immediate successors (except the exit point which has none).
-Associated with each point is a simple statement, which may be of one of
-the following kinds:
-
-\begin{itemize}
-\item
-{\tt HeadInst}
-
-The statement associated with the unique entry point of a control-flow graph.
-
-It serves as a placeholder for:
-
-\begin{itemize}
-\item
-the (possibly empty) list of the method's {\it argument variables} of reference type;
-in the case of an instance method, this list is non-empty,
-with the $0^{th}$ element being the distinguished implicit {\tt this}
-argument variable.
-\item
-the (possibly empty) list of the method's {\it temporary variables} of reference type.
-\end{itemize}
-
-Argument variables and temporary variables of primitive type are
-elided due to \Chord's emphasis on the analysis of objects.
-
-Argument variables and temporary variables are distinct, namely,
-the set of a method's {\it local variables} (of reference type)
-is the disjoint union of the sets of the method's argument variables (of reference type) and
-temporary variables (of reference type).
-
-\item
-{\tt TailInst}
-
-The statement associated with the unique exit point of a control-flow graph.
-
-It serves as a placeholder for 
-the (possibly empty) list of the method's {\it return variables} of reference type.
-
-Return variables of primitive type are
-elided due to \Chord's emphasis on the analysis of objects.
-
-The set of a method's return variables (of reference type) is a subset of the set
-of the method's local variables (of reference type).
-
-A Java method has at most one return variable but
-\Chord's intermediate representation allows multiple return variables
-to enable compiling away exceptions, loops, synchronized blocks, etc.
-
-\item
-{\tt AcqLockInst}
-
-A statement of the form {\tt monitorenter v} where
-{\tt v} is a local variable of reference type.
-
-\item
-{\tt RelLockInst}
-
-A statement of the form {\tt monitorexit v} where {\tt v}
-is a local variable of reference type.
-
-\item
-{\tt StrValAsgnInst}
-
-A statement of the form {\tt v = "..."} where {\tt v} is
-a local variable of any reference type that can store a string
-constant.  It is typically (but not necessarily) of type
-{\tt java.lang.String}.
-
-\item
-{\tt NilValAsgnInst}
-
-A statement of the form {\tt v = null} where {\tt v}
-is a local variable of reference type.
-
-\item
-{\tt StatFldRefInst}
-
-A statement of one of the following forms:
-
-{\tt v = f} \\
-{\tt f = v} \\
-{\tt * = f} \\
-{\tt f = *}
-
-where {\tt v} is a local variable of reference type and
-{\tt f} is a static field.
-
-The {\tt *} denotes an expression of primitive type in the
-original program which is elided due to \Chord's emphasis on the
-analysis of objects.
-
-\item
-{\tt InstFldRefInst}
-
-A statement of one of the following forms:
-
-{\tt v = b.f} \\
-{\tt b.f = v} \\
-{\tt * = b.f} \\
-{\tt b.f = *}
-
-where {\tt v} and {\tt b} are local variables of reference
-type and {\tt f} is an instance field.
-
-The {\tt *} denotes an expression of primitive type in the
-original program which is elided due to \Chord's emphasis on the
-analysis of objects.
-
-\item
-{\tt SkipInst}
-
-A no-op statement.
-
-Due to \Chord's emphasis on the analysis of objects, each statement
-in the original program that does not access
-any value of reference type (i.e., only accesses values of primitive
-type) is represented by this statement.
-
-\item
-{\tt InvkInst}
-
-A method invocation statement of the form
-{\tt r1,...,rn = m(a1,...,ak)} where
-{\tt r1}, ..., {\tt rn} (return variables) and
-{\tt a1}, ..., {\tt ak} (argument variables)
-are local variables of reference type.
-
-Argument variables and return variables of primitive type are
-elided due to \Chord's emphasis on the analysis of objects.
-
-A method invocation statement in Java has at most
-one return variable but
-\Chord's intermediate representation allows multiple return variables
-to enable compiling away exceptions, loops, synchronized blocks, etc.
-
-\item
-{\tt ObjVarAsgnInst}
-
-A statement of the form {\tt v1 = v2} where {\tt v1}
-and {\tt v2} are local variables of (possibly different) reference types.
-
-\item
-{\tt ObjValAsgnInst}
-
-A statement of the form {\tt v = new t} where {\tt t} is a
-concrete reference type and
-{\tt v} is a local variable of a (possibly different) reference type.
-
-\item
-{\tt ClsVarAsgnInst}
-
-A statement of the form {\tt v = b.getClass()} where
-{\tt v} is a local variable of type {\tt java.lang.Class}
-and {\tt b} is a local variable of reference type.
-
-The reason this kind of statement is modeled is because
-\Chord\ has an emphasis on concurrency and many Java programs
-hold locks on the runtime class of an object
-(which is obtained by calling the native {\tt getClass}
-instance method of the object).
-
-\item
-{\tt AryElemRefInst}
-
-A statement of one of the following forms:
-
-{\tt v = b[*]}  \\
-{\tt b[*] = v} \\
-{\tt * = b[*]} \\
-{\tt b[*] = *}
-
-where {\tt v} and {\tt b} are local variables of reference
-type.
-
-The {\tt *} denotes an expression of primitive type in the
-original program which is elided due to \Chord's emphasis on the
-analysis of objects.
-\end{itemize}
-
-\texonly{\newpage}
-\section{Program Transformations}
-\label{sec-prog-transform}
-
-After building the intermediate representation of the
-program, \Chord\ performs
-the following source-to-source transformations 
-that simplify several program analyses:
-
-\begin{enumerate}
-\item
-{\bf Sync Removal}
-
-Removes all synchronized blocks from each method M in each class
-C in the program by generating a fresh static method (in class C)
-per synchronized block in the body of M.
-
-\item
-{\bf Loop Removal}
-
-Removes all loops from each method M in each class C in the program
-by generating a fresh static method (in class C)
-per loop in the body of M.
-
-\item
-{\bf Skip Removal}
-
-Removes all skip (i.e., no-op) statements from each method in each class
-in the program.
-Such statements arise primarily because \Chord, due to its emphasis
-on the analysis of objects, replaces each
-statement in the original program that does not access any
-value of reference type (i.e., only accesses values of primitive
-type) by a skip statement.
-
-\end{enumerate}
-
-\texonly{\newpage}
-\section{Task Graph Construction}
-\label{sec-tg}
-
-\Chord\ requires the following three system properties to construct the task graph:
-
-\begin{enumerate}
-\item
-{\tt chord.dlogmap.path}
-
-The path specifying the mapping from the name of each Datalog-based program
-analysis that must be included in the project to the location
-of the file defining that program analysis.
-
-Each path element must be the location of a file containing a
-list of entries (one per line) of the form:
-
-{\tt <name> <file-name>}
-
-The path elements must be separated by the system-specific path separator
-character (e.g., `:' on UNIX and `;' on Windows).
-
-Let \DlogMapEntries\ denote the set of all entries of the
-above form.
-
-\item
-{\tt chord.trgtmap.path}
-
-The path partially specifying the mapping from the name of each program domain,
-program relation, and Java-based program analysis that must be
-included in the project to the fully-qualified name of a subclass, either
-pre-defined in \Chord\ or defined by the user, of {\tt chord.Dom},
-{\tt chord.Ary*Rel}, and
-{\tt chord.Entity}, respectively, the instance fields of which define the
-state of the corresponding domain, relation, or analysis.
-The `{\tt *}' in {\tt chord.Ary*Rel} is 1, 2, ..., 6, or N for
-program relations of arity 1, 2, ..., 6, or $>$ 6, respectively.
-
-Each path element must be the location of a file containing a
-list of entries (one per line) of the form:
-
-{\tt <name> <class-name>}
-
-The path elements must be separated by the system-specific path separator
-character (e.g., `:` on UNIX and ';' on Windows).
-
-Let \TaskMapEntries\ denote the set of all entries of the
-above form.
-
-\item
-{\tt chord.taskmap.path}
-
-The path partially specifying the mapping from the name of each program domain,
-program relation, and Java-based program analysis that must be included in
-the project to the fully-qualified name of a
-subclass, either pre-defined in \Chord\ or defined by the user, of
-{\tt chord.Dom}, {\tt chord.Ary*Rel}, and {\tt chord.Entity}, respectively,
-which define the following instance methods:
-\begin{itemize}
-\item
-the {\tt execute} method specifying
-the computation that updates the state of the corresponding domain, relation,
-or analysis,
-\item
-the {\tt invalidate} method specifying the action to be taken when the
-state of the corresponding domain, relation, or analysis is invalidated.
-\item
-the {\tt getConsumedNames} method specifying names of any domains, relations,
-and Datalog-based or Java-based program analyses upon whose state the
-computation depends, and
-\item
-the {\tt getProducedNames} method specifying names of any domains, relations,
-and Datalog-based or Java-based program analyses whose state depends upon
-the computation.
-\end{itemize}
-
-Each path element must be the location of a file containing a
-list of entries (one per line) of the form:
-
-{\tt <name> <class-name>}
-
-The path elements must be separated by the system-specific path separator
-character (e.g., `:` on UNIX and ';' on Windows).
-
-Let \TrgtMapEntries\ denote the set of all entries of the above form.
-\end{enumerate}
-
-Then, \Chord\ uses the values of \DlogMapEntries, \TaskMapEntries, and \TrgtMapEntries\ 
-to do the following:
-
-\begin{enumerate}
-\item
-Compute {\tt Tasks\ :\ Set<chord.Entity>}, the set of all tasks,
-each of which is an object of (a subclass of) {\tt chord.Entity},
-as follows:
-
-\begin{enumerate}
-\item
-For each entry in \DlogMapEntries\ (i.e., a Datalog-based program analysis,
-there is a unique task, namely, a unique object of class {\tt chord.DatalogAnalysis}.
-
-\item
-For each entry in \TaskMapEntries\ (i.e., of the form {\tt <name> <class-name>} specifying
-a program domain, a program relation, or a Java-based program analysis), there is a
-unique task, namely, a unique object
-of the class named {\tt <class-name>} which must be (a subclass of) {\tt chord.Dom} in the
-case of a program domain, (a subclass of) {\tt chord.Ary*Rel} in the case of a
-program relation, and (a subclass of) {\tt chord.Entity} in the case of a Java-based
-program analysis.
-\end{enumerate}
-
-\item
-Compute the following maps as shown below:
-
-{\tt ConsumedNameToTasksMap\ :\ String -> Set<Task>},
-mapping each name to the set of (possibly empty) tasks that consume that name.
-
-{\tt NameToTaskMap\ :\ String -> Task},
-mapping each name to the unique task that produces that name.
-
-\Chord\ reports an error if a name is produced by no task or by more than one task.
-{\small 
-\begin{verbatim}
-for each (<task> in Tasks) {
-    case <task> came from item 1(a) above: {
-        // <task> corresponds to a Datalog-based program analysis
-        for each (<name> declared as input relation) {
-            add <task> to ConsumedNameToTasksMap(<name>)
-        }
-        for each (<name> declared as output relation) {
-            if (<name> is in domain of map NameToTaskMap)
-                error ``multiple tasks produce name <name>''
-            else
-                NameToTaskMap(<name>) = <task>
-        }
-    }
-    case <task> came from item 1(b) above: {
-        // <task> corresponds to a program domain, a program relation, or
-        // a Java-based program analysis
-        // Suppose <task> came from entry <name> <class-name> in
-        // TaskMapEntries
-        // i.e. <task> is an object of class named <class-name>
-        for each (<name2> in <task>.getConsumedNames())
-            add <task> to ConsumedNameToTaskMaps(<name2>)
-        for each (<name2> in <task>.getProducedNames() union { <name> })
-            if (<name2> is in domain of map NameToTaskMap)
-                error ``multiple tasks produce name <name>''
-            else
-                NameToTaskMap(<name2>) = <task>
-    }
-}
-\end{verbatim}
-}
-
-{\small 
-\begin{verbatim}
-for each (<name> in domain of map ConsumedNameToTasksMap) {
-    if (<name> is not in domain of map NameToTaskMap)
-        error ``no task produces name <name>''
-}
-\end{verbatim}
-}
-
-\item
-Compute \NameToTrgtMap : {\tt String -> Entity},
-mapping each name to a unique target, as below.
-\Chord\ reports an error if a name is mapped to multiple targets
-or to no target.
-
-{\small 
-\begin{verbatim}
-for each (entry <name> <class-name> in TaskMapEntries) {
-    if (<name> is in domain of map NameToTrgtMap)
-        error ``name <name> mapped to multiple targets''
-    NameToTrgtMap(<name>) = NameToTaskMap(<name>)
-}
-for each (entry <name> <class-name> in TrgtMapEntries) {
-    if (<name> is in domain of map NameToTrgtMap)
-        error ``name <name> mapped to multiple targets''
-    NameToTrgtMap(<name>) = a fresh object of class named <class-name>
-}
-\end{verbatim}
-}
-
-To reduce user burden in providing \TaskMapEntries\ and
-\TrgtMapEntries, \Chord\ implicitly extends \NameToTrgtMap\ using
-\DlogMapEntries\ as follows:
-
-{\small 
-\begin{verbatim}
-for each (<name> in domain of map NameToTaskMap) {
-    if (<name> is not in domain of map NameToTrgtMap) {
-        if (<name> is declared as a domain in some datalog file in DlogMapEntries)
-            NameToTrgtMap(<name>) = a fresh object of class chord.Dom
-        else if (<name> is declared as a (input or output) relation
-                in some datalog file in DlogMapEntries)
-            NameToTrgtMap(<name>) = a fresh object of class chord.Rel
-        else
-            error ``<name> not mapped to any target''
-    } else {
-        chord.Entity <trgt> = NameToTrgtMap(<name>)
-        if (<name> is declared as a domain in some datalog file in DlogMapEntries
-                and <trgt> is not of type (subclass of) chord.Dom)
-            error ``<name> has conflicting types''
-        else if (<name> is declared as a (input or output) relation
-                in some datalog file in DlogMapEntries
-                and <trgt> is not of type (subclass of) chord.Rel)
-            error ``<name> has conflicting types''
-    }
-}
-\end{verbatim}
-}
-
-\item
-Compute graph \TaskGraph\ as a graph whose:
-
-\begin{itemize}
-\item
-set of nodes is \Tasks\ and
-\item
-set of edges containing an edge from task {\tt <task1>} to task {\tt <task2>} 
-whenever there exists a name {\tt <name>} such that
-\NameToTaskMap({\tt <name>}) = {\tt <task1>} and
-\ConsumedNameToTasksMap({\tt <name>}) contains {\tt <task2>}.
-\end{itemize}
-
-The task graph in XML and HTML formats is dumped to files {\tt taskgraph.xml}
-and {\tt taskgraph.html} in the project's working directory specified using
-system property {\tt chord.work.dir}.
-
-\end{enumerate}
-
-\texonly{\newpage}
-\section{Project Organization}
-
-\Chord\ interprets system property {\tt chord.trgts}, if defined, as
-a space-separated list of names of program domains, program relations, and
-Datalog-based or Java-based program analyses, denoted \TrgtNames,
-to be computed in that order.
-
-To avoid redundant computation, \Chord\ maintains during its execution a set of
-up-to-date tasks, denoted \ValidTasks,
-which is a subset of the set of all tasks, denoted \Tasks.
-Set \ValidTasks\ is empty at startup.
-A future release of \Chord\ may provide the ability to retrieve a set of up-to-date tasks from
-a previous run of \Chord\ to prevent re-computing those tasks in subsequent runs.
-The following procedure is executed upon startup:
-
-{\small
-\begin{verbatim}
-for each (<name> in TrgtNames in order) {
-    if (<name> is not in domain of map NameToTaskMap) {
-        error ``name <name> is an invalid target name''
-    else {
-        chord.Entity <task> = NameToTaskMap(<name>)
-        executeTask(<task>)
-    }
-}
-\end{verbatim}
-}
-
-\noindent where procedure {\tt executeTask} is defined as follows:
-
-{\small
-\begin{verbatim}
-procedure executeTask(chord.Entity <task>) {
-    if (<task> is not in ValidTasks) {
-        for each (immed. pred. <task2> of <task> in TaskGraph)
-            executeTask(<task2>)
-        <task>.execute()
-        for each (immed. succ. <task2> of <task> in TaskGraph)
-            invalidateTask(<task2>)
-        add <task> to ValidTasks
-    }
-}
-\end{verbatim}
-}
-
-\noindent where procedure {\tt invalidateTask} is defined as follows:
-
-{\small
-\begin{verbatim}
-procedure invalidateTask(chord.Entity t) {
-    if (<task> is in ValidTasks) {
-        remove <task> from ValidTasks
-        <task>.invalidate()
-        for each (immed. succ. <task2> of <task> in TaskGraph)
-            invalidateTask(<task2>)
-    }
-}
-\end{verbatim}
-}
-
-\texonly{\newpage}
-\section{Chord System Properties}
-\label{sec-sys-prop}
-
-\Chord\ recognizes the following system properties:
-
-\begin{itemize}
-\item
-{\tt chord.out}
-
-The location of a file to which the standard output stream
-of \Chord's execution is redirected.
-
-It may be the same file to which the standard error stream is redirected (see
-the description of system property {\tt chord.err}).
-
-The file is created if it does not exist and its contents are overwritten if it exists.
-
-\item
-{\tt chord.err}
-
-The location of a file to which the standard error stream
-of \Chord's execution is redirected.
-
-It may be the same file to which the standard output stream is redirected (see
-the description of system property {\tt chord.out}).
-
-The file is created if it does not exist and its contents are overwritten if it exists.
-
-\item
-{\tt chord.program.builder.class}
-
-The fully-qualified name of the class to be used for building the
-intermediate representation of the program to be analyzed.
-It must be a concrete class with a nullary constructor that
-implements interface {\tt chord.program.builders.IProgramBuilder}.
-
-The default value of this property is {\tt chord.program.builders.SootProgramBuilder}
-which uses the Soot Java compiler framework \cite{Soot}.
-
-\item
-{\tt chord.serial.file}
-
-The location of a file from (resp.\ to) which
-the serialized form of the intermediate representation of the
-program to be analyzed is loaded (resp.\ stored).
-
-\item
-{\tt chord.main.class}
-
-The fully-qualified name of the main class (e.g., {\tt foo.Main}) of the
-program to be analyzed.
-
-This system property (and {\tt chord.class.path})
-must be defined if system property {\tt chord.serial.file}
-does not specify the file containing the serialized form of the intermediate representation
-of the program to be analyzed.
-
-\item
-{\tt chord.class.path}
-
-The application classpath of the program to be analyzed (as opposed to the
-JDK library classpath which consists of jar files such as {\tt rt.jar}
-which are implicitly appended to the program's application classpath to
-yield the complete program classpath).
-
-Each path element must be the location of a directory or a jar/zip file.
-
-The path elements must be separated by the system-specific path separator character
-(e.g., `:' on UNIX and `;' on Windows).
-
-This system property (and {\tt chord.main.class}) must
-be defined if system property {\tt chord.serial.file}
-does not specify the file containing the serialized form of the intermediate representation
-of the program to be analyzed.
-
-\item
-{\tt chord.src.path}
-
-The Java source path of the program to be analyzed.
-
-Each path element must be the location of a directory.
-
-The purpose of this path is to provide the location of the
-{\tt .java} files of the program being analyzed so that, for instance, the
-results of an analysis (which operates on an intermediate representation of the program
-extracted from {\tt .class} files) can be mapped to the corresponding {\tt .java}
-files and presented to the user for manual inspection.
-
-Each file having suffix {\tt .java} in a directory (or a sub-directory of a directory)
-specified in this path is regarded as a source file of the program.
-
-\item
-{\tt chord.dlog.analysis.path}
-
-The path specifying the mapping from the name of each Datalog-based program
-analysis that must be included in the project to the location
-of the file defining that program analysis.
-
-Each path element must be the location of a file containing a
-list of entries (one per line) of the form:
-
-{\tt <name> <file-name>}
-
-The path elements must be separated by the system-specific path separator character
-(e.g., `:' on UNIX and `;' on Windows).
-
-\item
-{\tt chord.java.analysis.path}
-
-The path specifying the mapping from the name of each program domain, program relation,
-and Java-based program analysis that must be included in the project to the
-fully-qualified name of a subclass, either pre-defined in \Chord\ or defined by the user, of
-{\tt chord.Dom}, {\tt chord.Ary*Rel}, and {\tt chord.Entity}, respectively, which
-define the following instance methods:
-\begin{itemize}
-\item
-the {\tt execute} method specifying
-the computation that updates the state of the corresponding domain, relation,
-or analysis, 
-\item
-the {\tt invalidate} method specifying the action to be taken when the
-state of the corresponding domain, relation, or analysis is invalidated.
-\item
-the {\tt getConsumedNames} method specifying names of any domains, relations,
-and analyses upon whose state the computation depends, and
-\item
-the {\tt getProducedNames} method specifying names of any domains, relations,
-and analyses whose state depends upon the computation.
-\end{itemize}
-
-Each path element must be the location of a file containing a
-list of entries (one per line) of the form:
-
-{\tt <name> <class-name>}
-
-The path elements must be separated by the system-specific path separator character
-(e.g., `:' on UNIX and `;' on Windows).
-
-\item
-{\tt chord.trgtmap.path}
-
-The path specifying the mapping from the name of each program domain,
-program relation, and Java-based program analysis that must be
-included in the project to the fully-qualified name of a subclass,
-either pre-defined in \Chord\ or defined by the user, of {\tt chord.Dom},
-{\tt chord.Ary*Rel}, and
-{\tt chord.Entity}, respectively, the instance fields of which define the
-state of the corresponding domain, relation, or analysis.
-
-Each path element in the list must be the location of a file containing a
-list of entries (one per line) of the form:
-
-{\tt <name> <class-name>}
-
-The path elements must be separated by the system-specific path separator character
-(e.g., `:` on UNIX and ';' on Windows).
-
-\item
-{\tt chord.analyses}
-
-A space-separated list of the names of program domains, program relations, and
-Java-based or Datalog-based program analyses that must be computed in that order.
-
-\item
-{\tt chord.ext.class.path}
-
-The classpath of user-defined program domains, user-defined program relations,
-and user-defined Java-based program analyses.
-It is implicitly appended to the classpath of pre-defined program domains,
-pre-defined program relations, and pre-defined Java-based program analyses
-to yield the complete runtime classpath of \Chord.
-
-Each path element must be the location of a directory or a jar/zip file.
-
-The path elements must be separated by the system-specific path separator character
-(e.g., `:' on UNIX and `;' on Windows).
-\end{itemize}
-
-\texonly{\newpage}
-\section{Pre-defined Domains}
-\label{sec-predef-domains}
-
-The following domains are pre-defined in \Chord:
-
-\begin{itemize}
-\item
-$\Cdom$
-
-The domain of all k-object-sensitive contexts.
-
-The $0^{th}$ element in this domain denotes a distinguished hypothetical
-context (denoted \texorhtml{$\epsilon$}{{\it epsilon}} in the literature) that is regarded as the
-lone abstract context in which certain static methods are analyzed.
-The main static method with signature {\tt void main(java.lang.String[])}
-has no callers and is therefore analyzed in this context.
-For simplicity, the k-object-sensitive may alias analysis provided in
-\Chord\ also analyzes each class initializer method in this context.
-
-\item
-$\Edom$
-
-The domain of all statements that access (read or write) an instance field,
-a static field, or an array element, namely, all statements
-represented by instances of (subclasses of) {\tt chord.program.insts.HeapInst}.
-
-\item
-$\Fdom$
-
-The domain of  all instance and static fields.
-
-The $0^{th}$ element in this domain denotes a distinguished hypothetical
-field that is regarded as accessed whenever an array element is
-accessed.
-
-\item
-$\Hdom$
-
-The domain of all object allocation statements, namely, all statements
-represented by instances of {\tt chord.program.insts.ObjValAsgnInst}.
-
-\item
-$\Idom$
-
-The domain of all method invocation statements, namely, all statements
-represented by instances of {\tt chord.program.insts.InvkInst}.
-
-\item
-$\Mdom$
-
-The domain of all instance and static methods.
-
-The $0^{th}$ element in this domain is the main method of the program.
-
-The $1^{st}$ element in this domain is the method with signature
-{\tt void start()} of class {\tt java.lang.Thread}, if this
-method is deemed reachable from the main method of the program.
-
-The above two methods are the root methods of the implicitly spawned
-main thread and each explicitly spawned child thread, respectively.
-Due to \Chord's emphasis on concurrency, these methods are
-referenced frequently by various pre-defined program analyses
-expressed in Datalog, and giving them special indices makes it
-convenient to reference them in those analyses.
-
-\item
-$\Pdom$
-
-The domain of all simple statements, namely, all statements
-represented by instances of (subclasses of) {\tt chord.program.insts.Inst}.
-
-The $0^{th}$ element in this domain is the statement at the unique
-entry point of the main method of the program.
-
-The statements of each method in the program are assigned
-contiguous indices in this domain, with the statements at the
-unique entry and exit points of each method being assigned
-the smallest and largest indices, respectively, of all
-indices assigned to statements in that method.
-
-\item
-$\Tdom$
-
-The domain of all types, including the types of classes, interfaces,
-arrays, and primitives.
-
-\item
-$\Vdom$
-
-The domain of all local variables.
-
-Each local variable declared in each block of each method is
-represented by a unique element in this domain, in particular,
-local variables that have the same name but are declared in
-different methods or in different blocks of the same method
-are represented by different elements in this domain.
-
-All local variables of the same method are assigned contiguous
-indices in this domain, in particular, the set of local
-variables of a method is the disjoint union of its argument
-variables and its temporary variables, and the argument
-variables are assigned contiguous indices in order followed by
-the temporary variables.
-
-\item
-$\Zdom$
-
-The domain of all argument and return variable positions.
-Let $N$ be the largest number or arguments or return variables of any
-method.  Then, this domain contains the $N$ elements $0, 1, ..., N-1$.
-\end{itemize}
-
-\texonly{\newpage}
-\section{Pre-defined Relations}
-\label{sec-predef-relations}
-
-The following relations are pre-defined in \Chord:
-
-\begin{itemize}
-
-\item
-$\cha : \Mdom \times \Tdom \times \Mdom$
-
-Relation containing each tuple $(m_1,t,m_2)$ such that $m_2$ is
-the target method of a method invocation statement whose {\it resolved method}
-is $m_1$ and whose receiver has run-time type $t$.
-It is named $\cha$ because it represents the relation
-computed by Class Hierarchy Analysis.
-
-\item
-$\sub : \Tdom \times \Tdom$
-
-Relation containing each tuple $(s,t)$ such that type $s$ is a
-subtype of type $t$.
-
-\item
-$\HT : \Hdom \times \Tdom$
-
-Relation containing each tuple $(h,t)$ such that object allocation
-statement $h$ allocates objects of type $t$.
-
-\item
-$\VT : \Vdom \times \Tdom$
-
-Relation containing each tuple $(v,t)$ such that local variable $v$
-has type $t$.
-
-\item
-$\MH : \Mdom \times \Hdom$
-
-Relation containing each tuple $(m,h)$ such that method $m$ contains
-object allocation statement $h$.
-
-\item
-$\MI : \Mdom \times \Idom$
-
-Relation containing each tuple $(m,i)$ such that method $m$ contains
-method invocation statement $i$.
-
-\item
-$\ME : \Mdom \times \Edom$
-
-Relation containing each tuple $(m,e)$ such that method $m$ contains
-statement $e$ that accesses (reads or writes) an instance field, a
-static field, or an array element.
-
-\item
-$\instM : \Mdom$
-
-Relation containing all instance (as opposed static) methods.
-
-\item
-$\statM : \Mdom$
-
-Relation containing all static (as opposed instance) methods.
-
-\item
-$\clinitM : \Mdom$
-
-Relation containing all class initializer methods, namely,
-methods having signature {\tt void <clinit>()}.
-
-\item
-$\IinvkArg : \Idom \times \Zdom \times \Vdom$
-
-Relation containing each tuple $(i,z,v)$ such that local variable $v$
-is the $z^{th}$ argument variable of method invocation statement $i$.
-
-\item
-$\IinvkRet : \Idom \times \Zdom \times \Vdom$
-
-Relation containing each tuple $(i,z,v)$ such that local variable $v$
-is the $z^{th}$ return variable of method invocation statement $i$.
-
-\item
-$\MmethArg : \Mdom \times \Zdom \times \Vdom$
-
-Relation containing each tuple $(m,z,v)$ such that local variable
-$v$ is the $z^{th}$ argument variable of method $m$.
-
-\item
-$\MmethRet : \Mdom \times \Zdom \times \Vdom$
-
-Relation containing each tuple $(m,z,v)$ such that local variable
-$v$ is the $z^{th}$ return variable of method $m$.
-
-\item
-$\MspcMethArg : \Mdom \times \Vdom$
-
-Relation containing each tuple $(m,v)$ such that local variable $v$ is
-the implicit {\tt this} argument variable of instance method $m$.
-
-\item
-
-$\MordMethArg : \Mdom \times \Zdom \times \Vdom$
-
-Relation containing each tuple $(m,z,v)$ such that local variable $v$
-is the $z^{th}$ argument variable of method $m$ other than the implicit
-{\tt this} argument variable (in the case in which $m$ is an instance
-method).
-
-\item
-$\MobjVarAsgnInst : \Mdom \times \Vdom \times \Vdom$
-
-Relation containing each tuple $(m,l,r)$ such that method $m$
-contains a copy statement of the form $l = r$ where $l$ and $r$ are
-local variables.
-
-\item
-$\MobjValAsgnInst : \Mdom \times \Vdom \times \Vdom$
-
-Relation containing each tuple $(m,l,h)$ such that method $m$ contains
-object allocation statement $h$ which assigns to local variable $l$.
-
-\item
-$\MclsVarAsgnInst : \Mdom \times \Vdom \times \Vdom$
-
-Relation containing each tuple $(m,l,b)$ such that method $m$
-contains a statement of the form $l = b.{\tt getClass()}$.
-
-\item
-$\MgetInstFldInst : \Mdom \times \Vdom \times \Fdom \times \Vdom$
-
-Relation containing each tuple $(m,b,f,r)$ such that method $m$
-contains a statement of the form $b.f = r$ which assigns to
-instance field or array element $f$ of the object denoted by local
-variable $b$ the value denoted by local variable $r$.
-
-\item
-$\MputInstFldInst : \Mdom \times \Vdom \times \Fdom \times \Vdom$
-
-Relation containing each tuple $(m,b,f,r)$ such that method $m$
-contains a statement of the form $b.f = r$ which writes the value
-denoted by local variable $r$ to instance field or array element
-$f$ of the object denoted by local variable $b$.
-
-\item
-$\MgetStatFldInst : \Mdom \times \Vdom \times \Fdom$
-
-Relation containing each tuple $(m,l,f)$ such that method $m$ contains
-a statement of the form $l = f$ that assigns to local variable $l$ the
-value denoted by static field $f$.
-
-\item
-$\MputStatFldInst : \Mdom \times \Fdom \times \Vdom$
-
-Relation containing each tuple $(m,f,r)$ such that method $m$ contains
-a statement of the form $f = r$ which assigns to static field $f$ the
-value denoted by local variable $r$.
-
-\item
-$\virtIM : \Idom \times \Mdom$
-
-Relation containing each tuple $(i,m)$ such that $m$ is the
-{\it resolved method} of {\tt invokevirtual} or {\tt invokeinterface} method
-invocation statement $i$.
-
-\item
-$\statIM : \Idom \times \Mdom$
-
-Relation containing each tuple $(i,m)$ such that $m$ is the
-{\it resolved method} of {\tt invokestatic} method invocation statement $i$.
-
-\item
-$\specIM : \Idom \times \Mdom$
-
-Relation containing each tuple $(i,m)$ such that $m$ is the
-{\it resolved method} of {\tt invokespecial} method invocation statement $i$.
-
-\item
-$\syncMV : \Mdom \times \Vdom$
-
-Relation containing each tuple $(m,v)$ such that method $m$ is
-synchronized on argument variable $v$.
-
-\item
-$\writeE : \Edom$
-
-Relation containing all statements that write (as opposed to read)
-an instance field, static field, or array element.
-
-\item
-$\EF : \Edom \times \Fdom$
-
-Relation containing each tuple $(e,f)$ such that statement $e$
-accesses (reads or writes) instance field, static field, or
-array element $f$.
-
-\item
-$\VH : \Vdom \times \Hdom$
-
-Relation containing each tuple $(v,h)$ such that local variable $v$
-may point to abstract object $h$, as determined by 0CFA-based may
-alias analysis.
-
-\item
-$\FH : \Fdom \times \Hdom$
-
-Relation containing each tuple $(f,h)$ such that static field $f$ may
-point to abstract object $h$, as determined by 0CFA-based may alias
-analysis.
-
-\item
-
-$\HFH : \Hdom \times \Fdom \times \Hdom$
-
-Relation containing each tuple $(h_1,f,h_2)$ such that instance field
-or array element $f$ of abstract object $h_1$ may point to abstract
-object $h_2$, as determined by 0CFA-based may alias analysis.
-
-\item
-$\IM : \Idom \times \Mdom$
-
-Relation containing each tuple $(i,m)$ such that method invocation
-statement $i$ may call method $m$ as determined by 0CFA-based may alias
-analysis.
-
-\item
-$\nestedLocks : \Cdom \times \Cdom \times \Mdom \times \Cdom \times \Mdom$
-
-Relation containing each tuple $(c,c_1,m_1,c_2,m_2)$ such that in the
-context-sensitive call graph computed by k-object-sensitive may
-alias analysis, abstract thread $c$ may first acquire the lock at
-the entry of method $m_1$ in abstract context $c_1$ and without
-releasing it proceed to acquire the lock at the entry of method
-$m_2$ in abstract context $c_2$.
-
-\item
-$\forkC : \Cdom$
-
-Relation containing all abstract contexts in which the
-{\tt start()} method of class {\tt java.lang.Thread} may
-be reachable from the main method, as determined by
-k-object-sensitive may alias analysis.
-
-\item
-$\CVC : \Cdom \times \Vdom \times \Cdom$
-
-Relation containing each tuple $(c_1,v,c_2)$ such that local variable
-$v$ may point to abstract object $c_2$ in abstract context $c_1$ of its
-containing method, as determined by k-object-sensitive may alias
-analysis.
-
-\item
-$\FC : \Fdom \times \Cdom$
-
-Relation containing each tuple $(f,c)$ such that static field $f$
-may point to abstract object $c$, as determined by k-object-sensitive
-may alias analysis.
-
-\item
-$\CFC : \Cdom \times \Fdom \times \Cdom$
-
-Relation containing each tuple $(c_1,f,c_2)$ such that instance field
-or array element $f$ of abstract object $c_1$ may point to abstract
-object $c_2$, as determined by k-object-sensitive may alias analysis.
-
-\item
-$\CICM : \Cdom \times \Idom \times \Cdom \times \Mdom$
-
-Relation containing each tuple $(c_1,i,c_2,m)$ such that method
-invocation statement $i$ in abstract context $c_1$ of its containing method
-may call method $m$ in abstract context $c_2$, as determined by
-k-object-sensitive may alias analysis.
-
-\item
-$\CHC : \Cdom \times \Hdom \times \Cdom$
-
-Relation containing each tuple $(c_1,h,c_2)$ such that the method
-containing object allocation statement $h$ may be reachable in
-abstract context $c_1$, and $c_2 = (h::c_1) + k$, where $k$ is the k
-value of statement $h$, provided as an argument to k-object-sensitive
-may alias analysis.
-
-\end{itemize}
-
-\texonly{\newpage}
-\section{Pre-defined Datalog-based Analyses}
-\label{sec-predef-datalog-analyses}
-
-\texonly{\newpage}
-\section{Pre-defined Java-based Analyses}
-\label{sec-predef-java-analyses}
-
-\texonly{\newpage}
-\bibliographystyle{plain}
-\bibliography{citations}
-
-\end{document}
-
-\newpage
-\section{Defining Tasks and Targets}
-
-\Chord\ uses the notion of {\it tasks} to express computations
-and the notion of {\it targets} to express the structure of
-inputs/outputs of computations.
-
-Tasks and targets are both implemented uniformly as {\it entities}.
-A generic entity is represented by class {\tt chord.Entity}
-depicted in Figure \ref{fig-entity}.
-
-\begin{figure}
-\begin{verbatim}
-public class Entity {
-    protected static final String[] emptyStringArray = new String[0];
-    // subclasses may override
-    public void execute() {
-        // do nothing
-    }
-    // subclasses may override
-    public String[] getConsumedTargetNames() {
-        return emptyStringArray;
-    }
-    // subclasses may override
-    public String[] getProducedTargetNames() {
-        return emptyStringArray;
-    }
-    ...
-}
-\end{verbatim}
-\caption{Snippet of class {\tt chord.Entity}}
-\label{fig-entity}
-\end{figure}
-
-Whenever an enntity is regarded as a task:
-
-\begin{itemize}
-\item
-The {\tt execute} method defines the computation that must be performed by the task.
-
-\item
-The {\tt getConsumedTargetNames} method specifies the names of targets
-that are inputs to (the computation performed by) the task.
-
-\item
-The {\tt getProducedTargetNames} method specifies the names of targets
-that are outputs of (the computation performed by) the task.
-\end{itemize}
-
-There are four kinds of entities:
-\begin{itemize}
-\item
-A program domain
-
-\item
-A program relation
-
-\item
-A program analysis expressed in Datalog
-
-\item
-A program analysis expressed in Java
-
-\end{itemize}
-
-We next describe each of these kinds of entities in detail.
-
-\subsection{Program Domains}
-
-A program domain is a set of related elements in a given program (e.g.,
-the set of all methods, the set of all object allocation statements, etc.).
-The structure of a generic program domain is represented by
-class {\tt chord.Dom} depicted in Figure \ref{fig-dom}.
-
-\begin{figure}
-\begin{verbatim}
-public class Dom<T> extends Entity {
-    public final IndexMap<T> map;
-    public Dom() {
-        this.map = new IndexMap<T>();
-    }
-    public void execute() {
-        populate();
-        saveToDomFile();
-    }
-    // subclasses cannot override
-    public final String[] getProducedTargetNames() {
-        return emptyStringArray;
-    }
-    // subclasses may override
-    protected void populate() {
-        // do nothing
-    }
-    ...
-\end{verbatim}
-\caption{Snippet of class {\tt chord.Dom}}
-\label{fig-dom}
-\end{figure}
-
-A program domain provides a map from each element in the domain
-to a unique integer in the range $[0..n-1]$ where $n$ is the size of
-the domain.
-This map is required by {\tt bddbddb} (a BDD-based Datalog solver).
-The {\tt execute} method is specialized to construct the map and then
-save the map to disk for use by {\tt bddbddb}.
-Subclasses must override the {\tt populate} method to construct the
-map as desired.
-
-It is not possible to override the {\tt getProducedTargetNames}
-method (note that it is declared {\tt final}) because a program domain
-task is a specialized task whose output is exactly one target, namely,
-the program domain it constructs.
-It is, however, possible to override the {\tt getConsumedTargetNames}
-method (inherited from superclass {\tt chord.Entity}) since constructing
-a program domain may require other targets (i.e., other program
-domains, program relations, or program analyses).
-
-Figure \ref{fig-domI} defines the program domain
-named $\Idom$ as the domain of all method invocation statements.
-Note from method {\tt getConsumedTargetNames} that it takes as input
-the domain named $\Mdom$ which is the domain of all methods.
-This is because the {\tt populate} method refers to that domain.
-Intuitively, the {\tt populate} method gives adjacent indices in
-domain $\Idom$ to method invocation statements contained in
-methods that have adjacent indices in domain $\Mdom$ are given
-adjacent indices in domain $\Idom$.
-
-\begin{figure}
-\begin{verbatim}
-public class DomI extends Dom<InvkInst> {
-    private IndexMap<Method> mapM;
-    private final String[] consumedTaskNames = new String[] { "M" };
-    public String[] getConsumedTargetNames() {
-        return consumedTaskNames;
-    }
-    protected void populate() {
-        DomM domM = (DomM) project.getEntity("M");
-        mapM = domM.map;
-        for (int m = 0; m < mapM.size(); m++) {
-            Method meth = mapM.get(m);
-            Body body = meth.body;
-            if (body == null)
-                continue;
-            Set<Inst> insts = body.cfg.getNodes();
-            for (Inst inst : insts) {
-                if (inst instanceof InvkInst)
-                    map.set((InvkInst) inst);
-            }
-        }
-    }
-    ...
-}
-\end{verbatim}
-\caption{Snippet of class {\tt chord.DomI}}
-\label{fig-domI}
-\end{figure}
-
-A list of pre-defined program domains is presented in Section \ref{sec-predef-domains}.
-
-\subsection{Program Relations}
-
-A program relation is a relation over one or more program domains.
-The structure of a generic program relation is represented by class
-{\tt chord.Rel} depicted in Figure \ref{fig-rel}.
-
-\begin{figure}
-\begin{verbatim}
-\end{verbatim}
-\caption{Snippet of class {\tt chord.Rel}}
-\label{fig-rel}
-\end{figure}
-
-It is not possible to override the {\tt getProducedTargetNames}
-method (note that it is declared {\tt final}) because a program relation
-task is a specialized task whose output is exactly one target, namely,
-the program relation it constructs.
-It is, however, possible to override the {\tt getConsumedTargetNames}
-method (inherited from superclass {\tt chord.Entity}) since constructing
-a program relation may require other targets (i.e., program
-domains, other program relations, or program analyses).
-
-A list of pre-defined program relations is presented in Section \ref{sec-predef-relations}.
-
-\subsection{Program Analyses in Datalog}
-
-A list of pre-defined program analyses in Datalog is presented in
-Section \ref{sec-predef-datalog-analyses}.
-
-\subsection{Program Analyses in Java}
-
-A list of pre-defined program analyses in Java is presented in
-Section \ref{sec-predef-java-analyses}.
+%chord.run.analyses
+%List of names of program analyses to be run in order.
+%chord.java.analysis.path	 path	 Classpath containing program analyses in Java (i.e. @Chord-annotated classes).	${chord.main.dir}/classes/
+%chord.dlog.analysis.path	 path	 Path of directories containing program analyses in Datalog (i.e. *.datalog and *.dlog files).	${chord.main.dir}/src/main/dlog/
+%chord.reuse.rels	 bool	 Construct program relations from BDDs stored on disk (from a previous run of Chord) whenever possible instead of re-computing them.	false
+%chord.publish.results	 bool	 Publish the results of program analyses in HTML.	true

### 2010-04-01 main/doc/writing_analysis.tex

\section{Writing a Program Analysis}
\label{sec:writing-analysis}

TODO

### 2010-04-01 main/doc/writing_dynamic_analysis.tex

\section{Writing a Dynamic Program Analysis}
\label{sec:writing-dynamic-analysis}

{\bf Step 1: Setting up the Instrumentation Scheme}

Determine the instrumentation scheme required by your dynamic analysis, that is,
the kind and format of events to be generated during an instrumented program's
execution. See class \verb+chord.instr.InstrScheme+ for the kinds of supported
events and their formats.

{\bf Step 2: Implementing the Analysis}

Create a class extending \verb+chord.project.DynamicAnalysis+ and overriding
the appropriate methods in it.

The only method that must be compulsorily overridden is method \verb+getInstrScheme()+,
which must return an instance of the instrumentation scheme chosen in Step 1 above,
plus each \verb+process<event>(<args>)+ method that corresponds to event \verb+<event>+
with format \verb+<args>+ enabled by the chosen instrumentation scheme.
All other methods are no-ops if not overridden.

A sample such class called \verb+MyDynamicAnalysis+ is shown below:

\begin{verbatim}
    import chord.project.DynamicAnalysis;
    import chord.instr.InstrScheme;

    // ***TODO***: analysis won't be recognized by Chord without this annotation
    @Chord(name=<name-of-analysis>)

    public class MyDynamicAnalysis extends DynamicAnalysis {
        InstrScheme scheme;

        public InstrScheme getInstrScheme() {
            if (scheme != null)
                return scheme;
            scheme = new InstrScheme();
            // ***TODO***: Choose (<event1>, <args1>), ... (<eventN>, <argsN>)
            // depending upon the kind and format of events required by this
            // dynamic analysis to be generated for this during an instrumented
            // program's execution.
            scheme.set<event1>(<args1>);
            ...
            scheme.set<eventN>(<argsN>);
            return scheme;
        }

        public void initAllPasses() {
            // ***TODO***: User code to be executed once and for all
            // before all instrumented program runs start.
        }

        public void doneAllPasses() {
            // ***TODO***: User code to be executed once and for all
            // after all instrumented program runs finish.
        }

        public void initPass() {
            // ***TODO***: User code to be executed once before each
            // instrumented program run starts.
        }

        public void donePass() {
            // ***TODO***: User code to be executed once after each
            // instrumented program run finishes.
        }

        // User-defined event handlers for this dynamic analysis.
        // No-ops if not overridden.
        public void process<event1>(<args1>) {
            // ***TODO***: User code for handling events of kind <event1> with format <args1>.
        }
        ...
        public void process<eventN>(<argsN>) {
            // ***TODO***: User code for handling events of kind <eventN> with format <argsN>.
        }
    }
\end{verbatim}

{\bf Step 3: Choosing or Implementing the Runtime Event Handler}

Determine the runtime event handler for your dynamic analysis. There are two kinds of runtime event
handlers, offline and online; see package \verb+chord.runtime+ for more details.

For most dynamic analyses, an offline event hander suffices, in which case you don't need to do
anything in this step (except that you may want to change the default values of system properties
\verb+chord.trace.pipe+ and \verb+chord.trace.block.size+ for performance; see
Section \ref{sec:chord-sysprops} for the meaning of these properties). Otherwise, you need to
create a class extending class \verb+chord.runtime.Runtime+ and override the appropriate methods
in it, and provide the fully-qualified name of the created class as the value of system property
\verb+chord.runtime.class+ (whose default value is the class \verb+chord.runtime.BufferedRuntime+).

{\bf Step 4: Compiling the Analysis}

Compile the analysis by placing the directory containing class \verb+MyDynamicAnalysis+ created
above in the path defined by system property \verb+chord.java.analysis.path+.

{\bf Step 5: Specifying Program Inputs}

Provide the IDs of program runs to be generated (say 1, 2, ..., M) and the command-line arguments to be
used for the program in each of those runs (say <args1>, ..., <argsM>) via system properties
\verb+chord.run.ids=1,2,...,N+ and \verb+chord.args.1=<args1>+, ..., \verb+chord.args.M=<argsM>+.

By default, \verb+chord.run.ids=0+ and \verb+chord.args.0=""+, that is, the program will be run only
once (using run ID 0) with no command-line arguments.

{\bf Step 6: Running the Analysis}

To run the analysis, set system property \verb+chord.run.analyses=<name-of-analysis>+ where
\verb+<name-of-analysis>+ is the name that was provided in the \verb+@Chord+ annotation for class
\verb+MyDynamicAnalysis+ created above.

### 2010-04-02 main/doc/building_scope.tex

\section{Building Analysis Scope}
\label{sec:building-scope}

Chord computes the analysis scope (i.e., reachable classes and methods) of a given Java program
either if property \code{chord.build.scope} is set to \code{true} or if some other task (e.g.,
a program analysis specified via property \code{chord.run.analyses}) demands it by
calling static method \code{chord.program.Program.v()}.

The algorithm used to compute the analysis scope is as follows.

\begin{itemize}
\item
If property \code{chord.reuse.scope} has value \code{true} and the files specified by properties
\code{chord.classes.file} and \code{chord.methods.file} exist,
then Chord regards those files as specifying the classes and methods, respectively,
to be regarded as reachable.  The format
of the classes file is a fully-qualified class name per line (e.g., \code{foo.bar.Main}).  The format
of the methods file is an entry of the form \code{<method name>:<method descriptor>@<class name>} per line,
specifying the method's name, the method's descriptor, and the method's declaring class
(e.g., \code{main:([Ljava/lang/String;)V@foo.bar.Main+}).
Chord throws a runtime exception if the declaring class of a method in the methods file is not
specified in the classes file.

\item
If property \code{chord.reuse.scope} has value \code{false} or the file specified by
property \code{chord.classes.file} or \code{chord.methods.file} does not exist,
then Chord computes the analysis scope
using the algorithm specified by property \code{chord.scope.kind} and then
writes the classes and methods deemed reachable by that algorithm to those files.

The possible legal values of property \code{chord.scope.kind} are \code{rta} and \code{dynamic}.
In both cases, Chord expects properties \code{chord.main.class} and \code{chord.class.path}
to be set.

\begin{itemize}
\item
The \code{rta} value instructs Chord to compute the analysis scope statically using Rapid Type Analysis.
\item
The \code{dynamic} value instructs Chord to compute the analysis scope dynamically, by running the program
and observing using JVMTI the classes that are loaded at run-time.
The number of times the program is run and the command-line arguments to be supplied to
the program in each run is specified by properties \code{chord.run.ids} and
\code{chord.args.<id>} for each run ID \code{<id>}.  By default, the program is run only once, using run ID \code{0},
and without any command-line arguments.
Only classes loaded in some run are regarded as reachable but {\it all} methods of each loaded class are regarded
as reachable regardless of whether they were invoked in the run.
The rationale behind this decision is to both reduce run-time overhead of JVMTI and to increase the
predictive power of program analyses performed using the computed analysis scope.
\end{itemize}
\end{itemize}

The default value of property \code{chord.reuse.scope} is \code{false} and that of
property \code{chord.scope.kind} is \code{rta}.
The default values of properties \code{chord.classes.file} and \code{chord.methods.file} are
\code{{chord.out.dir}/classes.txt} and \code{{chord.out.dir}/methods.txt}, respectively,
where property \code{chord.out.dir} defaults to \code{{chord.work.dir}/chord_output/},
and property \code{chord.work.dir} defaults to the current working directory.

### 2010-04-02 main/doc/chord_sysprops.tex

@@
 
-\begin{longtable}{|l|p{0.4in}|p{4.0in}|}
-%\caption{TODO} \\
-\hline
-{\bf Property name} & {\bf Type} & {\bf Description and default value if any} \\
-\endfirsthead
-\hline
-{\bf Property name} & {\bf Type} & {\bf Description and default value if any} \\
-\endhead
-\hline
-\hline
-\multicolumn{3}{|l|}{{\bf Chord resource properties}} \\
-\verb+chord.main.dir+ & location & \verb+main+ directory in Chord's installation. \\
-\hline
-\verb+chord.lib.dir+ & location & Directory containing libraries needed by Chord.
-\newline Default: \verb+${chord.main.dir}/lib/+ \\
-\hline
-\verb+chord.main.class.path+ & path & Classpath of Chord.
-It includes the path specified by property \verb+chord.java.analysis.path+
-to allow running user-defined program analyses.
-\newline Default: See \verb+${chord.main.dir}/build.xml+ \\
-\hline
-\verb+chord.bddbddb.class.path+ & path & Classpath of \verb+bddbddb+.
-\newline Default: See \verb+${chord.main.dir}/build.xml+ \\
-\hline
-\verb+chord.instr.agent.file+ & location &
-Shared library implementing Chord instrumentation agent. \\
-%\newline Default: \verb+${chord.main.dir}/lib/libchord_instr_agent.so+ \\
-\hline
-\verb+chord.javadoc.url+ & string & URL of the Javadoc location of program analyses.
-It is used when publishing targets (i.e., when property \verb+chord.publish.targets+
-is set to true).
-\newline Default: \verb+http://chord.stanford.edu/javadoc/+ \\
-\hline
-\hline
-\multicolumn{3}{|l|}{{\bf Chord boot properties}} \\
-\hline
-\verb+chord.work.dir+ & location & Working directory during Chord's execution.
-\newline Default: current working directory \\
-\hline
-\verb+chord.props.file+ & location & Properties file loaded by \verb+${chord.main.dir}/build.xml+.
-Any of the below properties, as well as other user-defined properties to be passed to Chord
-(e.g., for user-defined analyses), may be set in this file.  Each relative (instead of absolute)
-path element in the value of any property named \verb+chord.<...>.path+ set in this file is
-converted to an absolute path element with respect to the directory containing this file.
-\newline Default: \verb+${chord.work.dir}/chord.properties+ \\
-\hline
-\verb+chord.max.heap+ & string & Maximum memory size of JVM running Chord.
-\newline Default: \verb+1024m+ \\
-\hline 
-\verb+chord.max.stack+ & string & Maximum thread stack size of JVM running Chord.
-\newline Default: \verb+32m+ \\
-\hline
-\verb+chord.jvmargs+ & string & Arguments passed to JVM running Chord. \\
-%\newline Default: \verb+"-showversion -ea -Xmx${chord.max.heap} -Xss${chord.max.stack}"+ \\
-\hline
-\verb+chord.bddbddb.max.heap+ & string & Maximum memory size of JVM running bddbddb.
-\newline Default: \verb+1024m+ \\
-\hline
-\hline
-\multicolumn{3}{|l|}{{\bf Chord task properties}} \\
-\hline
-\verb+chord.build.scope+ & bool & Compute analysis scope (i.e., reachable classes and methods)
-using algorithm specified by properties \verb+chord.scope.kind+ and \verb+chord.reuse.scope+.
+\begin{enumerate}
+\item
+\code{chord.main.dir} \\
+{\bf Type:} location \\
+{\bf Description:} \code{main/} directory in Chord's installation.
+
+\item
+\code{chord.lib.dir} \\
+{\bf Type:} location \\
+{\bf Description:} Directory containing libraries needed by Chord. \\
+{\bf Default value:} \code{{chord.main.dir}/lib/}
+
+\item
+\code{chord.main.class.path} \\
+{\bf Type:} path \\
+{\bf Description:} Classpath of Chord.  It includes the path specified by property \code{chord.java.analysis.path} to allow running user-defined program analyses. \\
+{\bf Default value:} See \code{{chord.main.dir}/build.xml}
+
+\item
+\code{chord.bddbddb.class.path} \\
+{\bf Type:} path \\
+{\bf Description:} Classpath of \code{bddbddb}. \\
+{\bf Default value:} See \code{{chord.main.dir}/build.xml}
+
+\item
+\code{chord.instr.agent.file} \\
+{\bf Type:} location \\
+{\bf Description:} Shared library implementing Chord instrumentation agent. \\
+{\bf Default value:} \code{{chord.main.dir}/lib/libchord_instr_agent.so}
+
+\item
+\code{chord.javadoc.url} \\
+{\bf Type:} string \\
+{\bf Description:} URL of the Javadoc location of program analyses.  It is used when publishing targets (i.e., when property \code{chord.publish.targets} is set to true). \\
+{\bf Default value:} \code{http://chord.stanford.edu/javadoc/}
+
+\item
+\code{chord.work.dir} \\
+{\bf Type:} location \\
+{\bf Description:} Working directory during Chord's execution. \\
+{\bf Default value:} current working directory
+
+\item
+\code{chord.props.file} \\
+{\bf Type:} location \\
+{\bf Description:} Properties file loaded by \code{{chord.main.dir}/build.xml}.
+Any of the below properties, as well as other user-defined properties to be passed to Chord (e.g., for user-defined analyses), may be set in this file.
+Each relative (instead of absolute) path element in the value of any property named \code{chord.<...>.path} set in this file is converted to an absolute path element with respect to the directory containing this file.
+\\
+{\bf Default value:} \code{{chord.work.dir}/chord.properties}
+
+\item
+\code{chord.max.heap} \\
+{\bf Type:} string \\
+{\bf Description:} Maximum memory size of JVM running Chord. \\
+{\bf Default value:} \code{1024m}
+
+\item 
+\code{chord.max.stack} \\
+{\bf Type:} string \\
+{\bf Description:} Maximum thread stack size of JVM running Chord. \\
+{\bf Default value:} \code{32m}
+
+\item
+\code{chord.jvmargs} \\
+{\bf Type:} string \\
+{\bf Description:} Arguments passed to JVM running Chord. \\
+{\bf Default value:} \code{"-showversion -ea -Xmx{chord.max.heap} -Xss{chord.max.stack}"}
+
+\item
+\code{chord.bddbddb.max.heap} \\
+{\bf Type:} string \\
+{\bf Description:} Maximum memory size of JVM running \bddbddb. \\
+{\bf Default value:} \code{1024m}
+
+\item
+\code{chord.build.scope} \\
+ bool \\
+ Compute analysis scope (i.e., reachable classes and methods)
+using algorithm specified by properties \code{chord.scope.kind} and \code{chord.reuse.scope}.
 Program scope is computed regardless of the value of this property if another task (e.g., a
 program analysis to be run) needs it.
-\newline Default: \verb+false+ \\
-\hline
-\verb+chord.run.analyses+ & string list & List of names of program analyses to be run
+{\bf Default value:} \code{false} \\
+
+\item
+\code{chord.run.analyses} \\
+ string list \\
+ List of names of program analyses to be run
 in order.
-\newline Default: \verb+""+ \\
-\hline
-\verb+chord.print.rels+ & string list & List of names of program relations whose contents
-must be printed to files \verb+${chord.out.dir}/<...>.txt+ where \verb+<...>+ denotes the
-relation name. 
-\newline Default: \verb+""+ \\
-\hline
-\verb+chord.publish.targets+ & bool & Publish all targets defined by program analyses in
-paths specified by properties \verb+chord.java.analysis.path+ and
-\verb+chord.dlog.analysis.path+.
-\newline Default: \verb+false+ \\
-\hline
-\hline
-\multicolumn{3}{|l|}{{\bf Basic program properties}} \\
-\hline
-\verb+chord.main.class+ & class & Fully-qualified name of main class of given program
-(e.g., \verb+com.example.Main+). \\
-\hline
-\verb+chord.class.path+ & path & Classpath of given program. \verb+${sun.boot.class.path}+
+{\bf Default value:} \code{""}
+
+\item
+\code{chord.print.rels} \\
+{\bf Type:} string list \\
+{\bf Description:} List of names of program relations whose contents must be printed to files \code{{chord.out.dir}/<...>.txt} where \code{<...>} denotes the relation name.  \\
+{\bf Default value:} \code{""}
+
+\item
+\code{chord.publish.targets} \\
+ bool \\
+ Publish all targets defined by program analyses in
+paths specified by properties \code{chord.java.analysis.path} and
+\code{chord.dlog.analysis.path}.
+{\bf Default value:} \code{false} \\
+
+\item
+\code{chord.main.class} \\
+ class \\
+ Fully-qualified name of main class of given program
+(e.g., \code{com.example.Main}). \\
+
+\item
+\code{chord.class.path} \\
+{\bf Type:} path \\
+ Classpath of given program. \code{{sun.boot.class.path}}
 is implicitly appended to it.
-\newline Default: \verb+${sun.boot.class.path}+ \\
-\hline
-\verb+chord.src.path+ & path & Java source path of given program.
-\newline Default: \verb+""+ \\
-\hline
-\verb+chord.run.ids+ & string list & List of IDs to identify runs of given program.
-\newline Default: \verb+0+ \\
-\hline
-\verb+chord.args.XXX+ & string & Command-line argument string to be used for given
-program in run having ID \verb+XXX+.
-\newline Default: \verb+""+ \\
-\hline
-\verb+chord.runtime.jvmargs+ & string & Arguments to JVM when running given program.
-It is used when the program is executed to compute its scope (i.e., when property
-\verb+chord.scope.kind+ is set to \verb+dynamic+) as well as when the instrumented
-program is executed for dynamic analyses.
-\newline Default: \verb+"-ea -Xmx1024m"+ \\ 
-\hline
-\hline
-\multicolumn{3}{|l|}{{\bf Program scope properties}} \\ 
-\hline
-\verb+chord.scope.kind+ & [\verb+rta+ $|$ \verb+dynamic+] & Algorithm to compute
-analysis scope (i.e., reachable classes and methods).  Current choices include
-\verb+rta+ (static Rapid Type Analysis) and \verb+dynamic+ (dynamic analysis).
-\newline Default: \verb+rta+ \\
-\hline
-\verb+chord.reuse.scope+ & bool & Treat analysis scope as the classes and methods listed
-in files specified by properties \verb+chord.classes.file+ and \verb+chord.methods.file+,
-respectively. Property \verb+chord.scope.kind+ is ignored if this property is set to
-\verb+true+ and the two files exist.
-\newline Default: \verb+false+ \\
-\hline
-\verb+chord.check.exclude+ & string list & List of prefixes of names of classes and
+{\bf Default value:} \code{{sun.boot.class.path}}
+
+\item
+\code{chord.src.path} \\
+{\bf Type:} path \\
+ Java source path of given program. \\
+{\bf Default value:} \code{""}
+
+\item
+\code{chord.run.ids} \\
+ string list \\
+ List of IDs to identify runs of given program.
+{\bf Default value:} \code{0}
+
+\item
+\code{chord.args.XXX} \\
+ string \\
+ Command-line argument string to be used for given
+program in run having ID \code{XXX}.
+{\bf Default value:} \code{""}
+
+\item
+\code{chord.runtime.jvmargs} \\
+{\bf Type:} string \\
+{\bf Description:} Arguments to JVM when running given program.
+It is used when the program is executed to compute its scope (i.e., when property \code{chord.scope.kind} is set to \code{dynamic}) as well as when the instrumented program is executed for dynamic analyses.
+\\
+{\bf Default value:} \code{"-ea -Xmx1024m"}
+
+\item
+\code{chord.scope.kind} \\
+{\bf Type:} [\code{rta} $|$ \code{dynamic}] \\
+{\bf Description:} Algorithm to compute analysis scope (i.e., reachable classes and methods).  Current choices include \code{rta} (static Rapid Type Analysis) and \code{dynamic} (dynamic analysis). \\
+{\bf Default value:} \code{rta}
+
+\item
+\code{chord.reuse.scope} \\
+ bool \\
+ Treat analysis scope as the classes and methods listed
+in files specified by properties \code{chord.classes.file} and \code{chord.methods.file},
+respectively. Property \code{chord.scope.kind} is ignored if this property is set to
+\code{true} and the two files exist.
+{\bf Default value:} \code{false}
+
+\item
+\code{chord.check.exclude} \\
+ string list \\
+ List of prefixes of names of classes and
 packages to be excluded by program analyses.
 Interpretation of this property is analysis-specific. \\
-%\newline Default: \verb+"java.,sun.,com.sun.,com.ibm.,org.apache.harmony.,joeq.,jwutil."+ \\
-\hline
-\verb+chord.instr.exclude+ & string list & List of prefixes of names of classes and
-packages to be excluded from being instrumented.  Classes in package \verb+java.lang.ref+
-are not instrumented regardless of the value of this property. \\
-%\newline Default: \verb+"java.,sun.,com.sun.,com.ibm.,org.apache.harmony.,joeq.,jwutil."+ \\
-\hline
-\hline
-\multicolumn{3}{|l|}{{\bf Program analysis properties}} \\
-\hline
-\verb+chord.java.analysis.path+ & path & Classpath containing program analyses written
-in Java (i.e., \verb+@Chord+-annotated classes).
-\newline Default: \verb+${chord.main.dir}/classes/+ \\
-\hline
-\verb+chord.dlog.analysis.path+ & path & Path of directories containing program analyses
-written in Datalog (i.e., \verb+*.datalog+ and \verb+*.dlog+ files).
-\newline Default: \verb+${chord.main.dir}/src/dlog/+ \\
-\hline
-\verb+chord.reuse.rels+ & bool & Construct program relations from BDDs stored on disk
+{\bf Default value:} \code{"java.,sun.,com.sun.,com.ibm.,org.apache.harmony.,joeq.,jwutil."} \\
+
+\item
+\code{chord.instr.exclude} \\
+{\bf Type:} string list \\
+{\bf Description:} List of prefixes of names of classes and packages to be excluded from being instrumented.
+Classes in package \code{java.lang.ref} are not instrumented regardless of the value of this property.
+\\
+{\bf Default value:} \code{"java.,sun.,com.sun.,com.ibm.,org.apache.harmony.,joeq.,jwutil."}
+
+\item
+\code{chord.java.analysis.path} \\
+{\bf Type:} path \\
+{\bf Description:} Classpath containing program analyses written in Java (i.e., \code{@Chord}-annotated classes). \\
+{\bf Default value:} \code{{chord.main.dir}/classes/}
+
+\item
+\code{chord.dlog.analysis.path} \\
+{\bf Type:} path \\
+{\bf Description:} Path of directories containing program analyses written in Datalog (i.e., \code{*.datalog} and \code{*.dlog} files). \\
+{\bf Default value:} \code{{chord.main.dir}/src/dlog/}
+
+\item
+\code{chord.reuse.rels} \\
+ bool \\
+ Construct program relations from BDDs stored on disk
 (from a previous run of Chord) whenever possible instead of re-computing them.
-\newline Default: \verb+false+ \\
-\hline
-\verb+chord.publish.results+ & bool & Publish the results of program analyses in HTML.
+{\bf Default value:} \code{false} \\
+\item
+\code{chord.publish.results} \\
+ bool \\
+ Publish the results of program analyses in HTML.
 Interpretation of this property is analysis-specific.
-\newline Default: \verb+true+ \\
-\hline
-\hline
-\multicolumn{3}{|l|}{{\bf Program transformation properties}} \\
-\hline
-\verb+chord.ssa+ & bool & Do SSA transformation for all methods deemed reachable by the
+{\bf Default value:} \code{true} \\
+
+\item
+\code{chord.ssa} \\
+ bool \\
+ Do SSA transformation for all methods deemed reachable by the
 algorithm used to compute analysis scope.
-\newline Default: \verb+true+ \\
-\hline
-\hline
-\multicolumn{3}{|l|}{{\bf Chord debug properties}} \\
-\hline
-\verb+chord.bddbddb.noisy+ & bool & Make bddbddb output debug info when solving Datalog programs.
-\newline Default: \verb+false+ \\
-\hline
-\verb+chord.save.maps+ & bool & Write to file \verb+${chord.bddbddb.work.dir}/<...>.map+
-when saving program domain named \verb+<...>+ (useful for debugging Datalog programs using
-\verb+debug+ target of \verb+${chord.main.dir}/build.xml+).
-\newline Default: \verb+true+ \\
-\hline
-\hline
-\multicolumn{3}{|l|}{{\bf Chord instrumentation properties}} \\
-\hline
-\verb+chord.trace.pipe+ & bool & Implement the dynamic trace file as a POSIX pipe instead of a regular file.
-\newline Default: \verb+true+ \\
-\hline
-\verb+chord.trace.block.size+ & int & Number of bytes to read/write in a single operation from/to the dynamic trace file.
-\newline Default: \verb+4096+ \\
-\hline
-\verb+chord.runtime.class+ & class & Subclass of \verb+chord.project.Runtime+ used by dynamic program
-analyses to handle events generated during an instrumented program's execution.
-\newline Default: \verb+chord.project.BufferedRuntime+ \\
-\hline
-\hline
-\multicolumn{3}{|l|}{{\bf Chord output properties}} \\
-\hline
-\verb+chord.out.dir+ & location & Directory to which Chord dumps all files.
-\newline Default: \verb+${chord.work.dir}/chord_output/+ \\
-\hline
-\verb+chord.out.file+ & loc. & File to redirect standard output stream during Chord's execution.
-\newline Default: \verb+${chord.out.dir}/log.txt+ \\
-\hline
-\verb+chord.err.file+ & loc. & File to redirect standard error stream during Chord's execution.
-\newline Default: \verb+${chord.out.dir}/log.txt+ \\
-\hline
-\verb+chord.classes.file+ & loc. & File from/to which list of classes deemed reachable is read/written.
-\newline Default: \verb+${chord.out.dir}/classes.txt+ \\
-\hline
-\verb+chord.methods.file+ & loc. & File from/to which list of methods deemed reachable is read/written.
-\newline Default: \verb+${chord.out.dir}/methods.txt+ \\
-\hline
-\verb+chord.bddbddb.work.dir+ & loc. & Directory used by bddbddb as its input/output directory.
-\newline Default: \verb+${chord.out.dir}/bddbddb/+ \\
-\hline
-\verb+chord.boot.classes.dir+ & loc. & Directory containing instrumented JDK classes used by given program.
-\newline Default: \verb+${chord.out.dir}/boot_classes/+ \\
-\hline
-\verb+chord.user.classes.dir+ & loc. & Directory containing instrumented non-JDK classes of given program.
-\newline Default: \verb+${chord.out.dir}/user_classes/+  \\
-\hline
-\verb+chord.instr.scheme.file+ & loc. & File specifying kind and format of events in dynamic trace file.
-\newline Default: \verb+${chord.out.dir}/scheme.ser+ \\
-\hline
-\verb+chord.crude.trace.file+ & loc. & Crude dynamic trace file.
-\newline Default: \verb+${chord.out.dir}/crude_trace.txt+ \\
-\hline
-\verb+chord.final.trace.file+ & loc. & Final dynamic trace file.
-\newline Default: \verb+${chord.out.dir}/final_trace.txt+ \\
-\hline
-\end{longtable}
+{\bf Default value:} \code{true} \\
+
+\item
+\code{chord.bddbddb.noisy} \\
+ bool \\
+ Make \bddbddb\ output debug info when solving Datalog programs.
+{\bf Default value:} \code{false} \\
+
+\item
+\code{chord.save.maps} \\
+ bool \\
+ Write to file \code{{chord.bddbddb.work.dir}/<...>.map}
+when saving program domain named \code{<...>} (useful for debugging Datalog programs using
+\code{debug} target of \code{{chord.main.dir}/build.xml}).
+{\bf Default value:} \code{true} \\
+
+\item
+\code{chord.trace.pipe} \\
+ bool \\
+ Implement the dynamic trace file as a POSIX pipe instead of a regular file.
+{\bf Default value:} \code{true} \\
+
+\item
+\code{chord.trace.block.size} \\
+{\bf Type:} int \\
+{\bf Description:} Number of bytes to read/write in a single operation from/to the dynamic trace file.
+{\bf Default value:} \code{4096} \\
+
+\item
+\code{chord.runtime.class} \\
+{\bf Type:} class \\
+{\bf Description:} Subclass of \code{chord.project.Runtime} used by dynamic program analyses to handle events generated during an instrumented program's execution. \\
+{\bf Default value:} \code{chord.project.BufferedRuntime}
+
+\item
+\code{chord.out.dir} \\
+{\bf Type:} location \\
+{\bf Description:} Directory to which Chord dumps all files. \\
+{\bf Default value:} \code{{chord.work.dir}/chord_output/}
+
+\item
+\code{chord.out.file} \\
+{\bf Type:} location \\
+{\bf Description:} File to redirect standard output stream during Chord's execution. \\
+{\bf Default value:} \code{{chord.out.dir}/log.txt}
+
+\item
+\code{chord.err.file} \\
+{\bf Type:} location \\
+{\bf Description:} File to redirect standard error stream during Chord's execution. \\
+{\bf Default value:} \code{{chord.out.dir}/log.txt}
+
+\item
+\code{chord.classes.file} \\
+{\bf Type:} location \\
+{\bf Description:} File from/to which list of classes deemed reachable is read/written. \\
+{\bf Default value:} \code{{chord.out.dir}/classes.txt}
+
+\item
+\code{chord.methods.file} \\
+{\bf Type:} location \\
+{\bf Description:} File from/to which list of methods deemed reachable is read/written. \\
+{\bf Default value:} \code{{chord.out.dir}/methods.txt}
+
+\item
+\code{chord.bddbddb.work.dir} \\
+{\bf Type:} location \\
+{\bf Description:} Directory used by \bddbddb\ as its input/output directory. \\
+{\bf Default value:} \code{{chord.out.dir}/bddbddb/}
+
+\item
+\code{chord.boot.classes.dir} \\
+{\bf Type:} location \\
+{\bf Description:} Directory containing instrumented JDK classes used by given program. \\
+{\bf Default value:} \code{{chord.out.dir}/boot_classes/}
+
+\item
+\code{chord.user.classes.dir} \\
+{\bf Type:} location \\
+{\bf Description:} Directory containing instrumented non-JDK classes of given program. \\
+{\bf Default value:} \code{{chord.out.dir}/user_classes/}
+
+\item
+\code{chord.instr.scheme.file} \\
+{\bf Type:} location \\
+{\bf Description:} File specifying kind and format of events in dynamic trace file. \\
+{\bf Default value:} \code{{chord.out.dir}/scheme.ser}
+
+\item
+\code{chord.crude.trace.file} \\
+{\bf Type:} location \\
+{\bf Description:} Crude dynamic trace file. \\
+{\bf Default value:} \code{{chord.out.dir}/crude_trace.txt}
+
+\item
+\code{chord.final.trace.file} \\
+{\bf Type:} location \\
+{\bf Description:} Final dynamic trace file. \\
+{\bf Default value:} \code{{chord.out.dir}/final_trace.txt}
+\end{enumerate}

### 2010-04-02 main/doc/datarace.tex

\section{Datarace Analysis}
\label{sec:datarace}

Run the datarace analysis provided in Chord by running the following command in Chord's \code{main/} directory:

\begin{quote}
\begin{verbatim}
prompt> ant -Dchord.work.dir=<...> -Dchord.run.analyses=datarace-java run
\end{verbatim}
\end{quote}

\noindent where directory \code{<...>} contains a file named \code{chord.properties} which
defines properties \code{chord.main.class}, \code{chord.class.path}, and \code{chord.src.path}.
See Section \ref{sec:chord-sysprops} for the meaning of these properties.

Directory \code{main/examples/datarace_test/} provides a toy Java program on which you can run the datarace
analysis.
First run \code{ant} in that directory (in order to compile the program's \code{.java} files to
\code{.class} files) and then run the above command in Chord's \code{main/} directory
with \code{<...>} replaced by \code{examples/datarace_test/}.
Upon successful completion, the following files should be produced in directory
\code{main/examples/datarace_test/chord_output/}:

\begin{itemize}
\item
File \code{dataraces_by_fld.html}, listing all dataraces grouped by the field on which they occur; all
dataraces on the same instance field or the same static field are listed in the same group, and so are
all dataraces on array elements.
\item
File \code{dataraces_by_obj.html}, listing all dataraces grouped by the abstract object on whose field they occur;
dataraces on all static fields are listed in the same group, and so are dataraces on different
instance fields of the same abstract object.
\end{itemize}

### 2010-04-02 main/doc/program_setup.tex

@@
-\section{Setting up a Java Program for Analysis}
+\section{Setting up a Java Program}
 \label{sec:program-setup}
@@
 
-The above structure is typical: the program's Java source files are under \verb+src/+,
-its class files are under \verb+classes/+, the source and jar files of the libraries
-used by the program are under \verb+lib/src/+ and \verb+lib/jar/+, respectively.
+\noindent The above structure is typical: the program's Java source files are under \code{src/},
+its class files are under \code{classes/}, the source and jar files of the libraries
+used by the program are under \code{lib/src/} and \code{lib/jar/}, respectively.
 
-File \verb+chord.properties+ specifies properties to be passed to Chord (alternatively,
-these properties may be passed on the command-line, in format \verb+-D<name>=<value>+).
+File \code{chord.properties} specifies properties to be passed to Chord (alternatively,
+these properties may be passed on the command-line, in format \code{-D<name>=<value>}).
 A sample such file for the above program is as follows:
@@
 
-Each relative (instead of absolute) path element in the value of any property
-named \verb+chord.<...>.path+ defined in this file is converted to an absolute path element with respect
-to the directory containing this file, which in the above case is \verb+example/+.
+Each relative path element in the value of any property
+named \code{chord.<...>.path} defined in this file is converted to an absolute path element with respect
+to the directory containing this file, which in the above case is \code{example/}.
 
@@
 Here, we only describe those defined in the above
-sample \verb+chord.properties+ file:
+sample \code{chord.properties} file:
 
@@
 \item
-\verb+chord.main.class+ specifies the fully-qualified name of the main class of the
+\code{chord.main.class} specifies the fully-qualified name of the main class of the
 program.
 \item
-\verb+chord.class.path+ specifies the classpath of the program.
-The value of pre-defined property \verb+${sun.boot.class.path}+ is implicitly appended
+\code{chord.class.path} specifies the classpath of the program.
+The value of pre-defined property \code{sun.boot.class.path} is implicitly appended
 to it.
 \item
-\verb+chord.src.path+ specifies the Java source path of the program.
-All program analyses in Chord operate on Java bytecode (specified by \verb+chord.class.path+).
+\code{chord.src.path} specifies the Java source path of the program.
+All program analyses in Chord operate on Java bytecode (specified by \code{chord.class.path}).
 The only use of this property is to HTMLize the Java source files of the program so that the
@@
 \item
-\verb+chord.run.ids+ specifies a list of IDs to identify runs of the program.
+\code{chord.run.ids} specifies a list of IDs to identify runs of the program.
 It is used by dynamic program analyses to determine how many times the program must be run.
 An additional use of this property is to allow specifying the command-line arguments to use
-in the run having ID \verb+XXX+ via property \verb+chord.args.XXX+, as illustrated by
-properties \verb+chord.args.0+ and \verb+chord.args.1+ in the above example.
+in the run having ID \code{<id>} via property \code{chord.args.<id>}, as illustrated by
+properties \code{chord.args.0} and \code{chord.args.1} in the above example.
 \end{itemize}
 
-To run Chord on the above program, run the following command in Chord's \verb+main/+ directory:
+To run Chord on the above program, run the following command in Chord's \code{main/} directory:
 
@@
 
-where \verb+<...>+ denotes either the absolute path, or the path relative to Chord's main directory,
+\noindent where \code{<...>} denotes either the absolute path, or the path relative to Chord's \code{main/} directory,
 of the parent of directory {\tt example/}.
-System property \verb+chord.work.dir+ specifies the working directory during Chord's execution.
+System property \code{chord.work.dir} specifies the working directory during Chord's execution.
 
-The above command causes Chord to load the \verb+chord.properties+ file if present.
+The above command causes Chord to load the \code{chord.properties} file if present.
 The location of this file may be specified explicitly via property
-\verb+chord.props.file+ on the above command line.
-By default, it is \verb+${chord.work.dir}/chord.properties+.
+\code{chord.props.file} on the above command line.
+By default, it is \code{[chord.work.dir]/chord.properties}.
 
-In the above case, the above command does not do much beyond loading the sample \verb+chord.properties+ file.
+In the above case, Chord does not do much beyond loading the sample \code{chord.properties} file.
 For Chord to do something interesting, you need to set additional properties, either
-on the above command line or in the \verb+chord.properties+ file, specifying the task(s)
+on the above command line or in the \code{chord.properties} file, specifying the task(s)
 Chord must perform.
-All tasks are summarized in the ``Chord task properties" portion of the table in Section \ref{sec:chord-sysprops}.
+%All tasks are summarized in the ``Chord task properties" portion of the table in Section \ref{sec:chord-sysprops}.
 The two most common tasks, described next, are building program scope (Section \ref{sec:building-scope})

### 2010-04-02 main/doc/user_guide.tex

--- 2009-07-01 doc/user_guide.tex
@@
 \documentclass{article}
-\usepackage{hyperlatex}
+\usepackage{html}
+\usepackage{hyperref}
 
-%\topmargin=-0.5in
-\texonly{\oddsidemargin=0in}
-\texonly{\evensidemargin=0in}
-\texonly{\textwidth=5.6in}
-%\textheight=8.5in
+\oddsidemargin=0in
+\evensidemargin=0in
+\textwidth=6.0in
+\textheight=8.0in
 
-%TODO: specify use of BDD's under efficiency in section 1 (overview)
+\html{\def\code#1{{\tt {#1}}}}
+\latex{\def\code#1{\path{#1}}}
+\def\url#1{\htmladdnormallink{#1}{#1}}
+\def\Mdom{{\bf M}}
+\def\Idom{{\bf I}}
+\def\Vdom{{\bf V}}
+\def\Zdom{{\bf Z}}
+\def\Hdom{{\bf H}}
+\def\Edom{{\bf E}}
+\def\Fdom{{\bf F}}
+\def\Cdom{{\bf C}}
+\def\Tdom{{\bf T}}
+\def\Pdom{{\bf P}}
 
-\htmlmathitalic{0}
-\htmltitle{User Guide}
-\htmladdress{chord@cs.stanford.edu, \today}
+\def\instM{{\bf instM}}
+\def\statM{{\bf statM}}
+\def\MH{{\bf MH}}
+\def\MI{{\bf MI}}
+\def\ME{{\bf ME}}
 
-\providecommand\Mdom{{\bf M}}
-\providecommand\Idom{{\bf I}}
-\providecommand\Vdom{{\bf V}}
-\providecommand\Zdom{{\bf Z}}
-\providecommand\Hdom{{\bf H}}
-\providecommand\Edom{{\bf E}}
-\providecommand\Fdom{{\bf F}}
-\providecommand\Cdom{{\bf C}}
-\providecommand\Tdom{{\bf T}}
-\providecommand\Pdom{{\bf P}}
-\providecommand\Sdom{{\bf S}}
+\def\cha{{\bf cha}}
+\def\sub{{\bf sub}}
+\def\HT{{\bf HT}}
+\def\VT{{\bf VT}}
+\def\EF{{\bf EF}}
+\def\writeE{{\bf writeE}}
 
-\providecommand\instM{{\bf instM}}
-\providecommand\statM{{\bf statM}}
-\providecommand\MH{{\bf MH}}
-\providecommand\MI{{\bf MI}}
-\providecommand\ME{{\bf ME}}
+\def\virtIM{{\bf virtIM}}
+\def\statIM{{\bf statIM}}
+\def\specIM{{\bf specIM}}
 
-\providecommand\cha{{\bf cha}}
-\providecommand\sub{{\bf sub}}
-\providecommand\HT{{\bf HT}}
-\providecommand\VT{{\bf VT}}
-\providecommand\EF{{\bf EF}}
-\providecommand\writeE{{\bf writeE}}
+\def\MmethArg{{\bf MmethArg}}
+\def\MmethRet{{\bf MmethRet}}
+\def\IinvkArg{{\bf IinvkArg}}
+\def\IinvkRet{{\bf IinvkRet}}
+\def\MordMethArg{{\bf MordMethArg}}
+\def\MspcMethArg{{\bf MspcMethArg}}
 
-\providecommand\virtIM{{\bf virtIM}}
-\providecommand\statIM{{\bf statIM}}
-\providecommand\specIM{{\bf specIM}}
+\def\MobjVarAsgnInst{{\bf MobjVarAsgnInst}}
+\def\MclsVarAsgnInst{{\bf MclsVarAsgnInst}}
+\def\MobjValAsgnInst{{\bf MobjValAsgnInst}}
+\def\MgetInstFldInst{{\bf MgetInstFldInst}}
+\def\MputInstFldInst{{\bf MputInstFldInst}}
+\def\MgetStatFldInst{{\bf MgetStatFldInst}}
+\def\MputStatFldInst{{\bf MputStatFldInst}}
 
-\providecommand\MmethArg{{\bf MmethArg}}
-\providecommand\MmethRet{{\bf MmethRet}}
-\providecommand\IinvkArg{{\bf IinvkArg}}
-\providecommand\IinvkRet{{\bf IinvkRet}}
-\providecommand\MordMethArg{{\bf MordMethArg}}
-\providecommand\MspcMethArg{{\bf MspcMethArg}}
+\def\syncMV{{\bf syncMV}}
+\def\nestedLocks{{\bf nestedLocks}}
 
-\providecommand\MobjVarAsgnInst{{\bf MobjVarAsgnInst}}
-\providecommand\MclsVarAsgnInst{{\bf MclsVarAsgnInst}}
-\providecommand\MobjValAsgnInst{{\bf MobjValAsgnInst}}
-\providecommand\MgetInstFldInst{{\bf MgetInstFldInst}}
-\providecommand\MputInstFldInst{{\bf MputInstFldInst}}
-\providecommand\MgetStatFldInst{{\bf MgetStatFldInst}}
-\providecommand\MputStatFldInst{{\bf MputStatFldInst}}
+\def\VH{{\bf VH}}
+\def\FH{{\bf FH}}
+\def\HFH{{\bf HFH}}
+\def\IM{{\bf IM}}
 
-\providecommand\syncMV{{\bf syncMV}}
-\providecommand\nestedLocks{{\bf nestedLocks}}
+\def\forkC{{\bf forkC}}
 
-\providecommand\VH{{\bf VH}}
-\providecommand\FH{{\bf FH}}
-\providecommand\HFH{{\bf HFH}}
-\providecommand\IM{{\bf IM}}
+\def\CVC{{\bf CVC}}
+\def\FC{{\bf FC}}
+\def\CFC{{\bf CFC}}
+\def\CICM{{\bf CICM}}
+\def\CHC{{\bf CHC}}
 
-\providecommand\forkC{{\bf forkC}}
+\def\clinitM{{\bf clinitM}}
 
-\providecommand\CVC{{\bf CVC}}
-\providecommand\FC{{\bf FC}}
-\providecommand\CFC{{\bf CFC}}
-\providecommand\CICM{{\bf CICM}}
-\providecommand\CHC{{\bf CHC}}
+\def\bddbddb{{\tt bddbddb}}
+\def\Chord{Chord}
 
-\providecommand\clinitM{{\bf clinitM}}
+\def\NameToTrgtMap{{\tt NameToTrgtMap}}
+\def\NameToTaskMap{{\tt NameToTaskMap}}
+\def\ConsumedNameToTasksMap{{\tt ConsumedNameToTasksMap}}
+\def\TaskGraph{{\tt TaskGraph}}
+\def\TaskMapEntries{{\tt TaskMapEntries}}
+\def\TrgtMapEntries{{\tt TrgtMapEntries}}
+\def\DlogMapEntries{{\tt DlogMapEntries}}
+\def\Tasks{{\tt Tasks}}
+\def\Trgts{{\tt Trgts}}
+\def\ValidTasks{{\tt ValidTasks}}
+\def\TrgtNames{{\tt TrgtNames}}
 
-\providecommand\Chord{{Chord}}
-
-\providecommand\NameToTrgtMap{{\tt NameToTrgtMap}}
-\providecommand\NameToTaskMap{{\tt NameToTaskMap}}
-\providecommand\ConsumedNameToTasksMap{{\tt ConsumedNameToTasksMap}}
-\providecommand\TaskGraph{{\tt TaskGraph}}
-\providecommand\TaskMapEntries{{\tt TaskMapEntries}}
-\providecommand\TrgtMapEntries{{\tt TrgtMapEntries}}
-\providecommand\DlogMapEntries{{\tt DlogMapEntries}}
-\providecommand\Tasks{{\tt Tasks}}
-\providecommand\Trgts{{\tt Trgts}}
-\providecommand\ValidTasks{{\tt ValidTasks}}
-\providecommand\TrgtNames{{\tt TrgtNames}}
-
-\title{Chord: A Static Analysis Framework for Java}
-\author{Mayur Naik and Alex Aiken}
+\title{Chord: A Static and Dynamic Program Analysis Framework for Java}
+\author{Mayur Naik}
 \date{\today}
@@
 \maketitle
-\texonly{\tableofcontents}
 
-\texonly{\newpage}
-\section{Overview}
-
-\Chord\ is a static analysis framework for Java with an emphasis on the analysis of objects and concurrency.
-It has the following characteristics:
-\begin{enumerate}
-\item
-It is {\it stand-alone} in that it provides various off-the-shelf static analyses (e.g., may alias analysis, thread escape analysis, datarace analysis, deadlock analysis, etc.).
-\item
-It is {\it extensible} in that it allows users to build their own static analyses on top of the provided static analyses.
-\item
-It is {\it compositional} in that it enables different static analyses to coexist and interact with each other.
-\item
-It is {\it efficient} in that it
-caches results computed by each static analysis so that they can be reused by other static analyses without re-computation.
-\item
-It is {\it flexible} in that it allows each static analysis to be expressed either declaratively as a Datalog program or imperatively as a Java program.
-\end{enumerate}
-
-\texonly{\newpage}
-\section{Download and Installation}
-
-Before installing \Chord\ ensure that the following software is installed on your machine:
-
-\begin{itemize}
-\item
-JDK 5 or higher (e.g., from
-Sun\footnote{\xlink{{\tt http://java.sun.com/javase/downloads/index.jsp}}{http://java.sun.com/javase/downloads/index.jsp}}
-or IBM\footnote{\xlink{{\tt http://www.ibm.com/developerworks/java/jdk/}}{http://www.ibm.com/developerworks/java/jdk/}}).
-
-\item
-Apache Ant (a Java build tool)\footnote{\xlink{{\tt http://ant.apache.org/}}{http://ant.apache.org/}}
-\end{itemize}
-
-Download the latest version of \Chord\ from \xlink{{\tt http://code.google.com/p/jchord/}}{http://code.google.com/p/jchord/}
-(say {\tt chord-x.y.tar.gz}) and uncompress it in a suitable directory by running the following command:
-
-\begin{verbatim}
-tar xvzf chord-x.y.tar.gz
-\end{verbatim}
-
-This will create a directory structure with root directory {\tt chord-x.y/}.
-
-\texonly{\newpage}
-\section{Building Chord}
-
-The root directory {\tt chord-x.y/} contains a {\tt build.xml} file which is interpreted by
-Apache Ant (a Java build tool).
-To see the various targets available run the following command in the root directory:
-
-\begin{verbatim}
-ant -p
-\end{verbatim}
-
-The default target {\tt compile} compiles \Chord.  So simply run the following command in the
-root directory:
-
-\begin{verbatim}
-ant
-\end{verbatim}
-
-This will compile the \Chord\ {\tt .java} source files and store the {\tt .class} files
-under the {\tt classes/} sub-directory of the root directory.
-
-\texonly{\newpage}
-\section{Running Chord}
-
-Suppose the program to be analyzed has the following directory structure:
-
-\begin{verbatim}
-example/
-    src/
-        foo/
-            Main.java
-            ...
-    classes/
-        foo/
-            Main.class
-            ...
-    lib/
-        src/
-            taz/
-                ...
-        jars/
-            taz.jar
-    chord/
-        chord.properties
-\end{verbatim}
-
-The above structure is typical: the program's Java source files are under {\tt example/src/}, its class files
-are under {\tt example/classes/}, the source and jar files of the libraries used by the program are under
-{\tt example/lib/src/} and {\tt example/lib/jars/}, respectively, and finally {\tt example/chord/} is the directory
-to which Chord will output its intermediate and final results.
-
-Execute Chord on the above program by running the
-following command in Chord's root directory {\tt chord-x.y/}:
-
-\begin{verbatim}
-ant -Dchord.work.dir=.../example/chord run
-\end{verbatim}
-
-where ``{\tt ...}'' denotes either the absolute path, or the path relative to Chord's root directory,
-of the parent of directory {\tt example/}.
-Chord will read the system properties specified in the file named {\tt chord.properties}
-that it presumes exists in the directory specified
-by system property {\tt chord.work.dir}
-which in the above case is directory {\tt example/chord/}.
-A sample such file for the above program is as follows:
-
-\begin{verbatim}
-chord.main.class=foo.Main
-chord.class.path=../classes:../lib/classes/taz.jar
-chord.src.path=../src:../lib/src
-chord.out=log.txt
-chord.err=log.txt
-chord.serial.file=program.ser
-chord.java.analysis.path=${chord.home.dir}/classes/main:
-    ${chord.home.dir}/classes/downcast:${chord.home.dir}/src/dlog/monosite
-chord.dlog.analysis.path=${chord.home.dir}/src/dlog/main:
-    ${chord.home.dir}/src/dlog/downcast:${chord.home.dir}/src/dlog/monosite
-chord.analyses=downcast-java,monosite-java
-\end{verbatim}
-
-System property {\tt chord.home.dir} specifies Chord's root directory
-{\tt chord-x.y/}.
-Every non-absolute location used in the value of any system property defined in the file
-is taken to be relative to the directory specified by system
-property {\tt chord.work.dir}.
-
-Section \ref{sec-sys-prop} presents an exhaustive
-listing of system properties that are recognized by \Chord.
-Briefly:
-\begin{itemize}
-\item
-{\tt chord.main.class} specifies the main class of the program
-to be analyzed.
-\item
-{\tt chord.class.path} specifies the application classpath of the program to be analyzed (as opposed to the
-JDK library classpath which consists of jar files such as {\tt rt.jar}
-which are implicitly appended to the program's application classpath to yield the
-complete program classpath).
-\item
-{\tt chord.out} and {\tt chord.err} specify the files (typically one and the same) to which the standard output and
-standard error streams, respectively, must be redirected while running Chord.
-\item
-{\tt chord.serial.file} specifies the location of a file from (resp. to) which
-the serialized form of the intermediate representation of the program to be analyzed must be loaded (resp. stored).
-This option is provided because loading the serialized form is much faster than
-constructing the intermediate representation of the program each time.
-\item
-{\tt chord.java.analysis.path} and {\tt chord.dlog.analysis.path} 
-specify all program domains, program relations, and program analyses
-to be included in the project and the
-dependencies between them (see Section \ref{sec-tg}).
-In the above case, they specify certain
- program domains, program relations, and program analyses
-pre-defined in \Chord, in particular, all those in (1) Chord's standard library, (2) Chord's downcast safety
-analysis, and (3) Chord's monomorphic call site analysis.
-Users can also define their own
-program domains, program relations, and program analyses and include them in the project.
-\item
-{\tt chord.analyses} is a space- or comma-separated list of the names of program domains,
-program relations, and program analyses that must be run in that order.  In the above case,
-these are downcast safety analysis, denoted {\tt downcast-java}, and monomorphic call site analysis,
-denoted {\tt monosite-java}.
-Both are program analyses pre-defined in \Chord\ 
-but users can also include the names
-of program domains, program relations, and program analyses they have defined and wish to run.
-\end{itemize}
-
-\texonly{\newpage}
-\section{Specifying Program Scope}
-
-\Chord\ takes as input a complete (i.e., closed) program with a main method.
-The scope of the program must be specified in one of the
-following ways:
-
-\begin{itemize}
-\item
-Defining system property {\tt chord.serial.file} as
-the location of a file containing in serialized form the intermediate representation
-of the program generated by a previous run of \Chord.
-\item
-Defining system property {\tt chord.main.class} as the fully-qualified name
-of the program's main class (e.g., {\tt foo.Main}) and system property
-{\tt chord.class.path} as the program's application classpath.
-\end{itemize}
-
-The former option is provided because loading the serialized form
-is much faster than constructing the intermediate representation
-of the program each time.
-
-\Chord\ first checks whether system property {\tt chord.serial.file} is defined.
-If it is defined, then it further checks whether the file specified by it
-exists.
-If the file exists, then \Chord\ treats the file as a serialized form of
-the intermediate representation of the program to be analyzed, and loads it.
-
-If system property {\tt chord.serial.file} is not defined or if the file
-specified by the property does not exist, then \Chord\ requires system properties
-{\tt chord.main.class} and {\tt chord.class.path} 
-to be defined, and uses them to
-build an intermediate representation of the program (see Section \ref{sec-prog-build})
-and perform source-to-source transformations on that representation
-(see Section \ref{sec-prog-transform}) that simplify several program analyses.
-Finally, if system property {\tt chord.serial.file} was defined but
-specified a non-existent file, then \Chord\ creates the file and stores the
-transformed intermediate representation of the program into the file in
-serialized form so that it can be loaded in subsequent runs.
-
-Defining system property {\tt chord.src.path} is optional.
-It specifies the program's Java source path which provides the
-locations of the program's {\tt .java} files so that, for instance, the
-results of an analysis (which operates on the intermediate representation of
-the program extracted from {\tt .class} files) can be mapped to the
-program's {\tt .java} files and presented to the user for
-manual inspection.
-
-\texonly{\newpage}
-\section{Program Representation}
-\label{sec-prog-build}
-
-Given the program's program's main class via system property {\tt chord.main.class} and the
-program's application classpath via system property {\tt chord.class.path},
-\Chord\ uses system property {\tt chord.program.builder.class} to locate the
-builder to be used for building an intermediate representation of the program,
-namely, an object of class {\tt chord.program.Program}.
-By default, the value of this property is class {\tt chord.program.builders.SootProgramBuilder}
-which uses the Soot Java compiler framework \cite{Soot}.
-Any builder, however, must implement interface
-{\tt chord.program.builders.IProgramBuilder}.
-
-The intermediate representation
-elides values of primitive type (e.g., int, boolean, etc.) and operations
-on them (e.g., arithmetic and relational operations) in the original program.
-The current emphasis of \Chord\ is on the analysis of objects
-but the analysis of primitives may be supported in future
-releases.
-
-The intermediate representation represents the program as a set of types.
-Each type has a set of (instance or static) fields and a set of
-(instance or static) methods.
-Each method has an optional control-flow graph.
-The control-flow graph has a unique entry point and
-a unique exit point\footnote{Even the control-flow graph of a method with an infinite loop
-has an exit point because loops are transformed into tail-recursive methods
-(see Section \ref{sec-prog-transform}).}.
-The entry and exit points are always distinct.
-Each point in the control-flow graph has one or more immediate predecessors
-(except the entry point which has none) and, likewise,
-one or more immediate successors (except the exit point which has none).
-Associated with each point is a simple statement, which may be of one of
-the following kinds:
-
-\begin{itemize}
-\item
-{\tt HeadInst}
-
-The statement associated with the unique entry point of a control-flow graph.
-
-It serves as a placeholder for:
-
-\begin{itemize}
-\item
-the (possibly empty) list of the method's {\it argument variables} of reference type;
-in the case of an instance method, this list is non-empty,
-with the $0^{th}$ element being the distinguished implicit {\tt this}
-argument variable.
-\item
-the (possibly empty) list of the method's {\it temporary variables} of reference type.
-\end{itemize}
-
-Argument variables and temporary variables of primitive type are
-elided due to \Chord's emphasis on the analysis of objects.
-
-Argument variables and temporary variables are distinct, namely,
-the set of a method's {\it local variables} (of reference type)
-is the disjoint union of the sets of the method's argument variables (of reference type) and
-temporary variables (of reference type).
-
-\item
-{\tt TailInst}
-
-The statement associated with the unique exit point of a control-flow graph.
-
-It serves as a placeholder for 
-the (possibly empty) list of the method's {\it return variables} of reference type.
-
-Return variables of primitive type are
-elided due to \Chord's emphasis on the analysis of objects.
-
-The set of a method's return variables (of reference type) is a subset of the set
-of the method's local variables (of reference type).
-
-A Java method has at most one return variable but
-\Chord's intermediate representation allows multiple return variables
-to enable compiling away exceptions, loops, synchronized blocks, etc.
-
-\item
-{\tt AcqLockInst}
-
-A statement of the form {\tt monitorenter v} where
-{\tt v} is a local variable of reference type.
-
-\item
-{\tt RelLockInst}
-
-A statement of the form {\tt monitorexit v} where {\tt v}
-is a local variable of reference type.
-
-\item
-{\tt StrValAsgnInst}
-
-A statement of the form {\tt v = "..."} where {\tt v} is
-a local variable of any reference type that can store a string
-constant.  It is typically (but not necessarily) of type
-{\tt java.lang.String}.
-
-\item
-{\tt NilValAsgnInst}
-
-A statement of the form {\tt v = null} where {\tt v}
-is a local variable of reference type.
-
-\item
-{\tt StatFldRefInst}
-
-A statement of one of the following forms:
-
-{\tt v = f} \\
-{\tt f = v} \\
-{\tt * = f} \\
-{\tt f = *}
-
-where {\tt v} is a local variable of reference type and
-{\tt f} is a static field.
-
-The {\tt *} denotes an expression of primitive type in the
-original program which is elided due to \Chord's emphasis on the
-analysis of objects.
-
-\item
-{\tt InstFldRefInst}
-
-A statement of one of the following forms:
-
-{\tt v = b.f} \\
-{\tt b.f = v} \\
-{\tt * = b.f} \\
-{\tt b.f = *}
-
-where {\tt v} and {\tt b} are local variables of reference
-type and {\tt f} is an instance field.
-
-The {\tt *} denotes an expression of primitive type in the
-original program which is elided due to \Chord's emphasis on the
-analysis of objects.
-
-\item
-{\tt SkipInst}
-
-A no-op statement.
-
-Due to \Chord's emphasis on the analysis of objects, each statement
-in the original program that does not access
-any value of reference type (i.e., only accesses values of primitive
-type) is represented by this statement.
-
-\item
-{\tt InvkInst}
-
-A method invocation statement of the form
-{\tt r1,...,rn = m(a1,...,ak)} where
-{\tt r1}, ..., {\tt rn} (return variables) and
-{\tt a1}, ..., {\tt ak} (argument variables)
-are local variables of reference type.
-
-Argument variables and return variables of primitive type are
-elided due to \Chord's emphasis on the analysis of objects.
-
-A method invocation statement in Java has at most
-one return variable but
-\Chord's intermediate representation allows multiple return variables
-to enable compiling away exceptions, loops, synchronized blocks, etc.
-
-\item
-{\tt ObjVarAsgnInst}
-
-A statement of the form {\tt v1 = v2} where {\tt v1}
-and {\tt v2} are local variables of (possibly different) reference types.
-
-\item
-{\tt ObjValAsgnInst}
-
-A statement of the form {\tt v = new t} where {\tt t} is a
-concrete reference type and
-{\tt v} is a local variable of a (possibly different) reference type.
-
-\item
-{\tt ClsVarAsgnInst}
-
-A statement of the form {\tt v = b.getClass()} where
-{\tt v} is a local variable of type {\tt java.lang.Class}
-and {\tt b} is a local variable of reference type.
-
-The reason this kind of statement is modeled is because
-\Chord\ has an emphasis on concurrency and many Java programs
-hold locks on the runtime class of an object
-(which is obtained by calling the native {\tt getClass}
-instance method of the object).
-
-\item
-{\tt AryElemRefInst}
-
-A statement of one of the following forms:
-
-{\tt v = b[*]}  \\
-{\tt b[*] = v} \\
-{\tt * = b[*]} \\
-{\tt b[*] = *}
-
-where {\tt v} and {\tt b} are local variables of reference
-type.
-
-The {\tt *} denotes an expression of primitive type in the
-original program which is elided due to \Chord's emphasis on the
-analysis of objects.
-\end{itemize}
-
-\texonly{\newpage}
-\section{Program Transformations}
-\label{sec-prog-transform}
-
-After building the intermediate representation of the
-program, \Chord\ performs
-the following source-to-source transformations 
-that simplify several program analyses:
-
-\begin{enumerate}
-\item
-{\bf Sync Removal}
-
-Removes all synchronized blocks from each method M in each class
-C in the program by generating a fresh static method (in class C)
-per synchronized block in the body of M.
-
-\item
-{\bf Loop Removal}
-
-Removes all loops from each method M in each class C in the program
-by generating a fresh static method (in class C)
-per loop in the body of M.
-
-\item
-{\bf Skip Removal}
-
-Removes all skip (i.e., no-op) statements from each method in each class
-in the program.
-Such statements arise primarily because \Chord, due to its emphasis
-on the analysis of objects, replaces each
-statement in the original program that does not access any
-value of reference type (i.e., only accesses values of primitive
-type) by a skip statement.
-
-\end{enumerate}
-
-\texonly{\newpage}
-\section{Task Graph Construction}
-\label{sec-tg}
-
-\Chord\ requires the following three system properties to construct the task graph:
-
-\begin{enumerate}
-\item
-{\tt chord.dlogmap.path}
-
-The path specifying the mapping from the name of each Datalog-based program
-analysis that must be included in the project to the location
-of the file defining that program analysis.
-
-Each path element must be the location of a file containing a
-list of entries (one per line) of the form:
-
-{\tt <name> <file-name>}
-
-The path elements must be separated by the system-specific path separator
-character (e.g., `:' on UNIX and `;' on Windows).
-
-Let \DlogMapEntries\ denote the set of all entries of the
-above form.
-
-\item
-{\tt chord.trgtmap.path}
-
-The path partially specifying the mapping from the name of each program domain,
-program relation, and Java-based program analysis that must be
-included in the project to the fully-qualified name of a subclass, either
-pre-defined in \Chord\ or defined by the user, of {\tt chord.Dom},
-{\tt chord.Ary*Rel}, and
-{\tt chord.Entity}, respectively, the instance fields of which define the
-state of the corresponding domain, relation, or analysis.
-The `{\tt *}' in {\tt chord.Ary*Rel} is 1, 2, ..., 6, or N for
-program relations of arity 1, 2, ..., 6, or $>$ 6, respectively.
-
-Each path element must be the location of a file containing a
-list of entries (one per line) of the form:
-
-{\tt <name> <class-name>}
-
-The path elements must be separated by the system-specific path separator
-character (e.g., `:` on UNIX and ';' on Windows).
-
-Let \TaskMapEntries\ denote the set of all entries of the
-above form.
-
-\item
-{\tt chord.taskmap.path}
-
-The path partially specifying the mapping from the name of each program domain,
-program relation, and Java-based program analysis that must be included in
-the project to the fully-qualified name of a
-subclass, either pre-defined in \Chord\ or defined by the user, of
-{\tt chord.Dom}, {\tt chord.Ary*Rel}, and {\tt chord.Entity}, respectively,
-which define the following instance methods:
-\begin{itemize}
-\item
-the {\tt execute} method specifying
-the computation that updates the state of the corresponding domain, relation,
-or analysis,
-\item
-the {\tt invalidate} method specifying the action to be taken when the
-state of the corresponding domain, relation, or analysis is invalidated.
-\item
-the {\tt getConsumedNames} method specifying names of any domains, relations,
-and Datalog-based or Java-based program analyses upon whose state the
-computation depends, and
-\item
-the {\tt getProducedNames} method specifying names of any domains, relations,
-and Datalog-based or Java-based program analyses whose state depends upon
-the computation.
-\end{itemize}
-
-Each path element must be the location of a file containing a
-list of entries (one per line) of the form:
-
-{\tt <name> <class-name>}
-
-The path elements must be separated by the system-specific path separator
-character (e.g., `:` on UNIX and ';' on Windows).
-
-Let \TrgtMapEntries\ denote the set of all entries of the above form.
-\end{enumerate}
-
-Then, \Chord\ uses the values of \DlogMapEntries, \TaskMapEntries, and \TrgtMapEntries\ 
-to do the following:
-
-\begin{enumerate}
-\item
-Compute {\tt Tasks\ :\ Set<chord.Entity>}, the set of all tasks,
-each of which is an object of (a subclass of) {\tt chord.Entity},
-as follows:
-
-\begin{enumerate}
-\item
-For each entry in \DlogMapEntries\ (i.e., a Datalog-based program analysis,
-there is a unique task, namely, a unique object of class {\tt chord.DatalogAnalysis}.
-
-\item
-For each entry in \TaskMapEntries\ (i.e., of the form {\tt <name> <class-name>} specifying
-a program domain, a program relation, or a Java-based program analysis), there is a
-unique task, namely, a unique object
-of the class named {\tt <class-name>} which must be (a subclass of) {\tt chord.Dom} in the
-case of a program domain, (a subclass of) {\tt chord.Ary*Rel} in the case of a
-program relation, and (a subclass of) {\tt chord.Entity} in the case of a Java-based
-program analysis.
-\end{enumerate}
-
-\item
-Compute the following maps as shown below:
-
-{\tt ConsumedNameToTasksMap\ :\ String -> Set<Task>},
-mapping each name to the set of (possibly empty) tasks that consume that name.
-
-{\tt NameToTaskMap\ :\ String -> Task},
-mapping each name to the unique task that produces that name.
-
-\Chord\ reports an error if a name is produced by no task or by more than one task.
-{\small 
-\begin{verbatim}
-for each (<task> in Tasks) {
-    case <task> came from item 1(a) above: {
-        // <task> corresponds to a Datalog-based program analysis
-        for each (<name> declared as input relation) {
-            add <task> to ConsumedNameToTasksMap(<name>)
-        }
-        for each (<name> declared as output relation) {
-            if (<name> is in domain of map NameToTaskMap)
-                error ``multiple tasks produce name <name>''
-            else
-                NameToTaskMap(<name>) = <task>
-        }
-    }
-    case <task> came from item 1(b) above: {
-        // <task> corresponds to a program domain, a program relation, or
-        // a Java-based program analysis
-        // Suppose <task> came from entry <name> <class-name> in
-        // TaskMapEntries
-        // i.e. <task> is an object of class named <class-name>
-        for each (<name2> in <task>.getConsumedNames())
-            add <task> to ConsumedNameToTaskMaps(<name2>)
-        for each (<name2> in <task>.getProducedNames() union { <name> })
-            if (<name2> is in domain of map NameToTaskMap)
-                error ``multiple tasks produce name <name>''
-            else
-                NameToTaskMap(<name2>) = <task>
-    }
-}
-\end{verbatim}
-}
-
-{\small 
-\begin{verbatim}
-for each (<name> in domain of map ConsumedNameToTasksMap) {
-    if (<name> is not in domain of map NameToTaskMap)
-        error ``no task produces name <name>''
-}
-\end{verbatim}
-}
-
-\item
-Compute \NameToTrgtMap : {\tt String -> Entity},
-mapping each name to a unique target, as below.
-\Chord\ reports an error if a name is mapped to multiple targets
-or to no target.
-
-{\small 
-\begin{verbatim}
-for each (entry <name> <class-name> in TaskMapEntries) {
-    if (<name> is in domain of map NameToTrgtMap)
-        error ``name <name> mapped to multiple targets''
-    NameToTrgtMap(<name>) = NameToTaskMap(<name>)
-}
-for each (entry <name> <class-name> in TrgtMapEntries) {
-    if (<name> is in domain of map NameToTrgtMap)
-        error ``name <name> mapped to multiple targets''
-    NameToTrgtMap(<name>) = a fresh object of class named <class-name>
-}
-\end{verbatim}
-}
-
-To reduce user burden in providing \TaskMapEntries\ and
-\TrgtMapEntries, \Chord\ implicitly extends \NameToTrgtMap\ using
-\DlogMapEntries\ as follows:
-
-{\small 
-\begin{verbatim}
-for each (<name> in domain of map NameToTaskMap) {
-    if (<name> is not in domain of map NameToTrgtMap) {
-        if (<name> is declared as a domain in some datalog file in DlogMapEntries)
-            NameToTrgtMap(<name>) = a fresh object of class chord.Dom
-        else if (<name> is declared as a (input or output) relation
-                in some datalog file in DlogMapEntries)
-            NameToTrgtMap(<name>) = a fresh object of class chord.Rel
-        else
-            error ``<name> not mapped to any target''
-    } else {
-        chord.Entity <trgt> = NameToTrgtMap(<name>)
-        if (<name> is declared as a domain in some datalog file in DlogMapEntries
-                and <trgt> is not of type (subclass of) chord.Dom)
-            error ``<name> has conflicting types''
-        else if (<name> is declared as a (input or output) relation
-                in some datalog file in DlogMapEntries
-                and <trgt> is not of type (subclass of) chord.Rel)
-            error ``<name> has conflicting types''
-    }
-}
-\end{verbatim}
-}
-
-\item
-Compute graph \TaskGraph\ as a graph whose:
-
-\begin{itemize}
-\item
-set of nodes is \Tasks\ and
-\item
-set of edges containing an edge from task {\tt <task1>} to task {\tt <task2>} 
-whenever there exists a name {\tt <name>} such that
-\NameToTaskMap({\tt <name>}) = {\tt <task1>} and
-\ConsumedNameToTasksMap({\tt <name>}) contains {\tt <task2>}.
-\end{itemize}
-
-The task graph in XML and HTML formats is dumped to files {\tt taskgraph.xml}
-and {\tt taskgraph.html} in the project's working directory specified using
-system property {\tt chord.work.dir}.
-
-\end{enumerate}
-
-\texonly{\newpage}
-\section{Project Organization}
-
-\Chord\ interprets system property {\tt chord.trgts}, if defined, as
-a space-separated list of names of program domains, program relations, and
-Datalog-based or Java-based program analyses, denoted \TrgtNames,
-to be computed in that order.
-
-To avoid redundant computation, \Chord\ maintains during its execution a set of
-up-to-date tasks, denoted \ValidTasks,
-which is a subset of the set of all tasks, denoted \Tasks.
-Set \ValidTasks\ is empty at startup.
-A future release of \Chord\ may provide the ability to retrieve a set of up-to-date tasks from
-a previous run of \Chord\ to prevent re-computing those tasks in subsequent runs.
-The following procedure is executed upon startup:
-
-{\small
-\begin{verbatim}
-for each (<name> in TrgtNames in order) {
-    if (<name> is not in domain of map NameToTaskMap) {
-        error ``name <name> is an invalid target name''
-    else {
-        chord.Entity <task> = NameToTaskMap(<name>)
-        executeTask(<task>)
-    }
-}
-\end{verbatim}
-}
-
-\noindent where procedure {\tt executeTask} is defined as follows:
-
-{\small
-\begin{verbatim}
-procedure executeTask(chord.Entity <task>) {
-    if (<task> is not in ValidTasks) {
-        for each (immed. pred. <task2> of <task> in TaskGraph)
-            executeTask(<task2>)
-        <task>.execute()
-        for each (immed. succ. <task2> of <task> in TaskGraph)
-            invalidateTask(<task2>)
-        add <task> to ValidTasks
-    }
-}
-\end{verbatim}
-}
-
-\noindent where procedure {\tt invalidateTask} is defined as follows:
-
-{\small
-\begin{verbatim}
-procedure invalidateTask(chord.Entity t) {
-    if (<task> is in ValidTasks) {
-        remove <task> from ValidTasks
-        <task>.invalidate()
-        for each (immed. succ. <task2> of <task> in TaskGraph)
-            invalidateTask(<task2>)
-    }
-}
-\end{verbatim}
-}
-
-\texonly{\newpage}
-\section{Chord System Properties}
-\label{sec-sys-prop}
-
-\Chord\ recognizes the following system properties:
-
-\begin{itemize}
-\item
-{\tt chord.out}
-
-The location of a file to which the standard output stream
-of \Chord's execution is redirected.
-
-It may be the same file to which the standard error stream is redirected (see
-the description of system property {\tt chord.err}).
-
-The file is created if it does not exist and its contents are overwritten if it exists.
-
-\item
-{\tt chord.err}
-
-The location of a file to which the standard error stream
-of \Chord's execution is redirected.
-
-It may be the same file to which the standard output stream is redirected (see
-the description of system property {\tt chord.out}).
-
-The file is created if it does not exist and its contents are overwritten if it exists.
-
-\item
-{\tt chord.program.builder.class}
-
-The fully-qualified name of the class to be used for building the
-intermediate representation of the program to be analyzed.
-It must be a concrete class with a nullary constructor that
-implements interface {\tt chord.program.builders.IProgramBuilder}.
-
-The default value of this property is {\tt chord.program.builders.SootProgramBuilder}
-which uses the Soot Java compiler framework \cite{Soot}.
-
-\item
-{\tt chord.serial.file}
-
-The location of a file from (resp.\ to) which
-the serialized form of the intermediate representation of the
-program to be analyzed is loaded (resp.\ stored).
-
-\item
-{\tt chord.main.class}
-
-The fully-qualified name of the main class (e.g., {\tt foo.Main}) of the
-program to be analyzed.
-
-This system property (and {\tt chord.class.path})
-must be defined if system property {\tt chord.serial.file}
-does not specify the file containing the serialized form of the intermediate representation
-of the program to be analyzed.
-
-\item
-{\tt chord.class.path}
-
-The application classpath of the program to be analyzed (as opposed to the
-JDK library classpath which consists of jar files such as {\tt rt.jar}
-which are implicitly appended to the program's application classpath to
-yield the complete program classpath).
-
-Each path element must be the location of a directory or a jar/zip file.
-
-The path elements must be separated by the system-specific path separator character
-(e.g., `:' on UNIX and `;' on Windows).
-
-This system property (and {\tt chord.main.class}) must
-be defined if system property {\tt chord.serial.file}
-does not specify the file containing the serialized form of the intermediate representation
-of the program to be analyzed.
-
-\item
-{\tt chord.src.path}
-
-The Java source path of the program to be analyzed.
-
-Each path element must be the location of a directory.
-
-The purpose of this path is to provide the location of the
-{\tt .java} files of the program being analyzed so that, for instance, the
-results of an analysis (which operates on an intermediate representation of the program
-extracted from {\tt .class} files) can be mapped to the corresponding {\tt .java}
-files and presented to the user for manual inspection.
-
-Each file having suffix {\tt .java} in a directory (or a sub-directory of a directory)
-specified in this path is regarded as a source file of the program.
-
-\item
-{\tt chord.dlog.analysis.path}
-
-The path specifying the mapping from the name of each Datalog-based program
-analysis that must be included in the project to the location
-of the file defining that program analysis.
-
-Each path element must be the location of a file containing a
-list of entries (one per line) of the form:
-
-{\tt <name> <file-name>}
-
-The path elements must be separated by the system-specific path separator character
-(e.g., `:' on UNIX and `;' on Windows).
-
-\item
-{\tt chord.java.analysis.path}
-
-The path specifying the mapping from the name of each program domain, program relation,
-and Java-based program analysis that must be included in the project to the
-fully-qualified name of a subclass, either pre-defined in \Chord\ or defined by the user, of
-{\tt chord.Dom}, {\tt chord.Ary*Rel}, and {\tt chord.Entity}, respectively, which
-define the following instance methods:
-\begin{itemize}
-\item
-the {\tt execute} method specifying
-the computation that updates the state of the corresponding domain, relation,
-or analysis, 
-\item
-the {\tt invalidate} method specifying the action to be taken when the
-state of the corresponding domain, relation, or analysis is invalidated.
-\item
-the {\tt getConsumedNames} method specifying names of any domains, relations,
-and analyses upon whose state the computation depends, and
-\item
-the {\tt getProducedNames} method specifying names of any domains, relations,
-and analyses whose state depends upon the computation.
-\end{itemize}
-
-Each path element must be the location of a file containing a
-list of entries (one per line) of the form:
-
-{\tt <name> <class-name>}
-
-The path elements must be separated by the system-specific path separator character
-(e.g., `:' on UNIX and `;' on Windows).
-
-\item
-{\tt chord.trgtmap.path}
-
-The path specifying the mapping from the name of each program domain,
-program relation, and Java-based program analysis that must be
-included in the project to the fully-qualified name of a subclass,
-either pre-defined in \Chord\ or defined by the user, of {\tt chord.Dom},
-{\tt chord.Ary*Rel}, and
-{\tt chord.Entity}, respectively, the instance fields of which define the
-state of the corresponding domain, relation, or analysis.
-
-Each path element in the list must be the location of a file containing a
-list of entries (one per line) of the form:
-
-{\tt <name> <class-name>}
-
-The path elements must be separated by the system-specific path separator character
-(e.g., `:` on UNIX and ';' on Windows).
-
-\item
-{\tt chord.analyses}
-
-A space-separated list of the names of program domains, program relations, and
-Java-based or Datalog-based program analyses that must be computed in that order.
-
-\item
-{\tt chord.ext.class.path}
-
-The classpath of user-defined program domains, user-defined program relations,
-and user-defined Java-based program analyses.
-It is implicitly appended to the classpath of pre-defined program domains,
-pre-defined program relations, and pre-defined Java-based program analyses
-to yield the complete runtime classpath of \Chord.
-
-Each path element must be the location of a directory or a jar/zip file.
-
-The path elements must be separated by the system-specific path separator character
-(e.g., `:' on UNIX and `;' on Windows).
-\end{itemize}
-
-\texonly{\newpage}
-\section{Pre-defined Domains}
-\label{sec-predef-domains}
-
-The following domains are pre-defined in \Chord:
-
-\begin{itemize}
-\item
-$\Cdom$
-
-The domain of all k-object-sensitive contexts.
-
-The $0^{th}$ element in this domain denotes a distinguished hypothetical
-context (denoted \texorhtml{$\epsilon$}{{\it epsilon}} in the literature) that is regarded as the
-lone abstract context in which certain static methods are analyzed.
-The main static method with signature {\tt void main(java.lang.String[])}
-has no callers and is therefore analyzed in this context.
-For simplicity, the k-object-sensitive may alias analysis provided in
-\Chord\ also analyzes each class initializer method in this context.
-
-\item
-$\Edom$
-
-The domain of all statements that access (read or write) an instance field,
-a static field, or an array element, namely, all statements
-represented by instances of (subclasses of) {\tt chord.program.insts.HeapInst}.
-
-\item
-$\Fdom$
-
-The domain of  all instance and static fields.
-
-The $0^{th}$ element in this domain denotes a distinguished hypothetical
-field that is regarded as accessed whenever an array element is
-accessed.
-
-\item
-$\Hdom$
-
-The domain of all object allocation statements, namely, all statements
-represented by instances of {\tt chord.program.insts.ObjValAsgnInst}.
-
-\item
-$\Idom$
-
-The domain of all method invocation statements, namely, all statements
-represented by instances of {\tt chord.program.insts.InvkInst}.
-
-\item
-$\Mdom$
-
-The domain of all instance and static methods.
-
-The $0^{th}$ element in this domain is the main method of the program.
-
-The $1^{st}$ element in this domain is the method with signature
-{\tt void start()} of class {\tt java.lang.Thread}, if this
-method is deemed reachable from the main method of the program.
-
-The above two methods are the root methods of the implicitly spawned
-main thread and each explicitly spawned child thread, respectively.
-Due to \Chord's emphasis on concurrency, these methods are
-referenced frequently by various pre-defined program analyses
-expressed in Datalog, and giving them special indices makes it
-convenient to reference them in those analyses.
-
-\item
-$\Pdom$
-
-The domain of all simple statements, namely, all statements
-represented by instances of (subclasses of) {\tt chord.program.insts.Inst}.
-
-The $0^{th}$ element in this domain is the statement at the unique
-entry point of the main method of the program.
-
-The statements of each method in the program are assigned
-contiguous indices in this domain, with the statements at the
-unique entry and exit points of each method being assigned
-the smallest and largest indices, respectively, of all
-indices assigned to statements in that method.
-
-\item
-$\Tdom$
-
-The domain of all types, including the types of classes, interfaces,
-arrays, and primitives.
-
-\item
-$\Vdom$
-
-The domain of all local variables.
-
-Each local variable declared in each block of each method is
-represented by a unique element in this domain, in particular,
-local variables that have the same name but are declared in
-different methods or in different blocks of the same method
-are represented by different elements in this domain.
-
-All local variables of the same method are assigned contiguous
-indices in this domain, in particular, the set of local
-variables of a method is the disjoint union of its argument
-variables and its temporary variables, and the argument
-variables are assigned contiguous indices in order followed by
-the temporary variables.
-
-\item
-$\Zdom$
-
-The domain of all argument and return variable positions.
-Let $N$ be the largest number or arguments or return variables of any
-method.  Then, this domain contains the $N$ elements $0, 1, ..., N-1$.
-\end{itemize}
-
-\texonly{\newpage}
-\section{Pre-defined Relations}
-\label{sec-predef-relations}
-
-The following relations are pre-defined in \Chord:
-
-\begin{itemize}
-
-\item
-$\cha : \Mdom \times \Tdom \times \Mdom$
-
-Relation containing each tuple $(m_1,t,m_2)$ such that $m_2$ is
-the target method of a method invocation statement whose {\it resolved method}
-is $m_1$ and whose receiver has run-time type $t$.
-It is named $\cha$ because it represents the relation
-computed by Class Hierarchy Analysis.
-
-\item
-$\sub : \Tdom \times \Tdom$
-
-Relation containing each tuple $(s,t)$ such that type $s$ is a
-subtype of type $t$.
-
-\item
-$\HT : \Hdom \times \Tdom$
-
-Relation containing each tuple $(h,t)$ such that object allocation
-statement $h$ allocates objects of type $t$.
-
-\item
-$\VT : \Vdom \times \Tdom$
-
-Relation containing each tuple $(v,t)$ such that local variable $v$
-has type $t$.
-
-\item
-$\MH : \Mdom \times \Hdom$
-
-Relation containing each tuple $(m,h)$ such that method $m$ contains
-object allocation statement $h$.
-
-\item
-$\MI : \Mdom \times \Idom$
-
-Relation containing each tuple $(m,i)$ such that method $m$ contains
-method invocation statement $i$.
-
-\item
-$\ME : \Mdom \times \Edom$
-
-Relation containing each tuple $(m,e)$ such that method $m$ contains
-statement $e$ that accesses (reads or writes) an instance field, a
-static field, or an array element.
-
-\item
-$\instM : \Mdom$
-
-Relation containing all instance (as opposed static) methods.
-
-\item
-$\statM : \Mdom$
-
-Relation containing all static (as opposed instance) methods.
-
-\item
-$\clinitM : \Mdom$
-
-Relation containing all class initializer methods, namely,
-methods having signature {\tt void <clinit>()}.
-
-\item
-$\IinvkArg : \Idom \times \Zdom \times \Vdom$
-
-Relation containing each tuple $(i,z,v)$ such that local variable $v$
-is the $z^{th}$ argument variable of method invocation statement $i$.
-
-\item
-$\IinvkRet : \Idom \times \Zdom \times \Vdom$
-
-Relation containing each tuple $(i,z,v)$ such that local variable $v$
-is the $z^{th}$ return variable of method invocation statement $i$.
-
-\item
-$\MmethArg : \Mdom \times \Zdom \times \Vdom$
-
-Relation containing each tuple $(m,z,v)$ such that local variable
-$v$ is the $z^{th}$ argument variable of method $m$.
-
-\item
-$\MmethRet : \Mdom \times \Zdom \times \Vdom$
-
-Relation containing each tuple $(m,z,v)$ such that local variable
-$v$ is the $z^{th}$ return variable of method $m$.
-
-\item
-$\MspcMethArg : \Mdom \times \Vdom$
-
-Relation containing each tuple $(m,v)$ such that local variable $v$ is
-the implicit {\tt this} argument variable of instance method $m$.
-
-\item
-
-$\MordMethArg : \Mdom \times \Zdom \times \Vdom$
-
-Relation containing each tuple $(m,z,v)$ such that local variable $v$
-is the $z^{th}$ argument variable of method $m$ other than the implicit
-{\tt this} argument variable (in the case in which $m$ is an instance
-method).
-
-\item
-$\MobjVarAsgnInst : \Mdom \times \Vdom \times \Vdom$
-
-Relation containing each tuple $(m,l,r)$ such that method $m$
-contains a copy statement of the form $l = r$ where $l$ and $r$ are
-local variables.
-
-\item
-$\MobjValAsgnInst : \Mdom \times \Vdom \times \Vdom$
-
-Relation containing each tuple $(m,l,h)$ such that method $m$ contains
-object allocation statement $h$ which assigns to local variable $l$.
-
-\item
-$\MclsVarAsgnInst : \Mdom \times \Vdom \times \Vdom$
-
-Relation containing each tuple $(m,l,b)$ such that method $m$
-contains a statement of the form $l = b.{\tt getClass()}$.
-
-\item
-$\MgetInstFldInst : \Mdom \times \Vdom \times \Fdom \times \Vdom$
-
-Relation containing each tuple $(m,b,f,r)$ such that method $m$
-contains a statement of the form $b.f = r$ which assigns to
-instance field or array element $f$ of the object denoted by local
-variable $b$ the value denoted by local variable $r$.
-
-\item
-$\MputInstFldInst : \Mdom \times \Vdom \times \Fdom \times \Vdom$
-
-Relation containing each tuple $(m,b,f,r)$ such that method $m$
-contains a statement of the form $b.f = r$ which writes the value
-denoted by local variable $r$ to instance field or array element
-$f$ of the object denoted by local variable $b$.
-
-\item
-$\MgetStatFldInst : \Mdom \times \Vdom \times \Fdom$
-
-Relation containing each tuple $(m,l,f)$ such that method $m$ contains
-a statement of the form $l = f$ that assigns to local variable $l$ the
-value denoted by static field $f$.
-
-\item
-$\MputStatFldInst : \Mdom \times \Fdom \times \Vdom$
-
-Relation containing each tuple $(m,f,r)$ such that method $m$ contains
-a statement of the form $f = r$ which assigns to static field $f$ the
-value denoted by local variable $r$.
-
-\item
-$\virtIM : \Idom \times \Mdom$
-
-Relation containing each tuple $(i,m)$ such that $m$ is the
-{\it resolved method} of {\tt invokevirtual} or {\tt invokeinterface} method
-invocation statement $i$.
-
-\item
-$\statIM : \Idom \times \Mdom$
-
-Relation containing each tuple $(i,m)$ such that $m$ is the
-{\it resolved method} of {\tt invokestatic} method invocation statement $i$.
-
-\item
-$\specIM : \Idom \times \Mdom$
-
-Relation containing each tuple $(i,m)$ such that $m$ is the
-{\it resolved method} of {\tt invokespecial} method invocation statement $i$.
-
-\item
-$\syncMV : \Mdom \times \Vdom$
-
-Relation containing each tuple $(m,v)$ such that method $m$ is
-synchronized on argument variable $v$.
-
-\item
-$\writeE : \Edom$
-
-Relation containing all statements that write (as opposed to read)
-an instance field, static field, or array element.
-
-\item
-$\EF : \Edom \times \Fdom$
-
-Relation containing each tuple $(e,f)$ such that statement $e$
-accesses (reads or writes) instance field, static field, or
-array element $f$.
-
-\item
-$\VH : \Vdom \times \Hdom$
-
-Relation containing each tuple $(v,h)$ such that local variable $v$
-may point to abstract object $h$, as determined by 0CFA-based may
-alias analysis.
-
-\item
-$\FH : \Fdom \times \Hdom$
-
-Relation containing each tuple $(f,h)$ such that static field $f$ may
-point to abstract object $h$, as determined by 0CFA-based may alias
-analysis.
-
-\item
-
-$\HFH : \Hdom \times \Fdom \times \Hdom$
-
-Relation containing each tuple $(h_1,f,h_2)$ such that instance field
-or array element $f$ of abstract object $h_1$ may point to abstract
-object $h_2$, as determined by 0CFA-based may alias analysis.
-
-\item
-$\IM : \Idom \times \Mdom$
-
-Relation containing each tuple $(i,m)$ such that method invocation
-statement $i$ may call method $m$ as determined by 0CFA-based may alias
-analysis.
-
-\item
-$\nestedLocks : \Cdom \times \Cdom \times \Mdom \times \Cdom \times \Mdom$
-
-Relation containing each tuple $(c,c_1,m_1,c_2,m_2)$ such that in the
-context-sensitive call graph computed by k-object-sensitive may
-alias analysis, abstract thread $c$ may first acquire the lock at
-the entry of method $m_1$ in abstract context $c_1$ and without
-releasing it proceed to acquire the lock at the entry of method
-$m_2$ in abstract context $c_2$.
-
-\item
-$\forkC : \Cdom$
-
-Relation containing all abstract contexts in which the
-{\tt start()} method of class {\tt java.lang.Thread} may
-be reachable from the main method, as determined by
-k-object-sensitive may alias analysis.
-
-\item
-$\CVC : \Cdom \times \Vdom \times \Cdom$
-
-Relation containing each tuple $(c_1,v,c_2)$ such that local variable
-$v$ may point to abstract object $c_2$ in abstract context $c_1$ of its
-containing method, as determined by k-object-sensitive may alias
-analysis.
-
-\item
-$\FC : \Fdom \times \Cdom$
-
-Relation containing each tuple $(f,c)$ such that static field $f$
-may point to abstract object $c$, as determined by k-object-sensitive
-may alias analysis.
-
-\item
-$\CFC : \Cdom \times \Fdom \times \Cdom$
-
-Relation containing each tuple $(c_1,f,c_2)$ such that instance field
-or array element $f$ of abstract object $c_1$ may point to abstract
-object $c_2$, as determined by k-object-sensitive may alias analysis.
-
-\item
-$\CICM : \Cdom \times \Idom \times \Cdom \times \Mdom$
-
-Relation containing each tuple $(c_1,i,c_2,m)$ such that method
-invocation statement $i$ in abstract context $c_1$ of its containing method
-may call method $m$ in abstract context $c_2$, as determined by
-k-object-sensitive may alias analysis.
-
-\item
-$\CHC : \Cdom \times \Hdom \times \Cdom$
-
-Relation containing each tuple $(c_1,h,c_2)$ such that the method
-containing object allocation statement $h$ may be reachable in
-abstract context $c_1$, and $c_2 = (h::c_1) + k$, where $k$ is the k
-value of statement $h$, provided as an argument to k-object-sensitive
-may alias analysis.
-
-\end{itemize}
-
-\texonly{\newpage}
-\section{Pre-defined Datalog-based Analyses}
-\label{sec-predef-datalog-analyses}
-
-\texonly{\newpage}
-\section{Pre-defined Java-based Analyses}
-\label{sec-predef-java-analyses}
-
-\texonly{\newpage}
-\bibliographystyle{plain}
-\bibliography{citations}
+\input{whatis_chord}
+\newpage
+\input{download_and_installation}
+\newpage
+\input{program_setup}
+\newpage
+\input{building_scope}
+\newpage
+\input{running_analysis}
+\newpage
+\input{writing_analysis}
+\newpage
+\input{writing_dynamic_analysis}
+\newpage
+\input{instr_events}
+\newpage
+\input{writing_datalog_analysis}
+\newpage
+\input{tuning_datalog_analysis}
+\newpage
+\input{program_representation}
+\newpage
+\input{pointsto}
+\newpage
+\input{datarace}
+\newpage
+\input{deadlock}
+\newpage
+\input{chord_sysprops}
+\newpage
+\input{acknowledgments}
 
 \end{document}
-
-\newpage
-\section{Defining Tasks and Targets}
-
-\Chord\ uses the notion of {\it tasks} to express computations
-and the notion of {\it targets} to express the structure of
-inputs/outputs of computations.
-
-Tasks and targets are both implemented uniformly as {\it entities}.
-A generic entity is represented by class {\tt chord.Entity}
-depicted in Figure \ref{fig-entity}.
-
-\begin{figure}
-\begin{verbatim}
-public class Entity {
-    protected static final String[] emptyStringArray = new String[0];
-    // subclasses may override
-    public void execute() {
-        // do nothing
-    }
-    // subclasses may override
-    public String[] getConsumedTargetNames() {
-        return emptyStringArray;
-    }
-    // subclasses may override
-    public String[] getProducedTargetNames() {
-        return emptyStringArray;
-    }
-    ...
-}
-\end{verbatim}
-\caption{Snippet of class {\tt chord.Entity}}
-\label{fig-entity}
-\end{figure}
-
-Whenever an enntity is regarded as a task:
-
-\begin{itemize}
-\item
-The {\tt execute} method defines the computation that must be performed by the task.
-
-\item
-The {\tt getConsumedTargetNames} method specifies the names of targets
-that are inputs to (the computation performed by) the task.
-
-\item
-The {\tt getProducedTargetNames} method specifies the names of targets
-that are outputs of (the computation performed by) the task.
-\end{itemize}
-
-There are four kinds of entities:
-\begin{itemize}
-\item
-A program domain
-
-\item
-A program relation
-
-\item
-A program analysis expressed in Datalog
-
-\item
-A program analysis expressed in Java
-
-\end{itemize}
-
-We next describe each of these kinds of entities in detail.
-
-\subsection{Program Domains}
-
-A program domain is a set of related elements in a given program (e.g.,
-the set of all methods, the set of all object allocation statements, etc.).
-The structure of a generic program domain is represented by
-class {\tt chord.Dom} depicted in Figure \ref{fig-dom}.
-
-\begin{figure}
-\begin{verbatim}
-public class Dom<T> extends Entity {
-    public final IndexMap<T> map;
-    public Dom() {
-        this.map = new IndexMap<T>();
-    }
-    public void execute() {
-        populate();
-        saveToDomFile();
-    }
-    // subclasses cannot override
-    public final String[] getProducedTargetNames() {
-        return emptyStringArray;
-    }
-    // subclasses may override
-    protected void populate() {
-        // do nothing
-    }
-    ...
-\end{verbatim}
-\caption{Snippet of class {\tt chord.Dom}}
-\label{fig-dom}
-\end{figure}
-
-A program domain provides a map from each element in the domain
-to a unique integer in the range $[0..n-1]$ where $n$ is the size of
-the domain.
-This map is required by {\tt bddbddb} (a BDD-based Datalog solver).
-The {\tt execute} method is specialized to construct the map and then
-save the map to disk for use by {\tt bddbddb}.
-Subclasses must override the {\tt populate} method to construct the
-map as desired.
-
-It is not possible to override the {\tt getProducedTargetNames}
-method (note that it is declared {\tt final}) because a program domain
-task is a specialized task whose output is exactly one target, namely,
-the program domain it constructs.
-It is, however, possible to override the {\tt getConsumedTargetNames}
-method (inherited from superclass {\tt chord.Entity}) since constructing
-a program domain may require other targets (i.e., other program
-domains, program relations, or program analyses).
-
-Figure \ref{fig-domI} defines the program domain
-named $\Idom$ as the domain of all method invocation statements.
-Note from method {\tt getConsumedTargetNames} that it takes as input
-the domain named $\Mdom$ which is the domain of all methods.
-This is because the {\tt populate} method refers to that domain.
-Intuitively, the {\tt populate} method gives adjacent indices in
-domain $\Idom$ to method invocation statements contained in
-methods that have adjacent indices in domain $\Mdom$ are given
-adjacent indices in domain $\Idom$.
-
-\begin{figure}
-\begin{verbatim}
-public class DomI extends Dom<InvkInst> {
-    private IndexMap<Method> mapM;
-    private final String[] consumedTaskNames = new String[] { "M" };
-    public String[] getConsumedTargetNames() {
-        return consumedTaskNames;
-    }
-    protected void populate() {
-        DomM domM = (DomM) project.getEntity("M");
-        mapM = domM.map;
-        for (int m = 0; m < mapM.size(); m++) {
-            Method meth = mapM.get(m);
-            Body body = meth.body;
-            if (body == null)
-                continue;
-            Set<Inst> insts = body.cfg.getNodes();
-            for (Inst inst : insts) {
-                if (inst instanceof InvkInst)
-                    map.set((InvkInst) inst);
-            }
-        }
-    }
-    ...
-}
-\end{verbatim}
-\caption{Snippet of class {\tt chord.DomI}}
-\label{fig-domI}
-\end{figure}
-
-A list of pre-defined program domains is presented in Section \ref{sec-predef-domains}.
-
-\subsection{Program Relations}
-
-A program relation is a relation over one or more program domains.
-The structure of a generic program relation is represented by class
-{\tt chord.Rel} depicted in Figure \ref{fig-rel}.
-
-\begin{figure}
-\begin{verbatim}
-\end{verbatim}
-\caption{Snippet of class {\tt chord.Rel}}
-\label{fig-rel}
-\end{figure}
-
-It is not possible to override the {\tt getProducedTargetNames}
-method (note that it is declared {\tt final}) because a program relation
-task is a specialized task whose output is exactly one target, namely,
-the program relation it constructs.
-It is, however, possible to override the {\tt getConsumedTargetNames}
-method (inherited from superclass {\tt chord.Entity}) since constructing
-a program relation may require other targets (i.e., program
-domains, other program relations, or program analyses).
-
-A list of pre-defined program relations is presented in Section \ref{sec-predef-relations}.
-
-\subsection{Program Analyses in Datalog}
-
-A list of pre-defined program analyses in Datalog is presented in
-Section \ref{sec-predef-datalog-analyses}.
-
-\subsection{Program Analyses in Java}
-
-A list of pre-defined program analyses in Java is presented in
-Section \ref{sec-predef-java-analyses}.

### 2010-04-02 main/doc/whatis_chord.tex

\section{What is Chord?}
\label{sec:whatis-chord}

\Chord\ is a static and dynamic program analysis framework for Java.
It has the following key characteristics:

\begin{itemize}
\item
Stand-alone: various off-the-shelf program analyses are provided (e.g., may-alias, thread-escape, datarace, deadlock, etc.).
\item
Extensible: users can build their own program analyses on top of the provided ones.
\item
Compositional: each program analysis can be written independently and yet made to interact in complex ways with other program analyses.
\item
Efficient: results computed by each program analysis are cached for reuse by other program analyses without re-computation.
\item
Flexible: a broad range of program analyses can be expressed, including those
written imperatively in Java or declaratively in Datalog, summary-based as well as cloning-based context-sensitive analyses,
iterative refinement-based analyses, client-driven analyses, and combined static and dynamic analyses.
\end{itemize}

\noindent \Chord\ is intended for use by a broad range of users:

\begin{itemize}
\item
Program analysis writers: Program analysis researchers wanting to implement and evaluate new program analysis algorithms.
\item
Program analysis appliers: Researchers with possibly a limited program analysis background wanting to build applications on
top of program analyses in Chord used as black boxes.
\item
Software engineers: Programmers wanting to use program analyses in Chord to assist with software development and testing.
\end{itemize}

\noindent \Chord\ is intended to work on a variety of platforms, including Linux, Windows/Cygwin, and MacOS.
It has been tested most extensively on Linux.
It is open source software distributed under the \htmladdnormallink{New BSD License}{http://www.opensource.org/licenses/bsd-license.php}.
Contributions from users are welcome and encouraged.
The project website is located at \url{http://code.google.com/p/jchord/}.

### 2010-04-02 main/doc/writing_dynamic_analysis.tex

@@
 \label{sec:writing-dynamic-analysis}
+
+Follow the following steps to write your own dynamic analysis.
 
@@
 the kind and format of events to be generated during an instrumented program's
-execution. See class \verb+chord.instr.InstrScheme+ for the kinds of supported
-events and their formats.
+execution. See Section \ref{sec:instr-events} 
+for the kinds of supported events and their formats.
 
@@
 
-Create a class extending \verb+chord.project.DynamicAnalysis+ and overriding
+Create a class extending \code{chord.project.DynamicAnalysis} and override
 the appropriate methods in it.
-
-The only method that must be compulsorily overridden is method \verb+getInstrScheme()+,
+The only method that must be compulsorily overridden is method \code{getInstrScheme()},
 which must return an instance of the instrumentation scheme chosen in Step 1 above,
-plus each \verb+process<event>(<args>)+ method that corresponds to event \verb+<event>+
-with format \verb+<args>+ enabled by the chosen instrumentation scheme.
+plus each \code{process<event>(<args>)} method that corresponds to event \code{<event>}
+with format \code{<args>} enabled by the chosen instrumentation scheme.
 All other methods are no-ops if not overridden.
 
-A sample such class called \verb+MyDynamicAnalysis+ is shown below:
+A sample such class called \code{MyDynamicAnalysis} is shown below:
 
@@
         public void process<event1>(<args1>) {
-            // ***TODO***: User code for handling events of kind <event1> with format <args1>.
+            // ***TODO***: User code for handling events of kind
+            // <event1> with format <args1>.
         }
@@
         public void process<eventN>(<argsN>) {
-            // ***TODO***: User code for handling events of kind <eventN> with format <argsN>.
+            // ***TODO***: User code for handling events of kind
+            // <eventN> with format <argsN>.
         }
@@
 Determine the runtime event handler for your dynamic analysis. There are two kinds of runtime event
-handlers, offline and online; see package \verb+chord.runtime+ for more details.
+handlers, offline and online; see package \code{chord.runtime} for more details.
 
 For most dynamic analyses, an offline event hander suffices, in which case you don't need to do
-anything in this step (except that you may want to change the default values of system properties
-\verb+chord.trace.pipe+ and \verb+chord.trace.block.size+ for performance; see
-Section \ref{sec:chord-sysprops} for the meaning of these properties). Otherwise, you need to
-create a class extending class \verb+chord.runtime.Runtime+ and override the appropriate methods
-in it, and provide the fully-qualified name of the created class as the value of system property
-\verb+chord.runtime.class+ (whose default value is the class \verb+chord.runtime.BufferedRuntime+).
+anything in this step.
+Otherwise, you need to create a class extending \code{chord.runtime.Runtime} and override the
+appropriate methods in it, and provide the fully-qualified name of the created class as the value of property
+\code{chord.runtime.class} (whose default value is \code{chord.runtime.BufferedRuntime}).
 
@@
 
-Compile the analysis by placing the directory containing class \verb+MyDynamicAnalysis+ created
-above in the path defined by system property \verb+chord.java.analysis.path+.
+Compile the analysis by placing the directory containing class \code{MyDynamicAnalysis} created
+above in the path defined by property \code{chord.java.analysis.path}.
 
@@
 Provide the IDs of program runs to be generated (say 1, 2, ..., M) and the command-line arguments to be
-used for the program in each of those runs (say <args1>, ..., <argsM>) via system properties
-\verb+chord.run.ids=1,2,...,N+ and \verb+chord.args.1=<args1>+, ..., \verb+chord.args.M=<argsM>+.
-
-By default, \verb+chord.run.ids=0+ and \verb+chord.args.0=""+, that is, the program will be run only
+used for the program in each of those runs (say \code{<args1>}, ..., \code{<argsM>}) via properties
+\code{chord.run.ids=1,2,...,N} and \code{chord.args.1=<args1>}, ..., \code{chord.args.M=<argsM>}.
+By default, \code{chord.run.ids=0} and \code{chord.args.0=""}, that is, the program will be run only
 once (using run ID 0) with no command-line arguments.
@@
 
-To run the analysis, set system property \verb+chord.run.analyses=<name-of-analysis>+ where
-\verb+<name-of-analysis>+ is the name that was provided in the \verb+@Chord+ annotation for class
-\verb+MyDynamicAnalysis+ created above.
+To run the analysis, set property \code{chord.run.analyses} to \code{<name-of-analysis>}
+(recall that \code{<name-of-analysis>} is the name provided in the \code{@Chord} annotation for class
+\code{MyDynamicAnalysis} created above).
+
+%(except that you may want to change the default values of properties
+%\code{chord.trace.pipe} and \code{chord.trace.block.size} for performance, or the default values
+%of properties \code{chord.trace.pipe} and \code{chord.reuse.trace} for determinism across; see
+%Section \ref{sec:chord-sysprops} for the meaning of these properties).

### 2010-04-02 main/doc/instr_events.tex

\section{Instrumentation Events}
\label{sec:instr-events}

Chord allows a given Java program to be instrumented to generate the
following kinds of events during its execution.

\begin{enumerate}
\item
Method entry and exit. \\
{\bf Event:} \code{ENTER_METHOD} $m$ $t$ \\
{\bf Desc.:} Generated after thread $t$ enters method $m$ respectively. \\
{\bf Event:} \code{LEAVE_METHOD} $m$ $t$ \\
{\bf Desc.:} Generated before thread $t$ leaves method $m$ respectively.

\item
Loop entry and exit. \\
{\bf Event:} \code{ENTER_LOOP} $w$ $t$ \\
{\bf Desc.:} Generated before thread $t$ enters loop $w$ respectively. \\
{\bf Event:} \code{LEAVE_LOOP} $w$ $t$ \\
{\bf Desc.:} Generated after thread $t$ leaves loop $w$ respectively.

\item
Object allocation \\
{\bf Event:} \code{BEF_NEW} $h$ $t$ \\
{\bf Desc.:} Generated only in the crude trace before thread $t$ executes a \code{new} bytecode instruction at program point $h$. \\
{\bf Event:} \code{AFT_NEW} $h$ $t$ $o$ \\
{\bf Desc.:} Generated only in the crude trace after thread $t$ executes a \code{new} bytecode instruction at program point $h$ and allocates new object $o$. \\
%Note: Both above events are visible only in the crude trace and not in the final trace. They are generated only if argument \code{hasObj} of method {@link #setNewAndNewArrayEvent} is true.
{\bf Event:} \code{NEW} $h$ $t$ $o$ \\
{\bf Desc.:} Generated after thread $t$ executes a \code{new} bytecode instruction at program point $h$ and allocates new object $o$. \\
{\bf Event:} \code{NEW_ARRAY} $h$ $t$ $o$ \\
{\bf Desc.:} Generated after thread $t$ executes a \code{newarray} bytecode instruction at program point $h$ and allocates new object $o$.

\item
Getstatic primitive \\
{\bf Event:} \code{GETSTATIC_PRIMITIVE} $e$ $t$ $f$ \\
{\bf Desc.:} Generated after thread $t$ reads primitive-typed static field $f$ at program point $e$.

\item
Getstatic reference \\
{\bf Event:} \code{GETSTATIC_REFERENCE} $e$ $t$ $f$ $o$ \\
{\bf Desc.:} Generated after thread $t$ reads object $o$ from reference-typed static field $f$ at program point $e$.

\item
Putstatic primitive \\
{\bf Event:} \code{PUTSTATIC_PRIMITIVE} $e$ $t$ $f$ \\
{\bf Desc.:} Generated after thread $t$ writes primitive-typed static field $f$ at program point $e$.

\item
Putstatic reference \\
{\bf Event:} \code{PUTSTATIC_REFERENCE} $e$ $t$ $f$ $o$ \\
{\bf Desc.:} Generated after thread $t$ writes object $o$ to reference-typed static field $f$ at program point $e$.

\item
Getfield primitive \\
{\bf Event:} \code{GETFIELD_PRIMITIVE} $e$ $t$ $b$ $f$ \\
{\bf Desc.:} Generated after thread $t$ reads primitive-typed instance field $f$ of object $b$ at program point $e$.

\item
Getfield reference \\
{\bf Event:} \code{GETFIELD_REFERENCE} $e$ $t$ $b$ $f$ $o$ \\
{\bf Desc.:} Generated after thread $t$ reads object $o$ from reference-typed instance field $f$ of object $b$ at program point $e$.

\item
Putfield primitive \\
{\bf Event:} \code{PUTFIELD_PRIMITIVE} $e$ $t$ $b$ $f$ \\
{\bf Desc.:} Generated after thread $t$ writes primitive-typed instance field $f$ of object $b$ at program point $e$.

\item
Putfield reference \\
{\bf Event:} \code{PUTFIELD_REFERENCE} $e$ $t$ $b$ $f$ $o$ \\
{\bf Desc.:} Generated after thread $t$ writes object $o$ to reference-typed instance field $f$ of object $b$ at program point $e$.

\item
Aload primitive \\
{\bf Event:} \code{ALOAD_PRIMITIVE} $e$ $t$ $b$ $i$ \\
{\bf Desc.:} Generated after thread $t$ reads the primitive-typed element at index $i$ of array object $b$ at program point $e$.

\item
Aload reference \\
{\bf Event:} \code{ALOAD_REFERENCE} $e$ $t$ $b$ $i$ $o$ \\
{\bf Desc.:} Generated after thread $t$ reads object $o$ from the reference-typed element at index $i$ of array object $b$ at program point $e$.

\item
Astore primitive \\
{\bf Event:} \code{ASTORE_PRIMITIVE} $e$ $t$ $b$ $i$ \\
{\bf Desc.:} Generated after thread $t$ writes the primitive-typed element at index $i$ of array object $b$ at program point $e$.

\item
Astore reference \\
{\bf Event:} \code{ASTORE_REFERENCE} $e$ $t$ $b$ $i$ $o$ \\
{\bf Desc.:} Generated after thread $t$ writes object $o$ to the reference-typed element at index $i$ of array object $b$ at program point $e$.

\item
Method call \\
{\bf Event:} \code{METHOD_CALL} $i$ $t$ \\
{\bf Desc.:} Generated before or after thread $t$ executes the method invocation statement at program point $i$.

\item
Thread start call \\
{\bf Event:} \code{THREAD_START} $i$ $t$ $o$ \\
{\bf Desc.:} Generated before thread $t$ calls the \code{start()} method of \code{java.lang.Thread} at program point $i$ and spawns a thread $o$.

\item
Thread join call \\
{\bf Event:} \code{THREAD_JOIN} $i$ $t$ $o$ \\
{\bf Desc.:} Generated before thread $t$ calls the \code{join()} method of \code{java.lang.Thread} at program point i to join with thread $o$.

\item
Lock acquire \\
{\bf Event:} \code{ACQUIRE_LOCK} $l$ $t$ $o$ \\
{\bf Desc.:} Generated after thread $t$ executes a statement of the form monitorenter $o$ or enters a method synchronized on $o$ at program point $l$.

\item
Lock release \\
{\bf Event:} \code{RELEASE_LOCK} $r$ $t$ $o$ \\
{\bf Desc.:} Generated before thread t executes a statement of the form monitorexit o or leaves a method synchronized on o at program point r.

\item
Thread wait call \\
{\bf Event:} \code{WAIT} $i$ $t$ $o$ \\
{\bf Desc.:} Generated before thread $t$ calls the \code{wait()} method of \code{java.lang.Object} at program point $i$ on object $o$.

\item
Thread notify call \\
{\bf Event:} \code{NOTIFY} $i$ $t$ $o$ \\
{\bf Desc.:} Generated before thread $t$ calls the \code{notify()} or \code{notifyAll()} method of \code{java.lang.Object} at program point $i$ on object $o$.
\end{enumerate}

### 2010-04-03 main/doc/building_scope.tex

@@
 specifying the method's name, the method's descriptor, and the method's declaring class
-(e.g., \code{main:([Ljava/lang/String;)V@foo.bar.Main+}).
+(e.g., \code{main:([Ljava/lang/String;)V@foo.bar.Main}).
 Chord throws a runtime exception if the declaring class of a method in the methods file is not
@@
 The \code{rta} value instructs Chord to compute the analysis scope statically using Rapid Type Analysis.
+
 \item
@@
 The default values of properties \code{chord.classes.file} and \code{chord.methods.file} are
-\code{{chord.out.dir}/classes.txt} and \code{{chord.out.dir}/methods.txt}, respectively,
-where property \code{chord.out.dir} defaults to \code{{chord.work.dir}/chord_output/},
+\code{[chord.out.dir]/classes.txt} and \code{[chord.out.dir]/methods.txt}, respectively,
+where property \code{chord.out.dir} defaults to \code{[chord.work.dir]/chord_output/},
 and property \code{chord.work.dir} defaults to the current working directory.

### 2010-04-03 main/doc/chord_sysprops.tex

@@
 {\bf Description:} Directory containing libraries needed by Chord. \\
-{\bf Default value:} \code{{chord.main.dir}/lib/}
+{\bf Default value:} \code{[chord.main.dir]/lib/}
 
@@
 {\bf Description:} Classpath of Chord.  It includes the path specified by property \code{chord.java.analysis.path} to allow running user-defined program analyses. \\
-{\bf Default value:} See \code{{chord.main.dir}/build.xml}
+{\bf Default value:} See \code{[chord.main.dir]/build.xml}
 
@@
 {\bf Description:} Classpath of \code{bddbddb}. \\
-{\bf Default value:} See \code{{chord.main.dir}/build.xml}
+{\bf Default value:} See \code{[chord.main.dir]/build.xml}
 
@@
 {\bf Description:} Shared library implementing Chord instrumentation agent. \\
-{\bf Default value:} \code{{chord.main.dir}/lib/libchord_instr_agent.so}
+{\bf Default value:} \code{[chord.main.dir]/lib/libchord_instr_agent.so}
 
@@
 {\bf Type:} location \\
-{\bf Description:} Properties file loaded by \code{{chord.main.dir}/build.xml}.
+{\bf Description:} Properties file loaded by \code{[chord.main.dir]/build.xml}.
 Any of the below properties, as well as other user-defined properties to be passed to Chord (e.g., for user-defined analyses), may be set in this file.
@@
 \\
-{\bf Default value:} \code{{chord.work.dir}/chord.properties}
+{\bf Default value:} \code{[chord.work.dir]/chord.properties}
 
@@
 {\bf Description:} Arguments passed to JVM running Chord. \\
-{\bf Default value:} \code{"-showversion -ea -Xmx{chord.max.heap} -Xss{chord.max.stack}"}
+{\bf Default value:} \code{"-showversion} \code{-ea} \code{-Xmx[chord.max.heap]} \code{-Xss[chord.max.stack]"}
 
@@
 \code{chord.build.scope} \\
- bool \\
- Compute analysis scope (i.e., reachable classes and methods)
-using algorithm specified by properties \code{chord.scope.kind} and \code{chord.reuse.scope}.
-Program scope is computed regardless of the value of this property if another task (e.g., a
-program analysis to be run) needs it.
-{\bf Default value:} \code{false} \\
+{\bf Type:} bool \\
+{\bf Description:} Compute analysis scope (i.e., reachable classes and methods) using algorithm specified by properties \code{chord.scope.kind} and \code{chord.reuse.scope}.
+Program scope is computed regardless of the value of this property if another task (e.g., a program analysis to be run) needs it. \\
+{\bf Default value:} \code{false}
 
@@
 \code{chord.run.analyses} \\
- string list \\
- List of names of program analyses to be run
-in order.
+{\bf Type:} string list \\
+{\bf Description:} List of names of program analyses to be run in order. \\
 {\bf Default value:} \code{""}
@@
 {\bf Type:} string list \\
-{\bf Description:} List of names of program relations whose contents must be printed to files \code{{chord.out.dir}/<...>.txt} where \code{<...>} denotes the relation name.  \\
+{\bf Description:} List of names of program relations whose contents must be printed to files \code{[chord.out.dir]/<...>.txt} where \code{<...>} denotes the relation name.  \\
 {\bf Default value:} \code{""}
@@
 \code{chord.publish.targets} \\
- bool \\
- Publish all targets defined by program analyses in
-paths specified by properties \code{chord.java.analysis.path} and
-\code{chord.dlog.analysis.path}.
-{\bf Default value:} \code{false} \\
+{\bf Type:} bool \\
+{\bf Description:} Publish all targets defined by program analyses in paths specified by properties \code{chord.java.analysis.path} and \code{chord.dlog.analysis.path}. \\
+{\bf Default value:} \code{false}
 
@@
 \code{chord.main.class} \\
- class \\
- Fully-qualified name of main class of given program
-(e.g., \code{com.example.Main}). \\
+{\bf Type:} class \\
+{\bf Description:} Fully-qualified name of main class of given program (e.g., \code{com.example.Main}).
 
@@
 {\bf Type:} path \\
- Classpath of given program. \code{{sun.boot.class.path}}
-is implicitly appended to it.
-{\bf Default value:} \code{{sun.boot.class.path}}
+{\bf Description:} Classpath of given program. \code{[sun.boot.class.path]} is implicitly appended to it. \\
+{\bf Default value:} \code{[sun.boot.class.path]}
 
@@
 {\bf Type:} path \\
- Java source path of given program. \\
+{\bf Description:} Java source path of given program. \\
 {\bf Default value:} \code{""}
@@
 \code{chord.run.ids} \\
- string list \\
- List of IDs to identify runs of given program.
+{\bf Type:} string list \\
+{\bf Description:} List of IDs to identify runs of given program. \\
 {\bf Default value:} \code{0}
@@
 \item
-\code{chord.args.XXX} \\
- string \\
- Command-line argument string to be used for given
-program in run having ID \code{XXX}.
+\code{chord.args.<id>} \\
+{\bf Type:} string \\
+{\bf Description:} Command-line argument string to be used for given program in run having ID \code{<id>}. \\
 {\bf Default value:} \code{""}
@@
 \code{chord.reuse.scope} \\
- bool \\
- Treat analysis scope as the classes and methods listed
-in files specified by properties \code{chord.classes.file} and \code{chord.methods.file},
-respectively. Property \code{chord.scope.kind} is ignored if this property is set to
-\code{true} and the two files exist.
+{\bf Type:} bool \\
+{\bf Description:} Treat analysis scope as the classes and methods listed in files specified by properties \code{chord.classes.file} and \code{chord.methods.file}, respectively.
+Property \code{chord.scope.kind} is ignored if this property is set to \code{true} and the two files exist.
+\\
 {\bf Default value:} \code{false}
@@
 \code{chord.check.exclude} \\
- string list \\
- List of prefixes of names of classes and
-packages to be excluded by program analyses.
-Interpretation of this property is analysis-specific. \\
-{\bf Default value:} \code{"java.,sun.,com.sun.,com.ibm.,org.apache.harmony.,joeq.,jwutil."} \\
+{\bf Type:} string list \\
+{\bf Description:} List of prefixes of names of classes and packages to be excluded by program analyses.  Interpretation of this property is analysis-specific. \\
+{\bf Default value:} \code{"java.,sun.,com.sun.,com.ibm.,org.apache.harmony.,joeq.,jwutil."}
 
@@
 {\bf Description:} Classpath containing program analyses written in Java (i.e., \code{@Chord}-annotated classes). \\
-{\bf Default value:} \code{{chord.main.dir}/classes/}
+{\bf Default value:} \code{[chord.main.dir]/classes/}
 
@@
 {\bf Description:} Path of directories containing program analyses written in Datalog (i.e., \code{*.datalog} and \code{*.dlog} files). \\
-{\bf Default value:} \code{{chord.main.dir}/src/dlog/}
+{\bf Default value:} \code{[chord.main.dir]/src/dlog/}
 
@@
 \code{chord.reuse.rels} \\
- bool \\
- Construct program relations from BDDs stored on disk
-(from a previous run of Chord) whenever possible instead of re-computing them.
-{\bf Default value:} \code{false} \\
+{\bf Type:} bool \\
+{\bf Description:} Construct program relations from BDDs stored on disk (from a previous run of Chord) whenever possible instead of re-computing them. \\
+{\bf Default value:} \code{false}
+
 \item
 \code{chord.publish.results} \\
- bool \\
- Publish the results of program analyses in HTML.
-Interpretation of this property is analysis-specific.
-{\bf Default value:} \code{true} \\
+{\bf Type:} bool \\
+{\bf Description:} Publish the results of program analyses in HTML.  Interpretation of this property is analysis-specific. \\
+{\bf Default value:} \code{true}
 
@@
 \code{chord.ssa} \\
- bool \\
- Do SSA transformation for all methods deemed reachable by the
-algorithm used to compute analysis scope.
-{\bf Default value:} \code{true} \\
+{\bf Type:} bool \\
+{\bf Description:} Do SSA transformation for all methods deemed reachable by the algorithm used to compute analysis scope. \\
+{\bf Default value:} \code{true}
 
@@
 \code{chord.bddbddb.noisy} \\
- bool \\
- Make \bddbddb\ output debug info when solving Datalog programs.
-{\bf Default value:} \code{false} \\
+{\bf Type:} bool \\
+{\bf Description:} Make \bddbddb\ output debug info when solving Datalog programs. \\
+{\bf Default value:} \code{false}
 
@@
 \code{chord.save.maps} \\
- bool \\
- Write to file \code{{chord.bddbddb.work.dir}/<...>.map}
-when saving program domain named \code{<...>} (useful for debugging Datalog programs using
-\code{debug} target of \code{{chord.main.dir}/build.xml}).
-{\bf Default value:} \code{true} \\
+{\bf Type:} bool \\
+{\bf Description:} Write to file \code{[chord.bddbddb.work.dir]/<...>.map} when saving program domain named \code{<...>} (useful for debugging Datalog programs using \code{debug} target of \code{[chord.main.dir]/build.xml}).
+\\
+{\bf Default value:} \code{true}
 
@@
 \code{chord.trace.pipe} \\
- bool \\
- Implement the dynamic trace file as a POSIX pipe instead of a regular file.
-{\bf Default value:} \code{true} \\
+{\bf Type:} bool \\
+{\bf Description:} Implement the dynamic trace file as a POSIX pipe instead of a regular file. \\
+{\bf Default value:} \code{true}
 
@@
 {\bf Type:} int \\
-{\bf Description:} Number of bytes to read/write in a single operation from/to the dynamic trace file.
-{\bf Default value:} \code{4096} \\
+{\bf Description:} Number of bytes to read/write in a single operation from/to the dynamic trace file. \\
+{\bf Default value:} \code{4096}
 
@@
 \item
-\code{chord.out.dir} \\
+\code[chord.out.dir] \\
 {\bf Type:} location \\
 {\bf Description:} Directory to which Chord dumps all files. \\
-{\bf Default value:} \code{{chord.work.dir}/chord_output/}
+{\bf Default value:} \code{[chord.work.dir]/chord_output/}
 
@@
 {\bf Description:} File to redirect standard output stream during Chord's execution. \\
-{\bf Default value:} \code{{chord.out.dir}/log.txt}
+{\bf Default value:} \code{[chord.out.dir]/log.txt}
 
@@
 {\bf Description:} File to redirect standard error stream during Chord's execution. \\
-{\bf Default value:} \code{{chord.out.dir}/log.txt}
+{\bf Default value:} \code{[chord.out.dir]/log.txt}
 
@@
 {\bf Description:} File from/to which list of classes deemed reachable is read/written. \\
-{\bf Default value:} \code{{chord.out.dir}/classes.txt}
+{\bf Default value:} \code{[chord.out.dir]/classes.txt}
 
@@
 {\bf Description:} File from/to which list of methods deemed reachable is read/written. \\
-{\bf Default value:} \code{{chord.out.dir}/methods.txt}
+{\bf Default value:} \code{[chord.out.dir]/methods.txt}
 
@@
 {\bf Description:} Directory used by \bddbddb\ as its input/output directory. \\
-{\bf Default value:} \code{{chord.out.dir}/bddbddb/}
+{\bf Default value:} \code{[chord.out.dir]/bddbddb/}
 
@@
 {\bf Description:} Directory containing instrumented JDK classes used by given program. \\
-{\bf Default value:} \code{{chord.out.dir}/boot_classes/}
+{\bf Default value:} \code{[chord.out.dir]/boot_classes/}
 
@@
 {\bf Description:} Directory containing instrumented non-JDK classes of given program. \\
-{\bf Default value:} \code{{chord.out.dir}/user_classes/}
+{\bf Default value:} \code{[chord.out.dir]/user_classes/}
 
@@
 {\bf Description:} File specifying kind and format of events in dynamic trace file. \\
-{\bf Default value:} \code{{chord.out.dir}/scheme.ser}
+{\bf Default value:} \code{[chord.out.dir]/scheme.ser}
 
@@
 {\bf Description:} Crude dynamic trace file. \\
-{\bf Default value:} \code{{chord.out.dir}/crude_trace.txt}
+{\bf Default value:} \code{[chord.out.dir]/crude_trace.txt}
 
@@
 {\bf Description:} Final dynamic trace file. \\
-{\bf Default value:} \code{{chord.out.dir}/final_trace.txt}
+{\bf Default value:} \code{[chord.out.dir]/final_trace.txt}
 \end{enumerate}

### 2010-04-09 main/doc/building_scope.tex
documentation

@@
 
-Chord computes the analysis scope (i.e., reachable classes and methods) of a given Java program
+Chord computes the analysis scope (i.e., reachable classes and methods) of the given program
 either if property \code{chord.build.scope} is set to \code{true} or if some other task (e.g.,
 a program analysis specified via property \code{chord.run.analyses}) demands it by
-calling static method \code{chord.program.Program.v()}.
+calling method \code{chord.program.Program.v()}.
 

### 2010-04-15 main/doc/building_scope.tex
updated documentation

@@
 
-Chord computes the analysis scope (i.e., reachable classes and methods) of the given program
+\Chord\ computes the analysis scope (i.e., reachable classes and methods) of the given program
 either if property \code{chord.build.scope} is set to \code{true} or if some other task (e.g.,
@@
 \code{chord.classes.file} and \code{chord.methods.file} exist,
-then Chord regards those files as specifying the classes and methods, respectively,
+then \Chord\ regards those files as specifying the classes and methods, respectively,
 to be regarded as reachable.  The format
@@
 (e.g., \code{main:([Ljava/lang/String;)V@foo.bar.Main}).
-Chord throws a runtime exception if the declaring class of a method in the methods file is not
-specified in the classes file.
+
+The default value of property \code{chord.reuse.scope} is \code{false}, and
+the default values of properties \code{chord.classes.file} and \code{chord.methods.file} are
+\code{[chord.out.dir]/classes.txt} and \code{[chord.out.dir]/methods.txt}, respectively,
+where property \code{chord.out.dir} defaults to \code{[chord.work.dir]/chord_output/},
+and property \code{chord.work.dir} defaults to the current working directory.
 
@@
 property \code{chord.classes.file} or \code{chord.methods.file} does not exist,
-then Chord computes the analysis scope
-using the algorithm specified by property \code{chord.scope.kind} and then
-writes the classes and methods deemed reachable by that algorithm to those files.
+then \Chord\ computes the analysis scope
+using the algorithm specified by property \code{chord.scope.kind} and
+then writes the classes and methods deemed reachable by that algorithm to those files.
 
-The possible legal values of property \code{chord.scope.kind} are \code{rta} and \code{dynamic}.
-In both cases, Chord expects properties \code{chord.main.class} and \code{chord.class.path}
+The possible legal values of property \code{chord.scope.kind} are [\code{dynamic}$|$\code{rta}$|$\code{rta_reflect}$|$\code{cha}].
+In each case, \Chord\ at least expects properties \code{chord.main.class} and \code{chord.class.path}
 to be set.
@@
 \item
-The \code{rta} value instructs Chord to compute the analysis scope statically using Rapid Type Analysis.
-
-\item
-The \code{dynamic} value instructs Chord to compute the analysis scope dynamically, by running the program
+The \code{dynamic} value instructs \Chord\ to compute the analysis scope dynamically, by running the program
 and observing using JVMTI the classes that are loaded at run-time.
@@
 predictive power of program analyses performed using the computed analysis scope.
+
+\item
+The \code{rta} value instructs \Chord\ to compute the analysis scope statically using Rapid Type Analysis (RTA).
+In this case, no attempt is made to resolve reflection.
+
+RTA is an iterative fixed-point algorithm.  It maintains a set of reachable methods $M$.
+The initial iteration starts by assuming that only the main method in the main class is reachable
+(\Chord\ also handles class initializer methods but we ignore them here for brevity; we also ignore
+the set of reachable classes maintained besides the set of reachable methods).
+All object allocation sites $H$ contained in methods in $M$ are deemed
+reachable (i.e., control-flow within method bodies is ignored).  Whenever a dynamically-dispatching
+method call site (i.e., an invokevirtual or invokeinterface site) with receiver of static
+type $t$ is encountered in a method in $M$, only subtypes of $t$ whose objects are allocated at some site in $H$
+are considered to determine the possible target methods, and each such target
+method is added to $M$.  The process terminates when no more methods can be added.
+
+RTA is a relatively inexpensive and precise algorithm in practice.  Its key shortcoming is that it makes
+no attempt to resolve reflection, which is rampant in real-world Java programs, and can therefore be
+unsound (i.e., underestimate the set of reachable classes and methods).
+The next option attempts to overcome this problem.
+
+\item
+The \code{rta_reflect} value instructs \Chord\ to compute the analysis scope statically using Rapid Type Analysis
+and, moreover, to resolve a common reflection pattern:
+
+\begin{quote}
+\begin{verbatim}
+String s = ...;
+Class c = Class.forName(s);
+Object o = c.newInstance();
+T t = (T) o;
+\end{verbatim}
+\end{quote}
+
+This analysis is identical to RTA except that it additionally inspects every cast statement in the program,
+such as the last statement in the above snippet,
+and queries the class hierarchy to find all concrete classes that subclass \code{T} (if \code{T} is a class)
+or that implement \code{T} (if \code{T} is an interface).
+\Chord\ allows users to control which classes are included in the class hierarchy (see below).
+
+\item
+The \code{cha} value instructs \Chord\ to compute the analysis scope statically using Class Hierarchy Analysis (CHA).
+The key difference between CHA and RTA is that for invokevirtual and invokeinterface sites with receiver of
+static type $t$, CHA considers {\it all} subtypes of $t$ in the class hierarchy to determine the possible
+target methods, whereas RTA restricts them to types of objects allocated in methods deemed reachable so far.
+As a result, CHA is highly imprecise in practice, and also expensive since it grossly overestimates the set
+of reachable classes and methods.
+Nevertheless, \Chord\ allows users to control which classes are included in the class hierarchy (see below)
+and thereby control the precision and cost of CHA.
+
 \end{itemize}
+The default value of property \code{chord.scope.kind} is \code{rta}.
 \end{itemize}
 
-The default value of property \code{chord.reuse.scope} is \code{false} and that of
-property \code{chord.scope.kind} is \code{rta}.
-The default values of properties \code{chord.classes.file} and \code{chord.methods.file} are
-\code{[chord.out.dir]/classes.txt} and \code{[chord.out.dir]/methods.txt}, respectively,
-where property \code{chord.out.dir} defaults to \code{[chord.work.dir]/chord_output/},
-and property \code{chord.work.dir} defaults to the current working directory.
+The class hierarchy is built if property \code{chord.scope.kind} has value \code{rta_reflect} or \code{cha}.
+Users can control which classes are included in building the class hierarchy.
+\Chord\ first constructs the entire classpath of the given program by concatenating in order the following
+classpaths:
+
+\begin{enumerate}
+\item
+The boot classpath, specified by property \code{sun.boot.class.path}.
+\item
+The library extensions classpath, comprising all jar files in directory \code{[java.home]/lib/dir/}.
+\item
+\Chord's classpath, specified by property \code{chord.main.class.path}.
+\item
+The classpath of user-defined Java analyses, specified by user-defined property \code{chord.java.analysis.path}
+(it is empty by default).
+\item
+The classpath of the given program, specified by user-defined property \code{chord.class.path}
+(it is empty by default).
+\end{enumerate}
+
+All classes in the entire classpath (resulting from items 1--5 above) are included in the class hierarchy
+with the following exceptions:
+\begin{itemize}
+\item
+Duplicate classes, i.e., classes with the same name occurring in more than one classpath element;
+in this case, all occurrences except the first are excluded. 
+\item
+All classes in \Chord's classpath are excluded, i.e., all classes in the classpath specified by property \code{chord.main.class.path},
+such as those in packages and sub-packages of \code{chord}, \code{joeq}, \code{net.sf.bddbddb}, \code{net.sf.javabdd}, \code{javassist}, \code{gnu.trove}, \code{net.sf.saxon}, etc.
+\item
+If property \code{chord.ch.kind} has value \code{dynamic}, then \Chord\ runs the given program and
+observes the set of all classes the JVM loads; any class not in this set is excluded.
+By default, property \code{chord.ch.dynamic} has value \code{static}.
+\item
+If the superclass of a class $c$ is missing or if an interface implemented/extended by a class/interface
+$c$ is missing, where ``missing" means that it is either not in the classpath resulting from items 1--5 above
+or it is excluded by one of these rules, then $c$ itself is excluded.  Note that this rule is recursive, e.g.,
+if $c$ has superclass $b$ which in turn has superclass $a$, and $a$ is missing, then both $b$ and $c$ are
+excluded.
+\end{itemize}
+
+%TODO: describe chord.scope.exclude?
+
+%\begin{tabular}
+%chord.scope.kind,chord.ch.dynamic & runs program? & \# reachable classes & \# reachable methods & running time \\
+%dynamic,-         & yes &    314 &  4,491 &   25s \\
+%cha,true          & yes &    427 &  1,532 &   28s \\
+%rta,-             &  no &    849 &  4,836 &    9s \\
+%rta_reflect,true  & yes &    849 &  4,836 &   34s \\
+%rta_reflect,false &  no &  9,871 & 58,726 &  7m6s \\
+%cha,false         &  no & 14,121 & 74,613 & 4m11s
+%\end{tabular}

### 2010-04-15 main/doc/program_setup.tex
updated documentation

@@
 \item
-\code{chord.class.path} specifies the classpath of the program.
-The value of pre-defined property \code{sun.boot.class.path} is implicitly appended
-to it.
+\code{chord.class.path} specifies the application-specific classpath of the program
+(the standard library classpath is implicitly included).
 \item
 \code{chord.src.path} specifies the Java source path of the program.
-All program analyses in Chord operate on Java bytecode (specified by \code{chord.class.path}).
+All program analyses in Chord operate on Java bytecode.
 The only use of this property is to HTMLize the Java source files of the program so that the
@@
 Chord must perform.
-%All tasks are summarized in the ``Chord task properties" portion of the table in Section \ref{sec:chord-sysprops}.
-The two most common tasks, described next, are building program scope (Section \ref{sec:building-scope})
-and running program analyses (Section \ref{sec:running-analysis}).
+All tasks are summarized in Section \ref{sec:chord-sysprops:task}.
+The two most common tasks, described next, are building the analysis scope of the given program (Section \ref{sec:building-scope})
+and running program analyses on the given program (Section \ref{sec:running-analysis}).

### 2010-04-15 main/doc/building_scope.tex
documentation update

@@
 The class hierarchy is built if property \code{chord.scope.kind} has value \code{rta_reflect} or \code{cha}.
-Users can control which classes are included in building the class hierarchy.
+Users can control which classes are included in building the class hierarchy by setting property \code{chord.ch.kind},
+whose possible legal values are [\code{static}$|$\code{dynamic}].
 \Chord\ first constructs the entire classpath of the given program by concatenating in order the following
@@
 observes the set of all classes the JVM loads; any class not in this set is excluded.
-By default, property \code{chord.ch.dynamic} has value \code{static}.
+By default, property \code{chord.ch.kind} has value \code{static}.
 \item

### 2010-08-06 main/doc/bdd_representation.tex
added documentation on BDDs

\section{BDD Representation}
\label{sec:bdd-representation}

Each domain containing $N$ elements is assigned $log_2 N$ BDD variables in the underlying BDD factory with contiguous IDs.
For instance,
domain {\tt F0} containing [2048..4096) elements will be assigned 12 variables with IDs (say) 59,60,61,62,63,64,65,66,67,68,69,70 and
domain {\tt Z0} containing [8..16) elements will be assigned 4 variables with IDs (say) 105,106,107,108.

If two domains are uninterleaved in the declared domain order in a Datalog program (i.e., {\tt \_} is used instead of {\tt x} between them),
then the BDD variables assigned to those domains are ordered in reverse order in the underlying BDD factory.
For instance, the BDD variable order corresponding to the declared domain order {\tt F0\_Z0} is (in level2var format)
``70,69,68,67,66,65,64,63,62,61,60,59,108,107,106,105".

If two domains are interleaved in the declared domain order in a Datalog program (i.e., {\tt x} is used instead of {\tt \_} between them),
then the BDD variables assigned to those domains are still ordered in reverse order of IDs in the underlying BDD factory,
but they are also interleaved.
For instance, the BDD variable order corresponding to the declared domain order {\tt F0xZ0} is (in level2var format)
``70,108,69,107,68,106,67,105,66,65,64,63,62,61,60,59".

Each BDD variable is at a unique ``level" which is its 0-based position in the BDD variable order in the underlying BDD factory.

We will next illustrate the format of a BDD stored on disk (in a .bdd file) using the following example:

\begin{verbatim}
# V0:16 H0:12
# 14 15 16 17 18 19 20 21 22 23 24 25 26 27 28 29
# 82 83 84 85 86 87 88 89 90 91 92 93
28489 134
39 36 33 30 27 24 21 18 15 12 9 6 3 0 81 79 77 75 73 71 69 67 65 63 61 59 57 55  \
   53 51 82 80 78 76 74 72 70 68 66 64 62 60 58 56 54 52 37 34 31 28 25 22 19 16 \
   13 10 7 4 1 117 116 115 114 113 112 111 110 109 108 107 106 50 49 48 47 46 45 \
   44 43 42 41 40 105 104 103 102 101 100 99 98 97 96 95 94 93 92 91 90 89 88 87 \
   86 85 84 83 133 132 131 130 129 128 127 126 125 124 123 122 121 120 119 118   \
   38 35 32 29 26 23 20 17 14 11 8 5 2
287 83 0 1
349123 84 287 0
349138 85 0 349123
...
\end{verbatim}

The first comment line indicates the domains in the relation (in the above case, {\tt V0} and {\tt H0},
represented using 16 and 12 unique BDD variables, respectively).

If there are $N$ domains, there are $N$ following comment lines, each specifying the
BDD variables assigned to the corresponding domain.

The following line has two numbers: the number of nodes in the represented BDD (28489 in this case) and the number of variables
in the BDD factory from which the BDD was dumped to disk.  Note that the number of variables (134 in this case) is not
necessarily the number of variables in the represented BDD (16+12=28 in this case) though it is guaranteed to be
greater than or equal to it.

The following line specifies the BDD variable order in var2level format.  In this case, the specified order subsumes
{\tt V0\_H0} (notice that levels ``81 79 77 75 73 71 69 67 65 63 61 59 57 55 53 51", which are at positions ``14 15 ... 28 29" 
in the specified order are lower than levels ``105 104 103 102 101 100 99 98 97 96 95 94" which are at positions
``82 83 .. 92 93").

Each of the following lines specifies a unique node in the represented BDD; it has format ``X V L H" where:
\begin{itemize}
\item X is the ID of the BDD node
\item V is the ID of the bdd variable labeling that node (unless it is 0 or 1, in which case it represents a leaf node)
\item L is the ID of the BDD node's low child
\item H is the ID of the BDD node's high child
\end{itemize}

The order of these lines specifying BDD nodes is such that the lines specifying the BDD nodes in the L and H entries
of each BDD node are guaranteed to occur before the line specifying that BDD node (for instance, the L entry 287 on
the second line and the R entry 349123 on the third line are IDs of BDD nodes specified on the first and second lines,
respectively).

Terminology note:
The {\it support} of a BDD {\tt b} is another BDD {\tt r} = {\tt b.support()}
that represents all the variables used in {\tt b}.
The support BDD {\tt r} is a linear tree each of whose nodes contains a separate variable,
the low branch is 0, and high branch is the node representing the next variable.
To list all the variables used in a BDD {\tt b} use {\tt r.scanSet()}.
Needless to say, {\tt scanSet()} simply walks along the high branches
starting at the root of BDD {\tt r}.

### 2010-08-14 main/doc/chord_sysprops.tex
updated documentation

@@
+% TODO: write some text at beginning of each subsection
+% TODO: do cross-referencing (i.e. put "See Also" for properties used by multiple subsections)
+
 \section{Chord System Properties}
@@
 
-The following properties are recognized by Chord.
+Users can use system properties to control \Chord's functionality.
+The following properties are recognized by \Chord.
 The separator for list-valued properties can be either a blank space, a comma, a colon, or a semi-colon.
+Notation {\tt [<...>]} is used in this section to denote the value of the property named {\tt <...>}.
 
-\begin{enumerate}
+\subsection{Basic Program Properties}
+
+This section describes basic properties of the given program that are required by program analyses in \Chord,
+such as its main class, the location(s) of its class files and Java source files, and command-line arguments
+to be used to run the program.
+
+\begin{itemize}
+\item
+\code{chord.main.class} \\
+{\bf Type:} class \\
+{\bf Description:} Fully-qualified name of the main class of the given program (e.g., \code{com.example.Main}). \\
+{\bf Note:} This property is required by virtually every task \Chord\ performs.
+
+\item
+\code{chord.class.path} \\
+{\bf Type:} path \\
+{\bf Description:} Classpath of the given program.
+It does not need to include boot classes (i.e., classes in \code{[sun.boot.class.path]}) or standard extensions (i.e., classes in {\tt .jar} files in directory \code{[java.home]/lib/ext/}).
+\\
+{\bf Default value:} {\tt ""} \\
+{\bf Note:} This property is required by virtually every task \Chord\ performs.
+
+\item
+\code{chord.src.path} \\
+{\bf Type:} path \\
+{\bf Description:} Java source path of the given program. \\
+{\bf Default value:} {\tt ""}  \\
+{\bf Note:} \Chord\ analyzes only Java bytecode, not Java source code.
+This property is used only by the task of converting Java source files into HTML files by program analyses that need to present their analysis results at the Java source code level (by calling method \code{chord.program.Program.getProgram().HTMLizeJavaSrcFiles())}.
+
+\item
+\code{chord.run.ids} \\
+{\bf Type:} string list \\
+{\bf Description:} List of IDs to identify runs of the given program. \\
+{\bf Default value:} {\tt 0} \\
+{\bf Note:} This property is used only when \Chord\ executes the given program, namely, when it is asked to compute the program scope dynamically (i.e., when \code{[chord.scope.kind]=dynamic}) or when it is asked to execute a dynamic program analysis.
+
+\item
+\code{chord.args.<id>} \\
+{\bf Type:} string \\
+{\bf Description:} Command-line arguments string to be used for the given program in the run having ID {\tt <id>}. \\
+{\bf Default value:} {\tt ""} \\
+{\bf Note:} This property is used only when \Chord\ executes the given program, namely, when it is asked to compute the program scope dynamically (i.e., when \code{[chord.scope.kind]=dynamic}) or when it is asked to execute a dynamic program analysis.
+
+\item
+\code{chord.runtime.jvmargs} \\
+{\bf Type:} string \\
+{\bf Description:} Arguments to JVM which runs the given program. \\
+{\bf Default value:} {\tt "-ea -Xmx1024m"} \\
+{\bf Note:} This property is used only when \Chord\ executes the given program, namely, when it is asked to compute the program scope dynamically (i.e., when \code{[chord.scope.kind]=dynamic}) or when it is asked to execute a dynamic program analysis.
+\end{itemize}
+
+\subsection{Chord Task Properties}
+\label{sec:chord-sysprops:task}
+
+This section describes properties that specify what task(s) \Chord\ must perform, 
+such as computing the scope of the given program or running program analyses on the given program.
+
+\begin{itemize}
+\item
+\code{chord.build.scope} \\
+{\bf Type:} bool \\
+{\bf Description:} Compute the scope (i.e., reachable classes and methods) of the given program using the algorithm specified by properties \code{chord.scope.kind} and \code{chord.reuse.scope}. \\
+{\bf Default value:} {\tt false} \\
+{\bf Note:} The program scope is computed regardless of the value of this property if another task (e.g., a program analysis specified via property \code{chord.run.analyses}) demands it by calling method \code{chord.program.Program.getProgram()}.
+
+\item
+\code{chord.run.analyses} \\
+{\bf Type:} string list \\
+{\bf Description:} List of names of program analyses to be run in order. \\
+{\bf Default value:} {\tt ""} \\
+{\bf Note:} If the analysis is written in Java, its name is specified via statement {\tt name=<...>} in its {\tt @Chord} annotation, and is the name of the class itself if this statement is missing.
+If the analysis is written in Datalog, its name is specified via a line of the form ``{\tt \# name=<...>}", and is the absolute location of the file itself if this line is missing.
+
+\item
+\code{chord.print.classes} \\
+{\bf Type:} string list \\
+{\bf Description:} List of classes in scope whose intermediate representation to print to standard output. \\
+{\bf Default value:} {\tt ""} \\
+{\bf Note:} In class names, use `.' instead of `/', and use {\tt \#} instead of the dollar character.
+
+\item
+\code{chord.print.methods} \\
+{\bf Type:} string list \\
+{\bf Description:} List of methods in scope whose intermediate representation to print to standard output. \\
+{\bf Default value:} {\tt ""} \\
+{\bf Note:} Specify each method in format \code{<mname>:<mdesc>@<cname>} where {\tt <mname>} is the method's name, {\tt <mdesc>} is the method's descriptor, and {\tt <cname>} is the name of the method's declaring class.
+In {\tt <cname>}, use `.' instead of `/', and use {\tt \#} instead of the dollar character.
+
+\item
+\code{chord.print.all.classes} \\
+{\bf Type:} bool \\
+{\bf Description:} Print intermediate representation of all classes in scope to standard output. \\
+{\bf Default value:} {\tt false}
+
+\item
+\code{chord.print.rels} \\
+{\bf Type:} string list \\
+{\bf Description:} List of names of program relations whose contents must be printed to files \code{[chord.out.dir]/<...>.txt} where {\tt <...>} denotes the relation name. \\
+{\bf Default value:} {\tt ""} \\
+{\bf Note:} This task must be used with caution as certain program relations, albeit represented compactly as BDDs, may contain a large number (e.g., billions) of tuples, resulting in voluminous output when printed to a file.
+See Section \ref{sec:tuning-datalog-analysis} for a more efficient way to query the contents of program relations (namely, by using the {\tt debug} target provided in file {\tt build.xml} in \Chord's {\tt main/} directory).
+
+\item
+\code{chord.publish.targets} \\
+{\bf Type:} bool \\
+{\bf Description:} Create files \code{targets_sortby_name.html}, \code{targets_sortby_kind.html}, and \code{targets_sortby_producers.html} in directory \code{[chord.out.dir]}, publishing all targets defined by program analyses in paths \code{[chord.java.analysis.path]} and \code{[chord.dlog.analysis.path]}.
+\\
+{\bf Default value:} {\tt false}
+\end{itemize}
+
+\subsection{Chord Boot Properties}
+
+This section describes properties of the JVM running \Chord, such as the working directory and various memory limits.
+
+\begin{itemize}
+\item
+\code{chord.work.dir} \\
+{\bf Type:} location \\
+{\bf Description:} Working directory during \Chord's execution. \\
+{\bf Default value:} current working directory
+
+\item
+\code{chord.props.file} \\
+{\bf Type:} location \\
+{\bf Description:} Properties file loaded by \code{[chord.main.dir]/build.xml}.
+Any of the below properties, as well as other user-defined properties to be passed to \Chord\ (e.g., for user-defined analyses), may be set in this file.
+Each relative (instead of absolute) path element in the value of any property named \code{chord.<...>.path} set in this file is converted to an absolute path element with respect to the directory containing this file.
+\\
+{\bf Default value:} \code{[chord.work.dir]/chord.properties}
+
+\item
+\code{chord.max.heap} \\
+{\bf Type:} string \\
+{\bf Description:} Maximum memory size of JVM running \Chord. \\
+{\bf Default value:} {\tt 1024m}
+
+\item 
+\code{chord.max.stack} \\
+{\bf Type:} string \\
+{\bf Description:} Maximum thread stack size of JVM running \Chord. \\
+{\bf Default value:} {\tt 32m}
+
+\item
+\code{chord.jvmargs} \\
+{\bf Type:} string \\
+{\bf Description:} Arguments to JVM running \Chord. \\
+{\bf Default value:} \code{"-showversion} \code{-ea} \code{-Xmx[chord.max.heap]} \code{-Xss[chord.max.stack]"}
+
+\item
+\code{chord.bddbddb.max.heap} \\
+{\bf Type:} string \\
+{\bf Description:} Maximum memory size of JVM running \bddbddb. \\
+{\bf Default value:} {\tt 1024m} \\
+{\bf Note:} Each program analysis written in Datalog is run in a separate JVM because there may be multiple invocations of the Datalog solver \bddbddb\ in a single run of \Chord\ and it is difficult to reset the global state of \bddbddb\ on each invocation.
+\end{itemize}
+
+\subsection{Program Scope Properties}
+
+This section describes properties that specify how the analysis scope is computed.
+See Section \ref{sec:building-scope} for more details.
+
+\begin{itemize}
+\item
+\code{chord.reuse.scope} \\
+{\bf Type:} bool \\
+{\bf Description:} Treat analysis scope as the classes and methods listed in files specified by properties \code{chord.classes.file} and \code{chord.methods.file}, respectively.
+Property \code{chord.scope.kind} is ignored if this property is set to {\tt true} and the two files exist.
+\\
+{\bf Default value:} {\tt false}
+
+\item
+\code{chord.scope.kind} \\
+{\bf Type:} [\code{dynamic}$|$\code{rta}$|$\code{cha}] \\
+{\bf Description:} Algorithm to compute analysis scope (i.e., reachable classes and methods).
+Current choices include {\tt dynamic} (dynamic analysis), {\tt rta} (static Rapid Type Analysis), and {\tt cha} (static Class Hierarchy Analysis).
+\\
+{\bf Default value:} {\tt rta}
+
+\item
+\code{chord.ch.kind} \\
+{\bf Type:} [\code{static}$|$\code{dynamic}] \\
+{\bf Description:} Algorithm to build the class hierarchy.
+If it is {\tt dynamic}, then the given program is executed and classes not loaded by the JVM while running the program are excluded while building the class hierarchy.
+This property is relevant only if \code{chord.scope.kind} is {\tt cha} since only this scope computing algorithm queries the class hierarchy.
+\\
+{\bf Default value:} {\tt static}
+
+% TODO: mention that below two properties are also used by instrumentor to decide which classes to exclude from instrumentation
+
+\item
+\code{chord.scope.exclude.std} \\
+{\bf Type:} string list \\
+{\bf Description:} List of prefixes of names of classes inside the standard library whose method bodies must be treated as no-ops by static scope building algorithms. \\
+{\bf Default value:} \code{[chord.main.class.path]}
+
+\item
+\code{chord.scope.exclude.ext} \\
+{\bf Type:} string list \\
+{\bf Description:} List of prefixes of names of classes outside the standard library whose method bodies must be treated as no-ops by static scope building algorithms. \\
+{\bf Default value:} {\tt ""}
+
+\item
+\code{chord.scope.exclude} \\
+{\bf Type:} string list \\
+{\bf Description:} List of prefixes of names of classes whose method bodies must be treated as no-ops by static scope building algorithms. \\
+{\bf Default value:} \code{"[chord.main.class.path],[chord.scope.exclude.ext]"}
+\end{itemize}
+
+\subsection{Program Analysis Properties}
+
+This section describes properties regarding program analyses executed by \Chord.
+
+\begin{itemize}
+\item
+\code{chord.java.analysis.path} \\
+{\bf Type:} path \\
+{\bf Description:} Classpath containing program analyses written in Java (i.e., {\tt @Chord}-annotated classes). \\
+{\bf Default value:} \code{[chord.main.dir]/classes/}
+
+\item
+\code{chord.dlog.analysis.path} \\
+{\bf Type:} path \\
+{\bf Description:} Path of directories containing program analyses written in Datalog (i.e., files with suffix {\tt .dlog} or {\tt .datalog}). \\
+{\bf Default value:} \code{[chord.main.dir]/src/dlog/:[chord.main.dir]/../extra/src/dlog/}
+
+\item
+\code{chord.reuse.rels} \\
+{\bf Type:} bool \\
+{\bf Description:} Construct program relations from BDDs stored on disk (from a previous run of \Chord) whenever possible instead of re-computing them. \\
+{\bf Default value:} {\tt false} 
+% TODO: tell when this doesn't work
+
+\item
+\code{chord.publish.results} \\
+{\bf Type:} bool \\
+{\bf Description:} Publish the results of program analyses in HTML.  Interpretation of this property is analysis-specific. \\
+{\bf Default value:} {\tt true}
+
+\code{chord.check.exclude.std} \\
+{\bf Type:} string list \\
+{\bf Description:} List of prefixes of names of classes and packages inside the standard library to be excluded by program analyses.  Interpretation of this property is analysis-specific. \\
+{\bf Default value:} \code{"sun.,com.sun.,com.ibm.jvm.,com.ibm.oti.,com.ibm.misc.,org.apache.harmony.,joeq.,jwutil.,java.,javax."}
+
+\item
+\code{chord.check.exclude.ext} \\
+{\bf Type:} string list \\
+{\bf Description:} List of prefixes of names of classes and packages outside the standard library to be excluded by program analyses.  Interpretation of this property is analysis-specific. \\
+{\bf Default value:} {\tt ""}
+
+\item
+\code{chord.check.exclude} \\
+{\bf Type:} string list \\
+{\bf Description:} List of prefixes of names of classes and packages to be excluded by program analyses.  Interpretation of this property is analysis-specific. \\
+{\bf Default value:} \code{"[chord.check.exclude.std],[chord.check.exclude.ext]"}
+\end{itemize}
+
+\subsection{Program Transformation Properties}
+
+This section describes properties regarding program transformations performed by \Chord.
+
+\begin{itemize}
+\item
+\code{chord.ssa} \\
+{\bf Type:} bool \\
+{\bf Description:} Do SSA transformation for all methods deemed reachable by the algorithm used to compute analysis scope. \\
+{\bf Default value:} {\tt true} 
+\end{itemize}
+
+\subsection{Chord Debug Properties}
+
+This section describes properties that specify the amount of debug information that \Chord\ must produce during execution.
+
+\begin{itemize}
+\item
+\code{chord.verbose} \\
+{\bf Type:} bool \\
+{\bf Description:} Produce more verbose output during \Chord's execution. \\
+{\bf Default value:} {\tt false}
+
+\item
+\code{chord.bddbddb.noisy} \\
+{\bf Type:} bool \\
+{\bf Description:} Produce more verbose output during the Datalog solver \bddbddb's execution. \\
+{\bf Default value:} {\tt false}
+
+\item
+\code{chord.save.maps} \\
+{\bf Type:} bool \\
+{\bf Description:} Write to file \code{[chord.bddbddb.work.dir]/<...>.map} when saving program domain named {\tt <...>}. \\
+{\bf Default value:} {\tt true}  \\
+{\bf Note:} This functionality is useful for debugging Datalog programs using the {\tt debug} target provided in file {\tt build.xml} in \Chord's {\tt main/} directory (see Section \ref{sec:tuning-datalog-analysis}).
+\end{itemize}
+
+\subsection{Chord Instrumentation Properties}
+
+This section describes properties regarding execution of instrumented programs for dynamic analyses.
+
+% TODO: mention chord.scope.exclude* here
+
+\begin{itemize}
+\item
+\code{chord.reuse.traces} \\
+{\bf Type:} bool \\
+{\bf Description:} Reuse dynamic traces computed by a previous run of \Chord\ if they exist. \\
+{\bf Default value:} {\tt false} \\
+{\bf Note:} Property \code{chord.trace.kind} must be set to {\tt full} if this property is set to {\tt true}.
+
+\item
+\code{chord.trace.kind} \\
+{\bf Type:} [\code{none}$|$\code{full}$|$\code{pipe}] \\
+{\bf Description:} The medium to store dynamic traces (none, regular file, or POSIX pipe). \\
+{\bf Default value:} {\tt full}
+
+\item
+\code{chord.trace.block.size} \\
+{\bf Type:} int \\
+{\bf Description:} Number of bytes to read/write in a single operation from/to the dynamic trace file. \\
+{\bf Default value:} {\tt 4096}
+
+\item
+\code{chord.max.cons.size} \\
+{\bf Type:} int \\
+{\bf Description:} Maximum number of bytes over which events generated during the execution of any constructor in the given program may span. \\
+{\bf Default value:} 50000000 \\
+{\bf Note:} This property is relevant only for dynamic analyses which want events of the form {\tt NEW} $h$ $t$ $o$ to be generated (see Section \ref{sec:instr-events}).
+The problem with generating such events at run-time is that the ID $o$ of the object freshly created by thread $t$ at object allocation site $h$ cannot be instrumented until the object is fully initialized (i.e., its constructor has finished executing).
+Hence, \Chord\ first generates a ``crude dynamic trace", which has events of the form {\tt BEF\_NEW} $h$ $t$ and {\tt AFT\_NEW} $h$ $t$ $o$ generated before and after the execution of the constructor, respectively.
+A subsequent pass generates a ``final dynamic trace", which replaces the {\tt BEF\_NEW} $h$ $t$ events by {\tt NEW} $h$ $t$ $o$ and discards the {\tt AFT\_NEW} $h$ $t$ $o$ events.
+For this purpose, however, \Chord\ must buffer all events generated between the {\tt BEF\_NEW} and {\tt AFT\_NEW} events, and this property specifies the number of bytes over which these events may span.
+If the actual number of bytes exceeds the value specified by this property (e.g., if the constructor throws an exception and the {\tt AFT\_NEW} event is not generated at all), then \Chord\ simply generates event {\tt NEW} $h$ $i$ $0$ (i.e., it treats the created object as having ID 0, which is the ID also used for {\tt null}).
+
+\item
+\code{chord.dynamic.haltonerr} \\
+{\bf Type:} bool \\
+{\bf Description:} Whether or not to terminate Chord if the given program terminates abnormally during dynamic analysis. \\
+{\bf Default value:} true
+
+\item
+\code{chord.dynamic.timeout} \\
+{\bf Type:} int \\
+{\bf Description:} The amount of time, in milliseconds, after which to kill the process running the given program during dynamic analysis, or -1 if the process must never be killed. \\
+{\bf Default value:} -1
+
+\item
+\code{chord.dynamic.silent} \\
+{\bf Type:} bool \\
+{\bf Description:} Whether or not to emit warnings while running the given program during dynamic analysis. \\
+{\bf Default value:} false
+\end{itemize}
+
+\subsection{Chord Output Properties}
+
+This section describes properties specifying the names of files and directories output by Chord.
+Most users will not want to alter the default values of these properties.
+
+\begin{itemize}
+\item
+\code{chord.out.dir} \\
+{\bf Type:} location \\
+{\bf Description:} Absolute location of the directory to which Chord dumps all files. \\
+{\bf Default value:} \code{[chord.work.dir]/chord_output/}
+
+\item
+\code{chord.out.file} \\
+{\bf Type:} location \\
+{\bf Description:} Absolute location of the file to which the standard output stream is redirected during Chord's execution. \\
+{\bf Default value:} \code{[chord.out.dir]/log.txt}
+
+\item
+\code{chord.err.file} \\
+{\bf Type:} location \\
+{\bf Description:} Absolute location of the file to which the standard error stream is redirected during Chord's execution. \\
+{\bf Default value:} \code{[chord.out.dir]/log.txt}
+
+\item
+\code{chord.classes.file} \\
+{\bf Type:} location \\
+{\bf Description:} Absolute location of the file from/to which list of classes deemed reachable is read/written. \\
+{\bf Default value:} \code{[chord.out.dir]/classes.txt}
+
+\item
+\code{chord.methods.file} \\
+{\bf Type:} location \\
+{\bf Description:} Absolute location of the file from/to which list of methods deemed reachable is read/written. \\
+{\bf Default value:} \code{[chord.out.dir]/methods.txt}
+
+\item
+\code{chord.bddbddb.work.dir} \\
+{\bf Type:} location \\
+{\bf Description:} Absolute location of the directory used by the Datalog solver \bddbddb\ as its input/output directory (namely, for program domain files {\tt *.dom} and {\tt *.map}, and program relation files {\tt *.bdd}).
+\\
+{\bf Default value:} \code{[chord.out.dir]/bddbddb/}
+
+\item
+\code{chord.boot.classes.dir} \\
+{\bf Type:} location \\
+{\bf Description:} Absolute location of the directory from/to which instrumented JDK classes used by the given program are read/written by dynamic program analyses. \\
+{\bf Default value:} \code{[chord.out.dir]/boot_classes/}
+
+\item
+\code{chord.user.classes.dir} \\
+{\bf Type:} location \\
+{\bf Description:} Absolute location of the directory from/to which instrumented non-JDK classes of the given program are read/written by dynamic program analyses. \\
+{\bf Default value:} \code{[chord.out.dir]/user_classes/}
+
+\item
+\code{chord.instr.scheme.file} \\
+{\bf Type:} location \\
+{\bf Description:} Absolute location of the file specifying the kind and format of events in dynamic trace files. \\
+{\bf Default value:} \code{[chord.out.dir]/scheme.ser}
+
+\item
+\code{chord.trace.file} \\
+{\bf Type:} location \\
+{\bf Description:} Absolute location of the dynamic trace file. \\
+{\bf Default value:} \code{[chord.out.dir]/trace}  \\
+{\bf Note:} \code{_full_verN.txt} or \code{_pipe_verN.txt} is appended to the name of the file, depending upon whether it is a regular file or a POSIX pipe, respectively, where {\tt N} is the version of the file (multiple versions are maintained if the trace is transformed by filters defined by the dynamic analysis; 0 is the final version).
+If \code{chord.reuse.traces} is set to {\tt true}, then \code{_full_verN_runM.txt} is appended to the name of the file, where {\tt M} is the run ID.
+\end{itemize}
+
+\subsection{Chord Resource Properties}
+\label{sec:chord-sysprops:resource}
+
+This section describes properties specifying classpaths and locations of runtime libraries used by Chord.
+Ordinarily, users must not alter the default values of these properties.
+
+\begin{itemize}
 \item
@@
 {\bf Type:} location \\
-{\bf Description:} \code{main/} directory in Chord's installation.
+{\bf Description:} Absolute location of the {\tt main/} directory in Chord's installation.
 
@@
 {\bf Type:} path \\
-{\bf Description:} Classpath of \code{bddbddb}. \\
+{\bf Description:} Classpath of {\tt bddbddb}. \\
 {\bf Default value:} See \code{[chord.main.dir]/build.xml}
-
-\item
-\code{chord.instr.agent.file} \\
-{\bf Type:} location \\
-{\bf Description:} Shared library implementing Chord instrumentation agent. \\
-{\bf Default value:} \code{[chord.main.dir]/lib/libchord_instr_agent.so}
 
@@
 {\bf Type:} string \\
-{\bf Description:} URL of the Javadoc location of program analyses.  It is used when publishing targets (i.e., when property \code{chord.publish.targets} is set to true). \\
-{\bf Default value:} \code{http://chord.stanford.edu/javadoc/}
-
-\item
-\code{chord.work.dir} \\
-{\bf Type:} location \\
-{\bf Description:} Working directory during Chord's execution. \\
-{\bf Default value:} current working directory
-
-\item
-\code{chord.props.file} \\
-{\bf Type:} location \\
-{\bf Description:} Properties file loaded by \code{[chord.main.dir]/build.xml}.
-Any of the below properties, as well as other user-defined properties to be passed to Chord (e.g., for user-defined analyses), may be set in this file.
-Each relative (instead of absolute) path element in the value of any property named \code{chord.<...>.path} set in this file is converted to an absolute path element with respect to the directory containing this file.
-\\
-{\bf Default value:} \code{[chord.work.dir]/chord.properties}
-
-\item
-\code{chord.max.heap} \\
-{\bf Type:} string \\
-{\bf Description:} Maximum memory size of JVM running Chord. \\
-{\bf Default value:} \code{1024m}
-
-\item 
-\code{chord.max.stack} \\
-{\bf Type:} string \\
-{\bf Description:} Maximum thread stack size of JVM running Chord. \\
-{\bf Default value:} \code{32m}
-
-\item
-\code{chord.jvmargs} \\
-{\bf Type:} string \\
-{\bf Description:} Arguments passed to JVM running Chord. \\
-{\bf Default value:} \code{"-showversion} \code{-ea} \code{-Xmx[chord.max.heap]} \code{-Xss[chord.max.stack]"}
-
-\item
-\code{chord.bddbddb.max.heap} \\
-{\bf Type:} string \\
-{\bf Description:} Maximum memory size of JVM running \bddbddb. \\
-{\bf Default value:} \code{1024m}
-
-\item
-\code{chord.build.scope} \\
-{\bf Type:} bool \\
-{\bf Description:} Compute analysis scope (i.e., reachable classes and methods) using algorithm specified by properties \code{chord.scope.kind} and \code{chord.reuse.scope}.
-Program scope is computed regardless of the value of this property if another task (e.g., a program analysis to be run) needs it. \\
-{\bf Default value:} \code{false}
-
-\item
-\code{chord.run.analyses} \\
-{\bf Type:} string list \\
-{\bf Description:} List of names of program analyses to be run in order. \\
-{\bf Default value:} \code{""}
-
-\item
-\code{chord.print.rels} \\
-{\bf Type:} string list \\
-{\bf Description:} List of names of program relations whose contents must be printed to files \code{[chord.out.dir]/<...>.txt} where \code{<...>} denotes the relation name.  \\
-{\bf Default value:} \code{""}
-
-\item
-\code{chord.publish.targets} \\
-{\bf Type:} bool \\
-{\bf Description:} Publish all targets defined by program analyses in paths specified by properties \code{chord.java.analysis.path} and \code{chord.dlog.analysis.path}. \\
-{\bf Default value:} \code{false}
-
-\item
-\code{chord.main.class} \\
-{\bf Type:} class \\
-{\bf Description:} Fully-qualified name of main class of given program (e.g., \code{com.example.Main}).
-
-\item
-\code{chord.class.path} \\
-{\bf Type:} path \\
-{\bf Description:} Classpath of given program. \code{[sun.boot.class.path]} is implicitly appended to it. \\
-{\bf Default value:} \code{[sun.boot.class.path]}
-
-\item
-\code{chord.src.path} \\
-{\bf Type:} path \\
-{\bf Description:} Java source path of given program. \\
-{\bf Default value:} \code{""}
-
-\item
-\code{chord.run.ids} \\
-{\bf Type:} string list \\
-{\bf Description:} List of IDs to identify runs of given program. \\
-{\bf Default value:} \code{0}
-
-\item
-\code{chord.args.<id>} \\
-{\bf Type:} string \\
-{\bf Description:} Command-line argument string to be used for given program in run having ID \code{<id>}. \\
-{\bf Default value:} \code{""}
-
-\item
-\code{chord.runtime.jvmargs} \\
-{\bf Type:} string \\
-{\bf Description:} Arguments to JVM when running given program.
-It is used when the program is executed to compute its scope (i.e., when property \code{chord.scope.kind} is set to \code{dynamic}) as well as when the instrumented program is executed for dynamic analyses.
-\\
-{\bf Default value:} \code{"-ea -Xmx1024m"}
-
-\item
-\code{chord.scope.kind} \\
-{\bf Type:} [\code{rta} $|$ \code{dynamic}] \\
-{\bf Description:} Algorithm to compute analysis scope (i.e., reachable classes and methods).  Current choices include \code{rta} (static Rapid Type Analysis) and \code{dynamic} (dynamic analysis). \\
-{\bf Default value:} \code{rta}
-
-\item
-\code{chord.reuse.scope} \\
-{\bf Type:} bool \\
-{\bf Description:} Treat analysis scope as the classes and methods listed in files specified by properties \code{chord.classes.file} and \code{chord.methods.file}, respectively.
-Property \code{chord.scope.kind} is ignored if this property is set to \code{true} and the two files exist.
-\\
-{\bf Default value:} \code{false}
-
-\item
-\code{chord.check.exclude} \\
-{\bf Type:} string list \\
-{\bf Description:} List of prefixes of names of classes and packages to be excluded by program analyses.  Interpretation of this property is analysis-specific. \\
-{\bf Default value:} \code{"java.,sun.,com.sun.,com.ibm.,org.apache.harmony.,joeq.,jwutil."}
-
-\item
-\code{chord.instr.exclude} \\
-{\bf Type:} string list \\
-{\bf Description:} List of prefixes of names of classes and packages to be excluded from being instrumented.
-Classes in package \code{java.lang.ref} are not instrumented regardless of the value of this property.
-\\
-{\bf Default value:} \code{"java.,sun.,com.sun.,com.ibm.,org.apache.harmony.,joeq.,jwutil."}
-
-\item
-\code{chord.java.analysis.path} \\
-{\bf Type:} path \\
-{\bf Description:} Classpath containing program analyses written in Java (i.e., \code{@Chord}-annotated classes). \\
-{\bf Default value:} \code{[chord.main.dir]/classes/}
-
-\item
-\code{chord.dlog.analysis.path} \\
-{\bf Type:} path \\
-{\bf Description:} Path of directories containing program analyses written in Datalog (i.e., \code{*.datalog} and \code{*.dlog} files). \\
-{\bf Default value:} \code{[chord.main.dir]/src/dlog/}
-
-\item
-\code{chord.reuse.rels} \\
-{\bf Type:} bool \\
-{\bf Description:} Construct program relations from BDDs stored on disk (from a previous run of Chord) whenever possible instead of re-computing them. \\
-{\bf Default value:} \code{false}
-
-\item
-\code{chord.publish.results} \\
-{\bf Type:} bool \\
-{\bf Description:} Publish the results of program analyses in HTML.  Interpretation of this property is analysis-specific. \\
-{\bf Default value:} \code{true}
-
-\item
-\code{chord.ssa} \\
-{\bf Type:} bool \\
-{\bf Description:} Do SSA transformation for all methods deemed reachable by the algorithm used to compute analysis scope. \\
-{\bf Default value:} \code{true}
-
-\item
-\code{chord.bddbddb.noisy} \\
-{\bf Type:} bool \\
-{\bf Description:} Make \bddbddb\ output debug info when solving Datalog programs. \\
-{\bf Default value:} \code{false}
-
-\item
-\code{chord.save.maps} \\
-{\bf Type:} bool \\
-{\bf Description:} Write to file \code{[chord.bddbddb.work.dir]/<...>.map} when saving program domain named \code{<...>} (useful for debugging Datalog programs using \code{debug} target of \code{[chord.main.dir]/build.xml}).
-\\
-{\bf Default value:} \code{true}
-
-\item
-\code{chord.trace.pipe} \\
-{\bf Type:} bool \\
-{\bf Description:} Implement the dynamic trace file as a POSIX pipe instead of a regular file. \\
-{\bf Default value:} \code{true}
-
-\item
-\code{chord.trace.block.size} \\
-{\bf Type:} int \\
-{\bf Description:} Number of bytes to read/write in a single operation from/to the dynamic trace file. \\
-{\bf Default value:} \code{4096}
-
-\item
-\code{chord.runtime.class} \\
-{\bf Type:} class \\
-{\bf Description:} Subclass of \code{chord.project.Runtime} used by dynamic program analyses to handle events generated during an instrumented program's execution. \\
-{\bf Default value:} \code{chord.project.BufferedRuntime}
-
-\item
-\code[chord.out.dir] \\
-{\bf Type:} location \\
-{\bf Description:} Directory to which Chord dumps all files. \\
-{\bf Default value:} \code{[chord.work.dir]/chord_output/}
-
-\item
-\code{chord.out.file} \\
-{\bf Type:} location \\
-{\bf Description:} File to redirect standard output stream during Chord's execution. \\
-{\bf Default value:} \code{[chord.out.dir]/log.txt}
-
-\item
-\code{chord.err.file} \\
-{\bf Type:} location \\
-{\bf Description:} File to redirect standard error stream during Chord's execution. \\
-{\bf Default value:} \code{[chord.out.dir]/log.txt}
-
-\item
-\code{chord.classes.file} \\
-{\bf Type:} location \\
-{\bf Description:} File from/to which list of classes deemed reachable is read/written. \\
-{\bf Default value:} \code{[chord.out.dir]/classes.txt}
-
-\item
-\code{chord.methods.file} \\
-{\bf Type:} location \\
-{\bf Description:} File from/to which list of methods deemed reachable is read/written. \\
-{\bf Default value:} \code{[chord.out.dir]/methods.txt}
-
-\item
-\code{chord.bddbddb.work.dir} \\
-{\bf Type:} location \\
-{\bf Description:} Directory used by \bddbddb\ as its input/output directory. \\
-{\bf Default value:} \code{[chord.out.dir]/bddbddb/}
-
-\item
-\code{chord.boot.classes.dir} \\
-{\bf Type:} location \\
-{\bf Description:} Directory containing instrumented JDK classes used by given program. \\
-{\bf Default value:} \code{[chord.out.dir]/boot_classes/}
-
-\item
-\code{chord.user.classes.dir} \\
-{\bf Type:} location \\
-{\bf Description:} Directory containing instrumented non-JDK classes of given program. \\
-{\bf Default value:} \code{[chord.out.dir]/user_classes/}
-
-\item
-\code{chord.instr.scheme.file} \\
-{\bf Type:} location \\
-{\bf Description:} File specifying kind and format of events in dynamic trace file. \\
-{\bf Default value:} \code{[chord.out.dir]/scheme.ser}
-
-\item
-\code{chord.crude.trace.file} \\
-{\bf Type:} location \\
-{\bf Description:} Crude dynamic trace file. \\
-{\bf Default value:} \code{[chord.out.dir]/crude_trace.txt}
-
-\item
-\code{chord.final.trace.file} \\
-{\bf Type:} location \\
-{\bf Description:} Final dynamic trace file. \\
-{\bf Default value:} \code{[chord.out.dir]/final_trace.txt}
-\end{enumerate}
+{\bf Description:} URL of the Javadoc location of program analyses.  It is used when publishing targets (i.e., when property \code{chord.publish.targets} is set to {\tt true}). \\
+{\bf Default value:} \code{http://jchord.googlecode.com/svn/wiki/javadoc/}
+\end{itemize}

### 2010-08-14 main/doc/datarace.tex
updated documentation

@@
 
-Run the datarace analysis provided in Chord by running the following command in Chord's \code{main/} directory:
+Run the datarace analysis provided in \Chord\ by running the following command in \Chord's {\tt main/} directory:
 
@@
 
-\noindent where directory \code{<...>} contains a file named \code{chord.properties} which
+\noindent where directory {\tt <...>} contains a file named \code{chord.properties} which
 defines properties \code{chord.main.class}, \code{chord.class.path}, and \code{chord.src.path}.
@@
 analysis.
-First run \code{ant} in that directory (in order to compile the program's \code{.java} files to
-\code{.class} files) and then run the above command in Chord's \code{main/} directory
-with \code{<...>} replaced by \code{examples/datarace_test/}.
+First run {\tt ant} in that directory (in order to compile the program's {\tt .java} files to
+{\tt .class} files) and then run the above command in Chord's {\tt main/} directory
+with {\tt <...>} replaced by \code{examples/datarace_test/}.
 Upon successful completion, the following files should be produced in directory

### 2010-08-14 main/doc/download_and_installation.tex
updated documentation

\section{Download and Installation}
\label{sec:download-and-installation}

Ensure that the following software is installed on your machine:

\begin{itemize}
\item
JDK 6 or higher, e.g. from \htmladdnormallink{IBM}{http://www.ibm.com/developerworks/java/jdk/} or
\htmladdnormallink{Sun}{http://java.sun.com/javase/downloads/index.jsp}
\item
\htmladdnormallink{Apache Ant}{http://ant.apache.org/}, a Java-based build tool
\item
a C++ compiler, e.g. \htmladdnormallink{GCC}{http://gcc.gnu.org/}
\item
a Make utility, e.g. \htmladdnormallink{GNU Make}{http://www.gnu.org/software/make/}
\item
\htmladdnormallink{Cygwin}{http://www.cygwin.com/}, if it is a Windows machine
\end{itemize}

Download a \Chord\ source release from \url{http://code.google.com/p/jchord/downloads/list} or
the latest source from the SVN repository at \url{http://code.google.com/p/jchord/source/checkout}.

Directory {\tt main/} contains a {\tt build.xml} file which is interpreted by Apache Ant.
To see the various targets available, run the following command in that directory:

\begin{quote}
\begin{verbatim}
prompt> ant
\end{verbatim}
\end{quote}

\noindent To compile \Chord, run the following command in the same directory:

\begin{quote}
\begin{verbatim}
prompt> ant compile
\end{verbatim}
\end{quote}

\noindent This will compile the following:
\begin{itemize}
\item
the Java source code of \Chord\ from \code{main/src/java/} (and from \code{extra/src/java/} if that directory is present) to Java bytecode in \code{main/classes/}.
\item
the C source code of BDD library \htmladdnormallink{BuDDy}{http://buddy.sourceforge.net/} from \code{main/src/bdd/} to a shared library in \code{main/lib/} (\code{libbuddy.so} on Linux, \code{buddy.dll} on Windows, and \code{libbuddy.dylib} on MacOS); this library is needed for executing program analyses written in Datalog using \htmladdnormallink{bddbddb}{http://bddbddb.sourceforge.net/} (a BDD-based Datalog solver).
\item
the C++ source code of the Chord instrumentation agent from \code{main/src/agent/} to a shared library in \code{main/lib/} (\code{libchord_instr_agent.so} on all architectures); this agent is needed for computing program scope dynamically and for executing dynamic program analyses.
\end{itemize}

### 2010-08-14 main/doc/program_setup.tex
updated documentation

@@
 
-\noindent The above structure is typical: the program's Java source files are under \code{src/},
-its class files are under \code{classes/}, the source and jar files of the libraries
+\noindent The above structure is typical: the program's Java source files are under {\tt src/},
+its class files are under {\tt classes/}, the source and jar files of the libraries
 used by the program are under \code{lib/src/} and \code{lib/jar/}, respectively.
@@
 An additional use of this property is to allow specifying the command-line arguments to use
-in the run having ID \code{<id>} via property \code{chord.args.<id>}, as illustrated by
+in the run having ID {\tt <id>} via property \code{chord.args.<id>}, as illustrated by
 properties \code{chord.args.0} and \code{chord.args.1} in the above example.
@@
 
-To run Chord on the above program, run the following command in Chord's \code{main/} directory:
+To run Chord on the above program, run the following command in Chord's {\tt main/} directory:
 
@@
 
-\noindent where \code{<...>} denotes either the absolute path, or the path relative to Chord's \code{main/} directory,
+\noindent where {\tt <...>} denotes either the absolute path, or the path relative to Chord's {\tt main/} directory,
 of the parent of directory {\tt example/}.
@@
 All tasks are summarized in Section \ref{sec:chord-sysprops:task}.
-The two most common tasks, described next, are building the analysis scope of the given program (Section \ref{sec:building-scope})
-and running program analyses on the given program (Section \ref{sec:running-analysis}).
+The two most common tasks, described next, are computing the scope of the given program (Section \ref{sec:building-scope})
+and running analyses on the given program (Section \ref{sec:running-analysis}).

### 2010-08-14 main/doc/troubleshooting.tex
updated documentation

\section{Troubleshooting}
\label{sec:troubleshooting}

This section describes known common problems with running \Chord\ and how to fix or work around them.
We assume that the following command was run in \Chord's {\tt main/} directory:

\begin{quote}
\begin{verbatim}
prompt> ant -Dchord.work.dir=<...> <sys-props> run
\end{verbatim}
\end{quote}

\noindent where {\tt <sys-props>} denotes a whitespace-separated list of zero of more property definitions
of the form {\tt -D<key>=<value>} (see Section \ref{sec:chord-sysprops} for an exhaustive listing of all
properties recognized by \Chord).

Abnormal termination of the above command typically produces output of the form:

\begin{quote}
\begin{verbatim}
Buildfile: build.xml

props-file:

pathconvert:

run:
     [java] java version "1.6.0"
     [java] ... // JVM version details omitted
     [java] 
     [java] Redirecting stdout to file: <...>/chord_output/log.txt
     [java] Redirecting stderr to file: <...>/chord_output/log.txt
     [java] Java Result: 1

BUILD FAILED
<...>/main/build.xml:325: Chord terminated abnormally; see above file(s) for details.
\end{verbatim}
\end{quote}

The first step is to inspect the file(s) reported above to which stdout and stderr were redirected.
We next list common errors and how to fix or work around them.

\begin{enumerate}
\item
\end{enumerate}

### 2010-08-14 main/doc/tuning_datalog_analysis.tex
updated documentation

\section{Tuning a Datalog Analysis}
\label{sec:tuning-datalog-analysis}

There are several tricks analysis writers can try to improve the
performance of \bddbddb, the Datalog solver used by \Chord, often
by several orders of magnitude:

\begin{enumerate}
\item
Set system properties \verb+noisy=yes+, \verb+tracesolve=yes+, and \verb+fulltracesolve=yes+
while running \bddbddb\ and observe which rule gets ``stuck" (i.e., takes several seconds to solve).
\verb+fulltracesolve+ is seldom useful, but \verb+noisy+ and \verb+tracesolve+ is
often very useful.  Once you identify the rule that is getting stuck, it
will also tell you which relations and which domains used in that rule
and which operation on them is taking a long time to solve.  Then try
to fix the slowdown problem with that rule by either simplifying the
relations involved (e.g., breaking their set of attributes into two if
possible) or changing the relative order of the domains of those
relations in the BDD variable order (doing the latter alone
frequently causes the problem to go away).

\item
Once you have ensured that none of the rules is getting ``stuck",
you will notice that some rules are applied too many times, and so
although each application of the rule itself isn't taking too much
time, the cumulative time for the rule is too much.  After finishing
solving a Datalog analysis, \bddbddb\ prints how long each rule took to
solve (both in terms of the number of times it was applied and the
cumulative time it took).  It sorts the rules in the order of the
cumulative time.  You need to focus on the rules that took the most
time to solve (they will be at the bottom of the list).  Assuming you
removed the problem of rules getting ``stuck", the rules will roughly
be in the order of the number of times they were applied.  Here is an
example:

\begin{verbatim}
OUT> Rule VH(u:V0,h:H0) :- VV(u:V0,v:V1), VH(v:V1,h:H0), VHfilter
(u:V0,h:H0).
OUT>    Updates: 2871
OUT>    Time: 6798 ms
OUT>    Longest Iteration: 0 (0 ms)
OUT> Rule IM(i:I0,m:M0) :- reachableI(i:I0), specIMV(i:I0,m:M0,v:V0), VH(v:V0,_:H0).
OUT>    Updates: 5031
OUT>    Time: 6972 ms
OUT>    Longest Iteration: 0 (0 ms)
\end{verbatim}

Notice that the second rule was applied 5031 times whereas the first
was applied 2871 times.  More importantly, the second rule took 6972
milliseconds in all, compared to 6798 for the first rule.  Hence, you
should focus on the second rule first, and try to speed it up.  This
means that you should focus only on relations IM, reachableI, specIMV,
and VH, and the domains I0, M0, V0, and H0.  Any changes you make that
do not affect these relations and domains are unlikely to make your
solving faster.  In general, look at the last few rules, not just the
last one, and try to identify the "sub-program" of the Datalog program
that seems problematic, and then focus on speeding up just that sub-
program.

\item
Tweak the BDD variable ordering in the Datalog file (the line that
starts with \verb+bddvarorder+).  This is one of the most
effective ways of making \bddbddb\ run faster.  As explained above, use
\verb+noisy+ and \verb+tracesolve+ to find out which pairs of domains in the
bddvarorder are responsible for the slowdown, and try to change their
relative ordering (note that you can use either `\_' or `x' between a pair
of domains, and the latter is commutative).

\item
You can add the \verb+.split+ keyword at the end of certain rules as a
hint to \bddbddb\ to break up those rules into simpler ones that can be
solved faster.  You can also set property \verb+split_all_rules=yes+ as shorthand
for splitting all rules without adding the \verb+.split+ keyword to any of
them, though I seldom find splitting all rules helpful.

\item
You can yourself break down rules by creating intermediate relations (the more
relations you have on the RHS of a rule the slower it takes to solve
that rule).  This is another very effective way to make \bddbddb\ run
faster.

\item
Try breaking down a single Datalog program into two programs.  Of
course, you cannot separate mutually-recursive rules into two
different programs, but if you unnecessarily club
together rules that could have gone into different programs, then they
can put conflicting demands on \bddbddb\ (e.g., on \verb+bddvarorder+).
So if rule 2 uses the result of rule 1 and rule 1 does not use the result of
rule 2, then put rule 1 and rule 2 in two different Datalog programs.

\item
Observe the sizes of the BDDs representing the relations that are
input and output.  \bddbddb\ tells both the number of tuples in each
relation and the number of "nodes" in the BDD.  Try changing the
bddvarorder for the domains of the relation, and observe how the
number of "nodes" in the bdd for that relation change.  You will
notice that some orderings do remarkably better than others.  Then
note down these orderings as invariants that you will not violate as
you tweak other things.

\item
The relative ordering of values *within* domains (e.g.
in domains named \verb+M+, \verb+H+, \verb+C+, etc. in Chord) affects the
solving time of \bddbddb, but
I've never tried changing this and studying its effect.  It might be
worth trying.  For instance, John Whaley's PLDI'04 paper describes a
specific way in which he numbers contexts (in domain \verb+C+) and that it was
fundamental to the speedup of his infinity-CFA points-to analysis.

\item
Finally, it is worth emphasizing that BDDs are not magic.
If your algorithm itself is fundamentally hard to scale, then BDDs are
unlikely to help you a whole lot.  Secondly, many things are awkward to
encode as integers (e.g., the abstract contexts in the domain \verb+C+ 
in Chord) or as Datalog rules.
For instance, I've noticed that summary-based context-sensitive program
analyses are hard to express in Datalog.  The may-happen-in-parallel
analysis provided in Chord shows a relatively simple kind of summary-based
analysis that uses the Reps-Horwitz-Sagiv tabulation algorithm.  But this
is as far as I could get---more complicated summary-based algorithms are
best written in Java itself instead of Datalog.
\end{enumerate}

### 2010-08-14 main/doc/writing_datalog_analysis.tex
updated documentation

\section{Writing a Datalog Analysis}
\label{sec:writing-datalog-analysis}

Chord interprets each file with extension {\tt .dlog} or {\tt .datalog} in the path specified by
property \code{chord.dlog.analysis.path} as a program analysis expressed in Datalog.
A Datalog analysis declares a bunch of input/output program relations,
each over one or more program domains, and provides a bunch of rules (constraints)
specifying how to compute the output relations from the input relations.
An example of such an analysis is as follows:

\begin{quote}
\begin{verbatim}
# name=datarace-dlog

# Program domains
.include "E.dom"
.include "F.dom"
.include "T.dom"

# BDD variable order
.bddvarorder E0xE1_T0_T1_F0

# Input/intermediate/output program relations
field(e:E0,f:F0) input
write(e:E0) input
reach(t:T0,e:E0) input
alias(e1:E0,e2:E1) input
escape(e:E0) input
unguarded(t1:T0,e1:E0,t2:T1,e2:E1) input
hasWrite(e1:E0,e2:E1)
candidate(e1:E0,e2:E1) 
datarace(t1:T0,e1:E0, t2:T1,e2:E1) output

# Analysis constraints
hasWrite(e1,e2) :- write(e1).
hasWrite(e1,e2) :- write(e2).
candidate(e1,e2) :- field(e1,f), field(e2,f), hasWrite(e1,e2), e1 <= e2.
datarace(t1,e1,t2,e2) :- candidate(e1,e2), reach(t1,e1), reach(t2,e2), \
    alias(e1,e2), escape(e1), escape(e2), unguarded(t1,e1,t2,e2).
\end{verbatim}
\end{quote}

Any line that begins with a {\tt \#} is regarded a comment, except a
line of the form {\tt \#} {\tt name=<...>}, which specifies the name
{\tt <...>} of the Datalog analysis.
Each Datalog analysis is expected to have exactly one such line.
The above Datalog analysis is named {\tt datarace-dlog}.
In Chord, all Datalog analysis names have suffix
{\tt -dlog} and all Java analysis names have suffix {\tt -java}, but
this is merely a convention that user-defined analyses are free to deviate from.
The name of each analysis, written in Datalog or in Java, is expected to be
unique across all analyses in scope (i.e., across all analyses in paths
specified by properties \code{chord.dlog.analysis.path} and \code{chord.java.analysis.path}).

The {\tt .include "<...>.dom"} lines specify each program domain named
{\tt <...>} that is needed by the Datalog analysis, i.e., each domain over which
any program relation that is input/output by the Datalog analysis is defined. 
The declaration of each such relation specifies the domain of each of the relation's attributes.
If the same domain appears in multiple attributes of a relation then
contiguous integers starting from 0 must be used to distinguish them; for instance,
in the above example, {\tt candidate} is a binary relation, both of whose
attributes have domain E, and they are distinguished as E0 and E1.

Each relation is represented symbolically (as opposed to explicitly)
using a graph-based data structure called a Binary Decision Diagram (BDD for short).
Each domain containing N elements is assigned log2(N) BDD variables.
The size of a BDD and the efficiency of operations on it depends heavily
on the order of these BDD variables.
The {\tt .bddvarorder <...>} line in the Datalog analysis enables the Datalog
analysis writer to specify this order.
It must list all domains along with their numerical suffixes, separated
by {\tt \_} or {\tt x}.
Using a {\tt \_} between two domains, such as {\tt T0\_T1}, means that the BDD variables assigned
to domain {\tt T0} precede those assigned to domain {\tt T1} in the BDD variable
order for this Datalog analysis.
Using a {\tt x} between two domains, such as {\tt E0xE1}, means that the
BDD variables assigned to domains {\tt E0} and {\tt E1}
will be interleaved in the BDD variable order for this Datalog analysis.
See Section \ref{sec:bdd-representation} for more details on BDD representations.

Each Datalog analysis rule is a Horn clause of the form
{\tt R(t) :- R1(t1), ..., Rn(tn)}
meaning that if relations {\tt R1}, ..., {\tt Rn} contain tuples {\tt t1}, ..., {\tt tn}
respectively, then relation {\tt R} contains tuple {\tt t}.
A backslash may be used at the end of a line to break long rules for readability.
The Datalog analysis solver \bddbddb\ used in \Chord\ does not apply any
sophisticated optimizations to simplify the rules; besides the BDD variable order,
the manner in which these rules are expressed heavily affects the performance of
the solver.  For instance, an important manual optimization involves breaking down
long rules into multiple shorter rules communicating via intermediate relations.
See Section \ref{sec:tuning-datalog-analysis} for hints on tuning the performance
of Datalog analyses.

### 2010-08-14 main/doc/writing_dynamic_analysis.tex
updated documentation

@@
-\section{Writing a Dynamic Program Analysis}
+\section{Writing a Dynamic Analysis}
 \label{sec:writing-dynamic-analysis}
@@
 which must return an instance of the instrumentation scheme chosen in Step 1 above,
-plus each \code{process<event>(<args>)} method that corresponds to event \code{<event>}
-with format \code{<args>} enabled by the chosen instrumentation scheme.
+plus each \code{process<event>(<args>)} method that corresponds to event {\tt <event>}
+with format {\tt <args>} enabled by the chosen instrumentation scheme.
 All other methods are no-ops if not overridden.

### 2010-08-15 main/doc/chord_sysprops.tex
minor (added documentation + fixed javadoc tags)

@@
 \subsection{Program Scope Properties}
+\label{sec:scope-props}
 
@@
 \code{chord.scope.kind} \\
-{\bf Type:} [\code{dynamic}$|$\code{rta}$|$\code{cha}] \\
+{\bf Type:} {\tt [dynamic|rta|cha]} \\
 {\bf Description:} Algorithm to compute analysis scope (i.e., reachable classes and methods).
@@
 \code{chord.ch.kind} \\
-{\bf Type:} [\code{static}$|$\code{dynamic}] \\
+{\bf Type:} {\tt [static|dynamic]} \\
 {\bf Description:} Algorithm to build the class hierarchy.
@@
 \subsection{Chord Instrumentation Properties}
+\label{sec:instr-props}
 
@@
 \code{chord.trace.kind} \\
-{\bf Type:} [\code{none}$|$\code{full}$|$\code{pipe}] \\
+{\bf Type:} {\tt [none|full|pipe]} \\
 {\bf Description:} The medium to store dynamic traces (none, regular file, or POSIX pipe). \\

### 2010-08-15 main/doc/deadlock.tex
minor (added documentation + fixed javadoc tags)

@@
 
-Run the deadlock analysis provided in Chord by running the following command in Chord's \verb+main/+ directory:
+Run the deadlock analysis provided in Chord by running the following command in Chord's {\tt main/} directory:
 
-\begin{quote}
 \begin{verbatim}
@@
 \end{verbatim}
-\end{quote}
 
-where directory \verb+<...>+ contains a file named \verb+chord.properties+ which 
-defines properties \verb+chord.main.class+, \verb+chord.class.path+, and \verb+chord.src.path+.
+\noindent where directory {\tt <...>} contains a file named \code{chord.properties} which
+defines properties \code{chord.main.class}, \code{chord.class.path}, and \code{chord.src.path}.
 See Section \ref{sec:chord-sysprops} for the meaning of these properties.
 
-Directory \verb+main/examples/deadlock_test/+ provides a toy Java program on which you can run the deadlock
-analysis by first running \verb+ant+ in that directory (in order to compile the program's \verb+.java+ files to
-\verb+.class+ files) and then running the following command in Chord's \verb+main/+ directory:
-
-\begin{quote}
-\begin{verbatim}
-prompt> ant -Dchord.work.dir=examples/deadlock_test/ -Dchord.run.analyses=deadlock-java run
-\end{verbatim}
-\end{quote}
-
-Upon successful completion, the file \verb+deadlocks.html+ should be produced in directory
-\verb+main/examples/deadlock_test/chord_output/+.
+Directory \code{main/examples/deadlock_test/} provides a toy Java program on which you can run
+the deadlock analysis.
+First run {\tt ant} in that directory (in order to compile the program's {\tt .java} files to
+{\tt .class} files) and then run the above command in Chord's {\tt main/} directory with
+{\tt <...>} replaced by \code{examples/deadlock_test/}.
+Upon successful completion, the file \code{deadlocks.html} should be produced in directory
+\code{main/examples/deadlock_test/chord_output/}.

### 2010-08-15 main/src/java/chord/analyses/thread/DomA.java

@@ -35,7 +35,7 @@ import chord.doms.DomM;
  * <tt>java.lang.Thread</tt> and <tt>c</tt> is an abstract context
  * of that method.
  * 
- * @see chord.analyses.thread.ForkCtxtsAnalysis
+ * @see chord.analyses.thread.ThreadsAnalysis
  * 
  * @author Mayur Naik (mhn@cs.stanford.edu)
  */

### 2010-08-31 main/doc/chord_sysprops.tex
documentation changes

@@
 {\bf Note:} \Chord\ analyzes only Java bytecode, not Java source code.
-This property is used only by the task of converting Java source files into HTML files by program analyses that need to present their analysis results at the Java source code level (by calling method \code{chord.program.Program.getProgram().HTMLizeJavaSrcFiles())}.
+This property is used only by the task of converting Java source files into HTML files by program analyses that need to present their analysis results at the Java source code level (by calling method \code{chord.program.Program.g().HTMLizeJavaSrcFiles())}.
 
@@
 {\bf Default value:} {\tt false} \\
-{\bf Note:} The program scope is computed regardless of the value of this property if another task (e.g., a program analysis specified via property \code{chord.run.analyses}) demands it by calling method \code{chord.program.Program.getProgram()}.
+{\bf Note:} The program scope is computed regardless of the value of this property if another task (e.g., a program analysis specified via property \code{chord.run.analyses}) demands it by calling method \code{chord.program.Program.g()}.
 
@@
 \code{chord.verbose} \\
-{\bf Type:} bool \\
-{\bf Description:} Produce more verbose output during \Chord's execution. \\
-{\bf Default value:} {\tt false}
+{\bf Type:} int in the range [0..3] \\
+{\bf Description:} Control the verbosity of logging during \Chord's execution. \\
+{\bf Default value:} {\tt 2}
 
 \item
-\code{chord.bddbddb.noisy} \\
-{\bf Type:} bool \\
-{\bf Description:} Produce more verbose output during the Datalog solver \bddbddb's execution. \\
-{\bf Default value:} {\tt false}
+\code{chord.bddbddb.verbose} \\
+{\bf Type:} int in the range [0..5] \\
+{\bf Description:} Control the verbosity of logging during the execution of JavaBDD (BDD library) and \bddbddb\ (Datalog solver). \\
+{\bf Default value:} {\tt 0}
 

### 2010-09-01 main/doc/chord_sysprops.tex
replaced chord.publish.results -> chord.print.results and chord.publish.targets -> chord.print.project

@@
 \item
-\code{chord.publish.targets} \\
+\code{chord.print.project} \\
 {\bf Type:} bool \\
-{\bf Description:} Create files \code{targets_sortby_name.html}, \code{targets_sortby_kind.html}, and \code{targets_sortby_producers.html} in directory \code{[chord.out.dir]}, publishing all targets defined by program analyses in paths \code{[chord.java.analysis.path]} and \code{[chord.dlog.analysis.path]}.
+{\bf Description:} Create files \code{targets_sortby_name.html}, \code{targets_sortby_kind.html}, and \code{targets_sortby_producers.html} in directory \code{[chord.out.dir]}, publishing all tasks and targets defined by program analyses in paths \code{[chord.java.analysis.path]} and \code{[chord.dlog.analysis.path]}.
 \\
@@
 \item
-\code{chord.publish.results} \\
+\code{chord.print.results} \\
 {\bf Type:} bool \\
-{\bf Description:} Publish the results of program analyses in HTML.  Interpretation of this property is analysis-specific. \\
+{\bf Description:} Print the results of program analyses in HTML.  Interpretation of this property is analysis-specific. \\
 {\bf Default value:} {\tt true}
@@
 {\bf Type:} string \\
-{\bf Description:} URL of the Javadoc location of program analyses.  It is used when publishing targets (i.e., when property \code{chord.publish.targets} is set to {\tt true}). \\
+{\bf Description:} URL of the Javadoc location of program analyses.  It is used when printing the project (i.e., when property \code{chord.print.project} is set to {\tt true}). \\
 {\bf Default value:} \code{http://jchord.googlecode.com/svn/wiki/javadoc/}

### 2010-09-06 main/doc/chord_sysprops.tex
minor
added new property chord.analysis.exclude to exclude specified tasks; also improved error messages regarding ambiguous tasks in both classic and modern projects

@@
 \item
+\code{chord.analysis.exclude} \\
+{\bf Type:} string list \\
+{\bf Descrption:} Names of program analyses written in Java or Datalog to be excluded from the project. \\
+{\bf Default value:} {\tt ""}
+
+\item
 \code{chord.reuse.rels} \\
@@
 {\bf Default value:} -1
-
-\item
-\code{chord.dynamic.silent} \\
-{\bf Type:} bool \\
-{\bf Description:} Whether or not to emit warnings while running the given program during dynamic analysis. \\
-{\bf Default value:} false
 \end{itemize}

### 2010-09-06 main/src/java/chord/project/TaskParser.java

@@ -180,6 +180,10 @@ public class TaskParser {
 			if (Config.verbose > 2) Messages.log(ANON_JAVA_TASK, className);
 			name = className;
 		}
+		for (String name2 : Config.analysisExcludeAry) {
+			if (name.equals(name2))
+				return;
+		}
 		DlogAnalysis dlogTask = nameToDlogTaskMap.get(name);
 		if (dlogTask != null) {
 			redefinedJavaTask(className, name, dlogTask.getFileName());
@@ -250,6 +254,10 @@ public class TaskParser {
 		if (name == null) {
 			if (Config.verbose > 2) Messages.log(ANON_DLOG_TASK, fileName);
 			name = fileName;
+		}
+		for (String name2 : Config.analysisExcludeAry) {
+			if (name.equals(name2))
+				return;
 		}
 		DlogAnalysis dlogTask = nameToDlogTaskMap.get(name);
 		if (dlogTask != null) {

### 2010-10-27 main/doc/running_analysis.tex
write some more documentation

@@
 
-TODO
+\subsection{Running an analysis}
+Running a Chord analysis is straightforward.
+The property  \texttt{chord.run.analyses} is a list of analyses to be run.
+Chord will compute the listed dependencies for each of these relations and run these prerequisites before running the analysis you specify.
 
-%chord.run.analyses
+\subsection{The dependence graph in classic projects}
+
+Chord has two kinds of projects: ``Classic'' and ``Modern''.  Despite the names, classic is effectively the standard at this point.
+In the classic model, there is an implicit dependency graph between analyses.
+For a datalog analysis (program), the ``input'' and ``output'' keywords in defining a relation specify the requirements and outputs of an analysis.
+For a Java analysis, the \texttt{@Chord} annotation does so.
+See the section ``writing a Java analysis'' for details.
+
+\subsection{Printing the results}
+
+\texttt{chord.print.rels} is a list of relations to dump to disk.
+Note that these relations are printed in a somewhat unfriendly format; this is more useful for debugging than for interpreting programs.
+Note also that listing a relation here does not cause it to be computed; you will get an error at the end of your Chord execution if you ask it to print a relation that has not been computed.
+
 %List of names of program analyses to be run in order.
@@
 %chord.reuse.rels	 bool	 Construct program relations from BDDs stored on disk (from a previous run of Chord) whenever possible instead of re-computing them.	false
-%chord.publish.results	 bool	 Publish the results of program analyses in HTML.	true
+%chord.print.results	 bool	 Print the results of program analyses in HTML.	true

### 2010-11-30 main/doc/writing_dynamic_analysis.tex
minor documentation fixes

@@
 
-{\bf Step 1: Setting up the Instrumentation Scheme}
+\subsection{Implementing the Analysis}
 
-Determine the instrumentation scheme required by your dynamic analysis, that is,
-the kind and format of events to be generated during an instrumented program's
-execution. See Section \ref{sec:instr-events} 
-for the kinds of supported events and their formats.
-
-{\bf Step 2: Implementing the Analysis}
-
-Create a class extending \code{chord.project.DynamicAnalysis} and override
+Create a class extending \code{chord.project.analyses.DynamicAnalysis} and override
 the appropriate methods in it.
-The only method that must be compulsorily overridden is method \code{getInstrScheme()},
-which must return an instance of the instrumentation scheme chosen in Step 1 above,
+The only methods that must be compulsorily overridden are method \code{getInstrScheme()},
+which must return an instance of the ``instrumentation scheme" required by
+your dynamic analysis (i.e., the kind and format of events to be generated during an
+instrumented program's execution)
 plus each \code{process<event>(<args>)} method that corresponds to event {\tt <event>}
 with format {\tt <args>} enabled by the chosen instrumentation scheme.
-All other methods are no-ops if not overridden.
+See Section \ref{sec:instr-events} for the kinds of supported events and their formats.
 
-A sample such class called \code{MyDynamicAnalysis} is shown below:
+A sample such class \code{MyDynamicAnalysis} is shown below:
 
 \begin{verbatim}
-    import chord.project.DynamicAnalysis;
+    import chord.project.analyses.DynamicAnalysis;
     import chord.instr.InstrScheme;
@@
     @Chord(name=<name-of-analysis>)    
-
     public class MyDynamicAnalysis extends DynamicAnalysis {
@@
 
+        @Override
         public InstrScheme getInstrScheme() {
@@
 
+        @Override
         public void initAllPasses() {
@@
 
+        @Override
         public void doneAllPasses() {
@@
 
+        @Override
         public void initPass() {
@@
 
+        @Override
         public void donePass() {
@@
         // No-ops if not overridden.
+        @Override
         public void process<event1>(<args1>) {
@@
         ...
+        @Override
         public void process<eventN>(<argsN>) {
@@
 
-{\bf Step 3: Choosing or Implementing the Runtime Event Handler}
+\subsection{Configuring the Analysis}
 
-Determine the runtime event handler for your dynamic analysis. There are two kinds of runtime event
-handlers, offline and online; see package \code{chord.runtime} for more details.
+You can change the default values of various properties for configuring your dynamic analysis;
+see Sections \ref{sec:scope-props} and \ref{sec:instr-props}.
+For instance:
+\begin{itemize}
+\item
+You can set property \code{chord.scope.kind} to {\tt dynamic} so that the program scope
+is computed dynamically (i.e., by running the program) instead of statically.
+\item
+You can exclude certain classes (e.g., JDK classes) from being instrumented by setting
+properties \code{chord.std.scope.exclude}, \code{chord.ext.scope.exclude}, and
+\code{chord.scope.exclude}.
+\item
+You can choose between online (i.e. load-time) and offline bytecode instrumentation by
+setting property \code{chord.instr.kind} to {\tt online} or {\tt offline}.
+\item
+You can require the event-generating and event-handling JVMs to be one and the same (by setting
+property \code{chord.trace.kind} to {\tt none}),
+or to be separate (by setting property {\tt chord.trace.kind} to {\tt full} or {\tt pipe}, depending upon
+whether you want the two JVMs to exchange events by a regular file or a POSIX pipe, respectively).
+Using a single JVM can cause correctness/performance issues if event-handling Java code itself is instrumented
+(e.g., say the event-handling code uses class \code{java.util.ArrayList} which is not excluded from program scope).
+Using separate JVMs prevents such issues since the event-handling JVM runs uninstrumented bytecode (only
+the event-generating JVM runs instrumented bytecode).
+If a regular file is used to exchange events between the two JVMs, then the JVMs run serially:
+the event-generating JVM first runs to completion, dumps the entire dynamic trace to the regular file,
+and then the event-handling JVM processes the dynamic trace.
+If a POSIX pipe is used to exchange events between the two JVMs, then the JVMs run in lockstep.
+Obviously, a pipe is more efficient for very long traces, but it 
+not portable (e.g., it does not currently work on Windows/Cygwin), and the traces cannot be reused
+across Chord runs (see the following item).
+\item
+You can reuse dynamic traces from a previous \Chord\ run by setting property
+\code{chord.reuse.traces} to {\tt true}.  In this case, you must also set property
+{\tt chord.trace.kind} to {\tt full}.
+\item
+You can set property \code{chord.dynamic.haltonerr} to {\tt false} to prevent \Chord\ from terminating
+even if the program on which dynamic analysis is being performed crashes.
+\end{itemize}
 
-For most dynamic analyses, an offline event hander suffices, in which case you don't need to do
-anything in this step.
-Otherwise, you need to create a class extending \code{chord.runtime.Runtime} and override the
-appropriate methods in it, and provide the fully-qualified name of the created class as the value of property
-\code{chord.runtime.class} (whose default value is \code{chord.runtime.BufferedRuntime}).
+\Chord\ offers much more flexibility in crafting dynamic analyses.
+You can define your own instrumentor (by subclassing \code{chord.instr.CoreInstrumentor}
+instead of using the default \code{chord.instr.Instrumentor}) and your own event handler (by subclassing \code{chord.runtime.CoreEventHandler}
+instead of using the default \code{chord.runtime.EventHandler}).
+You can ask the dynamic analysis to use your custom instrumentor and/or your custom event handler by overriding
+methods \code{getInstrumentor()} and \code{getEventHandler()}, respectively, defined in \code{chord.project.analyses.CoreDynamicAnalysis}.
+Finally, you can define your own dynamic analysis template by subclassing \code{chord.project.analyses.CoreDynamicAnalysis}
+instead of subclassing the default \code{chord.project.analyses.DynamicAnalysis}.
 
-{\bf Step 4: Compiling the Analysis}
+\subsection{Compiling and Running the Analysis}
 
@@
 above in the path defined by property \code{chord.java.analysis.path}.
-
-{\bf Step 5: Specifying Program Inputs}
 
@@
 
-{\bf Step 6: Running the Analysis}
-
 To run the analysis, set property \code{chord.run.analyses} to \code{<name-of-analysis>}
@@
 \code{MyDynamicAnalysis} created above).
-
-%(except that you may want to change the default values of properties
-%\code{chord.trace.pipe} and \code{chord.trace.block.size} for performance, or the default values
-%of properties \code{chord.trace.pipe} and \code{chord.reuse.trace} for determinism across; see
-%Section \ref{sec:chord-sysprops} for the meaning of these properties).

### 2010-12-13 main/doc/building_scope.tex
updated documentation

--- 2010-04-09 main/doc/building_scope.tex
@@
-\section{Building Analysis Scope}
+\section{Computing Program Scope}
 \label{sec:building-scope}
 
-Chord computes the analysis scope (i.e., reachable classes and methods) of the given program
-either if property \code{chord.build.scope} is set to \code{true} or if some other task (e.g.,
-a program analysis specified via property \code{chord.run.analyses}) demands it by
-calling method \code{chord.program.Program.v()}.
+\Chord\ computes the scope (i.e., reachable methods) of the given
+program either if property \code{chord.build.scope} is set to {\tt
+  true} or if some other task (e.g., a program analysis specified via
+property \code{chord.run.analyses}) demands it by calling method
+\code{chord.program.Program.g()}.
 
-The algorithm used to compute the analysis scope is as follows.
+Users can set the following properties to control scope computation:
+
+\begin{itemize}
+\item \code{chord.main.class}
+\item \code{chord.class.path}
+\item \code{chord.reuse.scope}
+\item \code{chord.methods.file}
+\item \code{chord.reflect.file}
+\item \code{chord.scope.kind}
+\item \code{chord.ch.kind}
+\item \code{chord.reflect.kind}
+\item \code{chord.std.scope.exclude}
+\item \code{chord.ext.scope.exclude}
+\item \code{chord.scope.exclude}
+\end{itemize}
+
+\subsection{Reusing Program Scope}
+
+If property \code{chord.reuse.scope} has value {\tt true} and both
+files specified by properties \code{chord.methods.file} and
+\code{chord.reflect.file} exist, then \Chord\ regards those files as
+specifying which methods to consider reachable and how to resolve
+reflection, respectively.
+
+The format of the file specified by property \code{chord.methods.file}
+is a list of zero or more lines, where each line is of the form
+\code{mname:mdesc@cname}
+specifying the method's name {\tt mname}, the method's descriptor
+{\tt mdesc}, and the method's declaring class {\tt cname} (e.g.,
+\code{main:([Ljava/lang/String;)V@foo.bar.Main}).
+
+The format of the file specified by property \code{chord.reflect.file}
+is of the form:
+\begin{verbatim}
+# resolvedClsForNameSites
+...
+# resolvedObjNewInstSites
+...
+# resolvedConNewInstSites
+...
+# resolvedAryNewInstSites
+...
+\end{verbatim}
+where each of the above ``{\tt ...}'' is a list of zero or more lines, where
+each line is of the form
+\code{bci!mname:mdesc@cname->type1,type2,...typeN}
+meaning the call site at bytecode offset {\tt bci} in the
+method denoted by {\tt mname:mdesc@cname} may resolve to any of
+reference types {\tt type1}, {\tt type2}, ..., {\tt typeN}.
+The meaning of the above four sections is as follows.
+\begin{itemize}
+\item {\tt resolvedClsForNameSites} lists
+each call to static method {\tt forName(String)} defined in class
+{\tt java.lang.Class}, along with a list of the types of the named
+classes.
+\item {\tt resolvedObjNewInstSites} lists
+each call to instance method {\tt newInstance()} defined in class
+{\tt java.lang.Class}, along with a list of the types of the
+instantiated classes.
+\item {\tt resolvedConNewInstSites} lists
+each call to instance method {\tt newInstance(Object[])} defined in class
+{\tt java.lang.reflect.Constructor}, along with a list of the types of the
+instantiated classes.
+\item {\tt resolvedAryNewInstSites} lists
+each call to instance method {\tt newInstance(Class,int)} defined in class
+{\tt java.lang.reflect.Array}, along with a list of the types of the
+instantiated classes.
+\end{itemize}
+The default value of property \code{chord.reuse.scope} is {\tt false}.
+The default value of properties \code{chord.methods.file} and
+\code{chord.reflect.file} is \code{[chord.out.dir]/methods.txt} and
+\code{[chord.out.dir]/reflect.txt}, respectively.
+Property \code{chord.out.dir} defaults to
+\code{[chord.work.dir]/chord_output/} and property
+\code{chord.work.dir} defaults to the working directory during Chord's
+execution.
+
+\subsection{Scope Construction Algorithms}
+
+If property \code{chord.reuse.scope} has value {\tt false} or the
+files specified by properties \code{chord.methods.file} or
+\code{chord.reflect.file} do not exist, then \Chord\ computes program
+scope using the algorithm specified by property
+\code{chord.scope.kind} and then writes the list of methods deemed
+reachable and the reflection resolved by that algorithm to the files
+specified by properties \code{chord.methods.file} and
+\code{chord.reflect.file}, respectively.
+
+The possible legal values of property \code{chord.scope.kind} are
+[\code{dynamic}$|$\code{rta}$|$\code{cha}].  In each case, \Chord\ at
+least expects properties \code{chord.main.class} and
+\code{chord.class.path} to be set.
 
@@
 \item
-If property \code{chord.reuse.scope} has value \code{true} and the files specified by properties
-\code{chord.classes.file} and \code{chord.methods.file} exist,
-then Chord regards those files as specifying the classes and methods, respectively,
-to be regarded as reachable.  The format
-of the classes file is a fully-qualified class name per line (e.g., \code{foo.bar.Main}).  The format
-of the methods file is an entry of the form \code{<method name>:<method descriptor>@<class name>} per line,
-specifying the method's name, the method's descriptor, and the method's declaring class
-(e.g., \code{main:([Ljava/lang/String;)V@foo.bar.Main}).
-Chord throws a runtime exception if the declaring class of a method in the methods file is not
-specified in the classes file.
+The {\tt dynamic} value instructs \Chord\ to compute program scope
+dynamically, by running the program and observing using JVMTI the
+classes that are loaded at run-time.  The number of times the program
+is run and the command-line arguments to be supplied to the program in
+each run is specified by properties \code{chord.run.ids} and
+\code{chord.args.<id>} for each run ID {\tt <id>}.  By default, the
+program is run only once, using run ID {\tt 0}, and without any
+command-line arguments.  Only classes loaded in some run are regarded
+as reachable but {\it all} methods of each loaded class are regarded
+as reachable regardless of whether they were invoked in the run.  The
+rationale behind this decision is to both reduce run-time overhead of
+JVMTI and to increase the predictive power of program analyses
+performed using the computed program scope.
 
 \item
-If property \code{chord.reuse.scope} has value \code{false} or the file specified by
-property \code{chord.classes.file} or \code{chord.methods.file} does not exist,
-then Chord computes the analysis scope
-using the algorithm specified by property \code{chord.scope.kind} and then
-writes the classes and methods deemed reachable by that algorithm to those files.
+The {\tt rta} value instructs \Chord\ to compute program scope
+statically using Rapid Type Analysis (RTA).
 
-The possible legal values of property \code{chord.scope.kind} are \code{rta} and \code{dynamic}.
-In both cases, Chord expects properties \code{chord.main.class} and \code{chord.class.path}
-to be set.
+RTA is an iterative fixed-point algorithm.  It maintains a set of
+reachable methods $M$.  The initial iteration starts by assuming that
+only the main method in the main class is reachable (\Chord\ also
+handles class initializer methods but we ignore them here for brevity;
+we also ignore the set of reachable classes maintained besides the set
+of reachable methods).  All object allocation sites $H$ contained in
+methods in $M$ are deemed reachable (i.e., control-flow within method
+bodies is ignored).  Whenever a dynamically-dispatching method call
+site (i.e., an invokevirtual or invokeinterface site) with receiver of
+static type $t$ is encountered in a method in $M$, only subtypes of
+$t$ whose objects are allocated at some site in $H$ are considered to
+determine the possible target methods, and each such target method is
+added to $M$.  The process terminates when no more methods can be
+added.
 
+%RTA is a relatively inexpensive and precise algorithm in practice.  Its key shortcoming is that it makes
+%no attempt to resolve reflection, which is rampant in real-world Java programs, and can therefore be
+%unsound (i.e., underestimate the set of reachable classes and methods).
+%The next option attempts to overcome this problem.
+%
+%\item
+%The \code{rta_reflect} value instructs \Chord\ to compute program scope statically using Rapid Type Analysis
+%and, moreover, to resolve a common reflection pattern:
+%
+%\begin{quote}
+%\begin{verbatim}
+%String s = ...;
+%Class c = Class.forName(s);
+%Object o = c.newInstance();
+%T t = (T) o;
+%\end{verbatim}
+%\end{quote}
+%
+%This analysis is identical to RTA except that it additionally inspects every cast statement in the program,
+%such as the last statement in the above snippet,
+%and queries the class hierarchy to find all concrete classes that subclass \code{T} (if \code{T} is a class)
+%or that implement \code{T} (if \code{T} is an interface).
+%\Chord\ allows users to control which classes are included in the class hierarchy (see below).
+%
+\item
+The {\tt cha} value instructs \Chord\ to compute program scope
+statically using Class Hierarchy Analysis (CHA).
+
+The key difference between CHA and RTA is that for invokevirtual and
+invokeinterface sites with receiver of static type $t$, CHA considers
+{\it all} subtypes of $t$ in the class hierarchy to determine the
+possible target methods, whereas RTA restricts them to types of
+objects allocated in methods deemed reachable so far.  As a result,
+CHA is highly imprecise in practice, and also expensive since it
+grossly overestimates the set of reachable classes and methods.
+Nevertheless, \Chord\ allows users to control which classes are
+included in the class hierarchy (see below) and thereby control the
+precision and cost of CHA.
+
+\end{itemize}
+The default value of property \code{chord.scope.kind} is {\tt rta}.
+
+\subsection{Class Hierarchy Construction}
+
+The class hierarchy is built if property \code{chord.scope.kind} has
+value {\tt cha}.  Users can control which classes are included in
+building the class hierarchy by setting property \code{chord.ch.kind},
+whose possible legal values are [\code{static}$|$\code{dynamic}].
+\Chord\ first constructs the entire classpath of the given program by
+concatenating in order the following classpaths:
+
+\begin{enumerate}
+\item
+The boot classpath, specified by property \code{sun.boot.class.path}.
+\item
+The library extensions classpath, comprising all jar files in
+directory \code{[java.home]/lib/dir/}.
+\item
+\Chord's classpath, specified by property
+\code{chord.main.class.path}.  This includes the classpath of any
+user-defined Java analyses, specified by user-defined property
+\code{chord.java.analysis.path}, which is empty by default.
+\item
+The classpath of the given program, specified by user-defined property
+\code{chord.class.path}, which is empty by default.
+\end{enumerate}
+
+All classes in the entire classpath (resulting from items 1--4 above)
+are included in the class hierarchy with the following exceptions:
 \begin{itemize}
 \item
-The \code{rta} value instructs Chord to compute the analysis scope statically using Rapid Type Analysis.
-
+Duplicate classes, i.e., classes with the same name occurring in more
+than one classpath element; in this case, all occurrences except the
+first are excluded.
 \item
-The \code{dynamic} value instructs Chord to compute the analysis scope dynamically, by running the program
-and observing using JVMTI the classes that are loaded at run-time.
-The number of times the program is run and the command-line arguments to be supplied to
-the program in each run is specified by properties \code{chord.run.ids} and
-\code{chord.args.<id>} for each run ID \code{<id>}.  By default, the program is run only once, using run ID \code{0},
-and without any command-line arguments.
-Only classes loaded in some run are regarded as reachable but {\it all} methods of each loaded class are regarded
-as reachable regardless of whether they were invoked in the run.
-The rationale behind this decision is to both reduce run-time overhead of JVMTI and to increase the
-predictive power of program analyses performed using the computed analysis scope.
-\end{itemize}
+Any class whose name's prefix is specified in the value of property
+\code{chord.scope.exclude}.
+\item
+If property \code{chord.ch.kind} has value {\tt dynamic}, then
+\Chord\ runs the given program and observes the set of all classes the
+JVM loads; any class not in this set is excluded.  By default,
+property \code{chord.ch.kind} has value {\tt static}.
+\item
+If the superclass of a class C is missing or if an interface
+implemented/extended by a class/interface C is missing, where
+``missing" means that it is either not in the classpath resulting from
+items 1--4 above or it is excluded by one of these rules, then C
+itself is excluded.  Note that this rule is recursive, e.g., if C has
+superclass B which in turn has superclass A, and A is missing, then
+both B and C are excluded.
 \end{itemize}
 
-The default value of property \code{chord.reuse.scope} is \code{false} and that of
-property \code{chord.scope.kind} is \code{rta}.
-The default values of properties \code{chord.classes.file} and \code{chord.methods.file} are
-\code{[chord.out.dir]/classes.txt} and \code{[chord.out.dir]/methods.txt}, respectively,
-where property \code{chord.out.dir} defaults to \code{[chord.work.dir]/chord_output/},
-and property \code{chord.work.dir} defaults to the current working directory.
+\subsection{Scope Exclusion}
+
+%TODO: describe chord.scope.exclude?
+
+%\begin{tabular}
+%chord.scope.kind,chord.ch.dynamic & runs program? & \# reachable classes & \# reachable methods & running time \\
+%dynamic,-         & yes &    314 &  4,491 &   25s \\
+%cha,true          & yes &    427 &  1,532 &   28s \\
+%rta,-             &  no &    849 &  4,836 &    9s \\
+%rta_reflect,true  & yes &    849 &  4,836 &   34s \\
+%rta_reflect,false &  no &  9,871 & 58,726 &  7m6s \\
+%cha,false         &  no & 14,121 & 74,613 & 4m11s
+%\end{tabular}

### 2010-12-13 main/doc/chord_sysprops.tex
updated documentation

@@
 {\bf Type:} bool \\
-{\bf Description:} Compute the scope (i.e., reachable classes and methods) of the given program using the algorithm specified by properties \code{chord.scope.kind} and \code{chord.reuse.scope}. \\
+{\bf Description:} Compute the scope of the given program using the algorithm specified by properties \code{chord.scope.kind} and \code{chord.reuse.scope}. \\
 {\bf Default value:} {\tt false} \\
@@
 
-This section describes properties that specify how the analysis scope is computed.
+This section describes properties that specify how the program scope is computed.
 See Section \ref{sec:building-scope} for more details.
@@
 {\bf Type:} bool \\
-{\bf Description:} Treat analysis scope as the classes and methods listed in files specified by properties \code{chord.classes.file} and \code{chord.methods.file}, respectively.
+{\bf Description:} Compute program scope using the information in files specified by properties \code{chord.methods.file} and \code{chord.reflect.file}.
 Property \code{chord.scope.kind} is ignored if this property is set to {\tt true} and the two files exist.
@@
 {\bf Type:} {\tt [dynamic|rta|cha]} \\
-{\bf Description:} Algorithm to compute analysis scope (i.e., reachable classes and methods).
-Current choices include {\tt dynamic} (dynamic analysis), {\tt rta} (static Rapid Type Analysis), and {\tt cha} (static Class Hierarchy Analysis).
-\\
+{\bf Description:} Algorithm to compute program scope.  Current choices include {\tt dynamic} (dynamic analysis), {\tt rta} (Rapid Type Analysis), and {\tt cha} (Class Hierarchy Analysis). \\
 {\bf Default value:} {\tt rta}
@@
 \item
-\code{chord.scope.exclude.std} \\
+\code{chord.std.scope.exclude} \\
 {\bf Type:} string list \\
-{\bf Description:} List of prefixes of names of classes inside the standard library whose method bodies must be treated as no-ops by static scope building algorithms. \\
+{\bf Description:} List of prefixes of names of classes inside the standard library whose method bodies must be treated as no-ops. \\
 {\bf Default value:} \code{[chord.main.class.path]}
@@
 \item
-\code{chord.scope.exclude.ext} \\
+\code{chord.ext.scope.exclude} \\
 {\bf Type:} string list \\
-{\bf Description:} List of prefixes of names of classes outside the standard library whose method bodies must be treated as no-ops by static scope building algorithms. \\
+{\bf Description:} List of prefixes of names of classes outside the standard library whose method bodies must be treated as no-ops. \\
 {\bf Default value:} {\tt ""}
@@
 {\bf Type:} string list \\
-{\bf Description:} List of prefixes of names of classes whose method bodies must be treated as no-ops by static scope building algorithms. \\
-{\bf Default value:} \code{"[chord.main.class.path],[chord.scope.exclude.ext]"}
+{\bf Description:} List of prefixes of names of classes whose method bodies must be treated as no-ops. \\
+{\bf Default value:} \code{"[chord.std.scope.exclude],[chord.ext.scope.exclude]"}
 \end{itemize}
@@
 {\bf Description:} Path of directories containing program analyses written in Datalog (i.e., files with suffix {\tt .dlog} or {\tt .datalog}). \\
-{\bf Default value:} \code{[chord.main.dir]/src/dlog/:[chord.main.dir]/../extra/src/dlog/}
+{\bf Default value:} \code{[chord.main.dir]/src/}
 
@@
 
-\code{chord.check.exclude.std} \\
+\code{chord.std.check.exclude} \\
 {\bf Type:} string list \\
@@
 \item
-\code{chord.check.exclude.ext} \\
+\code{chord.ext.check.exclude} \\
 {\bf Type:} string list \\
@@
 {\bf Description:} List of prefixes of names of classes and packages to be excluded by program analyses.  Interpretation of this property is analysis-specific. \\
-{\bf Default value:} \code{"[chord.check.exclude.std],[chord.check.exclude.ext]"}
+{\bf Default value:} \code{"[chord.std.check.exclude],[chord.ext.check.exclude]"}
 \end{itemize}
@@
 {\bf Type:} bool \\
-{\bf Description:} Do SSA transformation for all methods deemed reachable by the algorithm used to compute analysis scope. \\
+{\bf Description:} Do SSA transformation for all methods deemed reachable by the algorithm used to compute program scope. \\
 {\bf Default value:} {\tt true} 

### 2010-12-13 main/doc/program_setup.tex
updated documentation

@@
 
-\begin{quote}
 \begin{verbatim}
@@
 \end{verbatim}
-\end{quote}
 
-\noindent The above structure is typical: the program's Java source files are under {\tt src/},
-its class files are under {\tt classes/}, the source and jar files of the libraries
-used by the program are under \code{lib/src/} and \code{lib/jar/}, respectively.
+\noindent The above structure is typical: the program's Java source
+files are under {\tt src/}, its class files are under {\tt classes/},
+the source and jar files of the libraries used by the program are
+under \code{lib/src/} and \code{lib/jar/}, respectively.
 
-File \code{chord.properties} specifies properties to be passed to Chord (alternatively,
-these properties may be passed on the command-line, in format \code{-D<name>=<value>}).
-A sample such file for the above program is as follows:
+File \code{chord.properties} specifies properties to be passed to
+Chord (alternatively, these properties may be passed on the
+command-line, in format \code{-D<name>=<value>}).  A sample such file
+for the above program is as follows:
 
-\begin{quote}
 \begin{verbatim}
@@
 \end{verbatim}
-\end{quote}
 
-Each relative path element in the value of any property
-named \code{chord.<...>.path} defined in this file is converted to an absolute path element with respect
-to the directory containing this file, which in the above case is \code{example/}.
+Each relative path element in the value of any property named
+\code{chord.<...>.path} defined in this file is converted to an
+absolute path element with respect to the directory containing this
+file, which in the above case is \code{example/}.
 
-Section \ref{sec:chord-sysprops} presents an exhaustive
-listing of properties recognized by \Chord.
-Here, we only describe those defined in the above
-sample \code{chord.properties} file:
+Section \ref{sec:chord-sysprops} presents an exhaustive listing of
+properties recognized by \Chord.  Here, we only describe those defined
+in the above sample \code{chord.properties} file:
 
@@
 \item
-\code{chord.main.class} specifies the fully-qualified name of the main class of the
-program.
+\code{chord.main.class} specifies the fully-qualified name of the main
+class of the program.
 \item
-\code{chord.class.path} specifies the application-specific classpath of the program
-(the standard library classpath is implicitly included).
+\code{chord.class.path} specifies the application-specific classpath
+of the program (the standard library classpath is implicitly
+included).
 \item
 \code{chord.src.path} specifies the Java source path of the program.
-All program analyses in Chord operate on Java bytecode.
-The only use of this property is to HTMLize the Java source files of the program so that the
-results of program analyses can be reported at the Java source code level.
+All program analyses in Chord operate on Java bytecode.  The only use
+of this property is to HTMLize the Java source files of the program so
+that the results of program analyses can be reported at the Java
+source code level.
 \item
-\code{chord.run.ids} specifies a list of IDs to identify runs of the program.
-It is used by dynamic program analyses to determine how many times the program must be run.
-An additional use of this property is to allow specifying the command-line arguments to use
-in the run having ID {\tt <id>} via property \code{chord.args.<id>}, as illustrated by
-properties \code{chord.args.0} and \code{chord.args.1} in the above example.
+\code{chord.run.ids} specifies a list of IDs to identify runs of the
+program.  It is used by dynamic program analyses to determine how many
+times the program must be run.  An additional use of this property is
+to allow specifying the command-line arguments to use in the run
+having ID {\tt <id>} via property \code{chord.args.<id>}, as
+illustrated by properties \code{chord.args.0} and \code{chord.args.1}
+in the above example.
 \end{itemize}
 
-To run Chord on the above program, run the following command in Chord's {\tt main/} directory:
+To run Chord on the above program, run the following command in
+Chord's {\tt main/} directory:
 
-\begin{quote}
 \begin{verbatim}
@@
 \end{verbatim}
-\end{quote}
 
-\noindent where {\tt <...>} denotes either the absolute path, or the path relative to Chord's {\tt main/} directory,
-of the parent of directory {\tt example/}.
-System property \code{chord.work.dir} specifies the working directory during Chord's execution.
+\noindent where {\tt <...>} denotes the absolute path of the parent of
+directory {\tt example/}.  Property \code{chord.work.dir} specifies
+the working directory during Chord's execution.
 
-The above command causes Chord to load the \code{chord.properties} file if present.
-The location of this file may be specified explicitly via property
-\code{chord.props.file} on the above command line.
-By default, it is \code{[chord.work.dir]/chord.properties}.
+The above command causes Chord to load the \code{chord.properties}
+file if present.  The location of this file may be specified
+explicitly via property \code{chord.props.file} on the above command
+line.  By default, it is \code{[chord.work.dir]/chord.properties}.
 
-In the above case, Chord does not do much beyond loading the sample \code{chord.properties} file.
-For Chord to do something interesting, you need to set additional properties, either
-on the above command line or in the \code{chord.properties} file, specifying the task(s)
-Chord must perform.
-All tasks are summarized in Section \ref{sec:chord-sysprops:task}.
-The two most common tasks, described next, are computing the scope of the given program (Section \ref{sec:building-scope})
-and running analyses on the given program (Section \ref{sec:running-analysis}).
+In the above case, Chord does not do much beyond loading the sample
+\code{chord.properties} file.  For Chord to do something interesting,
+you need to set additional properties, either on the above command
+line or in the \code{chord.properties} file, specifying the task(s)
+Chord must perform.  All tasks are summarized in Section
+\ref{sec:chord-sysprops:task}.  The two most common tasks, described
+next, are computing the scope of the given program (Section
+\ref{sec:building-scope}) and running analyses on the given program
+(Section \ref{sec:running-analysis}).

### 2010-12-13 main/doc/whatis_chord.tex
updated documentation

@@
 
-\Chord\ is a static and dynamic program analysis framework for Java.
-It has the following key characteristics:
+\Chord\ is a program analysis platform for Java designed with the
+primary goal of helping advance the state-of-the-art in the field by
+enabling researchers to productively design, implement, and evaluate a
+broad variety of program analyses. To this end, Chord has the
+following key characteristics:
 
@@
 \item
-Stand-alone: various off-the-shelf program analyses are provided (e.g., may-alias, thread-escape, datarace, deadlock, etc.).
+Stand-alone: it provides various off-the-shelf analyses (e.g., various
+call-graph analyses; points-to analyses; thread-escape analysis;
+static slicing; static and dynamic concurrency analyses for finding
+races, deadlocks, and atomicity violations; etc.)
 \item
-Extensible: users can build their own program analyses on top of the provided ones.
+Extensible: it allows users to build their own analyses atop the
+provided analyses.
 \item
-Compositional: each program analysis can be written independently and yet made to interact in complex ways with other program analyses.
+Compositional: it allows users to write each analysis independently
+yet allow it to interact in complex ways with other analyses by
+specifying lightweight data- and control-dependencies.
 \item
-Efficient: results computed by each program analysis are cached for reuse by other program analyses without re-computation.
+Efficient: It executes analyses in a demand-driven fashion, caches
+results computed by each analysis for reuse by other analyses without
+re-computation, and can execute analyses without dependencies between
+them in parallel.
 \item
-Flexible: a broad range of program analyses can be expressed, including those
-written imperatively in Java or declaratively in Datalog, summary-based as well as cloning-based context-sensitive analyses,
-iterative refinement-based analyses, client-driven analyses, and combined static and dynamic analyses.
+Deterministic: It guarantees that the result is the same regardless of
+the order in which different analyses are executed; moreover, results
+can be shared across different runs.
+\item
+Flexible: It allows a broad range of analyses to be expressed,
+including both static and dynamic analyses, analyses written
+imperatively in Java or declaratively in Datalog, summary-based as
+well as cloning-based inter-procedural context-sensitive analyses,
+iterative refinement-based analyses, client-driven analyses, and
+combined static/dynamic analyses.
 \end{itemize}
 
-\noindent \Chord\ is intended for use by a broad range of users:
-
-\begin{itemize}
-\item
-Program analysis writers: Program analysis researchers wanting to implement and evaluate new program analysis algorithms.
-\item
-Program analysis appliers: Researchers with possibly a limited program analysis background wanting to build applications on
-top of program analyses in Chord used as black boxes.
-\item
-Software engineers: Programmers wanting to use program analyses in Chord to assist with software development and testing.
-\end{itemize}
-
-\noindent \Chord\ is intended to work on a variety of platforms, including Linux, Windows/Cygwin, and MacOS.
-It has been tested most extensively on Linux.
-It is open source software distributed under the \htmladdnormallink{New BSD License}{http://www.opensource.org/licenses/bsd-license.php}.
-Contributions from users are welcome and encouraged.
-The project website is located at \url{http://code.google.com/p/jchord/}.
+\noindent \Chord\ is intended to work on a variety of platforms,
+including Linux, Windows/Cygwin, and MacOS.  It has been tested most
+extensively on Linux.  It is open source software distributed under
+the \htmladdnormallink{New BSD
+  License}{http://www.opensource.org/licenses/bsd-license.php}.
+Improvements by users are welcome and encouraged.  The project website
+is \url{http://berkeley.intel-research.net/chord/} and the software
+website is \url{http://code.google.com/p/jchord/}.

### 2010-12-13 main/doc/building_scope.tex
migrated documentation from latex2html to hyperlatex

@@
 
-\Chord\ computes the scope (i.e., reachable methods) of the given
+Chord computes the scope (i.e., reachable methods) of the given
 program either if property \code{chord.build.scope} is set to {\tt
@@
 files specified by properties \code{chord.methods.file} and
-\code{chord.reflect.file} exist, then \Chord\ regards those files as
+\code{chord.reflect.file} exist, then Chord regards those files as
 specifying which methods to consider reachable and how to resolve
@@
 files specified by properties \code{chord.methods.file} or
-\code{chord.reflect.file} do not exist, then \Chord\ computes program
+\code{chord.reflect.file} do not exist, then Chord computes program
 scope using the algorithm specified by property
@@
 The possible legal values of property \code{chord.scope.kind} are
-[\code{dynamic}$|$\code{rta}$|$\code{cha}].  In each case, \Chord\ at
+[\code{dynamic}$|$\code{rta}$|$\code{cha}].  In each case, Chord at
 least expects properties \code{chord.main.class} and
@@
 \item
-The {\tt dynamic} value instructs \Chord\ to compute program scope
+The {\tt dynamic} value instructs Chord to compute program scope
 dynamically, by running the program and observing using JVMTI the
@@
 \item
-The {\tt rta} value instructs \Chord\ to compute program scope
+The {\tt rta} value instructs Chord to compute program scope
 statically using Rapid Type Analysis (RTA).
@@
 reachable methods $M$.  The initial iteration starts by assuming that
-only the main method in the main class is reachable (\Chord\ also
+only the main method in the main class is reachable (Chord also
 handles class initializer methods but we ignore them here for brevity;
@@
 %\item
-%The \code{rta_reflect} value instructs \Chord\ to compute program scope statically using Rapid Type Analysis
+%The \code{rta_reflect} value instructs Chord to compute program scope statically using Rapid Type Analysis
 %and, moreover, to resolve a common reflection pattern:
@@
 %or that implement \code{T} (if \code{T} is an interface).
-%\Chord\ allows users to control which classes are included in the class hierarchy (see below).
+%Chord allows users to control which classes are included in the class hierarchy (see below).
 %
 \item
-The {\tt cha} value instructs \Chord\ to compute program scope
+The {\tt cha} value instructs Chord to compute program scope
 statically using Class Hierarchy Analysis (CHA).
@@
 grossly overestimates the set of reachable classes and methods.
-Nevertheless, \Chord\ allows users to control which classes are
+Nevertheless, Chord allows users to control which classes are
 included in the class hierarchy (see below) and thereby control the
@@
 whose possible legal values are [\code{static}$|$\code{dynamic}].
-\Chord\ first constructs the entire classpath of the given program by
+Chord first constructs the entire classpath of the given program by
 concatenating in order the following classpaths:
@@
 \item
-\Chord's classpath, specified by property
+Chord's classpath, specified by property
 \code{chord.main.class.path}.  This includes the classpath of any
@@
 If property \code{chord.ch.kind} has value {\tt dynamic}, then
-\Chord\ runs the given program and observes the set of all classes the
+Chord runs the given program and observes the set of all classes the
 JVM loads; any class not in this set is excluded.  By default,

### 2010-12-13 main/doc/download_and_installation.tex
minor (added documentation + fixed javadoc tags)
migrated documentation from latex2html to hyperlatex

@@
 \item
-JDK 6 or higher, e.g. from \htmladdnormallink{IBM}{http://www.ibm.com/developerworks/java/jdk/} or
-\htmladdnormallink{Sun}{http://java.sun.com/javase/downloads/index.jsp}
+JDK 5 or higher, e.g. from
+\xlink{IBM}{http://www.ibm.com/developerworks/java/jdk/} or
+\xlink{Sun}{http://java.sun.com/javase/downloads/index.jsp}
 \item
-\htmladdnormallink{Apache Ant}{http://ant.apache.org/}, a Java-based build tool
+\xlink{Apache Ant}{http://ant.apache.org/}, a Java-based build tool
 \item
-a C++ compiler, e.g. \htmladdnormallink{GCC}{http://gcc.gnu.org/}
+a C++ compiler, e.g. \xlink{GCC}{http://gcc.gnu.org/}
 \item
-a Make utility, e.g. \htmladdnormallink{GNU Make}{http://www.gnu.org/software/make/}
+a Make utility, e.g. \xlink{GNU Make}{http://www.gnu.org/software/make/}
 \item
-\htmladdnormallink{Cygwin}{http://www.cygwin.com/}, if it is a Windows machine
+\xlink{Cygwin}{http://www.cygwin.com/}, if it is a Windows machine
 \end{itemize}
 
-Download a \Chord\ source release from \url{http://code.google.com/p/jchord/downloads/list} or
-the latest source from the SVN repository at \url{http://code.google.com/p/jchord/source/checkout}.
+Download the latest Chord source code from the SVN repository:
 
-Directory {\tt main/} contains a {\tt build.xml} file which is interpreted by Apache Ant.
-To see the various targets available, run the following command in that directory:
+\begin{verbatim}
+    prompt> svn checkout http://jchord.googlecode.com/svn/trunk/ chord
+\end{verbatim}
+For many users, it may be sufficient to only checkout directory {\tt
+  trunk/main/} instead of the entire {\tt trunk/}.  Other directories
+of interest are {\tt trunk/test/}, which contains regression tests,
+and {\tt trunk/libsrc/}, which contains the source code of software
+used by Chord (Joeq, Javassist, bddbddb, JavaBDD, and jwutil).  All
+other directories under {\tt trunk/} are projects built atop Chord
+that might be of interest to some users.
 
-\begin{quote}
+Directory {\tt main/} contains a {\tt build.xml} file which is
+interpreted by Apache Ant.  To see the various targets available, run
+the following command in that directory:
+
 \begin{verbatim}
@@
 \end{verbatim}
-\end{quote}
 
-\noindent To compile \Chord, run the following command in the same directory:
+\noindent To compile Chord, run the following command in the same
+directory:
 
-\begin{quote}
 \begin{verbatim}
@@
 \end{verbatim}
-\end{quote}
 
@@
 \item
-the Java source code of \Chord\ from \code{main/src/java/} (and from \code{extra/src/java/} if that directory is present) to Java bytecode in \code{main/classes/}.
+the Java source code of Chord from \code{main/src/} to Java bytecode
+in \code{main/classes/}.
 \item
-the C source code of BDD library \htmladdnormallink{BuDDy}{http://buddy.sourceforge.net/} from \code{main/src/bdd/} to a shared library in \code{main/lib/} (\code{libbuddy.so} on Linux, \code{buddy.dll} on Windows, and \code{libbuddy.dylib} on MacOS); this library is needed for executing program analyses written in Datalog using \htmladdnormallink{bddbddb}{http://bddbddb.sourceforge.net/} (a BDD-based Datalog solver).
+the C source code of BDD library
+\xlink{BuDDy}{http://buddy.sourceforge.net/} from
+\code{main/bdd/} to a shared library in \code{main/lib/}
+(\code{libbuddy.so} on Linux, \code{buddy.dll} on Windows, and
+\code{libbuddy.dylib} on MacOS); this library is needed for executing
+program analyses written in Datalog using
+\xlink{bddbddb}{http://bddbddb.sourceforge.net/} (a
+BDD-based Datalog solver).
 \item
-the C++ source code of the Chord instrumentation agent from \code{main/src/agent/} to a shared library in \code{main/lib/} (\code{libchord_instr_agent.so} on all architectures); this agent is needed for computing program scope dynamically and for executing dynamic program analyses.
+the C++ source code of the Chord instrumentation agent from
+\code{main/agent/} to a shared library in \code{main/lib/}
+(\code{libchord_instr_agent.so} on all architectures); this agent is
+needed for computing program scope dynamically and for executing
+dynamic program analyses.
 \end{itemize}

### 2010-12-13 main/doc/tuning_datalog_analysis.tex
migrated documentation from latex2html to hyperlatex

@@
 There are several tricks analysis writers can try to improve the
-performance of \bddbddb, the Datalog solver used by \Chord, often
-by several orders of magnitude:
+performance of bddbddb, the Datalog solver used by Chord, often
+by several orders of magnitude.
+Try these tricks by running the following command:
+
+\begin{verbatim}
+    prompt> ant -Ddlog.file=<file> -Dwork.dir=<dir> solve
+\end{verbatim}
+where {\tt <file>} denotes the file defining the Datalog analysis
+to be tuned, and {\tt <dir>} is the directory containing the
+program domains ({\tt *.dom} files) and program relations ({\tt *.bdd} files)
+consumed by the analysis (this is by default the
+\code{chord_output/bddbddb/} directory generated
+by a previous run of Chord.
 
@@
 \item
-Set system properties \verb+noisy=yes+, \verb+tracesolve=yes+, and \verb+fulltracesolve=yes+
-while running \bddbddb\ and observe which rule gets ``stuck" (i.e., takes several seconds to solve).
-\verb+fulltracesolve+ is seldom useful, but \verb+noisy+ and \verb+tracesolve+ is
+Set properties \verb+noisy=yes+, \verb+tracesolve=yes+, and \verb+fulltracesolve=yes+
+on the above command line and observe which rule gets ``stuck" (i.e., takes several seconds to solve).
+\verb+fulltracesolve+ is seldom useful, but \verb+noisy+ and \verb+tracesolve+ are
 often very useful.  Once you identify the rule that is getting stuck, it
-will also tell you which relations and which domains used in that rule
-and which operation on them is taking a long time to solve.  Then try
-to fix the slowdown problem with that rule by either simplifying the
-relations involved (e.g., breaking their set of attributes into two if
-possible) or changing the relative order of the domains of those
-relations in the BDD variable order (doing the latter alone
-frequently causes the problem to go away).
+will also tell you which relations and which domains used in that rule,
+and which operation on them, is taking a long time to solve.  Then try
+to fix the problem with that rule by doing either or both of the following:
+\begin{itemize}
+\item
+Break down the rule into multiple rules by creating intermediate relations (the more
+relations you have on the RHS of a rule the slower it generally takes to solve
+that rule).
+\item
+Change the relative order of the domains of those
+relations in the BDD variable order
+(note that you can use either `\_' or `x' between a pair of domains).
+\end{itemize}
 
@@
 time, the cumulative time for the rule is too much.  After finishing
-solving a Datalog analysis, \bddbddb\ prints how long each rule took to
+solving a Datalog analysis, bddbddb prints how long each rule took to
 solve (both in terms of the number of times it was applied and the
@@
 solving faster.  In general, look at the last few rules, not just the
-last one, and try to identify the "sub-program" of the Datalog program
+last one, and try to identify the ``sub-analysis" of the Datalog analysis
 that seems problematic, and then focus on speeding up just that sub-
-program.
-
-\item
-Tweak the BDD variable ordering in the Datalog file (the line that
-starts with \verb+bddvarorder+).  This is one of the most
-effective ways of making \bddbddb\ run faster.  As explained above, use
-\verb+noisy+ and \verb+tracesolve+ to find out which pairs of domains in the
-bddvarorder are responsible for the slowdown, and try to change their
-relative ordering (note that you can use either `\_' or `x' between a pair
-of domains, and the latter is commutative).
+analysis.
 
@@
 You can add the \verb+.split+ keyword at the end of certain rules as a
-hint to \bddbddb\ to break up those rules into simpler ones that can be
+hint to bddbddb to decompose those rules into simpler ones that can be
 solved faster.  You can also set property \verb+split_all_rules=yes+ as shorthand
@@
 \item
-You can yourself break down rules by creating intermediate relations (the more
-relations you have on the RHS of a rule the slower it takes to solve
-that rule).  This is another very effective way to make \bddbddb\ run
-faster.
-
-\item
-Try breaking down a single Datalog program into two programs.  Of
-course, you cannot separate mutually-recursive rules into two
-different programs, but if you unnecessarily club
-together rules that could have gone into different programs, then they
-can put conflicting demands on \bddbddb\ (e.g., on \verb+bddvarorder+).
+You can try to decompose a single Datalog analysis file into two separate Datalog analysis
+files.  Of course, you cannot separate mutually-recursive rules into two
+different analyses, but if you unnecessarily club
+together rules that could have gone into different analyses, then they
+can put conflicting demands on bddbddb (e.g., on the BDD variable order).
 So if rule 2 uses the result of rule 1 and rule 1 does not use the result of
-rule 2, then put rule 1 and rule 2 in two different Datalog programs.
+rule 2, then put rule 1 and rule 2 in separate Datalog analyses.
 
@@
 Observe the sizes of the BDDs representing the relations that are
-input and output.  \bddbddb\ tells both the number of tuples in each
-relation and the number of "nodes" in the BDD.  Try changing the
-bddvarorder for the domains of the relation, and observe how the
-number of "nodes" in the bdd for that relation change.  You will
-notice that some orderings do remarkably better than others.  Then
-note down these orderings as invariants that you will not violate as
+input and output.  bddbddb prints both the number of tuples in each
+relation and the number of nodes in the BDD.  Try changing the
+BDD variable order for the domains of the relation, and observe how the
+number of nodes in the BDD for that relation change.  You will
+notice that some orders perform remarkably better than others.  Then
+note down these orders as invariants that you will not violate as
 you tweak other things.
@@
 \item
-The relative ordering of values *within* domains (e.g.
+The relative order of values *within* domains (e.g.,
 in domains named \verb+M+, \verb+H+, \verb+C+, etc. in Chord) affects the
-solving time of \bddbddb, but
+performance of bddbddb, but
 I've never tried changing this and studying its effect.  It might be
@@
 specific way in which he numbers contexts (in domain \verb+C+) and that it was
-fundamental to the speedup of his infinity-CFA points-to analysis.
+fundamental to the speedup of his ``infinity"-CFA points-to analysis.
 
@@
 unlikely to help you a whole lot.  Secondly, many things are awkward to
-encode as integers (e.g., the abstract contexts in the domain \verb+C+ 
+encode as integers (e.g., the abstract contexts in domain \verb+C+ 
 in Chord) or as Datalog rules.

### 2010-12-13 main/doc/whatis_chord.tex
migrated documentation from latex2html to hyperlatex

@@
 
-\Chord\ is a program analysis platform for Java designed with the
+Chord is a program analysis platform for Java designed with the
 primary goal of helping advance the state-of-the-art in the field by
@@
 
-\noindent \Chord\ is intended to work on a variety of platforms,
+\noindent Chord is intended to work on a variety of platforms,
 including Linux, Windows/Cygwin, and MacOS.  It has been tested most
 extensively on Linux.  It is open source software distributed under
-the \htmladdnormallink{New BSD
-  License}{http://www.opensource.org/licenses/bsd-license.php}.
+the \xlink{New BSD License}{http://www.opensource.org/licenses/bsd-license.php}.
 Improvements by users are welcome and encouraged.  The project website

### 2010-12-13 main/doc/writing_dynamic_analysis.tex
migrated documentation from latex2html to hyperlatex

@@
 \item
-You can reuse dynamic traces from a previous \Chord\ run by setting property
+You can reuse dynamic traces from a previous Chord run by setting property
 \code{chord.reuse.traces} to {\tt true}.  In this case, you must also set property
@@
 \item
-You can set property \code{chord.dynamic.haltonerr} to {\tt false} to prevent \Chord\ from terminating
+You can set property \code{chord.dynamic.haltonerr} to {\tt false} to prevent Chord from terminating
 even if the program on which dynamic analysis is being performed crashes.
@@
 
-\Chord\ offers much more flexibility in crafting dynamic analyses.
+Chord offers much more flexibility in crafting dynamic analyses.
 You can define your own instrumentor (by subclassing \code{chord.instr.CoreInstrumentor}

### 2010-12-18 main/doc/bdd_representation.tex
cleaned up and updated documentation

@@
-\section{BDD Representation}
+\section{BDD representation}
 \label{sec:bdd-representation}
 
-Each domain containing $N$ elements is assigned $log_2 N$ BDD variables in the underlying BDD factory with contiguous IDs.
+Each domain containing N elements is assigned log2(N) BDD variables in the underlying BDD factory with contiguous IDs.
 For instance,
-domain {\tt F0} containing [2048..4096) elements will be assigned 12 variables with IDs (say) 59,60,61,62,63,64,65,66,67,68,69,70 and
+domain {\tt F0} containing [128..256) elements will be assigned 8 variables with IDs (say) 63,64,65,66,67,68,69,70 and
 domain {\tt Z0} containing [8..16) elements will be assigned 4 variables with IDs (say) 105,106,107,108.
@@
 For instance, the BDD variable order corresponding to the declared domain order {\tt F0\_Z0} is (in level2var format)
-``70,69,68,67,66,65,64,63,62,61,60,59,108,107,106,105".
+``70,69,68,67,66,65,64,63,108,107,106,105".
 
@@
 For instance, the BDD variable order corresponding to the declared domain order {\tt F0xZ0} is (in level2var format)
-``70,108,69,107,68,106,67,105,66,65,64,63,62,61,60,59".
+``70,108,69,107,68,106,67,105,66,65,64,63".
 

### 2010-12-18 main/doc/chord_sysprops.tex
cleaned up and updated documentation

@@
 
-\section{Chord System Properties}
+\section{Chord system properties}
 \label{sec:chord-sysprops}
 
-Users can use system properties to control \Chord's functionality.
-The following properties are recognized by \Chord.
+Users can use system properties to control Chord's functionality.
+The following properties are recognized by Chord.
 The separator for list-valued properties can be either a blank space, a comma, a colon, or a semi-colon.
@@
 
-This section describes basic properties of the given program that are required by program analyses in \Chord,
+This section describes basic properties of the given program that are required by program analyses in Chord,
 such as its main class, the location(s) of its class files and Java source files, and command-line arguments
@@
 {\bf Description:} Fully-qualified name of the main class of the given program (e.g., \code{com.example.Main}). \\
-{\bf Note:} This property is required by virtually every task \Chord\ performs.
+{\bf Note:} This property is required by virtually every task Chord performs.
 
@@
 {\bf Default value:} {\tt ""} \\
-{\bf Note:} This property is required by virtually every task \Chord\ performs.
+{\bf Note:} This property is required by virtually every task Chord performs.
 
@@
 {\bf Default value:} {\tt ""}  \\
-{\bf Note:} \Chord\ analyzes only Java bytecode, not Java source code.
+{\bf Note:} Chord analyzes only Java bytecode, not Java source code.
 This property is used only by the task of converting Java source files into HTML files by program analyses that need to present their analysis results at the Java source code level (by calling method \code{chord.program.Program.g().HTMLizeJavaSrcFiles())}.
@@
 {\bf Default value:} {\tt 0} \\
-{\bf Note:} This property is used only when \Chord\ executes the given program, namely, when it is asked to compute the program scope dynamically (i.e., when \code{[chord.scope.kind]=dynamic}) or when it is asked to execute a dynamic program analysis.
+{\bf Note:} This property is used only when Chord executes the given program, namely, when it is asked to compute the program scope dynamically (i.e., when \code{[chord.scope.kind]=dynamic}) or when it is asked to execute a dynamic program analysis.
 
@@
 {\bf Default value:} {\tt ""} \\
-{\bf Note:} This property is used only when \Chord\ executes the given program, namely, when it is asked to compute the program scope dynamically (i.e., when \code{[chord.scope.kind]=dynamic}) or when it is asked to execute a dynamic program analysis.
+{\bf Note:} This property is used only when Chord executes the given program, namely, when it is asked to compute the program scope dynamically (i.e., when \code{[chord.scope.kind]=dynamic}) or when it is asked to execute a dynamic program analysis.
 
@@
 {\bf Default value:} {\tt "-ea -Xmx1024m"} \\
-{\bf Note:} This property is used only when \Chord\ executes the given program, namely, when it is asked to compute the program scope dynamically (i.e., when \code{[chord.scope.kind]=dynamic}) or when it is asked to execute a dynamic program analysis.
+{\bf Note:} This property is used only when Chord executes the given program, namely, when it is asked to compute the program scope dynamically (i.e., when \code{[chord.scope.kind]=dynamic}) or when it is asked to execute a dynamic program analysis.
 \end{itemize}
@@
 
-This section describes properties that specify what task(s) \Chord\ must perform, 
+This section describes properties that specify what task(s) Chord must perform, 
 such as computing the scope of the given program or running program analyses on the given program.
@@
 {\bf Note:} This task must be used with caution as certain program relations, albeit represented compactly as BDDs, may contain a large number (e.g., billions) of tuples, resulting in voluminous output when printed to a file.
-See Section \ref{sec:tuning-datalog-analysis} for a more efficient way to query the contents of program relations (namely, by using the {\tt debug} target provided in file {\tt build.xml} in \Chord's {\tt main/} directory).
+See Section \ref{sec:tuning-datalog-analysis} for a more efficient way to query the contents of program relations (namely, by using the {\tt debug} target provided in file {\tt build.xml} in Chord's {\tt main/} directory).
 
@@
 
-This section describes properties of the JVM running \Chord, such as the working directory and various memory limits.
+This section describes properties of the JVM running Chord, such as the working directory and various memory limits.
 
@@
 {\bf Type:} location \\
-{\bf Description:} Working directory during \Chord's execution. \\
+{\bf Description:} Working directory during Chord's execution. \\
 {\bf Default value:} current working directory
@@
 {\bf Description:} Properties file loaded by \code{[chord.main.dir]/build.xml}.
-Any of the below properties, as well as other user-defined properties to be passed to \Chord\ (e.g., for user-defined analyses), may be set in this file.
+Any of the below properties, as well as other user-defined properties to be passed to Chord (e.g., for user-defined analyses), may be set in this file.
 Each relative (instead of absolute) path element in the value of any property named \code{chord.<...>.path} set in this file is converted to an absolute path element with respect to the directory containing this file.
@@
 {\bf Type:} string \\
-{\bf Description:} Maximum memory size of JVM running \Chord. \\
+{\bf Description:} Maximum memory size of JVM running Chord. \\
 {\bf Default value:} {\tt 1024m}
@@
 {\bf Type:} string \\
-{\bf Description:} Maximum thread stack size of JVM running \Chord. \\
+{\bf Description:} Maximum thread stack size of JVM running Chord. \\
 {\bf Default value:} {\tt 32m}
@@
 {\bf Type:} string \\
-{\bf Description:} Arguments to JVM running \Chord. \\
+{\bf Description:} Arguments to JVM running Chord. \\
 {\bf Default value:} \code{"-showversion} \code{-ea} \code{-Xmx[chord.max.heap]} \code{-Xss[chord.max.stack]"}
@@
 {\bf Type:} string \\
-{\bf Description:} Maximum memory size of JVM running \bddbddb. \\
+{\bf Description:} Maximum memory size of JVM running bddbddb. \\
 {\bf Default value:} {\tt 1024m} \\
-{\bf Note:} Each program analysis written in Datalog is run in a separate JVM because there may be multiple invocations of the Datalog solver \bddbddb\ in a single run of \Chord\ and it is difficult to reset the global state of \bddbddb\ on each invocation.
+{\bf Note:} Each program analysis written in Datalog is run in a separate JVM because there may be multiple invocations of the Datalog solver bddbddb in a single run of Chord and it is difficult to reset the global state of bddbddb on each invocation.
 \end{itemize}
@@
 {\bf Type:} {\tt [dynamic|rta|cha]} \\
-{\bf Description:} Algorithm to compute program scope.  Current choices include {\tt dynamic} (dynamic analysis), {\tt rta} (Rapid Type Analysis), and {\tt cha} (Class Hierarchy Analysis). \\
+{\bf Description:} Algorithm to compute program scope.  The choices are {\tt dynamic} (dynamic analysis), {\tt rta} (Rapid Type Analysis), and {\tt cha} (Class Hierarchy Analysis). \\
 {\bf Default value:} {\tt rta}
@@
 {\bf Type:} string list \\
-{\bf Description:} List of prefixes of names of classes inside the standard library whose method bodies must be treated as no-ops. \\
+{\bf Description:} Partial list of prefixes of names of classes whose declared methods must be treated as no-ops.  Conventionally, this property is used to specify classes to be
+excluded from the scope of all programs to be analyzed, e.g., classes in the JDK standard library.  \\
 {\bf Default value:} \code{[chord.main.class.path]}
@@
 {\bf Type:} string list \\
-{\bf Description:} List of prefixes of names of classes outside the standard library whose method bodies must be treated as no-ops. \\
+{\bf Description:} Partial list of prefixes of names of classes whose declared methods must be treated as no-ops.  Conventionally, this property is used to specify classes to be
+excluded from the scope of specific programs to be analyzed. \\
 {\bf Default value:} {\tt ""}
@@
 {\bf Type:} string list \\
-{\bf Description:} List of prefixes of names of classes whose method bodies must be treated as no-ops. \\
+{\bf Description:} Complete list of prefixes of names of classes whose declared methods must be treated as no-ops. \\
 {\bf Default value:} \code{"[chord.std.scope.exclude],[chord.ext.scope.exclude]"}
@@
 
-This section describes properties regarding program analyses executed by \Chord.
+This section describes properties regarding program analyses executed by Chord.
 
@@
 {\bf Type:} bool \\
-{\bf Description:} Construct program relations from BDDs stored on disk (from a previous run of \Chord) whenever possible instead of re-computing them. \\
+{\bf Description:} Construct program relations from BDDs stored on disk (from a previous run of Chord) whenever possible instead of re-computing them. \\
 {\bf Default value:} {\tt false} 
@@
 
+\item
 \code{chord.std.check.exclude} \\
@@
 
-This section describes properties regarding program transformations performed by \Chord.
+This section describes properties regarding program transformations performed by Chord.
 
@@
 
-This section describes properties that specify the amount of debug information that \Chord\ must produce during execution.
+This section describes properties that specify the amount of debug information that Chord must produce during execution.
 
@@
 \code{chord.verbose} \\
-{\bf Type:} int in the range [0..3] \\
-{\bf Description:} Control the verbosity of logging during \Chord's execution. \\
+{\bf Type:} int in the range [0..6] \\
+{\bf Description:} Control the verbosity of logging during Chord's execution. \\
 {\bf Default value:} {\tt 2}
-
-\item
-\code{chord.bddbddb.verbose} \\
-{\bf Type:} int in the range [0..5] \\
-{\bf Description:} Control the verbosity of logging during the execution of JavaBDD (BDD library) and \bddbddb\ (Datalog solver). \\
-{\bf Default value:} {\tt 0}
 
@@
 {\bf Default value:} {\tt true}  \\
-{\bf Note:} This functionality is useful for debugging Datalog programs using the {\tt debug} target provided in file {\tt build.xml} in \Chord's {\tt main/} directory (see Section \ref{sec:tuning-datalog-analysis}).
+{\bf Note:} This functionality is useful for debugging Datalog programs using the {\tt debug} target provided in file {\tt build.xml} in Chord's {\tt main/} directory (see Section \ref{sec:tuning-datalog-analysis}).
 \end{itemize}
@@
 \item
-\code{chord.reuse.traces} \\
-{\bf Type:} bool \\
-{\bf Description:} Reuse dynamic traces computed by a previous run of \Chord\ if they exist. \\
-{\bf Default value:} {\tt false} \\
-{\bf Note:} Property \code{chord.trace.kind} must be set to {\tt full} if this property is set to {\tt true}.
+\code{chord.instr.kind} \\
+{\bf Type:} {\tt [offline|online]} \\
+{\bf Description:} The kind of bytecode instrumentation.  The choices are offline and load-time. \\
+{\bf Default value:} {\tt offline} \\
 
@@
 \item
-\code{chord.max.cons.size} \\
-{\bf Type:} int \\
-{\bf Description:} Maximum number of bytes over which events generated during the execution of any constructor in the given program may span. \\
-{\bf Default value:} 50000000 \\
-{\bf Note:} This property is relevant only for dynamic analyses which want events of the form {\tt NEW} $h$ $t$ $o$ to be generated (see Section \ref{sec:instr-events}).
-The problem with generating such events at run-time is that the ID $o$ of the object freshly created by thread $t$ at object allocation site $h$ cannot be instrumented until the object is fully initialized (i.e., its constructor has finished executing).
-Hence, \Chord\ first generates a ``crude dynamic trace", which has events of the form {\tt BEF\_NEW} $h$ $t$ and {\tt AFT\_NEW} $h$ $t$ $o$ generated before and after the execution of the constructor, respectively.
-A subsequent pass generates a ``final dynamic trace", which replaces the {\tt BEF\_NEW} $h$ $t$ events by {\tt NEW} $h$ $t$ $o$ and discards the {\tt AFT\_NEW} $h$ $t$ $o$ events.
-For this purpose, however, \Chord\ must buffer all events generated between the {\tt BEF\_NEW} and {\tt AFT\_NEW} events, and this property specifies the number of bytes over which these events may span.
-If the actual number of bytes exceeds the value specified by this property (e.g., if the constructor throws an exception and the {\tt AFT\_NEW} event is not generated at all), then \Chord\ simply generates event {\tt NEW} $h$ $i$ $0$ (i.e., it treats the created object as having ID 0, which is the ID also used for {\tt null}).
+\code{chord.reuse.traces} \\
+{\bf Type:} bool \\
+{\bf Description:} Reuse dynamic traces computed by a previous run of Chord if they exist. \\
+{\bf Default value:} {\tt false} \\
+{\bf Note:} Property \code{chord.trace.kind} must be set to {\tt full} if this property is set to {\tt true}.
 
@@
 {\bf Default value:} -1
+
+\item
+\code{chord.max.cons.size} \\
+{\bf Type:} int \\
+{\bf Description:} Maximum number of bytes over which events generated during the execution of any constructor in the given program may span. \\
+{\bf Default value:} 50000000 \\
+{\bf Note:} This property is relevant only for dynamic analyses which want events of the form {\tt BEF\_NEW} $h$ $t$ $o$ to be generated (see Section \ref{sec:instr-events}).
+The problem with generating such events at run-time is that the ID $o$ of the object freshly created by thread $t$ at object allocation site $h$ cannot be instrumented until the object is fully initialized (i.e., its constructor has finished executing).
+Hence, Chord first generates a ``crude dynamic trace", which has events of the form {\tt BEF\_NEW} $h$ $t$ and {\tt AFT\_NEW} $h$ $t$ $o$ generated before and after the execution of the constructor, respectively.
+A subsequent pass generates a ``final dynamic trace", which replaces each {\tt BEF\_NEW} $h$ $t$ event by {\tt BEF\_NEW} $h$ $t$ $o$.
+For this purpose, however, Chord must buffer all events generated between the {\tt BEF\_NEW} and {\tt AFT\_NEW} events, and this property specifies the number of bytes over which these events may span.
+If the actual number of bytes exceeds the value specified by this property (e.g., if the constructor throws an exception and the {\tt AFT\_NEW} event is not generated at all), then Chord simply generates event {\tt BEF\_NEW} $h$ $i$ $0$ (i.e., it treats the created object as having ID 0, which is the ID also used for {\tt null}).
 \end{itemize}
@@
 \item
-\code{chord.classes.file} \\
+\code{chord.reflect.file} \\
 {\bf Type:} location \\
-{\bf Description:} Absolute location of the file from/to which list of classes deemed reachable is read/written. \\
-{\bf Default value:} \code{[chord.out.dir]/classes.txt}
+{\bf Description:} Absolute location of the file from/to which resolved reflection information is read/written. \\
+{\bf Default value:} \code{[chord.out.dir]/reflect.txt}
 
@@
 \item
+\code{chord.classes.file} \\
+{\bf Type:} location \\
+{\bf Description:} Absolute location of the file from/to which list of classes deemed reachable is read/written. \\
+{\bf Default value:} \code{[chord.out.dir]/classes.txt}
+
+\item
 \code{chord.bddbddb.work.dir} \\
 {\bf Type:} location \\
-{\bf Description:} Absolute location of the directory used by the Datalog solver \bddbddb\ as its input/output directory (namely, for program domain files {\tt *.dom} and {\tt *.map}, and program relation files {\tt *.bdd}).
+{\bf Description:} Absolute location of the directory used by the Datalog solver bddbddb as its input/output directory (namely, for program domain files {\tt *.dom} and {\tt *.map}, and program relation files {\tt *.bdd}).
 \\

### 2010-12-18 main/doc/datarace.tex
cleaned up and updated documentation

@@
-\section{Datarace Analysis}
+\section{Datarace analysis}
 \label{sec:datarace}
 
-Run the datarace analysis provided in \Chord\ by running the following command in \Chord's {\tt main/} directory:
+Run the datarace analysis provided in Chord by running the following
+command in Chord's {\tt main/} directory:
 
-\begin{quote}
 \begin{verbatim}
@@
 \end{verbatim}
-\end{quote}
 
-\noindent where directory {\tt <...>} contains a file named \code{chord.properties} which
-defines properties \code{chord.main.class}, \code{chord.class.path}, and \code{chord.src.path}.
-See Section \ref{sec:chord-sysprops} for the meaning of these properties.
+\noindent where {\tt <...>} is the absolute path of a directory
+containing a file named \code{chord.properties} which defines
+properties \code{chord.main.class}, \code{chord.class.path}, and
+\code{chord.src.path}.  See Section \ref{sec:chord-sysprops} for the
+meaning of these properties.
 
-Directory \code{main/examples/datarace_test/} provides a toy Java program on which you can run the datarace
-analysis.
-First run {\tt ant} in that directory (in order to compile the program's {\tt .java} files to
-{\tt .class} files) and then run the above command in Chord's {\tt main/} directory
-with {\tt <...>} replaced by \code{examples/datarace_test/}.
-Upon successful completion, the following files should be produced in directory
+Directory \code{main/examples/datarace_test/} provides a toy Java
+program on which you can run the datarace analysis.  First run {\tt
+  ant} in that directory (in order to compile the program's {\tt
+  .java} files to {\tt .class} files) and then run the above command
+in Chord's {\tt main/} directory with {\tt <...>} replaced by
+\code{/.../examples/datarace_test/}.  Upon successful completion, the
+following files should be produced in directory
 \code{main/examples/datarace_test/chord_output/}:
@@
 \item
-File \code{dataraces_by_fld.html}, listing all dataraces grouped by the field on which they occur; all
-dataraces on the same instance field or the same static field are listed in the same group, and so are
-all dataraces on array elements.
+File \code{dataraces_by_fld.html}, listing all dataraces grouped by
+the field on which they occur; all dataraces on the same instance
+field or the same static field are listed in the same group, and so
+are all dataraces on array elements.
 \item
-File \code{dataraces_by_obj.html}, listing all dataraces grouped by the abstract object on whose field they occur;
-dataraces on all static fields are listed in the same group, and so are dataraces on different
+File \code{dataraces_by_obj.html}, listing all dataraces grouped by
+the abstract object on whose field they occur; dataraces on all static
+fields are listed in the same group, and so are dataraces on different
 instance fields of the same abstract object.

### 2010-12-18 main/doc/deadlock.tex
cleaned up and updated documentation

@@
-\section{Deadlock Analysis}
+\section{Deadlock analysis}
 \label{sec:deadlock}
 
-Run the deadlock analysis provided in Chord by running the following command in Chord's {\tt main/} directory:
+Run the deadlock analysis provided in Chord by running the following
+command in Chord's {\tt main/} directory:
 
@@
 
-\noindent where directory {\tt <...>} contains a file named \code{chord.properties} which
-defines properties \code{chord.main.class}, \code{chord.class.path}, and \code{chord.src.path}.
-See Section \ref{sec:chord-sysprops} for the meaning of these properties.
+\noindent where {\tt <...>} is the absolute path of a directory
+containing a file named \code{chord.properties} which defines
+properties \code{chord.main.class}, \code{chord.class.path}, and
+\code{chord.src.path}.  See Section \ref{sec:chord-sysprops} for the
+meaning of these properties.
 
-Directory \code{main/examples/deadlock_test/} provides a toy Java program on which you can run
-the deadlock analysis.
-First run {\tt ant} in that directory (in order to compile the program's {\tt .java} files to
-{\tt .class} files) and then run the above command in Chord's {\tt main/} directory with
-{\tt <...>} replaced by \code{examples/deadlock_test/}.
-Upon successful completion, the file \code{deadlocks.html} should be produced in directory
+Directory \code{main/examples/deadlock_test/} provides a toy Java
+program on which you can run the deadlock analysis.  First run {\tt
+  ant} in that directory (in order to compile the program's {\tt
+  .java} files to {\tt .class} files) and then run the above command
+in Chord's {\tt main/} directory with {\tt <...>} replaced by
+\code{/.../examples/deadlock_test/}.  Upon successful completion, the
+file \code{deadlocks.html} should be produced in directory
 \code{main/examples/deadlock_test/chord_output/}.

### 2010-12-18 main/doc/instr_events.tex
cleaned up and updated documentation

@@
-\section{Instrumentation Events}
+\section{Instrumentation events}
 \label{sec:instr-events}
@@
 Method entry and exit. \\
-{\bf Event:} \code{ENTER_METHOD} $m$ $t$ \\
-{\bf Desc.:} Generated after thread $t$ enters method $m$ respectively. \\
-{\bf Event:} \code{LEAVE_METHOD} $m$ $t$ \\
-{\bf Desc.:} Generated before thread $t$ leaves method $m$ respectively.
+{\bf Event:} \code{ENTER_METHOD} m t \\
+{\bf Desc.:} Generated after thread t enters method m respectively. \\
+{\bf Event:} \code{LEAVE_METHOD} m t \\
+{\bf Desc.:} Generated before thread t leaves method m respectively.
 
@@
 Loop entry and exit. \\
-{\bf Event:} \code{ENTER_LOOP} $w$ $t$ \\
-{\bf Desc.:} Generated before thread $t$ enters loop $w$ respectively. \\
-{\bf Event:} \code{LEAVE_LOOP} $w$ $t$ \\
-{\bf Desc.:} Generated after thread $t$ leaves loop $w$ respectively.
+{\bf Event:} \code{ENTER_LOOP} w t \\
+{\bf Desc.:} Generated before thread t enters loop w respectively. \\
+{\bf Event:} \code{LEAVE_LOOP} w t \\
+{\bf Desc.:} Generated after thread t leaves loop w respectively.
 
@@
 Object allocation \\
-{\bf Event:} \code{BEF_NEW} $h$ $t$ \\
-{\bf Desc.:} Generated only in the crude trace before thread $t$ executes a \code{new} bytecode instruction at program point $h$. \\
-{\bf Event:} \code{AFT_NEW} $h$ $t$ $o$ \\
-{\bf Desc.:} Generated only in the crude trace after thread $t$ executes a \code{new} bytecode instruction at program point $h$ and allocates new object $o$. \\
-%Note: Both above events are visible only in the crude trace and not in the final trace. They are generated only if argument \code{hasObj} of method {@link #setNewAndNewArrayEvent} is true.
-{\bf Event:} \code{NEW} $h$ $t$ $o$ \\
-{\bf Desc.:} Generated after thread $t$ executes a \code{new} bytecode instruction at program point $h$ and allocates new object $o$. \\
-{\bf Event:} \code{NEW_ARRAY} $h$ $t$ $o$ \\
-{\bf Desc.:} Generated after thread $t$ executes a \code{newarray} bytecode instruction at program point $h$ and allocates new object $o$.
+{\bf Event:} \code{BEF_NEW} h t \\
+{\bf Desc.:} Generated before thread t executes a \code{new} bytecode instruction at program point h and allocates new object o. \\
+{\bf Event:} \code{AFT_NEW} h t o \\
+{\bf Desc.:} Generated after thread t executes a \code{new} bytecode instruction at program point h and allocates new object o. \\
+{\bf Event:} \code{NEW_ARRAY} h t o \\
+{\bf Desc.:} Generated after thread t executes a \code{newarray} bytecode instruction at program point h and allocates new object o.
 
@@
 Getstatic primitive \\
-{\bf Event:} \code{GETSTATIC_PRIMITIVE} $e$ $t$ $f$ \\
-{\bf Desc.:} Generated after thread $t$ reads primitive-typed static field $f$ at program point $e$.
+{\bf Event:} \code{GETSTATIC_PRIMITIVE} e t f \\
+{\bf Desc.:} Generated after thread t reads primitive-typed static field f at program point e.
 
@@
 Getstatic reference \\
-{\bf Event:} \code{GETSTATIC_REFERENCE} $e$ $t$ $f$ $o$ \\
-{\bf Desc.:} Generated after thread $t$ reads object $o$ from reference-typed static field $f$ at program point $e$.
+{\bf Event:} \code{GETSTATIC_REFERENCE} e t f o \\
+{\bf Desc.:} Generated after thread t reads object o from reference-typed static field f at program point e.
 
@@
 Putstatic primitive \\
-{\bf Event:} \code{PUTSTATIC_PRIMITIVE} $e$ $t$ $f$ \\
-{\bf Desc.:} Generated after thread $t$ writes primitive-typed static field $f$ at program point $e$.
+{\bf Event:} \code{PUTSTATIC_PRIMITIVE} e t f \\
+{\bf Desc.:} Generated after thread t writes primitive-typed static field f at program point e.
 
@@
 Putstatic reference \\
-{\bf Event:} \code{PUTSTATIC_REFERENCE} $e$ $t$ $f$ $o$ \\
-{\bf Desc.:} Generated after thread $t$ writes object $o$ to reference-typed static field $f$ at program point $e$.
+{\bf Event:} \code{PUTSTATIC_REFERENCE} e t f o \\
+{\bf Desc.:} Generated after thread t writes object o to reference-typed static field f at program point e.
 
@@
 Getfield primitive \\
-{\bf Event:} \code{GETFIELD_PRIMITIVE} $e$ $t$ $b$ $f$ \\
-{\bf Desc.:} Generated after thread $t$ reads primitive-typed instance field $f$ of object $b$ at program point $e$.
+{\bf Event:} \code{GETFIELD_PRIMITIVE} e t b f \\
+{\bf Desc.:} Generated after thread t reads primitive-typed instance field f of object b at program point e.
 
@@
 Getfield reference \\
-{\bf Event:} \code{GETFIELD_REFERENCE} $e$ $t$ $b$ $f$ $o$ \\
-{\bf Desc.:} Generated after thread $t$ reads object $o$ from reference-typed instance field $f$ of object $b$ at program point $e$.
+{\bf Event:} \code{GETFIELD_REFERENCE} e t b f o \\
+{\bf Desc.:} Generated after thread t reads object o from reference-typed instance field f of object b at program point e.
 
@@
 Putfield primitive \\
-{\bf Event:} \code{PUTFIELD_PRIMITIVE} $e$ $t$ $b$ $f$ \\
-{\bf Desc.:} Generated after thread $t$ writes primitive-typed instance field $f$ of object $b$ at program point $e$.
+{\bf Event:} \code{PUTFIELD_PRIMITIVE} e t b f \\
+{\bf Desc.:} Generated after thread t writes primitive-typed instance field f of object b at program point e.
 
@@
 Putfield reference \\
-{\bf Event:} \code{PUTFIELD_REFERENCE} $e$ $t$ $b$ $f$ $o$ \\
-{\bf Desc.:} Generated after thread $t$ writes object $o$ to reference-typed instance field $f$ of object $b$ at program point $e$.
+{\bf Event:} \code{PUTFIELD_REFERENCE} e t b f o \\
+{\bf Desc.:} Generated after thread t writes object o to reference-typed instance field f of object b at program point e.
 
@@
 Aload primitive \\
-{\bf Event:} \code{ALOAD_PRIMITIVE} $e$ $t$ $b$ $i$ \\
-{\bf Desc.:} Generated after thread $t$ reads the primitive-typed element at index $i$ of array object $b$ at program point $e$.
+{\bf Event:} \code{ALOAD_PRIMITIVE} e t b i \\
+{\bf Desc.:} Generated after thread t reads the primitive-typed element at index i of array object b at program point e.
 
@@
 Aload reference \\
-{\bf Event:} \code{ALOAD_REFERENCE} $e$ $t$ $b$ $i$ $o$ \\
-{\bf Desc.:} Generated after thread $t$ reads object $o$ from the reference-typed element at index $i$ of array object $b$ at program point $e$.
+{\bf Event:} \code{ALOAD_REFERENCE} e t b i o \\
+{\bf Desc.:} Generated after thread t reads object o from the reference-typed element at index i of array object b at program point e.
 
@@
 Astore primitive \\
-{\bf Event:} \code{ASTORE_PRIMITIVE} $e$ $t$ $b$ $i$ \\
-{\bf Desc.:} Generated after thread $t$ writes the primitive-typed element at index $i$ of array object $b$ at program point $e$.
+{\bf Event:} \code{ASTORE_PRIMITIVE} e t b i \\
+{\bf Desc.:} Generated after thread t writes the primitive-typed element at index i of array object b at program point e.
 
@@
 Astore reference \\
-{\bf Event:} \code{ASTORE_REFERENCE} $e$ $t$ $b$ $i$ $o$ \\
-{\bf Desc.:} Generated after thread $t$ writes object $o$ to the reference-typed element at index $i$ of array object $b$ at program point $e$.
+{\bf Event:} \code{ASTORE_REFERENCE} e t b i o \\
+{\bf Desc.:} Generated after thread t writes object o to the reference-typed element at index i of array object b at program point e.
 
@@
 Method call \\
-{\bf Event:} \code{METHOD_CALL} $i$ $t$ \\
-{\bf Desc.:} Generated before or after thread $t$ executes the method invocation statement at program point $i$.
+{\bf Event:} \code{BEF_METHOD_CALL} i t o \\
+{\bf Desc.:} Generated before thread t executes the method invocation statement at program point i with this argument as object o. \\
+{\bf Event:} \code{AFT_METHOD_CALL} i t o \\
+{\bf Desc.:} Generated after thread t executes the method invocation statement at program point i with this argument as object o. \\
+{\bf Note:} These events are not generated before/after constructor calls; use \code{BEF_NEW} and \code{AFT_NEW} events for that purpose.
 
@@
 Thread start call \\
-{\bf Event:} \code{THREAD_START} $i$ $t$ $o$ \\
-{\bf Desc.:} Generated before thread $t$ calls the \code{start()} method of \code{java.lang.Thread} at program point $i$ and spawns a thread $o$.
+{\bf Event:} \code{THREAD_START} i t o \\
+{\bf Desc.:} Generated before thread t calls the \code{start()} method of \code{java.lang.Thread} at program point i and spawns a thread o.
 
@@
 Thread join call \\
-{\bf Event:} \code{THREAD_JOIN} $i$ $t$ $o$ \\
-{\bf Desc.:} Generated before thread $t$ calls the \code{join()} method of \code{java.lang.Thread} at program point i to join with thread $o$.
+{\bf Event:} \code{THREAD_JOIN} i t o \\
+{\bf Desc.:} Generated before thread t calls the \code{join()} method of \code{java.lang.Thread} at program point i to join with thread o.
 
@@
 Lock acquire \\
-{\bf Event:} \code{ACQUIRE_LOCK} $l$ $t$ $o$ \\
-{\bf Desc.:} Generated after thread $t$ executes a statement of the form monitorenter $o$ or enters a method synchronized on $o$ at program point $l$.
+{\bf Event:} \code{ACQUIRE_LOCK} l t o \\
+{\bf Desc.:} Generated after thread t executes a statement of the form monitorenter o or enters a method synchronized on o at program point l.
 
@@
 Lock release \\
-{\bf Event:} \code{RELEASE_LOCK} $r$ $t$ $o$ \\
+{\bf Event:} \code{RELEASE_LOCK} r t o \\
 {\bf Desc.:} Generated before thread t executes a statement of the form monitorexit o or leaves a method synchronized on o at program point r.
@@
 Thread wait call \\
-{\bf Event:} \code{WAIT} $i$ $t$ $o$ \\
-{\bf Desc.:} Generated before thread $t$ calls the \code{wait()} method of \code{java.lang.Object} at program point $i$ on object $o$.
+{\bf Event:} \code{WAIT} i t o \\
+{\bf Desc.:} Generated before thread t calls the \code{wait()} method of \code{java.lang.Object} at program point i on object o.
 
@@
 Thread notify call \\
-{\bf Event:} \code{NOTIFY} $i$ $t$ $o$ \\
-{\bf Desc.:} Generated before thread $t$ calls the \code{notify()} or \code{notifyAll()} method of \code{java.lang.Object} at program point $i$ on object $o$.
+{\bf Event:} \code{NOTIFY} i t o \\
+{\bf Desc.:} Generated before thread t calls the \code{notify()} or \code{notifyAll()} method of \code{java.lang.Object} at program point i on object o.
 \end{enumerate}

### 2010-12-18 main/doc/writing_datalog_analysis.tex
cleaned up and updated documentation

@@
-\section{Writing a Datalog Analysis}
+\section{Writing a Datalog analysis}
 \label{sec:writing-datalog-analysis}
@@
 A backslash may be used at the end of a line to break long rules for readability.
-The Datalog analysis solver \bddbddb\ used in \Chord\ does not apply any
+The Datalog analysis solver bddbddb used in Chord does not apply any
 sophisticated optimizations to simplify the rules; besides the BDD variable order,

### 2011-03-04 main/doc/building_scope.tex
documentation

@@
-\section{Computing Program Scope}
+\section{Computing program scope}
 \label{sec:building-scope}
@@
 
-\subsection{Reusing Program Scope}
+\subsection{Scope reuse}
 
@@
 
-\subsection{Scope Construction Algorithms}
+\subsection{Scope construction algorithms}
 
@@
 statically using Rapid Type Analysis (RTA).
-
 RTA is an iterative fixed-point algorithm.  It maintains a set of
@@
 
-%RTA is a relatively inexpensive and precise algorithm in practice.  Its key shortcoming is that it makes
-%no attempt to resolve reflection, which is rampant in real-world Java programs, and can therefore be
-%unsound (i.e., underestimate the set of reachable classes and methods).
-%The next option attempts to overcome this problem.
-%
-%\item
-%The \code{rta_reflect} value instructs Chord to compute program scope statically using Rapid Type Analysis
-%and, moreover, to resolve a common reflection pattern:
-%
-%\begin{quote}
-%\begin{verbatim}
-%String s = ...;
-%Class c = Class.forName(s);
-%Object o = c.newInstance();
-%T t = (T) o;
-%\end{verbatim}
-%\end{quote}
-%
-%This analysis is identical to RTA except that it additionally inspects every cast statement in the program,
-%such as the last statement in the above snippet,
-%and queries the class hierarchy to find all concrete classes that subclass \code{T} (if \code{T} is a class)
-%or that implement \code{T} (if \code{T} is an interface).
-%Chord allows users to control which classes are included in the class hierarchy (see below).
-%
+%RTA is a relatively inexpensive and precise algorithm in practice.
+%Its key shortcoming is that it makes no attempt to resolve
+%reflection, which is rampant in real-world Java programs, and can
+%therefore be unsound (i.e., underestimate the set of reachable
+%classes and methods).  The next option attempts to overcome this
+%problem.  \item The \code{rta_reflect} value instructs Chord to
+%compute program scope statically using Rapid Type Analysis and,
+%moreover, to resolve a common reflection
+%pattern: \begin{quote} \begin{verbatim} String s = ...; Class c =
+%Class.forName(s); Object o = c.newInstance(); T t = (T)
+%o; \end{verbatim} \end{quote} This analysis is identical to RTA
+%except that it additionally inspects every cast statement in the
+%program, such as the last statement in the above snippet, and queries
+%the class hierarchy to find all concrete classes that subclass
+%\code{T} (if \code{T} is a class) or that implement \code{T} (if
+%\code{T} is an interface).  Chord allows users to control which
+%classes are included in the class hierarchy (see Section
+%\ref{sec:cha}).
 \item
@@
 statically using Class Hierarchy Analysis (CHA).
-
 The key difference between CHA and RTA is that for invokevirtual and
@@
 Nevertheless, Chord allows users to control which classes are
-included in the class hierarchy (see below) and thereby control the
+included in the class hierarchy (see \ref{sec:cha}) and thereby control the
 precision and cost of CHA.
@@
 
-\subsection{Class Hierarchy Construction}
+\subsection{Class hierarchy construction}
+\label{sec:cha}
 
@@
 
-\subsection{Scope Exclusion}
+\subsection{Scope exclusion}
 
-%TODO: describe chord.scope.exclude?
+For tractability, Chord can be instructed to treat the bodies of certain
+methods as no-ops.  Chord provides three properties for this purpose.
+The value of each of these properties is a comma-separated list of prefixes
+of names of classes all of whose declared methods must be treated as no-ops:
+
+\begin{itemize}
+\item
+Property \code{chord.std.scope.exclude} conventionally specifies
+classes to be excluded from the scope of {\it all} programs to be
+analyzed, e.g., classes in the JDK standard library.  Its default
+value is \code{[chord.main.class.path]}, i.e., the classes comprising
+Chord itself.
+\item
+Property \code{chord.ext.scope.exclude} conventionally specifies
+classes to be excluded from the scope of specific programs to be
+analyzed.  Its default value is the empty list.
+\item
+Property \code{chord.scope.exclude} specifies the final list of
+classes to be excluded from scope.  Its default value is simply the
+concatenation of the values of properties
+\code{chord.std.scope.exclude} and \code{chord.ext.scope.exclude}.
+\end{itemize}
 

### 2011-03-04 main/doc/download_and_installation.tex
documentation

@@
-\section{Download and Installation}
+\section{Download and installation}
 \label{sec:download-and-installation}
@@
 \item
-JDK 5 or higher, e.g. from
-\xlink{IBM}{http://www.ibm.com/developerworks/java/jdk/} or
-\xlink{Sun}{http://java.sun.com/javase/downloads/index.jsp}
+A JVM with JDK 5 or higher, e.g.
+\xlink{IBM J9}{http://www.ibm.com/developerworks/java/jdk/} or
+\xlink{Oracle HotSpot}{http://www.oracle.com/technetwork/java/javase/downloads/index.html}
 \item
 \xlink{Apache Ant}{http://ant.apache.org/}, a Java-based build tool
-\item
-a C++ compiler, e.g. \xlink{GCC}{http://gcc.gnu.org/}
-\item
-a Make utility, e.g. \xlink{GNU Make}{http://www.gnu.org/software/make/}
-\item
-\xlink{Cygwin}{http://www.cygwin.com/}, if it is a Windows machine
 \end{itemize}
 
-Download the latest Chord source code from the SVN repository:
+{\bf Note:} If you intend to use Chord for dynamic analysis, then
+IBM J9 is highly recommended, as it allows you to instrument
+the entire JDK; using any other JVM will likely require excluding
+large parts of the JDK from being instrumented.
+
+Obtain the latest stable release of Chord from \url{http://jchord.googlecode.com/checkout/}
+or the latest development snapshot from the SVN repository:
 
@@
 \begin{verbatim}
+    prompt> ant help
+\end{verbatim}
+
+\noindent To compile Chord, simply run the following command in the same
+directory:
+
+\begin{verbatim}
     prompt> ant 
@@
 
-\noindent To compile Chord, run the following command in the same
-directory:
-
-\begin{verbatim}
-    prompt> ant compile
-\end{verbatim}
-
-\noindent This will compile the following:
+\noindent This will compile
+the Java source code of Chord from \code{main/src/} to Java bytecode
+in \code{main/classes/}.
+Additionally:
 \begin{itemize}
 \item
-the Java source code of Chord from \code{main/src/} to Java bytecode
-in \code{main/classes/}.
-\item
-the C source code of BDD library
+If system property \code{chord.use.buddy} is set to true (default is false),
+then the C source code of BDD library
 \xlink{BuDDy}{http://buddy.sourceforge.net/} from
-\code{main/bdd/} to a shared library in \code{main/lib/}
+\code{main/bdd/} will be compiled to a shared library in \code{main/lib/}
 (\code{libbuddy.so} on Linux, \code{buddy.dll} on Windows, and
-\code{libbuddy.dylib} on MacOS); this library is needed for executing
+\code{libbuddy.dylib} on MacOS); this library is used for executing
 program analyses written in Datalog using
@@
 \item
-the C++ source code of the Chord instrumentation agent from
-\code{main/agent/} to a shared library in \code{main/lib/}
+If system property \code{chord.use.jvmti} is set to true (default is false),
+then the C++ source code of the Chord instrumentation agent from
+\code{main/agent/} will be compiled to a shared library in \code{main/lib/}
 (\code{libchord_instr_agent.so} on all architectures); this agent is
-needed for computing program scope dynamically and for executing
+used for computing program scope dynamically and for executing
 dynamic program analyses.
 \end{itemize}
+The above system properties can be set either in file \code{main/chord.properties}
+(recommended) or on the command line as follows:
+
+\begin{verbatim}
+    prompt> ant -Dchord.use.buddy=true -Dchord.use.jvmti=true
+\end{verbatim}

### 2011-03-04 main/doc/whatis_chord.tex
documentation

@@
 
-Chord is a program analysis platform for Java designed with the
-primary goal of helping advance the state-of-the-art in the field by
-enabling researchers to productively design, implement, and evaluate a
-broad variety of program analyses. To this end, Chord has the
-following key characteristics:
+Chord is a program analysis platform that enables 
+productively design, implement, evaluate, and combine a
+broad variety of program analyses for Java bytecode. It has the
+following key features:
 
@@
 \item
-Stand-alone: it provides various off-the-shelf analyses (e.g., various
+{\bf Stand-alone:} It provides various off-the-shelf analyses (e.g., various
 call-graph analyses; points-to analyses; thread-escape analysis;
@@
 races, deadlocks, and atomicity violations; etc.)
+
 \item
-Extensible: it allows users to build their own analyses atop the
-provided analyses.
-\item
-Compositional: it allows users to write each analysis independently
-yet allow it to interact in complex ways with other analyses by
-specifying lightweight data- and control-dependencies.
-\item
-Efficient: It executes analyses in a demand-driven fashion, caches
-results computed by each analysis for reuse by other analyses without
-re-computation, and can execute analyses without dependencies between
-them in parallel.
-\item
-Deterministic: It guarantees that the result is the same regardless of
-the order in which different analyses are executed; moreover, results
-can be shared across different runs.
-\item
-Flexible: It allows a broad range of analyses to be expressed,
+{\bf Flexible:} It allows users to express a broad range of analyses,
 including both static and dynamic analyses, analyses written
@@
 combined static/dynamic analyses.
+
+\item
+{\bf Efficient:} It executes analyses in a demand-driven fashion, caches
+results computed by each analysis for reuse by other analyses without
+re-computation, and can execute analyses without dependencies between
+them in parallel.
+
+\item
+{\bf Deterministic:} It guarantees that the result is the same regardless of
+the order in which different analyses are executed; moreover, results
+can be shared across different runs.
 \end{itemize}
@@
 Improvements by users are welcome and encouraged.  The project website
-is \url{http://berkeley.intel-research.net/chord/} and the software
-website is \url{http://code.google.com/p/jchord/}.
+is \url{http://chord.stanford.edu/} and the software development website is
+\url{http://jchord.googlecode.com/}.

### 2011-03-10 main/doc/chord_sysprops.tex
added more documentation

@@
 
-\section{Chord system properties}
+\section{Chord System Properties}
 \label{sec:chord-sysprops}
 
-Users can use system properties to control Chord's functionality.
+Users can use system properties to control Chord's execution.
 The following properties are recognized by Chord.
@@
 
-\subsection{Basic Program Properties}
+\paragraph{Program Properties}
 
-This section describes basic properties of the given program that are required by program analyses in Chord,
-such as its main class, the location(s) of its class files and Java source files, and command-line arguments
-to be used to run the program.
+This section describes basic properties of the given program, such as
+its main class, the location(s) of its class files and Java source
+files, and command-line arguments to be used to run the program.
 
-\begin{itemize}
-\item
-\code{chord.main.class} \\
+\noindent \code{chord.work.dir}
+\begin{quote}
+{\bf Type:} location \\
+{\bf Description:} Working directory during Chord's execution.  This is
+usually the top-level directory of the given program. \\
+{\bf Default value:} current working directory
+\end{quote}
+
+\noindent \code{chord.props.file}
+\begin{quote}
+{\bf Type:} location \\
+{\bf Description:} Properties file loaded by Chord at the
+beginning before doing anything else.  Any of the below properties may
+be defined in this file to avoid defining them on the command line
+(using the \code{-D<key>=<value>} format) every time Chord is run.
+Each relative file/directory name in the value of any property defined
+in this file is treated relative to Chord's working directory (which
+is specified by property \code{chord.work.dir}). \\
+{\bf Default value:} \code{[chord.work.dir]/chord.properties}
+\end{quote}
+
+\noindent \code{chord.main.class}
+\begin{quote}
 {\bf Type:} class \\
-{\bf Description:} Fully-qualified name of the main class of the given program (e.g., \code{com.example.Main}). \\
-{\bf Note:} This property is required by virtually every task Chord performs.
+{\bf Description:} Fully-qualified name of the main class of the given program (e.g., \code{com.example.Main}).
+\end{quote}
 
-\item
-\code{chord.class.path} \\
+\noindent \code{chord.class.path}
+\begin{quote}
 {\bf Type:} path \\
-{\bf Description:} Classpath of the given program.
-It does not need to include boot classes (i.e., classes in \code{[sun.boot.class.path]}) or standard extensions (i.e., classes in {\tt .jar} files in directory \code{[java.home]/lib/ext/}).
-\\
-{\bf Default value:} {\tt ""} \\
-{\bf Note:} This property is required by virtually every task Chord performs.
+{\bf Description:} Classpath of the given program.  It does not need to include
+boot classes (i.e., classes in \code{[sun.boot.class.path]}) or
+standard extensions (i.e., classes in {\tt .jar} files in directory
+\code{[java.home]/lib/ext/}). \\
+{\bf Default value:} {\tt ""}
+\end{quote}
 
-\item
-\code{chord.src.path} \\
+\noindent \code{chord.src.path}
+\begin{quote}
 {\bf Type:} path \\
@@
 This property is used only by the task of converting Java source files into HTML files by program analyses that need to present their analysis results at the Java source code level (by calling method \code{chord.program.Program.g().HTMLizeJavaSrcFiles())}.
+\end{quote}
 
-\item
-\code{chord.run.ids} \\
+\noindent \code{chord.run.ids}
+\begin{quote}
 {\bf Type:} string list \\
@@
 {\bf Note:} This property is used only when Chord executes the given program, namely, when it is asked to compute the program scope dynamically (i.e., when \code{[chord.scope.kind]=dynamic}) or when it is asked to execute a dynamic program analysis.
+\end{quote}
 
-\item
-\code{chord.args.<id>} \\
+\noindent \code{chord.args.<id>}
+\begin{quote}
 {\bf Type:} string \\
@@
 {\bf Note:} This property is used only when Chord executes the given program, namely, when it is asked to compute the program scope dynamically (i.e., when \code{[chord.scope.kind]=dynamic}) or when it is asked to execute a dynamic program analysis.
+\end{quote}
 
-\item
-\code{chord.runtime.jvmargs} \\
+\noindent \code{chord.runtime.jvmargs}
+\begin{quote}
 {\bf Type:} string \\
@@
 {\bf Note:} This property is used only when Chord executes the given program, namely, when it is asked to compute the program scope dynamically (i.e., when \code{[chord.scope.kind]=dynamic}) or when it is asked to execute a dynamic program analysis.
-\end{itemize}
+\end{quote}
 
-\subsection{Chord Task Properties}
-\label{sec:chord-sysprops:task}
+\paragraph{Analysis Scope Properties}
 
-This section describes properties that specify what task(s) Chord must perform, 
-such as computing the scope of the given program or running program analyses on the given program.
+This section describes properties that specify how the analysis scope of the given program is computed.
+See Section \ref{sec:building-scope} for more details.
 
-\begin{itemize}
-\item
-\code{chord.build.scope} \\
+\noindent \code{chord.scope.kind}
+\begin{quote}
+{\bf Type:} {\tt [dynamic|rta|cha]} \\
+{\bf Description:} Algorithm to compute program scope.  The choices are {\tt dynamic} (dynamic analysis), {\tt rta} (Rapid Type Analysis), and {\tt cha} (Class Hierarchy Analysis). \\
+{\bf Default value:} {\tt rta}
+\end{quote}
+
+\noindent \code{chord.reflect.kind}
+\begin{quote}
+{\bf Type:} {\tt [none|static|dynamic|static_cast]} \\
+{\bf Description:} TODO \\
+{\bf Default value:} {\tt none}
+\end{quote}
+
+\noindent \code{chord.ch.kind}
+\begin{quote}
+{\bf Type:} {\tt [static|dynamic]} \\
+{\bf Description:} Algorithm to build the class hierarchy.
+If it is {\tt dynamic}, then the given program is executed and classes not loaded by the JVM while running the program are excluded while building the class hierarchy.
+\\
+{\bf Default value:} {\tt static} \\
+{\bf Note:} This property is relevant only if \code{chord.scope.kind} is {\tt cha} since only this scope computing algorithm queries the class hierarchy. 
+\end{quote}
+
+\noindent \code{chord.ssa}
+\begin{quote}
+{\bf Type:} bool  \\
+{\bf Description:} Do SSA transformation for all methods deemed reachable by the algorithm used to compute program scope. \\
+{\bf Default value:} {\tt true}
+\end{quote}
+
+% TODO: mention that below two properties are also used by instrumentor to decide which classes to exclude from instrumentation
+
+\noindent \code{chord.std.scope.exclude}
+\begin{quote}
+{\bf Type:} string list \\
+{\bf Description:} Partial list of prefixes of names of classes whose declared methods must be treated as no-ops.
+Conventionally, this property is used to specify classes to be excluded from the scope of all programs to be analyzed, e.g., classes in the JDK standard library.
+\\
+{\bf Default value:} {\tt ""}
+\end{quote}
+
+\noindent \code{chord.ext.scope.exclude}
+\begin{quote}
+{\bf Type:} string list \\
+{\bf Description:} Partial list of prefixes of names of classes whose declared methods must be treated as no-ops.
+Conventionally, this property is used to specify classes to be excluded from the scope of specific programs to be analyzed.
+\\
+{\bf Default value:} {\tt ""}
+\end{quote}
+
+\noindent \code{chord.scope.exclude}
+\begin{quote}
+{\bf Type:} string list \\
+{\bf Description:} Complete list of prefixes of names of classes whose declared methods must be treated as no-ops. \\
+{\bf Default value:} \code{"[chord.std.scope.exclude],[chord.ext.scope.exclude]"}
+\end{quote}
+
+\noindent \code{chord.std.check.exclude}
+\begin{quote}
+{\bf Type:} string list \\
+{\bf Description:} List of prefixes of names of classes and packages inside the standard library to be excluded by program analyses.  Interpretation of this property is analysis-specific. \\
+{\bf Default value:} \code{"java.,javax.,sun.,com.sun.,com.ibm.,} \code{org.apache.harmony."}
+\end{quote}
+
+\noindent \code{chord.ext.check.exclude}
+\begin{quote}
+{\bf Type:} string list \\
+{\bf Description:} List of prefixes of names of classes and packages outside the standard library to be excluded by program analyses.  Interpretation of this property is analysis-specific. \\
+{\bf Default value:} {\tt ""}
+\end{quote}
+
+\noindent \code{chord.check.exclude}
+\begin{quote}
+{\bf Type:} string list \\
+{\bf Description:} List of prefixes of names of classes and packages to be excluded by program analyses.  Interpretation of this property is analysis-specific. \\
+{\bf Default value:} \code{"[chord.std.check.exclude],[chord.ext.check.exclude]"}
+\end{quote}
+
+\paragraph{Functionality Properties}
+
+This section describes properties that dictate what Chord outputs during execution.
+
+\noindent \code{chord.build.scope}
+\begin{quote}
 {\bf Type:} bool \\
@@
 {\bf Note:} The program scope is computed regardless of the value of this property if another task (e.g., a program analysis specified via property \code{chord.run.analyses}) demands it by calling method \code{chord.program.Program.g()}.
+\end{quote}
 
-\item
-\code{chord.run.analyses} \\
+\noindent \code{chord.run.analyses}
+\begin{quote}
 {\bf Type:} string list  \\
@@
 If the analysis is written in Datalog, its name is specified via a line of the form ``{\tt \# name=<...>}", and is the absolute location of the file itself if this line is missing.
+\end{quote}
 
-\item
-\code{chord.print.classes} \\
-{\bf Type:} string list \\
-{\bf Description:} List of classes in scope whose intermediate representation to print to standard output. \\
-{\bf Default value:} {\tt ""} \\
-{\bf Note:} In class names, use `.' instead of `/', and use {\tt \#} instead of the dollar character.
-
-\item
-\code{chord.print.methods} \\
+\noindent \code{chord.print.methods}
+\begin{quote}
 {\bf Type:} string list  \\
@@
 In {\tt <cname>}, use `.' instead of `/', and use {\tt \#} instead of the dollar character.
+\end{quote}
 
-\item
-\code{chord.print.all.classes} \\
+\noindent \code{chord.print.classes}
+\begin{quote}
+{\bf Type:} string list  \\
+{\bf Description:} List of classes in scope whose intermediate representation to print to standard output. \\
+{\bf Default value:} {\tt ""} \\
+{\bf Note:} In class names, use `.' instead of `/', and use {\tt \#} instead of the dollar character. 
+\end{quote}
+
+\noindent \code{chord.print.all.classes}
+\begin{quote}
 {\bf Type:} bool \\
@@
 {\bf Default value:} {\tt false}
+\end{quote}
 
-\item
-\code{chord.print.rels} \\
+\noindent \code{chord.print.rels}
+\begin{quote}
 {\bf Type:} string list  \\
@@
 {\bf Default value:} {\tt ""} \\
-{\bf Note:} This task must be used with caution as certain program relations, albeit represented compactly as BDDs, may contain a large number (e.g., billions) of tuples, resulting in voluminous output when printed to a file.
+{\bf Note:} This functionality must be used with caution as certain program relations, albeit represented compactly as BDDs, may contain a large number (e.g., billions) of tuples, resulting in voluminous output when printed to a file.
 See Section \ref{sec:tuning-datalog-analysis} for a more efficient way to query the contents of program relations (namely, by using the {\tt debug} target provided in file {\tt build.xml} in Chord's {\tt main/} directory).
+\end{quote}
 
-\item
-\code{chord.print.project} \\
+\noindent \code{chord.print.project}
+\begin{quote}
 {\bf Type:} bool \\
@@
 {\bf Default value:} {\tt false}
-\end{itemize}
+\end{quote}
 
-\subsection{Chord Boot Properties}
+\noindent \code{chord.print.results}
+\begin{quote}
+{\bf Type:} bool \\
+{\bf Description:} Print the results of program analyses in HTML.  Interpretation of this property is analysis-specific.  \\
+{\bf Default value:} {\tt true}
+\end{quote}
 
-This section describes properties of the JVM running Chord, such as the working directory and various memory limits.
+\noindent \code{chord.verbose}
+\begin{quote}
+{\bf Type:} int in the range [0..6]  \\
+{\bf Description:} Control the verbosity of logging during Chord's execution.  \\
+{\bf Default value:} {\tt 2}
+\end{quote}
 
-\begin{itemize}
-\item
-\code{chord.work.dir} \\
-{\bf Type:} location \\
-{\bf Description:} Working directory during Chord's execution. \\
-{\bf Default value:} current working directory
-
-\item
-\code{chord.props.file} \\
-{\bf Type:} location \\
-{\bf Description:} Properties file loaded by \code{[chord.main.dir]/build.xml}.
-Any of the below properties, as well as other user-defined properties to be passed to Chord (e.g., for user-defined analyses), may be set in this file.
-Each relative (instead of absolute) path element in the value of any property named \code{chord.<...>.path} set in this file is converted to an absolute path element with respect to the directory containing this file.
-\\
-{\bf Default value:} \code{[chord.work.dir]/chord.properties}
-
-\item
-\code{chord.max.heap} \\
-{\bf Type:} string \\
-{\bf Description:} Maximum memory size of JVM running Chord. \\
-{\bf Default value:} {\tt 1024m}
-
-\item 
-\code{chord.max.stack} \\
-{\bf Type:} string \\
-{\bf Description:} Maximum thread stack size of JVM running Chord. \\
-{\bf Default value:} {\tt 32m}
-
-\item
-\code{chord.jvmargs} \\
-{\bf Type:} string \\
-{\bf Description:} Arguments to JVM running Chord. \\
-{\bf Default value:} \code{"-showversion} \code{-ea} \code{-Xmx[chord.max.heap]} \code{-Xss[chord.max.stack]"}
-
-\item
-\code{chord.bddbddb.max.heap} \\
-{\bf Type:} string \\
-{\bf Description:} Maximum memory size of JVM running bddbddb. \\
-{\bf Default value:} {\tt 1024m} \\
-{\bf Note:} Each program analysis written in Datalog is run in a separate JVM because there may be multiple invocations of the Datalog solver bddbddb in a single run of Chord and it is difficult to reset the global state of bddbddb on each invocation.
-\end{itemize}
-
-\subsection{Program Scope Properties}
-\label{sec:scope-props}
-
-This section describes properties that specify how the program scope is computed.
-See Section \ref{sec:building-scope} for more details.
-
-\begin{itemize}
-\item
-\code{chord.reuse.scope} \\
-{\bf Type:} bool \\
-{\bf Description:} Compute program scope using the information in files specified by properties \code{chord.methods.file} and \code{chord.reflect.file}.
-Property \code{chord.scope.kind} is ignored if this property is set to {\tt true} and the two files exist.
-\\
-{\bf Default value:} {\tt false}
-
-\item
-\code{chord.scope.kind} \\
-{\bf Type:} {\tt [dynamic|rta|cha]} \\
-{\bf Description:} Algorithm to compute program scope.  The choices are {\tt dynamic} (dynamic analysis), {\tt rta} (Rapid Type Analysis), and {\tt cha} (Class Hierarchy Analysis). \\
-{\bf Default value:} {\tt rta}
-
-\item
-\code{chord.ch.kind} \\
-{\bf Type:} {\tt [static|dynamic]} \\
-{\bf Description:} Algorithm to build the class hierarchy.
-If it is {\tt dynamic}, then the given program is executed and classes not loaded by the JVM while running the program are excluded while building the class hierarchy.
-This property is relevant only if \code{chord.scope.kind} is {\tt cha} since only this scope computing algorithm queries the class hierarchy.
-\\
-{\bf Default value:} {\tt static}
-
-% TODO: mention that below two properties are also used by instrumentor to decide which classes to exclude from instrumentation
-
-\item
-\code{chord.std.scope.exclude} \\
-{\bf Type:} string list \\
-{\bf Description:} Partial list of prefixes of names of classes whose declared methods must be treated as no-ops.  Conventionally, this property is used to specify classes to be
-excluded from the scope of all programs to be analyzed, e.g., classes in the JDK standard library.  \\
-{\bf Default value:} \code{[chord.main.class.path]}
-
-\item
-\code{chord.ext.scope.exclude} \\
-{\bf Type:} string list \\
-{\bf Description:} Partial list of prefixes of names of classes whose declared methods must be treated as no-ops.  Conventionally, this property is used to specify classes to be
-excluded from the scope of specific programs to be analyzed. \\
-{\bf Default value:} {\tt ""}
-
-\item
-\code{chord.scope.exclude} \\
-{\bf Type:} string list \\
-{\bf Description:} Complete list of prefixes of names of classes whose declared methods must be treated as no-ops. \\
-{\bf Default value:} \code{"[chord.std.scope.exclude],[chord.ext.scope.exclude]"}
-\end{itemize}
-
-\subsection{Program Analysis Properties}
+\paragraph{Project Properties}
 
@@
 
-\begin{itemize}
-\item
-\code{chord.java.analysis.path} \\
+\noindent \code{chord.classic}
+\begin{quote}
+{\bf Type:} bool \\
+{\bf Description:} TODO \\
+{\bf Default value:} TODO
+\end{quote}
+
+\noindent \code{chord.std.java.analysis.path}
+\begin{quote}
+{\bf Type:} path \\
+{\bf Description:} TODO \\
+{\bf Default value:} TODO
+\end{quote}
+
+\noindent \code{chord.ext.java.analysis.path}
+\begin{quote}
+{\bf Type:} path \\
+{\bf Description:} TODO \\
+{\bf Default value:} TODO
+\end{quote}
+
+\noindent \code{chord.java.analysis.path}
+\begin{quote}
 {\bf Type:} path \\
@@
 {\bf Default value:} \code{[chord.main.dir]/classes/}
+\end{quote}
 
-\item
-\code{chord.dlog.analysis.path} \\
+\noindent \code{chord.std.dlog.analysis.path}
+\begin{quote}
+{\bf Type:} path \\
+{\bf Description:} TODO \\
+{\bf Default value:} TODO
+\end{quote}
+
+\noindent \code{chord.ext.dlog.analysis.path}
+\begin{quote}
+{\bf Type:} path \\
+{\bf Description:} TODO \\
+{\bf Default value:} TODO
+\end{quote}
+
+\noindent \code{chord.dlog.analysis.path}
+\begin{quote}
 {\bf Type:} path  \\
@@
 {\bf Default value:} \code{[chord.main.dir]/src/}
+\end{quote}
 
-\item
-\code{chord.analysis.exclude} \\
-{\bf Type:} string list \\
-{\bf Descrption:} Names of program analyses written in Java or Datalog to be excluded from the project. \\
-{\bf Default value:} {\tt ""}
-
-\item
-\code{chord.reuse.rels} \\
-{\bf Type:} bool \\
-{\bf Description:} Construct program relations from BDDs stored on disk (from a previous run of Chord) whenever possible instead of re-computing them. \\
-{\bf Default value:} {\tt false} 
-% TODO: tell when this doesn't work
-
-\item
-\code{chord.print.results} \\
-{\bf Type:} bool \\
-{\bf Description:} Print the results of program analyses in HTML.  Interpretation of this property is analysis-specific. \\
-{\bf Default value:} {\tt true}
-
-\item
-\code{chord.std.check.exclude} \\
-{\bf Type:} string list \\
-{\bf Description:} List of prefixes of names of classes and packages inside the standard library to be excluded by program analyses.  Interpretation of this property is analysis-specific. \\
-{\bf Default value:} \code{"sun.,com.sun.,com.ibm.jvm.,com.ibm.oti.,com.ibm.misc.,org.apache.harmony.,joeq.,jwutil.,java.,javax."}
-
-\item
-\code{chord.ext.check.exclude} \\
-{\bf Type:} string list \\
-{\bf Description:} List of prefixes of names of classes and packages outside the standard library to be excluded by program analyses.  Interpretation of this property is analysis-specific. \\
-{\bf Default value:} {\tt ""}
-
-\item
-\code{chord.check.exclude} \\
-{\bf Type:} string list \\
-{\bf Description:} List of prefixes of names of classes and packages to be excluded by program analyses.  Interpretation of this property is analysis-specific. \\
-{\bf Default value:} \code{"[chord.std.check.exclude],[chord.ext.check.exclude]"}
-\end{itemize}
-
-\subsection{Program Transformation Properties}
-
-This section describes properties regarding program transformations performed by Chord.
-
-\begin{itemize}
-\item
-\code{chord.ssa} \\
-{\bf Type:} bool \\
-{\bf Description:} Do SSA transformation for all methods deemed reachable by the algorithm used to compute program scope. \\
-{\bf Default value:} {\tt true} 
-\end{itemize}
-
-\subsection{Chord Debug Properties}
-
-This section describes properties that specify the amount of debug information that Chord must produce during execution.
-
-\begin{itemize}
-\item
-\code{chord.verbose} \\
-{\bf Type:} int in the range [0..6] \\
-{\bf Description:} Control the verbosity of logging during Chord's execution. \\
-{\bf Default value:} {\tt 2}
-
-\item
-\code{chord.save.maps} \\
-{\bf Type:} bool \\
-{\bf Description:} Write to file \code{[chord.bddbddb.work.dir]/<...>.map} when saving program domain named {\tt <...>}. \\
-{\bf Default value:} {\tt true}  \\
-{\bf Note:} This functionality is useful for debugging Datalog programs using the {\tt debug} target provided in file {\tt build.xml} in Chord's {\tt main/} directory (see Section \ref{sec:tuning-datalog-analysis}).
-\end{itemize}
-
-\subsection{Chord Instrumentation Properties}
-\label{sec:instr-props}
+\paragraph{Instrumentation Properties}
 
@@
 % TODO: mention chord.scope.exclude* here
+\noindent \code{chord.use.jvmti}
+\begin{quote}
+{\bf Type:} bool \\
+{\bf Description:} TODO \\
+{\bf Default value:} TODO
+\end{quote}
 
-\begin{itemize}
-\item
-\code{chord.instr.kind} \\
+\noindent \code{chord.instr.kind}
+\begin{quote}
 {\bf Type:} {\tt [offline|online]}  \\
 {\bf Description:} The kind of bytecode instrumentation.  The choices are offline and load-time.  \\
-{\bf Default value:} {\tt offline} \\
+{\bf Default value:} {\tt offline}
+\end{quote}
 
-\item
-\code{chord.trace.kind} \\
+\noindent \code{chord.trace.kind}
+\begin{quote}
 {\bf Type:} {\tt [none|full|pipe]}  \\
@@
 {\bf Default value:} {\tt full}
+\end{quote}
 
-\item
-\code{chord.trace.block.size} \\
+\noindent \code{chord.trace.block.size}
+\begin{quote}
 {\bf Type:} int \\
@@
 {\bf Default value:} {\tt 4096}
+\end{quote}
 
-\item
-\code{chord.reuse.traces} \\
-{\bf Type:} bool \\
-{\bf Description:} Reuse dynamic traces computed by a previous run of Chord if they exist. \\
-{\bf Default value:} {\tt false} \\
-{\bf Note:} Property \code{chord.trace.kind} must be set to {\tt full} if this property is set to {\tt true}.
-
-\item
-\code{chord.dynamic.haltonerr} \\
+\noindent \code{chord.dynamic.haltonerr}
+\begin{quote}
 {\bf Type:} bool \\
@@
 {\bf Default value:} true
+\end{quote}
 
-\item
-\code{chord.dynamic.timeout} \\
+\noindent \code{chord.dynamic.timeout}
+\begin{quote}
 {\bf Type:} int  \\
 {\bf Description:} The amount of time, in milliseconds, after which to kill the process running the given program during dynamic analysis, or -1 if the process must never be killed. \\
-{\bf Default value:} -1
+{\bf Default value:} {\tt -1}
+\end{quote}
 
-\item
-\code{chord.max.cons.size} \\
+\noindent \code{chord.max.cons.size}
+\begin{quote}
 {\bf Type:} int \\
 {\bf Description:} Maximum number of bytes over which events generated during the execution of any constructor in the given program may span. \\
-{\bf Default value:} 50000000 \\
+{\bf Default value:} {\tt 50000000} \\
 {\bf Note:} This property is relevant only for dynamic analyses which want events of the form {\tt BEF\_NEW} $h$ $t$ $o$ to be generated (see Section \ref{sec:instr-events}).
@@
 If the actual number of bytes exceeds the value specified by this property (e.g., if the constructor throws an exception and the {\tt AFT\_NEW} event is not generated at all), then Chord simply generates event {\tt BEF\_NEW} $h$ $i$ $0$ (i.e., it treats the created object as having ID 0, which is the ID also used for {\tt null}).
-\end{itemize}
+\end{quote}
 
-\subsection{Chord Output Properties}
+\paragraph{Caching Properties}
+
+TODO
+
+\noindent \code{chord.reuse.scope}
+\begin{quote}
+{\bf Type:} bool \\
+{\bf Description:} Compute program scope using the information in files specified by properties \code{chord.methods.file} and \code{chord.reflect.file}. \\
+{\bf Default value:} {\tt false} \\
+{\bf Note:} Property \code{chord.scope.kind} is ignored if this property is set to {\tt true} and the two files exist. 
+\end{quote}
+
+\noindent \code{chord.reuse.rels}
+\begin{quote}
+{\bf Type:} bool  \\
+{\bf Description:} Construct program relations from BDDs stored on disk (from a previous run of Chord) whenever possible instead of re-computing them. \\
+{\bf Default value:} {\tt false}
+\end{quote}
+
+\noindent \code{chord.reuse.traces}
+\begin{quote}
+{\bf Type:} bool \\
+{\bf Description:} Reuse dynamic traces computed by a previous run of Chord if they exist.  Property \code{chord.trace.kind} must be set to {\tt full} if this property is set to {\tt true}. \\
+{\bf Default value:} {\tt false}
+\end{quote}
+
+\paragraph{Chord JVM Properties}
+
+TODO
+
+\noindent \code{java.class.path}
+\begin{quote}
+{\bf Type:} path \\
+{\bf Description:} TODO \\
+{\bf Default value:} TODO
+\end{quote}
+
+\noindent \code{user.dir}
+\begin{quote}
+{\bf Type:} location
+{\bf Description:} TODO \\
+{\bf Default value:} TODO
+\end{quote}
+
+\noindent \code{chord.max.heap}
+\begin{quote}
+{\bf Type:} string \\
+{\bf Description:} Maximum memory size of JVM running Chord. \\
+{\bf Default value:} {\tt 1024m}
+\end{quote}
+
+\noindent \code{chord.max.stack}
+\begin{quote}
+{\bf Type:} string \\
+{\bf Description:} Maximum thread stack size of JVM running Chord. \\
+{\bf Default value:} {\tt 32m}
+\end{quote}
+
+\noindent \code{chord.jvmargs}
+\begin{quote}
+{\bf Type:} string \\
+{\bf Description:} Arguments to JVM running Chord. \\
+{\bf Default value:} \code{"-showversion} \code{-ea} \code{-Xmx[chord.max.heap]} \code{-Xss[chord.max.stack]"}
+\end{quote}
+
+\paragraph{BDD Properties}
+
+TODO
+
+\noindent \code{chord.use.buddy}
+\begin{quote}
+{\bf Type:} bool \\
+{\bf Description:} TODO \\
+{\bf Default value:} TODO
+\end{quote}
+
+\noindent \code{chord.bddbddb.max.heap}
+\begin{quote}
+{\bf Type:} string \\
+{\bf Description:} Maximum memory size of JVM running Datalog solver bddbddb. \\
+{\bf Default value:} {\tt 1024m} \\
+{\bf Note:}  Each program analysis written in Datalog is run in a separate JVM.
+This is primarily because there may be multiple invocations of bddbddb in a single run of Chord and it is difficult to reset the global state of bddbddb on each invocation.
+\end{quote}
+
+\paragraph{Output Location Properties}
 
@@
 
-\begin{itemize}
-\item
-\code{chord.out.dir} \\
+\noindent \code{chord.out.file}
+\begin{quote}
+{\bf Type:} location \\
+{\bf Description:} Absolute location of the file to which the standard output stream is redirected during Chord's execution. \\
+{\bf Default value:} \code{null}
+\end{quote}
+
+\noindent \code{chord.err.file}
+\begin{quote}
+{\bf Type:} location  \\
+{\bf Description:} Absolute location of the file to which the standard error stream is redirected during Chord's execution. \\
+{\bf Default value:} \code{null}
+\end{quote}
+
+\noindent \code{chord.out.dir}
+\begin{quote}
 {\bf Type:} location \\
@@
 {\bf Default value:} \code{[chord.work.dir]/chord_output/}
+\end{quote}
 
-\item
-\code{chord.out.file} \\
-{\bf Type:} location \\
-{\bf Description:} Absolute location of the file to which the standard output stream is redirected during Chord's execution. \\
-{\bf Default value:} \code{[chord.out.dir]/log.txt}
-
-\item
-\code{chord.err.file} \\
-{\bf Type:} location \\
-{\bf Description:} Absolute location of the file to which the standard error stream is redirected during Chord's execution. \\
-{\bf Default value:} \code{[chord.out.dir]/log.txt}
-
-\item
-\code{chord.reflect.file} \\
+\noindent \code{chord.reflect.file}
+\begin{quote}
 {\bf Type:} location  \\
@@
 {\bf Default value:} \code{[chord.out.dir]/reflect.txt}
+\end{quote}
 
-\item
-\code{chord.methods.file} \\
+\noindent \code{chord.methods.file}
+\begin{quote}
 {\bf Type:} location  \\
@@
 {\bf Default value:} \code{[chord.out.dir]/methods.txt}
+\end{quote}
 
-\item
-\code{chord.classes.file} \\
+\noindent \code{chord.classes.file}
+\begin{quote}
 {\bf Type:} location  \\
@@
 {\bf Default value:} \code{[chord.out.dir]/classes.txt}
+\end{quote}
 
-\item
-\code{chord.bddbddb.work.dir} \\
+\noindent \code{chord.bddbddb.work.dir}
+\begin{quote}
 {\bf Type:} location \\
@@
 {\bf Default value:} \code{[chord.out.dir]/bddbddb/}
+\end{quote}
 
-\item
-\code{chord.boot.classes.dir} \\
+\noindent \code{chord.boot.classes.dir}
+\begin{quote}
 {\bf Type:} location \\
@@
 {\bf Default value:} \code{[chord.out.dir]/boot_classes/}
+\end{quote}
 
-\item
-\code{chord.user.classes.dir} \\
+\noindent \code{chord.user.classes.dir}
+\begin{quote}
 {\bf Type:} location  \\
@@
 {\bf Default value:} \code{[chord.out.dir]/user_classes/}
+\end{quote}
 
-\item
-\code{chord.instr.scheme.file} \\
+\noindent \code{chord.instr.scheme.file}
+\begin{quote}
 {\bf Type:} location \\
@@
 {\bf Default value:} \code{[chord.out.dir]/scheme.ser}
+\end{quote}
 
-\item
-\code{chord.trace.file} \\
+\noindent \code{chord.trace.file}
+\begin{quote}
 {\bf Type:} location  \\
@@
 {\bf Default value:} \code{[chord.out.dir]/trace} \\
-{\bf Note:} \code{_full_verN.txt} or \code{_pipe_verN.txt} is appended to the name of the file, depending upon whether it is a regular file or a POSIX pipe, respectively, where {\tt N} is the version of the file (multiple versions are maintained if the trace is transformed by filters defined by the dynamic analysis; 0 is the final version).
+{\bf Note:} Suffix \code{_full_verN.txt} or \code{_pipe_verN.txt} is appended to the name of the file, depending upon whether it is a regular file or a POSIX pipe, respectively, where {\tt N} is the version of the file (multiple versions are maintained if the trace is transformed by filters defined by the dynamic analysis; 0 is the final version).
 If \code{chord.reuse.traces} is set to {\tt true}, then \code{_full_verN_runM.txt} is appended to the name of the file, where {\tt M} is the run ID.
-\end{itemize}
+\end{quote}
 
-\subsection{Chord Resource Properties}
-\label{sec:chord-sysprops:resource}
+%\item
+%\code{chord.main.dir}
+% location
+%Absolute location of the {\tt main/} directory in Chord's installation.
 
-This section describes properties specifying classpaths and locations of runtime libraries used by Chord.
-Ordinarily, users must not alter the default values of these properties.
-
-\begin{itemize}
-\item
-\code{chord.main.dir} \\
-{\bf Type:} location \\
-{\bf Description:} Absolute location of the {\tt main/} directory in Chord's installation.
-
-\item
-\code{chord.lib.dir} \\
-{\bf Type:} location \\
-{\bf Description:} Directory containing libraries needed by Chord. \\
-{\bf Default value:} \code{[chord.main.dir]/lib/}
-
-\item
-\code{chord.main.class.path} \\
-{\bf Type:} path \\
-{\bf Description:} Classpath of Chord.  It includes the path specified by property \code{chord.java.analysis.path} to allow running user-defined program analyses. \\
-{\bf Default value:} See \code{[chord.main.dir]/build.xml}
-
-\item
-\code{chord.bddbddb.class.path} \\
-{\bf Type:} path \\
-{\bf Description:} Classpath of {\tt bddbddb}. \\
-{\bf Default value:} See \code{[chord.main.dir]/build.xml}
-
-\item
-\code{chord.javadoc.url} \\
-{\bf Type:} string \\
-{\bf Description:} URL of the Javadoc location of program analyses.  It is used when printing the project (i.e., when property \code{chord.print.project} is set to {\tt true}). \\
-{\bf Default value:} \code{http://jchord.googlecode.com/svn/wiki/javadoc/}
-\end{itemize}
+%\item
+%\code{chord.save.maps}
+% bool
+%Write to file \code{[chord.bddbddb.work.dir]/<...>.map} when saving program domain named {\tt <...>}.
+%{\tt true}
+%{\bf Note:} This functionality is useful for debugging Datalog programs using the {\tt debug} target provided in file {\tt build.xml} in Chord's {\tt main/} directory (see Section \ref{sec:tuning-datalog-analysis}).

### 2011-03-10 main/doc/deadlock.tex
added more documentation

@@
-\section{Deadlock analysis}
+\section{Deadlock Analysis}
 \label{sec:deadlock}
 
-Run the deadlock analysis provided in Chord by running the following
+Run Chord's static deadlock analysis by running the following
 command in Chord's {\tt main/} directory:
@@
 
-\noindent where {\tt <...>} is the absolute path of a directory
+\noindent where {\tt <...>} denotes the absolute or relative path of a directory
 containing a file named \code{chord.properties} which defines
@@
 in Chord's {\tt main/} directory with {\tt <...>} replaced by
-\code{/.../examples/deadlock_test/}.  Upon successful completion, the
+\code{examples/deadlock_test/}.  Upon successful completion, the
 file \code{deadlocks.html} should be produced in directory
-\code{main/examples/deadlock_test/chord_output/}.
+\code{examples/deadlock_test/chord_output/}.

### 2011-03-10 main/doc/download_and_installation.tex
added more documentation

--- 2010-08-14 main/doc/download_and_installation.tex
@@
 
-Ensure that the following software is installed on your machine:
+You can obtain pre-built binaries as well as the source code of Chord.
+The following sections explain both options in detail.
 
+\subsection {Binary Installation}
+\label{sec:binary-install}
+
+To obtain Chord's pre-built binaries, download file \xlink{chord-bin-2.0.tar.gz}{http://jchord.googlecode.com/files/chord-bin-2.0.tar.gz}.
+It contains the following files:
+
+\begin{enumerate}
+\item
+\code{chord.jar}, which contains the class files of Chord and of libraries used by Chord.
+\item
+\code{libbuddy.so}, \code{buddy.dll}, and \code{libbuddy.dylib}: you can keep one of these files
+depending upon whether you intend to run Chord on Linux, Windows/Cygwin, or MacOS, respectively.
+These files are needed only if you want to use the BuDDy BDD library for executing program
+analyses written in Datalog.
+\item
+\code{libchord_instr_agent.so}: this file is needed only if you want to use the JVMTI-based bytecode
+instrumentation agent for executing dynamic program analyses in Chord.
+\end{enumerate}
+
+\noindent Novice users can ignore items (2) and (3) above until they become more familiar with Chord.
+The provided binaries in those items might not be compatible with your machine, in which case you
+will either have to forgo using them (with hardly any noticeable difference in Chord's functionality)
+or you will have to build them yourself (see Section \ref{sec:source-install}).
+
+\subsection{Source Installation}
+\label{sec:source-install}
+
+To obtain Chord's source code, download the following files:
+
+\begin{enumerate}
+\item
+Mandatory: file \xlink{chord-src-2.0.tar.gz}{http://jchord.googlecode.com/files/chord-src-2.0.tar.gz}, which contains Chord's source code and the jars of libraries used by Chord.
+\item
+Optional: file \xlink{chord-libsrc-2.0.tar.gz}{http://jchord.googlecode.com/files/chord-libsrc-2.0.tar.gz}, which contains the source
+code of libraries used by Chord (e.g., joeq, javassist, bddbddb, etc.)
+\end{enumerate}
+
+\noindent Alternatively, you can obtain the latest development snapshot from the SVN repository:
+
+\begin{verbatim}
+    prompt> svn checkout http://jchord.googlecode.com/svn/trunk/ chord
+\end{verbatim}
+
+\noindent Instead of checking out the entire \code{trunk/}, which contains several sub-directories, you can check out
+specific sub-directories:
 \begin{itemize}
-\item
-JDK 6 or higher, e.g. from \htmladdnormallink{IBM}{http://www.ibm.com/developerworks/java/jdk/} or
-\htmladdnormallink{Sun}{http://java.sun.com/javase/downloads/index.jsp}
-\item
-\htmladdnormallink{Apache Ant}{http://ant.apache.org/}, a Java-based build tool
-\item
-a C++ compiler, e.g. \htmladdnormallink{GCC}{http://gcc.gnu.org/}
-\item
-a Make utility, e.g. \htmladdnormallink{GNU Make}{http://www.gnu.org/software/make/}
-\item
-\htmladdnormallink{Cygwin}{http://www.cygwin.com/}, if it is a Windows machine
+\item \code{main/} contains Chord's source code and the jars of libraries used by Chord.
+\item \code{libsrc/} contains the source code of libraries used by Chord (e.g., joeq, javassist, bddbddb, etc.).
+\item \code{test/} contains Chord's regression tests.
+\item many more; these might eventually move into \code{main/}.
 \end{itemize}
 
-Download a \Chord\ source release from \url{http://code.google.com/p/jchord/downloads/list} or
-the latest source from the SVN repository at \url{http://code.google.com/p/jchord/source/checkout}.
-
-Directory {\tt main/} contains a {\tt build.xml} file which is interpreted by Apache Ant.
-To see the various targets available, run the following command in that directory:
-
-\begin{quote}
-\begin{verbatim}
-prompt> ant
-\end{verbatim}
-\end{quote}
-
-\noindent To compile \Chord, run the following command in the same directory:
-
-\begin{quote}
-\begin{verbatim}
-prompt> ant compile
-\end{verbatim}
-\end{quote}
-
-\noindent This will compile the following:
-\begin{itemize}
-\item
-the Java source code of \Chord\ from \code{main/src/java/} (and from \code{extra/src/java/} if that directory is present) to Java bytecode in \code{main/classes/}.
-\item
-the C source code of BDD library \htmladdnormallink{BuDDy}{http://buddy.sourceforge.net/} from \code{main/src/bdd/} to a shared library in \code{main/lib/} (\code{libbuddy.so} on Linux, \code{buddy.dll} on Windows, and \code{libbuddy.dylib} on MacOS); this library is needed for executing program analyses written in Datalog using \htmladdnormallink{bddbddb}{http://bddbddb.sourceforge.net/} (a BDD-based Datalog solver).
-\item
-the C++ source code of the Chord instrumentation agent from \code{main/src/agent/} to a shared library in \code{main/lib/} (\code{libchord_instr_agent.so} on all architectures); this agent is needed for computing program scope dynamically and for executing dynamic program analyses.
-\end{itemize}
+\noindent Files \code{chord-2.0-src.tar.gz} and \code{chord-2.0-libsrc.tar.gz} mentioned above are essentially stable releases of
+the \code{main/} and \code{libsrc/} directories, respectively.

### 2011-03-10 main/doc/setting_up.tex
added more documentation

--- 2010-12-13 main/doc/program_setup.tex
@@
 \section{Setting up a Java Program}
-\label{sec:program-setup}
+\label{sec:setting-up}
 
@@
 
-File \code{chord.properties} specifies properties to be passed to
-Chord (alternatively, these properties may be passed on the
-command-line, in format \code{-D<name>=<value>}).  A sample such file
-for the above program is as follows:
+To run Chord on the above program, run the following command in
+Chord's {\tt main/} directory:
+
+\begin{verbatim}
+    prompt> ant -Dchord.work.dir=<...>/example run
+\end{verbatim}
+
+\noindent where \code{<...>} denotes the absolute or relative path of the
+directory containing the above \code{example/} directory.
+Property \code{chord.work.dir} specifies the working directory during Chord's execution.
+Chord searches for a file named \code{chord.properties} under that working directory, and
+loads upfront all properties defined in that file, if it exists.  Alternatively, these properties
+may be defined on the command-line, using the \code{-D<key>=<value>} format.
+A sample \code{chord.properties} file for the above program is as follows:
 
@@
     chord.class.path=classes:lib/jar/taz.jar
-    chord.src.path=src:../lib/src
+    chord.src.path=src:lib/src
     chord.run.ids=0,1
@@
 
-Each relative path element in the value of any property named
-\code{chord.<...>.path} defined in this file is converted to an
-absolute path element with respect to the directory containing this
-file, which in the above case is \code{example/}.
+Each relative file/directory name in the value of any property
+defined in this file (e.g., the \code{lib/src} directory name in the value of
+property \code{chord.src.path} above) is treated relative to Chord's working directory
+(which is specified by property \code{chord.work.dir}).
 
 Section \ref{sec:chord-sysprops} presents an exhaustive listing of
-properties recognized by \Chord.  Here, we only describe those defined
+properties recognized by Chord.  Here, we only describe those defined
 in the above sample \code{chord.properties} file:
@@
 \item
-\code{chord.class.path} specifies the application-specific classpath
+\code{chord.class.path} specifies the application classpath
 of the program (the standard library classpath is implicitly
@@
 
-To run Chord on the above program, run the following command in
-Chord's {\tt main/} directory:
-
-\begin{verbatim}
-    prompt> ant -Dchord.work.dir=<...>/example run
-\end{verbatim}
-
-\noindent where {\tt <...>} denotes the absolute path of the parent of
-directory {\tt example/}.  Property \code{chord.work.dir} specifies
-the working directory during Chord's execution.
-
-The above command causes Chord to load the \code{chord.properties}
-file if present.  The location of this file may be specified
-explicitly via property \code{chord.props.file} on the above command
-line.  By default, it is \code{[chord.work.dir]/chord.properties}.
-
-In the above case, Chord does not do much beyond loading the sample
+In the above case, Chord does not do much beyond loading the above
 \code{chord.properties} file.  For Chord to do something interesting,
@@
 Chord must perform.  All tasks are summarized in Section
-\ref{sec:chord-sysprops:task}.  The two most common tasks, described
-next, are computing the scope of the given program (Section
+\ref{sec:func-props}.  The two most common tasks, described
+next, are computing the analysis scope of the given program (Section
 \ref{sec:building-scope}) and running analyses on the given program

### 2011-03-10 main/doc/whatis_chord.tex
added more documentation

@@
 
-Chord is a program analysis platform that enables 
+Chord is a program analysis platform that enables users to
 productively design, implement, evaluate, and combine a
@@
 races, deadlocks, and atomicity violations; etc.)
-
 \item
@@
 combined static/dynamic analyses.
-
 \item
@@
 them in parallel.
-
 \item

### 2011-03-10 main/doc/writing_analysis.tex
added more documentation

@@
-\section{Writing a Program Analysis}
+\section{Writing an Analysis}
 \label{sec:writing-analysis}
 
-TODO
+Chord is built around the notion of modeling a program with relations.
+A Chord analysis is a program component that operates on these relations.
+It may consume zero or more relations, do some work on them, and then output zero or more relations.
+It may also print output in some other format, perhaps human-readable text or HTML.
+``Relation'' here has the same meaning it does in the context of relational databases: a set of tuples, containing elements of one or more domains.  
+%For example, the relation \texttt{reachableM(m:M)} identifies a subset of
+
+An analysis can be written in one of two ways: either as a Java class or as a datalog program.
+
+\subsection{Dependency information}
+
+Chord has two kinds of projects: ``Classic'' and ``Modern''.  Despite the names, classic is effectively the standard at this point.
+In the classic model, each analysis specifies the domains/relations it consumes and that it produces.
+This defines an implicit dependency graph between analyses.
+For a datalog analysis, the ``input'' and ``output'' keywords on relations specify the requirements and results of an analysis.
+For a Java analysis, the \texttt{@Chord} annotation does so.
+
+\subsection{Domains}
+
+In Chord, a domain is represented as a list of possible values. 
+The value types must be comparable and hashable.
+Strings work, as do integers, \texttt{joeq} quads, fields, types, and so forth. 
+%
+% What are the required types?
+
+Both domains and relations have names. The built-in relations are usually identified by a single capital letter.
+Chord by default includes code for constructing a number of domains.
+
+It is sometimes useful to construct new domains in the middle of an analysis, perhaps as a way of compactly describing some entities that required nontrivial work to find.
+Domains, for now, must be defined in Java.
+One key requirement is that construction must be deterministic: if the elements in a domain can be reordered from execution to execution on the same program, this can result in subtle and hard-to-track-down bugs.
+
+\begin{itemize}
+\item
+B is the domain of basic blocks.
+\item
+C is the domain of abstract calling contexts.
+\item
+E is the domain of program points that read or write a field.
+\item
+F is the domain of fields inside a class.
+\item
+H is the domain of allocation sites.
+\item
+I is the domain of invocation statements (method calls)
+\item
+L is the domain of lock acquisition points.
+\item
+M is the domain of program methods.
+\item
+P is the domain of simple statements. %SAY MORE HERE?
+\item
+R is the domain of lock release points.
+\item
+T is the domain of types (classes and primitives).
+\item
+V is the domain of reference-typed variables.
+\item
+W is the domain of loop head statements. %SAY MORE HERE?
+\item
+Z is a domain of integers corresponding to argument positions.The size of the domain is chosen to 
+accommodate the method with the greatest number of arguments.
+\item
+ZZ is another domain of integers, this one of fixed size and available for other uses.
+\end{itemize}
+
+\subsection{Core relations}
+
+Chord includes code to build a very large number of relations. These standard relations
+are very useful in writing your own analyses.
+For example, relation \texttt{sub} includes all pairs of types \texttt{(t1, t2)} where \texttt{t2} is a subclass of \texttt{t1}.
+Most of these standard relations are in the package \texttt{chord.rels}.
+
+\subsection{Java analyses}
+
+A Java analysis should extend JavaAnalysis or ProgramRel.  It's usually easiest if an analysis
+only produces one relation, in which case extending ProgramRel is the best approach.
+If you extend ProgramRel, all the work of creating and saving the relation is taken care of for you;
+all you need to do is add tuples to the relation.Typically you do this by calling
+\texttt{super.add(...)} with the tuple you wish to add to the relation.
+For forther convenience, you can implement one of the Visitor interfaces defined in 
+\texttt{chord.program.visitors}.  This will force you to implement some \texttt{visit} methods that iterate over
+all elements in some domain, such as all methods or all invocation points.
+
+Sometimes, you need to create two different relations at the same time. In this case, the right approach is to
+extend JavaAnalysis, and explicitly create and save the relations on your own.
+\textbf{Should reference JavaDoc.  Should make sure JavaDoc explains what you need to do to fill and save a relation}.
+
+Regardless of which of these approaches you take, every Java analysis must have an \texttt{@Chord} annotation.
+This annotation is a set of named fields.
+At a minimum, this annotation must specify the name of the analysis.
+If you extend ProgramRel, the name of the analysis will be treated as the same as that of the generated relation.
+For example, \texttt{@Chord(name = "F")} is a valid annotation. 
+The name of the Java class that generates the relation is ignored by Chord -- only the annotation matters.
+
+\textbf{MAYUR, IS SIGN MANDATORY?   IF NOT, WHEN CAN IT BE OMITTED?}
+
+There are also other annotation fields available.
+You may specify the dependencies of your analysis via \texttt{consumes = \{ "xyz"\}}.
+
+If your analysis does not extend ProgramRel, you should also specify the relations that your 
+analysis produces. You do this via the \texttt{produces} relation.
+
+%Should mention the signs fields.
+%
+%
+
+\noindent  The default value of property \code{chord.java.analysis.path}
+is \code{chord.std.java.analysis.path}:\code{chord.ext.java.analysis.path},
+where the defalt value of \code{chord.std.java.analysis.path} is the location of file \code{chord.jar}, and
+the default value of \code{chord.ext.java.analysis.path} is empty (users defining their own Java analyses are expected to
+set the latter property).
+
+\noindent  The default value of property \code{chord.dlog.analysis.path}
+is \code{chord.std.dlog.analysis.path}:\code{chord.ext.java.analysis.path},
+where the defalt value of \code{chord.std.java.analysis.path} is the location of file \code{chord.jar}, and
+the default value of \code{chord.ext.java.analysis.path} is empty (users defining their own Java analyses are expected to
+set the latter property).

### 2011-03-10 main/doc/writing_datalog_analysis.tex
added more documentation

@@
-\section{Writing a Datalog analysis}
+\section{Writing a Datalog Analysis}
 \label{sec:writing-datalog-analysis}
@@
 Any line that begins with a {\tt \#} is regarded a comment, except a
-line of the form {\tt \#} {\tt name=<...>}, which specifies the name
+line of the form ``{\tt \#} {\tt name=<...>}", which specifies the name
 {\tt <...>} of the Datalog analysis.
@@
 {\tt -dlog} and all Java analysis names have suffix {\tt -java}, but
-this is merely a convention that user-defined analyses are free to deviate from.
+this is merely a convention that analysis writers are free to deviate from.
 The name of each analysis, written in Datalog or in Java, is expected to be

### 2011-03-11 main/doc/chord_sysprops.tex
refactoring: moved chord.analyses.XYZ (where XZY is field, type, method, etc.) and all other factual domains/relations to chord.analyses.facts.XYZ to reduce cluttering in package chord.analyses

@@
 {\bf Type:} {\tt [offline|online]}  \\
-{\bf Description:} The kind of bytecode instrumentation.  The choices are offline and load-time.  \\
+{\bf Description:} The kind of bytecode instrumentation.  The choices are offline and online (load-time).  \\
 {\bf Default value:} {\tt offline}
@@
 \begin{quote}
-{\bf Type:} {\tt [none|full|pipe]}  \\
-{\bf Description:} The medium to store dynamic traces (none, regular file, or POSIX pipe).  \\
+{\bf Type:} {\tt [full|pipe]}  \\
+{\bf Description:} The medium by which an event-generating JVM and an event-handling JVM communicate in a dynamic analysis.  The choices are regular file and POSIX pipe.  \\
 {\bf Default value:} {\tt full} 
@@
 {\bf Type:} int \\
-{\bf Description:} Number of bytes to read/write in a single operation from/to the dynamic trace file. \\
+{\bf Description:} Number of bytes to read/write in a single operation from/to the event file used by a multi-JVM dynamic analysis. \\
 {\bf Default value:} {\tt 4096}

### 2011-03-11 main/src/chord/analyses/argret/RelIinvkArg.java

@@ -1,50 +1,0 @@
-/*
- * Copyright (c) 2008-2010, Intel Corporation.
- * Copyright (c) 2006-2007, The Trustees of Stanford University.
- * All rights reserved.
- * Licensed under the terms of the New BSD License.
- */
-package chord.analyses.argret;
-
-import joeq.Compiler.Quad.Quad;
-import joeq.Compiler.Quad.Operand.ParamListOperand;
-import joeq.Compiler.Quad.Operand.RegisterOperand;
-import joeq.Compiler.Quad.Operator.Invoke;
-import joeq.Compiler.Quad.RegisterFactory.Register;
-
-import chord.analyses.invk.DomI;
-import chord.analyses.var.DomV;
-import chord.project.Chord;
-import chord.project.analyses.ProgramRel;
-
-/**
- * Relation containing each tuple (i,z,v) such that local variable v
- * is the zth argument variable of method invocation statement i.
- *
- * @author Mayur Naik (mhn@cs.stanford.edu)
- */
-@Chord(
-	name = "IinvkArg",
-	sign = "I0,Z0,V1:I0_V1_Z0"
-)
-public class RelIinvkArg extends ProgramRel {
-	public void fill() {
-		DomI domI = (DomI) doms[0];
-		DomV domV = (DomV) doms[2];
-		int numI = domI.size();
-		for (int iIdx = 0; iIdx < numI; iIdx++) {
-			Quad q = (Quad) domI.get(iIdx);
-			ParamListOperand l = Invoke.getParamList(q);
-			int numArgs = l.length();
-			for (int zIdx = 0; zIdx < numArgs; zIdx++) {
-				RegisterOperand vo = l.get(zIdx);
-				Register v = vo.getRegister();
-				if (v.getType().isReferenceType()) {
-					int vIdx = domV.indexOf(v);
-					assert (vIdx >= 0);
-					add(iIdx, zIdx, vIdx);
-				}
-			}
-		}
-	}
-}

### 2011-03-11 main/src/chord/analyses/argret/RelIinvkRet.java

@@ -1,47 +1,0 @@
-/*
- * Copyright (c) 2008-2010, Intel Corporation.
- * Copyright (c) 2006-2007, The Trustees of Stanford University.
- * All rights reserved.
- * Licensed under the terms of the New BSD License.
- */
-package chord.analyses.argret;
-
-import joeq.Compiler.Quad.Quad;
-import joeq.Compiler.Quad.Operand.RegisterOperand;
-import joeq.Compiler.Quad.Operator.Invoke;
-import joeq.Compiler.Quad.RegisterFactory.Register;
-
-import chord.analyses.invk.DomI;
-import chord.analyses.var.DomV;
-import chord.project.Chord;
-import chord.project.analyses.ProgramRel;
-
-/**
- * Relation containing each tuple (i,z,v) such that local variable v
- * is the zth return variable of method invocation statement i.
- *
- * @author Mayur Naik (mhn@cs.stanford.edu)
- */
-@Chord(
-	name = "IinvkRet",
-	sign = "I0,Z0,V0:I0_V0_Z0"
-)
-public class RelIinvkRet extends ProgramRel {
-	public void fill() {
-		DomI domI = (DomI) doms[0];
-		DomV domV = (DomV) doms[2];
-		int numI = domI.size();
-		for (int iIdx = 0; iIdx < numI; iIdx++) {
-			Quad q = (Quad) domI.get(iIdx);
-			RegisterOperand vo = Invoke.getDest(q);
-			if (vo != null) {
-				Register v = vo.getRegister();
-				if (v.getType().isReferenceType()) {
-					int vIdx = domV.indexOf(v);
-					assert (vIdx >= 0);
-					add(iIdx, 0, vIdx);
-				}
-			}
-		}
-	}
-}

### 2011-03-11 main/src/chord/analyses/argret/RelMmethRet.java

@@ -1,50 +1,0 @@
-/*
- * Copyright (c) 2008-2010, Intel Corporation.
- * Copyright (c) 2006-2007, The Trustees of Stanford University.
- * All rights reserved.
- * Licensed under the terms of the New BSD License.
- */
-package chord.analyses.argret;
-
-import joeq.Class.jq_Class;
-import joeq.Class.jq_Method;
-import joeq.Compiler.Quad.Operand;
-import joeq.Compiler.Quad.Quad;
-import joeq.Compiler.Quad.Operand.RegisterOperand;
-import joeq.Compiler.Quad.Operator.Return;
-import joeq.Compiler.Quad.RegisterFactory.Register;
-
-import chord.program.visitors.IReturnInstVisitor;
-import chord.project.Chord;
-import chord.project.analyses.ProgramRel;
-
-/**
- * Relation containing each tuple (m,z,v) such that local variable
- * v is the zth return variable of method m.
- *
- * @author Mayur Naik (mhn@cs.stanford.edu)
- */
-@Chord(
-	name = "MmethRet",
-	sign = "M0,Z0,V1:M0_V1_Z0"
-)
-public class RelMmethRet extends ProgramRel
-		implements IReturnInstVisitor {
-	private static Integer ZERO = new Integer(0);
-	private jq_Method ctnrMethod;
-	public void visit(jq_Class c) { }
-	public void visit(jq_Method m) {
-		ctnrMethod = m;
-	}
-	public void visitReturnInst(Quad q) {
-		Operand rx = Return.getSrc(q);
-		// note: rx is null if this method returns void
-		if (rx instanceof RegisterOperand) {
-			RegisterOperand ro = (RegisterOperand) rx;
-			if (ro.getType().isReferenceType()) {
-				Register v = ro.getRegister();
-				add(ctnrMethod, ZERO, v);
-			}
-		}
-	}
-}

### 2011-03-11 main/src/chord/analyses/basicblock/RelSuccBB.java

@@ -1,41 +1,0 @@
-/*
- * Copyright (c) 2008-2010, Intel Corporation.
- * Copyright (c) 2006-2007, The Trustees of Stanford University.
- * All rights reserved.
- * Licensed under the terms of the New BSD License.
- */
-package chord.analyses.basicblock;
-
-import java.util.List;
-
-import joeq.Compiler.Quad.BasicBlock;
-import chord.program.Program;
-import chord.project.Chord;
-import chord.project.analyses.ProgramRel;
-import chord.analyses.basicblock.DomB;
-
-/**
- * Relation containing each pair of basic blocks (b1,b2) in each method
- * such that b1 is immediate postdominator of b2.
- *
- * @author Mayur Naik (mhn@cs.stanford.edu)
- */
-@Chord(
-	name = "succBB",
-	sign = "B0,B1:B0xB1"
-)
-public class RelSuccBB extends ProgramRel {
-	public void fill() {
-		DomB domB = (DomB) doms[0];
-		int numB = domB.size();
-		for (int bIdx = 0; bIdx < numB; bIdx++) {
-			BasicBlock bb = domB.get(bIdx);
-			List<BasicBlock> succs = bb.getSuccessorsList();
-			for (BasicBlock bb2 : succs) {
-				int bIdx2 = domB.indexOf(bb2);
-				assert (bIdx2 >= 0);
-				add(bIdx, bIdx2);
-			}
-		}
-	}
-}

### 2011-03-11 main/src/chord/analyses/deadlock/DeadlockAnalysis.java

@@ -35,10 +35,10 @@ import chord.analyses.alias.DomO;
 import chord.analyses.alias.DomC;
 import chord.bddbddb.Rel.RelView;
 import chord.analyses.thread.DomA;
-import chord.analyses.lock.DomL;
-import chord.analyses.alloc.DomH;
-import chord.analyses.invk.DomI;
-import chord.analyses.method.DomM;
+import chord.analyses.facts.alloc.DomH;
+import chord.analyses.facts.invk.DomI;
+import chord.analyses.facts.lock.DomL;
+import chord.analyses.facts.method.DomM;
 import chord.util.SetUtils;
 import chord.util.tuple.object.Pair;
 import chord.util.tuple.object.Trio;

### 2011-03-11 main/src/chord/analyses/facts/argret/DomK.java

@@ -1,0 +1,29 @@
+/*
+ * Copyright (c) 2008-2010, Intel Corporation.
+ * Copyright (c) 2006-2007, The Trustees of Stanford University.
+ * All rights reserved.
+ * Licensed under the terms of the New BSD License.
+ */
+package chord.analyses.facts.argret;
+
+import chord.project.Chord;
+import chord.project.analyses.ProgramDom;
+/**
+ * Another domain of integers. Unlike domZ, has fixed size, unconnected
+ * to max_args.
+ * 
+ * Size is set via option chord.domK.size. Default is 32.
+ *
+ */
+@Chord(
+	name = "K"
+  )
+public class DomK extends ProgramDom<Integer> {
+  public static final int MAXZ = Integer.getInteger("chord.domK.size", 32);
+  
+  @Override
+  public void fill() {
+	for (int i = 0; i < MAXZ; i++)
+	  getOrAdd(new Integer(i));  
+  }
+}

### 2011-03-11 main/src/chord/analyses/facts/argret/RelIinvkArg0.java

@@ -1,0 +1,50 @@
+/*
+ * Copyright (c) 2008-2010, Intel Corporation.
+ * Copyright (c) 2006-2007, The Trustees of Stanford University.
+ * All rights reserved.
+ * Licensed under the terms of the New BSD License.
+ */
+package chord.analyses.facts.argret;
+
+import joeq.Compiler.Quad.Quad;
+import joeq.Compiler.Quad.Operand.ParamListOperand;
+import joeq.Compiler.Quad.Operand.RegisterOperand;
+import joeq.Compiler.Quad.Operator.Invoke;
+import joeq.Compiler.Quad.RegisterFactory.Register;
+
+import chord.analyses.facts.invk.DomI;
+import chord.analyses.facts.var.DomV;
+import chord.project.Chord;
+import chord.project.analyses.ProgramRel;
+
+/**
+ * Relation containing each tuple (i,v) such that local variable v
+ * is the 0th argument variable of method invocation statement i,
+ * if i has >= 0 arguments.
+ *
+ * @author Mayur Naik (mhn@cs.stanford.edu)
+ */
+@Chord(
+	name = "IinvkArg0",
+	sign = "I0,V1:I0_V1"
+)
+public class RelIinvkArg0 extends ProgramRel {
+	public void fill() {
+		DomI domI = (DomI) doms[0];
+		DomV domV = (DomV) doms[1];
+		int numI = domI.size();
+		for (int iIdx = 0; iIdx < numI; iIdx++) {
+			Quad q = (Quad) domI.get(iIdx);
+			ParamListOperand l = Invoke.getParamList(q);
+			if (l.length() > 0) {
+				RegisterOperand vo = l.get(0);
+				Register v = vo.getRegister();
+				if (v.getType().isReferenceType()) {
+					int vIdx = domV.indexOf(v);
+					assert (vIdx >= 0);
+					add(iIdx, vIdx);
+				}
+			}
+		}
+	}
+}

### 2011-03-11 main/src/chord/analyses/facts/argret/RelIinvkRet.java

@@ -1,0 +1,47 @@
+/*
+ * Copyright (c) 2008-2010, Intel Corporation.
+ * Copyright (c) 2006-2007, The Trustees of Stanford University.
+ * All rights reserved.
+ * Licensed under the terms of the New BSD License.
+ */
+package chord.analyses.facts.argret;
+
+import joeq.Compiler.Quad.Quad;
+import joeq.Compiler.Quad.Operand.RegisterOperand;
+import joeq.Compiler.Quad.Operator.Invoke;
+import joeq.Compiler.Quad.RegisterFactory.Register;
+
+import chord.analyses.facts.invk.DomI;
+import chord.analyses.facts.var.DomV;
+import chord.project.Chord;
+import chord.project.analyses.ProgramRel;
+
+/**
+ * Relation containing each tuple (i,z,v) such that local variable v
+ * is the zth return variable of method invocation statement i.
+ *
+ * @author Mayur Naik (mhn@cs.stanford.edu)
+ */
+@Chord(
+	name = "IinvkRet",
+	sign = "I0,Z0,V0:I0_V0_Z0"
+)
+public class RelIinvkRet extends ProgramRel {
+	public void fill() {
+		DomI domI = (DomI) doms[0];
+		DomV domV = (DomV) doms[2];
+		int numI = domI.size();
+		for (int iIdx = 0; iIdx < numI; iIdx++) {
+			Quad q = (Quad) domI.get(iIdx);
+			RegisterOperand vo = Invoke.getDest(q);
+			if (vo != null) {
+				Register v = vo.getRegister();
+				if (v.getType().isReferenceType()) {
+					int vIdx = domV.indexOf(v);
+					assert (vIdx >= 0);
+					add(iIdx, 0, vIdx);
+				}
+			}
+		}
+	}
+}

### 2011-03-11 main/src/chord/analyses/facts/argret/RelMmethArg.java

@@ -1,0 +1,51 @@
+/*
+ * Copyright (c) 2008-2010, Intel Corporation.
+ * Copyright (c) 2006-2007, The Trustees of Stanford University.
+ * All rights reserved.
+ * Licensed under the terms of the New BSD License.
+ */
+package chord.analyses.facts.argret;
+
+import joeq.Class.jq_Method;
+import joeq.Compiler.Quad.ControlFlowGraph;
+import joeq.Compiler.Quad.RegisterFactory;
+import joeq.Compiler.Quad.RegisterFactory.Register;
+
+import chord.analyses.facts.method.DomM;
+import chord.analyses.facts.var.DomV;
+import chord.project.Chord;
+import chord.project.analyses.ProgramRel;
+
+/**
+ * Relation containing each tuple (m,z,v) such that local variable
+ * v is the zth argument variable of method m.
+ *
+ * @author Mayur Naik (mhn@cs.stanford.edu)
+ */
+@Chord(
+	name = "MmethArg",
+	sign = "M0,Z0,V0:M0_V0_Z0"
+)
+public class RelMmethArg extends ProgramRel {
+	public void fill() {
+		DomM domM = (DomM) doms[0];
+		DomV domV = (DomV) doms[2];
+		int numM = domM.size();
+		for (int mIdx = 0; mIdx < numM; mIdx++) {
+			jq_Method m = domM.get(mIdx);
+			if (m.isAbstract())
+				continue;
+			ControlFlowGraph cfg = m.getCFG();
+			RegisterFactory rf = cfg.getRegisterFactory();
+			int numArgs = m.getParamTypes().length;
+			for (int zIdx = 0; zIdx < numArgs; zIdx++) {
+				Register v = rf.get(zIdx);
+				if (v.getType().isReferenceType()) {
+					int vIdx = domV.indexOf(v);
+					assert (vIdx >= 0);
+					add(mIdx, zIdx, vIdx);
+				}
+			}
+		}
+	}
+}

### 2011-03-11 main/src/chord/analyses/facts/basicblock/RelPostDomBB.java

@@ -1,0 +1,98 @@
+/*
+ * Copyright (c) 2008-2010, Intel Corporation.
+ * Copyright (c) 2006-2007, The Trustees of Stanford University.
+ * All rights reserved.
+ * Licensed under the terms of the New BSD License.
+ */
+package chord.analyses.facts.basicblock;
+
+import chord.util.ArraySet;
+import java.util.Map;
+import java.util.HashMap;
+import java.util.Set;
+
+import joeq.Class.jq_Method;
+import joeq.Class.jq_Class;
+import joeq.Compiler.Quad.BasicBlock;
+import joeq.Compiler.Quad.ControlFlowGraph;
+import chord.program.visitors.IMethodVisitor;
+import chord.program.Program;
+import chord.project.Chord;
+import chord.project.analyses.ProgramRel;
+import joeq.Util.Templates.List;
+
+/**
+ * Relation containing each pair of basic blocks (b1,b2) in each method
+ * such that b1 is immediate postdominator of b2.
+ *
+ * @author Mayur Naik (mhn@cs.stanford.edu)
+ */
+@Chord(
+	name = "postDomBB",
+	sign = "B0,B1:B0xB1"
+)
+public class RelPostDomBB extends ProgramRel implements IMethodVisitor {
+	private final Map<BasicBlock, Set<BasicBlock>> pdomMap =
+		new HashMap<BasicBlock, Set<BasicBlock>>();
+	public void visit(jq_Class c) { }
+	public void visit(jq_Method m) {
+		if (m.isAbstract())
+			return;
+		// System.out.println("VISITING: " + m);
+		pdomMap.clear();
+		ControlFlowGraph cfg = m.getCFG();
+		BasicBlock exit = cfg.exit();
+		Set<BasicBlock> exitSet = new ArraySet<BasicBlock>(1);
+		exitSet.add(exit);
+		pdomMap.put(exit, exitSet);
+		List.BasicBlock rpo = cfg.reversePostOrder(cfg.entry());
+		int n = rpo.size();
+		Set<BasicBlock> initSet = new ArraySet<BasicBlock>(n);
+		for (int i = 0; i < n; i++) {
+			BasicBlock bb = rpo.getBasicBlock(i);
+			initSet.add(bb);
+		}
+		for (int i = 0; i < n; i++) {
+			BasicBlock bb = rpo.getBasicBlock(i);
+			if (bb != exit)
[…]

### 2011-03-11 main/src/chord/analyses/facts/field/RelStatTF.java

@@ -1,0 +1,35 @@
+/*
+ * Copyright (c) 2008-2010, Intel Corporation.
+ * Copyright (c) 2006-2007, The Trustees of Stanford University.
+ * All rights reserved.
+ * Licensed under the terms of the New BSD License.
+ */
+package chord.analyses.facts.field;
+
+import joeq.Class.jq_Field;
+import joeq.Class.jq_Class;
+import chord.program.visitors.IFieldVisitor;
+import chord.project.Chord;
+import chord.project.analyses.ProgramRel;
+
+/**
+ * Relation containing each tuple (t,f) such that f is a
+ * static field defined in type t.
+ *
+ * @author Mayur Naik (mhn@cs.stanford.edu)
+ */
+@Chord(
+	name = "staticTF",
+	sign = "T0,F0:F0_T0"
+)
+public class RelStatTF extends ProgramRel implements IFieldVisitor {
+	private jq_Class ctnrClass;
+	public void visit(jq_Class c) {
+		ctnrClass = c;
+	}
+	public void visit(jq_Field f) {
+		if (f.isStatic()) {
+			add(ctnrClass, f);
+		}
+	}
+}

### 2011-03-11 main/src/chord/analyses/facts/heapacc/RelEF.java

@@ -1,0 +1,40 @@
+/*
+ * Copyright (c) 2008-2010, Intel Corporation.
+ * Copyright (c) 2006-2007, The Trustees of Stanford University.
+ * All rights reserved.
+ * Licensed under the terms of the New BSD License.
+ */
+package chord.analyses.facts.heapacc;
+
+import joeq.Class.jq_Field;
+import joeq.Compiler.Quad.Quad;
+import chord.analyses.facts.field.DomF;
+import chord.program.Program;
+import chord.project.Chord;
+import chord.project.analyses.ProgramRel;
+
+/**
+ * Relation containing each tuple (e,f) such that statement e
+ * accesses (reads or writes) instance field, static field, or
+ * array element f.
+ *
+ * @author Mayur Naik (mhn@cs.stanford.edu)
+ */
+@Chord(
+	name = "EF",
+	sign = "E0,F0:F0_E0"
+)
+public class RelEF extends ProgramRel {
+	public void fill() {
+		DomE domE = (DomE) doms[0];
+		DomF domF = (DomF) doms[1];
+		int numE = domE.size();
+		for (int eIdx = 0; eIdx < numE; eIdx++) {
+			Quad e = domE.get(eIdx);
+			jq_Field f = e.getField();
+			int fIdx = domF.indexOf(f);
+			assert (fIdx >= 0);
+			add(eIdx, fIdx);
+		}
+	}
+}

### 2011-03-11 main/src/chord/analyses/facts/heapacc/RelEV.java

@@ -1,0 +1,59 @@
+/*
+ * Copyright (c) 2008-2010, Intel Corporation.
+ * Copyright (c) 2006-2007, The Trustees of Stanford University.
+ * All rights reserved.
+ * Licensed under the terms of the New BSD License.
+ */
+package chord.analyses.facts.heapacc;
+
+import joeq.Compiler.Quad.Operator;
+import joeq.Compiler.Quad.Quad;
+import joeq.Compiler.Quad.Operand.RegisterOperand;
+import joeq.Compiler.Quad.Operator.ALoad;
+import joeq.Compiler.Quad.Operator.AStore;
+import joeq.Compiler.Quad.Operator.Getfield;
+import joeq.Compiler.Quad.Operator.Putfield;
+import joeq.Compiler.Quad.RegisterFactory.Register;
+import chord.analyses.facts.var.DomV;
+import chord.project.Chord;
+import chord.project.analyses.ProgramRel;
+
+/**
+ * Relation containing each tuple (e,v) such that statement e
+ * accesses (reads or writes) an instance field or array element
+ * of an object denoted by local variable v.
+ *
+ * @author Mayur Naik (mhn@cs.stanford.edu)
+ */
+@Chord(
+	name = "EV",
+	sign = "E0,V0:E0_V0"
+)
+public class RelEV extends ProgramRel {
+	public void fill() {
+		DomE domE = (DomE) doms[0];
+		DomV domV = (DomV) doms[1];
+		int numE = domE.size();
+		for (int eIdx = 0; eIdx < numE; eIdx++) {
+			Quad q = (Quad) domE.get(eIdx);
+			Operator op = q.getOperator();
+			RegisterOperand bo;
+			if (op instanceof ALoad) {
+				bo = (RegisterOperand) ALoad.getBase(q);
+			} else if (op instanceof Getfield) {
+				bo = (RegisterOperand) Getfield.getBase(q);
+			} else if (op instanceof AStore) {
+				bo = (RegisterOperand) AStore.getBase(q);
+			} else if (op instanceof Putfield) {
+				bo = (RegisterOperand) Putfield.getBase(q);
+			} else
+				bo = null;
+			if (bo != null) {
+				Register b = bo.getRegister();
+				int vIdx = domV.indexOf(b);
+				assert (vIdx >= 0);
+				add(eIdx, vIdx);
+			}
+		}
+	}
+}

### 2011-03-11 main/src/chord/analyses/facts/heapacc/RelPE.java

@@ -1,0 +1,38 @@
+/*
+ * Copyright (c) 2008-2010, Intel Corporation.
+ * Copyright (c) 2006-2007, The Trustees of Stanford University.
+ * All rights reserved.
+ * Licensed under the terms of the New BSD License.
+ */
+package chord.analyses.facts.heapacc;
+
+import joeq.Compiler.Quad.Quad;
+import chord.analyses.facts.point.DomP;
+import chord.project.Chord;
+import chord.project.analyses.ProgramRel;
+
+/**
+ * Relation containing each tuple (p,e) such that the statement
+ * at program point p is a heap access statement e that
+ * accesses (reads or writes) an instance field, a static field,
+ * or an array element.
+ *
+ * @author Mayur Naik (mhn@cs.stanford.edu)
+ */
+@Chord(
+	name = "PE",
+	sign = "P0,E0:E0_P0"
+)
+public class RelPE extends ProgramRel {
+	public void fill() {
+		DomP domP = (DomP) doms[0];
+		DomE domE = (DomE) doms[1];
+		int numE = domE.size();
+		for (int eIdx = 0; eIdx < numE; eIdx++) {
+			Quad e = (Quad) domE.get(eIdx);
+			int pIdx = domP.indexOf(e);
+			assert (pIdx >= 0);
+			add(pIdx, eIdx);
+		}
+	}
+}

### 2011-03-11 main/src/chord/analyses/facts/heapacc/RelWriteE.java

@@ -1,0 +1,36 @@
+/*
+ * Copyright (c) 2008-2010, Intel Corporation.
+ * Copyright (c) 2006-2007, The Trustees of Stanford University.
+ * All rights reserved.
+ * Licensed under the terms of the New BSD License.
+ */
+package chord.analyses.facts.heapacc;
+
+import joeq.Compiler.Quad.Operator;
+import joeq.Compiler.Quad.Quad;
+import chord.program.Program;
+import chord.project.Chord;
+import chord.project.analyses.ProgramRel;
+
+/**
+ * Relation containing all statements that write to an instance
+ * field, static field, or array element.
+ * 
+ * @author Mayur Naik (mhn@cs.stanford.edu)
+ */
+@Chord(
+	name = "writeE",
+	sign = "E0"
+)
+public class RelWriteE extends ProgramRel {
+	public void fill() {
+		DomE domE = (DomE) doms[0];
+		int numE = domE.size();
+		for (int eIdx = 0; eIdx < numE; eIdx++) {
+			Quad e = (Quad) domE.get(eIdx);
+			Operator op = e.getOperator();
+			if (op.isWrHeapInst())
+				add(eIdx);
+		}
+	}
+}

### 2011-03-11 main/src/chord/analyses/facts/inst/RelMgetClassInvkInst.java

@@ -1,0 +1,60 @@
+/*
+ * Copyright (c) 2008-2010, Intel Corporation.
+ * Copyright (c) 2006-2007, The Trustees of Stanford University.
+ * All rights reserved.
+ * Licensed under the terms of the New BSD License.
+ */
+package chord.analyses.facts.inst;
+
+import java.util.List;
+
+import joeq.Class.jq_Class;
+import joeq.Class.jq_Method;
+import joeq.Compiler.Quad.Quad;
+import joeq.Compiler.Quad.Operator.Invoke;
+import joeq.Compiler.Quad.Operand.RegisterOperand;
+
+import chord.analyses.facts.invk.DomI;
+import chord.analyses.facts.method.DomM;
+import chord.program.visitors.IInvokeInstVisitor;
+import chord.program.Program;
+import chord.project.Chord;
+import chord.project.analyses.ProgramRel;
+
+/**
+ * Relation containing each tuple (m,i) such that method m contains a
+ * statement i invoking instance method getClass() defined in class
+ * java.lang.Object.
+ *
+ * @author Mayur Naik (mhn@cs.stanford.edu)
+ */
+@Chord(
+	name = "MgetClassInvkInst",
+	sign = "M0,I0:M0xI0"
+)
+public class RelMgetClassInvkInst extends ProgramRel
+		implements IInvokeInstVisitor {
+	private DomM domM;
+	private DomI domI;
+	private jq_Method ctnrMethod;
+	public void init() {
+		domM = (DomM) doms[0];
+		domI = (DomI) doms[1];
+	}
+	public void visit(jq_Class c) { }
+	public void visit(jq_Method m) {
+		ctnrMethod = m;
+	}
+	public void visitInvokeInst(Quad q) {
+		jq_Method meth = Invoke.getMethod(q).getMethod();
+		if (meth.getName().toString().equals("getClass") &&
+			meth.getDesc().toString().equals("()Ljava/lang/Class;") &&
+			meth.getDeclaringClass().getName().equals("java.lang.Object")) {
+			int mIdx = domM.indexOf(ctnrMethod);
+			assert (mIdx >= 0);
+			int iIdx = domI.indexOf(q);
+			assert (iIdx >= 0);
+			add(mIdx, iIdx);
+		}
+	}
+}

### 2011-03-11 main/src/chord/analyses/facts/inst/RelPgetStatFldInst.java

@@ -1,0 +1,64 @@
+/*
+ * Copyright (c) 2008-2010, Intel Corporation.
+ * Copyright (c) 2006-2007, The Trustees of Stanford University.
+ * All rights reserved.
+ * Licensed under the terms of the New BSD License.
+ */
+package chord.analyses.facts.inst;
+
+import joeq.Class.jq_Class;
+import joeq.Class.jq_Field;
+import joeq.Class.jq_Method;
+import joeq.Compiler.Quad.Operator;
+import joeq.Compiler.Quad.Quad;
+import joeq.Compiler.Quad.Operand.RegisterOperand;
+import joeq.Compiler.Quad.Operand.FieldOperand;
+import joeq.Compiler.Quad.Operator.Getstatic;
+import joeq.Compiler.Quad.RegisterFactory.Register;
+import chord.analyses.facts.field.DomF;
+import chord.analyses.facts.point.DomP;
+import chord.analyses.facts.var.DomV;
+import chord.program.visitors.IHeapInstVisitor;
+import chord.project.Chord;
+import chord.project.analyses.ProgramRel;
+
+/**
+ * Relation containing each tuple (p,v,f) such that the statement
+ * at program point p is of the form <tt>v = f</tt>.
+ *
+ * @author Mayur Naik (mhn@cs.stanford.edu)
+ */
+@Chord(
+	name = "PgetStatFldInst",
+	sign = "P0,V0,F0:F0_P0_V0"
+)
+public class RelPgetStatFldInst extends ProgramRel
+		implements IHeapInstVisitor {
+	private DomP domP;
+	private DomV domV;
+	private DomF domF;
+	public void init() {
+		domP = (DomP) doms[0];
+		domV = (DomV) doms[1];
+		domF = (DomF) doms[2];
+	}
+	public void visit(jq_Class c) { }
+	public void visit(jq_Method m) { }
+	public void visitHeapInst(Quad q) {
+		Operator op = q.getOperator();
+		if (op instanceof Getstatic) {
+			jq_Field f = Getstatic.getField(q).getField();
+			if (f.getType().isReferenceType()) {
+				RegisterOperand lo = Getstatic.getDest(q);
+				Register l = lo.getRegister();
+				int pIdx = domP.indexOf(q);
+				assert (pIdx >= 0);
+				int lIdx = domV.indexOf(l);
+				assert (lIdx >= 0);
[…]

### 2011-03-11 main/src/chord/analyses/facts/inst/RelPputStatFldInst.java

@@ -1,0 +1,68 @@
+/*
+ * Copyright (c) 2008-2010, Intel Corporation.
+ * Copyright (c) 2006-2007, The Trustees of Stanford University.
+ * All rights reserved.
+ * Licensed under the terms of the New BSD License.
+ */
+package chord.analyses.facts.inst;
+
+import joeq.Class.jq_Method;
+import joeq.Class.jq_Class;
+import joeq.Class.jq_Field;
+import joeq.Compiler.Quad.Operator;
+import joeq.Compiler.Quad.Operator.Putstatic;
+import joeq.Compiler.Quad.Quad;
+import joeq.Compiler.Quad.Operand;
+import joeq.Compiler.Quad.Operand.RegisterOperand;
+import joeq.Compiler.Quad.Operand.FieldOperand;
+import joeq.Compiler.Quad.RegisterFactory.Register;
+import chord.analyses.facts.field.DomF;
+import chord.analyses.facts.point.DomP;
+import chord.analyses.facts.var.DomV;
+import chord.program.visitors.IHeapInstVisitor;
+import chord.project.Chord;
+import chord.project.analyses.ProgramRel;
+
+/**
+ * Relation containing each tuple (p,f,v) such that the statement
+ * at program point p is of the form <tt>f = v</tt>.
+ *
+ * @author Mayur Naik (mhn@cs.stanford.edu)
+ */
+@Chord(
+	name = "PputStatFldInst",
+	sign = "P0,F0,V0:F0_P0_V0"
+)
+public class RelPputStatFldInst extends ProgramRel
+		implements IHeapInstVisitor {
+	private DomP domP;
+	private DomF domF;
+	private DomV domV;
+	public void init() {
+		domP = (DomP) doms[0];
+		domF = (DomF) doms[1];
+		domV = (DomV) doms[2];
+	}
+	public void visit(jq_Class c) { }
+	public void visit(jq_Method m) { }
+	public void visitHeapInst(Quad q) {
+		Operator op = q.getOperator();
+		if (op instanceof Putstatic) {
+			jq_Field f = Putstatic.getField(q).getField();
+			if (f.getType().isReferenceType()) {
+				Operand rx = Putstatic.getSrc(q);
+				if (rx instanceof RegisterOperand) {
+					RegisterOperand ro = (RegisterOperand) rx;
+					Register r = ro.getRegister();
[…]

### 2011-03-11 main/src/chord/analyses/facts/invk/RelInstI.java

@@ -1,0 +1,36 @@
+/*
+ * Copyright (c) 2008-2010, Intel Corporation.
+ * Copyright (c) 2006-2007, The Trustees of Stanford University.
+ * All rights reserved.
+ * Licensed under the terms of the New BSD License.
+ */
+package chord.analyses.facts.invk;
+
+import joeq.Class.jq_Method;
+import joeq.Compiler.Quad.Quad;
+import joeq.Compiler.Quad.Operator.Invoke;
+import chord.project.Chord;
+import chord.project.analyses.ProgramRel;
+
+/**
+ * Relation containing all method invocation statements whose
+ * target is an instance method.
+ * 
+ * @author Mayur Naik (mhn@cs.stanford.edu)
+ */
+@Chord(
+	name = "instI",
+	sign = "I0"
+)
+public class RelInstI extends ProgramRel {
+	public void fill() {
+		DomI domI = (DomI) doms[0];
+		int numI = domI.size();
+		for (int iIdx = 0; iIdx < numI; iIdx++) {
+			Quad q = (Quad) domI.get(iIdx);
+			jq_Method m = Invoke.getMethod(q).getMethod();
+			if (!m.isStatic())
+				add(iIdx);
+		}
+	}
+}

### 2011-03-11 main/src/chord/analyses/facts/method/RelExtraEntryPoints.java

@@ -1,0 +1,155 @@
+/*
+ * Copyright (c) 2010, Ariel Rabkin and Intel Corporation.
+ * You may use and modify this code, subject to the license agreement 
+ * in the "COPYING" file associated with this project.
+ * 
+ */
+
+package chord.analyses.facts.method;
+
+import java.io.*;
+import java.util.LinkedHashSet;
+import java.util.Set;
+import joeq.Class.jq_Class;
+import joeq.Class.jq_Method;
+import joeq.Class.jq_Type;
+import chord.program.ClassHierarchy;
+import chord.program.Program;
+import chord.project.Chord;
+import chord.project.analyses.ProgramRel;
+
+@Chord(
+	name = "MentryPoints",
+	sign = "M0:M0"
+)
+/**
+ * A relation over domain M containing additional entry points for the program.
+ * The values of this relation are derived from the file indicated by property 
+ * chord.entrypoints.file.
+ * 
+ * File should be a list whose entries are class names, interface names, or fully qualified method names.
+ *  (A fully qualified method name is of the form <name>:<desc>@<classname>.)
+ *  If a concrete class is listed, all non-private methods of that class will be added as entry points.
+ *  If an interface or abstract is listed, all public declared methods of that interaface/class in
+ *  all concrete subclasses will be added as entry points.
+ */
+public class RelExtraEntryPoints extends ProgramRel {
+
+	public static String extraMethodsFile = System.getProperty("chord.entrypoints.file");
+	public static String extraMethodsList = System.getProperty("chord.entrypoints");
+	static LinkedHashSet<jq_Method> methods;
+
+	@Override
+	public void fill() {
+		Iterable<jq_Method> publicMethods =  slurpMList(Program.g().getClassHierarchy());
+
+		for(jq_Method m: publicMethods) {
+			super.add(m);
+		}
+	}
+
+
+
+	public static Iterable<jq_Method> slurpMList(ClassHierarchy ch) {
+		if(methods == null)
+			methods = new LinkedHashSet<jq_Method>();
+			else
+				return methods;
+
+		if(extraMethodsList != null) {
+			String[] entries = extraMethodsList.split(",");
+			for(String s: entries)
+				processLine(s, ch);
+		}
+
+		try {
+			if(extraMethodsFile != null) {
+				String s = null;
[…]

### 2011-03-11 main/src/chord/analyses/facts/method/RelStatTM.java

@@ -1,0 +1,33 @@
+/*
+ * Copyright (c) 2008-2010, Intel Corporation.
+ * Copyright (c) 2006-2007, The Trustees of Stanford University.
+ * All rights reserved.
+ * Licensed under the terms of the New BSD License.
+ */
+package chord.analyses.facts.method;
+
+import joeq.Class.jq_Class;
+import joeq.Class.jq_Method;
+import chord.program.visitors.IMethodVisitor;
+import chord.project.Chord;
+import chord.project.analyses.ProgramRel;
+
+/**
+ * Relation containing each tuple (t,m) such that m is a
+ * static method defined in type t.
+ * 
+ * @author Mayur Naik (mhn@cs.stanford.edu)
+ */
+@Chord(
+	name = "staticTM",
+	sign = "T0,M0:M0_T0"
+)
+public class RelStatTM extends ProgramRel implements IMethodVisitor {
+	public void visit(jq_Class c) { }
+	public void visit(jq_Method m) {
+		if (m.isStatic()) {
+			jq_Class t = m.getDeclaringClass();
+			add(t, m);
+		}
+	}
+}

### 2011-03-11 main/src/chord/analyses/facts/point/RelMPtail.java

@@ -1,0 +1,37 @@
+/*
+ * Copyright (c) 2008-2010, Intel Corporation.
+ * Copyright (c) 2006-2007, The Trustees of Stanford University.
+ * All rights reserved.
+ * Licensed under the terms of the New BSD License.
+ */
+package chord.analyses.facts.point;
+
+import joeq.Class.jq_Class;
+import joeq.Class.jq_Method;
+import joeq.Compiler.Quad.EntryOrExitBasicBlock;
+import joeq.Compiler.Quad.ControlFlowGraph;
+import chord.program.visitors.IMethodVisitor;
+import chord.project.Chord;
+import chord.project.analyses.ProgramRel;
+
+/**
+ * Relation containing each tuple (m,p) such that statement p is
+ * the unique exit basic block of method m.
+ *
+ * @author Mayur Naik (mhn@cs.stanford.edu)
+ */
+@Chord(
+	name = "MPtail",
+	sign = "M0,P0:M0xP0"
+)
+public class RelMPtail extends ProgramRel
+		implements IMethodVisitor {
+	public void visit(jq_Class c) { }
+	public void visit(jq_Method m) {
+		if (m.isAbstract())
+			return;
+		ControlFlowGraph cfg = m.getCFG();
+		EntryOrExitBasicBlock bx = cfg.exit();
+		add(m, bx);
+	}
+}

### 2011-03-11 main/src/chord/analyses/facts/type/DomT.java

@@ -1,0 +1,44 @@
+/*
+ * Copyright (c) 2008-2010, Intel Corporation.
+ * Copyright (c) 2006-2007, The Trustees of Stanford University.
+ * All rights reserved.
+ * Licensed under the terms of the New BSD License.
+ */
+package chord.analyses.facts.type;
+
+import joeq.Class.jq_Type;
+import joeq.Class.jq_Class;
+import chord.program.Program;
+import chord.project.Chord;
+import chord.project.analyses.ProgramDom;
+import chord.util.IndexSet;
+
+/**
+ * Domain of classes.
+ * 
+ * @author Mayur Naik (mhn@cs.stanford.edu)
+ */
+@Chord(
+	name = "T"
+)
+public class DomT extends ProgramDom<jq_Type> {
+	public void fill() {
+		Program program = Program.g();
+		IndexSet<jq_Type> types = program.getTypes();
+		for (jq_Type t : types)
+			add(t);
+	}
+	public String toXMLAttrsString(jq_Type t) {
+		String name = t.getName();
+		String file;
+		if (t instanceof jq_Class) {
+			jq_Class c = (jq_Class) t;
+			file = c.getSourceFileName();
+		} else
+			file = "";
+		int line = 0;  // TODO
+		return "name=\"" + name +
+			"\" file=\"" + file +
+			"\" line=\"" + line + "\"";
+	}
+}

### 2011-03-11 main/src/chord/analyses/facts/type/RelClassT.java

@@ -1,0 +1,29 @@
+/*
+ * Copyright (c) 2008-2010, Intel Corporation.
+ * Copyright (c) 2006-2007, The Trustees of Stanford University.
+ * All rights reserved.
+ * Licensed under the terms of the New BSD License.
+ */
+package chord.analyses.facts.type;
+
+import joeq.Class.jq_Class;
+import chord.program.visitors.IClassVisitor;
+import chord.project.Chord;
+import chord.project.analyses.ProgramRel;
+
+/**
+ * Relation containing each (concrete or abstract) class type
+ * (as opposed to interface types, primitive types, etc.).
+ *
+ * @author Mayur Naik (mhn@cs.stanford.edu)
+ */
+@Chord(
+	name = "classT",
+	sign = "T0"
+)
+public class RelClassT extends ProgramRel implements IClassVisitor {
+	public void visit(jq_Class c) {
+		if (!c.isInterface())
+			add(c);
+	}
+}

### 2011-03-11 main/src/chord/analyses/facts/type/RelInterfaceT.java

@@ -1,0 +1,29 @@
+/*
+ * Copyright (c) 2008-2010, Intel Corporation.
+ * Copyright (c) 2006-2007, The Trustees of Stanford University.
+ * All rights reserved.
+ * Licensed under the terms of the New BSD License.
+ */
+package chord.analyses.facts.type;
+
+import joeq.Class.jq_Class;
+import chord.program.visitors.IClassVisitor;
+import chord.project.Chord;
+import chord.project.analyses.ProgramRel;
+
+/**
+ * Relation containing each interface type.
+ *
+ * @author Mayur Naik (mhn@cs.stanford.edu)
+ */
+@Chord(
+	name = "interfaceT",
+	sign = "T0"
+)
+public class RelInterfaceT extends ProgramRel
+		implements IClassVisitor {
+	public void visit(jq_Class c) {
+		if (c.isInterface())
+			add(c);
+	}
+}

### 2011-03-11 main/src/chord/analyses/facts/var/RelMV.java

@@ -1,0 +1,35 @@
+/*
+ * Copyright (c) 2008-2010, Intel Corporation.
+ * Copyright (c) 2006-2007, The Trustees of Stanford University.
+ * All rights reserved.
+ * Licensed under the terms of the New BSD License.
+ */
+package chord.analyses.facts.var;
+
+import joeq.Class.jq_Class;
+import joeq.Class.jq_Method;
+import chord.project.Chord;
+import chord.project.analyses.ProgramRel;
+import joeq.Compiler.Quad.RegisterFactory.Register;
+import chord.analyses.facts.var.DomV;
+
+/**
+ * Relation containing each tuple (m,v) such that method m
+ * declares local variable v, that is, v is either an
+ * argument or temporary variable of m.
+ *
+ * @author Mayur Naik (mhn@cs.stanford.edu)
+ */
+@Chord(
+	name = "MV",
+	sign = "M0,V0:M0_V0"
+)
+public class RelMV extends ProgramRel {
+	public void fill() {
+		DomV domV = (DomV) doms[1];
+		for (Register v : domV) {
+			jq_Method m = domV.getMethod(v);
+			add(m, v);
+		}
+	}
+}

### 2011-03-11 main/src/chord/analyses/facts/var/RelVT.java

@@ -1,0 +1,109 @@
+/*
+ * Copyright (c) 2008-2010, Intel Corporation.
+ * Copyright (c) 2006-2007, The Trustees of Stanford University.
+ * All rights reserved.
+ * Licensed under the terms of the New BSD License.
+ */
+package chord.analyses.facts.var;
+
+import java.util.Map;
+import java.util.HashMap;
+
+import chord.program.Program;
+import chord.program.visitors.IMethodVisitor;
+import chord.project.Chord;
+import chord.project.analyses.ProgramRel;
+import chord.project.analyses.ProgramDom;
+import chord.util.CollectionUtils;
+import chord.util.ArraySet;
+import chord.project.Chord;
+import joeq.Compiler.Quad.RegisterFactory;
+import joeq.Compiler.Quad.BasicBlock;
+import joeq.Compiler.Quad.ControlFlowGraph;
+import joeq.Compiler.Quad.Operator;
+import joeq.Compiler.Quad.Quad;
+import joeq.Compiler.Quad.Operand;
+import joeq.Compiler.Quad.Operand.RegisterOperand;
+import joeq.Compiler.Quad.Operand.ParamListOperand;
+import joeq.Compiler.Quad.RegisterFactory.Register;
+import joeq.Class.jq_Class;
+import joeq.Class.jq_Method;
+import joeq.Class.jq_Type;
+import joeq.Class.jq_Reference;
+import joeq.Util.Templates.ListIterator;
+
+/**
+ * Relation containing each tuple (v,t) such that local variable v
+ * of reference type has type t.
+ * If SSA is used (system property chord.ssa is set to true) then it is
+ * guaranteed that each local variable v has a unique type t.
+ * 
+ * These types will usually be more precise than "Object".
+ *
+ * @author Mayur Naik (mhn@cs.stanford.edu)
+ */
+@Chord(
+	name = "VT",
+	sign = "V0,T0:T0_V0"
+)
+public class RelVT extends ProgramRel implements IMethodVisitor {
+	private jq_Reference javaLangObject;
+	public void init() {
+		javaLangObject = Program.g().getClass("java.lang.Object");
+		assert (javaLangObject != null);
+	}
+	public void visit(jq_Class c) { }
+	public void visit(jq_Method m) {
+		if (m.isAbstract())
+			return;
+		ControlFlowGraph cfg = m.getCFG();
[…]

### 2011-03-11 main/src/chord/analyses/heapacc/DomE.java

@@ -1,66 +1,0 @@
-/*
- * Copyright (c) 2008-2010, Intel Corporation.
- * Copyright (c) 2006-2007, The Trustees of Stanford University.
- * All rights reserved.
- * Licensed under the terms of the New BSD License.
- */
-package chord.analyses.heapacc;
-
-import joeq.Class.jq_Class;
-import joeq.Class.jq_Method;
-import joeq.Compiler.Quad.Quad;
-import joeq.Compiler.Quad.Operator;
-import joeq.Compiler.Quad.Operator.Getfield;
-import joeq.Compiler.Quad.Operator.Putfield;
-import joeq.Compiler.Quad.Operand.RegisterOperand;
-import chord.program.visitors.IHeapInstVisitor;
-import chord.analyses.method.DomM;
-import chord.project.Chord;
-import chord.project.ClassicProject;
-import chord.project.Config;
-import chord.project.analyses.ProgramDom;
-
-/**
- * Domain of statements that access (read or write) an
- * instance field, a static field, or an array element.
- * 
- * @author Mayur Naik (mhn@cs.stanford.edu)
- */
-@Chord(
-	name = "E",
-	consumes = { "M" }
-)
-public class DomE extends ProgramDom<Quad> implements IHeapInstVisitor {
-	protected DomM domM;
-	public void init() {
-		domM = (DomM) (Config.classic ?
-			ClassicProject.g().getTrgt("M") : consumes[0]);
-	}
-	public void visit(jq_Class c) { }
-	public void visit(jq_Method m) { }
-	public void visitHeapInst(Quad q) {
-		Operator op = q.getOperator();
-		if (op instanceof Getfield) {
-			if (!(Getfield.getBase(q) instanceof RegisterOperand))
-				return;
-		}
-		if (op instanceof Putfield) {
-			if (!(Putfield.getBase(q) instanceof RegisterOperand))
-				return;
-		}
-		add(q);
-	}
-	public String toUniqueString(Quad q) {
-		return q.toByteLocStr();
-	}
-	public String toXMLAttrsString(Quad q) {
-		Operator op = q.getOperator();
-		jq_Method m = q.getMethod();
-		String file = m.getDeclaringClass().getSourceFileName();
-		int line = q.getLineNumber();
-		int mIdx = domM.indexOf(m);
-		return "file=\"" + file + "\" " + "line=\"" + line + "\" " +
[…]

### 2011-03-11 main/src/chord/analyses/heapacc/RelEV.java

@@ -1,59 +1,0 @@
-/*
- * Copyright (c) 2008-2010, Intel Corporation.
- * Copyright (c) 2006-2007, The Trustees of Stanford University.
- * All rights reserved.
- * Licensed under the terms of the New BSD License.
- */
-package chord.analyses.heapacc;
-
-import joeq.Compiler.Quad.Operator;
-import joeq.Compiler.Quad.Quad;
-import joeq.Compiler.Quad.Operand.RegisterOperand;
-import joeq.Compiler.Quad.Operator.ALoad;
-import joeq.Compiler.Quad.Operator.AStore;
-import joeq.Compiler.Quad.Operator.Getfield;
-import joeq.Compiler.Quad.Operator.Putfield;
-import joeq.Compiler.Quad.RegisterFactory.Register;
-import chord.analyses.var.DomV;
-import chord.project.Chord;
-import chord.project.analyses.ProgramRel;
-
-/**
- * Relation containing each tuple (e,v) such that statement e
- * accesses (reads or writes) an instance field or array element
- * of an object denoted by local variable v.
- *
- * @author Mayur Naik (mhn@cs.stanford.edu)
- */
-@Chord(
-	name = "EV",
-	sign = "E0,V0:E0_V0"
-)
-public class RelEV extends ProgramRel {
-	public void fill() {
-		DomE domE = (DomE) doms[0];
-		DomV domV = (DomV) doms[1];
-		int numE = domE.size();
-		for (int eIdx = 0; eIdx < numE; eIdx++) {
-			Quad q = (Quad) domE.get(eIdx);
-			Operator op = q.getOperator();
-			RegisterOperand bo;
-			if (op instanceof ALoad) {
-				bo = (RegisterOperand) ALoad.getBase(q);
-			} else if (op instanceof Getfield) {
-				bo = (RegisterOperand) Getfield.getBase(q);
-			} else if (op instanceof AStore) {
-				bo = (RegisterOperand) AStore.getBase(q);
-			} else if (op instanceof Putfield) {
-				bo = (RegisterOperand) Putfield.getBase(q);
-			} else
-				bo = null;
-			if (bo != null) {
-				Register b = bo.getRegister();
-				int vIdx = domV.indexOf(b);
-				assert (vIdx >= 0);
-				add(eIdx, vIdx);
-			}
-		}
-	}
-}

### 2011-03-11 main/src/chord/analyses/heapacc/RelPE.java

@@ -1,38 +1,0 @@
-/*
- * Copyright (c) 2008-2010, Intel Corporation.
- * Copyright (c) 2006-2007, The Trustees of Stanford University.
- * All rights reserved.
- * Licensed under the terms of the New BSD License.
- */
-package chord.analyses.heapacc;
-
-import joeq.Compiler.Quad.Quad;
-import chord.analyses.point.DomP;
-import chord.project.Chord;
-import chord.project.analyses.ProgramRel;
-
-/**
- * Relation containing each tuple (p,e) such that the statement
- * at program point p is a heap access statement e that
- * accesses (reads or writes) an instance field, a static field,
- * or an array element.
- *
- * @author Mayur Naik (mhn@cs.stanford.edu)
- */
-@Chord(
-	name = "PE",
-	sign = "P0,E0:E0_P0"
-)
-public class RelPE extends ProgramRel {
-	public void fill() {
-		DomP domP = (DomP) doms[0];
-		DomE domE = (DomE) doms[1];
-		int numE = domE.size();
-		for (int eIdx = 0; eIdx < numE; eIdx++) {
-			Quad e = (Quad) domE.get(eIdx);
-			int pIdx = domP.indexOf(e);
-			assert (pIdx >= 0);
-			add(pIdx, eIdx);
-		}
-	}
-}

### 2011-03-11 main/src/chord/analyses/heapacc/RelStatFldE.java

@@ -1,38 +1,0 @@
-/*
- * Copyright (c) 2008-2010, Intel Corporation.
- * Copyright (c) 2006-2007, The Trustees of Stanford University.
- * All rights reserved.
- * Licensed under the terms of the New BSD License.
- */
-package chord.analyses.heapacc;
-
-import joeq.Compiler.Quad.Operator;
-import joeq.Compiler.Quad.Quad;
-import joeq.Compiler.Quad.Operator.Getstatic;
-import joeq.Compiler.Quad.Operator.Putstatic;
-import chord.project.Chord;
-import chord.project.analyses.ProgramRel;
-
-/**
- * Relation containing each statement that accesses (reads or writes)
- * a static field (as opposed to an instance field or array element).
- *
- * @author Mayur Naik (mhn@cs.stanford.edu)
- */
-@Chord(
-	name = "statFldE",
-	sign = "E0"
-)
-public class RelStatFldE extends ProgramRel {
-	public void fill() {
-		DomE domE = (DomE) doms[0];
-		int numE = domE.size();
-		for (int eIdx = 0; eIdx < numE; eIdx++) {
-			Quad e = (Quad) domE.get(eIdx);
-			Operator op = e.getOperator();
-			if (op instanceof Getstatic || op instanceof Putstatic) {
-				add(eIdx);
-			}
-		}
-	}
-}

### 2011-03-11 main/src/chord/analyses/inst/RelMcheckCastInst.java

@@ -1,72 +1,0 @@
-/*
- * Copyright (c) 2008-2010, Intel Corporation.
- * Copyright (c) 2006-2007, The Trustees of Stanford University.
- * All rights reserved.
- * Licensed under the terms of the New BSD License.
- */
-package chord.analyses.inst;
-
-import joeq.Class.jq_Class;
-import joeq.Class.jq_Reference;
-import joeq.Class.jq_Method;
-import joeq.Class.jq_Type;
-import joeq.Compiler.Quad.Operand;
-import joeq.Compiler.Quad.Quad;
-import joeq.Compiler.Quad.Operand.RegisterOperand;
-import joeq.Compiler.Quad.Operator.CheckCast;
-import joeq.Compiler.Quad.RegisterFactory.Register;
-import chord.analyses.method.DomM;
-import chord.analyses.type.DomT;
-import chord.analyses.var.DomV;
-import chord.program.visitors.IMoveInstVisitor;
-import chord.project.Chord;
-import chord.project.analyses.ProgramRel;
-import chord.program.visitors.ICastInstVisitor;
-
-/**
- * Relation containing each tuple (m,v1,t,v2) such that method m
- * contains a statement of the form <tt>v1 = (t) v2</tt>.
- *
- * @author Omer Tripp (omertrip@post.tau.ac.il)
- */
-@Chord(
-	name = "McheckCastInst",
-	sign = "M0,V0,T0,V1:M0_T0_V0xV1"
-)
-public class RelMcheckCastInst extends ProgramRel
-		implements ICastInstVisitor {
-	private jq_Method ctnrMethod;
-	private DomM domM;
-	private DomV domV;
-	private DomT domT;
-	public void init() {
-		domM = (DomM) doms[0];
-		domV = (DomV) doms[1];
-		domT = (DomT) doms[2];
-	}
-	public void visit(jq_Class c) { }
-	public void visit(jq_Method m) {
-		ctnrMethod = m;
-	}
-	public void visitCastInst(Quad q) {
-		Operand rx = CheckCast.getSrc(q);
-		if (rx instanceof RegisterOperand) {
-			jq_Type t = CheckCast.getType(q).getType();
-			if (t instanceof jq_Reference) {
-				RegisterOperand ro = (RegisterOperand) rx;
-				Register r = ro.getRegister();
-				RegisterOperand lo = CheckCast.getDest(q);
[…]

### 2011-03-11 main/src/chord/analyses/inst/RelPobjValAsgnInst.java

@@ -1,54 +1,0 @@
-/*
- * Copyright (c) 2008-2010, Intel Corporation.
- * Copyright (c) 2006-2007, The Trustees of Stanford University.
- * All rights reserved.
- * Licensed under the terms of the New BSD License.
- */
-package chord.analyses.inst;
-
-import joeq.Compiler.Quad.Operator;
-import joeq.Compiler.Quad.Quad;
-import joeq.Compiler.Quad.Operand.RegisterOperand;
-import joeq.Compiler.Quad.Operator.New;
-import joeq.Compiler.Quad.Operator.NewArray;
-import joeq.Compiler.Quad.RegisterFactory.Register;
-import chord.analyses.alloc.DomH;
-import chord.analyses.point.DomP;
-import chord.analyses.var.DomV;
-import chord.project.Chord;
-import chord.project.analyses.ProgramRel;
-
-/**
- * Relation containing each tuple (p,v,h) such that the statement
- * at program point p is an object allocation statement h which
- * assigns to local variable v.
- *
- * @author Mayur Naik (mhn@cs.stanford.edu)
- */
-@Chord(
-	name = "PobjValAsgnInst",
-	sign = "P0,V0,H0:P0_V0_H0"
-)
-public class RelPobjValAsgnInst extends ProgramRel {
-	public void fill() {
-		DomP domP = (DomP) doms[0];
-		DomV domV = (DomV) doms[1];
-		DomH domH = (DomH) doms[2];
-		int numH = domH.size();
-		for (int hIdx = 1; hIdx < numH; hIdx++) {
-			Quad h = (Quad) domH.get(hIdx);
-			int pIdx = domP.indexOf(h);
-			assert (pIdx >= 0);
-			Operator op = h.getOperator();
-			RegisterOperand vo;
-			if (op instanceof New)
-				vo = New.getDest(h);
-			else
-				vo = NewArray.getDest(h);
-			Register v = vo.getRegister();
-			int vIdx = domV.indexOf(v);
-			assert (vIdx >= 0);
-			add(pIdx, vIdx, hIdx);
-		}
-	}
-}

### 2011-03-11 main/src/chord/analyses/invk/DomI.java

@@ -1,53 +1,0 @@
-/*
- * Copyright (c) 2008-2010, Intel Corporation.
- * Copyright (c) 2006-2007, The Trustees of Stanford University.
- * All rights reserved.
- * Licensed under the terms of the New BSD License.
- */
-package chord.analyses.invk;
-
-import joeq.Class.jq_Class;
-import joeq.Class.jq_Method;
-import joeq.Compiler.Quad.Operator;
-import joeq.Compiler.Quad.Quad;
-import chord.program.visitors.IInvokeInstVisitor;
-import chord.analyses.method.DomM;
-import chord.project.Chord;
-import chord.project.Config;
-import chord.project.ClassicProject;
-import chord.project.analyses.ProgramDom;
-
-/**
- * Domain of method invocation statements.
- * 
- * @author Mayur Naik (mhn@cs.stanford.edu)
- */
-@Chord(
-	name = "I",
-	consumes = { "M" }
-)
-public class DomI extends ProgramDom<Quad> implements IInvokeInstVisitor {
-	protected DomM domM;
-	public void init() {
-		domM = (DomM) (Config.classic ?
-			ClassicProject.g().getTrgt("M") : consumes[0]);
-	}
-	public void visit(jq_Class c) { }
-	public void visit(jq_Method m) { }
-	public void visitInvokeInst(Quad q) {
-		add(q);
-	}
-	public String toUniqueString(Quad q) {
-		return q.toByteLocStr();
-	}
-	public String toXMLAttrsString(Quad q) {
-		Operator op = q.getOperator();
-		jq_Method m = q.getMethod();
-		String file = m.getDeclaringClass().getSourceFileName();
-		int line = q.getLineNumber();
-		int mIdx = domM.indexOf(m);
-		return "file=\"" + file + "\" " + "line=\"" + line + "\" " +
-			"Mid=\"M" + mIdx + "\"" +
-			" rdwr=\"" + (op.isWrHeapInst() ? "Wr" : "Rd") + "\"";
-	}
-}

### 2011-03-11 main/src/chord/analyses/lock/RelPL.java

@@ -1,38 +1,0 @@
-/*
- * Copyright (c) 2008-2010, Intel Corporation.
- * Copyright (c) 2006-2007, The Trustees of Stanford University.
- * All rights reserved.
- * Licensed under the terms of the New BSD License.
- */
-package chord.analyses.lock;
-
-import joeq.Compiler.Quad.Inst;
-import chord.analyses.point.DomP;
-import chord.project.Chord;
-import chord.project.analyses.ProgramRel;
-
-/**
- * Relation containing each tuple (p,e) such that the statement
- * at program point p is a heap access statement e that
- * accesses (reads or writes) an instance field, a static field,
- * or an array element.
- *
- * @author Mayur Naik (mhn@cs.stanford.edu)
- */
-@Chord(
-	name = "PL",
-	sign = "P0,L0:P0xL0"
-)
-public class RelPL extends ProgramRel {
-	public void fill() {
-		DomP domP = (DomP) doms[0];
-		DomL domL = (DomL) doms[1];
-		int numL = domL.size();
-		for (int lIdx = 0; lIdx < numL; lIdx++) {
-			Inst i = domL.get(lIdx);
-			int pIdx = domP.indexOf(i);
-			assert (pIdx >= 0);
-			add(pIdx, lIdx);
-		}
-	}
-}

### 2011-03-11 main/src/chord/analyses/method/RelInstM.java

@@ -1,30 +1,0 @@
-/*
- * Copyright (c) 2008-2010, Intel Corporation.
- * Copyright (c) 2006-2007, The Trustees of Stanford University.
- * All rights reserved.
- * Licensed under the terms of the New BSD License.
- */
-package chord.analyses.method;
-
-import joeq.Class.jq_Class;
-import joeq.Class.jq_Method;
-import chord.program.visitors.IMethodVisitor;
-import chord.project.Chord;
-import chord.project.analyses.ProgramRel;
-
-/**
- * Relation containing all instance (as opposed to static) methods.
- *
- * @author Mayur Naik (mhn@cs.stanford.edu)
- */
-@Chord(
-	name = "instM",
-	sign = "M0"
-)
-public class RelInstM extends ProgramRel implements IMethodVisitor {
-	public void visit(jq_Class c) { }
-	public void visit(jq_Method m) {
-		if (!m.isStatic())
-			add(m);
-	}
-}

### 2011-03-11 main/src/chord/analyses/method/RelStatM.java

@@ -1,30 +1,0 @@
-/*
- * Copyright (c) 2008-2010, Intel Corporation.
- * Copyright (c) 2006-2007, The Trustees of Stanford University.
- * All rights reserved.
- * Licensed under the terms of the New BSD License.
- */
-package chord.analyses.method;
-
-import joeq.Class.jq_Class;
-import joeq.Class.jq_Method;
-import chord.program.visitors.IMethodVisitor;
-import chord.project.Chord;
-import chord.project.analyses.ProgramRel;
-
-/**
- * Relation containing all static (as opposed to instance) methods.
- *
- * @author Mayur Naik (mhn@cs.stanford.edu)
- */
-@Chord(
-	name = "statM",
-	sign = "M0"
-)
-public class RelStatM extends ProgramRel implements IMethodVisitor {
-	public void visit(jq_Class c) { }
-	public void visit(jq_Method m) {
-		if (m.isStatic())
-			add(m);
-	}
-}

### 2011-03-11 main/src/chord/analyses/method/RelSyncM.java

@@ -1,30 +1,0 @@
-/*
- * Copyright (c) 2008-2010, Intel Corporation.
- * Copyright (c) 2006-2007, The Trustees of Stanford University.
- * All rights reserved.
- * Licensed under the terms of the New BSD License.
- */
-package chord.analyses.method;
-
-import joeq.Class.jq_Class;
-import joeq.Class.jq_Method;
-import chord.program.visitors.IMethodVisitor;
-import chord.project.Chord;
-import chord.project.analyses.ProgramRel;
-
-/**
- * Relation containing all synchronized methods.
- *
- * @author Mayur Naik (mhn@cs.stanford.edu)
- */
-@Chord(
-	name = "syncM",
-	sign = "M0"
-)
-public class RelSyncM extends ProgramRel implements IMethodVisitor {
-	public void visit(jq_Class c) { }
-	public void visit(jq_Method m) {
-		if (m.isSynchronized())
-			add(m);
-	}
-}

### 2011-03-11 main/src/chord/analyses/reflect/ConNewInstAnalysis.java

@@ -14,9 +14,9 @@ import joeq.Class.jq_Class;
 import joeq.Class.jq_InstanceMethod;
 import joeq.Compiler.Quad.Quad;
 
-import chord.analyses.invk.DomI;
-import chord.analyses.alloc.DomH;
-import chord.analyses.method.DomM;
+import chord.analyses.facts.alloc.DomH;
+import chord.analyses.facts.invk.DomI;
+import chord.analyses.facts.method.DomM;
 import chord.project.ClassicProject;
 import chord.program.Program;
 import chord.project.Chord;

### 2011-03-11 main/src/chord/analyses/reflect/ObjNewInstAnalysis.java

@@ -14,9 +14,9 @@ import joeq.Class.jq_Class;
 import joeq.Class.jq_NameAndDesc;
 import joeq.Compiler.Quad.Quad;
 
-import chord.analyses.invk.DomI;
-import chord.analyses.alloc.DomH;
-import chord.analyses.method.DomM;
+import chord.analyses.facts.alloc.DomH;
+import chord.analyses.facts.invk.DomI;
+import chord.analyses.facts.method.DomM;
 import chord.program.Program;
 import chord.project.Chord;
 import chord.project.ClassicProject;

### 2011-03-11 main/src/chord/analyses/reflect/RelClsForNameIT.java

@@ -13,8 +13,8 @@ import joeq.Class.jq_Method;
 import joeq.Compiler.Quad.Quad;
 import joeq.Compiler.Quad.Operator.Invoke;
 
-import chord.analyses.invk.DomI;
-import chord.analyses.type.DomT;
+import chord.analyses.facts.invk.DomI;
+import chord.analyses.facts.type.DomT;
 import chord.program.Program;
 import chord.project.Chord;
 import chord.project.analyses.ProgramRel;

### 2011-03-11 main/src/chord/analyses/type/DomT.java

@@ -1,44 +1,0 @@
-/*
- * Copyright (c) 2008-2010, Intel Corporation.
- * Copyright (c) 2006-2007, The Trustees of Stanford University.
- * All rights reserved.
- * Licensed under the terms of the New BSD License.
- */
-package chord.analyses.type;
-
-import joeq.Class.jq_Type;
-import joeq.Class.jq_Class;
-import chord.program.Program;
-import chord.project.Chord;
-import chord.project.analyses.ProgramDom;
-import chord.util.IndexSet;
-
-/**
- * Domain of classes.
- * 
- * @author Mayur Naik (mhn@cs.stanford.edu)
- */
-@Chord(
-	name = "T"
-)
-public class DomT extends ProgramDom<jq_Type> {
-	public void fill() {
-		Program program = Program.g();
-		IndexSet<jq_Type> types = program.getTypes();
-		for (jq_Type t : types)
-			add(t);
-	}
-	public String toXMLAttrsString(jq_Type t) {
-		String name = t.getName();
-		String file;
-		if (t instanceof jq_Class) {
-			jq_Class c = (jq_Class) t;
-			file = c.getSourceFileName();
-		} else
-			file = "";
-		int line = 0;  // TODO
-		return "name=\"" + name +
-			"\" file=\"" + file +
-			"\" line=\"" + line + "\"";
-	}
-}

### 2011-03-11 main/src/chord/analyses/type/RelClassT.java

@@ -1,29 +1,0 @@
-/*
- * Copyright (c) 2008-2010, Intel Corporation.
- * Copyright (c) 2006-2007, The Trustees of Stanford University.
- * All rights reserved.
- * Licensed under the terms of the New BSD License.
- */
-package chord.analyses.type;
-
-import joeq.Class.jq_Class;
-import chord.program.visitors.IClassVisitor;
-import chord.project.Chord;
-import chord.project.analyses.ProgramRel;
-
-/**
- * Relation containing each (concrete or abstract) class type
- * (as opposed to interface types, primitive types, etc.).
- *
- * @author Mayur Naik (mhn@cs.stanford.edu)
- */
-@Chord(
-	name = "classT",
-	sign = "T0"
-)
-public class RelClassT extends ProgramRel implements IClassVisitor {
-	public void visit(jq_Class c) {
-		if (!c.isInterface())
-			add(c);
-	}
-}

### 2011-03-11 main/src/chord/analyses/type/RelInterfaceT.java

@@ -1,29 +1,0 @@
-/*
- * Copyright (c) 2008-2010, Intel Corporation.
- * Copyright (c) 2006-2007, The Trustees of Stanford University.
- * All rights reserved.
- * Licensed under the terms of the New BSD License.
- */
-package chord.analyses.type;
-
-import joeq.Class.jq_Class;
-import chord.program.visitors.IClassVisitor;
-import chord.project.Chord;
-import chord.project.analyses.ProgramRel;
-
-/**
- * Relation containing each interface type.
- *
- * @author Mayur Naik (mhn@cs.stanford.edu)
- */
-@Chord(
-	name = "interfaceT",
-	sign = "T0"
-)
-public class RelInterfaceT extends ProgramRel
-		implements IClassVisitor {
-	public void visit(jq_Class c) {
-		if (c.isInterface())
-			add(c);
-	}
-}

### 2011-03-11 main/src/chord/analyses/type/RelScopeExcludedT.java

@@ -1,20 +1,0 @@
-package chord.analyses.type;
-
-import joeq.Class.jq_Class;
-import chord.program.visitors.IClassVisitor;
-import chord.project.Chord;
-import chord.project.Config;
-import chord.project.analyses.ProgramRel;
-import chord.util.Utils;
-
-@Chord(
-		name = "scopeExcludedT",
-		sign = "T0"
-	)
-public class RelScopeExcludedT extends ProgramRel implements IClassVisitor {
-	public void visit(jq_Class c) {
-		if (Utils.prefixMatch(c.getName(), Config.scopeExcludeAry))
-			add(c);
-	}
-
-}

### 2011-03-11 main/src/chord/analyses/var/DomV.java

@@ -1,71 +1,0 @@
-/*
- * Copyright (c) 2008-2010, Intel Corporation.
- * Copyright (c) 2006-2007, The Trustees of Stanford University.
- * All rights reserved.
- * Licensed under the terms of the New BSD License.
- */
-package chord.analyses.var;
-
-import java.util.HashMap;
-import java.util.Map;
-import java.util.List;
-
-import joeq.Class.jq_Class;
-import joeq.Class.jq_Method;
-import joeq.Compiler.Quad.RegisterFactory.Register;
-import chord.program.visitors.IMethodVisitor;
-import chord.project.Chord;
-import chord.project.analyses.ProgramDom;
-import joeq.Compiler.Quad.RegisterFactory;
-import joeq.Compiler.Quad.BasicBlock;
-import joeq.Compiler.Quad.ControlFlowGraph;
-import joeq.Compiler.Quad.Operator;
-import joeq.Compiler.Quad.Quad;
-import joeq.Compiler.Quad.Operand;
-import joeq.Compiler.Quad.Operand.RegisterOperand;
-import joeq.Compiler.Quad.Operand.ParamListOperand;
-import joeq.Class.jq_Type;
-import joeq.Util.Templates.ListIterator;
-
-/**
- * Domain of local variables of reference type.
- * <p>
- * Each local variable declared in each block of each method is
- * represented by a unique element in this domain.  Local variables
- * that have the same name but are declared in different methods
- * or in different blocks of the same method are represented by
- * different elements in this domain.
- * <p>
- * The set of local variables of a method is the disjoint union of
- * its argument variables and its temporary variables.  All local
- * variables of the same method are assigned contiguous indices in
- * this domain.  The argument variables are assigned contiguous
- * indices in order followed by the temporary variables.
- * 
- * @author Mayur Naik (mhn@cs.stanford.edu)
- */
-@Chord(
-	name = "V"
-)
-public class DomV extends ProgramDom<Register> implements IMethodVisitor {
-	private Map<Register, jq_Method> varToMethodMap;
-	public void init() {
-		varToMethodMap = new HashMap<Register, jq_Method>();
-	}
-	public jq_Method getMethod(Register v) {
-		return varToMethodMap.get(v);
-	}
-	public void visit(jq_Class c) { }
-	public void visit(jq_Method m) {
-		if (m.isAbstract())
-			return;
-		List<Register> vars = m.getLiveRefVars();
[…]

### 2011-03-11 main/src/chord/program/RTA.java

@@ -56,7 +56,7 @@ import chord.project.Config;
 import chord.program.reflect.CastBasedStaticReflect;
 import chord.program.reflect.DynamicReflectResolver;
 import chord.program.reflect.StaticReflectResolver;
-import chord.analyses.method.RelExtraEntryPoints;
+import chord.analyses.facts.method.RelExtraEntryPoints;
 import chord.util.IndexSet;
 import chord.util.Timer;
 import chord.util.ArraySet;
@@ -74,7 +74,7 @@ public class RTA {
 	private static final String MAIN_CLASS_NOT_DEFINED =
 		"ERROR: Property chord.main.class must be set to specify the main class of program to be analyzed.";
 	private static final String MAIN_METHOD_NOT_FOUND =
-		"ERROR: Could not find main class `%s` or main method in that class.";
+		"ERROR: Could not find main class '%s' or main method in that class.";
 	private static final String METHOD_NOT_FOUND_IN_SUBTYPE =
 		"WARN: Expected instance method %s in class %s implementing/extending interface/class %s.";
 

### 2011-03-11 main/src/chord/program/reflect/CastBasedStaticReflect.java

@@ -10,7 +10,7 @@ import java.util.*;
 
 import chord.program.ClassHierarchy;
 import chord.program.Program;
-import chord.analyses.method.RelScopeExcludedM;
+import chord.analyses.facts.method.RelScopeExcludedM;
 import chord.util.ArraySet;
 import chord.util.IndexSet;
 import chord.util.tuple.object.Pair;

### 2011-03-11 main/src/chord/project/Boot.java

@@ -70,17 +70,17 @@ import chord.util.ProcessExecutor;
  */
 public class Boot {
 	private static final String CHORD_JAR_NOT_FOUND =
-		"ERROR: Expected Chord to be loaded from chord.jar instead of from `%s`.";
+		"ERROR: Boot: Expected Chord to be loaded from chord.jar instead of from '%s'.";
     private static final String USER_DIR_AS_CHORD_WORK_DIR =
-        "WARN: Property chord.work.dir not set; using value of user.dir `%s` instead.";
+        "WARN: Boot: Property chord.work.dir not set; using value of user.dir '%s' instead.";
     private static final String CHORD_MAIN_DIR_UNDEFINED =
-        "ERROR: Property chord.main.dir must be set to location of directory named 'main' in your Chord installation.";
+        "ERROR: Boot: Property chord.main.dir must be set to location of directory named 'main' in your Chord installation.";
     private static final String CHORD_MAIN_DIR_NOT_FOUND =
-        "ERROR: Directory `%s` specified by property chord.main.dir not found.";
+        "ERROR: Boot: Directory '%s' specified by property chord.main.dir not found.";
     private static final String CHORD_WORK_DIR_UNDEFINED =
-        "ERROR: Property chord.work.dir must be set to location of working directory desired during Chord's execution.";
+        "ERROR: Boot: Property chord.work.dir must be set to location of working directory desired during Chord's execution.";
     private static final String CHORD_WORK_DIR_NOT_FOUND =
-        "ERROR: Directory `%s` specified by property chord.work.dir not found.";
+        "ERROR: Boot: Directory '%s' specified by property chord.work.dir not found.";
 
 	public static void main(String[] args) throws Throwable {
 		String chordJarFile = getChordJarFile();

### 2011-03-11 main/src/chord/project/analyses/DynamicAnalysis.java

@@ -21,8 +21,8 @@ import java.util.Collections;
 import joeq.Class.jq_Method;
 import joeq.Compiler.Quad.BasicBlock;
 import joeq.Compiler.Quad.ControlFlowGraph;
-import chord.analyses.basicblock.DomB;
-import chord.analyses.method.DomM;
+import chord.analyses.facts.basicblock.DomB;
+import chord.analyses.facts.method.DomM;
 import chord.instr.EventKind;
 import chord.instr.InstrScheme;
 import chord.instr.Instrumentor;

### 2011-03-12 main/doc/acknowledgments.tex
more documentation

@@
 \item
-\href{http://joeq.sourceforge.net/}{Joeq}, a Java compiler framework
+\xlink{Joeq}{http://joeq.sourceforge.net/}, a Java compiler framework
 \item
-\href{http://www.csg.is.titech.ac.jp/~chiba/javassist/}{Javassist},
+\xlink{Javassist}{http://www.csg.is.titech.ac.jp/~chiba/javassist/},
 a Java bytecode manipulation tool
 \item
-\href{ttp://bddbddb.sourceforge.net/}{bddbddb}, a BDD-based Datalog solver
+\xlink{bddbddb}{http://bddbddb.sourceforge.net/}, a BDD-based Datalog solver
 \end{itemize}
@@
 \item
-\href{http://ant-contrib.sourceforge.net/}{Ant-Contrib},
+\xlink{Ant-Contrib}{http://ant-contrib.sourceforge.net/},
 a collection of useful Ant tasks  
 \item
-\href{http://buddy.sourceforge.net/}{BuDDy}, a BDD library
+\xlink{BuDDy}{http://buddy.sourceforge.net/}, a BDD library
 \item
-\href{http://trove4j.sourceforge.net/}{GNU Trove},
+\xlink{GNU Trove}{http://trove4j.sourceforge.net/},
 a primitive collections library for Java
 \item
-\href{http://www.java2html.com/}{Java2HTML} and
-\href{http://www.java2html.de/}{Java2Html}, Java to HTML tools
+\xlink{Java2HTML}{http://www.java2html.com/} and
+\xlink{Java2Html}{http://www.java2html.de/}, Java-to-HTML tools
 \item
-\href{http://saxon.sourceforge.net/}{Saxon}, an XSLT processor
+\xlink{Saxon}{http://saxon.sourceforge.net/}, an XSLT processor
 \end{itemize}

### 2011-03-12 main/doc/building_chord.tex
more documentation

\section{Building Chord}
\label{sec:building-chord}

Compiling Chord's source installation requires the following software:

\begin{itemize}
\item
A JVM with JDK 5 or higher, e.g. \xlink{IBM J9}{http://www.ibm.com/developerworks/java/jdk/} or
\xlink{Oracle HotSpot}{http://www.oracle.com/technetwork/java/javase/downloads/index.html}.
\item
\xlink{Apache Ant}{http://ant.apache.org/}, a Java build tool.
\end{itemize}

Directory {\tt main/} contains a file named {\tt build.xml} which is
interpreted by Apache Ant.  To see the various targets available, run
the following command in that directory:

\begin{verbatim}
    prompt> ant
\end{verbatim}

To compile Chord, run the following command in the same directory:

\begin{verbatim}
    prompt> ant compile
\end{verbatim}

This will compile Chord's Java sources from \code{main/src/} to class files
in \code{main/classes/}, as well as build a jar file \code{main/chord.jar} containing
these class files as well as the class files in the jars of libraries that are used by Chord
and are provided under \code{main/lib/} (e.g., \code{joeq.jar}, \code{javassist.jar},
\code{bddbddb.jar}, etc.).
Additionally:
\begin{itemize}
\item
If system property \code{chord.use.buddy} is set to \code{true} (default is \code{false}),
then the C source code of the \xlink{BuDDy}{http://buddy.sourceforge.net/} BDD library
from \code{main/bdd/} will be compiled to a shared library in \code{main/}
(\code{libbuddy.so} on Linux, \code{buddy.dll} on Windows, and
\code{libbuddy.dylib} on MacOS); this library is used for executing
program analyses written in Datalog
using \xlink{bddbddb}{http://bddbddb.sourceforge.net/} (a BDD-based Datalog solver).
\item
If system property \code{chord.use.jvmti} is set to \code{true} (default is \code{false}),
then the C++ source code of the JVMTI-based bytecode instrumentation agent from
\code{main/agent/} will be compiled to a shared library in \code{main/}
(\code{libchord_instr_agent.so} on all architectures); this agent is
used for computing program scope dynamically and for executing
dynamic program analyses.
\end{itemize}
The above system properties can be defined in file \code{main/chord.properties}
or on the command-line of ant (using the \code{-D<key>=<value>} format).  If either of these
properties is set to \code{true}, then you will also need a utility like GNU Make (to run the
\code{Makefile}'s under \code{main/bdd/} and \code{main/agent/}) and a C++ compiler.

### 2011-03-12 main/doc/computing_scope.tex
more documentation

--- 2011-03-04 main/doc/building_scope.tex
@@
-\section{Computing program scope}
-\label{sec:building-scope}
+\section{Computing Analysis Scope}
+\label{sec:computing-scope}
 
-Chord computes the scope (i.e., reachable methods) of the given
+Chord computes the analysis scope (i.e., reachable methods) of the given
 program either if property \code{chord.build.scope} is set to {\tt
@@
 \code{chord.program.Program.g()}.
-
-Users can set the following properties to control scope computation:
-
-\begin{itemize}
-\item \code{chord.main.class}
-\item \code{chord.class.path}
-\item \code{chord.reuse.scope}
-\item \code{chord.methods.file}
-\item \code{chord.reflect.file}
-\item \code{chord.scope.kind}
-\item \code{chord.ch.kind}
-\item \code{chord.reflect.kind}
-\item \code{chord.std.scope.exclude}
-\item \code{chord.ext.scope.exclude}
-\item \code{chord.scope.exclude}
-\end{itemize}
+The following sections describe Chord's analysis scope 
+computation in more detail.
+%Users can set the following properties to control scope computation:
+%
+%\begin{itemize}
+%\item \code{chord.main.class}
+%\item \code{chord.class.path}
+%\item \code{chord.reuse.scope}
+%\item \code{chord.methods.file}
+%\item \code{chord.reflect.file}
+%\item \code{chord.scope.kind}
+%\item \code{chord.ch.kind}
+%\item \code{chord.reflect.kind}
+%\item \code{chord.std.scope.exclude}
+%\item \code{chord.ext.scope.exclude}
+%\item \code{chord.scope.exclude}
+%\end{itemize}
+%
+%The meaning of these properties is explained on demand in
 
@@
 least expects properties \code{chord.main.class} and
-\code{chord.class.path} to be set.
+\code{chord.class.path} to be set to the fully-qualified name of the
+program's main class (e.g., \code{com.example.Main}) and the
+program's application classpath, respectively.
 
@@
 \item
-The {\tt dynamic} value instructs Chord to compute program scope
-dynamically, by running the program and observing using JVMTI the
+The {\tt dynamic} value instructs Chord to compute analysis scope
+dynamically, by running the program and observing the
 classes that are loaded at run-time.  The number of times the program
@@
 as reachable regardless of whether they were invoked in the run.  The
-rationale behind this decision is to both reduce run-time overhead of
-JVMTI and to increase the predictive power of program analyses
-performed using the computed program scope.
+rationale behind this decision is to both reduce the run-time instrumentation
+overhead and increase the predictive power of program analyses
+performed using the computed analysis scope.
 
 \item
-The {\tt rta} value instructs Chord to compute program scope
+The {\tt rta} value instructs Chord to compute analysis scope
 statically using Rapid Type Analysis (RTA).
@@
 %problem.  \item The \code{rta_reflect} value instructs Chord to
-%compute program scope statically using Rapid Type Analysis and,
+%compute analysis scope statically using Rapid Type Analysis and,
 %moreover, to resolve a common reflection
@@
 \item
-The {\tt cha} value instructs Chord to compute program scope
+The {\tt cha} value instructs Chord to compute analysis scope
 statically using Class Hierarchy Analysis (CHA).
@@
 Nevertheless, Chord allows users to control which classes are
-included in the class hierarchy (see \ref{sec:cha}) and thereby control the
+included in the class hierarchy (see Section \ref{sec:cha}) and thereby control the
 precision and cost of CHA.
@@
 \item
-Chord's classpath, specified by property
-\code{chord.main.class.path}.  This includes the classpath of any
-user-defined Java analyses, specified by user-defined property
-\code{chord.java.analysis.path}, which is empty by default.
-\item
-The classpath of the given program, specified by user-defined property
-\code{chord.class.path}, which is empty by default.
+The application classpath of the given program, specified by property \code{chord.class.path},
+which is empty by default.
 \end{enumerate}
 
-All classes in the entire classpath (resulting from items 1--4 above)
+All classes in the entire classpath (resulting from items 1--3 above)
 are included in the class hierarchy with the following exceptions:
@@
 ``missing" means that it is either not in the classpath resulting from
-items 1--4 above or it is excluded by one of these rules, then C
+items 1--3 above or it is excluded by one of these rules, then C
 itself is excluded.  Note that this rule is recursive, e.g., if C has
@@
 analyzed, e.g., classes in the JDK standard library.  Its default
-value is \code{[chord.main.class.path]}, i.e., the classes comprising
-Chord itself.
+value is the empty list.
 \item
@@
 Property \code{chord.scope.exclude} specifies the final list of
-classes to be excluded from scope.  Its default value is simply the
-concatenation of the values of properties
-\code{chord.std.scope.exclude} and \code{chord.ext.scope.exclude}.
+classes to be excluded from scope.  Its default value is
+\code{[chord.std.scope.exclude]:[chord.ext.scope.exclude]}.
 \end{itemize}

### 2011-03-12 main/doc/deadlock.tex
more documentation

@@
 
-\noindent where {\tt <...>} denotes the absolute or relative path of a directory
+where {\tt <...>} denotes the absolute or relative path of a directory
 containing a file named \code{chord.properties} which defines

### 2011-03-12 main/doc/running_chord.tex
more documentation

\section{Running Chord}
\label{sec:running-chord}

Running Chord requires a JVM with JDK 5 or higher.
There are two equivalent ways to run Chord.
One way is to run the following command in the \code{main/} directory:

\begin{verbatim}
    prompt> ant -D<key1>=<value1> ... -D<keyN>=<valueN> run
\end{verbatim}

The above requires Apache Ant to be installed.  The alternative, which
does not require Apache Ant, is to invoke Chord directly as follows:

\begin{verbatim}
    prompt> java -cp <...>/chord.jar -D<key1>=<value1> ... -D<keyN>=<valueN> chord.project.Boot
\end{verbatim}

where \code{<...>} denotes the absolute or relative path of the directory containing file \code{chord.jar};
that directory is also expected to contain any other binaries in Chord's installation
(e.g., \code{libbuddy.so} and \code{libchord_instr_agent.so}).

### 2011-03-12 main/doc/setting_up.tex
more documentation

@@
 
-\noindent The above structure is typical: the program's Java source
+The above structure is typical: the program's Java source
 files are under {\tt src/}, its class files are under {\tt classes/},
@@
 
-\noindent where \code{<...>} denotes the absolute or relative path of the
+where \code{<...>} denotes the absolute or relative path of the
 directory containing the above \code{example/} directory.
@@
 line or in the \code{chord.properties} file, specifying the task(s)
-Chord must perform.  All tasks are summarized in Section
-\ref{sec:func-props}.  The two most common tasks, described
+Chord must perform.  All tasks are summarized under ``Functionality Properties"
+in Section \ref{sec:chord-sysprops}.  The two most common tasks, described
 next, are computing the analysis scope of the given program (Section
-\ref{sec:building-scope}) and running analyses on the given program
+\ref{sec:computing-scope}) and running analyses on the given program
 (Section \ref{sec:running-analysis}).

### 2011-03-12 main/doc/troubleshooting.tex
more documentation

@@
 
-\noindent where {\tt <sys-props>} denotes a whitespace-separated list of zero of more property definitions
+where {\tt <sys-props>} denotes a whitespace-separated list of zero of more property definitions
 of the form {\tt -D<key>=<value>} (see Section \ref{sec:chord-sysprops} for an exhaustive listing of all

### 2011-03-12 main/doc/writing_dynamic_analysis.tex
more documentation

@@
 You can change the default values of various properties for configuring your dynamic analysis;
-see Sections \ref{sec:scope-props} and \ref{sec:instr-props}.
+see ``Analysis Scope Properties" and ``Instrumentation Properties" in Section \ref{sec:chord-sysprops}.
 For instance:
@@
 Compile the analysis by placing the directory containing class \code{MyDynamicAnalysis} created
-above in the path defined by property \code{chord.java.analysis.path}.
+above in the path defined by property \code{chord.ext.java.analysis.path}.
 
@@
 \code{MyDynamicAnalysis} created above).
+
+{\bf Note:} The IBM J9 JVM on Linux is highly recommended if you intend to use Chord for
+dynamic program analysis, as it allows you to instrument the entire JDK; using any other
+platform will likely require excluding large parts of the JDK from being instrumented.
+Additionally, if you intend to use online (load-time) bytecode instrumentation in
+your dynamic program analysis, then you will need JDK 6 or higher, since this functionality
+requires the \code{java.lang.instrument} API with class retransformation support (the latter
+support is available only in JDK 6 and higher).

### 2011-03-14 main/doc/analyses.tex
new documentation

--- 2011-03-10 main/doc/writing_analysis.tex
@@
-\section{Writing an Analysis}
-\label{sec:writing-analysis}
+\chapter{Writing and Running Analyses}
+\label{chap:analyses}
 
@@
 
-\subsection{Dependency information}
+\section{Dependency information}
 
@@
 
-\subsection{Domains}
+\section{Domains}
 
@@
 
-\subsection{Core relations}
+\section{Core relations}
 
@@
 
-\subsection{Java analyses}
+\section{Java analyses}
 
@@
 
-\noindent  The default value of property \code{chord.java.analysis.path}
+The default value of property \code{chord.java.analysis.path}
 is \code{chord.std.java.analysis.path}:\code{chord.ext.java.analysis.path},
@@
 
-\noindent  The default value of property \code{chord.dlog.analysis.path}
+The default value of property \code{chord.dlog.analysis.path}
 is \code{chord.std.dlog.analysis.path}:\code{chord.ext.java.analysis.path},
@@
 set the latter property).
+
+\section{Running an Analysis}
+\label{sec:running-analysis}
+
+Running a Chord analysis is straightforward.
+The property  \texttt{chord.run.analyses} is a list of analyses to be run.
+Chord will compute the listed dependencies for each of these relations and run these prerequisites before running the analysis you specify.
+
+Chord has two kinds of projects: ``Classic'' and ``Modern''.  Despite the names, classic is effectively the standard at this point.
+In the classic model, there is an implicit dependency graph between analyses.
+For a datalog analysis (program), the ``input'' and ``output'' keywords in defining a relation specify the requirements and outputs of an analysis.
+For a Java analysis, the \texttt{@Chord} annotation does so.
+See the section ``writing a Java analysis'' for details.
+
+\texttt{chord.print.rels} is a list of relations to dump to disk.
+Note that these relations are printed in a somewhat unfriendly format; this is more useful for debugging than for interpreting programs.
+Note also that listing a relation here does not cause it to be computed; you will get an error at the end of your Chord execution if you ask it to print a relation that has not been computed.
+
+%List of names of program analyses to be run in order.
+%chord.java.analysis.path    path    Classpath containing program analyses in Java (i.e. @Chord-annotated classes). ${chord.main.dir}/classes/
+%chord.dlog.analysis.path    path    Path of directories containing program analyses in Datalog (i.e. *.datalog and *.dlog files).  ${chord.main.dir}/src/main/dlog/
+%chord.reuse.rels    bool    Construct program relations from BDDs stored on disk (from a previous run of Chord) whenever possible instead of re-computing them.    false
+%chord.print.results     bool    Print the results of program analyses in HTML. true

### 2011-03-14 main/doc/computing_scope.tex
new documentation

@@
-\section{Computing Analysis Scope}
-\label{sec:computing-scope}
+\chapter{Computing Analysis Scope}
+\label{chap:computing-scope}
 
-Chord computes the analysis scope (i.e., reachable methods) of the given
-program either if property \code{chord.build.scope} is set to {\tt
-  true} or if some other task (e.g., a program analysis specified via
-property \code{chord.run.analyses}) demands it by calling method
-\code{chord.program.Program.g()}.
+A pre-requisite to analyzing a Java program using any program analysis framework,
+including Chord, is to compute the {\it analysis scope}: the set of methods
+of the program to be analyzed.  Several scope construction algorithms (so-called
+call-graph algorithms) exist in the literature that differ in scalability (i.e., how large
+a program they can handle with the available resources) and precision (i.e., how small a
+set of methods they deem is reachable).
+
+Chord implements several standard scope construction algorithms.  Besides scalability and precision,
+an additional metric of these algorithms in Chord that can be controlled by users is usability,
+which concerns aspects such as excluding certain methods from being analyzed even if they
+are reachable, and modeling Java features such as reflection,
+dynamic class loading, and native methods.
+These features affect which methods are
+reachable but, in general, they cannot be modeled soundly by any program analysis framework.  The best
+a framework can do is
+ provide stubs for commonly-used native methods in the standard JDK library (e.g., the \code{arraycopy}
+method of class \code{java.lang.System}), 
+offer users a range of options on how to resolve reflection (e.g.,
+an option might be running the program and observing how reflection is resolved), etc.
+
+Chord computes the analysis scope of the given
+program either if property \code{chord.build.scope} is set to {\tt true} or if some other task
+(e.g., a program analysis specified via property \code{chord.run.analyses}) demands it.
 The following sections describe Chord's analysis scope 
-computation in more detail.
+computation in more detail.  Section \ref{sec:scope-reuse} describes how to reuse the analysis scope
+computed in a previous run of Chord for a given program.  Section \ref{sec:scope-algos} describes
+Chord's analysis scope construction algorithms.  Finally, Section \ref{sec:scope-exclude} describes how users
+can exclude certain classes from the analysis scope.
+
 %Users can set the following properties to control scope computation:
@@
 
-\subsection{Scope reuse}
+\section{Scope Reuse}
+\label{sec:scope-reuse}
 
@@
 \code{[chord.out.dir]/reflect.txt}, respectively.
-Property \code{chord.out.dir} defaults to
-\code{[chord.work.dir]/chord_output/} and property
-\code{chord.work.dir} defaults to the working directory during Chord's
-execution.
+Property \code{chord.out.dir} denotes the output directory of Chord;
+its default value is \code{[chord.work.dir]/chord_output/}.
+Property \code{chord.work.dir} denotes the working directory during
+Chord's execution; its default value is the current directory.
 
-\subsection{Scope construction algorithms}
+\section{Scope Construction Algorithms}
+\label{sec:scope-algos}
 
@@
 files specified by properties \code{chord.methods.file} or
-\code{chord.reflect.file} do not exist, then Chord computes program
+\code{chord.reflect.file} do not exist, then Chord computes analysis
 scope using the algorithm specified by property
@@
 
-The possible legal values of property \code{chord.scope.kind} are
-[\code{dynamic}$|$\code{rta}$|$\code{cha}].  In each case, Chord at
+The possible values of property \code{chord.scope.kind} are
+[\code{dynamic}$|$\code{rta}$|$\code{cha}] (the default value is {\tt rta}).
+The following subsections describe the scope construction algorithm
+that Chord runs in each of these three cases.
+In each case, Chord at
 least expects properties \code{chord.main.class} and
@@
 
-\begin{itemize}
-\item
-The {\tt dynamic} value instructs Chord to compute analysis scope
+\subsection{Dynamic Scope Construction}
+
+If property \code{chord.scope.kind} has value {\tt dynamic}, then Chord
+computes analysis scope
 dynamically, by running the program and observing the
@@
 
-\item
-The {\tt rta} value instructs Chord to compute analysis scope
-statically using Rapid Type Analysis (RTA).
+\subsection{RTA Scope Construction}
+
+If property \code{chord.scope.kind} has value {\tt rta}, then Chord
+computes analysis scope statically using Rapid Type Analysis (RTA).
 RTA is an iterative fixed-point algorithm.  It maintains a set of
@@
 %\ref{sec:cha}).
-\item
-The {\tt cha} value instructs Chord to compute analysis scope
-statically using Class Hierarchy Analysis (CHA).
+
+\subsection{CHA Scope Construction}
+
+If property \code{chord.scope.kind} has value {\tt cha}, then Chord
+computes analysis scope statically using Class Hierarchy Analysis (CHA).
 The key difference between CHA and RTA is that for invokevirtual and
@@
 Nevertheless, Chord allows users to control which classes are
-included in the class hierarchy (see Section \ref{sec:cha}) and thereby control the
-precision and cost of CHA.
+included in the class hierarchy, and thereby control the
+precision and cost of CHA, by setting property \code{chord.ch.kind},
+whose possible values are [\code{static}$|$\code{dynamic}] (the default
+value is {\tt static}).
 
-\end{itemize}
-The default value of property \code{chord.scope.kind} is {\tt rta}.
-
-\subsection{Class hierarchy construction}
-\label{sec:cha}
-
-The class hierarchy is built if property \code{chord.scope.kind} has
-value {\tt cha}.  Users can control which classes are included in
-building the class hierarchy by setting property \code{chord.ch.kind},
-whose possible legal values are [\code{static}$|$\code{dynamic}].
 Chord first constructs the entire classpath of the given program by
@@
 Any class whose name's prefix is specified in the value of property
-\code{chord.scope.exclude}.
+\code{chord.scope.exclude} (see Section \ref{sec:scope-exclude}).
 \item
@@
 Chord runs the given program and observes the set of all classes the
-JVM loads; any class not in this set is excluded.  By default,
-property \code{chord.ch.kind} has value {\tt static}.
+JVM loads; any class not in this set is excluded.
 \item
@@
 
-\subsection{Scope exclusion}
+\section{Scope Exclusion}
+\label{sec:scope-exclude}
 
-For tractability, Chord can be instructed to treat the bodies of certain
-methods as no-ops.  Chord provides three properties for this purpose.
+Chord can be instructed to exclude certain classes in a given program 
+from being analyzed.
+This functionality might be desirable, for instance, if the given program
+contains a larger framework (e.g., Hadoop or Android) which
+must not be analyzed.
+Chord provides three properties for this purpose.
 The value of each of these properties is a comma-separated list of prefixes
-of names of classes all of whose declared methods must be treated as no-ops:
+of names of classes.  Chord treats the body of each method defined in
+each such class as a no-op.
 
@@
 \item
-Property \code{chord.std.scope.exclude} conventionally specifies
+Property \code{chord.std.scope.exclude} is intended to specify
 classes to be excluded from the scope of {\it all} programs to be
@@
 \item
-Property \code{chord.ext.scope.exclude} conventionally specifies
+Property \code{chord.ext.scope.exclude} is intended to specify
 classes to be excluded from the scope of specific programs to be
@@
 \end{itemize}
+
+{\bf Note:} The value of each of the above properties is a list of {\it prefixes},
+not {\it regular expressions}.  A valid value is \code{java.,com.sun.},
+but not \code{java.*,com.sun.*}.
 

### 2011-03-14 main/doc/datalog_analysis.tex
new documentation

\chapter{Datalog Analysis}
\label{chap:datalog-analysis}

A common way to rapidly prototype program analyses in Chord is using a declarative
logic-programming language called Datalog.  This chapter describes everything about
Datalog analysis in Chord.
Section \ref{sec:writing-datalog-analysis} explains how to write a Datalog analysis,
Section \ref{sec:running-datalog-analysis} explains how to run it, and
Section \ref{sec:tuning-datalog-analysis} explains how to tune its performance.
Finally, Section \ref{sec:bdd} explains the representation of BDDs (Binary Decision Diagrams)
which are a data structure central to executing Datalog analyses.

\section{Writing a Datalog Analysis}
\label{sec:writing-datalog-analysis}

A Datalog analysis declares a bunch of input/output program relations,
each over one or more program domains, and provides a bunch of rules (constraints)
specifying how to compute the output relations from the input relations.
An example of such an analysis is as follows:

\begin{verbatim}
    # name=datarace-dlog

    # Program domains
    .include "E.dom"
    .include "F.dom"
    .include "T.dom"

    # BDD variable order
    .bddvarorder E0xE1_T0_T1_F0

    # Input/intermediate/output program relations
    field(e:E0,f:F0) input
    write(e:E0) input
    reach(t:T0,e:E0) input
    alias(e1:E0,e2:E1) input
    escape(e:E0) input
    unguarded(t1:T0,e1:E0,t2:T1,e2:E1) input
    hasWrite(e1:E0,e2:E1)
    candidate(e1:E0,e2:E1) 
    datarace(t1:T0,e1:E0, t2:T1,e2:E1) output

    # Analysis constraints
    hasWrite(e1,e2) :- write(e1).
    hasWrite(e1,e2) :- write(e2).
    candidate(e1,e2) :- field(e1,f), field(e2,f), hasWrite(e1,e2), e1 <= e2.
    datarace(t1,e1,t2,e2) :- candidate(e1,e2), reach(t1,e1), reach(t2,e2), \
        alias(e1,e2), escape(e1), escape(e2), unguarded(t1,e1,t2,e2).
\end{verbatim}

Any line that begins with a {\tt \#} is regarded a comment, except a
line of the form ``{\tt \#} {\tt name=<...>}", which specifies the name
{\tt <...>} of the Datalog analysis.
Each Datalog analysis is expected to have exactly one such line.
The above Datalog analysis is named {\tt datarace-dlog}.
In Chord, all Datalog analysis names have suffix
{\tt -dlog} and all Java analysis names have suffix {\tt -java}, but
this is merely a convention that analysis writers are free to deviate from.
The name of each analysis, written in Datalog or in Java, is expected to be
unique across all analyses in scope (i.e., across all analyses in paths
specified by properties \code{chord.dlog.analysis.path} and \code{chord.java.analysis.path}).

The {\tt .include "<...>.dom"} lines specify each program domain named
{\tt <...>} that is needed by the Datalog analysis, i.e., each domain over which
any program relation that is input/output by the Datalog analysis is defined. 
The declaration of each such relation specifies the domain of each of the relation's attributes.
If the same domain appears in multiple attributes of a relation then
contiguous integers starting from 0 must be used to distinguish them; for instance,
in the above example, {\tt candidate} is a binary relation, both of whose
attributes have domain E, and they are distinguished as E0 and E1.

Each relation is represented symbolically (as opposed to explicitly)
using a graph-based data structure called a Binary Decision Diagram (BDD for short).
Each domain containing N elements is assigned log2(N) BDD variables.
The size of a BDD and the efficiency of operations on it depends heavily
on the order of these BDD variables.
The {\tt .bddvarorder <...>} line in the Datalog analysis enables the Datalog
analysis writer to specify this order.
It must list all domains along with their numerical suffixes, separated
by {\tt \_} or {\tt x}.
Using a {\tt \_} between two domains, such as {\tt T0\_T1}, means that the BDD variables assigned
to domain {\tt T0} precede those assigned to domain {\tt T1} in the BDD variable
order for this Datalog analysis.
Using a {\tt x} between two domains, such as {\tt E0xE1}, means that the
BDD variables assigned to domains {\tt E0} and {\tt E1}
will be interleaved in the BDD variable order for this Datalog analysis.
See Section \ref{chap:bdd-representation} for more details on BDD representations.

Each Datalog analysis rule is a Horn clause of the form
{\tt R(t) :- R1(t1), ..., Rn(tn)}
meaning that if relations {\tt R1}, ..., {\tt Rn} contain tuples {\tt t1}, ..., {\tt tn}
respectively, then relation {\tt R} contains tuple {\tt t}.
A backslash may be used at the end of a line to break long rules for readability.
The Datalog analysis solver bddbddb used in Chord does not apply any
sophisticated optimizations to simplify the rules; besides the BDD variable order,
the manner in which these rules are expressed heavily affects the performance of
the solver.  For instance, an important manual optimization involves breaking down
long rules into multiple shorter rules communicating via intermediate relations.
See Section \ref{sec:tuning-datalog-analysis} for hints on tuning the performance
of Datalog analyses.

\section{Running a Datalog Analysis}
\label{sec:running-datalog-analysis}

Chord interprets each file with extension {\tt .dlog} or {\tt .datalog} in the path specified by
property \code{chord.dlog.analysis.path} as a program analysis expressed in Datalog.

TODO

\section{Tuning a Datalog Analysis}
\label{sec:tuning-datalog-analysis}

There are several tricks analysis writers can try to improve the
performance of bddbddb, the Datalog solver used by Chord, often
by several orders of magnitude.
Try these tricks by running the following command:

\begin{verbatim}
    prompt> ant -Ddlog.file=<file> -Dwork.dir=<dir> solve
\end{verbatim}
where {\tt <file>} denotes the file defining the Datalog analysis
to be tuned, and {\tt <dir>} is the directory containing the
program domains ({\tt *.dom} files) and program relations ({\tt *.bdd} files)
consumed by the analysis (this is by default the
\code{chord_output/bddbddb/} directory generated
by a previous run of Chord.

\begin{enumerate}
\item
Set properties \verb+noisy=yes+, \verb+tracesolve=yes+, and \verb+fulltracesolve=yes+
on the above command line and observe which rule gets ``stuck" (i.e., takes several seconds to solve).
\verb+fulltracesolve+ is seldom useful, but \verb+noisy+ and \verb+tracesolve+ are
often very useful.  Once you identify the rule that is getting stuck, it
will also tell you which relations and which domains used in that rule,
and which operation on them, is taking a long time to solve.  Then try
to fix the problem with that rule by doing either or both of the following:
\begin{itemize}
\item
Break down the rule into multiple rules by creating intermediate relations (the more
relations you have on the RHS of a rule the slower it generally takes to solve
that rule).
\item
Change the relative order of the domains of those
relations in the BDD variable order
(note that you can use either `\_' or `x' between a pair of domains).
\end{itemize}

\item
Once you have ensured that none of the rules is getting ``stuck",
you will notice that some rules are applied too many times, and so
although each application of the rule itself isn't taking too much
time, the cumulative time for the rule is too much.  After finishing
solving a Datalog analysis, bddbddb prints how long each rule took to
solve (both in terms of the number of times it was applied and the
cumulative time it took).  It sorts the rules in the order of the
cumulative time.  You need to focus on the rules that took the most
time to solve (they will be at the bottom of the list).  Assuming you
removed the problem of rules getting ``stuck", the rules will roughly
be in the order of the number of times they were applied.  Here is an
example:

\begin{verbatim}
OUT> Rule VH(u:V0,h:H0) :- VV(u:V0,v:V1), VH(v:V1,h:H0), VHfilter
(u:V0,h:H0).
OUT>    Updates: 2871
OUT>    Time: 6798 ms
OUT>    Longest Iteration: 0 (0 ms)
OUT> Rule IM(i:I0,m:M0) :- reachableI(i:I0), specIMV(i:I0,m:M0,v:V0), VH(v:V0,_:H0).
OUT>    Updates: 5031
OUT>    Time: 6972 ms
OUT>    Longest Iteration: 0 (0 ms)
\end{verbatim}

Notice that the second rule was applied 5031 times whereas the first
was applied 2871 times.  More importantly, the second rule took 6972
milliseconds in all, compared to 6798 for the first rule.  Hence, you
should focus on the second rule first, and try to speed it up.  This
means that you should focus only on relations IM, reachableI, specIMV,
and VH, and the domains I0, M0, V0, and H0.  Any changes you make that
do not affect these relations and domains are unlikely to make your
solving faster.  In general, look at the last few rules, not just the
last one, and try to identify the ``sub-analysis" of the Datalog analysis
that seems problematic, and then focus on speeding up just that sub-
analysis.

\item
You can add the \verb+.split+ keyword at the end of certain rules as a
hint to bddbddb to decompose those rules into simpler ones that can be
solved faster.  You can also set property \verb+split_all_rules=yes+ as shorthand
for splitting all rules without adding the \verb+.split+ keyword to any of
them, though I seldom find splitting all rules helpful.

\item
You can try to decompose a single Datalog analysis file into two separate Datalog analysis
files.  Of course, you cannot separate mutually-recursive rules into two
different analyses, but if you unnecessarily club
together rules that could have gone into different analyses, then they
can put conflicting demands on bddbddb (e.g., on the BDD variable order).
So if rule 2 uses the result of rule 1 and rule 1 does not use the result of
rule 2, then put rule 1 and rule 2 in separate Datalog analyses.

\item
Observe the sizes of the BDDs representing the relations that are
input and output.  bddbddb prints both the number of tuples in each
relation and the number of nodes in the BDD.  Try changing the
BDD variable order for the domains of the relation, and observe how the
number of nodes in the BDD for that relation change.  You will
notice that some orders perform remarkably better than others.  Then
note down these orders as invariants that you will not violate as
you tweak other things.

\item
The relative order of values *within* domains (e.g.,
in domains named \verb+M+, \verb+H+, \verb+C+, etc. in Chord) affects the
performance of bddbddb, but
I've never tried changing this and studying its effect.  It might be
worth trying.  For instance, John Whaley's PLDI'04 paper describes a
specific way in which he numbers contexts (in domain \verb+C+) and that it was
fundamental to the speedup of his ``infinity"-CFA points-to analysis.

\item
Finally, it is worth emphasizing that BDDs are not magic.
If your algorithm itself is fundamentally hard to scale, then BDDs are
unlikely to help you a whole lot.  Secondly, many things are awkward to
encode as integers (e.g., the abstract contexts in domain \verb+C+ 
in Chord) or as Datalog rules.
For instance, I've noticed that summary-based context-sensitive program
analyses are hard to express in Datalog.  The may-happen-in-parallel
analysis provided in Chord shows a relatively simple kind of summary-based
analysis that uses the Reps-Horwitz-Sagiv tabulation algorithm.  But this
is as far as I could get---more complicated summary-based algorithms are
best written in Java itself instead of Datalog.
\end{enumerate}

\section{BDD Representation}
\label{sec:bdd}

Each domain containing N elements is assigned log2(N) BDD variables in the underlying BDD factory with contiguous IDs.
For instance,
domain {\tt F0} containing [128..256) elements will be assigned 8 variables with IDs (say) 63,64,65,66,67,68,69,70 and
domain {\tt Z0} containing [8..16) elements will be assigned 4 variables with IDs (say) 105,106,107,108.

If two domains are uninterleaved in the declared domain order in a Datalog program (i.e., {\tt \_} is used instead of {\tt x} between them),
then the BDD variables assigned to those domains are ordered in reverse order in the underlying BDD factory.
For instance, the BDD variable order corresponding to the declared domain order {\tt F0\_Z0} is (in level2var format)
``70,69,68,67,66,65,64,63,108,107,106,105".

If two domains are interleaved in the declared domain order in a Datalog program (i.e., {\tt x} is used instead of {\tt \_} between them),
then the BDD variables assigned to those domains are still ordered in reverse order of IDs in the underlying BDD factory,
but they are also interleaved.
For instance, the BDD variable order corresponding to the declared domain order {\tt F0xZ0} is (in level2var format)
``70,108,69,107,68,106,67,105,66,65,64,63".

Each BDD variable is at a unique ``level" which is its 0-based position in the BDD variable order in the underlying BDD factory.

We will next illustrate the format of a BDD stored on disk (in a .bdd file) using the following example:

\begin{verbatim}
# V0:16 H0:12
# 14 15 16 17 18 19 20 21 22 23 24 25 26 27 28 29
# 82 83 84 85 86 87 88 89 90 91 92 93
28489 134
39 36 33 30 27 24 21 18 15 12 9 6 3 0 81 79 77 75 73 71 69 67 65 63 61 59 57 55  \
   53 51 82 80 78 76 74 72 70 68 66 64 62 60 58 56 54 52 37 34 31 28 25 22 19 16 \
   13 10 7 4 1 117 116 115 114 113 112 111 110 109 108 107 106 50 49 48 47 46 45 \
   44 43 42 41 40 105 104 103 102 101 100 99 98 97 96 95 94 93 92 91 90 89 88 87 \
   86 85 84 83 133 132 131 130 129 128 127 126 125 124 123 122 121 120 119 118   \
   38 35 32 29 26 23 20 17 14 11 8 5 2
287 83 0 1
349123 84 287 0
349138 85 0 349123
...
\end{verbatim}

The first comment line indicates the domains in the relation (in the above case, {\tt V0} and {\tt H0},
represented using 16 and 12 unique BDD variables, respectively).

If there are $N$ domains, there are $N$ following comment lines, each specifying the
BDD variables assigned to the corresponding domain.

The following line has two numbers: the number of nodes in the represented BDD (28489 in this case) and the number of variables
in the BDD factory from which the BDD was dumped to disk.  Note that the number of variables (134 in this case) is not
necessarily the number of variables in the represented BDD (16+12=28 in this case) though it is guaranteed to be
greater than or equal to it.

The following line specifies the BDD variable order in var2level format.  In this case, the specified order subsumes
{\tt V0\_H0} (notice that levels ``81 79 77 75 73 71 69 67 65 63 61 59 57 55 53 51", which are at positions ``14 15 ... 28 29" 
in the specified order are lower than levels ``105 104 103 102 101 100 99 98 97 96 95 94" which are at positions
``82 83 .. 92 93").

Each of the following lines specifies a unique node in the represented BDD; it has format ``X V L H" where:
\begin{itemize}
\item X is the ID of the BDD node
\item V is the ID of the bdd variable labeling that node (unless it is 0 or 1, in which case it represents a leaf node)
\item L is the ID of the BDD node's low child
\item H is the ID of the BDD node's high child
\end{itemize}

The order of these lines specifying BDD nodes is such that the lines specifying the BDD nodes in the L and H entries
of each BDD node are guaranteed to occur before the line specifying that BDD node (for instance, the L entry 287 on
the second line and the R entry 349123 on the third line are IDs of BDD nodes specified on the first and second lines,
respectively).

Note on Terminology: 
The {\it support} of a BDD {\tt b} is another BDD {\tt r} = {\tt b.support()}
that represents all the variables used in {\tt b}.
The support BDD {\tt r} is a linear tree each of whose nodes contains a separate variable,
the low branch is 0, and high branch is the node representing the next variable.
To list all the variables used in a BDD {\tt b} use {\tt r.scanSet()}.
Needless to say, {\tt scanSet()} simply walks along the high branches
starting at the root of BDD {\tt r}.

### 2011-03-14 main/doc/getting_started.tex
new documentation

\chapter{Getting Started}
\label{chap:getting-started}

This chapter describes how to download, install, and run Chord.
Section \ref{sec:downloading-binaries} describes how to obtain pre-built binaries of Chord.
Section \ref{sec:downloading-sources} describes how to obtain the source code of Chord
and Section \ref{sec:downloading-sources} explains how to build it.
Finally, Section \ref{sec:running-chord} describes how to run Chord.

\section{Downloading Binaries}
\label{sec:downloading-binaries}

To obtain Chord's pre-built binaries, download file \xlink{chord-bin-2.0.tar.gz}{http://jchord.googlecode.com/files/chord-bin-2.0.tar.gz}.
It contains the following files:

\begin{enumerate}
\item
\code{chord.jar}, which contains the class files of Chord and of libraries used by Chord.
\item
\code{libbuddy.so}, \code{buddy.dll}, and \code{libbuddy.dylib}: you can keep one of these files
depending upon whether you intend to run Chord on Linux, Windows/Cygwin, or MacOS, respectively.
These files are needed only if you want to use the BuDDy BDD library for executing program
analyses written in Datalog.
\item
\code{libchord_instr_agent.so}: this file is needed only if you want to use the JVMTI-based bytecode
instrumentation agent for executing dynamic program analyses in Chord.
\end{enumerate}

Novice users can ignore items (2) and (3) until they become more familiar with Chord.
The binaries mentioned in items (2) and (3) might not be compatible with your machine, in which case you
can either forgo using them (with hardly any noticeable difference in functionality),
or you can download the sources (see Section \ref{sec:downloading-sources}) and build them yourself
(see Section \ref{sec:compiling-sources}).

\section{Downloading Source Code}
\label{sec:downloading-sources}

To obtain Chord's source code, download the following files:

\begin{enumerate}
\item
Mandatory: file \xlink{chord-src-2.0.tar.gz}{http://jchord.googlecode.com/files/chord-src-2.0.tar.gz}, which contains Chord's source code and the jars of libraries used by Chord.
\item
Optional: file \xlink{chord-libsrc-2.0.tar.gz}{http://jchord.googlecode.com/files/chord-libsrc-2.0.tar.gz}, which contains the source
code of libraries used by Chord (e.g., joeq, javassist, bddbddb, etc.)
\end{enumerate}

Alternatively, you can obtain the latest development snapshot from the SVN repository:

\begin{verbatim}
    prompt> svn checkout http://jchord.googlecode.com/svn/trunk/ chord
\end{verbatim}

Instead of checking out the entire \code{trunk/}, which contains several sub-directories, you can check out
specific sub-directories:
\begin{itemize}
\item \code{main/} contains Chord's source code and the jars of libraries used by Chord.
\item \code{libsrc/} contains the source code of libraries used by Chord (e.g., joeq, javassist, bddbddb, etc.).
\item \code{test/} contains Chord's regression tests.
\item many more; these might eventually move into \code{main/}.
\end{itemize}

Files \code{chord-2.0-src.tar.gz} and \code{chord-2.0-libsrc.tar.gz} mentioned above are essentially stable releases of
the \code{main/} and \code{libsrc/} directories, respectively.

\section{Compiling the Source Code}
\label{sec:compiling-sources}

Compiling Chord's source code requires the following software:

\begin{itemize}
\item
A JVM with JDK 5 or higher, e.g. \xlink{IBM J9}{http://www.ibm.com/developerworks/java/jdk/} or
\xlink{Oracle HotSpot}{http://www.oracle.com/technetwork/java/javase/downloads/index.html}.
\item
\xlink{Apache Ant}{http://ant.apache.org/}, a Java build tool.
\end{itemize}

Directory {\tt main/} contains a file named {\tt build.xml} which is
interpreted by Apache Ant.  To see the various targets available, run
the following command in that directory:

\begin{verbatim}
    prompt> ant
\end{verbatim}

To compile Chord, run the following command in the same directory:

\begin{verbatim}
    prompt> ant compile
\end{verbatim}

This will compile Chord's Java sources from \code{main/src/} to class files
in \code{main/classes/}, as well as build a jar file \code{main/chord.jar} containing
these class files as well as the class files in the jars of libraries that are used by Chord
and are provided under \code{main/lib/} (e.g., \code{joeq.jar}, \code{javassist.jar},
\code{bddbddb.jar}, etc.).
Additionally:
\begin{itemize}
\item
If system property \code{chord.use.buddy} is set to \code{true},
then the C source code of the \xlink{BuDDy}{http://buddy.sourceforge.net/} BDD library
from \code{main/bdd/} will be compiled to a shared library in \code{main/}
(\code{libbuddy.so} on Linux, \code{buddy.dll} on Windows, and
\code{libbuddy.dylib} on MacOS); this library is used for executing
program analyses written in Datalog
using \xlink{bddbddb}{http://bddbddb.sourceforge.net/} (a BDD-based Datalog solver).
\item
If system property \code{chord.use.jvmti} is set to \code{true},
then the C++ source code of the JVMTI-based bytecode instrumentation agent from
\code{main/agent/} will be compiled to a shared library in \code{main/}
(\code{libchord_instr_agent.so} on all architectures); this agent is
used for computing program scope dynamically and for executing
dynamic program analyses.
\end{itemize}
Properties \code{chord.use.buddy} and \code{chord.use.jvmti} are defined in file \code{main/chord.properties}.  The
default value of both these properties is \code{false}.
If either of them is set to \code{true}, then you will also need a utility like GNU Make (to run the
\code{Makefile}'s under \code{main/bdd/} and \code{main/agent/}) and a C++ compiler.

\section{Running Chord}
\label{sec:running-chord}

Running Chord requires a JVM with JDK 5 or higher.
There are two equivalent ways to run Chord.
One way is to run the following command in the \code{main/} directory:

\begin{verbatim}
    prompt> ant -D<key1>=<value1> ... -D<keyN>=<valueN> run
\end{verbatim}

The above requires Apache Ant (a Java build tool) to be installed on your machine.  The alternative, which
does not require Apache Ant, is to run the following command:

\begin{verbatim}
    prompt> java -cp <...>/chord.jar -D<key1>=<value1> ... -D<keyN>=<valueN> chord.project.Boot
\end{verbatim}

where \code{<...>} denotes the absolute or relative path of the directory containing file \code{chord.jar};
that directory is also expected to contain any other binaries in Chord's installation
(e.g., \code{libbuddy.so} and \code{libchord_instr_agent.so}).

Each ``\code{-D<key>=<value>}" argument above sets the system property named \code{<key>} to value
\code{<value>}.  The only way to specify inputs to Chord is via system properties; there is no
command-line argument processing.
Chapter\ref{chap:properties} describes all system properties recognized by Chord.

### 2011-03-14 main/doc/predefined_analyses.tex
new documentation

--- 2010-12-18 main/doc/datarace.tex
@@
-\section{Datarace analysis}
-\label{sec:datarace}
+\chapter{Predefined Analyses}
+\label{chap:predefined-analyses}
 
-Run the datarace analysis provided in Chord by running the following
+This chapter describes various predefined analyses in Chord.
+
+\section{Static Datarace Analysis}
+
+To run Chord's static datarace analysis, run the following
 command in Chord's {\tt main/} directory:
@@
 
-\noindent where {\tt <...>} is the absolute path of a directory
+where {\tt <...>} denotes the absolute or relative path of a directory
 containing a file named \code{chord.properties} which defines
 properties \code{chord.main.class}, \code{chord.class.path}, and
-\code{chord.src.path}.  See Section \ref{sec:chord-sysprops} for the
+\code{chord.src.path}.  See Chapter \ref{chap:properties} for the
 meaning of these properties.
@@
 in Chord's {\tt main/} directory with {\tt <...>} replaced by
-\code{/.../examples/datarace_test/}.  Upon successful completion, the
+\code{examples/datarace_test/}.  Upon successful completion, the
 following files should be produced in directory
-\code{main/examples/datarace_test/chord_output/}:
+\code{examples/datarace_test/chord_output/}:
 
@@
 \end{itemize}
+
+\section{Static Deadlock Analysis}
+
+To run Chord's static deadlock analysis,  run the following
+command in Chord's {\tt main/} directory:
+
+\begin{verbatim}
+    prompt> ant -Dchord.work.dir=<...> -Dchord.run.analyses=deadlock-java run
+\end{verbatim}
+
+where {\tt <...>} denotes the absolute or relative path of a directory
+containing a file named \code{chord.properties} which defines
+properties \code{chord.main.class}, \code{chord.class.path}, and
+\code{chord.src.path}.  See Chapter \ref{chap:properties} for the
+meaning of these properties.
+
+Directory \code{main/examples/deadlock_test/} provides a toy Java
+program on which you can run the deadlock analysis.  First run {\tt
+  ant} in that directory (in order to compile the program's {\tt
+  .java} files to {\tt .class} files) and then run the above command
+in Chord's {\tt main/} directory with {\tt <...>} replaced by
+\code{examples/deadlock_test/}.  Upon successful completion, the
+file \code{deadlocks.html} should be produced in directory
+\code{examples/deadlock_test/chord_output/}.
+
+\section{May-Alias and Call-Graph Analyses}
+
+TODO

### 2011-03-14 main/doc/properties.tex
new documentation

--- 2011-03-11 main/doc/chord_sysprops.tex
@@
-% TODO: write some text at beginning of each subsection
-% TODO: do cross-referencing (i.e. put "See Also" for properties used by multiple subsections)
+\chapter{Chord Properties}
+\label{chap:properties}
 
-\section{Chord System Properties}
-\label{sec:chord-sysprops}
+The only way to specify inputs to Chord is by means of system properties.
+There is no command-line argument processing in Chord and any command-line arguments are ignored.
+Section \ref{sec:properties-setting} explains how to set properties and
+Section \ref{sec:properties-meaning} explains the meaning of properties recognized by Chord.
 
-Users can use system properties to control Chord's execution.
+\section{How to Set Properties}
+\label{sec:properties-setting}
+
+A property can be passed to Chord in any of several ways.
+The reason for providing multiple ways is to provide users shorthand ways for defining properties
+once and for all for a particular input Java program, or even once and for all across all Chord runs. 
+The following are the different ways by which a property can be passed to Chord in decreasing order of precedence:
+\begin{enumerate}
+\item
+On the command-line via the \code{-D<key>=<value>} format.
+
+Use this option to specify properties specific to the current run of Chord.
+
+Typical usage of this option is by running the following command in
+the \code{main/} directory of Chord's installation:
+\begin{verbatim}
+    prompt> ant -D<key1>=<value1> ... -D<keyN>=<valueN> run
+\end{verbatim}
+
+\item
+Via the file specified by property \code{chord.props.file}.
+
+Use this option to specify once and for all properties of the program to
+be analyzed (e.g., property \code{chord.main.class} specifying the name of that program's main class, property \code{chord.class.path}
+specifying the application classpath of that program, etc.).
+
+The default value of this property is \code{[chord.work.dir]/chord.properties} and typically
+does not need to be changed by users.
+Property \code{chord.work.dir} specifies the directory in which Chord will run.
+The default value of this property is the current directory but users must typically set
+it on the command line, as follows:
+
+\begin{verbatim}
+    prompt> ant -Dchord.work.dir=<...> run
+\end{verbatim}
+where ``\code{<...>}" denotes the absolute or relative path of the top-level directory of
+the program to be analyzed, which contains a file named \code{chord.properties} that users
+must create themselves and populate with properties that are recognized by chord and
+are specific to that program.
+
+\item
+Via file \code{main/chord.properties}.
+
+Use this option to specify once and for all properties you would like to hold in every run of Chord
+(e.g., property \code{chord.max.heap} specifying the maximum runtime heap size to be used by the
+JVM running Chord).
+\end{enumerate}
+
+\section{Meaning of Properties}
+\label{sec:properties-meaning}
+
 The following properties are recognized by Chord.
@@
 
-\paragraph{Program Properties}
+\subsection{Java Program Properties} 
+\label{sec:program-props}
 
-This section describes basic properties of the given program, such as
+This section describes basic properties of the given Java program, such as
 its main class, the location(s) of its class files and Java source
 files, and command-line arguments to be used to run the program.
+\\[10pt]
 
-\noindent \code{chord.work.dir}
+\code{chord.work.dir}
 \begin{quote}
@@
 
-\noindent \code{chord.props.file}
+\code{chord.props.file}
 \begin{quote}
@@
 
-\noindent \code{chord.main.class}
+\code{chord.main.class}
 \begin{quote}
@@
 
-\noindent \code{chord.class.path}
+\code{chord.class.path}
 \begin{quote}
@@
 
-\noindent \code{chord.src.path}
+\code{chord.src.path}
 \begin{quote}
@@
 
-\noindent \code{chord.run.ids}
+\code{chord.run.ids}
 \begin{quote}
@@
 
-\noindent \code{chord.args.<id>}
+\code{chord.args.<id>}
 \begin{quote}
@@
 
-\noindent \code{chord.runtime.jvmargs}
+\code{chord.runtime.jvmargs}
 \begin{quote}
@@
 
-\paragraph{Analysis Scope Properties}
+\subsection{Analysis Scope Properties}
+\label{sec:scope-props}
 
 This section describes properties that specify how the analysis scope of the given program is computed.
-See Section \ref{sec:building-scope} for more details.
+See Chapter \ref{chap:computing-scope} for more details.
+\\[10pt]
 
-\noindent \code{chord.scope.kind}
+\code{chord.scope.kind}
 \begin{quote}
@@
 
-\noindent \code{chord.reflect.kind}
+\code{chord.reflect.kind}
 \begin{quote}
-{\bf Type:} {\tt [none|static|dynamic|static_cast]} \\
+{\bf Type:} {\tt [none|static|dynamic|static\_cast]} \\
 {\bf Description:} TODO \\
@@
 
-\noindent \code{chord.ch.kind}
+\code{chord.ch.kind}
 \begin{quote}
@@
 
-\noindent \code{chord.ssa}
+\code{chord.ssa}
 \begin{quote}
@@
 
-\noindent \code{chord.std.scope.exclude}
+\code{chord.std.scope.exclude}
 \begin{quote}
@@
 
-\noindent \code{chord.ext.scope.exclude}
+\code{chord.ext.scope.exclude}
 \begin{quote}
@@
 
-\noindent \code{chord.scope.exclude}
+\code{chord.scope.exclude}
 \begin{quote}
@@
 
-\noindent \code{chord.std.check.exclude}
+\code{chord.std.check.exclude}
 \begin{quote}
@@
 
-\noindent \code{chord.ext.check.exclude}
+\code{chord.ext.check.exclude}
 \begin{quote}
@@
 
-\noindent \code{chord.check.exclude}
+\code{chord.check.exclude}
 \begin{quote}
@@
 
-\paragraph{Functionality Properties}
+\subsection{Functionality Properties}
+\label{sec:func-props}
 
-This section describes properties that dictate what Chord outputs during execution.
+This section describes properties that dictate what task(s) Chord must perform.
+\\[10pt]
 
-\noindent \code{chord.build.scope}
+\code{chord.build.scope}
 \begin{quote}
@@
 
-\noindent \code{chord.run.analyses}
+\code{chord.run.analyses}
 \begin{quote}
@@
 
-\noindent \code{chord.print.methods}
+\code{chord.print.methods}
 \begin{quote}
@@
 
-\noindent \code{chord.print.classes}
+\code{chord.print.classes}
 \begin{quote}
@@
 
-\noindent \code{chord.print.all.classes}
+\code{chord.print.all.classes}
 \begin{quote}
@@
 
-\noindent \code{chord.print.rels}
+\code{chord.print.rels}
 \begin{quote}
@@
 
-\noindent \code{chord.print.project}
+\code{chord.print.project}
 \begin{quote}
@@
 
-\noindent \code{chord.print.results}
+\code{chord.print.results}
 \begin{quote}
@@
 
-\noindent \code{chord.verbose}
+\code{chord.verbose}
 \begin{quote}
@@
 
-\paragraph{Project Properties}
+\subsection{Project Properties}
+\label{sec:project-props}
 
 This section describes properties regarding program analyses executed by Chord.
+\\[10pt]
 
-\noindent \code{chord.classic}
+\code{chord.classic}
 \begin{quote}
@@
 
-\noindent \code{chord.std.java.analysis.path}
+\code{chord.std.java.analysis.path}
 \begin{quote}
@@
 
-\noindent \code{chord.ext.java.analysis.path}
+\code{chord.ext.java.analysis.path}
 \begin{quote}
@@
 
-\noindent \code{chord.java.analysis.path}
+\code{chord.java.analysis.path}
 \begin{quote}
@@
 
-\noindent \code{chord.std.dlog.analysis.path}
+\code{chord.std.dlog.analysis.path}
 \begin{quote}
@@
 
-\noindent \code{chord.ext.dlog.analysis.path}
+\code{chord.ext.dlog.analysis.path}
 \begin{quote}
@@
 
-\noindent \code{chord.dlog.analysis.path}
+\code{chord.dlog.analysis.path}
 \begin{quote}
@@
 
-\paragraph{Instrumentation Properties}
+\subsection{Instrumentation Properties}
+\label{sec:instr-props}
 
 This section describes properties regarding execution of instrumented programs for dynamic analyses.
+\\[10pt]
 
 % TODO: mention chord.scope.exclude* here
-\noindent \code{chord.use.jvmti}
+\code{chord.use.jvmti}
 \begin{quote}
@@
 
-\noindent \code{chord.instr.kind}
+\code{chord.instr.kind}
 \begin{quote}
@@
 
-\noindent \code{chord.trace.kind}
+\code{chord.trace.kind}
 \begin{quote}
@@
 
-\noindent \code{chord.trace.block.size}
+\code{chord.trace.block.size}
 \begin{quote}
@@
 
-\noindent \code{chord.dynamic.haltonerr}
+\code{chord.dynamic.haltonerr}
 \begin{quote}
@@
 
-\noindent \code{chord.dynamic.timeout}
+\code{chord.dynamic.timeout}
 \begin{quote}
@@
 
-\noindent \code{chord.max.cons.size}
+\code{chord.max.cons.size}
 \begin{quote}
@@
 
-\paragraph{Caching Properties}
+\subsection{Caching Properties}
+\label{sec:caching-props}
 
 TODO
+\\[10pt]
 
-\noindent \code{chord.reuse.scope}
+\code{chord.reuse.scope}
 \begin{quote}
@@
 
-\noindent \code{chord.reuse.rels}
+\code{chord.reuse.rels}
 \begin{quote}
@@
 
-\noindent \code{chord.reuse.traces}
+\code{chord.reuse.traces}
 \begin{quote}
@@
 
-\paragraph{Chord JVM Properties}
+\subsection{Chord JVM Properties}
+\label{sec:jvm-props}
 
 TODO
+\\[10pt]
 
-\noindent \code{java.class.path}
+\code{java.class.path}
 \begin{quote}
@@
 
-\noindent \code{user.dir}
+\code{user.dir}
 \begin{quote}
@@
 
-\noindent \code{chord.max.heap}
+\code{chord.max.heap}
 \begin{quote}
@@
 
-\noindent \code{chord.max.stack}
+\code{chord.max.stack}
 \begin{quote}
@@
 
-\noindent \code{chord.jvmargs}
+\code{chord.jvmargs}
 \begin{quote}
@@
 
-\paragraph{BDD Properties}
+\subsection{BDD Properties}
+\label{sec:bdd-props}
 
 TODO
+\\[10pt]
 
-\noindent \code{chord.use.buddy}
+\code{chord.use.buddy}
 \begin{quote}
@@
 
-\noindent \code{chord.bddbddb.max.heap}
+\code{chord.bddbddb.max.heap}
 \begin{quote}
@@
 
-\paragraph{Output Location Properties}
+\subsection{Output Location Properties}
+\label{sec:output-props}
 
@@
 Most users will not want to alter the default values of these properties.
+\\[10pt]
 
-\noindent \code{chord.out.file}
+\code{chord.out.file}
 \begin{quote}
@@
 
-\noindent \code{chord.err.file}
+\code{chord.err.file}
 \begin{quote}
@@
 
-\noindent \code{chord.out.dir}
+\code{chord.out.dir}
 \begin{quote}
@@
 
-\noindent \code{chord.reflect.file}
+\code{chord.reflect.file}
 \begin{quote}
@@
 
-\noindent \code{chord.methods.file}
+\code{chord.methods.file}
 \begin{quote}
@@
 
-\noindent \code{chord.classes.file}
+\code{chord.classes.file}
 \begin{quote}
@@
 
-\noindent \code{chord.bddbddb.work.dir}
+\code{chord.bddbddb.work.dir}
 \begin{quote}
@@
 
-\noindent \code{chord.boot.classes.dir}
+\code{chord.boot.classes.dir}
 \begin{quote}
@@
 
-\noindent \code{chord.user.classes.dir}
+\code{chord.user.classes.dir}
 \begin{quote}
@@
 
-\noindent \code{chord.instr.scheme.file}
+\code{chord.instr.scheme.file}
 \begin{quote}
@@
 
-\noindent \code{chord.trace.file}
+\code{chord.trace.file}
 \begin{quote}

### 2011-03-14 main/doc/example.tex
new documentation

\chapter{An Example Run of Chord}
\label{chap:example}

This chapter describes how to setup a Java program and run predefined analyses
in Chord on it.

Suppose the Java program to be analyzed has the following directory structure:

\begin{verbatim}
    example/
        src/
            foo/
                Main.java
                ...
        classes/
            foo/
                Main.class
                ...
        lib/
            src/
                taz/
                    ...
            jar/
                taz.jar
        chord.properties
\end{verbatim}

The above structure is typical: the program's Java source
files are under {\tt src/}, its class files are under {\tt classes/},
the source and jar files of the libraries used by the program are
under \code{lib/src/} and \code{lib/jar/}, respectively.

To run Chord on the above program, run the following command in
Chord's {\tt main/} directory:

\begin{verbatim}
    prompt> ant -Dchord.work.dir=<...>/example run
\end{verbatim}

where \code{<...>} denotes the absolute or relative path of the
directory containing the above \code{example/} directory.
Property \code{chord.work.dir} specifies the working directory during Chord's execution.
Chord searches for a file named \code{chord.properties} under that directory, and
loads upfront all properties defined in that file, if it exists.  Alternatively, these properties
may be defined on the command-line, using the \code{-D<key>=<value>} format.
A sample \code{chord.properties} file for the above program is as follows:

\begin{verbatim}
    chord.main.class=foo.Main
    chord.class.path=classes:lib/jar/taz.jar
    chord.src.path=src:lib/src
    chord.run.ids=0,1
    chord.args.0="-thread 1 -n 10"
    chord.args.1="-thread 2 -n 50"
\end{verbatim}

Each relative file/directory name in the value of any property
defined in this file (e.g., the \code{lib/src} directory name in the value of
property \code{chord.src.path} above) is treated relative to Chord's working directory
(which is specified by property \code{chord.work.dir}).

Chapter \ref{chap:properties} presents an exhaustive listing of
properties recognized by Chord.  Here, we only describe those defined
in the above sample \code{chord.properties} file:

\begin{itemize}
\item
\code{chord.main.class} specifies the fully-qualified name of the main
class of the program.
\item
\code{chord.class.path} specifies the application classpath
of the program (the standard library classpath is implicitly
included).
\item
\code{chord.src.path} specifies the Java source path of the program.
All program analyses in Chord operate on Java bytecode.  The only use
of this property is to HTMLize the Java source files of the program so
that the results of program analyses can be reported at the Java
source code level.
\item
\code{chord.run.ids} specifies a list of IDs to identify runs of the
program.  It is used by dynamic program analyses to determine how many
times the program must be run.  An additional use of this property is
to allow specifying the command-line arguments to use in the run
having ID {\tt <id>} via property \code{chord.args.<id>}, as
illustrated by properties \code{chord.args.0} and \code{chord.args.1}
in the above example.
\end{itemize}

In the above case, Chord does not do much beyond loading the above
\code{chord.properties} file.  For Chord to do something interesting,
additional properties must be set that specify the task(s)
Chord must perform.  All tasks are summarized in Section \ref{sec:func-props}.
The most common task is to run one or more analyses on an input Java program.
Chord implements many standard analyses, including various may-alias and
call-graph analyses, thread-escape analysis, various concurrency analyses,
etc. Many of these analyses are described in Chapter \ref{chap:predefined-analyses}.
Moreover, users can extend these analyses or define their own ones, as
explained in Chapter \ref{chap:analyses}.
Broadly, there are two kinds of analyses in Chord:

\begin{itemize}
\item
Those expressed imperatively in Java. They are in \code{*.java} files under
directory \code{main/src/chord/analyses/}, and are of the form:

\begin{verbatim}
    @Chord(
        name=<JAVA ANALYSIS NAME>
        ...
    )
    public class XYZ extends ... {
        ...
    }
\end{verbatim}

\item

Those expressed declaratively in Datalog.  They are in \code{*.dlog} files under
directory \code{main/src/chord/analyses/}, and are of the form:

\begin{verbatim}
    # name=<DLOG ANALYSIS NAME>

    # Program domains
    .include ...
    .include ...

    # BDD variable order
    .bddvarorder ...
\end{verbatim}
\end{itemize}

To run one or more such analyses, run the following command in Chord's \code{main/} directory:

\begin{verbatim}
    prompt> ant -Dchord.work.dir=<PROGRAM DIR NAME> -Dchord.run.analyses=<name1>,...,<nameN> run
\end{verbatim}

where \code{<PROGRAM DIR NAME>} specifies an example Java program as described above,
and each of \code{<name1>}, ..., \code{<nameN>} is the name of an analysis to run,
such as \code{<JAVA ANALYSIS NAME>} and \code{<DLOG ANALYSIS NAME>} in the Java and
Datalog analyses depicted above.

For instance, to run a basic may-alias and call-graph analysis (called 0CFA),
run the following command in Chord's main directory:

\begin{verbatim}
    prompt> ant -Dchord.work.dir=<PROGRAM DIR NAME> -Dchord.run.analyses=cipa-0cfa-dlog run
\end{verbatim}

This instructs Chord to run the Datalog analysis named \code{cipa-0cfa-dlog},
which happens to be defined in file \code{main/src/chord/analyses/alias/cipa_0cfa.dlog}.

The output of the above command is of the form:

\begin{verbatim}
    Buildfile: build.xml

    run:
         [java] Chord run initiated at: Mar 13, 2011 10:31:08 PM
         [java] ENTER: cipa-0cfa-dlog
         [java] ENTER: T
         [java] ENTER: RTA
         [java] Iteration: 0
         [java] Iteration: 1
         [java] Iteration: 2
         [java] LEAVE: RTA
         [java] SAVING dom T size: 1386
         [java] LEAVE: T
         [java] ENTER: F
         [java] SAVING dom F size: 4120
         [java] LEAVE: F
         ...
         [java] ENTER: MputStatFldInst
         [java] SAVING rel MputStatFldInst size: 739
         [java] LEAVE: MputStatFldInst
         [java] ENTER: statIM
         [java] SAVING rel statIM size: 3359
         [java] LEAVE: statIM
         [java] Starting command: 'java ... net.sf.bddbddb.Solver ...\chord_analyses_alias_cipa_0cfa.dlog'
         [java] Relation VH: 541 nodes, 449.0 elements (V0,H0)
         [java] Relation FH: 137 nodes, 8.0 elements (H0,F0)
         [java] Relation HFH: 199 nodes, 35.0 elements (H0,F0,H1)
         [java] Relation IM: 590 nodes, 69.0 elements (I0,M0)
         [java] Finished command: 'java ... net.sf.bddbddb.Solver ...\chord_analyses_alias_cipa_0cfa.dlog'
         [java] LEAVE: cipa-0cfa-dlog
         [java] Chord run completed at: Mar 13, 2011 10:31:36 PM
         [java] Total time: 00:00:27:671 hh:mm:ss:ms

    BUILD SUCCESSFUL
    Total time: 28 seconds
\end{verbatim}

Notice that although Chord is only asked to run the 0CFA analysis on the command line, it automatically
runs many other analyses before finally running the desired analysis and terminating.
Each analysis in Chord is written modularly, independent of other analyses, along with lightweight
annotations specifying the inputs and outputs of the analysis.  Chord's runtime automatically computes
producer-consumer relationships between analyses (e.g., determines which analysis produces as output a
result that is needed as input by another analysis).  Before running a desired analysis
(such as 0CFA in the above example), Chord recursively runs other analyses until the
inputs to the desired analysis have been computed; it finally runs the desired analysis to produce
the outputs of that analysis.

In the above example, the 0CFA analysis consumes {\it program relations} 
such as \code{MputStatFldInst} and \code{statIM}, each of which is produced by a separate Java analysis with the
corresponding name, and produces program relations \code{VH}, \code{FH}, \code{HFH}, and \code{IM}.

The program relations consumed by the 0CFA analysis contain basic program facts.  For instance,
\code{MputStatFldInst} is a relation containing each tuple (m,f,v) such that method m in the input Java
program contains a putstatic instruction of the form ``f = v", while \code{statIM} is a relation
containing each tuple (i,m) such that m is the target method of invokestatic instruction i.

The program relations produced by the 0CFA analysis represent points-to information and the
call graph of the input Java program as computed by the analysis.
Specifically, relations \code{VH}, \code{FH}, and \code{HFH} represent points-to information 
for local variables, static fields, and instance fields, respectively,
while relation \code{IM} represents the call graph, namely, it contain each tuple (i,m) such that
m is a possible target method of call site i.

Metavariables m, f, i, and v above range over entities in so-called {\it program domains} M, F, I, and V, respectively.
A program domain is a set of entities of a certain kind in the input Java program.  For instance, M is the domain representing
the set of all methods in the input Java program, F is the domain representing the set of all fields, I is the domain representing
the set of all call sites in methods in M, and V is the domain representing the set of all local variables of reference type in methods in M.
Each of these domains is produced by a separate Java analysis with the corresponding name.
Notice that the analyses producing these domains run upfront because these domains
are consumed by the analyses that produce relations such as \code{MputStatFldInst} and
\code{statIM}, which in turn are consumed by the desired 0CFA analysis.

During execution, Chord dumps intermediate and final results to files in the
directory specified by property \code{chord.out.dir}, whose default value is \code{[chord.work.dir]/chord_output/}
and typically does not need to be changed by users.
For the above example, this directory is \code{<PROGRAM DIR NAME>/chord_output/}.

The verbosity of Chord's output above is controlled by property \code{chord.verbose}, whose
default value is 1.  At verbosity level 0, the above command produces less voluminous output
of the form:

\begin{verbatim}
	Buildfile: build.xml

	run:
	     [java] Chord run initiated at: Mar 13, 2011 10:35:01 PM
	     [java] Chord run completed at: Mar 13, 2011 10:35:28 PM
	     [java] Total time: 00:00:26:297 hh:mm:ss:ms

    BUILD SUCCESSFUL
    Total time: 26 seconds
\end{verbatim}

### 2011-03-14 main/doc/whatis_chord.tex
more documentation
new documentation

@@
-\section{What is Chord?}
-\label{sec:whatis-chord}
+\chapter{What is Chord?}
+\label{chap:whatis-chord}
 
@@
 \item
+
 {\bf Stand-alone:} It provides various off-the-shelf analyses (e.g., various
-call-graph analyses; points-to analyses; thread-escape analysis;
-static slicing; static and dynamic concurrency analyses for finding
+may-alias and call-graph analyses; thread-escape analysis;
+static slicing analysis; static and dynamic concurrency analyses for finding
 races, deadlocks, and atomicity violations; etc.)
+
 \item
+
 {\bf Flexible:} It allows users to express a broad range of analyses,
@@
 combined static/dynamic analyses.
+
 \item
+
 {\bf Efficient:} It executes analyses in a demand-driven fashion, caches
@@
 them in parallel.
+
 \item
+
 {\bf Deterministic:} It guarantees that the result is the same regardless of
@@
 
-\noindent Chord is intended to work on a variety of platforms,
+Chord is intended to work on a variety of platforms,
 including Linux, Windows/Cygwin, and MacOS.  It has been tested most

### 2011-03-14 main/doc/dynamic_analysis.tex
more documentation

--- 2011-03-12 main/doc/writing_dynamic_analysis.tex
@@
+\chapter{Dynamic Analysis}
+\label{chap:dynamic-analysis}
+
+This chapter describes everything about dynamic analysis in Chord.
+Section \ref{sec:writing-dynamic-analysis} describes how to write a dynamic analysis,
+Section \ref{sec:running-dynamic-analysis} describes how to compile and run it,
+and Section \ref{sec:instr-events} describes common dynamic analysis events
+supported in Chord.
+
 \section{Writing a Dynamic Analysis}
@@
 Follow the following steps to write your own dynamic analysis.
-
-\subsection{Implementing the Analysis}
 
@@
 
-\subsection{Configuring the Analysis}
+\section{Compiling and Running a Dynamic Analysis}
+\label{sec:running-dynamic-analysis}
+
+Compile the analysis by placing the directory containing class \code{MyDynamicAnalysis} created
+above in the path defined by property \code{chord.ext.java.analysis.path}.
+
+Provide the IDs of program runs to be generated (say 1, 2, ..., M) and the command-line arguments to be
+used for the program in each of those runs (say \code{<args1>}, ..., \code{<argsM>}) via properties
+\code{chord.run.ids=1,2,...,N} and \code{chord.args.1=<args1>}, ..., \code{chord.args.M=<argsM>}.
+By default, \code{chord.run.ids=0} and \code{chord.args.0=""}, that is, the program will be run only
+once (using run ID 0) with no command-line arguments.
+
+To run the analysis, set property \code{chord.run.analyses} to \code{<name-of-analysis>}
+(recall that \code{<name-of-analysis>} is the name provided in the \code{@Chord} annotation for class
+\code{MyDynamicAnalysis} created above).
+
+{\bf Note:} The IBM J9 JVM on Linux is highly recommended if you intend to use Chord for
+dynamic program analysis, as it allows you to instrument the entire JDK; using any other
+platform will likely require excluding large parts of the JDK from being instrumented.
+Additionally, if you intend to use online (load-time) bytecode instrumentation in
+your dynamic program analysis, then you will need JDK 6 or higher, since this functionality
+requires the \code{java.lang.instrument} API with class retransformation support (the latter
+support is available only in JDK 6 and higher).
 
 You can change the default values of various properties for configuring your dynamic analysis;
-see ``Analysis Scope Properties" and ``Instrumentation Properties" in Section \ref{sec:chord-sysprops}.
+see Section \ref{sec:scope-props} and Section \ref{sec:instr-props} in Chapter \ref{chap:properties}.
 For instance:
@@
 
-\subsection{Compiling and Running the Analysis}
+\section{Common Dynamic Analysis Events}
+\label{sec:instr-events}
 
-Compile the analysis by placing the directory containing class \code{MyDynamicAnalysis} created
-above in the path defined by property \code{chord.ext.java.analysis.path}.
+Chord provides support for instrumenting common dynamic analysis events.
+The below table describes these events.
 
-Provide the IDs of program runs to be generated (say 1, 2, ..., M) and the command-line arguments to be
-used for the program in each of those runs (say \code{<args1>}, ..., \code{<argsM>}) via properties
-\code{chord.run.ids=1,2,...,N} and \code{chord.args.1=<args1>}, ..., \code{chord.args.M=<argsM>}.
-By default, \code{chord.run.ids=0} and \code{chord.args.0=""}, that is, the program will be run only
-once (using run ID 0) with no command-line arguments.
-
-To run the analysis, set property \code{chord.run.analyses} to \code{<name-of-analysis>}
-(recall that \code{<name-of-analysis>} is the name provided in the \code{@Chord} annotation for class
-\code{MyDynamicAnalysis} created above).
-
-{\bf Note:} The IBM J9 JVM on Linux is highly recommended if you intend to use Chord for
-dynamic program analysis, as it allows you to instrument the entire JDK; using any other
-platform will likely require excluding large parts of the JDK from being instrumented.
-Additionally, if you intend to use online (load-time) bytecode instrumentation in
-your dynamic program analysis, then you will need JDK 6 or higher, since this functionality
-requires the \code{java.lang.instrument} API with class retransformation support (the latter
-support is available only in JDK 6 and higher).
+\begin{table}
+\begin{center}
+\htmlattributes*{table}{border="1" cellpadding="10"}
+\begin{tabular}{|l|p{4.3in}|} 
+\hline
+{\bf Event Kind} & {\bf Description} \\
+\hline
+EnterMainMethod(\bt) & After thread \bt\ enters method \bm\ (in domain M).
+\\
+\hline
+EnterMethod(\bm, \bt) & After thread \bt\ enters method \bm\ (in domain M).
+\\
+\hline
+LeaveMethod(\bm, \bt) & Before thread \bt\ leaves method \bm\ (in domain M).
+\\
+\hline
+EnterLoop(\bw, \bt) & Before thread \bt\ begins loop \bw\ (in domain W).
+\\
+\hline
+LoopIteration(\bw, \bt) & Before thread \bt\ starts a new iteration of loop \bw\ (in domain W).
+\\
+\hline
+LeaveLoop(\bw, \bt) & After thread \bt\ finishes loop \bw\ (in domain W).
+\\
+\hline
+BasicBlock(\bb, \bt) & Before thread \bt\ enters basic block \bb\ (in domain B).
+\\
+\hline
+Quad(\bp, \bt) & Before thread \bt\ executes quad at program point \bp\ (in domain P).
+\\
+\hline
+BefMethodCall(\bi, \bt, \bo) & 
+Before thread \bt\ executes the method invocation statement at program point \bi\ (in domain I) with this argument as object \bo. 
+{\bf Note:} This event is not generated before constructor calls; use the BefNew event.
+\\
+\hline
+AftMethodCall(\bi, \bt, \bo) &
+After thread \bt\ executes the method invocation statement at program point \bi\ (in domain I) with this argument as object \bo. 
+{\bf Note:} This event is not generated after constructor calls; use the AftNew event.
+\\
+\hline
+BefNew(\bh, \bt, \bo) & Before thread \bt\ executes a \code{new} bytecode instruction and allocates fresh object \bo\ at program point \bh\ (in domain H).
+\\
+\hline
+AftNew(\bh, \bt, \bo) & After thread \bt\ executes a \code{new} bytecode instruction and allocates fresh object \bo\ at program point \bh\ (in domain H).
+\\
+\hline
+NewArray(\bh, \bt, \bo) & After thread \bt\ executes a \code{newarray} bytecode instruction and allocates fresh object \bo\ at program point \bh\ (in domain H).
+\\
+\hline
+GetstaticPrimitive(\be, \bt, \bg) & After thread \bt\ reads primitive-typed static field \bg\ (in domain F) at program point \be\ (in domain E).
+\\
+\hline
+GetstaticReference(\be, \bt, \bg, \bo) & After thread \bt\ reads object \bo\ from reference-typed static field \bg\ (in domain F) at program point \be\ (in domain E).
+\\
+\hline
+PutstaticPrimitive(\be, \bt, \bg) & After thread \bt\ writes primitive-typed static field \bg\ (in domain F) at program point \be\ (in domain E).
+\\
+\hline
+PutstaticReference(\be, \bt, \bg, \bo) & After thread \bt\ writes object \bo\ to reference-typed static field \bg\ (in domain F) at program point \be\ (in domain E).
+\\
+\hline
+GetfieldPrimitive(\be, \bt, \bb, \bg) & After thread \bt\ reads primitive-typed instance field \bg\ (in domain F) of object \bb\ at program point \be\ (in domain E).
+\\
+\hline
+GetfieldReference(\be, \bt, \bb, \bg, \bo) & After thread \bt\ reads object \bo\ from reference-typed instance field \bg\ (in domain F) of object \bb\ at program point \be\ (in domain E).
+\\
+\hline
+PutfieldPrimitive(\be, \bt, \bb, \bg) & After thread \bt\ writes primitive-typed instance field \bg\ (in domain F) of object \bb\ at program point \be\ (in domain E).
+\\
+\hline
+PutfieldReference(\be, \bt, \bb, \bg, \bo) & After thread \bt\ writes object \bo\ to reference-typed instance field \bg\ (in domain F) of object \bb\ at program point \be\ (in domain E).
+\\
+\hline
+AloadPrimitive(\be, \bt, \bb, \bi) & After thread \bt\ reads the primitive-typed element at index \bi\ of array object \bb\ at program point \be\ (in domain E).
+\\
+\hline
+AloadReference(\be, \bt, \bb, \bi, \bo) & After thread \bt\ reads object \bo\ from the reference-typed element at index \bi\ of array object \bb\ at program point \be\ (in domain E).
+\\
+\hline
+AstorePrimitive(\be, \bt, \bb, \bi) & After thread \bt\ writes the primitive-typed element at index \bi\ of array object \bb\ at program point \be\ (in domain E).
+\\
+\hline
+AstoreReference(\be, \bt, \bb, \bi, \bo) & After thread \bt\ writes object \bo\ to the reference-typed element at index \bi\ of array object \bb\ at program point \be\ (in domain E).
+\\
+\hline
+%ReturnPrimitive(\bp, \bt) & Not yet supported.
+%\\
+%ReturnReference(\bp, \bt) & Not yet supported.
+%\\
+%ExplicitThrow(\bp, \bt, \bo) & Not yet supported.
+%\\
+%ImplicitThrow(\bp, \bt, \bo) & Not yet supported.
+%\\
+ThreadStart(\bi, \bt, \bo) & Before thread \bt\ calls the \code{start()} method of \code{java.lang.Thread} at program point \bi\ (in domain I) and spawns a thread \bo.
+\\
+\hline
+ThreadJoin(\bi, \bt, \bo) & Before thread \bt\ calls the \code{join()} method of \code{java.lang.Thread} at program point \bi\ (in domain I) to join with thread \bo.
+\\
+\hline
+AcquireLock(\bl, \bt, \bo) & After thread \bt\ executes a statement of the form ``monitorenter \bo'' or enters a method synchronized on \bo\ at program point \bl\ (in domain L).
+\\
+\hline
+ReleaseLock(\br, \bt, \bo) & Before thread \bt\ executes a statement of the form ``monitorexit \bo'' or leaves a method synchronized on \bo\ at program point \br\ (in domain R).
+\\
+\hline
+Wait(\bi, \bt, \bo) & Before thread \bt\ calls the \code{wait()} method of \code{java.lang.Object} on object \bo\ at program point \bi\ (in domain I).
+\\
+\hline
+Notify(\bi, \bt, \bo) & Before thread \bt\ calls the \code{notify()} or \code{notifyAll()} method of \code{java.lang.Object} on object \bo\ at program point \bi\ (in domain I).
+\T \\
+\hline
+\end{tabular}
+\end{center}
+\end{table}

### 2011-03-14 main/doc/getting_started.tex
more documentation

@@
 depending upon whether you intend to run Chord on Linux, Windows/Cygwin, or MacOS, respectively.
-These files are needed only if you want to use the BuDDy BDD library for executing program
-analyses written in Datalog.
+These files are needed only if you want BDD library BuDDy to be used when 
+the BDD-based Datalog solver bddbddb in Chord runs analyses written in Datalog.
 \item
-\code{libchord_instr_agent.so}: this file is needed only if you want to use the JVMTI-based bytecode
-instrumentation agent for executing dynamic program analyses in Chord.
+\code{libchord_instr_agent.so}: this file is needed only if you want the JVMTI-based bytecode
+instrumentation agent to be used when Chord runs dynamic analyses.
 \end{enumerate}
@@
 
-\begin{enumerate}
+\begin{itemize}
 \item
@@
 code of libraries used by Chord (e.g., joeq, javassist, bddbddb, etc.)
-\end{enumerate}
+\end{itemize}
 
@@
 specific sub-directories:
+
 \begin{itemize}
-\item \code{main/} contains Chord's source code and the jars of libraries used by Chord.
-\item \code{libsrc/} contains the source code of libraries used by Chord (e.g., joeq, javassist, bddbddb, etc.).
-\item \code{test/} contains Chord's regression tests.
-\item many more; these might eventually move into \code{main/}.
+\item
+\code{main/} contains Chord's source code and the jars of libraries used by Chord.
+\item
+\code{libsrc/} contains the source code of libraries used by Chord (e.g., joeq, javassist, bddbddb, etc.).
+\item
+\code{test/} contains Chord's regression tests.
+\item
+many more; these might eventually move into \code{main/}.
 \end{itemize}
@@
 Additionally:
+
 \begin{itemize}
 \item
+
 If system property \code{chord.use.buddy} is set to \code{true},
-then the C source code of the \xlink{BuDDy}{http://buddy.sourceforge.net/} BDD library
+then the C source code of BDD library \xlink{BuDDy}{http://buddy.sourceforge.net/} 
 from \code{main/bdd/} will be compiled to a shared library in \code{main/}
 (\code{libbuddy.so} on Linux, \code{buddy.dll} on Windows, and
-\code{libbuddy.dylib} on MacOS); this library is used for executing
-program analyses written in Datalog
-using \xlink{bddbddb}{http://bddbddb.sourceforge.net/} (a BDD-based Datalog solver).
+\code{libbuddy.dylib} on MacOS); this library is used 
+by BDD-based Datalog solver \xlink{bddbddb}{http://bddbddb.sourceforge.net/} 
+in Chord for running analyses written in Datalog.
+
 \item
+
 If system property \code{chord.use.jvmti} is set to \code{true},
@@
 (\code{libchord_instr_agent.so} on all architectures); this agent is
-used for computing program scope dynamically and for executing
-dynamic program analyses.
+used in Chord for computing analysis scope dynamically and for running
+dynamic analyses.
 \end{itemize}
+
 Properties \code{chord.use.buddy} and \code{chord.use.jvmti} are defined in file \code{main/chord.properties}.  The
 default value of both these properties is \code{false}.
-If either of them is set to \code{true}, then you will also need a utility like GNU Make (to run the
+If you set either of them to \code{true}, then you will also need a utility like GNU Make (to run the
 \code{Makefile}'s under \code{main/bdd/} and \code{main/agent/}) and a C++ compiler.
@@
 
-Each ``\code{-D<key>=<value>}" argument above sets the system property named \code{<key>} to value
-\code{<value>}.  The only way to specify inputs to Chord is via system properties; there is no
+Each ``\code{-D<key>=<value>}" argument above sets the system property named \code{<key>} to
+the value denoted by \code{<value>}.  The only way to specify inputs to Chord is via system properties; there is no
 command-line argument processing.

### 2011-03-14 main/doc/properties.tex
more documentation

@@
 The following are the different ways by which a property can be passed to Chord in decreasing order of precedence:
+
 \begin{enumerate}
 \item
+
 On the command-line via the \code{-D<key>=<value>} format.
@@
 
-Typical usage of this option is by running the following command in
-the \code{main/} directory of Chord's installation:
+Typical usage of this option is by running the following command in Chord's \code{main/} directory:
 \begin{verbatim}
@@
 \item
+
 Via the file specified by property \code{chord.props.file}.
@@
 \item
+
 Via file \code{main/chord.properties}.
@@
 Use this option to specify once and for all properties you would like to hold in every run of Chord
-(e.g., property \code{chord.max.heap} specifying the maximum runtime heap size to be used by the
+(e.g., property \code{chord.max.heap} specifying the maximum heap memory size to be used by the
 JVM running Chord).
@@
 
-This section describes basic properties of the given Java program, such as
+This section describes properties of the Java program to be analyzed, such as
 its main class, the location(s) of its class files and Java source
@@
 {\bf Description:} Working directory during Chord's execution.  This is
-usually the top-level directory of the given program. \\
+usually the top-level directory of the input Java program. \\
 {\bf Default value:} current working directory
@@
 {\bf Type:} class \\
-{\bf Description:} Fully-qualified name of the main class of the given program (e.g., \code{com.example.Main}).
+{\bf Description:} Fully-qualified name of the main class of the input Java program (e.g., \code{com.example.Main}).
 \end{quote}
@@
 {\bf Type:} path \\
-{\bf Description:} Classpath of the given program.  It does not need to include
+{\bf Description:} Classpath of the input Java program.  It does not need to include
 boot classes (i.e., classes in \code{[sun.boot.class.path]}) or
-standard extensions (i.e., classes in {\tt .jar} files in directory
-\code{[java.home]/lib/ext/}). \\
+standard extensions (i.e., classes in jar files in directory \code{[java.home]/lib/ext/}). \\
 {\bf Default value:} {\tt ""}
@@
 {\bf Type:} path \\
-{\bf Description:} Java source path of the given program. \\
+{\bf Description:} Source path of the input Java program. \\
 {\bf Default value:} {\tt ""} \\
 {\bf Note:} Chord analyzes only Java bytecode, not Java source code.
-This property is used only by the task of converting Java source files into HTML files by program analyses that need to present their analysis results at the Java source code level (by calling method \code{chord.program.Program.g().HTMLizeJavaSrcFiles())}.
+This property is used only by the task of converting Java source files into HTML files by analyses that need to present their analysis results at the Java source code level (by calling method \code{chord.program.Program.g().HTMLizeJavaSrcFiles())}.
 \end{quote}
@@
 {\bf Type:} string list \\
-{\bf Description:} List of IDs to identify runs of the given program. \\
+{\bf Description:} List of IDs to identify runs of the input Java program. \\
 {\bf Default value:} {\tt 0} \\
-{\bf Note:} This property is used only when Chord executes the given program, namely, when it is asked to compute the program scope dynamically (i.e., when \code{[chord.scope.kind]=dynamic}) or when it is asked to execute a dynamic program analysis.
+{\bf Note:} This property is used only when Chord runs the input Java program, namely, when it is asked to compute the analysis scope dynamically (i.e., when \code{[chord.scope.kind]=dynamic}) or when it is asked to run a dynamic analysis.
 \end{quote}
@@
 {\bf Type:} string \\
-{\bf Description:} Command-line arguments string to be used for the given program in the run having ID {\tt <id>}. \\
+{\bf Description:} Command-line arguments string to be used for the input Java program in the run having ID {\tt <id>}. \\
 {\bf Default value:} {\tt ""} \\
-{\bf Note:} This property is used only when Chord executes the given program, namely, when it is asked to compute the program scope dynamically (i.e., when \code{[chord.scope.kind]=dynamic}) or when it is asked to execute a dynamic program analysis.
+{\bf Note:} This property is used only when Chord runs the input Java program, namely, when it is asked to compute the analysis scope dynamically (i.e., when \code{[chord.scope.kind]=dynamic}) or when it is asked to run a dynamic analysis.
 \end{quote}
@@
 {\bf Type:} string \\
-{\bf Description:} Arguments to JVM which runs the given program. \\
+{\bf Description:} Arguments to JVM which runs the input Java program. \\
 {\bf Default value:} {\tt "-ea -Xmx1024m"} \\
-{\bf Note:} This property is used only when Chord executes the given program, namely, when it is asked to compute the program scope dynamically (i.e., when \code{[chord.scope.kind]=dynamic}) or when it is asked to execute a dynamic program analysis.
+{\bf Note:} This property is used only when Chord runs the input Java program, namely, when it is asked to compute the analysis scope dynamically (i.e., when \code{[chord.scope.kind]=dynamic}) or when it is asked to run a dynamic analysis.
 \end{quote}
@@
 
-This section describes properties that specify how the analysis scope of the given program is computed.
+This section describes properties that specify how the analysis scope of the input Java program is computed.
 See Chapter \ref{chap:computing-scope} for more details.
@@
 {\bf Type:} {\tt [dynamic|rta|cha]} \\
-{\bf Description:} Algorithm to compute program scope.  The choices are {\tt dynamic} (dynamic analysis), {\tt rta} (Rapid Type Analysis), and {\tt cha} (Class Hierarchy Analysis). \\
-{\bf Default value:} {\tt rta}
+{\bf Description:} Algorithm to compute analysis scope.  The choices are {\tt dynamic} (dynamic analysis), {\tt rta} (Rapid Type Analysis), and {\tt cha} (Class Hierarchy Analysis). \\
+{\bf Default value:} {\tt rta} \\
+{\bf Note:} This property is ignored if property \code{chord.reuse.scope} is set to {\tt true} and the files specified by properties \code{chord.methods.file} and \code{chord.reflect.file} exist. 
 \end{quote}
@@
 {\bf Type:} {\tt [static|dynamic]} \\
-{\bf Description:} Algorithm to build the class hierarchy.
-If it is {\tt dynamic}, then the given program is executed and classes not loaded by the JVM while running the program are excluded while building the class hierarchy.
-\\
+{\bf Description:} Algorithm to build the class hierarchy.  If it is {\tt dynamic}, then the input Java program is executed
+and classes not loaded by the JVM while running the program are excluded while building the class hierarchy. \\
 {\bf Default value:} {\tt static} \\
-{\bf Note:} This property is relevant only if \code{chord.scope.kind} is {\tt cha} since only this scope computing algorithm queries the class hierarchy. 
+{\bf Note:} This property is relevant only if \code{chord.scope.kind} is {\tt cha} since only this
+scope computing algorithm queries the class hierarchy. 
 \end{quote}
@@
 {\bf Type:} bool  \\
-{\bf Description:} Do SSA transformation for all methods deemed reachable by the algorithm used to compute program scope. \\
+{\bf Description:} Do SSA (Static Single Assignment) transformation of the bodies of all methods deemed reachable by the algorithm used to compute analysis scope. \\
 {\bf Default value:} {\tt true}
@@
 {\bf Type:} string list \\
-{\bf Description:} Partial list of prefixes of names of classes whose declared methods must be treated as no-ops.
-Conventionally, this property is used to specify classes to be excluded from the scope of all programs to be analyzed, e.g., classes in the JDK standard library.
-\\
+{\bf Description:} Partial list of prefixes of names of classes, typically inside the JDK standard library, whose methods must be treated as no-ops. \\
 {\bf Default value:} {\tt ""}
@@
 {\bf Type:} string list \\
-{\bf Description:} Partial list of prefixes of names of classes whose declared methods must be treated as no-ops.
-Conventionally, this property is used to specify classes to be excluded from the scope of specific programs to be analyzed.
-\\
+{\bf Description:} Partial list of prefixes of names of classes, typically outside the JDK standard library, whose methods must be treated as no-ops. \\
 {\bf Default value:} {\tt ""}
@@
 {\bf Type:} string list \\
-{\bf Description:} Complete list of prefixes of names of classes whose declared methods must be treated as no-ops. \\
+{\bf Description:} Complete list of prefixes of names of classes whose methods must be treated as no-ops. \\
 {\bf Default value:} \code{"[chord.std.scope.exclude],[chord.ext.scope.exclude]"}
@@
 {\bf Type:} string list \\
-{\bf Description:} List of prefixes of names of classes and packages inside the standard library to be excluded by program analyses.  Interpretation of this property is analysis-specific. \\
+{\bf Description:} Partial list of prefixes of names of classes, typically inside the JDK standard library, to be excluded by analyses.  Interpretation of this property is analysis-specific. \\
 {\bf Default value:} \code{"java.,javax.,sun.,com.sun.,com.ibm.,} \code{org.apache.harmony."}
@@
 {\bf Type:} string list \\
-{\bf Description:} List of prefixes of names of classes and packages outside the standard library to be excluded by program analyses.  Interpretation of this property is analysis-specific. \\
+{\bf Description:} Partial list of prefixes of names of classes, typically outside the JDK standard library, to be excluded by analyses.  Interpretation of this property is analysis-specific. \\
 {\bf Default value:} {\tt ""}
@@
 {\bf Type:} string list \\
-{\bf Description:} List of prefixes of names of classes and packages to be excluded by program analyses.  Interpretation of this property is analysis-specific. \\
+{\bf Description:} Complete list of prefixes of names of classes to be excluded by analyses.  Interpretation of this property is analysis-specific. \\
 {\bf Default value:} \code{"[chord.std.check.exclude],[chord.ext.check.exclude]"}
@@
 {\bf Type:} bool \\
-{\bf Description:} Compute the scope of the given program using the algorithm specified by properties \code{chord.scope.kind} and \code{chord.reuse.scope}. \\
+{\bf Description:} Compute the analysis scope of the input Java program using the algorithm. \\
 {\bf Default value:} {\tt false} \\
-{\bf Note:} The program scope is computed regardless of the value of this property if another task (e.g., a program analysis specified via property \code{chord.run.analyses}) demands it by calling method \code{chord.program.Program.g()}.
+{\bf Note:} The analysis scope is computed regardless of the value of this property if another task (e.g., an analysis specified via property \code{chord.run.analyses}) demands it.
 \end{quote}
@@
 {\bf Type:} string list  \\
-{\bf Description:} List of names of program analyses to be run in order. \\
+{\bf Description:} List of names of analyses to be run in order. \\
 {\bf Default value:} {\tt ""} \\
-{\bf Note:} If the analysis is written in Java, its name is specified via statement {\tt name=<...>} in its {\tt @Chord} annotation, and is the name of the class itself if this statement is missing.
-If the analysis is written in Datalog, its name is specified via a line of the form ``{\tt \# name=<...>}", and is the absolute location of the file itself if this line is missing.
+{\bf Note:} If the analysis is written in Java, its name is specified via statement {\tt name=<...>} in its {\tt @Chord} annotation.
+If the analysis is written in Datalog, its name is specified via a line of the form ``{\tt \# name=<...>}".
 \end{quote}
@@
 {\bf Type:} string list  \\
-{\bf Description:} List of methods in scope whose intermediate representation to print to standard output. \\
+{\bf Description:} List of methods whose intermediate representation to print to standard output. \\
 {\bf Default value:} {\tt ""} \\
@@
 {\bf Type:} string list  \\
-{\bf Description:} List of classes in scope whose intermediate representation to print to standard output. \\
+{\bf Description:} List of classes whose intermediate representation to print to standard output. \\
 {\bf Default value:} {\tt ""} \\
@@
 {\bf Default value:} {\tt ""} \\
-{\bf Note:} This functionality must be used with caution as certain program relations, albeit represented compactly as BDDs, may contain a large number (e.g., billions) of tuples, resulting in voluminous output when printed to a file.
-See Section \ref{sec:tuning-datalog-analysis} for a more efficient way to query the contents of program relations (namely, by using the {\tt debug} target provided in file {\tt build.xml} in Chord's {\tt main/} directory).
+{\bf Note:} This functionality must be used with caution as certain program relations, albeit represented compactly as BDDs, may contain a large number (e.g., millions) of tuples, resulting in voluminous output when printed in explicit form to a text file.
+See Section \ref{sec:tuning-datalog-analysis} for a more efficient way to query the contents of program relations (namely, by using the {\tt debug} target provided in file {\tt build.xml} in Chord's \code{main/} directory).
 \end{quote}
@@
 {\bf Type:} bool \\
-{\bf Description:} Create files \code{targets_sortby_name.html}, \code{targets_sortby_kind.html}, and \code{targets_sortby_producers.html} in directory \code{[chord.out.dir]}, publishing all tasks and targets defined by program analyses in paths \code{[chord.java.analysis.path]} and \code{[chord.dlog.analysis.path]}.
+{\bf Description:} Create files \code{targets_sortby_name.html}, \code{targets_sortby_kind.html}, and \code{targets_sortby_producers.html} in directory \code{[chord.out.dir]}, publishing all tasks and targets defined by analyses in paths \code{[chord.java.analysis.path]} and \code{[chord.dlog.analysis.path]}.
 \\
@@
 {\bf Type:} bool \\
-{\bf Description:} Print the results of program analyses in HTML.  Interpretation of this property is analysis-specific.  \\
+{\bf Description:} Print the results of analyses in HTML.  Interpretation of this property is analysis-specific.  \\
 {\bf Default value:} {\tt true}
@@
 \begin{quote}
-{\bf Type:} int in the range [0..6]  \\
-{\bf Description:} Control the verbosity of logging during Chord's execution.  \\
-{\bf Default value:} {\tt 2}
+{\bf Type:} int in the range [0..5]  \\
+{\bf Description:} Control the verbosity of messages during Chord's execution.  \\
+{\bf Default value:} {\tt 1}
 \end{quote}
@@
 
-This section describes properties regarding program analyses executed by Chord.
+This section describes properties regarding analyses executed by Chord.
 \\[10pt]
@@
 {\bf Type:} path \\
-{\bf Description:} TODO \\
-{\bf Default value:} TODO
+{\bf Description:} Partial classpath of analyses written in Java (i.e., {\tt @Chord}-annotated classes).
+Conventionally, it includes all such analyses that are predefined in Chord.  \\
+{\bf Default value:} The absolute path of file \code{chord.jar}
 \end{quote}
@@
 {\bf Type:} path \\
-{\bf Description:} TODO \\
-{\bf Default value:} TODO
+{\bf Description:} Partial classpath of analyses written in Java (i.e., {\tt @Chord}-annotated classes).
+Conventionally, it includes all such external analyses that users want to run. \\
+{\bf Default value:} {\tt ""}
 \end{quote}
@@
 {\bf Type:} path \\
-{\bf Description:} Classpath containing program analyses written in Java (i.e., {\tt @Chord}-annotated classes).  \\
-{\bf Default value:} \code{[chord.main.dir]/classes/}
+{\bf Description:} Complete classpath of analyses written in Java (i.e., {\tt @Chord}-annotated classes). \\
+{\bf Default value:} \code{[chord.std.java.analysis.path]:[chord.ext.java.analysis.path]}
 \end{quote}
@@
 {\bf Type:} path \\
-{\bf Description:} TODO \\
-{\bf Default value:} TODO
+{\bf Description:} Partial classpath of analyses written in Datalog (i.e., files with suffix {\tt .dlog} or {\tt .datalog}).
+Conventionally, it includes all such analyses that are predefined in Chord.  \\
+{\bf Default value:} The absolute path of file \code{chord.jar}
 \end{quote}
@@
 {\bf Type:} path \\
-{\bf Description:} TODO \\
-{\bf Default value:} TODO
+{\bf Description:} Partial classpath of analyses written in Datalog (i.e., files with suffix {\tt .dlog} or {\tt .datalog}).
+Conventionally, it includes all such external analyses that users want to run. \\
+{\bf Default value:} {\tt ""}
 \end{quote}
@@
 {\bf Type:} path  \\
-{\bf Description:} Path of directories containing program analyses written in Datalog (i.e., files with suffix {\tt .dlog} or {\tt .datalog}). \\
-{\bf Default value:} \code{[chord.main.dir]/src/}
+{\bf Description:} Complete classpath of analyses written in Datalog (i.e., files with suffix {\tt .dlog} or {\tt .datalog}). \\
+{\bf Default value:} \code{[chord.std.dlog.analysis.path]:[chord.ext.dlog.analysis.path]}
 \end{quote}
@@
 
-This section describes properties regarding execution of instrumented programs for dynamic analyses.
+This section describes properties regarding bytecode instrumentation and dynamic analysis.
 \\[10pt]
@@
 {\bf Type:} bool \\
-{\bf Description:} TODO \\
-{\bf Default value:} TODO
+{\bf Description:} Whether the JVMTI-based bytecode instrumentation agent from \code{main/agent/} must be used for running dynamic analyses. \\
+{\bf Default value:} \code{false}
 \end{quote}
@@
 {\bf Type:} int \\
-{\bf Description:} Number of bytes to read/write in a single operation from/to the event file used by a multi-JVM dynamic analysis. \\
+{\bf Description:} Number of bytes to read/write in a single operation from/to the event trace file in a multi-JVM dynamic analysis. \\
 {\bf Default value:} {\tt 4096}
@@
 {\bf Type:} bool \\
-{\bf Description:} Whether or not to terminate Chord if the given program terminates abnormally during dynamic analysis. \\
+{\bf Description:} Whether to terminate Chord if the input Java program terminates abnormally during dynamic analysis. \\
 {\bf Default value:} true
@@
 
-TODO
+This section describes properties that specify what must be reused by Chord, if available, from previous runs instead of recomputing.
 \\[10pt]
@@
 {\bf Type:} bool \\
-{\bf Description:} Compute program scope using the information in files specified by properties \code{chord.methods.file} and \code{chord.reflect.file}. \\
+{\bf Description:} Compute analysis scope using the information in files specified by properties \code{chord.methods.file} and \code{chord.reflect.file}, if both of those files exist. \\
 {\bf Default value:} {\tt false} \\
@@
 {\bf Type:} bool  \\
-{\bf Description:} Construct program relations from BDDs stored on disk (from a previous run of Chord) whenever possible instead of re-computing them. \\
+{\bf Description:} Load each desired program relation named \code{<name>} from the BDD stored in file \code{[chord.bddbddb.work.dir]/<name>.bdd}, if the file exists. \\
 {\bf Default value:} {\tt false}
@@
 {\bf Type:} bool \\
-{\bf Description:} Reuse dynamic traces computed by a previous run of Chord if they exist.  Property \code{chord.trace.kind} must be set to {\tt full} if this property is set to {\tt true}. \\
+{\bf Description:} Reuse event traces stored in file(s) \code{chord.trace.file]_full_ver0_runM.txt} for dynamic analysis, if those files exist,
+where \code{M} ranges over run IDs specified by property \code{chord.run.ids}.
+Property \code{chord.trace.kind} must be set to {\tt full} if this property is set to {\tt true}. \\
 {\bf Default value:} {\tt false}
@@
 
-TODO
+This section describes properties regarding the JVM that runs Chord.
 \\[10pt]
-
-\code{java.class.path}
-\begin{quote}
-{\bf Type:} path \\
-{\bf Description:} TODO \\
-{\bf Default value:} TODO
-\end{quote}
-
-\code{user.dir}
-\begin{quote}
-{\bf Type:} location
-{\bf Description:} TODO \\
-{\bf Default value:} TODO
-\end{quote}
 
@@
 {\bf Type:} string \\
-{\bf Description:} Maximum memory size of JVM running Chord. \\
+{\bf Description:} Maximum heap memory size of the JVM running Chord. \\
 {\bf Default value:} {\tt 1024m}
@@
 {\bf Type:} string \\
-{\bf Description:} Maximum thread stack size of JVM running Chord. \\
+{\bf Description:} Maximum thread stack size of the JVM running Chord. \\
 {\bf Default value:} {\tt 32m}
@@
 {\bf Type:} string \\
-{\bf Description:} Arguments to JVM running Chord. \\
+{\bf Description:} Arguments to the JVM running Chord. \\
 {\bf Default value:} \code{"-showversion} \code{-ea} \code{-Xmx[chord.max.heap]} \code{-Xss[chord.max.stack]"}
@@
 
-TODO
+This section describes properties concerning BDD-based Datalog solver bddbddb that is used by Chord to run analyses written in Datalog.
 \\[10pt]
@@
 {\bf Type:} bool \\
-{\bf Description:} TODO \\
-{\bf Default value:} TODO
+{\bf Description:} Whether BDD library BuDDy from \code{main/bdd/} must be used by bddbddb. \\
+{\bf Default value:} \code{false}
 \end{quote}
@@
 {\bf Type:} string \\
-{\bf Description:} Maximum memory size of JVM running Datalog solver bddbddb. \\
+{\bf Description:} Maximum heap memory size of JVM running bddbddb. \\
 {\bf Default value:} {\tt 1024m} \\
-{\bf Note:}  Each program analysis written in Datalog is run in a separate JVM.
-This is primarily because there may be multiple invocations of bddbddb in a single run of Chord and it is difficult to reset the global state of bddbddb on each invocation.
+{\bf Note:} bddbddb is invoked in a separate JVM for each analysis written in Datalog that is executed.
+This is primarily because multiple Datalog analyses may be executed in a single run of Chord,
+resulting in multiple invocations of bddbddb, and it is difficult to reset the state of bddbddb on each invocation. 
 \end{quote}
@@
 This section describes properties specifying the names of files and directories output by Chord.
-Most users will not want to alter the default values of these properties.
+Most users will not need to alter the default values of these properties.
 \\[10pt]
@@
 {\bf Type:} location \\
-{\bf Description:} Absolute location of the directory used by the Datalog solver bddbddb as its input/output directory (namely, for program domain files {\tt *.dom} and {\tt *.map}, and program relation files {\tt *.bdd}).
+{\bf Description:} Absolute location of the directory used by BDD-based Datalog solver bddbddb as its input/output directory (namely, for program domain files {\tt *.dom} and {\tt *.map}, and program relation files {\tt *.bdd}).
 \\
@@
 {\bf Type:} location \\
-{\bf Description:} Absolute location of the directory from/to which instrumented JDK classes used by the given program are read/written by dynamic program analyses. \\
+{\bf Description:} Absolute location of the directory from/to which instrumented classes of the input Java program inside the JDK standard library are read/written by dynamic analyses. \\
 {\bf Default value:} \code{[chord.out.dir]/boot_classes/}
@@
 {\bf Type:} location  \\
-{\bf Description:} Absolute location of the directory from/to which instrumented non-JDK classes of the given program are read/written by dynamic program analyses. \\
+{\bf Description:} Absolute location of the directory from/to which instrumented classes of the input Java program outside the JDK standard library are read/written by dynamic analyses. \\
 {\bf Default value:} \code{[chord.out.dir]/user_classes/}
@@
 {\bf Type:} location \\
-{\bf Description:} Absolute location of the file specifying the kind and format of events in dynamic trace files. \\
+{\bf Description:} Absolute location of the file specifying the kind and format of events in trace files used by dynamic analyses. \\
 {\bf Default value:} \code{[chord.out.dir]/scheme.ser}
@@
 {\bf Type:} location  \\
-{\bf Description:} Absolute location of the dynamic trace file.  \\
+{\bf Description:} Absolute location of trace files used by dynamic analyses. \\
 {\bf Default value:} \code{[chord.out.dir]/trace} \\

### 2011-03-15 main/doc/dynamic_analysis.tex
more documentation

@@
 
-A sample such class \code{MyDynamicAnalysis} is shown below:
+A sample such class \code{MyDynamicAnalysis} is shown in Figure \ref{fig:dynamic-analysis}.
 
+\begin{figure}
+{\small
 \begin{verbatim}
@@
         InstrScheme scheme;
-
         @Override
@@
         }
-
         @Override
@@
         }
-
         @Override
         public void initPass() {
-            // ***TODO***: User code to be executed once before each
-            // instrumented program run starts.
+            // ***TODO***: User code to be executed once before each instrumented program run starts.
         }
-
         @Override
         public void donePass() {
-            // ***TODO***: User code to be executed once after each
-            // instrumented program run finishes.
+            // ***TODO***: User code to be executed once after each instrumented program run finishes.
         }
-
-        // User-defined event handlers for this dynamic analysis.
-        // No-ops if not overridden.
         @Override
         public void process<event1>(<args1>) {
-            // ***TODO***: User code for handling events of kind
-            // <event1> with format <args1>.
+            // ***TODO***: User code for handling events of kind <event1> with format <args1>.
         }
@@
         public void process<eventN>(<argsN>) {
-            // ***TODO***: User code for handling events of kind
-            // <eventN> with format <argsN>.
+            // ***TODO***: User code for handling events of kind <eventN> with format <argsN>.
         }
@@
 \end{verbatim}
+}
+\label{fig:dynamic-analysis}
+\caption{Dynamic Analysis Template.}
+\end{figure}
 
@@
 BefMethodCall(\bi, \bt, \bo) & 
-Before thread \bt\ executes the method invocation statement at program point \bi\ (in domain I) with this argument as object \bo. 
-{\bf Note:} This event is not generated before constructor calls; use the BefNew event.
+\begin{tabular}{p{4.3in}}
+Before thread \bt\ executes the method invocation statement at program point \bi\ (in domain I) with this argument as object \bo. \\
+{\bf Note:} Not generated before constructor calls; use BefNew event.
+\end{tabular}
 \\
@@
 AftMethodCall(\bi, \bt, \bo) &
-After thread \bt\ executes the method invocation statement at program point \bi\ (in domain I) with this argument as object \bo. 
-{\bf Note:} This event is not generated after constructor calls; use the AftNew event.
+\begin{tabular}{p{4.3in}}
+After thread \bt\ executes the method invocation statement at program point \bi\ (in domain I) with this argument as object \bo. \\
+{\bf Note:} Not generated after constructor calls; use AftNew event.
+\end{tabular}
 \\

### 2011-03-15 main/doc/example.tex
more documentation

@@
 \begin{verbatim}
-    prompt> ant -Dchord.work.dir=<PROGRAM DIR NAME> -Dchord.run.analyses=<name1>,...,<nameN> run
+    prompt> ant -Dchord.work.dir=<PROGRAM DIR> -Dchord.run.analyses=<ANALYSIS NAME> run
 \end{verbatim}
 
-where \code{<PROGRAM DIR NAME>} specifies an example Java program as described above,
-and each of \code{<name1>}, ..., \code{<nameN>} is the name of an analysis to run,
-such as \code{<JAVA ANALYSIS NAME>} and \code{<DLOG ANALYSIS NAME>} in the Java and
+where {\tt <PROGRAM DIR>} denotes a directory containing a Java program as described above,
+and {\tt <ANALYSIS NAME>} denotes the name of an analysis to run,
+such as {\tt <JAVA ANALYSIS NAME>} and {\tt <DLOG ANALYSIS NAME>} in the Java and
 Datalog analyses depicted above.
@@
 For instance, to run a basic may-alias and call-graph analysis (called 0CFA),
-run the following command in Chord's main directory:
+run the following command in Chord's \code{main/} directory:
 
 \begin{verbatim}
-    prompt> ant -Dchord.work.dir=<PROGRAM DIR NAME> -Dchord.run.analyses=cipa-0cfa-dlog run
+    prompt> ant -Dchord.work.dir=<PROGRAM DIR> -Dchord.run.analyses=cipa-0cfa-dlog run
 \end{verbatim}
@@
          [java] LEAVE: statIM
-         [java] Starting command: 'java ... net.sf.bddbddb.Solver ...\chord_analyses_alias_cipa_0cfa.dlog'
+         [java] Starting command: 'java ... chord_analyses_alias_cipa_0cfa.dlog'
          [java] Relation VH: 541 nodes, 449.0 elements (V0,H0)
@@
          [java] Relation IM: 590 nodes, 69.0 elements (I0,M0)
-         [java] Finished command: 'java ... net.sf.bddbddb.Solver ...\chord_analyses_alias_cipa_0cfa.dlog'
+         [java] Finished command: 'java ... chord_analyses_alias_cipa_0cfa.dlog'
          [java] LEAVE: cipa-0cfa-dlog
@@
 The program relations consumed by the 0CFA analysis contain basic program facts.  For instance,
-\code{MputStatFldInst} is a relation containing each tuple (m,f,v) such that method m in the input Java
-program contains a putstatic instruction of the form ``f = v", while \code{statIM} is a relation
-containing each tuple (i,m) such that m is the target method of invokestatic instruction i.
+\code{MputStatFldInst} is a relation containing each tuple ($m$,$f$,$v$) such that method $m$ in the input Java
+program contains a putstatic instruction of the form ``$f$ = $v$", while \code{statIM} is a relation
+containing each tuple ($i$,$m$) such that $m$ is the target method of invokestatic instruction $i$.
 
@@
 for local variables, static fields, and instance fields, respectively,
-while relation \code{IM} represents the call graph, namely, it contain each tuple (i,m) such that
-m is a possible target method of call site i.
+while relation \code{IM} represents the call graph, namely, it contain each tuple ($i$,$m$) such that
+$m$ is a possible target method of call site $i$.
 
-Metavariables m, f, i, and v above range over entities in so-called {\it program domains} M, F, I, and V, respectively.
-A program domain is a set of entities of a certain kind in the input Java program.  For instance, M is the domain representing
-the set of all methods in the input Java program, F is the domain representing the set of all fields, I is the domain representing
-the set of all call sites in methods in M, and V is the domain representing the set of all local variables of reference type in methods in M.
+Metavariables $m$, $f$, $i$, and $v$ above range over entities in so-called {\it program domains} \code{M},
+\code{F}, \code{I}, and \code{V}, respectively.
+A program domain is a set of entities of a certain kind in the input Java program.  For instance,
+\code{M} is the domain representing
+the set of all methods in the input Java program, \code{F} is the domain representing the set of all fields,
+\code{I} is the domain representing
+the set of all call sites in methods in \code{M}, and \code{V} is the domain representing the set of all local variables of reference type in methods in \code{M}.
 Each of these domains is produced by a separate Java analysis with the corresponding name.
@@
 and typically does not need to be changed by users.
-For the above example, this directory is \code{<PROGRAM DIR NAME>/chord_output/}.
+For the above example, this directory is {\tt <PROGRAM DIR>}\code{/chord_output/}.
 

### 2011-03-15 main/doc/properties.tex
more documentation

@@
 \end{verbatim}
-where ``\code{<...>}" denotes the absolute or relative path of the top-level directory of
+where \code{<...>} denotes the absolute or relative path of the top-level directory of
 the program to be analyzed, which contains a file named \code{chord.properties} that users
@@
 its main class, the location(s) of its class files and Java source
-files, and command-line arguments to be used to run the program.
-\\[10pt]
+files, and command-line arguments to be used when running the program.
 
@@
 See Chapter \ref{chap:computing-scope} for more details.
-\\[10pt]
 
@@
 This section describes properties that dictate what task(s) Chord must perform.
-\\[10pt]
 
@@
 This section describes properties regarding analyses executed by Chord.
-\\[10pt]
 
@@
 This section describes properties regarding bytecode instrumentation and dynamic analysis.
-\\[10pt]
 
@@
 This section describes properties that specify what must be reused by Chord, if available, from previous runs instead of recomputing.
-\\[10pt]
 
@@
 This section describes properties regarding the JVM that runs Chord.
-\\[10pt]
 
@@
 This section describes properties concerning BDD-based Datalog solver bddbddb that is used by Chord to run analyses written in Datalog.
-\\[10pt]
 
@@
 Most users will not need to alter the default values of these properties.
-\\[10pt]
 

### 2011-03-15 main/doc/datalog_analysis.tex
more documentation

@@
 A common way to rapidly prototype program analyses in Chord is using a declarative
-logic-programming language called Datalog.  This chapter describes everything about
+logic-programming language called Datalog.  This chapter describes all aspects of
 Datalog analysis in Chord.
@@
 specifying how to compute the output relations from the input relations.
-An example of such an analysis is as follows:
+An example of such an analysis is shown in Figure \ref{fig:datalog-analysis}.
 
+\begin{figure}
+{\small
 \begin{verbatim}
@@
 \end{verbatim}
+}
+\label{fig:datalog-analysis}
+\caption{An Example Datalog Analysis.}
+\end{figure}
 
@@
 will be interleaved in the BDD variable order for this Datalog analysis.
-See Section \ref{chap:bdd-representation} for more details on BDD representations.
+See Section \ref{sec:bdd} for more details on BDD representations.
 

### 2011-03-17 main/doc/predefined_analyses.tex
add material describing predefined points-to analyses

@@
+\xname{predefined_analyses}
 \chapter{Predefined Analyses}
@@
 properties \code{chord.main.class}, \code{chord.class.path}, and
-\code{chord.src.path}.  See Chapter \ref{chap:properties} for the
+\code{chord.src.path}.  See Section \ref{sec:program-props} for the
 meaning of these properties.
@@
 Directory \code{main/examples/datarace_test/} provides a toy Java
-program on which you can run the datarace analysis.  First run {\tt
+program on which one can run the datarace analysis.  First run {\tt
   ant} in that directory (in order to compile the program's {\tt
@@
 Directory \code{main/examples/deadlock_test/} provides a toy Java
-program on which you can run the deadlock analysis.  First run {\tt
+program on which one can run the deadlock analysis.  First run {\tt
   ant} in that directory (in order to compile the program's {\tt
@@
 
-\section{May-Alias and Call-Graph Analyses}
+\section{Points-to and Call-Graph Analyses}
 
-TODO
+Chord offers several choices for computing points-to and call graphs for programs.
+The simplest of these is a context-insensitive analysis. We describe this first,
+because understanding the concepts behind it will help you understand the more
+sophisticated context-sensitive analysis.
+
+In general, context-sensitive approaches are more precise but run more slowly;
+the higher the degree of precision, the slower the analysis.
+
+\subsection{Context-insensitive points-to}
+
+The context-insenstive points-to is implemented in \code{analyses/alias/cipa-0cfa.dlog}.
+The analysis treats all objects allocated at the same site as interchangeable.
+Put another way, each allocation site defines a unique abstract object.
+Domain H is the set of all allocation sites and therefore also of all abstract objects.
+The analysis produces several output relations.
+
+\begin{description}
+\item[VH] contains tuples \texttt{(v,h)} if variable \texttt{v} can point to abstract object \texttt{h}.
+\item[MM] contains tuples \texttt{(m1,m2)} if method \texttt{m1} might call \texttt{m2}.
+\item[IM] contains tuples \texttt{(i,m)} if invocation \texttt{i} might call \texttt{m}.
+\item[reachableM] contains the set of potentially-reachable methods.
+\item[reachableT] contains the set of potentially-reachable types.
+\item[reachableI] contains the set of potentially-reachable invocation statements.
+\item[rootM] contains the set of potentially-reachable methods without a caller (such as \texttt{main} and static initializers).
+\item[FH] contains tuples \texttt{(f,h)} if static field \texttt{f} can point to abstract object \texttt{h}.
+\item[HFH] contains tuples \texttt{(b,f,h)} if instance field \texttt{f} of abstract object \texttt{b} can point to abstract object \texttt{h}.
+\end{description}
+
+\subsection{Context-sensitive}
+
+In a context-sensitive analysis, there is no longer just one abstract object per site.
+Rather, the set of objects a reference can point to depends on the \textit{context} of
+the method call containing the reference.
+Whereas a context-insensitive analysis talks about the domain of methods and the domain
+of allocation sites (M and H), a context-sensitive analysis talks about the domain of contexts, labelled C.
+Domain C refers to both abstract objects and also to abstract calling contexts. (As we will see, these
+are merged for a reason.)
+
+Chord has several context-sensitive analyses, but they all expose the same relations.
+
+\begin{description}
+\item[C] domain containing all abstract contexts.
+\item[CC] contains tuples \texttt{(c,c2)} such that \texttt{c2} is all but the last element of context \texttt{c}.
+\item[CH] contains tuples \texttt{(c,h)} such that object allocation site \texttt{h} is the last element of abstract context \texttt{c}.
+\item[CI] contains tuples \texttt{(c,i)} such that call site \texttt{i} is the last element of abstract context \texttt{c}.
+\item[CVC] contains tuples \texttt{(c,v,o)} such that local \texttt{v} might point to object \texttt{o} in context \texttt{c} of its declaring method.  \\
+Note that both \texttt{o} and \texttt{c} are elements of domain C.
+\item[CFC] contains tuples \texttt{(o1,f,o2)} such that instance field \texttt{f} of object \texttt{o1} might point to object \texttt{o2}.
+\item[FC] contains tuples \texttt{(f,o)} such that static field \texttt{f} may point to object \texttt{o}.
+\item[CICM] contains tuples \texttt{(c,i,c2,m)} if invocation \texttt{i} in context \texttt{c} can reach method 2 (in context \texttt{c2}) .
+\item[rootCM] contains tuples \texttt{(c,m)} such that method \texttt{m} is an entry method in context \texttt{c}.
+\item[reachableCM] contains tuples \texttt{(c,m)} such that method \texttt{m} can be called in context \texttt{c}.
+\end{description}
+
+{\bf Under Construction: need to explain object-sensitive vs context-sensitive, plus how to invoke.}

### 2011-03-17 main/doc/predefined_analyses.tex
extended ariel's documentation of points-to analyses

--- 2011-03-14 main/doc/predefined_analyses.tex
@@
 
+\section{Points-to and Call-Graph Analyses}
+
+Chord offers several choices for computing points-to and call-graph information of Java programs.
+In each of these choices, points-to and call-graph information is computed simultaneously
+(called ``on-the-fly call-graph construction" in the literature in contrast to ``ahead-of-time call-graph
+construction" in which the call graph is computed first followed by points-to information).
+On-the-fly approaches are more precise because, in a dynamically dispatching language like Java,
+as more points-to facts are discovered, more (dynamically dispatched) methods are deemed reachable,
+thereby growing the call graph; the code in these newly added methods in turn results in more
+points-to facts.
+
+Flow-insensitive analysis computes a single abstract heap whereas flow-sensitive analysis computes
+per-program-point abstract heaps.  Context-insensitive analysis analyzes each method at most once (i.e. in a single abstract context),
+whereas context-sensitive analyses potentially analyze each method multiple times, in different abstract contexts.
+Thus, flow- and context-sensitive analyses are more precise but less scalable than flow- and context-insensitive
+analyses, respectively.
+
+Flow-sensitive analysis does not offer much precision over flow-insensitive analysis in
+practice, especially in the absence of strong updates and in the presence of SSA (Static Single Assignment form),
+a program representation that renders a flow-insensitive analysis almost as precise as a flow-sensitive analysis.
+Since analyses in Chord currently perform only weak updates, and since they all operate on an SSA form of the
+input Java program by default, the rest of this section focuses only on flow-insensitive analysis, which is the predominant
+kind of points-to/call-graph analysis in Chord.
+
+We describe context-insensitive analysis first because understanding the concepts behind it will help understand
+the more sophisticated context-sensitive analyses.
+We first recall some relevant program domains:
+\begin{itemize}
+\item M is the domain of all methods.
+\item I is the domain of all method call sites.
+\item F is the domain of all (instance and static) fields.
+\item V is the domain of all local variables of reference type.
+\item H is the domain of all object allocation sites.
+\end{itemize}
+
+\subsection{Context-Insensitive Analysis}
+
+The context-insensitive points-to/call-graph analysis 
+treats each object allocation site as a separate abstract memory location; in other words, it
+can distinguish objects created at different sites but not those created at the same site.
+Additionally, it is field-sensitive, in that it distinguishes between different instance fields of the same object,
+but array-insensitive, in that it cannot distinguish between different elements of the same array;
+all array elements are modeled using a distinguished hypothetical instance field (that has index 0 in domain F).
+
+To run this analysis, run the following command in Chord's main directory:
+
+\begin{verbatim}
+    prompt> ant -Dchord.work.dir=<...> -Dchord.run.analyses=cipa-0cfa-dlog run
+\end{verbatim}
+
+This analysis outputs the following relations:
+
+%is implemented in \code{chord/analyses/alias/cipa-0cfa.dlog}.
+
+\begin{itemize}
+\item
+{\it Call-graph information:}
+\begin{itemize}
+\item rootM subset M contains the set of entry methods that may be reachable; this includes the program's main method
+as well as each static initializer method that may be reachable from the main method.
+\item reachableM subset M contains the set of methods that may be reachable from the program's main method.
+\item IM subset (I $\times$ M) contains tuples \texttt{(i,m)} such that call site \texttt{i} may call method \texttt{m}.
+%\item MM subset (M $\times$ M) contains tuples \texttt{(m1,m2)} such that method \texttt{m1} may call method \texttt{m2}.
+%\item reachableI subset I contains the set of call sites that may be reachable.
+%\item reachableT subset T contains the set of classes that may be reachable.
+\end{itemize}
+\item
+{\it Points-to information:}
+\begin{itemize}
+\item FH subset (F $\times$ H) contains tuples \texttt{(f,h)} such that static field (i.e. global variable) \texttt{f} may point to an object allocated at site \texttt{h}.
+\item VH subset (V $\times$ H) contains tuples \texttt{(v,h)} such that local variable \texttt{v} may point to an object allocated at site \texttt{h}.
+\item HFH subset (H $\times$ F $\times$ H) contains tuples \texttt{(h1,f,h2)} such that instance field \texttt{f} of some object allocated at
+site \texttt{h1} may point to some object allocated at site \texttt{h2}.
+\end{itemize}
+\end{itemize}
+
+\subsection{Context-Sensitive Analysis}
+
+In a context-sensitive analysis, there is no longer just one abstract memory location per object allocation site.
+Rather, the set of objects a reference can point to depends on the \textit{context} in which the method containing the
+reference is called.
+Whereas a context-insensitive analysis talks about the domain of methods (M) and the domain
+of allocation sites (H), a context-sensitive analysis talks about the domain of abstract contexts, labeled C.
+Elements of domain C contain both abstract calling contexts and abstract objects. (These are merged for reasons
+described below.)
+
+Chord has several context-sensitive analyses, but they all expose the same relations, which are
+described below:
+
+\begin{itemize}
+\item
+{\it Context information:}
+\begin{itemize}
+\item C is the domain of all abstract calling contexts and abstract objects.  Each element in this domain is a sequence of zero or more sites,
+where each site can be a call site or an object allocation site.  A sequence may have mixed call sites and object allocation sites.
+The most significant site (i.e. the first site) in each sequence is called the head; the remaining sub-sequence is called the tail.  The
+below three relations relate a sequence with its head and tail.
+\item CC subset (C $\times$ C) contains tuples \texttt{(c1,c2)} such that context \texttt{c2} is the tail of context \texttt{c1}.
+\item CH subset (C $\times$ H) contains tuples \texttt{(c,h)} such that object allocation site \texttt{h} is the head of context \texttt{c}.
+\item CI subset (C $\times$ I) contains tuples \texttt{(c,i)} such that call site \texttt{i} is the head of context \texttt{c}.
+\end{itemize}
+\item
+{\it Call-graph information:}
+\begin{itemize}
+\item rootCM subset (C $\times$ M) contains tuples \texttt{(c,m)} such that method \texttt{m} is an entry method in context \texttt{c}.
+\item reachableCM subset (C $\times$ M) contains tuples \texttt{(c,m)} such that method \texttt{m} may be reachable in context \texttt{c}.
+\item CICM subset (C $\times$ I $\times$ C $\times$ M) contains tuples \texttt{(c1,i,c2,m)} such that call site \texttt{i} in
+context \texttt{c1} may call method \texttt{m2} in context \texttt{c2}.
+\end{itemize}
+\item
+{\it Points-to information:}
+\begin{itemize}
+\item FC subset (F $\times$ C) contains tuples \texttt{(f,o)} such that static field (i.e. global variable) \texttt{f} may point to object \texttt{o}.
+\item CVC subset (C $\times$ V $\times$ C) contains tuples \texttt{(c,v,o)} such that local variable \texttt{v} may point to
+object \texttt{o} in context \texttt{c} of that variable's declaring method. Note that both \texttt{o} and \texttt{c} are elements of domain C.
+\item CFC subset (C $\times$ F $\times$ C) contains tuples \texttt{(o1,f,o2)} such that instance field \texttt{f} of object \texttt{o1} may point to object \texttt{o2}.
+\end{itemize}
+\end{itemize}
+
+{\bf Under Construction: explain object-sensitive vs context-sensitive, plus how to invoke.}
+
 \section{Static Datarace Analysis}
@@
 To run Chord's static datarace analysis, run the following
-command in Chord's {\tt main/} directory:
+command in Chord's main directory:
 
@@
 properties \code{chord.main.class}, \code{chord.class.path}, and
-\code{chord.src.path}.  See Chapter \ref{chap:properties} for the
+\code{chord.src.path}.  See Section \ref{sec:program-props} for the
 meaning of these properties.
 
-Directory \code{main/examples/datarace_test/} provides a toy Java
-program on which you can run the datarace analysis.  First run {\tt
+Directory \code{examples/datarace_test/} in Chord's main directory provides a toy Java
+program on which one can run the datarace analysis.  First run {\tt
   ant} in that directory (in order to compile the program's {\tt
   .java} files to {\tt .class} files) and then run the above command
-in Chord's {\tt main/} directory with {\tt <...>} replaced by
+in Chord's main directory with {\tt <...>} replaced by
 \code{examples/datarace_test/}.  Upon successful completion, the
@@
 To run Chord's static deadlock analysis,  run the following
-command in Chord's {\tt main/} directory:
+command in Chord's main directory:
 
@@
 properties \code{chord.main.class}, \code{chord.class.path}, and
-\code{chord.src.path}.  See Chapter \ref{chap:properties} for the
+\code{chord.src.path}.  See Section \ref{sec:program-props} for the
 meaning of these properties.
 
-Directory \code{main/examples/deadlock_test/} provides a toy Java
-program on which you can run the deadlock analysis.  First run {\tt
+Directory \code{examples/deadlock_test/} in Chord's main directory provides a toy Java
+program on which one can run the deadlock analysis.  First run {\tt
   ant} in that directory (in order to compile the program's {\tt
   .java} files to {\tt .class} files) and then run the above command
-in Chord's {\tt main/} directory with {\tt <...>} replaced by
+in Chord's main directory with {\tt <...>} replaced by
 \code{examples/deadlock_test/}.  Upon successful completion, the
@@
 \code{examples/deadlock_test/chord_output/}.
-
-\section{May-Alias and Call-Graph Analyses}
-
-TODO

### 2011-03-18 extra/sliver_pldi11/response.txt

We thank all reviewers for their useful comments; we will certainly
incorporate them into future versions of this paper.

REVIEWERS 1 and 4 ask about running time.  The relationship between the number
of input tuples (which we report as the main evaluation metric) and running
time is indeed obscured by using BDDs, but there is a definite correlation.  In
particular, pruning offers modest improvements in performance for small k
values, but the main win is that pruning allows us to even reach higher k
values at all (without pruning, the analysis just runs out of memory before
getting there).

REVIEWER 1: In Table 4, column k corresponds to using a k-limited abstraction.
We will improve the presentation of the results.

REVIEWER 4: The budget of the PR algorithm is determined by the number of
iterations.  In our experiments, we increased k until our analysis ran out
of memory (the horizontal axis of Figure 10 shows the k values which were
reached).

REVIEWER 4: to relate this paper with past work: Indeed, at a high-level, both
our paper and Sridharan-Bodik PLDI'06 refine an abstraction based on a client,
but there are some important differences: (i) their work is specific to pointer
analysis, whereas ours applies generally to any Datalog program; (ii) we also
provide rigorous theoretical guarantees on the precision of our pruning
approach.  As for Liang-Tripp-Naik POPL'11: (i) they focused on finding the
coarsest abstraction to prove a query by searching through the space of
abstractions---this was a scientific study, not a usable algorithm, whereas the
PR algorithm presented in this paper is very practical; (ii) we also are
effectively working with a much richer family of abstractions, where the
context-sensitivity can depend on the specific call chain rather than just the
allocation site.

REVIEWER 2:

To provide more motivation for the particular abstractions we used: First, we
worked with k-limited and class-based abstractions because these are standard
abstractions in pointer analysis which have traditionally faced scalability
issues.  On \alpha_i and \beta_i: usually \beta_i is \alpha_i composed with
another abstraction \tau.  One intuition is that \tau and \alpha_i should be
"complementary"; the effort spent pre-pruning using \tau should neither be
trivial nor redundant with \alpha_i.

Regarding how the correctness of Theorem 2 is related to soundness and
completeness: what we mean is any query we can prove directly using a fine
abstraction we can also prove with pruning (completeness) and vice-versa
(soundness).  When the fine abstraction is no abstraction (concrete), this is
the usual notion of soundness/completeness.

Figure 3 only shows computation of P(X) in the first iteration of the pruning
algorithm.  In this iteration, the red tuples pruned.  In the next iteration,
using a finer abstraction, common(G1,G2,3) will not be reachable, as desired.

REVIEWER 3:

Regarding the comment about system approximation and overapproximation: we do
not view pruning as system approximation but rather as program slicing: one
needs to be careful about what is pruned/sliced out to ensure soundness; for
example, if you prune away all the tuples, you can "prove" anything.

### 2011-03-18 extra/sliver_pldi11/abstract.tex

\begin{abstract}

Many static analyses do not scale as they are made more precise.  For example,
increasing the amount of context sensitivity in a $k$-limited analysis causes
the number of contexts to grow exponentially.  Iterative refinement techniques
can be employed to mitigate this growth by starting with a coarse abstraction and
only refining parts of the abstraction deemed relevant with respect to a client.

In this paper, we introduce a new way to use this client feedback to greatly
reduce the complexity of an analysis: {\em pruning}.  The pruned analysis is no
longer a valid abstraction, but we prove that we do not forfeit correctness and
that our analysis is still sound with respect to a given client.  In the
context of $k$-limited analysis, our approach amounts to keeping track of a set
of context prefix patterns.  By iteratively refining and pruning these
patterns, we show that our analysis increases scalability greatly.

\end{abstract}

### 2011-03-18 extra/sliver_pldi11/abstractions.tex

\Section{abstractionExamples}{Abstractions}

Recall from \refsec{preliminaries} that we separate the Datalog program that
computes the query from the abstraction.  In \refsec{klimited}, we described
the Datalog programs that computed the queries for our three clients based on
$\infty$-object-sensitivity.  We now describe the abstractions that use to
further abstract this analysis.

The basic abstraction (\refsec{klimitedAbstraction}) corresponds exactly to
$k$-object-sensitivity.  We then present two orthogonal variants: one that in
addition limits the repetition of allocation sites (\refsec{barelyAbstraction})
and one that further abstracts allocation sites using type information
(\refsec{classAbstraction}).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\Subsection{klimitedAbstraction}{$k$-limited abstraction}

The {\em $k$-limited abstraction}
essentially partitions chains based on their length $k$ prefix.
We first setup some notation.
%Let $\dom$ be a set of values (e.g., $\dom = \{ 0, 1 \}$).
For a chain $c \in \dom^*$, let $|c|$ denote the length of the chain.
Let $c[i]$ be the $i$-th element of $c$ (starting with index 1) and let
let $c[i..j]$ be the subchain $[c[i] \cdots c[j]]$
(boundary cases: $c[i..j] = [\,]$ if $i > j$ and $c[i..j] = c[i..|c|]$ if $j > |c|$).
For two chains $c_1, c_2$, let $c_1+c_2$ denote concatenation.
For a chain $c$, let $c*$ denote the set of all chains
with prefix $c$; formally:
\begin{align}
c* \eqdef \{ c' \times \H^* : c'[1..|c|] = c \}.
\end{align}
%It is important that $c*$ is a set.

For an integer truncation level $k \ge 0$, define the abstraction $\klimabs_k$ as follows:
\begin{align}
\label{eqn:klimitedAbstraction}
\klimabs_k(c) \eqdef
\begin{cases}
\{c\}     & \text{if $|c| < k$} \\
c[1..k]*  & \text{if $|c| \ge k$}. \\
\end{cases}
\end{align}
If the concrete chain $c$ is shorter than length $k$,
we map it to the singleton set $\{ c \}$;
otherwise, we map it to the set of chains that share the first $k$ elements.
It is easy to verify that $\klimabs_k$ as defined is a valid abstraction under \refdef{abstraction}.
\reffig{repeatingExample} shows an example.

\begin{figure}
\[
\arraycolsep=3.5pt
\begin{array}{llllllll}
\{\vec{0}\}    &            &              &            & \{\vec{1}\} \\
\{\vec{00}\}   &            & \{\vec{01}\} &            & \{\vec{10}\} &            & \{\vec{11}\} \\
\vec{000}*     & \vec{001}* & \vec{010}*   & \vec{011}* & \vec{100}*   & \vec{101}* & \vec{110}* & \vec{111}*
\end{array}
\]
\caption{\label{fig:repeatingExample} For the $k$-limited abstraction
with $\dom = \{ 0, 1 \}$ and $k = 3$, the equivalence classes defined by $\klimabs_k$.
}
\end{figure}

Traditionally, $k$-limited analyses
simply use a list of sites ``$c$'' to represent an abstract value,
not making a distinction between $\{c\}$ and $c*$;
however when we do pruning, this distinction is crucial.
Suppose that $k=3$.
Na\"ively, ``$\vec{01}$'' would just represent $\{\vec{01}\}$.
The problem comes when we prune.
If we prune ``$\vec{010}$'', then ``$\vec{01}$'' really represents $\{\vec{01}\}\cup \vec{010}*$
because any chain starting with $\vec{010}$ must now fall back on ``$\vec{01}$''.
This is unacceptable because pruning tuples should not change the abstraction.
Actually, the situation is worse, because (i) in fact no pruning
(in the sense that some chains are ignored)
would happen and (ii) the resulting analysis would be much more imprecise because
``$[\,]$'' is always be present for representing the context of the main method,
but this is a prefix of any chain.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\Subsection{barelyAbstraction}{Barely-repeating $k$-limited abstraction}

When we applied the $k$-limited abstraction empirically,
we noticed that a major reason why it did not scale was
the seemingly unnecessary combinatorial explosion associated chains formed
by cycling endlessly through the same allocation sites.
For $k$-CFA, this repetition corresponds to recursion.
For $k$-object-sensitivity, this corresponds to recursive allocation, as
illustrated in \reffig{recursion}.\footnote{
Incidentally, the example in the figure also gives an interesting example where
$k$-object-sensitivity for any finite $k$ (no matter how large) is less precise
than $\infty$-object-sensitivity.}
We therefore wish to define an abstraction that not only truncates at $k$ but also truncates
a chain when it starts repeating.

\begin{figure}
\begin{center}
\begin{minipage}{2in}
\begin{verbatim}
class A {
  f() {
0:  v = new A
    if (*) return v
    else return v.f()
  }
}
\end{verbatim}
\end{minipage}
\begin{minipage}{1in}
\begin{verbatim}
1:  x1 = new A
2:  x2 = new A
    y1 = x1.f()
    y2 = x2.f()
\end{verbatim}
\end{minipage}
\end{center}
\caption{\label{fig:recursion} An example illustrating the repetition of allocation sites.
The points-to set of {\tt y1} using $\infty$-object-sensitivity
is $\{\vec{01},\vec{001},\vec{0001},\dots\}$ (any positive number of zeros followed by a 1),
and the points-to set of {\tt y2} is $\{\vec{02},\vec{002},\vec{0002},\dots\}$.
These two sets are disjoint, but for any finite $k$, no matter how large,
$k$-object-sensitivity would conclude that both point to $\bzero_k*$, where $\bzero_k$ is a chain of $k$ zeros.
}
\end{figure}

For a sequence $c$, we say $c$ is {\em non-repeating} if all its elements are distinct.
We say $c$ is {\em barely-repeating} if $c$ excluding the last element ($c[1..|c|-1]$)
is {\em non-repeating}.  Note that a barely-repeating includes the non-repeating case.
Let $\delta(c)$ be the length of the longest prefix that is barely-repeating:
\begin{align}
\label{eqn:barelyLen}
\distinct(c) \eqdef \max_{m' : c[1..m'] \text{ is barely-repeating}} m'.
\end{align}
For example, $\distinct(\vec{10010}) = 3$ because $\vec{100}$ is barely-repeating, but $\vec{1001}$ is not.

Then we define the {\em barely-repeating $k$-limited abstraction} $\klimdabs_k$ as follows:
\begin{align}
\label{eqn:barelyAbstraction}
\klimdabs_k(c) & \eqdef \klimabs_{\min\{k,\delta(c)\}}(c),
\end{align}
\reffig{barelyExample} shows an example of $\klimdabs_k$.
It is not automatic that $\klimabs_k$ is a valid abstraction,
but we can check that it is:
\begin{proposition}
\label{prop:barelyAbstraction}
The function $\klimdabs_k$ defined in \refeqn{barelyAbstraction} is a valid abstraction
(\refdef{abstraction}).
\end{proposition}
\begin{proof}
We consider two cases: (i) for $\{c\} \in \range(\klimdabs_k)$,
we have $\klimdabs_k(c) = \{c\}$; and (ii) for any $c* \in \range(\klimdabs_k)$,
either $|c| = k$ or $c$ is barely-repeating; in either case,
we can check that any extension $c' \in c*$ will have $\klimabs_k(c') = c*$.
\end{proof}
Remark: one might wonder why we defined the abstraction using the barely-repeating criterion
as opposed to the simpler non-repeating criterion.
It turns out that using the latter in \refeqn{barelyLen} would not result in a valid abstraction.
If $\klimdabs_k$ were defined using the non-repeating criterion,
then $\klimdabs_3(\vec{00}) = \vec{0}*$.
But for $\vec{01} \in \vec{0}*$, we have $\klimdabs_3(\vec{01}) = \{\vec{01}\} \neq \vec{0}*$.

\begin{figure}
\[
\begin{array}{lllllll}
\{\vec{0}\}    &            &              &            & \{\vec{1}\} \\
\vec{00}*      &            & \{\vec{01}\} &            & \{\vec{10}\} &            & \vec{11}* \\
               &            & \vec{010}*   & \vec{011}* & \vec{100}*   & \vec{101}* &
\end{array}
\]
\caption{\label{fig:barelyExample} For the barely-repeating $k$-limited abstraction
for $\dom = \{0,1\}$ and $k = 3$, we show the equivalence classes under $\klimdabs_k$.
Compare this with the classes for the $k$-limited abstraction (\reffig{repeatingExample}).
Note that, for example,
$\vec{000}*$ and $\vec{001}*$ are collapsed into $\vec{00}*$ since $\vec{000}$ and $\vec{001}$ are not barely-repeating,
but $\vec{00}$ is.
}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\Subsection{classAbstraction}{Class-based abstraction}

We now introduce an abstraction that we will use in the pre-pruning step of the \PR\ algorithm.
We start by defining an equivalence relation over allocation sites $\dom$,
represented by a function $\hclass : \dom \mapsto \sP(\dom)$
mapping each allocation site $h \in \dom$ to its its equivalence class.
Given such such an $\hclass$, we could then extend it sequences by elementwise application:
\begin{align}
\hclass(c) = \hclass(c[1]) \times \cdots \times \hclass(c[|c|]), \quad c \in \dom^*.
\end{align}

To construct $\hclass$, we consider two sources of type information associated
with an allocation site, motivated by \cite{smaragdakis11context}:
\begin{align}
\typeIs(h)  &= \text{declaring type of allocation site $h$} \label{eqn:typeIs} \\
\typeHas(h) &= \text{type of class containing allocation site $h$} \label{eqn:typeHas}
\end{align}
Using these two functions, we can construct three equivalence relations,
$\hclassIs$, $\hclassHas$, and $\hclassIsHas$ as follows:
\begin{align}
\hclass_f(h) = \{ h' : f(h) = f(h') \}, \label{eqn:hclass}
\end{align}
with $f \in \{ \typeIs, \typeHas, \typeIs \times \typeHas \}$.

% Complementary
Now we have three options for $\hclass$ which are complementary to the
$k$-limited abstraction: $\hclass$ abstracts a chain by abstracting each site
uniformly; $\pi_k$ keeps all sites up to a length $k$ prefix precisely but
abstracts away the rest completely.  This complementarity is desirable
for pre-pruning.
However, $\hclass$ is not coarser than $\klimabs_k$, and therefore cannot
be used directly in the \PR\ algorithm.

However, we can compose $\hclass$ and $\klimabs_k$ to yield another abstraction which will be coarser than $\klimabs_k$.
But in what order should we compose?
We must take care because the composition of two equivalence relations is not necessarily an equivalence relation.
The following proposition shows which orderings are valid:
\begin{proposition}
The functions
(i) $\klimabs_k \circ \hclass$ and (ii) $\hclass \circ \klimabs_k$
are valid abstractions (see \refdef{abstraction}) and equivalent;
(iii) $\klimdabs_k \circ \hclass$ is also valid,
but (iv) $\hclass \circ \klimdabs_k$ is not.
\end{proposition}
\begin{proof}
For each of these four composed functions, each set $s$ in the range of the function
must be either of the form $s = w_1 \times \cdots \times w_m$ for $m < k$ (case 1)
or $s = w_1 \times \cdots \times w_m \times \dom^*$ for some $m \le k$ (case 2),
where $w_i \in \range(\hclass)$ for each $i = 1, \dots, m$.

For (i) and (ii), it is straightforward to check
$(\klimabs_k \circ \hclass)(c) = (\hclass \circ \klimabs_k)(c) = s$ for each $c \in s$.
Intuitively, the truncation ($\klimabs_k$) and coarsening ($\tau$) operate
independently and can be interchanged.

For (iii) and (iv), the two dimensions do not act independently;
the amount of truncation depends on the amount of coarsening:
the coarser $\hclass$ is, the more truncation one might need to limit repetitions.
Showing that $\klimdabs_k \circ \hclass$ is valid
proceeds in a similar manner to \refprop{barelyAbstraction}.
If $s$ falls under case 1, note that no $c \in s$ is repeating
because the $w_i$'s must be disjoint; therefore $\klimdabs_k(\tau(c)) = s$.
If $s$ falls under case 2, note that for any $c[1..m] \in s$ must be barely-repeating but any longer prefix is not,
and therefore, $\klimdabs_k(\tau(c)) = s$.

To show that (iv) is not an abstraction,
consider the following counterexample:
let $\dom = \{ 0,1,2 \}$,
and define $\hclass(v) = \dom$ for all $v \in \dom$
(there is one equivalence class).
Consider two elements $\vec{01}$ and $\vec{00}$ under the composed abstraction:
For $\vec{01}$, we have $\klimdabs_3(\vec{01}) = \{ \vec{01 }\}$,
so $\hclass(\klimdabs_3(\vec{01})) = \dom^2$;
for $\vec{00}$, we have $\klimdabs_3(\vec{00}) = \vec{00}*$,
so $\hclass(\klimdabs_3(\vec{00})) = \dom^2 \times \dom^*$.
But $\dom^2 \subsetneq \dom^2 \times \dom^*$ (notably,
the two sets are neither equal nor disjoint),
so $\hclass \circ \klimdabs_3$ does not define an equivalence relation.
\end{proof}

In light of this result, we will use the valid abstractions $\cklimabsk$ and $\cklimdabsk$,
which work by first applying $\hclass$ and then applying $\klimabs_k$ or $\klimdabs_k$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\Subsection{computation}{Computation}
%
%In order to use the PR algorithm (\reffig{pseudocode}),
%we need to implement the various steps efficiently.
%
%We now present techniques for computing 
%
%\paragraph{Pre-pruning step}
%
%Computing $\cklimdabsk$
%
%\paragraph{Pruning step}
%
%However, need to compute $\ext$.
%
%\paragraph{Refinement step}
%
%We use $\alpha_t = \klimdabs_{t+1}$
%and $\beta_t = \klimdabs_{t+1}$
%and we set $\hclass(h)$ to be the dynamic type of allocation site $h$.
%
%Computing $\klimdabs_{k+1}$
%
%Suppose we have we have a current $A_k \subset \range(\klimdabs_k)$.
%For $c \in \dom^*$.
%$\alpha_{k+1}(\{c\}) = $
%Given $c$, we consider

### 2011-03-18 extra/sliver_pldi11/conclusion.tex

\Section{conclusion}{Conclusion}

We have introduced pruning as a general way of scaling up static analyses
written as Datalog programs.  The key idea is to run an analysis using a coarse
abstraction, only keeping input tuples deemed relevant for a finer abstraction.
Theoretically, we showed that pruning is both sound and complete (our analysis
is valid and we lose no precision).  Empirically, we showed that our pruning
approach enables us to scale up analyses based on $k$-object-sensitivity much
more than previous approaches.

### 2011-03-18 extra/sliver_pldi11/experiments.tex

\Section{experiments}{Experiments}

In this section, we apply our \PR\ algorithm (\refsec{algorithm}) to
$k$-object-sensitivity for three clients:
downcast safety checking (\downcast),
monomorphic call site inference (\monosite),
and race detection (\race).  These analyses are described in \refsec{klimited}.
Our main empirical result is that across different clients and benchmarks,
pruning is very effective at curbing the exponential growth,
which allows us to run the analysis on very refined abstractions.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\Subsection{setup}{Setup}

% Benchmark/setup
Our experiments were performed using IBM J9VM 1.6.0 on 64-bit Linux machines.
All analyses were implemented in Chord,
an extensible program analysis framework for Java bytecode,\footnote{{\tt{http://code.google.com/p/jchord/}}}
which uses the BDD solver {\tt bddbddb} \cite{Whaley2007}.
We evaluated on \numBenchmarks\ Java benchmarks shown in \reftab{benchmarks}.
In each run, we allocated 8GB of memory and terminated the process when it ran out of memory.

\begin{table*}
\centering
\input figures/stats.tex
\caption{Benchmark characteristics:
the number of classes,
number of methods,
total number of bytecodes in these methods,
and number of allocation sites ($|\H|$)
deemed reachable by 0-CFA.
\label{tab:benchmarks}}
\end{table*}

% Algorithms
We experimented with various combinations of abstractions and refinement
algorithms (see \reftab{algorithms}).  As a baseline, we consider running an analysis
with a full abstraction $\alpha$ (denoted $\Full(\alpha)$).
For $\alpha$, we can either use $k$-limited abstractions
($(\klimabs_k)_{k=1}^\infty$), in which case we recover ordinary $k$-object-sensitivity,
or the barely-repeating variants ($(\klimdabs_k)_{k=1}^\infty$).
We also consider the site-based refinement algorithm of \cite{liang11minimal},
which considers a sequence of abstractions $\balpha = (\balpha_0, \balpha_1, \dots)$
but stops refining sites which have been deemed irrelevant (this algorithm is denoted $\Site(\balpha)$).

As for the new methods that we propose in this paper,
we have $\PR(\balpha)$, which uses a sequence of abstractions $\balpha$, which does no pre-pruning,
or $\PR(\balpha,\hclass)$, which performs pre-pruning using $\beta_t = \alpha_t
\circ \hclass$ for $t = 0, 1, 2, \dots$.  We consider three choices of $\hclass$
($\hclassIs,\hclassHas,\hclassIsHas$) which use different kinds of type information.

In our implementation of the \PR\ algorithm, we depart slightly from our
presentation.  Instead of maintaining the set of relevant input tuples as
determined by our $k$-limited Datalog program across iterations, we instead
maintain the allocation site chains which were involved in any relevant input
tuple.  We can modify the original Datalog program so that the \PR\ algorithm
is consistent with this implementation by introducing a new input relation
$\dl{active}(c)$ for $c \in \H^*$, encoding existing input relations as rules
consisting of zero source terms, and adding $\dl{active}(c)$ for each existing
rule that uses a variable $c$ which is an allocation site.  Now computing the
relevant input tuples corresponds exactly to computing the set of relevant
allocation site chains.

\begin{table}
\[
\begin{array}{ll}
\text{\bf Abstractions} \\
\Klimabs = (\klimabs_k)_{k=1}^\infty   & \defn{$k$-limited abstractions \refeqn{klimitedAbstraction}} \\
\Klimdabs = (\klimdabs_k)_{k=1}^\infty & \defn{barely-repeating $k$-limited abstractions \refeqn{barelyAbstraction}} \\
\hclassIs                              & \defn{abstraction using type of allocation site} \\
\hclassHas                             & \defn{abstraction using type of containing class} \\
\hclassIsHas                           & \defn{abstraction using both types} \\
\\
\text{\bf Algorithms} \\
\Full(\alpha)         & \defn{standard analysis using abstraction $\alpha$} \\
\Site(\balpha)        & \defn{site-based refinement \cite{liang11minimal} on abstractions $\balpha$} \\
\PR(\balpha)          & \defn{\PR\ algorithm using $\balpha$, no pre-pruning} \\
\PR(\balpha,\hclass)  & \defn{\PR\ algorithm using $\balpha$, using $\balpha\circ\hclass$ to pre-prune} \\
\end{array}
\]
\caption{\label{tab:algorithms} Shows the abstractions and algorithms that we evaluated empirically.
For example,
$\PR(\Klimdabs,\hclassIsHas)$
means running the \PR-algorithm on the barely-repeating $k$-limited abstraction ($\alpha_k = \klimdabs_k$),
using a composed abstraction based on the type of an allocation site ($\hclassIs$) 
and the type of the declaring class ($\hclassHas$) to do pre-pruning
(specifically, $\beta_k = \klimdabs_k \circ \hclassIsHas$.
}
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\Subsection{results}{Results}

\reffig{absSizeGraph} compares the number of input tuples needed by various
algorithms; we see that the non-pruning methods completely hit a wall, whereas
the pruning methods are able to continue increasing $k$ farther.  We also
observed that the running time of these analyses tracked the number of tuples
somewhat, so we do obtain speedups, but the differences are less pronounced.
One reason is that we are using BDDs, which can handle large numbers of tuples
so long as they are structured; pruning destroys some of this structure,
yielding less predictable running times.

\FigStar{figures/absSizeGraph}{0.35}{absSizeGraph}{For each client/benchmark pair,
we show the growth of the number of input tuples $|A_t'|$ across iterations.
Recall that $|A_t'|$ is the number of tuples fed into the Datalog program.
\reftab{algorithms} describes the four methods.
We see that the methods that prune drastically cut down the number of input tuples
by many orders of magnitude.
}

\reftab{sizeRatio} provides more on the quantitative impact of pruning in our
best instantiation of the \PR\ algorithm.  We see that pre-pruning has a
significant positive impact: we are able to perform pruning at a much cheaper
level (using types instead of allocation sites).  Many of the tuples are pruned
at this level, and the results of this pruning carry over to the original
$k$-limited abstraction.

\begin{table}
\centering
\small
\input figures/sizeRatio
\caption{\label{tab:sizeRatio} Shows the shrinking and growth of the
number of tuples during the various pruning and refinement operations (see \reffig{pseudocode}) for 
our best method $\PR(\Klimabs,\hclassIsHas)$ across all the clients and benchmarks, averaged across iterations.
The columns are as follows:
First, $\frac{|B_t|}{|A_t|}$ measures the number of tuples after projecting
down to the auxiliary abstraction $\beta_t$ for pre-pruning; note that running the analysis
using types instead of allocation sites is much cheaper.
Next, $\frac{|\tilde B_t|}{|B_t|}$ shows the fraction of abstract values kept during pre-pruning;
When we return from types to allocation sites, see that the effect of pre-pruning essentially carries over
($\frac{|A_t'|}{|A_t|}$).
Recall that we do the pruning step after pre-pruning;
$\frac{|\tilde A_t|}{|A_t|}$ measures the total fraction of chains kept during pruning.
Finally, $\frac{|A_{t+1}|}{|A_t|}$ measures the ratio between iterations,
which includes both pruning and refinement.  Note that there is still almost a three-fold
growth of the number of tuples (on average), but this growth would have been much more unmanageable
without the pruning.
}
\end{table}

So far, we have been using the $k$-limited abstraction.  In
\reffig{barelyRepeatingEffect}, we show that when we can increase the $k$
value, the barely-repeating $k$-limited abstraction is helpful, but in most
cases, it does not improve scalability.  The reason is that the
barely-repeating abstraction curbs refinement, but often, and quite
paradoxically, it is exactly refinement which enables us to prune more.

\FigStar{figures/barelyRepeatingEffect}{0.35}{barelyRepeatingEffect}{
Shows the 5 (out of 15) client/benchmark pairs for which using the
barely-repeating $k$-limited abstraction ($\Klimdabs$) allows one to increase
$k$ much more than the plain $k$-limited abstraction ($\Klimabs$).
On the other 10 client/benchmarks where $k$ cannot get very large,
limiting repetitions is actually slightly worse in terms of scalability.
}

Finally, \reftab{numUnproven} shows the effect on the number of queries proven.
While pruning enables us to increase $k$ much more than before, it turns out
that our particular analyses for these clients saturate quite quickly, so we
are only able to prove two more queries than using the non-pruning techniques.
On the surface, these findings seem to contradict \cite{liang10abstraction},
which showed a sharp increase in precision around $k=4$ for $k$-CFA.  However,
this discrepancy merely suggests that our flow-insensitive analyses are simply
limited: since \cite{liang10abstraction} obtains a lower bound, we know for
sure that low $k$ values are insufficient; the fact that we don't see an
increase in precision suggests that the non-$k$-related aspects of our analyses
are insufficient.  However, given that our pruning approach is general, it
would be interesting tackling other aspects such as flow-sensitivity.

\begin{table}
\small
\centering
\input figures/numUnproven
\caption{\label{tab:numUnproven} The number of unproven queries (unsafe
downcasts, polymorphic sites, races) for each of the clients and benchmarks
over the first five iterations.
All analyses obtain the exact results on iterations where they obtain an answer.
Bolded numbers are ones obtained by our best method $\PR(\Klimabs,\hclassIsHas)$
but not by the full algorithm.
We see that although pruning enables to increase $k$ by much more than before,
only for two the client/benchmark pairs do we get strictly more precise results.
This points out inherent limitations of this family of $k$-limited abstractions.
}
\end{table}

### 2011-03-18 extra/sliver_pldi11/introduction.tex

\Section{introduction}{Introduction}

% Problem: want precision, abstractions don't scale up, even with
% client-driven, etc.
Making a static analysis more precise requires increasing the complexity of the
supporting abstraction---in pointer analysis, by increasing the amount of
context/object sensitivity \cite{kcfa, kobj, MilanovaRountevRyder2005,
WhaleyLam2004, LhotakHendren2006, LhotakHendren2008}; or in model checking, by
adding more abstraction predicates \cite{graf97predicate,slam}.
However, the complexity of these analyses often grows exponentially as the
abstraction increases.
Much work has been done on curbing this exponential growth
(e.g., client-driven \cite{GuyerLin2003} and demand-driven
\cite{HeintzeTardieu2001} approaches in pointer analysis,
lazy abstraction \cite{henzinger02lazy,mcmillan06lazy} and other iterative refinement approaches in model checking).
The general theme of all these approaches is that the
complexity of the abstraction is increased only in parts deemed useful using
some feedback from a client query.

% Pruning idea: general
In this paper, we use {\em pruning}, a different approach from
the above selected refinement techniques for making static analyses scale.
Our approach works on any analysis that can be expressed as a Datalog program.
A Datalog program takes a set of input tuples (determined by the abstraction)
and derives new tuples via a set of inference rules; the answer to a client
query corresponds to whether a designated tuple can be derived.  The key idea
is to identify input tuples which are provably irrelevant for deriving the
query tuple and prune these away.  Selected refinement techniques
attempt to keep the set of input tuples small by not refining some of them; we
keep the set small by removing some of them entirely.

% Soundness
Pruning can be a dangerous affair though.  With selected refinement, at any
point we are still performing static analysis with respect to an abstraction
and therefore have soundness guarantees.  However, once we start pruning input
tuples, we are no longer running a valid static analysis and thus do not
automatically inherit soundness guarantees.  Nonetheless, we prove that our
method is {\em sound with respect to the client}.

% Completeness
While soundness is trivial for selected refinement but requires some argument for pruning,
the situation is reversed for {\em completeness}, that is, the guarantee that
an analysis is as precise as some target.  For selected refinement,
it is difficult to argue that the chosen coarser abstraction is as precise as if we
had refined everything.  However, with pruning, we are working directly with
the more complex abstraction, and thus automatically inherit its completeness
guarantees.  By removing input tuples, we can only prove more queries.

% Refinement algorithm
%To capitalize on the idea of {\em pruning}, we propose an algorithm, which we
%call Refinement-Pruning (\PR).  The idea is to start with a coarse abstraction
%and prune the input tuples before moving to a more expensive abstraction.

% Slivers
We apply our pruning technique to the $k$-object-sensitivity abstraction
\cite{kobj}, where objects on the heap are abstracted into a chain of
allocation sites.  Pruning in this setting corresponds to maintaining a subset
of these chains.  Although the main contribution of the paper is the general
pruning technique, we also make the following contributions specific to
$k$-limited pointer analysis which arise because we are pruning: First, we show
that we need a more careful treatment of the $k$-object-sensitivity abstraction
than before due to pruning.  Second, we introduce an abstraction which limits
chains to length $k$ and also truncates to avoid having too many repeating
sites; this is an effective way to increase $k$ further (possible now due to
pruning) without getting bogged down by recursion.  Third, we show that an
abstraction that replaces allocation sites by types (a generalization of
\cite{smaragdakis11context}) is effective for pruning.

% Experiments
We ran our experiments on \numBenchmarks\ Java benchmarks using three clients that
depend heavily on having an accurate pointer analysis: downcast safety
checking, monomorphic call site inference, and race detection.
We show that with pruning, our \PR\ algorithm enables us to perform $k$-object-sensitivity
analysis with a substantially much finer abstraction (larger $k$)
compared to a full $k$-object-sensitive analysis or even using the selected
refinement strategy of \cite{liang11minimal}.
In some cases, the non-pruning approaches hit a wall around $k=3$ but the \PR\
algorithm is able to go well beyond $k=10$.

### 2011-03-18 extra/sliver_pldi11/klimited.tex

\Section{klimited}{$k$-limited Pointer Analysis}

% Basics
We now apply our general pruning approach to $k$-limited pointer analysis.
Each node in the control-flow graph of each method $m \in \M$ is associated with a
simple statement (e.g., $v_2 = v_1$).
We omit statements that have no effect on our analysis (e.g., operations on
data of primitive type).
For simplicity, we assume each method has a single argument and no return value.\footnote{
Our actual implementation is a straightforward extension of this simplified analysis
which handles multiple arguments, return values, class initializers, and
objects allocated through reflection.}
\reffig{aliasAnalysisRules} describes the Datalog program corresponding to this
analysis.

% Context, abstraction
Our analysis represents both contexts and abstract objects using a list of
allocation sites, which we will call {\em chains}.  However, note that these
chains are never truncated in the Datalog program.  Therefore, running the
Datalog program (ignoring the fact that it might not terminate) is performing
an $\infty$-object-sensitivity analysis.  Recall that we separate the
computation of the query in Datalog with the abstraction which is applied to
the input tuples (of course, our Datalog program is itself an approximation to
the concrete semantics---it is flow-insensitive, does not handle primitive
data, etc.).  \refsec{abstractions} will describe further abstractions of this
analysis.

Our analysis computes the reachable methods ($\reachM$), reachable statements ($\reachP$), and
points-to sets of local variables ($\ptsV$), each with the associated context;
the context-insensitive points-to sets of static fields ($\ptsG$) and
heap graph ($\heap$); and a context-sensitive call graph ($\cg$).

We briefly describe the analysis rules in Datalog.  Rule (1)
states that the main method $\mmain$ is reachable in a distinguished context $[]$.
Rule (2) states that a target method of a reachable call site is also reachable.
Rule (3) states that every statement in a reachable method is also reachable.
Rules (4) through (9) implement the transfer function associated with each kind of
statement.
%Rule (10) populate the call graph and
%rule (11) propagate the points-to set from the argument
%of a call site to the formal argument of each target method.
Rule (10) analyzes the target method
$m$ in a separate context $o$ for each abstract object $o$ to which the distinguished
{\tt this} argument of method $m$ points, and rule (11) sets the points-to set
of the {\tt this} argument of method $m$ in context $o$ to the singleton $\{ o \}$.

\begin{figure*}
\begin{minipage}{3.3in}
{\bf Domains}: \\
\[
\begin{array}{rrcl}
\mbox{{(method)}}                  & m   &\in & \M = \{ \mmain, ... \}\\
\mbox{{(local variable)}}          & v   &\in & \V \\
\mbox{{(global variable)}}         & g   &\in & \G  \\
\mbox{{(object field)}}            & f   &\in & \F  \\
\mbox{{(method call site)}}        & i   &\in & \I \\
\mbox{{(allocation site)}}         & h   &\in & \H \\
\mbox{{(statement)}}               & p   &\in & \P \\
\mbox{{(method context)}}          & c   &\in& \C = \H^* \\
\mbox{{(abstract object)}}         & o   &\in& \O = \H^* \\
\end{array}
\]
\[
\begin{array}{ll}
p ::= & v = \new\,h \bigmid v_2 = v_1 \bigmid g = v \bigmid v = g \bigmid \\
      & v_2.f = v_1 \bigmid v_2 = v_1.f  \bigmid i(v)
\end{array}
\]
\end{minipage}
\begin{minipage}{3.7in}
{\bf Input relations}: \\
\[ \begin{array}{lcll}
\body     &\subset& \M \times \P & \text{(method contains statement)} \\
\trgt     &\subset& \I \times \M & \text{(call site resolves to method)} \\
\argI     &\subset& \I \times \V & \text{(call site's argument variable)} \\
\argM     &\subset& \M \times \V & \text{(method's formal argument variable)} \\
\ext      &\subset& \S \times \C \times \C & \text{(extend context with site)} \\
          & =     & \multicolumn{2}{l}{\{ (h, c, [h]+c) : h\in\H, c\in\C \}} \\
\end{array}
\]
{\bf Output relations}: \\
\[ \begin{array}{lcll}
\reachM   &\subset& \C \times \M & \text{(reachable methods)} \\
\reachP   &\subset& \C \times \P & \text{(reachable statements)} \\
\ptsV     &\subset& \C \times \V \times \O & \text{(points-to sets of local variables)} \\
\ptsG     &\subset& \G \times \O & \text{(points-to sets of static fields)} \\
\heap     &\subset& \O \times \F \times \O & \text{(heap graph)} \\
\cg       &\subset& \C \times \I \times \C \times \M & \text{(call graph)}
\end{array}
\]
\end{minipage}
{\bf Rules}: \\
\[
\begin{array}{lclr}
\reachM([],\mmain). &           &  & (1) \\
\reachM(c,m)              &\dlogUpdate& \cg(*,*,c,m). & (2) \\
\reachP(c,p)              &\dlogUpdate& \reachM(c,m),\ \body(m,p). & (3)
\\[8pt]
\ptsV(c,v,o)              &\dlogUpdate& \reachP(c, v = \new\,h),\ \ext(h, c, o). & (4) \\
\ptsV(c,v_2,o)            &\dlogUpdate& \reachP(c, v_2 = v_1),\ \ptsV(c,v_1,o). & (5) \\
\ptsG(g,o)                &\dlogUpdate& \reachP(c, g = v),\ \ptsV(c,v,o). & (6) \\
\ptsV(c,v,o)              &\dlogUpdate& \reachP(c, v = g),\ \ptsG(g,o). & (7) \\
\heap(o_2,f,o_1)          &\dlogUpdate& \reachP(c, v_2.f = v_1),\ \ptsV(c,v_1,o_1),\ \ptsV(c,v_2,o_2). & (8) \\
\ptsV(c,v_2,o_2)          &\dlogUpdate& \reachP(c, v_2 = v_1.f),\ \ptsV(c,v_1,o_1),\ \heap(o_1,f,o_2). & (9)
\\[8pt]
\cg(c,i,o,m)            &\dlogUpdate& \reachP(c,i),\ \trgt(i,m),\ \argI(i,v),\ \ptsV(c,v,o). & (10) \\
\ptsV(c,v,c)             &\dlogUpdate& \reachM(c,m),\ \argM(m,v). & (11)
\end{array}
\]
%\centering
\caption{Datalog implementation of our $k$-object-sensitivity points-to analysis with call-graph construction.
Our abstraction $\ba$ affects the analysis solely through $\ext$,
which specifies that when we prepend $s$ to $c$, we truncate the resulting sequence to length $\ba_{s}$.
}
\label{fig:aliasAnalysisRules}
\end{figure*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\Subsection{clients}{Clients}

The core pointer analysis is used by three clients, which each define a set of queries.

\paragraph{Monomorphic call site detection}

Monomorphic call sites are dynamically dispatched call sites with at most one
target method.  These can be transformed into statically dispatched ones which
are cheaper to execute.
For each call site $i \in \I$ whose target is a virtual method,
we create a query $\dl{poly}(i)$ asking whether $i$ is polymorphic.
This query can be computed with the following rule:
\begin{align}
\dl{poly}(i) \dlogUpdate \cg(*,i,*,m_1), \cg(*,i,*,m_2), m_1 \neq m_2.
\end{align}

\paragraph{Downcast safety checking}

A safe downcast is one that cannot fail because the object to which the
downcast is applied is guaranteed to be a subtype of the target type.
Therefore, safe downcasts obviate the need for run-time cast checking.
We create a query for each downcast---statement of the form $v_1 = v_2$ where the
declared type of $v_2$ is not a subtype of the declared type of $v_1$.
The query can be computed with the following rule:
\begin{align}
& \dl{unsafe}(v_1,v_2) \dlogUpdate \ptsV(*,v_2,o), \dl{typeO}(o,t_2), \dl{typeV}(v_1,t_1), \nonumber \\
& \hspace{1in} \neg\dl{subtype}(t_1,t_2).
\end{align}
Here, $\dl{typeV}$ is a relation on a variable and its declared type
and $\dl{typeO}$ is a relation on an abstract object and its type
(computed by inspecting the initial allocation site of $o$).

\paragraph{Race detection}

In race detection, each query consists of a pair of heap-accessing statements
of the same field in which at least one statement is a write.  We used the
static race detector of \cite{NaikAikenWhaley2006}, which declares a
$(p_1,p_2)$ pair as racing if both statements may be reachable, may access
thread-escaping data, may point to the same object, and may happen in parallel.
All four components rely heavily on the context- and object-sensitive points-to
analysis.

### 2011-03-18 extra/sliver_pldi11/preliminaries.tex

\Section{preliminaries}{Preliminaries}

The pruning approach that we present in this paper works on Datalog, a general
language that has been used recently to declaratively express static analyses \cite{Whaley2007,bravenboer}.  Normally, one thinks of
the Datalog program as encapsulating the abstraction, but it will be useful for us
to decouple the two.\footnote{The reason is that we want to make theoretical
statements comparing the same Datalog program across different
abstractions.}  We first define Datalog (\refsec{datalog}), which describes the
computation of a query (think more in the direction of concrete semantics).
Then, we focus on the abstraction (\refsec{abstractions}), which only interacts
with the Datalog program via the input tuples.
Throughout this section, we will use \reffig{graphExample} as a running example.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\begin{center} {\bf Graph Example} \end{center}
{\bf Input relations}:
\[
\begin{array}{ll}
\edge(g,i,j) & \defn{edge from node $i$ to node $j$ in graph $g$} \\
\head(c,i)   & \defn{first element of array $c$ is $i$} \\
\ext(i,c,c') & \defn{$i$ prepended to $c$ yields $c'$: $c' = [i]+c$} \\
\end{array}
\]
{\bf Rules}:
\[
\begin{array}{lcl}
\path(g,[\initNode]). & \\
\path(g,c')           &\dlogUpdate& \path(g,c), \head(c,i), \\
                      &           & \edge(g, i, j), \ext(j, c, c'). \\
\common(g_1,g_2,i)    &\dlogUpdate& \path(g_1,c), \path(g_2,c), \head(c,i). \\
\end{array}
\]
{\bf Query tuple}: $\xo = \common(\Ga,\Gb,3)$. \\
\\
{\bf Constants}: $\sC = \{ \Ga, \Gb, 0, 1, 2, 3, \vec{0}, \vek{01}, \dots \}$.
\\
\\
\includegraphics[scale=0.5]{figures/graphExample}
\caption{\label{fig:graphExample}
A simple example illustrating Datalog:
Suppose we have two graphs $\Ga$ and $\Gb$ defined on the same set of nodes $\{0,1,2,3\}$,
and we want to compute the query tuple $\common(\Ga,\Gb,3)$,
asking whether the two graphs have a common path from node $0$ to node $3$.
Given the input tuples encoding the graph,
the Datalog program computes a set of derived tuples from the rules.
In this case, the absence of $\common(\Ga,\Gb,3)$ means the query is false.
}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\Subsection{datalog}{Datalog}

A {\em Datalog program} consists of
a set of {\em constants} $\sC$ (e.g., $\initNode,\vec{03} \in \sC$),
a set of {\em variables} $\sV$ (e.g., $i, j \in \sV$), and
a set of {\em relations} $\sR$ (e.g., $\edge \in \sR$).

A {\em term} $t$ consists of a relation $t.r \in \sR$ and a list of arguments $t.\ba$,
where each argument $t.a_i$ is either a variable or a constant,
(that is, $t.a_i \in \sV \cup \sC$) for $i = 1, \dots, |t.\ba|$.
We will write a term in any of the following three equivalent ways:
\begin{align}
t \quad\equiv\quad t.r(t.\ba) \quad\equiv\quad t.r(t.a_1, \dots, t.a_{|t.\ba|}).
\end{align}
For example, $\ext(j,c,c')$ is a term.
We call a term whose arguments are all constants a {\em tuple}
(e.g., $\ext(\initNode, [\,], \vec{\initNode})$).
Note that the tuple includes the relation as well as the arguments.
We let $\xo$ denote a designated {\em query tuple} (e.g., $\common()$),
whose truth value we want to determine.

Let $\sZ$ denote the set of rules,
where each {\em rule} $z \in \sZ$ consists of a target term $z.t$ and a set of source terms $z.\bs$.
We write a rule if $z.t = t$ and $z.\bs = \{ s_1, \dots, s_k \}$ as
\begin{align}
t \dlogUpdate s_1, \dots, s_n.
\end{align}

An {\em assignment} is a function $f : \sV \mapsto \sC$ which maps variables to
constants.  To simplify notation later, we extend an assignment $f$ so that it
can be applied (i) to constants ($f(c) = c$ for $c \in \sC$), and (ii) to terms by
replacing the variables in the term with constants ($f(t) = t.r(f(t.a_1), \dots, f(t.a_{|t.\ba|}))$).

\begin{figure}[ht]
%{\bf Notation:} \\
\[
\begin{array}{ll}
\xo                           & \defn{designated query tuple} \\
\sC                           & \defn{set of concrete values} \\
\bD(X)                        & \defn{derivations using input tuples $X$} \\
\bP(X) \subset X              & \defn{set of relevant input tuples} \\
\alpha : \sC \mapsto \sP(\sC) & \defn{abstraction, maps to equivalence class} \\
A_k \subset \sP(\sC)          & \defn{abstract input tuples after $k$ iterations} \\
\tilde A_k = \bP(A_k)         & \defn{relevant abstract input tuples} \\
\end{array}
\]
\caption{\label{fig:notation} Notation.}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\Subsection{derivations}{Derivations}
\paragraph{Derivations}

Intuitively, a Datalog program takes a set of input tuples and derives new
tuples.  To formalize this computation, we define the notation of a derivation.

A {\em derivation} (of the query $\xo$) with respect to a set of input tuples
$X$ is a sequence $\bx = (x_1, \dots, x_n)$ such that
\begin{enumerate}
\item [(i)] for each $i = 1, \dots, n$, we have $x_i \in X$; or
there exists a set of indices $J$ such that $(J,i)$ satisfies the following conditions:
$j < i$ for each $j \in J$, and
there is a rule $z \in \sZ$ and an assignment $f$
such that $f(z.t) = x_i$ and
$\{ x_j : j \in J \} = \{ f(s) : s \in z.\bs \}$.
\item [(ii)] $x_n = \xo$; and
\item [(iii)] 
for each $j = 1, \dots, n\!-\!1$,
there exists $J$ such that $j \in J$ and an index $i$
such that $(J,i)$ satisfies the conditions in (i).
\end{enumerate}
Define $\bD(X)$ to be the set of all derivations with respect to the input tuples $X$.

% Explain
Condition (i) says that each tuple in a derivation
should either be given as an input tuple ($x_i \in X$)
or be the result of some rule $z \in \sZ$.
Condition (ii) says that $\xo$ was finally produced via some sequence of rule applications.
Condition (iii) says that the derivation of $x_n$ is minimal; each tuple was
used in some way; as we will see later, this condition is important because it allow us to
define the set of relevant tuples for pruning.

A Datalog program computes $\bD(X)$.
We say that the query $\xo$ is false (proven) if and only if $\bD(X) = \emptyset$.
Although the query is the ultimate quantity of interest,
the Datalog program can be used to provide more information that is useful for pruning.
Specifically, we define $\bP(X)$ to be the subset of the input tuples $X$
which were used in some derivation:
\begin{align}
\label{eqn:Pdef}
\bP(X) \eqdef \{ x \in X : x \in \bx \in \bD(X) \}.
\end{align}
We call $\bP(X)$ the set of {\em relevant input tuples}.  As we will see later,
any tuple not in this set can be safely pruned.
In fact, $\bP(X)$ also tells us whether the query is true or false.
In particular, $\bD(X) = \emptyset$ if and only if $\bP(X) = \emptyset$.
This equivalence suggests that proving and pruning are intimately related;
in some sense, proving the query is just pruning away the query tuple.
In the remainder of the paper, we will make heavy use of $\bP$ as the principal
proving/pruning operation.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph{Computation}

We can compute $\bP(X)$ in Datalog by using the Datalog program transformation technique described in 
\cite{liang11minimal}; this allows us to use off-the-shelf Datalog solvers which have already been optimized.
Specifically, we define a set of new relations $\sR' = \{ r' : r \in \sR \}$.
For a term $t = t.r(t.\ba)$ we let $t' = t.r'(t.\ba)$ be the term that using the corresponding new relation.
We then add the following new Datalog rules:
\begin{align}
\xo' &\dlogUpdate \xo. \label{eqn:revBase} \\
s'   &\dlogUpdate z.t', z.\bs. \label{eqn:indBase}
\end{align}
The key is that a tuple $x'$ can be derived by the new Datalog program if and only if $x \in \bP(X)$.
These two rules construct $\bP(X)$ recursively:
The base case \refeqn{revBase} states that the query tuple $\xo \in \bP(X)$.
The recursive case \refeqn{indBase} states that if $x \in \bP(X)$ and a rule $z$ (with some assignment $f$)
was used to produce $x$, then for every source term $s \in z.\bs$ of that rule,
we also have $f(s) \in \bP(X)$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\Subsection{abstractions}{Abstractions}

Given a Datalog program, we define an {\em abstraction} as a equivalence relation over
the constants $\sC$ (concrete values).  In particular, we represent it as the projection
function which maps each element to its equivalence class (its abstract value):
\begin{definition}
\label{def:abstraction}
An abstraction is a function $\alpha : \sC \to \sP(\sC)$
such that for each set $s \in \range(\alpha)$,
we have $\alpha(c) = s$ for all $c \in s$.
\end{definition}
We use the natural partial order on abstractions,
where $\alpha_1 \preceq \alpha_2$ iff $\alpha_1(c) \supset \alpha_2(c)$ for all $c$,
that is, $\alpha_2$ is finer than $\alpha_1$.

% Example
In the context of our example (\reffig{graphExample}),
we could define an abstraction
that maps a sequence $c$ to the set of sequences that have the same first element as $c$
(e.g., $\alpha([10]) = \{ [1], [10], [11], \dots \}$).
Applied to a tuple, we get
\begin{align}
& \alpha(\ext(1, [0], [10])) = \\
& \quad\quad \ext(1, \{[0], [00], [01], \dots\}, \{[1], [10], [11], \dots\}). \nonumber
\end{align}
This abstraction is a simplified version of the abstractions we use in our
$k$-limited analyses (see \refsec{abstractions}).

% Extend to sets
We extend $\alpha$ to sets of constants (which we also call abstract values):
\begin{align}
\alpha(s) = \{ \alpha(c) : c \in s \}, \quad s \in \sP(\sC),
\end{align}
which returns a set of abstract values.
This allows us to naturally define compositions of abstractions
by flattening sets of abstract values:
\begin{align}
(\alpha_1 \circ \alpha_2)(c) \eqdef \cup_{s \in \alpha_1(\alpha_2(c))} s.
\end{align}
An important property of equivalence relations is
that if $\alpha_1 \preceq \alpha_2$, then $\alpha_1 \circ \alpha_2 = \alpha_1$
(applying a finer abstraction first has no impact).
Note that in general for arbitrary $\alpha_1,\alpha_2$,
the composition $\alpha_1 \circ \alpha_2$ is not an abstraction;
it is something that we must check when we consider compositions in
\refsec{abstractions}.

% Extend to tuples
We also extend $\alpha$ to concrete tuples $x$ and sets of concrete tuples $X$:
\begin{align}
\alpha(x) &= x.r(\alpha(x.a_1), \dots, \alpha(x.a_{|x.\ba|})), \\
\alpha(X) &= \{ \alpha(x) : x \in X \}.
\end{align}
Here, $\alpha(x)$ is an abstract tuple (one where the arguments are abstract values)
and $\alpha(X)$ is a set of abstract tuples.
Finally, we extend $\alpha$ to abstract tuples $b$ and sets of abstract tuples $B$:
\begin{align}
\alpha(b) &= \{ b.r(s_1, \dots, s_{|b.\ba|}) : \forall i, s_i \in \alpha(b.a_i) \}, \label{eqn:alphab} \\
\alpha(B) &= \cup_{b \in B} \alpha(b). \label{eqn:alphaB}
\end{align}
\refeqn{alphab} applies the abstraction function to each component and takes the cross product over the possible abstract values,
yielding a set of abstract tuples;
\refeqn{alphaB} aggregates these sets of abstract tuples.

Given an abstraction $\alpha$,
we can run an abstract version of the Datalog program to compute an abstract answer to the query tuple.
We do this by applying the abstraction to the concrete input tuples $X$,
to produce a set of abstract input tuples $\alpha(X)$.
We then can feed these tuples into the Datalog program,
which is oblivious to whether the tuples are abstract or concrete;
conceptually, the Datalog program now operates on the constants $\sP(\sC)$ rather than $\sC$.
\reffig{graphDerivation} shows an example of performing this computation on the
graph example from \reffig{graphExample}.

We say the query is proven by $\alpha$ if $\bP(\alpha(X)) = \emptyset$,
and it is proven only if the query is actually false ($\bP(X) = \emptyset$).
This is the standard soundness property, summarized below (see \refapp{proofs} for the proof):
\begin{proposition}[Abstraction is sound]
\label{prop:soundness}
Let $\alpha$ be an abstraction and let $X$ be any set of input tuples.
If $\bP(\alpha(X)) = \emptyset$ (the query is false abstractly),
then $\bP(X) = \emptyset$ (the query is false concretely).
\end{proposition}

\FigStar{figures/graphDerivation}{0.3}{graphDerivation}{
Computation of $\bP(X)$ on the graph example from \reffig{graphExample}
under an abstraction which maps a path onto the set of paths with the same first element
(e.g., $\alpha(\vec{10}) = \{\vec{1}, \vec{10}, \vec{11}, \dots \} \eqdef \vec{1}*$).
Each abstract tuple is derived by a rule whose source terms corresponding to the incoming edges.
Relevant input tuples ($\bP(X)$, shown in green) are the ones which are reachable by following the edges backwards;
ones which are not are pruned ($X\backslash\bP(X)$, shown in red).
}

### 2011-03-18 extra/sliver_pldi11/related.tex

\Section{related}{Related Work}

% Larger space of abstractions
There is a wealth of literature which attempts to scale static analyses without
sacrificing precision.  One general theme is to work with a flexible family of
abstractions, which in principle allows us to conform to the needs of the
client.  Milanova et al.~\cite{kobj,MilanovaRountevRyder2005} also consider
abstractions, where each local variable can be independently treated context
sensitively or insensitively, and different $k$ values can be chosen for
different allocation sites.  Lhot\'{a}k and
Hendren~\cite{LhotakHendren2006,LhotakHendren2008} present Paddle, a
parametrized framework for BDD-based, $k$-limited alias analysis.
\cite{smaragdakis11context} scale up $k$-object-sensitivity one more level by
using types rather than allocation sites.  However, which parts of the
abstraction should be more refined is largely left up to the user.

% Client-driven
Client-driven approaches use feedback from a client query to determine what
parts of an abstraction to refine.
Plevyak and Chien~\cite{PlevyakChien1994} use a refinement-based algorithm for
type inference, where context sensitivity is driven by detecting type conflict.
Guyer and Lin~\cite{GuyerLin2003} present a pointer analysis for C
which detects loss of precision (e.g., at merge points) and introduce context sensitivity.
Our method for determining relevant input tuples is similar in spirit but more general.

Perhaps the closest to this work is the part of Liang et al.~\cite{liang11minimal}, which
uses the same technique to compute the set of relevant tuples.  However, what
is done with this information is quite different.  \cite{liang11minimal} merely
stops refining the irrelevant sites whereas we actually prune all irrelevant
tuples, thereby exploiting this information fuller.  As we saw in
\refsec{experiments}, this difference had major ramifications.
%An importance difference is that we are not

%Similar to iterative refinement client-driven approaches is lazy abstraction
%\cite{henzinger02lazy,mcmillan06lazy} from model checking, which

% Demand-driven
Demand-driven analyses \cite{HeintzeTardieu2001,ZhengRugina2008} do not refine
the abstraction but rather compute an analysis on an existing
abstraction faster.  Some approaches such as \cite{SridharanBodik2006} do both.
Our \PR\ algorithm has a flavor of the client-driven approach, in that we do
refine our abstraction, but also of the demand-driven approach, in that we do
not perform a full computation (in particular, ignoring tuples which were
pruned).

%In contrast to client-driven analyses, which compute an exhaustive solution of
%varying precision, demand-driven approaches compute a partial solution of fixed
%precision.  Heintze and Tardieu's demand-driven alias analysis for
%C~\cite{HeintzeTardieu2001} performs a provably optimal computation to
%determine the points-to sets of variables queried by the client. Their analysis
%is applied to call-graph construction in the presence of function pointers. A
%more recent alias analysis developed by Zheng and Rugina~\cite{ZhengRugina2008}
%uses a demand-driven algorithm that is capable of answering alias queries
%without constructing points-to sets.

% Model checking
%Lazy abstraction \cite{henzinger02lazy} is an abstraction refinement technique
%for model checking by which one starts with a small set of predicates and tries
%to prove a query.  If the proof fails, an abstract counterexample (derivation
%in our terminology) is generated, and predicates are refined on this path.
%One difference is that they refine one predicate at a time.
%
%This work has been extended to interpolants \cite{mcmillan06lazy}, which
%provide a richer (and as they show, more effective) class of abstraction
%predicates.

%Our family of abstractions is still $\sA = \{ \klimabs_k : k = 0, 1 \}$.
%Let $K : \H \to \{ 0, 1, 2, \dots \}$ be a function mapping each site to a $k$ value.
%In \cite{liang11minimal}, we considered the abstraction family:
%\begin{align}
%\klimabs_K(c) = \klimabs_{K(c[1])}(c).
%\end{align}

% Dynamic - lower bound
Pruning has been implemented in various settings.
\cite{vipindeep05pruning} use dynamic analysis to prune down set of paths and
then focus a static analysis on these paths.
\cite{spoon04ddp} use pruning for type inference in functional languages,
where pruning is simply a heuristic which shortcuts a search algorithm.
As a result, pruning can hurt precision.
One advantage of our pruning approach is that it comes with strong soundness
and completeness guarantees.

%This is another fruitful but orthogonal way to prune.  The intuition is that if
%a query was observed to be true dynamically, there is no point trying to prove
%it statically.
%For example, we can use the dynamic analysis framework of
%\cite{liang10abstraction} to obtain lower bounds, predicates which are
%certainly true.  If we observe dynamically observe that a variable $v$ points
%to an object with abstraction $a$, the static analysis need not even attempt to
%prove that $v$ does not point to $a$.

### 2011-03-18 extra/sliver_pldi11/abstractions.tex
fixed several typos

@@
 the Datalog programs that computed the queries for our three clients based on
-$\infty$-object-sensitivity.  We now describe the abstractions that use to
+$\infty$-object-sensitivity.  We now describe the abstractions that we use to
 further abstract this analysis.
@@
 For a chain $c \in \dom^*$, let $|c|$ denote the length of the chain.
-Let $c[i]$ be the $i$-th element of $c$ (starting with index 1) and let
+Let $c[i]$ be the $i$-th element of $c$ (starting with index 1) and
 let $c[i..j]$ be the subchain $[c[i] \cdots c[j]]$
@@
 \begin{align}
-c* \eqdef \{ c' \times \H^* : c'[1..|c|] = c \}.
+c* \eqdef \{ c' \in \H^* : c'[1..|c|] = c \}.
 \end{align}
@@
 would happen and (ii) the resulting analysis would be much more imprecise because
-``$[\,]$'' is always be present for representing the context of the main method,
+``$[\,]$'' is always present for representing the context of the main method,
 but this is a prefix of any chain.
@@
 \reffig{barelyExample} shows an example of $\klimdabs_k$.
-It is not automatic that $\klimabs_k$ is a valid abstraction,
-but we can check that it is:
+We show that $\klimdabs_k$ is a valid abstraction:
 \begin{proposition}
@@
 either $|c| = k$ or $c$ is barely-repeating; in either case,
-we can check that any extension $c' \in c*$ will have $\klimabs_k(c') = c*$.
+it is easy to see that any extension $c' \in c*$ will have $\klimdabs_k(c') = c*$.
 \end{proof}
@@
 mapping each allocation site $h \in \dom$ to its its equivalence class.
-Given such such an $\hclass$, we could then extend it sequences by elementwise application:
+Given such an $\hclass$, we could then extend it sequences by elementwise application:
 \begin{align}

### 2011-03-18 extra/sliver_pldi11/experiments.tex
fixed several typos

@@
 and the type of the declaring class ($\hclassHas$) to do pre-pruning
-(specifically, $\beta_k = \klimdabs_k \circ \hclassIsHas$.
+(specifically, $\beta_k = \klimdabs_k \circ \hclassIsHas$).
 }

### 2011-03-18 extra/sliver_pldi11/introduction.tex
fixed several typos

@@
 The general theme of all these approaches is that the
-complexity of the abstraction is increased only in parts deemed useful using
+complexity of the abstraction is increased only in parts deemed relevant using
 some feedback from a client query.
@@
 We apply our pruning technique to the $k$-object-sensitivity abstraction
-\cite{kobj}, where objects on the heap are abstracted into a chain of
+\cite{kobj}, where objects in the heap are abstracted using chains of
 allocation sites.  Pruning in this setting corresponds to maintaining a subset
@@
 We ran our experiments on \numBenchmarks\ Java benchmarks using three clients that
-depend heavily on having an accurate pointer analysis: downcast safety
+depend heavily on having a precise pointer analysis: downcast safety
 checking, monomorphic call site inference, and race detection.

### 2011-03-18 extra/sliver_pldi11/preliminaries.tex
fixed several typos

@@
 Note that the tuple includes the relation as well as the arguments.
-We let $\xo$ denote a designated {\em query tuple} (e.g., $\common()$),
+We let $\xo$ denote a designated {\em query tuple} (e.g., $\common(\Ga,\Gb,3)$),
 whose truth value we want to determine.
@@
 Condition (ii) says that $\xo$ was finally produced via some sequence of rule applications.
-Condition (iii) says that the derivation of $x_n$ is minimal; each tuple was
+Condition (iii) says that the derivation of $x_n$ is minimal, that is, each tuple was
 used in some way; as we will see later, this condition is important because it allow us to
@@
 (e.g., $\alpha([10]) = \{ [1], [10], [11], \dots \}$).
-Applied to a tuple, we get
-\begin{align}
-& \alpha(\ext(1, [0], [10])) = \\
-& \quad\quad \ext(1, \{[0], [00], [01], \dots\}, \{[1], [10], [11], \dots\}). \nonumber
-\end{align}
 This abstraction is a simplified version of the abstractions we use in our
-$k$-limited analyses (see \refsec{abstractions}).
+$k$-limited analyses (see \refsec{abstractionExamples}).
 
@@
 it is something that we must check when we consider compositions in
-\refsec{abstractions}.
+\refsec{algorithm}.
 
@@
 and $\alpha(X)$ is a set of abstract tuples.
+For example:
+\begin{align}
+& \alpha(\ext(1, [0], [10])) = \\
+& \quad\quad \ext(1, \{[0], [00], [01], \dots\}, \{[1], [10], [11], \dots\}). \nonumber
+\end{align}
 Finally, we extend $\alpha$ to abstract tuples $b$ and sets of abstract tuples $B$:
@@
 (e.g., $\alpha(\vec{10}) = \{\vec{1}, \vec{10}, \vec{11}, \dots \} \eqdef \vec{1}*$).
-Each abstract tuple is derived by a rule whose source terms corresponding to the incoming edges.
+Each abstract tuple is derived by a rule whose source terms correspond to the incoming edges.
 Relevant input tuples ($\bP(X)$, shown in green) are the ones which are reachable by following the edges backwards;

### 2011-03-18 extra/sliver_pldi11/preliminaries.tex

@@
 The pruning approach that we present in this paper works on Datalog, a general
-language that has been used recently to declaratively express static analyses \cite{Whaley2007,bravenboer}.  Normally, one thinks of
-the Datalog program as encapsulating the abstraction, but it will be useful for us
-to decouple the two.\footnote{The reason is that we want to make theoretical
+language that has been used to express static analyses declaratively \cite{Whaley2007,bravenboer}.  Normally,
+these analyses are defined using one monolithic Datalog program which
+computes an abstract answer to a query.
+For us, it will be convenient to consider the Datalog program, which computes the concrete answer to a query,
+as distinct from the abstraction, which is applied on the Datalog program to generate an abstract answer.
+This separation will allow us to make theoretical
 statements comparing the same Datalog program across different
-abstractions.}  We first define Datalog (\refsec{datalog}), which describes the
+abstractions.
+
+We first define Datalog (\refsec{datalog}), which describes the
 computation of a query (think more in the direction of concrete semantics).
@@
 
-Intuitively, a Datalog program takes a set of input tuples and derives new
+A Datalog program takes a set of input tuples and derives new
 tuples.  To formalize this computation, we define the notation of a derivation.
@@
 such that $f(z.t) = x_i$ and
-$\{ x_j : j \in J \} = \{ f(s) : s \in z.\bs \}$.
+$\{ x_j : j \in J \} = \{ f(s) : s \in z.\bs \}$;
 \item [(ii)] $x_n = \xo$; and
@@
 Condition (ii) says that $\xo$ was finally produced via some sequence of rule applications.
-Condition (iii) says that the derivation of $x_n$ is minimal, that is, each tuple was
-used in some way; as we will see later, this condition is important because it allow us to
-define the set of relevant tuples for pruning.
+Condition (iii) says that in the derivation of $x_n$, every tuple is relevant for deriving $\xo$.
+%As we will see later, this condition is important because it allow us to
+%define the set of irrelevant tuples that can be pruned.
 
 A Datalog program computes $\bD(X)$.
-We say that the query $\xo$ is false (proven) if and only if $\bD(X) = \emptyset$.
+We say that the query $\xo$ is false (proven) if and only if $\bD(X)$ is empty.
 Although the query is the ultimate quantity of interest,
-the Datalog program can be used to provide more information that is useful for pruning.
+the Datalog program can be used to provide more information, which will be useful for pruning.
 Specifically, we define $\bP(X)$ to be the subset of the input tuples $X$
@@
 We can compute $\bP(X)$ in Datalog by using the Datalog program transformation technique described in 
-\cite{liang11minimal}; this allows us to use off-the-shelf Datalog solvers which have already been optimized.
+\cite{liang11minimal}.
+This allows us to use off-the-shelf Datalog solvers which have already been optimized
+to compute $\bP(X)$.
 Specifically, we define a set of new relations $\sR' = \{ r' : r \in \sR \}$.
-For a term $t = t.r(t.\ba)$ we let $t' = t.r'(t.\ba)$ be the term that using the corresponding new relation.
+For a term $t = t.r(t.\ba)$ we let $t' = t.r'(t.\ba)$ be the term that uses the corresponding new relation.
 We then add the following new Datalog rules:
 \begin{align}
-\xo' &\dlogUpdate \xo. \label{eqn:revBase} \\
-s'   &\dlogUpdate z.t', z.\bs. \label{eqn:indBase}
+\xo' &\dlogUpdate \xo, \label{eqn:revBase} \\
+s'   &\dlogUpdate z.t', z.\bs \quad \text{for each $z \in \sZ$ and $s \in z.\bs$.} \label{eqn:indBase}
 \end{align}
-The key is that a tuple $x'$ can be derived by the new Datalog program if and only if $x \in \bP(X)$.
+The key point is that a tuple $x'$ is derived by the new Datalog program if and only if $x \in \bP(X)$.
 These two rules construct $\bP(X)$ recursively:

### 2011-03-19 extra/sliver_pldi11/preliminaries.tex

@@
 
-The pruning approach that we present in this paper works on Datalog, a general
-language that has been used to express static analyses declaratively \cite{Whaley2007,bravenboer}.  Normally,
-these analyses are defined using one monolithic Datalog program which
-computes an abstract answer to a query.
-For us, it will be convenient to consider the Datalog program, which computes the concrete answer to a query,
-as distinct from the abstraction, which is applied on the Datalog program to generate an abstract answer.
-This separation will allow us to make theoretical
-statements comparing the same Datalog program across different
+Our pruning technique works on Datalog, a general language which can be used to
+express static analyses declaratively \cite{Whaley2007,bravenboer}.  Normally,
+these analyses are encoded by one monolithic Datalog program which computes an
+abstract answer to a query.  For us, it will be convenient to consider the
+Datalog program, which computes the concrete\footnote{We refer to this computation as ``concrete''
+to contrast with abstract computation we will consider later,
+but note that this ``concrete'' computation
+could already contain some level of abstraction.
+For example, the Datalog program might correspond to $\infty$-object-sensitivity
+without abstraction and $k$-object-sensitivity with abstraction.
+}
+answer to a query, as distinct
+from the abstraction, which is applied on the Datalog program to compute an
+abstract answer.  This separation allow us to make theoretical statements
+comparing the behavior of the same Datalog program across different
 abstractions.
 
-We first define Datalog (\refsec{datalog}), which describes the
-computation of a query (think more in the direction of concrete semantics).
-Then, we focus on the abstraction (\refsec{abstractions}), which only interacts
-with the Datalog program via the input tuples.
+We first define Datalog and the computation of a concrete query
+(\refsec{datalog}).  Then, we focus on the abstraction (\refsec{abstractions}),
+which interacts with the Datalog program by transforming the input tuples.
 Throughout this section, we will use \reffig{graphExample} as a running example.
@@
 \\
+%{\bf Abstraction}:
+%$\alpha_k(c) = \{ c' : \text{$c$ and $c'$ have same length $k$ prefix}\}$
+%\\
+%\\
 \includegraphics[scale=0.5]{figures/graphExample}
@@
 the Datalog program computes a set of derived tuples from the rules.
-In this case, the absence of $\common(\Ga,\Gb,3)$ means the query is false.
+In this case, the absence of $\common(\Ga,\Gb,3)$ from the derived tuples means the query is false (proven).
 }
@@
 
-\begin{figure}[ht]
+\begin{figure}[t]
 %{\bf Notation:} \\
@@
 proving/pruning operation.
+In our graph example, $\bD(X) = \bP(X) = \emptyset$,
+but as we will see later, this is not true if we apply an abstraction to $X$.
 
@@
 
-We can compute $\bP(X)$ in Datalog by using the Datalog program transformation technique described in 
+We can compute $\bP(X)$ in Datalog itself by using the Datalog program transformation technique described in 
 \cite{liang11minimal}.
-This allows us to use off-the-shelf Datalog solvers which have already been optimized
-to compute $\bP(X)$.
-Specifically, we define a set of new relations $\sR' = \{ r' : r \in \sR \}$.
+This allows us to use off-the-shelf optimized Datalog solvers to compute $\bP(X)$.
+The technique is as follows: Define a set of new relations $\sR' = \{ r' : r \in \sR \}$.
 For a term $t = t.r(t.\ba)$ we let $t' = t.r'(t.\ba)$ be the term that uses the corresponding new relation.
@@
 \end{align}
-The key point is that a tuple $x'$ is derived by the new Datalog program if and only if $x \in \bP(X)$.
+
 These two rules construct $\bP(X)$ recursively:
@@
 we also have $f(s) \in \bP(X)$.
+For example, the last rule of \reffig{graphExample} creates the following three rules:
+\begin{center}
+\scalebox{0.85}{
+$
+\begin{array}{lcl}
+\path'(g_1,c)    &\dlogUpdate& \common'(g_1,g_2,i), \path(g_1,c), \path(g_2,c), \head(c,i). \\
+\path'(g_2,c)    &\dlogUpdate& \common'(g_1,g_2,i), \path(g_1,c), \path(g_2,c), \head(c,i). \\
+\head'(c,i)      &\dlogUpdate& \common'(g_1,g_2,i), \path(g_1,c), \path(g_2,c), \head(c,i). \\
+\end{array}
+$
+}
+\end{center}
+
+The key fact is that a tuple $x'$ is derived by the new Datalog program if and only if $x \in \bP(X)$.
+Therefore, running the new Datalog program and examining tuples of the form $x'$ yields $\bP(X)$.
+In practice, we will not run $\bP$ on the concrete tuples $X$, but rather on an abstraction,
+which we consider next.
 
@@
 
-Given a Datalog program, we define an {\em abstraction} as a equivalence relation over
-the constants $\sC$ (concrete values).  In particular, we represent it as the projection
-function which maps each element to its equivalence class (its abstract value):
+Given a Datalog program, an {\em abstraction} is an equivalence relation over
+constants $\sC$.  In particular, we represent the abstraction as the projection
+function which maps each constant to its equivalence class
 \begin{definition}
@@
 \end{definition}
-We use the natural partial order on abstractions,
+We will refer to constants $\sC$ as the concrete values and $\range(\alpha)$ as
+the abstract values.
+
+Define the natural partial order on abstractions,
 where $\alpha_1 \preceq \alpha_2$ iff $\alpha_1(c) \supset \alpha_2(c)$ for all $c$,
@@
 
+%\Subsection{klimitedAbstraction}{$k$-limited abstraction}
+\paragraph{Example: $k$-limited abstraction}
+
+The main abstraction we will work with in this paper is the $k$-limited abstraction \cite{kcfa,kobj}.
+Our general theory does not depend on this particular choice, but we present
+the abstraction here so we can use it as a running example.
+
+First, we define some notation.
+Let $\dom$ be an arbitrary set;
+for the graph example of \reffig{graphExample}, let $\dom = \{ 0, 1, 2, 3 \}$;
+later, $\dom$ will be the set of allocation sites in a program.
+Define a {\em chain} to be a finite sequence in $\dom^*$.
+For a chain $c \in \dom^*$, let $|c|$ denote the length of the chain.
+Let $c[i]$ be the $i$-th element of $c$ (starting with index 1) and
+let $c[i..j]$ be the subchain $[c[i] \cdots c[j]]$
+(boundary cases: $c[i..j] = [\,]$ if $i > j$ and $c[i..j] = c[i..|c|]$ if $j > |c|$).
+For two chains $c_1$ and $c_2$, let $c_1+c_2$ denote their concatenation.
+For a chain $c$, let $c*$ denote the set of all chains
+with prefix $c$; formally:
+\begin{align}
+c* \eqdef \{ c' \in \H^* : c'[1..|c|] = c \}.
+\end{align}
+
+The {\em $k$-limited abstraction}
+partitions chains based on their length $k$ prefix.
+For an integer truncation level $k \ge 0$, define the $k$-limited abstraction $\klimabs_k$ as follows:
+\begin{align}
+\label{eqn:klimitedAbstraction}
+\klimabs_k(c) \eqdef
+\begin{cases}
+\{c\}     & \text{if $|c| < k$} \\
+c[1..k]*  & \text{if $|c| \ge k$}. \\
+\end{cases}
+\end{align}
+If the concrete chain $c$ is shorter than length $k$,
+we map it to the singleton set $\{ c \}$;
+otherwise, we map it to the set of chains that share the first $k$ elements.
+It is easy to verify that $\klimabs_k$ as defined is a valid abstraction under \refdef{abstraction}.
+
+For example, if $c = \vec{01}$,
+then we have that $\klimabs_1(c) = [0]* = \{ [0], [00], [01], [000], \dots \}$
+are the chains that start with $[0]$.
+As another example, \reffig{repeatingExample} shows the range of $\klimabs_3$.
+
+\begin{figure}
+\[
+\arraycolsep=3.5pt
+\begin{array}{llllllll}
+\{\vec{0}\}    &            &              &            & \{\vec{1}\} \\
+\{\vec{00}\}   &            & \{\vec{01}\} &            & \{\vec{10}\} &            & \{\vec{11}\} \\
+\vec{000}*     & \vec{001}* & \vec{010}*   & \vec{011}* & \vec{100}*   & \vec{101}* & \vec{110}* & \vec{111}*
+\end{array}
+\]
+\caption{\label{fig:repeatingExample} For the $k$-limited abstraction
+with $\dom = \{ 0, 1 \}$ and $k = 3$, the equivalence classes defined by $\klimabs_k$.
+}
+\end{figure}
+
+It is important that we represent $\{c\}$ and $c*$ as distinct abstract values.
+In contrast, traditional $k$-limited analyses are parametrized by a set $S$ of abstract values ``$c$'',
+where each abstract ``$c$'' represents the set of concrete chains whose longest
+matching prefix in $S$ is $c$.
+With this setup,
+every concrete chain would map to some abstract value regardless of $S$
+(note that we must have $\text{``$[\,]$''} \in S$).
+Therefore, pruning would be impossible.
+
+%Suppose that $k=3$.
+%Na\"ively, ``$\vec{01}$'' would just represent $\{\vec{01}\}$.
+%The problem comes when we prune.
+%If we prune ``$\vec{010}$'', then ``$\vec{01}$'' really represents $\{\vec{01}\}\cup \vec{010}*$
+%because any chain starting with $\vec{010}$ must now fall back on ``$\vec{01}$''.
+%This is unacceptable because pruning tuples should not change the abstraction.
+%Actually, the situation is worse, because (i) in fact no pruning
+%(in the sense that some chains are ignored)
+%would happen and (ii) the resulting analysis would be much more imprecise because
+%``$[\,]$'' is always present for representing the context of the main method,
+%but this is a prefix of any chain.
+
 % Example
-In the context of our example (\reffig{graphExample}),
-we could define an abstraction
-that maps a sequence $c$ to the set of sequences that have the same first element as $c$
-(e.g., $\alpha([10]) = \{ [1], [10], [11], \dots \}$).
-This abstraction is a simplified version of the abstractions we use in our
-$k$-limited analyses (see \refsec{abstractionExamples}).
+%In the context of our example (\reffig{graphExample}),
+%we could define an abstraction
+%that maps a sequence $c$ to the set of sequences that have the same first element as $c$
+%(e.g., $\alpha([10]) = \{ [1], [10], [11], \dots \}$).
+%This abstraction is a simplified version of the abstractions we use in our
+%$k$-limited analyses (see \refsec{abstractionExamples}).
+
+\paragraph{Extending the abstraction}
+
+Given an abstraction $\alpha$, it will be useful to extend the definition of
+$\alpha$ to not just concrete values, but also to abstract values,
+and (sets of) concrete/abstract tuples.
 
 % Extend to sets
-We extend $\alpha$ to sets of constants (which we also call abstract values):
+First, we extend $\alpha$ from concrete values $c$ to abstract values $s$:
 \begin{align}
-\alpha(s) = \{ \alpha(c) : c \in s \}, \quad s \in \sP(\sC),
+\alpha(s) \eqdef \{ \alpha(c) : c \in s \}, \quad s \in \sP(\sC),
 \end{align}
 which returns a set of abstract values.
-This allows us to naturally define compositions of abstractions
-by flattening sets of abstract values:
+This allows us to naturally define the composition of two abstractions.
+In particular, given two abstractions, $\alpha$ and $\beta$,
+define their composition to be:
 \begin{align}
-(\alpha_1 \circ \alpha_2)(c) \eqdef \cup_{s \in \alpha_1(\alpha_2(c))} s.
+(\alpha \circ \beta)(c) \eqdef \cup_{s \in \alpha(\beta(c))} s.
 \end{align}
-An important property of equivalence relations is
-that if $\alpha_1 \preceq \alpha_2$, then $\alpha_1 \circ \alpha_2 = \alpha_1$
-(applying a finer abstraction first has no impact).
-Note that in general for arbitrary $\alpha_1,\alpha_2$,
-the composition $\alpha_1 \circ \alpha_2$ is not an abstraction;
-it is something that we must check when we consider compositions in
-\refsec{algorithm}.
+Note that the composition $\alpha \circ \beta$ need not be an abstraction
+even if $\alpha_1$ and $\alpha_2$ are.\footnote{For example,
+suppose $\sC = \{1,2,3\}$;
+$\alpha(1) = \alpha(2) = \{1,2\}$, $\alpha(3) = \{ 3 \}$; and
+$\beta(1) = \beta(3) = \{1,3\}$, $\beta(2) = \{ 2 \}$.
+Then $(\alpha \circ \beta)(1) = \{1,2,3\}$ but $(\alpha \circ \beta)(2) = \{1,2\}$.
+Therefore, $\alpha \circ \beta$ is not a valid abstraction.
+}
+Therefore, when we compose abstractions in \refsec{algorithm},
+it will be important to check that the resulting compositions are valid abstractions.
+
+An important case in which compositions yield valid abstractions
+is when $\alpha \preceq \beta$ ($\beta$ is finer than $\alpha$).
+In this case, $\alpha \circ \beta = \alpha$, that is,
+applying a finer abstraction first has no impact.
 
 % Extend to tuples
-We also extend $\alpha$ to concrete tuples $x$ and sets of concrete tuples $X$:
+Next, we extend $\alpha$ to concrete tuples $x$ and sets of concrete tuples $X$
+in the natural way:
 \begin{align}
-\alpha(x) &= x.r(\alpha(x.a_1), \dots, \alpha(x.a_{|x.\ba|})), \\
-\alpha(X) &= \{ \alpha(x) : x \in X \}.
+\alpha(x) &\eqdef x.r(\alpha(x.a_1), \dots, \alpha(x.a_{|x.\ba|})), \\
+\alpha(X) &\eqdef \{ \alpha(x) : x \in X \}.
 \end{align}
@@
 \begin{align}
-& \alpha(\ext(1, [0], [10])) = \\
-& \quad\quad \ext(1, \{[0], [00], [01], \dots\}, \{[1], [10], [11], \dots\}). \nonumber
+\pi_1(\ext(1, [0], [10])) = \ext(1, [0]*, [1]*). \nonumber
 \end{align}
@@
 \begin{align}
-\alpha(b) &= \{ b.r(s_1, \dots, s_{|b.\ba|}) : \forall i, s_i \in \alpha(b.a_i) \}, \label{eqn:alphab} \\
-\alpha(B) &= \cup_{b \in B} \alpha(b). \label{eqn:alphaB}
+\alpha(b) &\eqdef \{ b.r(s_1, \dots, s_{|b.\ba|}) : \forall i, s_i \in \alpha(b.a_i) \}, \label{eqn:alphab} \\
+\alpha(B) &\eqdef \cup_{b \in B} \alpha(b). \label{eqn:alphaB}
 \end{align}
-\refeqn{alphab} applies the abstraction function to each component and takes the cross product over the possible abstract values,
+\refeqn{alphab} applies the abstraction function to each component and takes the cross product over the resulting abstract values,
 yielding a set of abstract tuples;
 \refeqn{alphaB} aggregates these sets of abstract tuples.
+For example:
+\begin{align}
+\pi_1(\ext(1, [00]*, [10]*)) = \ext(1, [0]*, [1]*). \nonumber
+\end{align}
+
+\paragraph{Using the abstraction}
 
 Given an abstraction $\alpha$,
-we can run an abstract version of the Datalog program to compute an abstract answer to the query tuple.
+we want to run the Datalog program to compute an abstract answer to the query.
 We do this by applying the abstraction to the concrete input tuples $X$,
 to produce a set of abstract input tuples $\alpha(X)$.
-We then can feed these tuples into the Datalog program,
-which is oblivious to whether the tuples are abstract or concrete;
-conceptually, the Datalog program now operates on the constants $\sP(\sC)$ rather than $\sC$.
+We then feed these tuples into the Datalog program to produce $\bP(\alpha(X))$.
+(Note that the Datalog program is oblivious to whether the tuples are abstract or concrete.)
 \reffig{graphDerivation} shows an example of performing this computation on the
-graph example from \reffig{graphExample}.
+graph example from \reffig{graphExample} with the $k$-limited abstraction $\pi_1$.
 
-We say the query is proven by $\alpha$ if $\bP(\alpha(X)) = \emptyset$,
-and it is proven only if the query is actually false ($\bP(X) = \emptyset$).
-This is the standard soundness property, summarized below (see \refapp{proofs} for the proof):
+We say the query is proven by $\alpha$ if $\bP(\alpha(X)) = \emptyset$.
+Because abstraction is sound,
+this happens only if the query is actually false ($\bP(X) = \emptyset$):
 \begin{proposition}[Abstraction is sound]
@@
 \end{proposition}
+See \refapp{proofs} for the proof.
 
 \FigStar{figures/graphDerivation}{0.3}{graphDerivation}{
-Computation of $\bP(X)$ on the graph example from \reffig{graphExample}
-under an abstraction which maps a path onto the set of paths with the same first element
-(e.g., $\alpha(\vec{10}) = \{\vec{1}, \vec{10}, \vec{11}, \dots \} \eqdef \vec{1}*$).
-Each abstract tuple is derived by a rule whose source terms correspond to the incoming edges.
-Relevant input tuples ($\bP(X)$, shown in green) are the ones which are reachable by following the edges backwards;
-ones which are not are pruned ($X\backslash\bP(X)$, shown in red).
+Computation of $\bP(\pi_1(X))$ on the graph example from \reffig{graphExample},
+where $X$ is the set of concrete input tuples, and $\pi_1$ is the $1$-limited abstraction
+which maps each path onto the set of paths with the same first element.
+%(e.g., $\alpha(\vec{10}) = \{\vec{1}, \vec{10}, \vec{11}, \dots \} \eqdef \vec{1}*$).
+In the figure,
+each abstract tuple is derived by a rule whose source terms correspond to the incoming edges.
+Relevant input tuples ($\bP(\pi_1(X))$, shown in green) are the ones which are reachable by following the edges backwards;
+ones which are not are pruned ($\pi_1(X)\backslash\bP(\pi_1(X))$, shown in red).
 }
+
+If $\alpha$ is coarse, $\bP(\alpha(X))$ will be imprecise;
+but if $\alpha$ is fine, $\bP(\alpha(X))$ will be expensive to compute.
+The next section shows how pruning can allow us to use a fine abstraction $\alpha$
+without incurring the full cost of computing $\bP(\alpha(X))$.

### 2011-03-19 extra/sliver_pldi11/theory.tex

\Section{theory}{General theory}

%In this section, we show how we can use pruning to obtain a sound analysis which
%is complete with respect to a given abstraction.
We first describe the core
idea behind pruning (\refsec{oneStep}) and then show how it can be used in the
Prune-Refine (\PR) algorithm (\refsec{algorithm}).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\Subsection{oneStep}{Pruning}

Recall that the central operation of a static analysis is $\bP$, which serves two roles: (i)
determining if the query is proven (when $\bP$ returns $\emptyset$); and (ii)
returning the subset of relevant input tuples.
The following theorem provides the key equation that drives everything in this paper (see \refapp{proofs} for the proof):

\begin{theorem}[Pruning is sound and complete]
\label{thm:master}
Let $\alpha$ and $\beta$ be two abstractions such that $\beta \preceq \alpha$
($\beta$ is coarser than $\alpha$).
Then for any set of concrete input tuples $X$, we have:
\begin{align}
\label{eqn:master}
\boxed{
\bP(\alpha(X)) = \bP(\alpha(X) \cap \alpha(\bP(\beta(X)))).
}
\end{align}
\end{theorem}

The left-hand side of \refeqn{master} corresponds to running the analysis with respect to $\alpha$.
The right-hand side corresponds to first pruning the input tuples $X$
with $\beta$ and then running the analysis with $\alpha$.
The theorem states that the two procedures obtain identical results.
The significance of this is that the right-hand side is often a much cheaper
way to compute the left-hand side.

Let us decipher \refeqn{master} a bit more.
On the right-hand side,
the abstract input tuples $\beta(X)$
are fed into the Datalog solver which computes $\bP(\beta(X))$,
which is the subset of input tuples, namely those that participate in any derivation of the abstract query tuple $\beta(\xo)$.
These relevant tuples are then refined via $\alpha$ to yield a set of tuples which are used to prune $\alpha(X)$.
The resulting subset is fed into the analysis $\bP$.
On the left-hand side, $\bP(\alpha(X))$ is the result of directly running the analysis
on the abstract tuples $\alpha(X)$ without pruning.

% Generalization
To obtain some intuition behind pruning,
consider the following simpler idea:
first run the analysis with $\beta$; if the query is proven, stop and declare {\em proven}; otherwise, we run the analysis
with $\alpha$ and output that answer.
It is easy to see that this two-step procedure returns the same answer as just running $\alpha$,
because $\beta \preceq \alpha$ implies that if $\beta$ proves the query, then so does $\alpha$
(\refprop{soundness}).
\refeqn{master} can be thought of as an extension of this basic idea:
instead of using $\beta$ to just determine where the query tuple is proven,
we obtain more information, namely the set of input tuples that are relevant.

% Complexity
The complexity of an analysis is largely determined by the number of input
tuples.  Traditionally, the abstraction alone determines the set of input
tuples and thus also the complexity of the analysis.  In our case, however, the set
of input tuples is pruned along the way, so the abstraction only partially
determines the complexity.  As we will see later, with sufficient pruning of
the input tuples, we can use a very refined abstraction at a low cost.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\Subsection{algorithm}{The Prune-Refine (\PR) algorithm}

\begin{figure}[t]
\begin{center} \scalebox{0.95}{\framebox{ \begin{minipage}{3.3in} % Begin model
\begin{center} Prune-Refine (\PR) Algorithm \end{center}
{\bf Input}: \\
\ind Sequence of abstractions: $\alpha_0 \preceq \alpha_1 \preceq \alpha_2 \preceq \cdots$ \\
\ind [Auxiliary abstractions: $\beta_t \preceq \alpha_t, t = 0, 1, 2, \dots$] \\
\ind $A_0 = \alpha_0(X)$, set of tuples \\
\\
For $t = 0, 1, 2, \dots$: \\
\ind [{\bf Pre-prune}: $A_t' \leftarrow A_t \cap \alpha_t(\bP(\beta_t(A_t)))$] \\
\ind {\bf Prune}: $\tilde A_t = \bP(A_t')$. If $\tilde A_t = \emptyset$: return {\em proven}. \\
\ind {\bf Refine}: $A_{t+1} = \alpha_{t+1}(\tilde A_t)$. \\
%\ind\ind\ind\ind\ind If $A_{t+1} = A_t$: return {\em not proven}.
\end{minipage} }} \end{center} % End model
\includegraphics[scale=0.5]{figures/algorithm}
\caption{\label{fig:pseudocode} The pseudocode and the schema for the \PR\ algorithm.
The algorithm maintains a set of (abstract) input tuples which could be
involved in some derivation of the query $\xo$ and attempts to prune down this set.
The basic version, which excludes parts in square brackets,
simply alternates between pruning and refining.  The full version includes
a pre-pruning step using auxiliary abstractions in order to further reduce the
number of tuples.
}
\end{figure}

We now turn \refthm{master} into a full algorithm, which we call the
Prune-Refine (\PR) algorithm.  \reffig{pseudocode} shows the pseudocode
of the algorithm and a diagram showing the computation of the various abstract
input tuples computed.

% Basic
This algorithm essentially applies \refeqn{master} repeatedly.
We first present a simplified version of the algorithm which
ignores the pre-pruning step (we take $A_t = A_t'$).
We are given a sequence of successively finer abstractions $\alpha_0, \alpha_1, \dots$,
and an input set of abstract tuples $A_0$, which is computed
under the initial abstraction $\alpha_0$.  Then the algorithm alternates between a pruning step
and a refining step, always maintaining only the abstract tuples
that could participate in a derivation of the query tuple $\xo$.
On iteration $t$, our current input tuples $A_t$ are first pruned to $\tilde A_t$ using $\bP$;
this is subsequently refined to $A_{t+1}$.
% Example
\reffig{graphIters} shows an example of running this algorithm on the graph
example from \reffig{graphExample}; the first pruning step is shown in
\reffig{graphDerivation}.

\Fig{figures/graphIters}{0.3}{graphIters}{
The abstract input tuples computed by
the \PR\ algorithm on the graph example from \reffig{graphExample} (without pre-pruning),
where the abstraction $\alpha_t$ maps $c$ to all paths that match $c$ on the first $t+1$ nodes.
Notation: $c*$ denotes the set of all paths with prefix $c$.
During the first pruning step, $\ext(2,\vec{0}*,\vec{2}*)$ is pruned from $A_0$,
yielding $\tilde A_0$.  In the refine step, we expand $\vec{1}*$ to $\{ \vec{1} \}$ and $\vec{10}*$,
resulting in an extra tuple.  In the second pruning step, we prove the query
(pruning everything).
}

% Pre-pruning
Now we discuss pre-pruning.  Pre-pruning requires the user to provide
another sequence of successively finer abstractions $\beta_0, \beta_1, \dots$
which are coarser than $\alpha_0, \alpha_1, \dots$, respectively.
These abstractions will also be used to prune the input tuples.
The idea is that before refining $A_t$ to $A_{t+1}$,
we perform two steps of pruning:
(i) first using $\beta_t$ during pre-pruning step, and then
(ii) using $\alpha_t$ during the main pruning step.
Pre-pruning does not affect the output of the pruning step ($\tilde A_t$);
the purpose is solely to make the pruning step faster.
One way to obtain the auxiliary abstractions is via
compositions with another abstraction $\hclass$: $\beta_t = \alpha_t \circ \hclass$.
Intuitively, $\hclass$ should neither be too coarse nor too fine.
If $\hclass$ is no abstraction, then pre-pruning is equivalent to running the pruning step;
if $\hclass$ is the trivial abstraction, pre-pruning will be fast but nothing will be pre-pruned.
A rule of thumb is that $\hclass$ should be complementary to $\alpha_t$
(we will see examples in \refsec{abstractions}).

\refthm{algorithm} states that the \PR\ algorithm is both
sound and complete.  In other words, pruning has no impact on our answer to a
query---we just compute it more efficiently.
The proof is given in \refapp{proofs} and uses \refthm{master}.

\begin{theorem}[Correctness of the \PR\ algorithm]
\label{thm:algorithm}
At iteration $t$,
the incrementally pruned abstraction $A_t$ is equivalent
to the full abstraction $\alpha_t(X)$ (formally, $\bP(\alpha_t(X)) = \bP(A_t)$).
If the algorithm returns ``proven,'' then $\bP(X) = \emptyset$ (the query is actually false).
\end{theorem}

### 2011-03-20 extra/sliver_pldi11/klimited.tex
first pass

@@
 % Basics
-We now apply our general pruning approach to $k$-limited pointer analysis.
+We now introduce our $k$-limited pointer analysis,
+on which we will apply the Prune-Refine algorithm.
 Each node in the control-flow graph of each method $m \in \M$ is associated with a
@@
 % Context, abstraction
-Our analysis represents both contexts and abstract objects using a list of
-allocation sites, which we will call {\em chains}.  However, note that these
-chains are never truncated in the Datalog program.  Therefore, running the
-Datalog program (ignoring the fact that it might not terminate) is performing
-an $\infty$-object-sensitivity analysis.  Recall that we separate the
-computation of the query in Datalog with the abstraction which is applied to
-the input tuples (of course, our Datalog program is itself an approximation to
-the concrete semantics---it is flow-insensitive, does not handle primitive
-data, etc.).  \refsec{abstractions} will describe further abstractions of this
-analysis.
+The analysis represents both contexts and abstract objects using chains of
+allocation sites ($\C = \H^*$).  Contexts are extended into new contexts via the $\ext$ relation,
+which prepends an allocation site to a chain (e.g., $\ext(3,[12],[312])$).
+Note that these chains are not truncated in the Datalog program, and therefore,
+running the Datalog program directly (ignoring the fact that it might not
+terminate) corresponds to performing an $\infty$-object-sensitivity analysis.
+
+Although this Datalog program itself is an approximation to the concrete
+program semantics---it is flow-insensitive, does not handle primitive data,
+etc., we will informally say that a client query computed with respect to this
+analysis yields a concrete answer.  In contrast, we obtain an abstract
+answer by computing the client query with respect to a $k$-limited abstraction,
+which we will discuss in \refsec{generalRelationship}.
+
+% Rules
+We now briefly describe the Datalog rules in \reffig{aliasAnalysisRules}.  Rule (1)
+states that the main method $\mmain$ is reachable in a distinguished context $[\,]$.
+Rule (2) states that a target method of a reachable call site is also reachable.
+Rule (3) states that every statement in a reachable method is also reachable.
+Rules (4) through (9) implement the transfer function associated with each kind of statement.
+Rule (10) analyzes the target method
+$m$ in a separate context $o$ for each abstract object $o$ to which the distinguished
+{\tt this} argument of method $m$ points, and rule (11) sets the points-to set
+of the {\tt this} argument of method $m$ in context $o$ to the singleton $\{ o \}$.
 
@@
 heap graph ($\heap$); and a context-sensitive call graph ($\cg$).
-
-We briefly describe the analysis rules in Datalog.  Rule (1)
-states that the main method $\mmain$ is reachable in a distinguished context $[]$.
-Rule (2) states that a target method of a reachable call site is also reachable.
-Rule (3) states that every statement in a reachable method is also reachable.
-Rules (4) through (9) implement the transfer function associated with each kind of
-statement.
-%Rule (10) populate the call graph and
-%rule (11) propagate the points-to set from the argument
-%of a call site to the formal argument of each target method.
-Rule (10) analyzes the target method
-$m$ in a separate context $o$ for each abstract object $o$ to which the distinguished
-{\tt this} argument of method $m$ points, and rule (11) sets the points-to set
-of the {\tt this} argument of method $m$ in context $o$ to the singleton $\{ o \}$.
 
@@
 \argM     &\subset& \M \times \V & \text{(method's formal argument variable)} \\
-\ext      &\subset& \S \times \C \times \C & \text{(extend context with site)} \\
+\ext      &\subset& \H \times \C \times \C & \text{(extend context with site)} \\
           & =     & \multicolumn{2}{l}{\{ (h, c, [h]+c) : h\in\H, c\in\C \}} \\
@@
 
-The core pointer analysis is used by three clients, which each define a set of queries.
+The core pointer analysis just described is used by three clients, which each
+define a set of queries.
 
@@
 analysis.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\Subsection{generalRelationship}{Relationship to general notation}
+
+% Relate 
+We now describe our analysis in \reffig{aliasAnalysisRules} in terms of our general notation presented in \refsec{preliminaries}.
+The set of concrete values $\sC$ is the union of all the domains
+(e.g.., allocation sites $\H = \{ 1, 2, 3, \dots \}$, abstract objects $\C = \H^*$, etc.).
+The input tuples $X$ are specified by the input relations
+(e.g., $X = \{ \body(\mmain, x = \new\,3), \ext(3,[12],[312]) \}$).
+Each of the three clients defines a set of possible query tuples, for example,
+$\xo = \dl{unsafe}(\dl{v4},\dl{v8})$ for downcast safety checking of an assignment $\dl{v4} = \dl{v8}$.
+
+Recall that $\bP(X)$ corresponds to obtaining an answer to a client query with respect to $\infty$-object-sensitivity.
+To obtain $k$-object-sensitivity,
+we first apply the $k$-limited abstraction $\pi_k$ to the input tuples
+and run the Datalog program on these abstract tuples ($\bP(\pi_k(X))$).
+Note that only the $\ext$ tuples are affected by the abstraction.

### 2011-03-20 extra/sliver_pldi11/related.tex
first pass

@@
 abstractions, which in principle allows us to conform to the needs of the
-client.  Milanova et al.~\cite{kobj,MilanovaRountevRyder2005} also consider
-abstractions, where each local variable can be independently treated context
-sensitively or insensitively, and different $k$ values can be chosen for
-different allocation sites.  Lhot\'{a}k and
+client.  Milanova et al.~\cite{kobj,MilanovaRountevRyder2005} consider
+abstractions where each local variable can be independently treated
+context-sensitively or context-insensitively, and different $k$ values can be
+chosen for different allocation sites.  Lhot\'{a}k and
 Hendren~\cite{LhotakHendren2006,LhotakHendren2008} present Paddle, a
 parametrized framework for BDD-based, $k$-limited alias analysis.
-\cite{smaragdakis11context} scale up $k$-object-sensitivity one more level by
-using types rather than allocation sites.  However, which parts of the
+\cite{smaragdakis11context} scale up $k$-object-sensitivity, increasing $k$ by one
+using types rather than allocation sites.  However, in all of this work, which parts of the
 abstraction should be more refined is largely left up to the user.
@@
 Plevyak and Chien~\cite{PlevyakChien1994} use a refinement-based algorithm for
-type inference, where context sensitivity is driven by detecting type conflict.
+type inference, where context-sensitivity is driven by detecting type conflicts.
 Guyer and Lin~\cite{GuyerLin2003} present a pointer analysis for C
-which detects loss of precision (e.g., at merge points) and introduce context sensitivity.
+which detects loss of precision (e.g., at merge points) and introduce context-sensitivity.
 Our method for determining relevant input tuples is similar in spirit but more general.
 
-Perhaps the closest to this work is the part of Liang et al.~\cite{liang11minimal}, which
-uses the same technique to compute the set of relevant tuples.  However, what
+Section~3.1 of Liang et al.~\cite{liang11minimal} (not the focus of that work)
+also computes the set of relevant tuples by running a transformed Datalog program.  However, what
 is done with this information is quite different.  \cite{liang11minimal} merely
 stops refining the irrelevant sites whereas we actually prune all irrelevant
-tuples, thereby exploiting this information fuller.  As we saw in
+tuples, thereby exploiting this information fully.  As we saw in
 \refsec{experiments}, this difference had major ramifications.
-%An importance difference is that we are not
 
@@
 Demand-driven analyses \cite{HeintzeTardieu2001,ZhengRugina2008} do not refine
-the abstraction but rather compute an analysis on an existing
-abstraction faster.  Some approaches such as \cite{SridharanBodik2006} do both.
-Our \PR\ algorithm has a flavor of the client-driven approach, in that we do
+the abstraction but rather try to compute an analysis on an existing
+abstraction more efficiently.
+Sridharan et al.~\cite{SridharanBodik2006} presents an algorithm
+which casts pointer analysis as a CFL-reachability problem and relaxes by
+introducing additional edges.
+
+Our Prune-Refine algorithm has a flavor of the client-driven approach, in that we do
 refine our abstraction, but also of the demand-driven approach, in that we do
 not perform a full computation (in particular, ignoring tuples which were
-pruned).
+pruned).  There are three important differences between the present work
+and the work described earlier:
+First, while most of that work is specific to pointer analysis, our Prune-Refine algorithm
+is generally applicable to any Datalog program.
+Second, while past work is based on selected iterative refinement,
+Prune-Refinement is based on generalizing the idea of program slicing.
+%Finally, we were able to prove rigorous theoretical guarantees that our algorithm is as precise
+%as a more expensive algorithm.
 
@@
 % Dynamic - lower bound
-Pruning has been implemented in various settings.
+Other forms of pruning have been implemented in various settings.
 \cite{vipindeep05pruning} use dynamic analysis to prune down set of paths and

### 2011-03-20 extra/sliver_pldi11/theory.tex
first pass

@@
 idea behind pruning (\refsec{oneStep}) and then show how it can be used in the
-Prune-Refine (\PR) algorithm (\refsec{algorithm}).
+Prune-Refine algorithm (\refsec{algorithm}).
 
@@
 with $\beta$ and then running the analysis with $\alpha$.
-The theorem states that the two procedures obtain identical results.
-The significance of this is that the right-hand side is often a much cheaper
-way to compute the left-hand side.
+The theorem states that the two procedures obtain identical results
+(the right-hand side is sound and complete with respect to the left-hand side).
+The significance of this is that the right-hand side is often much cheaper
+to compute than the left-hand side.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\Subsection{algorithm}{The Prune-Refine (\PR) algorithm}
+\Subsection{algorithm}{The Prune-Refine algorithm}
 
@@
 \begin{center} \scalebox{0.95}{\framebox{ \begin{minipage}{3.3in} % Begin model
-\begin{center} Prune-Refine (\PR) Algorithm \end{center}
+\begin{center} Prune-Refine algorithm \end{center}
 {\bf Input}: \\
@@
 \includegraphics[scale=0.5]{figures/algorithm}
-\caption{\label{fig:pseudocode} The pseudocode and the schema for the \PR\ algorithm.
+\caption{\label{fig:pseudocode} The pseudocode and the schema for the Prune-Refine algorithm.
 The algorithm maintains a set of (abstract) input tuples which could be
 involved in some derivation of the query $\xo$ and attempts to prune down this set.
-The basic version, which excludes parts in square brackets,
+The basic version of the algorithm, which excludes the lines in square brackets and the dotted boxes,
 simply alternates between pruning and refining.  The full version includes
-a pre-pruning step using auxiliary abstractions in order to further reduce the
+a pre-pruning step, which uses auxiliary abstractions to further reduce the
 number of tuples.
@@
 We now turn \refthm{master} into a full algorithm, which we call the
-Prune-Refine (\PR) algorithm.  \reffig{pseudocode} shows the pseudocode
+Prune-Refine algorithm.  \reffig{pseudocode} shows the pseudocode
 of the algorithm and a diagram showing the computation of the various abstract
@@
 ignores the pre-pruning step (we take $A_t = A_t'$).
-We are given a sequence of successively finer abstractions $\alpha_0, \alpha_1, \dots$,
+We are given a sequence of successively finer abstractions $\alpha_0, \alpha_1, \dots$
+(e.g., $\alpha_t = \pi_t$ for the $k$-limited abstractions)
 and an input set of abstract tuples $A_0$, which is computed
@@
 The abstract input tuples computed by
-the \PR\ algorithm on the graph example from \reffig{graphExample} (without pre-pruning),
-where the abstraction $\alpha_t$ maps $c$ to all paths that match $c$ on the first $t+1$ nodes.
-Notation: $c*$ denotes the set of all paths with prefix $c$.
+the Prune-Refine algorithm on the graph example from \reffig{graphExample} (without pre-pruning).
+We are using $k$-limited abstractions ($\alpha_t = \pi_t$).
 During the first pruning step, $\ext(2,\vec{0}*,\vec{2}*)$ is pruned from $A_0$,
-yielding $\tilde A_0$.  In the refine step, we expand $\vec{1}*$ to $\{ \vec{1} \}$ and $\vec{10}*$,
-resulting in an extra tuple.  In the second pruning step, we prove the query
+yielding $\tilde A_0$.  In the refinement step, we expand $\vec{1}*$ to $\{ \vec{1} \}$ and $\vec{10}*$.
+In the second pruning step, we prove the query
 (pruning everything).
@@
 we perform two steps of pruning:
-(i) first using $\beta_t$ during pre-pruning step, and then
-(ii) using $\alpha_t$ during the main pruning step.
+(i) first we use $\beta_t$ in a pre-pruning step;
+(ii) then we use $\alpha_t$ during the main pruning step.
 Pre-pruning does not affect the output of the pruning step ($\tilde A_t$);
 the purpose is solely to make the pruning step faster.
-One way to obtain the auxiliary abstractions is via
-compositions with another abstraction $\hclass$: $\beta_t = \alpha_t \circ \hclass$.
-Intuitively, $\hclass$ should neither be too coarse nor too fine.
-If $\hclass$ is no abstraction, then pre-pruning is equivalent to running the pruning step;
-if $\hclass$ is the trivial abstraction, pre-pruning will be fast but nothing will be pre-pruned.
-A rule of thumb is that $\hclass$ should be complementary to $\alpha_t$
-(we will see examples in \refsec{abstractions}).
 
-\refthm{algorithm} states that the \PR\ algorithm is both
+Pre-pruning requires a sequence of auxiliary abstractions $(\beta_t)$ which are coarser than the main abstractions $(\alpha_t)$.
+We can obtain auxiliary abstractions by composing the main abstractions
+with another abstraction $\hclass$ ($\beta_t = \alpha_t \circ \hclass$).
+We can use any $\hclass$ for which $\alpha_t \circ \hclass$ yields a valid abstraction,
+but the speedup we obtain from pre-pruning depends on the relationship between $\hclass$ and $\alpha_t$.
+If $\hclass$ is the total abstraction ($\hclass(c) = \sC$), then pre-pruning will be fast but nothing will be pre-pruned,
+so we get no speedup.
+If $\hclass$ is no abstraction ($\hclass(c) = c$), then pre-pruning is equivalent to just running the pruning step,
+so we again get no speedup.
+A good rule of thumb is that $\hclass$ should be ``complementary'' to $\alpha_t$
+(we will see some examples in \refsec{abstractions}).
+
+\refthm{algorithm} states that the Prune-Refine algorithm is both
 sound and complete.  In other words, pruning has no impact on our answer to a
@@
 
-\begin{theorem}[Correctness of the \PR\ algorithm]
+\begin{theorem}[Correctness of the Prune-Refine algorithm]
 \label{thm:algorithm}

### 2011-03-20 extra/sliver_pldi11/figures/results.rb

@@ -151,7 +151,7 @@ def absIterGraph(metric,logScale)
       g.yrotateAxisLabel(false)
       g.ysciNotation.xrange(0) if metric == 'Ap'
       g.yexpValue.ytickIncrValue(Math.log(10)) if logScale
-      g.xroundPlaces(0).xrange(1).xtickIncrValue([numIters/4,1].max)
+      g.xroundPlaces(0).xrange(0).xtickIncrValue([numIters/4,1].max)
       g.xlength(u(2.5)).ylength(u(2))
       g.colors(red, magenta, green, blue)
       nameGraphs << [CB(client, benchmark), g]
@@ -160,7 +160,7 @@ def absIterGraph(metric,logScale)
   #nameGraphs[0][1] = nameGraphs[1][1].getLegend
   arrangeGraphs(nameGraphs, $benchmarks.size)
 end
-printObj(:obj => absIterGraph('Ap',false).signature(147), :outPrefix => 'absSizeGraph')
+printObj(:obj => absIterGraph('Ap',false).signature(148), :outPrefix => 'absSizeGraph')
 #printObj(:obj => absIterGraph('Ap',true).signature(135), :outPrefix => 'absSizeGraph-log')
 #printObj(:obj => absIterGraph('time',false).signature(138), :outPrefix => 'timeGraph')
 #printObj(:obj => absIterGraph('time',true).signature(136), :outPrefix => 'timeGraph-log')

### 2011-03-20 extra/sliver_pldi11/appendix.tex

\appendix
\Section{proofs}{Proofs}

Instead of directly proving \refprop{soundness}, we state a more general
theorem which will be useful later:

\begin{theorem}[Soundness]
\label{thm:soundness}
Let $\alpha$ and $\beta$ be two abstractions with $\beta \preceq \alpha$ ($\beta$ is coarser),
and let $X$ be any set of input tuples.
For any derivation $\ba \in \bD(\alpha(X))$,
define $\bb = (b_1, \dots, b_{|\ba|})$ where each $b_i$ is the unique element in $\beta(a_i)$.
Then $\bb \in \bD(\beta(X))$.
\end{theorem}
\begin{proof}[Proof of \refthm{soundness}]
Define $A=\alpha(X)$ and $B=\beta(X)$.
Consider $\ba \in \bD(A)$ and let $\bb$ be as defined in the theorem.
For each position $i$,
we have two cases.
First, if $a_i \in A$, then $b_i \in (\beta \circ \alpha)(X) = \beta(X) = B$.
Otherwise, let $z \in \sZ$ be the rule and $J$ be the indices of the tuples used to derive $a_i$.
The same rule $z$ and the corresponding tuples $\{ b_j : j \in J \}$ can also be used to derive $b_i$.
Therefore, $\bb \in \bD(B)$.
\end{proof}

\begin{proof}[Proof of \refprop{soundness} (abstraction is sound)]
Apply \refthm{soundness} with $\beta = \alpha$ and $\alpha$ as the identity function (no abstraction).
\end{proof}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Before we prove \refthm{master}, we state a useful lemma.
\begin{lemma}[Pruning is idempotent]
\label{lem:Pidem}
For any set of tuples (concrete or abstract) $X$, $\bP(X) = \bP(\bP(X))$.
\end{lemma}
\begin{proof}[Proof]
Since $\bP(X) \subset X$ by definition and $\bP$ is monotonic,
we have $\bP(\bP(X)) \subset \bP(X)$.
For the other direction, let $x \in \bP(X)$.
Then $x$ is part of some derivation ($x \in \bx \in \bD(X)$).
All the input tuples of $\bx$ (those in $\bx\cap X$) are also in $\bP(X)$,
so $\bx \in \bD(\bP(X))$.  Therefore $x \in \bP(\bP(X))$.
\end{proof}

\begin{proof}[Proof of \refthm{master} (pruning is sound and complete)]
We define variables for the intermediate quantities in \refeqn{master}:
$A = \alpha(X)$ and $B = \beta(X)$, $\tilde B = \bP(B)$, $\tilde A = \alpha(\tilde B)$,
and $A' = A \cap \tilde A$.
We want to show that pruning is sound ($\bP(A) \subset \bP(A')$)
and complete ($\bP(A) \supset \bP(A')$).
Completeness follows directly because $A \supset A'$ and $\bP$ is monotonic (increasing
the number of input tuples can only increase the number of drived tuples).

Now we show soundness.
Let $a \in \bP(A)$.  By definition of $\bP$ (\refeqn{Pdef}), there is a derivation $\ba \in \bD(A)$ containing $a$.
For each $a_i \in \ba$, let $b_i$ be the unique element in $\beta(a_i)$ (a singleton set because $\beta \preceq \alpha$),
and let $\bb$ be the corresponding sequence constructed from the $b_i$s.
Since $\beta \preceq \alpha$, we have $\bb \in \bD(B)$ by \refthm{soundness},
and so each input tuple in $\bb$ is also in $\bP(B) = \tilde B$;
in particular, $b = \tilde B$ for $\beta(a) = \{ b \}$.
Since $\beta \preceq \alpha$, $a \in \alpha(b)$, and so $a \in \tilde A$.
We have thus shown that $\bP(A) \subset \tilde A$.
Finishing up, $\bP(A') = \bP(A \cap \tilde A) \supset \bP(A \cap \bP(A)) = \bP(\bP(A)) = \bP(A)$,
where the last equality follows from idempotence (\reflem{Pidem}).
\end{proof}

%Intuition: derive everything we could before
%\begin{theorem}[Relevance]
%\label{thm:relevance}
%Let $\alpha_1 \preceq \alpha_2$ be two abstractions ($\alpha_2$ is finer).
%Let $X$ be any set of tuples.
%Let $A_1 = \alpha_1(X)$ and $A_2 = \alpha_2(X)$ being the abstract tuples.
%Let $\tilde A_1 = \bP(A_1)$ be the relevant abstract values with respect to $\alpha_1$.
%Let $\tilde A_2 = \alpha_2(\tilde A_1)$ be the corresponding abstract values with respect to $\alpha_2$.
%Then for any sequence $(b_1, \dots, b_n)$ with $b_n = \xo$,
%\[ (b_1, \dots, b_n) \in \bD(A_2) \implies (b_1, \dots, b_n) \in \bD(\tilde A_2). \]
%Compactly, $\bD(\alpha_2(X)) \subset \bD(\alpha_2(\bP(\alpha_1(X))))$.
%\end{theorem}
%
%\begin{proof}[Proof of \refthm{relevance} (relevance)]
%Consider any derivation $(b_1, \dots, b_n) \in \bD(A_2)$ with $b_n = \xo$.
%For each $i = 1, \dots, n$, let $a_i = \alpha_1(b_i)$.
% By the soundness theorem, $(a_1, \dots, a_n) \in \bD(A_1)$.
%Because $(a_1, \dots, a_n)$ is a derivation of $a_n = \xo$,
% all these tuples are used, so
% each $a_i \in \bR(A_1) = \tilde A_1$.
%We now argue that $(b_1, \dots, b_n) \in \bD(\tilde A_2)$.
% For each $i = 1, \dots, n$:
%  If $b_i \in A_2$, then $a_i \in A_1$.  Since $a_i$ is part of a derivation of $\xo$ under $\alpha_1$,
%   we also have $a_i \in \tilde A_1$.
%   Note that $a_i \supset b_i$ because $\alpha_1 \preceq \alpha_2$.
%   Therefore, $b_i \subset a_i \subset \cup_{a \in \tilde A_1} a$.
%   Applying $\alpha_2$ to both sides yields $b_i \in \tilde A_2$.
%  If $b_i \not\in A_2$, it must have been derived due to some rule applied to $(J,i) \in P(b_1, \dots, b_n)$.
%   The same rule still applies.
%\end{proof}

We now show that the Prune-Refine algorithm is correct, which follows from a
straightforward application of \refthm{master}.

\begin{proof}[Proof of \refthm{algorithm} (correctness of the Prune-Refine algorithm)]
First, we argue that pre-pruning is correct.
For each iteration $t$, we invoke \refthm{master} with $\alpha = \alpha_t,
\beta = \beta_t$ and $X$ be such that $\alpha(X) = A_t$.
The result is that $\bP(A_t) = \bP(A_t')$, so without loss of generality,
we will assume $A_t = A_t'$ for the rest of the proof.

Now fix an iteration $t$.
We will show that $\bP(\alpha_t(X)) = \bP(A_t)$ by induction,
where the inductive hypothesis is $\bP(\alpha_t(X)) = \bP(\alpha_t(\tilde A_s))$.
For the base case ($s = -1$), we define $\tilde A_{-1} = \{X\}$, we get a tautology.
For the inductive case, apply the theorem with 
applied to $\beta = \alpha_s, \alpha = \alpha_t$ and $X$ such that $\alpha_s(X) = \alpha_s(\tilde A_{s-1})$,
we get that 
\[ \bP(\alpha_t(\tilde A_{s-1})) = \bP(\alpha_t(\bP(\alpha_s(\tilde A_{s-1})))) = \bP(\alpha_t(\tilde A_s)). \]
When $s = t-1$, we have $\bP(\alpha_t(X)) = \bP(\alpha_t(\tilde A_{t-1})) = \bP(A_t)$,
completing the claim.
Finally, if the algorithm returns {\em proven}, we have
\[ \emptyset = \bP(A_t) = \bF(\alpha_t(X)) \supset \bF(X). \]
%If the algorithm returns {\em not proven},
%then $A_t \subset \{ \{ x \} : x \in X \}$,
%so $\xo \in \bF(A_t)$ (not proven) implies $\xo \in \bF(X)$ (not proven) by monotonicity of $\bF$.
\end{proof}

### 2011-03-20 extra/sliver_pldi11/abstract.tex
fixed several typos

@@
 Many static analyses do not scale as they are made more precise.  For example,
-increasing the amount of context sensitivity in a $k$-limited analysis causes
+increasing the amount of context sensitivity in a $k$-limited pointer analysis causes
 the number of contexts to grow exponentially.  Iterative refinement techniques
-can be employed to mitigate this growth by starting with a coarse abstraction and
-only refining parts of the abstraction deemed relevant with respect to a client.
+can mitigate this growth by starting with a coarse abstraction and
+only refining parts of the abstraction that are deemed relevant with respect to a client.
 
-In this paper, we introduce a new way to use this client feedback to greatly
-reduce the complexity of an analysis: {\em pruning}.  The pruned analysis is no
+In this paper, we introduce a new technique called {\em pruning} that uses this client
+feedback to greatly reduce the complexity of an analysis.  The pruned analysis is no
 longer a valid abstraction, but we prove that we do not forfeit correctness and
-that our analysis is still sound with respect to a given client.  In the
-context of $k$-limited analysis, our approach amounts to keeping track of a set
+that our analysis is still sound with respect to a given client. 
+For $k$-limited pointer analysis, our approach amounts to keeping track of a set
 of context prefix patterns.  By iteratively refining and pruning these
-patterns, we show that our analysis increases scalability greatly.
+patterns, we show that our analysis can greatly increase scalability.
 

### 2011-03-20 extra/sliver_pldi11/appendix.tex

@@
 Completeness follows directly because $A \supset A'$ and $\bP$ is monotonic (increasing
-the number of input tuples can only increase the number of drived tuples).
+the number of input tuples can only increase the number of derived tuples).
 

### 2011-03-20 extra/sliver_pldi11/experiments.tex

--- 2011-03-18 extra/sliver_pldi11/experiments.tex
@@
 
-In this section, we apply our \PR\ algorithm (\refsec{algorithm}) to
-$k$-object-sensitivity for three clients:
+In this section, we apply the Prune-Refine algorithm (\refsec{algorithm}) to
+$k$-object-sensitivity for our three clients (\refsec{clients}):
 downcast safety checking (\downcast),
 monomorphic call site inference (\monosite),
-and race detection (\race).  These analyses are described in \refsec{klimited}.
+and race detection (\race).
 Our main empirical result is that across different clients and benchmarks,
-pruning is very effective at curbing the exponential growth,
-which allows us to run the analysis on very refined abstractions.
+pruning is effective at curbing the exponential growth,
+which allows us to run analyses on finer abstractions.
 
@@
 an extensible program analysis framework for Java bytecode,\footnote{{\tt{http://code.google.com/p/jchord/}}}
-which uses the BDD solver {\tt bddbddb} \cite{Whaley2007}.
+which uses the BDD Datalog solver {\tt bddbddb} \cite{Whaley2007}.
 We evaluated on \numBenchmarks\ Java benchmarks shown in \reftab{benchmarks}.
@@
 which considers a sequence of abstractions $\balpha = (\balpha_0, \balpha_1, \dots)$
-but stops refining sites which have been deemed irrelevant (this algorithm is denoted $\Site(\balpha)$).
+but stops refining sites which have been deemed irrelevant.  This algorithm is denoted $\Site(\balpha)$.
 
 As for the new methods that we propose in this paper,
-we have $\PR(\balpha)$, which uses a sequence of abstractions $\balpha$, which does no pre-pruning,
-or $\PR(\balpha,\hclass)$, which performs pre-pruning using $\beta_t = \alpha_t
+we have $\PR(\balpha)$, which corresponds to the Prune-Refine (\PR) algorithm using a sequence of abstractions $\balpha$ with no pre-pruning;
+and $\PR(\balpha,\hclass)$, which performs pre-pruning using $\beta_t = \alpha_t
 \circ \hclass$ for $t = 0, 1, 2, \dots$.  We consider three choices of $\hclass$
-($\hclassIs,\hclassHas,\hclassIsHas$) which use different kinds of type information.
+which use different kinds of type information ($\hclassIs,\hclassHas,\hclassIsHas$)
 
-In our implementation of the \PR\ algorithm, we depart slightly from our
+In our implementation of the Prune-Refine algorithm, we depart slightly from our
 presentation.  Instead of maintaining the set of relevant input tuples as
@@
 maintain the allocation site chains which were involved in any relevant input
-tuple.  We can modify the original Datalog program so that the \PR\ algorithm
+tuple.  This choice results in more conservative pruning, but reduces the
+amount of state that we have to keep.
+We can modify the original Datalog program so that the Prune-Refine algorithm
 is consistent with this implementation by introducing a new input relation
-$\dl{active}(c)$ for $c \in \H^*$, encoding existing input relations as rules
-consisting of zero source terms, and adding $\dl{active}(c)$ for each existing
-rule that uses a variable $c$ which is an allocation site.  Now computing the
-relevant input tuples corresponds exactly to computing the set of relevant
-allocation site chains.
+$\dl{active}(c)$ for $c \in \H^*$, encoding existing $\ext$ input relations as rules
+with zero source terms, and adding $\dl{active}(c)$ for each existing
+rule that uses a chain-valued variable.  Computing the
+relevant input tuples in this modified Datalog program corresponds exactly to
+computing the set of relevant allocation site chains.
 
@@
 \text{\bf Algorithms} \\
-\Full(\alpha)         & \defn{standard analysis using abstraction $\alpha$} \\
+\Full(\alpha)         & \defn{standard analysis using an abstraction $\alpha$} \\
 \Site(\balpha)        & \defn{site-based refinement \cite{liang11minimal} on abstractions $\balpha$} \\
@@
 $\PR(\Klimdabs,\hclassIsHas)$
-means running the \PR-algorithm on the barely-repeating $k$-limited abstraction ($\alpha_k = \klimdabs_k$),
+means running the Prune-Refine algorithm on the barely-repeating $k$-limited abstraction ($\alpha_k = \klimdabs_k$),
 using a composed abstraction based on the type of an allocation site ($\hclassIs$) 
 and the type of the declaring class ($\hclassHas$) to do pre-pruning
-(specifically, $\beta_k = \klimdabs_k \circ \hclassIsHas$.
+(specifically, $\beta_k = \klimdabs_k \circ \hclassIsHas$).
 }
@@
 
-\reffig{absSizeGraph} compares the number of input tuples needed by various
-algorithms; we see that the non-pruning methods completely hit a wall, whereas
-the pruning methods are able to continue increasing $k$ farther.  We also
-observed that the running time of these analyses tracked the number of tuples
-somewhat, so we do obtain speedups, but the differences are less pronounced.
-One reason is that we are using BDDs, which can handle large numbers of tuples
+We ran four analyses using $k$-limited abstractions, seeing
+how far we could increase $k$ until the analyses ran out of memory.
+For each analysis, we also measured the number of input tuples given to
+the Datalog solver; this quantity is denoted as $|A_t'|$ (see
+\reffig{pseudocode}).  In this section, the number of iterations $t$ is the same as the $k$ value.
+
+\reffig{absSizeGraph} plots the number of tuples $|A_t'|$ as a function of
+number of iterations $t$ for our four algorithms.  We see that the
+non-pruning algorithms completely hit a wall after a few iterations, with the
+number of tuples exploding exponentially.  On most benchmark-client pairs, the
+pruning methods are able to continue increasing $k$ much farther, but on several
+pairs, pruning only manages to increase $k$ by one beyond the non-pruning algorithms.
+We also observed that pruning does yield speedups, although these are
+less pronounced than the differences in the number of tuples.  Nonetheless, pruning
+overcomes the major bottleneck, which is that standard $k$-limited analyses run
+out of memory even for moderate $k$.  By curbing the growth of the
+number of tuples, pruning makes it possible to run some analyses at all.
+
+However, there are several caveats with pruning:
+First, we are using BDDs, which can actually handle large numbers of tuples
 so long as they are structured; pruning destroys some of this structure,
-yielding less predictable running times.
+yielding less predictable running times.  Second, pruning requires solving
+the transformed Datalog program for computing $\bP(X)$, which is more expensive than the original Datalog program.
+Finally, we have to solve the Datalog program several times, not just once.
+These three caveats also apply to the site-based refinement algorithm of \cite{liang11minimal},
+so pruning is a strict improvement over that.
 
@@
 
-\reftab{sizeRatio} provides more on the quantitative impact of pruning in our
-best instantiation of the \PR\ algorithm.  We see that pre-pruning has a
-significant positive impact: we are able to perform pruning at a much cheaper
-level (using types instead of allocation sites).  Many of the tuples are pruned
-at this level, and the results of this pruning carry over to the original
-$k$-limited abstraction.
+\reftab{sizeRatio} provides more details on the quantitative impact of pruning in our
+best instantiation of the Prune-Refine algorithm ($\PR(\bpi, \hclassIsHas)$).  We see that pre-pruning has a
+significant positive impact: we are able to perform eliminate many tuples by just operating on the coarser
+level of types rather than allocation sites (see the $\frac{|\tilde B_t|}{|B_t|}$ column).  Many of the tuples are pruned
+at the type level, and the results of this pruning carry over to the original
+$k$-limited abstraction (see the $\frac{|\tilde A_t'|}{|A_t|}$ column).
 
@@
 First, $\frac{|B_t|}{|A_t|}$ measures the number of tuples after projecting
-down to the auxiliary abstraction $\beta_t$ for pre-pruning; note that running the analysis
+down to the auxiliary abstraction $\beta_t = \klimdabs_t \circ \hclassIsHas$ for pre-pruning; note that running the analysis
 using types instead of allocation sites is much cheaper.
@@
 ($\frac{|A_t'|}{|A_t|}$).
-Recall that we do the pruning step after pre-pruning;
-$\frac{|\tilde A_t|}{|A_t|}$ measures the total fraction of chains kept during pruning.
+After pre-pruning, pruning kept $\frac{|\tilde A_t|}{|A_t|}$ of the chains.
 Finally, $\frac{|A_{t+1}|}{|A_t|}$ measures the ratio between iterations,
@@
 
-So far, we have been using the $k$-limited abstraction.  In
-\reffig{barelyRepeatingEffect}, we show that when we can increase the $k$
-value, the barely-repeating $k$-limited abstraction is helpful, but in most
+So far, we have been using the $k$-limited abstraction; we now compare this abstraction
+with the barely-repeating $k$-limited abstraction introduced in \refsec{barelyAbstraction}.
+As \reffig{barelyRepeatingEffect} shows, we found that when we can increase the $k$
+value, the barely-repeating $k$-limited abstraction reduces the number of tuples, but in most
 cases, it does not improve scalability.  The reason is that the
-barely-repeating abstraction curbs refinement, but often, and quite
-paradoxically, it is exactly refinement which enables us to prune more.
+barely-repeating abstraction curbs refinement, but often, and somewhat
+paradoxically, it is exactly the refinement which enables us to prune more.
 
@@
 limiting repetitions is actually slightly worse in terms of scalability.
+Three of the lines for $\Klimdabs$ stop early, not because the algorithm runs out of memory,
+but because the algorithm has actually converged and increasing $k$ would have no effect.
 }
@@
 While pruning enables us to increase $k$ much more than before, it turns out
-that our particular analyses for these clients saturate quite quickly, so we
-are only able to prove two more queries than using the non-pruning techniques.
+that our particular analyses for these clients saturate quite quickly, so
+over all the clients and benchmarks,
+we were only able to prove two queries more than using the non-pruning techniques.
 On the surface, these findings seem to contradict \cite{liang10abstraction},
@@
 are insufficient.  However, given that our pruning approach is general, it
-would be interesting tackling other aspects such as flow-sensitivity.
+would be interesting to tackle other aspects of program analysis such as
+flow-sensitivity.
 
@@
 All analyses obtain the exact results on iterations where they obtain an answer.
-Bolded numbers are ones obtained by our best method $\PR(\Klimabs,\hclassIsHas)$
-but not by the full algorithm.
-We see that although pruning enables to increase $k$ by much more than before,
-only for two the client/benchmark pairs do we get strictly more precise results.
+Bolded numbers refer to $k$ values reached by $\PR(\Klimabs,\hclassIsHas)$
+but not by any non-pruning algorithm.
+While pruning enables to increase $k$ more, we get strictly more precise results
+for only two of the client/benchmark pairs (\downcast/\hedc\ and \downcast/\lusearch).
 This points out inherent limitations of this family of $k$-limited abstractions.

### 2011-03-20 extra/sliver_pldi11/klimited.tex

@@
 %\centering
-\caption{Datalog implementation of our $k$-object-sensitivity points-to analysis with call-graph construction.
+\caption{Datalog implementation of our $k$-object-sensitivity pointer analysis with call-graph construction.
 Our abstraction $\ba$ affects the analysis solely through $\ext$,
@@
 thread-escaping data, may point to the same object, and may happen in parallel.
-All four components rely heavily on the context- and object-sensitive points-to
+All four components rely heavily on the context- and object-sensitive pointer
 analysis.

### 2011-03-20 extra/sliver_pldi11/related.tex

@@
 Hendren~\cite{LhotakHendren2006,LhotakHendren2008} present Paddle, a
-parametrized framework for BDD-based, $k$-limited alias analysis.
+parametrized framework for BDD-based, $k$-limited pointer analysis.
 \cite{smaragdakis11context} scale up $k$-object-sensitivity, increasing $k$ by one
@@
 %varying precision, demand-driven approaches compute a partial solution of fixed
-%precision.  Heintze and Tardieu's demand-driven alias analysis for
+%precision.  Heintze and Tardieu's demand-driven points-to analysis for
 %C~\cite{HeintzeTardieu2001} performs a provably optimal computation to

### 2011-03-20 extra/sliver_pldi11/introduction.tex

@@
 Making a static analysis more precise requires increasing the complexity of the
-supporting abstraction---in pointer analysis, by increasing the amount of
+underlying abstraction---in pointer analysis, by increasing the amount of
 context/object sensitivity \cite{kcfa, kobj, MilanovaRountevRyder2005,
@@
 However, the complexity of these analyses often grows exponentially as the
-abstraction increases.
+abstraction is refined.
 Much work has been done on curbing this exponential growth
 (e.g., client-driven \cite{GuyerLin2003} and demand-driven
-\cite{HeintzeTardieu2001} approaches in pointer analysis,
+\cite{HeintzeTardieu2001} approaches in pointer analysis;
 lazy abstraction \cite{henzinger02lazy,mcmillan06lazy} and other iterative refinement approaches in model checking).
-The general theme of all these approaches is that the
-complexity of the abstraction is increased only in parts deemed relevant using
-some feedback from a client query.
+We refer to these techniques as {\em selected refinement},
+where the main idea is to only refine an abstraction along components
+deemed relevant according to client feedback.
 
-% Pruning idea: general
-In this paper, we use {\em pruning}, a different approach from
-the above selected refinement techniques for making static analyses scale.
-Our approach works on any analysis that can be expressed as a Datalog program.
-A Datalog program takes a set of input tuples (determined by the abstraction)
-and derives new tuples via a set of inference rules; the answer to a client
-query corresponds to whether a designated tuple can be derived.  The key idea
-is to identify input tuples which are provably irrelevant for deriving the
-query tuple and prune these away.  Selected refinement techniques
-attempt to keep the set of input tuples small by not refining some of them; we
-keep the set small by removing some of them entirely.
+% Pruning setup
+In this paper, we introduce {\em pruning}, a new approach which represents a
+significant departure from existing selected refinement techniques.  Pruning is
+applicable to static analyses expressed as a set of inference rules where a
+program property of interest (a client query) is proven by the inability to
+derive a designated fact using the given rules.  For concreteness, assume that
+the static analysis is expressed as a Datalog program.  A Datalog program takes
+a set of input tuples and derives new tuples via a set of inference rules.
+These inference rules capture the abstract semantics of the static analysis and
+the evaluation of the client query using the analysis result; the input tuples encode the program we are
+analyzing and the abstraction we are using.  The program property is proven if the
+designated query tuple cannot be derived.
+
+% Pruning idea
+The key idea behind pruning is to identify input tuples which are provably
+irrelevant for deriving the query tuple and remove these tuples completely from
+analysis.  Consequently, when the abstraction is refined, only the relevant
+tuples are refined, resulting in potentially major computational savings.  It
+is helpful to think of pruning in terms of program slicing, where
+irrelevant parts of the program (irrelevant input tuples) are removed,
+resulting in a smaller program that is cheaper to analyze.  Existing selected
+refinement techniques attempt to keep the set of input tuples small by simply
+not refining some of them; pruning keeps the set small by removing some of them
+entirely.
 
 % Soundness
-Pruning can be a dangerous affair though.  With selected refinement, at any
-point we are still performing static analysis with respect to an abstraction
-and therefore have soundness guarantees.  However, once we start pruning input
-tuples, we are no longer running a valid static analysis and thus do not
-automatically inherit soundness guarantees.  Nonetheless, we prove that our
-method is {\em sound with respect to the client}.
+Pruning can be a dangerous affair though.  With selected refinement,
+we are always performing a static analysis with respect to an abstraction
+and therefore inherit the soundness guarantees of abstract interpretation.
+However, once we start pruning input tuples, we are no longer running a valid
+static analysis of the original program.  Soundness therefore is no longer automatic,
+though we do prove that our method is sound with respect to the given client.
 
 % Completeness
-While soundness is trivial for selected refinement but requires some argument for pruning,
-the situation is reversed for {\em completeness}, that is, the guarantee that
-an analysis is as precise as some target.  For selected refinement,
-it is difficult to argue that the chosen coarser abstraction is as precise as if we
-had refined everything.  However, with pruning, we are working directly with
-the more complex abstraction, and thus automatically inherit its completeness
-guarantees.  By removing input tuples, we can only prove more queries.
+While soundness is trivial for selected refinement but requires some
+argument for pruning, the situation is reversed for {\em completeness}.  By
+completeness, we mean that the analysis is as precise as
+if we had refined all the components of an abstraction.  Selected
+refinement only refines a subset of an abstraction, so it is unclear
+that the resulting abstraction is as precise as an abstraction obtained by
+refining all components.  However, with pruning, we conceptually work with the
+fully-refined abstraction; and by removing input tuples, we cannot prove fewer
+queries; thus, completeness is automatic.
 
 % Refinement algorithm
-%To capitalize on the idea of {\em pruning}, we propose an algorithm, which we
-%call Refinement-Pruning (\PR).  The idea is to start with a coarse abstraction
-%and prune the input tuples before moving to a more expensive abstraction.
+To capitalize on the idea of pruning, we propose an algorithm, which we
+call the {\em Prune-Refine algorithm}.  The idea is to start with a coarse abstraction
+and prune the irrelevant input tuples before refining the abstraction; the
+algorithm iterates until the query is proven or a specified computational budget
+is reached.  We prove that the Prune-Refine algorithm computes the same answers to client queries
+as directly using a refined abstraction without pruning, which would be precise
+but possibly infeasible.
 
 % Slivers
-We apply our pruning technique to the $k$-object-sensitivity abstraction
+We apply pruning to the $k$-object-sensitivity abstraction
 \cite{kobj}, where objects in the heap are abstracted using chains of
-allocation sites.  Pruning in this setting corresponds to maintaining a subset
-of these chains.  Although the main contribution of the paper is the general
+allocation sites; these chains are the input tuples we maintain.
+Although the main contribution of this paper is the general
 pruning technique, we also make the following contributions specific to
-$k$-limited pointer analysis which arise because we are pruning: First, we show
-that we need a more careful treatment of the $k$-object-sensitivity abstraction
-than before due to pruning.  Second, we introduce an abstraction which limits
-chains to length $k$ and also truncates to avoid having too many repeating
-sites; this is an effective way to increase $k$ further (possible now due to
-pruning) without getting bogged down by recursion.  Third, we show that an
-abstraction that replaces allocation sites by types (a generalization of
+$k$-limited pointer analysis which arise due to pruning: First, we show
+that we need a more careful treatment of the $k$-object-sensitivity abstraction.
+Second, we introduce a new abstraction which limits
+chains to length $k$ and truncates to avoid repeating
+allocation sites; this is an effective way to increase $k$ without getting bogged down by long chains created by recursion.
+Third, we show that an abstraction that
+replaces allocation sites by types (a generalization of
 \cite{smaragdakis11context}) is effective for pruning.
@@
 checking, monomorphic call site inference, and race detection.
-We show that with pruning, our \PR\ algorithm enables us to perform $k$-object-sensitivity
+We show that with pruning, our Prune-Refine algorithm enables us to perform $k$-object-sensitivity
 analysis with a substantially much finer abstraction (larger $k$)
@@
 refinement strategy of \cite{liang11minimal}.
-In some cases, the non-pruning approaches hit a wall around $k=3$ but the \PR\
+In a few cases, the non-pruning approaches hit a wall around $k=3$ but the Prune-Refine
 algorithm is able to go well beyond $k=10$.

### 2011-03-20 extra/sliver_pldi11/preliminaries.tex

@@
 express static analyses declaratively \cite{Whaley2007,bravenboer}.  Normally,
-these analyses are encoded by one monolithic Datalog program which computes an
-abstract answer to a query.  For us, it will be convenient to consider the
-Datalog program, which computes the concrete\footnote{We refer to this computation as ``concrete''
+these analyses and their underlying abstractions are encoded by one monolithic Datalog program
+which evaluates a query abstractly.
+For us, it will be convenient to consider the
+Datalog program, which evaluates a query concretely\footnote{We refer to this computation as ``concrete''
 to contrast with abstract computation we will consider later,
@@
 without abstraction and $k$-object-sensitivity with abstraction.
-}
-answer to a query, as distinct
-from the abstraction, which is applied on the Datalog program to compute an
-abstract answer.  This separation allow us to make theoretical statements
+}, as distinct
+from the abstraction, which is defined as a transformation on the input to
+the Datalog program, and makes it evaluate the query abstractly.  This separation
+allows us to make theoretical statements
 comparing the behavior of the same Datalog program across different
@@
 \]
-\caption{\label{fig:repeatingExample} For the $k$-limited abstraction
-with $\dom = \{ 0, 1 \}$ and $k = 3$, the equivalence classes defined by $\klimabs_k$.
+\caption{\label{fig:repeatingExample}
+The equivalence classes defined by the $k$-limited abstraction $\klimabs_k$
+with $\dom = \{ 0, 1 \}$ and $k = 3$,
+In other words, $\klimabs_k$ maps each chain $c \in \dom^*$ to one of the 14 abstract values above.
 }

### 2011-03-20 extra/sliver_pldi11/abstract.tex

--- 2011-03-18 extra/sliver_pldi11/abstract.tex
@@
 Many static analyses do not scale as they are made more precise.  For example,
-increasing the amount of context sensitivity in a $k$-limited analysis causes
-the number of contexts to grow exponentially.  Iterative refinement techniques
-can be employed to mitigate this growth by starting with a coarse abstraction and
-only refining parts of the abstraction deemed relevant with respect to a client.
+increasing the amount of context sensitivity in a $k$-limited pointer analysis causes
+the number of contexts to grow exponentially with $k$.  Iterative refinement techniques
+can mitigate this growth by starting with a coarse abstraction and
+only refining parts of the abstraction that are deemed relevant with respect to a given client.
 
-In this paper, we introduce a new way to use this client feedback to greatly
-reduce the complexity of an analysis: {\em pruning}.  The pruned analysis is no
-longer a valid abstraction, but we prove that we do not forfeit correctness and
-that our analysis is still sound with respect to a given client.  In the
-context of $k$-limited analysis, our approach amounts to keeping track of a set
-of context prefix patterns.  By iteratively refining and pruning these
-patterns, we show that our analysis increases scalability greatly.
+In this paper, we introduce a new technique called {\em pruning} that uses client
+feedback in a different way.  The basic idea is to use coarse
+abstractions to prune away parts of the program analysis deemed irrelevant for proving a
+client query, and then using finer abstractions on the sliced program analysis.
+%In contrast to existing iterative refinement techniques,
+%our analyses do not correspond to abstractions on the original program but rather one on a sliced program.
+%We prove that we do not forfeit correctness and
+%that our analysis is still sound with respect to a given client. 
+For a $k$-limited pointer analysis, this approach amounts to adaptively refining and pruning a set
+of prefix patterns representing the contexts relevant for the client.
+By pruning, we are able to scale up to much more expensive abstractions than before.
+We also prove that the pruned analysis is both sound and complete, that is, it yields the same results
+as an analysis that uses a more expensive abstraction directly without pruning.
 

### 2011-03-20 extra/sliver_pldi11/theory.tex

@@
 determining if the query is proven (when $\bP$ returns $\emptyset$); and (ii)
-returning the subset of relevant input tuples.
+returning the relevant input tuples.
 The following theorem provides the key equation that drives everything in this paper (see \refapp{proofs} for the proof):
@@
 consider the following simpler idea:
-first run the analysis with $\beta$; if the query is proven, stop and declare {\em proven}; otherwise, we run the analysis
+first run the analysis with $\beta$; if the query is proven, stop and declare {\em proven}; otherwise, run the analysis
 with $\alpha$ and output that answer.
-It is easy to see that this two-step procedure returns the same answer as just running $\alpha$,
-because $\beta \preceq \alpha$ implies that if $\beta$ proves the query, then so does $\alpha$
+It is easy to see that this two-step procedure returns the same answer as just running $\alpha$:
+Because $\beta \preceq \alpha$, if $\beta$ proves the query, then so does $\alpha$
 (\refprop{soundness}).
 \refeqn{master} can be thought of as an extension of this basic idea:
-instead of using $\beta$ to just determine where the query tuple is proven,
-we obtain more information, namely the set of input tuples that are relevant.
+instead of using $\beta$ to just determine whether the query tuple is proven,
+we obtain more information, namely the whole set of input tuples that are relevant.
 
@@
 (e.g., $\alpha_t = \pi_t$ for the $k$-limited abstractions)
-and an input set of abstract tuples $A_0$, which is computed
+and a set of input abstract tuples $A_0$, which is computed
 under the initial abstraction $\alpha_0$.  Then the algorithm alternates between a pruning step
-and a refining step, always maintaining only the abstract tuples
+and a refining step, maintaining only the abstract input tuples
 that could participate in a derivation of the query tuple $\xo$.
@@
 (ii) then we use $\alpha_t$ during the main pruning step.
-Pre-pruning does not affect the output of the pruning step ($\tilde A_t$);
-the purpose is solely to make the pruning step faster.
+%Pre-pruning does not affect the output of the pruning step ($\tilde A_t$);
+%the purpose is solely to make the pruning step faster.
 
 Pre-pruning requires a sequence of auxiliary abstractions $(\beta_t)$ which are coarser than the main abstractions $(\alpha_t)$.
-We can obtain auxiliary abstractions by composing the main abstractions
-with another abstraction $\hclass$ ($\beta_t = \alpha_t \circ \hclass$).
+A standard way to obtain auxiliary abstractions is by composing the main abstractions
+with another abstraction $\hclass$; formally, $\beta_t = \alpha_t \circ \hclass$.
 We can use any $\hclass$ for which $\alpha_t \circ \hclass$ yields a valid abstraction,
@@
 A good rule of thumb is that $\hclass$ should be ``complementary'' to $\alpha_t$
-(we will see some examples in \refsec{abstractions}).
+(we will see some examples in \refsec{abstractionExamples}).
 
 \refthm{algorithm} states that the Prune-Refine algorithm is both
-sound and complete.  In other words, pruning has no impact on our answer to a
-query---we just compute it more efficiently.
+sound and complete.  In other words, pruning has no impact on the answer to a
+query.
 The proof is given in \refapp{proofs} and uses \refthm{master}.
@@
 At iteration $t$,
-the incrementally pruned abstraction $A_t$ is equivalent
-to the full abstraction $\alpha_t(X)$ (formally, $\bP(\alpha_t(X)) = \bP(A_t)$).
-If the algorithm returns ``proven,'' then $\bP(X) = \emptyset$ (the query is actually false).
+using the incrementally pruned abstraction $A_t$ is equivalent
+to using the full abstraction $\alpha_t(X)$ in that $\bP(\alpha_t(X)) = \bP(A_t)$.
+Consequently, if the algorithm returns ``proven,'' then $\bP(X) = \emptyset$ (the query is actually false).
 \end{theorem}

### 2011-03-21 extra/sliver_pldi11/abstractions.tex

@@
 
-Recall from \refsec{preliminaries} that we separate the Datalog program that
-computes the query from the abstraction.  In \refsec{klimited}, we described
-the Datalog programs that computed the queries for our three clients based on
-$\infty$-object-sensitivity.  We now describe the abstractions that we use to
-further abstract this analysis.
+%Recall from \refsec{preliminaries} that we separate the Datalog program that
+%computes the query from the abstraction.  In \refsec{klimited}, we described
+%the Datalog programs that computed the queries for our three clients based on
+%$\infty$-object-sensitivity.  We now describe the abstractions that we use to
+%further abstract this analysis.
 
-The basic abstraction (\refsec{klimitedAbstraction}) corresponds exactly to
-$k$-object-sensitivity.  We then present two orthogonal variants: one that in
-addition limits the repetition of allocation sites (\refsec{barelyAbstraction})
+We have already defined the $k$-limited abstraction, which corresponds to
+$k$-object-sensitivity.  We now present two orthogonal variants of this basic abstraction: one that
+additionally limits the repetition of allocation sites (\refsec{barelyAbstraction})
 and one that further abstracts allocation sites using type information
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\Subsection{klimitedAbstraction}{$k$-limited abstraction}
-
-The {\em $k$-limited abstraction}
-essentially partitions chains based on their length $k$ prefix.
-We first setup some notation.
-%Let $\dom$ be a set of values (e.g., $\dom = \{ 0, 1 \}$).
-For a chain $c \in \dom^*$, let $|c|$ denote the length of the chain.
-Let $c[i]$ be the $i$-th element of $c$ (starting with index 1) and
-let $c[i..j]$ be the subchain $[c[i] \cdots c[j]]$
-(boundary cases: $c[i..j] = [\,]$ if $i > j$ and $c[i..j] = c[i..|c|]$ if $j > |c|$).
-For two chains $c_1, c_2$, let $c_1+c_2$ denote concatenation.
-For a chain $c$, let $c*$ denote the set of all chains
-with prefix $c$; formally:
-\begin{align}
-c* \eqdef \{ c' \in \H^* : c'[1..|c|] = c \}.
-\end{align}
-%It is important that $c*$ is a set.
-
-For an integer truncation level $k \ge 0$, define the abstraction $\klimabs_k$ as follows:
-\begin{align}
-\label{eqn:klimitedAbstraction}
-\klimabs_k(c) \eqdef
-\begin{cases}
-\{c\}     & \text{if $|c| < k$} \\
-c[1..k]*  & \text{if $|c| \ge k$}. \\
-\end{cases}
-\end{align}
-If the concrete chain $c$ is shorter than length $k$,
-we map it to the singleton set $\{ c \}$;
-otherwise, we map it to the set of chains that share the first $k$ elements.
-It is easy to verify that $\klimabs_k$ as defined is a valid abstraction under \refdef{abstraction}.
-\reffig{repeatingExample} shows an example.
-
-\begin{figure}
-\[
-\arraycolsep=3.5pt
-\begin{array}{llllllll}
-\{\vec{0}\}    &            &              &            & \{\vec{1}\} \\
-\{\vec{00}\}   &            & \{\vec{01}\} &            & \{\vec{10}\} &            & \{\vec{11}\} \\
-\vec{000}*     & \vec{001}* & \vec{010}*   & \vec{011}* & \vec{100}*   & \vec{101}* & \vec{110}* & \vec{111}*
-\end{array}
-\]
-\caption{\label{fig:repeatingExample} For the $k$-limited abstraction
-with $\dom = \{ 0, 1 \}$ and $k = 3$, the equivalence classes defined by $\klimabs_k$.
-}
-\end{figure}
-
-Traditionally, $k$-limited analyses
-simply use a list of sites ``$c$'' to represent an abstract value,
-not making a distinction between $\{c\}$ and $c*$;
-however when we do pruning, this distinction is crucial.
-Suppose that $k=3$.
-Na\"ively, ``$\vec{01}$'' would just represent $\{\vec{01}\}$.
-The problem comes when we prune.
-If we prune ``$\vec{010}$'', then ``$\vec{01}$'' really represents $\{\vec{01}\}\cup \vec{010}*$
-because any chain starting with $\vec{010}$ must now fall back on ``$\vec{01}$''.
-This is unacceptable because pruning tuples should not change the abstraction.
-Actually, the situation is worse, because (i) in fact no pruning
-(in the sense that some chains are ignored)
-would happen and (ii) the resulting analysis would be much more imprecise because
-``$[\,]$'' is always present for representing the context of the main method,
-but this is a prefix of any chain.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \Subsection{barelyAbstraction}{Barely-repeating $k$-limited abstraction}
 
-When we applied the $k$-limited abstraction empirically,
-we noticed that a major reason why it did not scale was
-the seemingly unnecessary combinatorial explosion associated chains formed
+When we applied the $k$-limited abstraction in practice,
+we noticed empirically that a major reason why it did not scale was
+the seemingly unnecessary combinatorial explosion associated with chains formed
 by cycling endlessly through the same allocation sites.
@@
 than $\infty$-object-sensitivity.}
-We therefore wish to define an abstraction that not only truncates at $k$ but also truncates
+We therefore wish to define an abstraction that not only truncates chains at length $k$ but also truncates
 a chain when it starts repeating.
@@
 and the points-to set of {\tt y2} is $\{\vec{02},\vec{002},\vec{0002},\dots\}$.
-These two sets are disjoint, but for any finite $k$, no matter how large,
-$k$-object-sensitivity would conclude that both point to $\bzero_k*$, where $\bzero_k$ is a chain of $k$ zeros.
+While these two sets are disjoint, if we use a $k$-limited abstraction for any finite $k$,
+we would conclude erroneously that both variables might point to $\bzero_k*$, where $\bzero_k$ is a chain of $k$ zeros.
+Incidentally, this demonstrates an intrinsic limitation of the $k$-object-sensitivity abstraction.
+Using the barely-repeating $k$-limited abstraction, we can increase $k$ while avoiding chains longer than $\vec{00}$
+since $\vec{00}$ is barely-repeating.
+This results in computational savings, and in this case, in no loss in precision.
 }
@@
 For a sequence $c$, we say $c$ is {\em non-repeating} if all its elements are distinct.
-We say $c$ is {\em barely-repeating} if $c$ excluding the last element ($c[1..|c|-1]$)
-is {\em non-repeating}.  Note that a barely-repeating includes the non-repeating case.
-Let $\delta(c)$ be the length of the longest prefix that is barely-repeating:
+We say $c$ is {\em barely-repeating} if (i) $c$ excluding the last element ($c[1..|c|-1]$)
+is {\em non-repeating} and (ii) the last element of $c$ is repeated earlier in $c$.
+%We say $c$ is {\em barely-repeating} if $c$ excluding the last element ($c[1..|c|-1]$) is {\em non-repeating}.
+%Note that a barely-repeating includes the non-repeating case.
+Let $\distinct(c)$ be the length of the longest prefix of $c$ that is barely-repeating, if it exists, and $\infty$
+otherwise:
 \begin{align}
 \label{eqn:barelyLen}
-\distinct(c) \eqdef \max_{m' : c[1..m'] \text{ is barely-repeating}} m'.
+\distinct(c) \eqdef
+\begin{cases} 
+\max_{m' : c[1..m'] \text{ is barely-repeating}} m' & \text{if $m'$ exists}, \\
+\infty & \text{otherwise}.
+\end{cases}
 \end{align}
+%\begin{align}
+%\label{eqn:barelyLen}
+%\distinct(c) \eqdef \max_{m' : c[1..m'] \text{ is barely-repeating}} m'
+%\end{align}
 For example, $\distinct(\vec{10010}) = 3$ because $\vec{100}$ is barely-repeating, but $\vec{1001}$ is not.
 
-Then we define the {\em barely-repeating $k$-limited abstraction} $\klimdabs_k$ as follows:
+We now define the {\em barely-repeating $k$-limited abstraction} $\klimdabs_k$ as follows:
 \begin{align}
 \label{eqn:barelyAbstraction}
-\klimdabs_k(c) & \eqdef \klimabs_{\min\{k,\delta(c)\}}(c),
+\klimdabs_k(c) & \eqdef \klimabs_{\min\{k,\distinct(c)\}}(c),
 \end{align}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\Subsection{classAbstraction}{Class-based abstraction}
+\Subsection{classAbstraction}{Type-based abstraction}
 
-We now introduce an abstraction that we will use in the pre-pruning step of the \PR\ algorithm.
+We now introduce an abstraction that we will use in the pre-pruning step of the Prune-Refine algorithm.
 We start by defining an equivalence relation over allocation sites $\dom$,
 represented by a function $\hclass : \dom \mapsto \sP(\dom)$
-mapping each allocation site $h \in \dom$ to its its equivalence class.
-Given such an $\hclass$, we could then extend it sequences by elementwise application:
+mapping each allocation site $h \in \dom$ to its equivalence class.
+In the graph example, we might have
+$\hclass(0) = \hclass(1) = \{ 0,1 \}$
+and $\hclass(2) = \hclass(3) = \{ 2,3 \}$.
+
+Given such a $\hclass$, we extend it to sequences by taking the cross product over elementwise applications:
 \begin{align}
@@
 \end{align}
+In the running example, $\hclass([02]) = \{ [02], [03], [12], [13] \}$.
 
-To construct $\hclass$, we consider two sources of type information associated
+To construct $\hclass$ for $k$-limited pointer analysis, we consider using two sources of type information associated
 with an allocation site, motivated by \cite{smaragdakis11context}:
@@
 \end{align}
-with $f \in \{ \typeIs, \typeHas, \typeIs \times \typeHas \}$.
+for $f \in \{ \typeIs, \typeHas, \typeIs \times \typeHas \}$.
 
 % Complementary
-Now we have three options for $\hclass$ which are complementary to the
-$k$-limited abstraction: $\hclass$ abstracts a chain by abstracting each site
-uniformly; $\pi_k$ keeps all sites up to a length $k$ prefix precisely but
-abstracts away the rest completely.  This complementarity is desirable
-for pre-pruning.
-However, $\hclass$ is not coarser than $\klimabs_k$, and therefore cannot
-be used directly in the \PR\ algorithm.
+Now we have three choices for $\hclass$: one that uses the declaring type ($\hclassIs$),
+one that uses the type of the containing class ($\hclassHas$), and one that uses both ($\hclassIsHas$).
+Recall that all three are complementary to the
+$k$-limited abstraction $\pi_k$: Specifically, $\hclass$ abstracts a chain by abstracting each site
+uniformly, whereas $\pi_k$ performs no abstraction on the first $k$ sites, but performs a total abstraction
+on the rest of the chain.  Recall that this complementarity is desirable for effective pre-pruning.
 
-However, we can compose $\hclass$ and $\klimabs_k$ to yield another abstraction which will be coarser than $\klimabs_k$.
-But in what order should we compose?
-We must take care because the composition of two equivalence relations is not necessarily an equivalence relation.
-The following proposition shows which orderings are valid:
+Since $\hclass$ is not coarser than $\klimabs_k$, we cannot
+use it directly in the Prune-Refine algorithm.
+We must compose $\hclass$ with $\klimabs_k$ or $\klimdabs_k$ to yield another
+abstraction which is coarser than $\klimabs_k$ or $\klimdabs_k$,
+respectively.  But in what order should we compose?
+We must be careful because the composition of two abstractions is not necessarily an abstraction.
+Fortunately, the following proposition shows which compositions are valid:
 \begin{proposition}
@@
 
-For (i) and (ii), it is straightforward to check
+For (i) and (ii), it is straightforward to check that
 $(\klimabs_k \circ \hclass)(c) = (\hclass \circ \klimabs_k)(c) = s$ for each $c \in s$.
@@
 let $\dom = \{ 0,1,2 \}$,
-and define $\hclass(v) = \dom$ for all $v \in \dom$
+and define $\hclass(h) = \dom$ for all $h \in \dom$
 (there is one equivalence class).
-Consider two elements $\vec{01}$ and $\vec{00}$ under the composed abstraction:
+Consider applying $\hclass \circ \klimdabs_3$ to two elements $\vec{01}$ and $\vec{00}$:
 For $\vec{01}$, we have $\klimdabs_3(\vec{01}) = \{ \vec{01 }\}$,
@@
 the two sets are neither equal nor disjoint),
-so $\hclass \circ \klimdabs_3$ does not define an equivalence relation.
+so $\hclass \circ \klimdabs_3$ does not define a valid abstraction.
 \end{proof}
 
+\noindent 
 In light of this result, we will use the valid abstractions $\cklimabsk$ and $\cklimdabsk$,
-which work by first applying $\hclass$ and then applying $\klimabs_k$ or $\klimdabs_k$.
+which work by first applying the type-based abstraction $\hclass$ and then applying $\klimabs_k$ or $\klimdabs_k$.
 

### 2011-03-21 extra/sliver_pldi11/conclusion.tex

@@
 
-We have introduced pruning as a general way of scaling up static analyses
-written as Datalog programs.  The key idea is to run an analysis using a coarse
-abstraction, only keeping input tuples deemed relevant for a finer abstraction.
+We have introduced pruning as a general technique for scaling up static analyses
+written in Datalog.  The basic idea is to run an analysis using a coarse
+abstraction, only keeping input tuples deemed relevant, and then using a finer abstraction on the remaining tuples.
 Theoretically, we showed that pruning is both sound and complete (our analysis
-is valid and we lose no precision).  Empirically, we showed that our pruning
-approach enables us to scale up analyses based on $k$-object-sensitivity much
+is valid and we lose no precision).  Empirically, we showed that pruning
+enables us to scale up analyses based on $k$-object-sensitivity much
 more than previous approaches.
+
+\acks
+We thank Mooly Sagiv and Hongseok Yang for discussion and useful feedback.
+We also thank the anonymous reviewers for their insightful comments.

### 2011-03-27 main/doc/architecture.tex
documentation

\xname{architecture}
\chapter{Architecture of Chord}
\label{chap:architecture}

This chapter presents the high-level architecture of Chord, depicted below, and describes its key components.

\begin{center}
\texonly{\includegraphics[scale=0.7]{chord_arch.png}}
\htmlonly{{\htmlimg{chord_arch.png}}}
\end{center}

\subsubsection*{Chord Properties}

All inputs to Chord are specified by means of system properties. 
Chapter \ref{chap:properties} describes how to set properties and the meaning of 
each property that is recognized by Chord.

The Java program to be analyzed is also specified via properties.
These properties are described in Section \ref{sec:program-props}.
Chord analyzes Java bytecode, not Java source code, and thus only requires
the program's class files.  Certain analyses, however, present
their results at the Java source code level, and thus require the program's
Java source files as well.

\subsubsection*{Java Program Representation}

Chord uses the \xlink{Joeq}{http://joeq.sourceforce.net/} Java compiler framework
to convert the Java bytecode of the input Java program, one class file at a time,
into a three-address-like intermediate program representation called {\it quadcode}
that is more suitable for analysis. Chapter \ref{chap:program-representation}
describes the quadcode representation in detail.

\subsubsection*{Computing Analysis Scope}

A pre-requisite to analyzing a Java program using any program analysis framework, including Chord, is
to compute the {\it analysis scope}: the set of methods of the program to be analyzed.  Chord implements
several standard scope construction algorithms from the literature that differ in aspects such as
scalability, precision, and usability for the problem at hand.  Chapter \ref{chap:computing-scope}
describes these algorithms in detail.

\subsubsection*{Writing and Running Analyses}

Chord provides various {\it analysis templates}: classes containing boilerplate code that can be
extended by users to rapidly prototype different kinds of analyses.  An example is class RHSAnalysis,
named after [Reps, Horowitz, and Sagiv 1995], which can be extended by users to write a summary-based inter-procedural
context-sensitive static analysis by merely specifying the abstract domain and intra-procedural transfer functions.
Another exampls is DynamicAnalysis, which can be extended by
users to write a dynamic analysis by merely specifying which of various provided events to instrument, and
the transfer functions for those events.

A distinctive aspect of Chord is that each analysis is written modularly, independent of other analyses, along with lightweight
annotations specifying the inputs and outputs of the analysis.
Users must provide a unique name to each analysis, which allows the desired analyses to be called from the
command-line, via a property named \code{chord.run.analyses}.
Chord's runtime automatically computes
producer-consumer relationships between analyses (e.g., determines which analysis produces as output a
result that is needed as input by another analysis).  Before running a desired analysis,
Chord recursively runs other analyses until the inputs to the desired analysis have been computed; it
finally runs the desired analysis to produce the outputs of that analysis.

Chord can be invoked in one of two modes: classic or modern (depending upon whether the value of property
\code{chord.classic} is \code{true} (default) or \code{false}, respectively).
These two modes defer in the semantics of dependencies between analyses.  In particular, the classic mode
is simpler to understand for novice users (the dependencies are only data dependencies) but has a sequential
runtime, whereas the modern mode is harder to understand (there are both data and control dependencies)
but has a parallel runtime that is capable of running analyses without dependencies between
them in parallel.  The parallel runtime is based on \xlink{Habanero-Java}{http://habanero.rice.edu/hj.html},
and the semantics of the dependencies between analyses is based on the
\xlink{Habanero Concurrent Collections (CnC)}{http://habanero.rice.edu/cnc.html} declarative parallel
programming model.

Chapter \ref{chap:analyses} describes the various analysis templates and how to connect disparate analyses
together via dependencies in either of the above two modes.

\subsubsection*{Dynamic Analysis}

Chord uses the \xlink{Javassist}{http://www.javassist.org/} Java bytecode manipulation
framework for instrumenting bytecode and doing dynamic analysis.
Chord offers the most versatile capabilities of any existing dynamic analysis framework
for Java, particularly the ability to instrument the entire JDK (including classes in
package java.lang).
Specifically, it includes support for:
\begin{itemize}
\item
offline as well as load-time instrumentation of Java bytecode;
\item
processing of dynamic analysis events online in the same JVM or offline in a different JVM with an uninstrumented JDK (the latter circumvents performance and correctness problems that can arise if a single JVM with an instrumented JDK is used to generate and handle events); and
\item
allowing the event-generating and event-handling JVMs to run either serially (by storing the entire trace of events to a regular file) or in parallel (by streaming the trace of events in a piped file).
\end{itemize}
Chapter \ref{chap:dynamic-analysis} describes all aspects of dynamic analysis in Chord.

\subsubsection*{Datalog Analysis}

An analysis in Chord can be written either imperatively in Java or declaratively in Datalog.
Chord uses the BDD-based Datalog solver \xlink{bddbddb}{http://bddbddb.sourceforge.net/} to
run analyes written in Datalog.
Chapter \ref{chap:datalog-analysis} describes all aspects of Datalog analysis.

%If the analyses are purely static, it suffices to specify the main class of the program
%(via property \code{chord.main.class}) and the application classpath of the program (via property \code{chord.class.path}).
%If the analyses need to present their results at the Java source level, then
%the path of the program's Java source files is also required (via property \code{chord.src.path}).
%Finally, if any analysis is dynamic, then the command-line arguments to be used for running the
%program must also be specified (via property \code{chord.run.ids} and properties \code{chord.args.<N>} for
%each item \code{<N>} in the value of property \code{chord.run.ids}).

### 2011-03-27 main/doc/computing_scope.tex
documentation

@@
+\xname{computing_scope}
 \chapter{Computing Analysis Scope}
@@
 The following sections describe Chord's analysis scope 
-computation in more detail.  Section \ref{sec:scope-reuse} describes how to reuse the analysis scope
+computation in detail.  Section \ref{sec:scope-reuse} describes how to reuse the analysis scope
 computed in a previous run of Chord for a given program.  Section \ref{sec:scope-algos} describes
@@
 is of the form:
+
+\begin{framed}
 \begin{verbatim}
@@
 \end{verbatim}
+\end{framed}
+
 where each of the above ``{\tt ...}'' is a list of zero or more lines, where
@@
 each call to static method {\tt forName(String)} defined in class
-{\tt java.lang.Class}, along with a list of the types of the named
+\code{java.lang.Class}, along with a list of the types of the named
 classes.
@@
 each call to instance method {\tt newInstance()} defined in class
-{\tt java.lang.Class}, along with a list of the types of the
+\code{java.lang.Class}, along with a list of the types of the
 instantiated classes.
@@
 each call to instance method {\tt newInstance(Object[])} defined in class
-{\tt java.lang.reflect.Constructor}, along with a list of the types of the
+\code{java.lang.reflect.Constructor}, along with a list of the types of the
 instantiated classes.
@@
 each call to instance method {\tt newInstance(Class,int)} defined in class
-{\tt java.lang.reflect.Array}, along with a list of the types of the
+\code{java.lang.reflect.Array}, along with a list of the types of the
 instantiated classes.

### 2011-03-27 main/doc/predefined_analyses.tex
documentation

@@
 We first recall some relevant program domains:
+
+\texonly{\newpage}
+
 \begin{itemize}
@@
 
+\begin{framed}
 \begin{verbatim}
-    prompt> ant -Dchord.work.dir=<...> -Dchord.run.analyses=cipa-0cfa-dlog run
+ant -Dchord.work.dir=<...> -Dchord.run.analyses=cipa-0cfa-dlog run
 \end{verbatim}
+\end{framed}
 
@@
 
+\begin{framed}
 \begin{verbatim}
-    prompt> ant -Dchord.work.dir=<...> -Dchord.run.analyses=datarace-java run
+ant -Dchord.work.dir=<...> -Dchord.run.analyses=datarace-java run
 \end{verbatim}
+\end{framed}
 
@@
 
+\begin{framed}
 \begin{verbatim}
-    prompt> ant -Dchord.work.dir=<...> -Dchord.run.analyses=deadlock-java run
+ant -Dchord.work.dir=<...> -Dchord.run.analyses=deadlock-java run
 \end{verbatim}
+\end{framed}
 

### 2011-03-27 main/doc/program_representation.tex
cleaned up and updated documentation
new documentation
documentation

@@
-\section{Program Representation}
-\label{sec:program-representation}
+\xname{program_representation}
+\chapter{Java Program Representation}
+\label{chap:program-representation}
 
-TODO
+Chord uses \xlink{Joeq}{http://joeq.sourceforge.net/} to translate
+Java bytecode, one class file at a time, into a three-address-like
+intermediate representation of the input Java program called {\it
+quadcode}.  This chapter describes how to pretty-print bytecode and
+quadcode, which is useful for debugging analyses and deciphering their
+output, as well as the API of Joeq and Chord that concerns bytecode
+and quadcode, which is useful for users writing their own analyses.
+
+\section{Pretty-Printing Bytecode and Quadcode}
+
+Consider the following Java program contained in
+file \code{examples/hello_world/src/test/HelloWorld.java} in Chord's
+main directory:
+\begin{framed}
+\begin{verbatim}
+package test;
+
+public class HelloWorld {
+    public static void main(String[] args) {
+        System.out.println("Hello World!");
+    }
+}
+\end{verbatim}
+\end{framed}
+
+First compile this program by running command \code{ant} in
+directory \code{examples/hello_world/}.
+
+To pretty-print the bytecode of a class, run the following command:
+\begin{framed}
+\begin{verbatim}
+javap -classpath <...> -private -verbose <CLASS NAME>
+\end{verbatim}
+\end{framed}
+
+where {\tt <CLASS NAME>} is the fully-qualified name of the class
+whose bytecode is to be printed (\code{test.HelloWorld} in our
+example) and \code{<...>} is the classpath of that class
+(\code{examples/hello_world/} in our example).  Program \code{javap}
+comes along with the JVM.  The output of the above command for our
+example is as follows:
+
+\begin{framed}
+\begin{verbatim}
+Compiled from "HelloWorld.java"
+public class test.HelloWorld extends java.lang.Object
+  SourceFile: "HelloWorld.java"
+  minor version: 0
+  major version: 49
+  Constant pool:
+const #1 = Method   #6.#20; //  java/lang/Object."<init>":()V
+const #2 = Field    #21.#22;    //  java/lang/System.out:Ljava/io/PrintStream;
+const #3 = String   #23;    //  Hello World!
+const #4 = Method   #24.#25;    //  java/io/PrintStream.println:(Ljava/lang/String;)V
+const #5 = class    #26;    //  test/HelloWorld
+const #6 = class    #27;    //  java/lang/Object
+const #7 = Asciz    <init>;
+const #8 = Asciz    ()V;
+const #9 = Asciz    Code;
+const #10 = Asciz   LineNumberTable;
+const #11 = Asciz   LocalVariableTable;
+const #12 = Asciz   this;
+const #13 = Asciz   Ltest/HelloWorld;;
+const #14 = Asciz   main;
+const #15 = Asciz   ([Ljava/lang/String;)V;
+const #16 = Asciz   args;
+const #17 = Asciz   [Ljava/lang/String;;
+const #18 = Asciz   SourceFile;
+const #19 = Asciz   HelloWorld.java;
+const #20 = NameAndType #7:#8;//  "<init>":()V
+const #21 = class   #28;    //  java/lang/System
+const #22 = NameAndType #29:#30;//  out:Ljava/io/PrintStream;
+const #23 = Asciz   Hello World!;
+const #24 = class   #31;    //  java/io/PrintStream
+const #25 = NameAndType #32:#33;//  println:(Ljava/lang/String;)V
+const #26 = Asciz   test/HelloWorld;
+const #27 = Asciz   java/lang/Object;
+const #28 = Asciz   java/lang/System;
+const #29 = Asciz   out;
+const #30 = Asciz   Ljava/io/PrintStream;;
+const #31 = Asciz   java/io/PrintStream;
+const #32 = Asciz   println;
+const #33 = Asciz   (Ljava/lang/String;)V;
+
+{
+public test.HelloWorld();
+  Code:
+   Stack=1, Locals=1, Args_size=1
+   0:   aload_0
+   1:   invokespecial   #1; //Method java/lang/Object."<init>":()V
+   4:   return
+  LineNumberTable:
+   line 3: 0
+  LocalVariableTable:
+   Start  Length  Slot  Name   Signature
+   0      5      0    this       Ltest/HelloWorld;
+
+public static void main(java.lang.String[]);
+  Code:
+   Stack=2, Locals=1, Args_size=1
+   0:   getstatic   #2; //Field java/lang/System.out:Ljava/io/PrintStream;
+   3:   ldc #3; //String Hello World!
+   5:   invokevirtual   #4; //Method java/io/PrintStream.println:(Ljava/lang/String;)V
+   8:   return
+  LineNumberTable:
+   line 5: 0
+   line 6: 8
+  LocalVariableTable:
+   Start  Length  Slot  Name   Signature
+   0      9      0    args       [Ljava/lang/String;
+}
+\end{verbatim}
+\end{framed}
+
+To pretty-print the quadcode of a class, run the following command:
+
+\begin{framed}
+\begin{verbatim}
+ant -Dchord.work.dir=<...> -Dchord.print.classes=<CLASS NAME> \
+    -Dchord.verbose=0 -Dchord.out.file=<FILE NAME> run
+\end{verbatim}
+\end{framed}
+
+where \code{<...>} is the path of a directory
+(\code{examples/hello_world/} in our example) that is expected to
+contain a file named \code{chord.properties} which defines
+properties \code{chord.main.class} and \code{chord.class.path}
+specifying the main class and the classpath of the input Java program
+(alternatively, these two properties can be defined directly on the
+above command line), {\tt <CLASS NAME>} is the name of the class whose
+quadcode is to be printed, and {\tt <...>/<FILE NAME>} is the name of
+a file to which the quadcode is to be dumped.  The output of the above
+command for our example is as follows:
+
+\begin{framed}
+\begin{verbatim}
+*** Class: test.HelloWorld
+Method: main:([Ljava/lang/String;)V@test.HelloWorld
+    0#1
+    5#3
+    5#2
+    8#4
+Control flow graph for main:([Ljava/lang/String;)V@test.HelloWorld:
+BB0 (ENTRY) (in: <none>, out: BB2)
+
+BB2 (in: BB0 (ENTRY), out: BB1 (EXIT))
+1: GETSTATIC_A T1, .out
+3: MOVE_A T2, AConst: "Hello World!"
+2: INVOKEVIRTUAL_V println:(Ljava/lang/String;)V@java.io.PrintStream, (T1, T2)
+4: RETURN_V
+
+BB1 (EXIT)  (in: BB2, out: <none>)
+
+Exception handlers: []
+Register factory: Registers: 3
+\end{verbatim}
+\end{framed}
+
+\section{Traversing Quadcode}
+
+\subsection{Program}
+
+\begin{framed}
+\begin{verbatim}
+package chord.program;
+
+class Program {
+    static Program g() {
+        // provides quadcode representation of program
+    }
+    public IndexSet<jq_Type> getTypes() {
+        // provides quadcode representation of all reachable types
+    }
+    public IndexSet<jq_Reference> getClasses() {
+        // provides quadcode representation of all reachable classes
+    }
+    public IndexSet<jq_Method> getMethods() {
+        // provides quadcode representation of all reachable methods
+    }
+    public Reflect getReflect() {
+        // provides information about resolved reflection
+    }
+    public jq_Reference getClass(String name) {
+        // provides quadcode representation of class 
+    }
+    public jq_Type getType(String name) {
+    }
+    public jq_Method getMainMethod() {
+    }
+    public jq_Method getMethod(String mName, String mDesc, String cName) {
+    }
+    public jq_Method getMethod(String sign) {
+    }
+    public Quad getQuad(MethodElem e) {
+    }
+    public Quad getQuad(MethodElem e, Class quadOpClass) {
+    }
+}    
+\end{verbatim}
+\end{framed}
+
+\subsection{Type}
+
+\begin{framed}
+\begin{verbatim}
+class jq_Type {
+}
+\end{verbatim}
+\end{framed}
+
+\subsection{Class}
+
+\begin{framed}
+\begin{verbatim}
+class jq_Reference {
+}
+\end{verbatim}
+\end{framed}
+
+\subsection{Field}
+
+\begin{framed}
+\begin{verbatim}
+class jq_Field {
+     public String getName();
+     public jq_Class getDeclaringClass();
+     public jq_Type getType() {
+     }
+     public String toString() {
+         // print this field in format fname:fdesc@cname where:
+         // - fname is the name of the field
+         // - fdesc is the descriptor of the field's type
+         // - cname is the name of the field's declaring class
+         // example: buf:[B@java.io.BufferedInputStream
+     }
+}
+\end{verbatim}
+\end{framed}
+
+\subsection{Method}
+
+\begin{framed}
+\begin{verbatim}
+class jq_Method {
+
+}
+\end{verbatim}
+\end{framed}
+
+Register Factory
+
+Control-Flow Graph
+
+Basic Block
+
+Quad
+
+\begin{framed}
+\begin{verbatim}
+Traversing a control-flow graph
+\end{verbatim}
+\end{framed}
+
+\section{Program Descriptors}
+
+MethodSign
+
+A field descriptor represents the type of a local variable or a
+(static or instance) field.  It is a series of characters generated by
+the grammar:
+
+\begin{framed}
+\begin{verbatim}
+FieldDescriptor : FieldType
+      FieldType : BaseType | ObjectType | ArrayType
+       BaseType : B | C | D | F | I | J | S | Z
+     ObjectType : L <classname> ;
+      ArrayType : [ ComponentType
+  ComponentType : FieldType
+\end{verbatim}
+\end{framed}
+
+The characters of \code{BaseType}, the {\tt L} and {\tt ;}
+of \code{ObjectType}, and the {\tt [} of
+\code{ArrayType} are all ASCII characters.
+The {\tt <classname>} represents a fully qualified class or interface
+name.  The interpretation of the field types is as shown in the below
+table:
+
+\begin{table}
+\htmlattributes*{table}{border=1 cellpadding=10}
+\begin{tabular}{l|l|l}
+BaseType Character	& Type	& Interpretation \\
+\hline
+{\tt B} &byte& signed byte \\
+{\tt C} &char& Unicode character \\
+{\tt D} &double& double-precision floating-point value \\
+{\tt F} &float& single-precision floating-point value \\
+{\tt I} &int& integer \\
+{\tt J} &long& long integer \\
+{\tt L<classname>;}	&reference& an instance of class {\tt <classname>} \\
+{\tt S}	&short& signed short \\
+{\tt Z}	&boolean& true or false \\
+{\tt [}	&reference& one array dimension
+\end{tabular}
+\end{table}
+
+For example, the descriptor of type {\tt int}
+ is simply {\tt I}.  The descriptor of an instance variable of
+ type \code{Object} is ``\code{Ljava/lang/Object;}''.  Note that the
+ internal form of the fully qualified name for class \code{Object} is
+ used. The descriptor of a multidimensional double array of type
+ ``\code{double[][][]}'' is ``\code{[[[D}''.
+
+A method descriptor represents the parameters that the method takes
+and the value that it returns:
+
+\begin{framed}
+\begin{verbatim}
+   MethodDescriptor : ( ParameterDescriptor* ) ReturnDescriptor
+ParameterDescriptor : FieldType
+   ReturnDescriptor : FieldType | V
+\end{verbatim}
+\end{framed}
+
+A parameter descriptor represents a parameter passed to a method.  A
+return descriptor represents the type of the value returned from a
+method.  The character \code{V} indicates that the method returns no
+value (its return type is void).  For example, the method descriptor
+for the method
+``\code{Object foo(int i, double d, Thread t)}'' is
+``\code{(IDLjava/lang/Thread;)Ljava/lang/Object;}''.
+Note that internal forms of the fully qualified names of \code{Thread} and
+\code{Object} are used in the method descriptor.  The method descriptor for
+\code{foo} is the same whether it is a static or an instance
+method. Although an instance method is passed \code{this}, a reference to the
+current class instance, in addition to its intended parameters, that
+fact is not reflected in the method descriptor.
+
+Chord uniquely represents each bytecode instruction in each method
+declared in each class of the program by a
+four-tuple \code{bci!mname:mdesc@cname} where \code{bci} is the
+(0-based) bytecode offset of the instruction in its containing method,
+whose name is \code{mname}, whose descriptor is \code{mname}, and
+whose declaring class is \code{cname}.  For
+instance, \code{8!main:[Ljava/lang/String;@test.HelloWorld} is the
+unique string representation of the \code{return} instruction in the
+main method above.

### 2011-03-29 main/doc/datalog_analysis.tex
more documentation

@@
+\xname{datalog_analysis}
 \chapter{Datalog Analysis}
@@
 specifying how to compute the output relations from the input relations.
-An example of such an analysis is shown in Figure \ref{fig:datalog-analysis}.
+An example of such an analysis is shown below:
 
-\begin{figure}
+\texonly{\newpage}
+
+\begin{framed}
 {\small
@@
 }
-\label{fig:datalog-analysis}
-\caption{An Example Datalog Analysis.}
-\end{figure}
+\end{framed}
 
@@
 
-TODO
+{\bf Under Construction}
 
@@
 
+\begin{framed}
 \begin{verbatim}
-    prompt> ant -Ddlog.file=<file> -Dwork.dir=<dir> solve
+ant -Ddlog.file=<file> -Dwork.dir=<dir> solve
 \end{verbatim}
+\end{framed}
+
 where {\tt <file>} denotes the file defining the Datalog analysis
@@
 
+\begin{framed}
 \begin{verbatim}
@@
 \end{verbatim}
+\end{framed}
 

### 2011-03-29 main/doc/user_guide.tex
added BDD representation documentation
cleaned up and updated documentation
added more documentation
more documentation
new documentation
documentation

@@
-\documentclass{article}
-\usepackage{html}
+\documentclass[openany]{book}
+\usepackage{hyperlatex}
 \usepackage{hyperref}
+\texonly{
+	\usepackage{graphicx}
+	\usepackage{framed}
+}
 
-\oddsidemargin=0in
-\evensidemargin=0in
-\textwidth=6.0in
+\texonly{
+% put borders around figures
+	%\usepackage{float}
+	%\floatstyle{boxed} 
+	%\restylefloat{figure}
+% set dimensions of page/text
+	\parindent=0in
+	\parskip=10pt
+	\oddsidemargin=-0.3in
+	\evensidemargin=-0.3in
+	\textwidth=7in
 	\textheight=8.0in
+% redefine certain commands
+	\renewcommand\code[1]{\path{#1}}
+	\renewcommand\xlink[2]{\href{#2}{#1}}
+}
 
-\html{\def\code#1{{\tt {#1}}}}
-\latex{\def\code#1{\path{#1}}}
-\def\url#1{\htmladdnormallink{#1}{#1}}
-\def\Mdom{{\bf M}}
-\def\Idom{{\bf I}}
-\def\Vdom{{\bf V}}
-\def\Zdom{{\bf Z}}
-\def\Hdom{{\bf H}}
-\def\Edom{{\bf E}}
-\def\Fdom{{\bf F}}
-\def\Cdom{{\bf C}}
-\def\Tdom{{\bf T}}
-\def\Pdom{{\bf P}}
+\htmlonly{
+	\newcommand\code[1]{{\tt #1}}
+	\newcommand\url[1]{\xlink{#1}{#1}}
+	\newenvironment{framed}{\xmlattributes*{table}{border=1 frame=box rules=none cellpadding=8}\begin{tabular}{l}}{\end{tabular}}
+}
 
-\def\instM{{\bf instM}}
-\def\statM{{\bf statM}}
-\def\MH{{\bf MH}}
-\def\MI{{\bf MI}}
-\def\ME{{\bf ME}}
+\newcommand\javadoc[2]{\xlink{\code{#1}}{http://chord.stanford.edu/javadoc/#2}}
 
-\def\cha{{\bf cha}}
-\def\sub{{\bf sub}}
-\def\HT{{\bf HT}}
-\def\VT{{\bf VT}}
-\def\EF{{\bf EF}}
-\def\writeE{{\bf writeE}}
+\newenvironment{mytable}[1]
+	{\htmlattributes*{table}{border=1 cellpadding=10}\begin{tabular}{#1}}
+	{\end{tabular}}
 
-\def\virtIM{{\bf virtIM}}
-\def\statIM{{\bf statIM}}
-\def\specIM{{\bf specIM}}
+\newcommand\bh{{\bf h}}
+\newcommand\bi{{\bf i}}
+\newcommand\bo{{\bf o}}
+\newcommand\bt{{\bf t}}
+\newcommand\bm{{\bf m}}
+\newcommand\be{{\bf e}}
+\newcommand\bg{{\bf f}}
+\newcommand\bb{{\bf b}}
+\newcommand\bl{{\bf l}}
+\newcommand\bp{{\bf p}}
+\newcommand\br{{\bf r}}
+\newcommand\bw{{\bf w}}
 
-\def\MmethArg{{\bf MmethArg}}
-\def\MmethRet{{\bf MmethRet}}
-\def\IinvkArg{{\bf IinvkArg}}
-\def\IinvkRet{{\bf IinvkRet}}
-\def\MordMethArg{{\bf MordMethArg}}
-\def\MspcMethArg{{\bf MspcMethArg}}
+\setcounter{htmldepth}{2}
+\setcounter{htmlautomenu}{3}
 
-\def\MobjVarAsgnInst{{\bf MobjVarAsgnInst}}
-\def\MclsVarAsgnInst{{\bf MclsVarAsgnInst}}
-\def\MobjValAsgnInst{{\bf MobjValAsgnInst}}
-\def\MgetInstFldInst{{\bf MgetInstFldInst}}
-\def\MputInstFldInst{{\bf MputInstFldInst}}
-\def\MgetStatFldInst{{\bf MgetStatFldInst}}
-\def\MputStatFldInst{{\bf MputStatFldInst}}
+\htmltitle{Chord}
+\htmladdress{mhn@cs.stanford.edu}
 
-\def\syncMV{{\bf syncMV}}
-\def\nestedLocks{{\bf nestedLocks}}
-
-\def\VH{{\bf VH}}
-\def\FH{{\bf FH}}
-\def\HFH{{\bf HFH}}
-\def\IM{{\bf IM}}
-
-\def\forkC{{\bf forkC}}
-
-\def\CVC{{\bf CVC}}
-\def\FC{{\bf FC}}
-\def\CFC{{\bf CFC}}
-\def\CICM{{\bf CICM}}
-\def\CHC{{\bf CHC}}
-
-\def\clinitM{{\bf clinitM}}
-
-\def\bddbddb{{\tt bddbddb}}
-\def\Chord{Chord}
-
-\def\NameToTrgtMap{{\tt NameToTrgtMap}}
-\def\NameToTaskMap{{\tt NameToTaskMap}}
-\def\ConsumedNameToTasksMap{{\tt ConsumedNameToTasksMap}}
-\def\TaskGraph{{\tt TaskGraph}}
-\def\TaskMapEntries{{\tt TaskMapEntries}}
-\def\TrgtMapEntries{{\tt TrgtMapEntries}}
-\def\DlogMapEntries{{\tt DlogMapEntries}}
-\def\Tasks{{\tt Tasks}}
-\def\Trgts{{\tt Trgts}}
-\def\ValidTasks{{\tt ValidTasks}}
-\def\TrgtNames{{\tt TrgtNames}}
-
-\title{Chord: A Static and Dynamic Program Analysis Framework for Java}
+\title{Chord: A Versatile Platform for Program Analysis}
 \author{Mayur Naik}
@@
 \input{whatis_chord}
-\newpage
-\input{download_and_installation}
-\newpage
-\input{program_setup}
-\newpage
-\input{building_scope}
-\newpage
-\input{running_analysis}
-\newpage
-\input{writing_analysis}
-\newpage
-\input{writing_dynamic_analysis}
-\newpage
-\input{instr_events}
-\newpage
-\input{writing_datalog_analysis}
-\newpage
-\input{tuning_datalog_analysis}
-\newpage
+\input{getting_started}
+\input{example}
+\input{architecture}
+\input{properties}
+\input{predefined_analyses}
+\input{analyses}
+\input{dynamic_analysis}
+\input{datalog_analysis}
+\input{computing_scope}
 \input{program_representation}
-\newpage
-\input{pointsto}
-\newpage
-\input{datarace}
-\newpage
-\input{deadlock}
-\newpage
-\input{chord_sysprops}
-\newpage
 \input{acknowledgments}

### 2011-03-29 main/doc/program_representation.tex
more documentation

@@
 \xname{program_representation}
-\chapter{Java Program Representation}
+\chapter{Program Representation}
 \label{chap:program-representation}
@@
 intermediate representation of the input Java program called {\it
-quadcode}.  This chapter describes how to pretty-print bytecode and
-quadcode, which is useful for debugging analyses and deciphering their
-output, as well as the API of Joeq and Chord that concerns bytecode
-and quadcode, which is useful for users writing their own analyses.
+quadcode}.  This chapter describes all aspects of quadcode and how it
+relates to bytecode.  It first explains how to pretty-print bytecode
+and quadcode (Section \ref{sec:pretty-printing}) which is useful for
+debugging analyses and deciphering their output.  The remaining
+sections describe the quadcode representation along with the API of
+Joeq and Chord for navigating it.  Briefly, the representation
+consists of a set of classes that may be loaded
+(Section \ref{sec:program}).  Each class has a set of fields and
+methods, collectively called members of the class
+(Section \ref{sec:class-members}).  The representation of a concrete
+method (Section \ref{sec:methods}) consists of a control-flow graph
+(CFG).  The representation of a CFG (Section \ref{sec:cfgs}) consists
+of a set of local variables and a set of basic blocks.  Each basic
+block contains zero or more primitive statements called quads
+(Section \ref{sec:quads}).  Finally, the most common way to traverse
+all quads is discussed (Section \ref{sec:traversing}).
 
-\section{Pretty-Printing Bytecode and Quadcode}
+\section{Pretty-Printing}
+\label{sec:pretty-printing}
 
@@
 
-To pretty-print the bytecode of a class, run the following command:
+To pretty-print the bytecode representation of a class, run the following command:
 \begin{framed}
@@
 
-where {\tt <CLASS NAME>} is the fully-qualified name of the class
-whose bytecode is to be printed (\code{test.HelloWorld} in our
-example) and \code{<...>} is the classpath of that class
-(\code{examples/hello_world/} in our example).  Program \code{javap}
-comes along with the JVM.  The output of the above command for our
-example is as follows:
+where:
+
+\begin{itemize}
+\item
+{\tt <CLASS NAME>} is the fully-qualified name of the class whose
+bytecode is to be printed (\code{test.HelloWorld} in our example).
+\item
+\code{<...>} is the classpath of that class (\code{examples/hello_world/} in our example).
+\end{itemize}
+
+Program \code{javap} comes along with the JVM.  The output of the
+above command for our example is as follows:
 
@@
 
-To pretty-print the quadcode of a class, run the following command:
+To pretty-print the quadcode representation of a class, run the following command:
 
@@
 
-where \code{<...>} is the path of a directory
-(\code{examples/hello_world/} in our example) that is expected to
+where:
+
+\begin{itemize}
+\item
+\code{<...>} is the path of a directory (\code{examples/hello_world/} in our example) that is expected to
 contain a file named \code{chord.properties} which defines
@@
 (alternatively, these two properties can be defined directly on the
-above command line), {\tt <CLASS NAME>} is the name of the class whose
-quadcode is to be printed, and {\tt <...>/<FILE NAME>} is the name of
-a file to which the quadcode is to be dumped.  The output of the above
-command for our example is as follows:
+above command line).
+\item
+{\tt <CLASS NAME>} is the name of the class whose quadcode is to be
+printed.
+\item
+{\tt <FILE NAME>} is the name of a file to which the quadcode is to be
+dumped.
+\end{itemize}
+
+The output of the above command for our example is as follows:
 
@@
 
-\section{Traversing Quadcode}
+\section{Whole Program}
+\label{sec:program}
 
-\subsection{Program}
+This and the following sections describe the quadcode representation
+along with the API of Joeq and Chord for navigating it.
+This API is contained in
+packages \javadoc{chord.program}{chord/program/package-tree.html}, \javadoc{joeq.Class}{joeq/Class/package-tree.html},
+and \javadoc{joeq.Compiler.Quad}{joeq/Compiler/Quad/package-tree.html}.
+
+The quadcode representation of the whole program is a unique global
+object of
+class \javadoc{chord.program.Program}{chord/program/Program.html}
+which can be obtained by calling static
+method \javadoc{chord.program.Program.g()}{chord/program/Program.html\#g()}.
+This class provides a rich API (in the form of public instance
+methods) to access various parts of the representation, most notably:
+
+\begin{mytable}{|l|l|}
+\hline
+\verb+IndexSet<jq_Type> getTypes()+ & All types referenced in classes that may be loaded. \\
+\hline
+\verb+IndexSet<jq_Reference> getClasses()+ & All classes that may be loaded.* \\
+\hline
+\verb+IndexSet<jq_Method> getMethods()+ & All methods that may be called. \T \\
+\hline
+\end{mytable}
+
+* Includes both classes/interfaces and array types, represented as objects
+of \code{jq_Class} and \code{jq_Array}, respectively; both these are
+subclasses of \code{jq_Reference}.
+
+See Chapter \ref{chap:computing-scope} for how Chord determines which
+classes may be loaded and which methods may be called.
+
+The quadcode representation of each type is a unique object of the
+appropriate subclass
+of \javadoc{joeq.Class.jq_Type}{joeq/Class/jq_Type.html} in the following
+hierarchy:
 
@@
 \begin{verbatim}
-package chord.program;
-
-class Program {
-    static Program g() {
-        // provides quadcode representation of program
-    }
-    public IndexSet<jq_Type> getTypes() {
-        // provides quadcode representation of all reachable types
-    }
-    public IndexSet<jq_Reference> getClasses() {
-        // provides quadcode representation of all reachable classes
-    }
-    public IndexSet<jq_Method> getMethods() {
-        // provides quadcode representation of all reachable methods
-    }
-    public Reflect getReflect() {
-        // provides information about resolved reflection
-    }
-    public jq_Reference getClass(String name) {
-        // provides quadcode representation of class 
-    }
-    public jq_Type getType(String name) {
-    }
-    public jq_Method getMainMethod() {
-    }
-    public jq_Method getMethod(String mName, String mDesc, String cName) {
-    }
-    public jq_Method getMethod(String sign) {
-    }
-    public Quad getQuad(MethodElem e) {
-    }
-    public Quad getQuad(MethodElem e, Class quadOpClass) {
-    }
-}    
+joeq.Class.jq_Type
+  |
+  |--- joeq.Class.jq_Primitive
+  |
+  |--- joeq.Class.jq_Reference
+            |
+            |--- joeq.Class.jq_Class
+            |
+            |--- joeq.Class.jq_Array
 \end{verbatim}
@@
 
-\subsection{Type}
+\section{Class Members}
+\label{sec:class-members}
 
-\begin{framed}
-\begin{verbatim}
-class jq_Type {
-}
-\end{verbatim}
-\end{framed}
+Each primitive type (e.g., boolean, int, etc.) is represented by a
+unique \code{jq_Primtive} object.  Each class and each interface type is
+represented by a unique \code{jq_Class} object.  Each array type is
+represented by a unique \code{jq_Array} object.
 
-\subsection{Class}
+Members (i.e., fields and methods) of the class/interface represented
+by an object of
+class \javadoc{joeq.Class.jq_Class}{joeq/Class/jq_Class.html} can be
+accessed using the following API provided by that class.
 
-\begin{framed}
-\begin{verbatim}
-class jq_Reference {
-}
-\end{verbatim}
-\end{framed}
+\begin{mytable}{|l|p{3.1in}|}
+\hline
+\verb+String getName()+ & Fully-qualified name of the class, e.g., ``\code{java.lang.String[]}''. \\
+\hline
+\verb+jq_InstanceField[] getDeclaredInstanceFields()+ & All instance fields declared in the class. \\
+\hline
+\verb+jq_StaticField[] getDeclaredStaticFields()+ & All static fields declared in the class. \\
+\hline
+\verb+jq_InstanceMethod[] getDeclaredInstanceMethods()+ & All instance methods declared in the class. \\
+\hline
+\verb+jq_StaticMethod[] getDeclaredStaticMethods()+ & All static methods declared in the class. \T \\
+\hline
+\end{mytable}
 
-\subsection{Field}
-
-\begin{framed}
-\begin{verbatim}
-class jq_Field {
-     public String getName();
-     public jq_Class getDeclaringClass();
-     public jq_Type getType() {
-     }
-     public String toString() {
-         // print this field in format fname:fdesc@cname where:
-         // - fname is the name of the field
-         // - fdesc is the descriptor of the field's type
-         // - cname is the name of the field's declaring class
-         // example: buf:[B@java.io.BufferedInputStream
-     }
-}
-\end{verbatim}
-\end{framed}
-
-\subsection{Method}
-
-\begin{framed}
-\begin{verbatim}
-class jq_Method {
-
-}
-\end{verbatim}
-\end{framed}
-
-Register Factory
-
-Control-Flow Graph
-
-Basic Block
-
-Quad
-
-\begin{framed}
-\begin{verbatim}
-Traversing a control-flow graph
-\end{verbatim}
-\end{framed}
-
-\section{Program Descriptors}
-
-MethodSign
+Chord uses format \code{mName:mDesc@cName}, described in
+class \javadoc{chord.program.MethodSign}{chord/program/MethodSign.html},
+to uniquely identify each field and each method in the input Java
+program, where
+\code{mName} denotes the name of the field/method,
+\code{mDesc} denotes the descriptor of the field/method (see below),
+and \code{cName} denotes the fully-qualified name of the class
+declaring the field/method.
+For instance, ``\code{main:[Ljava/lang/String;@test.HelloWorld}''
+uniquely identifies the main method in the example above.
+We next review field descriptors and method descriptors from the Java
+bytecode specification.
 
@@
 
-\begin{table}
-\htmlattributes*{table}{border=1 cellpadding=10}
-\begin{tabular}{l|l|l}
+\begin{mytable}{l|l|l}
+\hline
 BaseType Character	& Type	& Interpretation \\
@@
 {\tt Z}	&boolean& true or false \\
-{\tt [}	&reference& one array dimension
-\end{tabular}
-\end{table}
+{\tt [}	&reference& one array dimension \T \\
+\hline
+\end{mytable}
 
@@
 
-A method descriptor represents the parameters that the method takes
-and the value that it returns:
+A method descriptor represents the types of the arguments and 
+return result of a method:
 
@@
 
-A parameter descriptor represents a parameter passed to a method.  A
-return descriptor represents the type of the value returned from a
+A parameter descriptor represents the type of an argument of a method.
+A return descriptor represents the type of the return result of a
 method.  The character \code{V} indicates that the method returns no
-value (its return type is void).  For example, the method descriptor
-for the method
-``\code{Object foo(int i, double d, Thread t)}'' is
-``\code{(IDLjava/lang/Thread;)Ljava/lang/Object;}''.
-Note that internal forms of the fully qualified names of \code{Thread} and
-\code{Object} are used in the method descriptor.  The method descriptor for
-\code{foo} is the same whether it is a static or an instance
-method. Although an instance method is passed \code{this}, a reference to the
-current class instance, in addition to its intended parameters, that
-fact is not reflected in the method descriptor.
+value (its return type is void).
 
-Chord uniquely represents each bytecode instruction in each method
-declared in each class of the program by a
-four-tuple \code{bci!mname:mdesc@cname} where \code{bci} is the
-(0-based) bytecode offset of the instruction in its containing method,
-whose name is \code{mname}, whose descriptor is \code{mname}, and
-whose declaring class is \code{cname}.  For
-instance, \code{8!main:[Ljava/lang/String;@test.HelloWorld} is the
-unique string representation of the \code{return} instruction in the
-main method above.
+The method descriptor is the same whether it is a static or an instance
+method.  Although an instance method is passed \code{this}, a reference
+to the current class instance, in addition to its intended arguments,
+that fact is not reflected in the method descriptor.
+
+For example, the method descriptor
+for the method ``\code{Object foo(int i, double d, Thread t)}'' is
+``\code{(IDLjava/lang/Thread;)Ljava/lang/Object;}''.  Note that
+internal forms of the fully qualified names of \code{Thread} and
+\code{Object} are used in the method descriptor.
+
+\section{Methods}
+\label{sec:methods}
+
+The quadcode representation of each method is a unique object of
+class \javadoc{joeq.Class.jq_Method}{joeq/Class/jq_Method.html}.  Components of the method (e.g.,
+its control-flow graph) can be accessed using the following API provided
+by that class.
+
+\begin{mytable}{|l|p{4in}|}
+\hline
+\verb+String getName()+ & Name of the method. \\
+\hline
+\verb+String getDesc().toString()+ & Descriptor of the method, e.g., ``\code{(Ljava/lang/String;)V}''. \\
+\hline
+\verb+jq_Class getDeclaringClass()+ & Declaring class of the method. \\
+\hline
+\verb+ControlFlowGraph getCFG()+ & Control-flow graph of the method.* \\
+\hline
+\verb+int getLineNumber(int bci)+ & Line number of the given bytecode offset (-1 if not found). \\
+\hline
+\verb+Quad getQuad(int bci)+ & First quad at the given bytecode
+offset (null if not found). \\
+\hline
+\verb+Quad getQuad(int bci, Class kind)+ & First quad of the given
+kind at the given bytecode offset (null if not found). \\
+\hline
+\verb+Quad getQuad(int bci, Class[] kind)+ & First quad of any
+given kind at the given bytecode offset (null if not found). \\
+\hline
+\verb+String toString()+ & Unique identifier of the method in format \code{mName:mDesc@cName}. \T \\
+\hline
+\end{mytable}
+
+* The control-flow graph must not be asked if the method is abstract (which can be determined by calling instance method \code{isAbstract()} of \code{jq_Method}).
+
+\section{Control-Flow Graphs}
+\label{sec:cfgs}
+
+The control-flow graph (CFG) of each method consists of a set of local
+variables of the method, called a ``register factory'', and a graph
+whose nodes are basic blocks and whose edges denote flow of control
+between basic blocks.
+
+The CFG of each method is a unique object of
+class \javadoc{joeq.Compiler.Quad.ControlFlowGraph}{joeq/Compiler/Quad/ControlFlowGraph.html}.
+Components of the CFG can be accessed using the following API provided
+by that class.
+
+\begin{mytable}{|l|l|}
+\hline
+\verb+getRegisterFactory()+ & Set of all local variables. \\
+\hline
+\verb+EntryOrExitBasicBlock entry()+ & Unique entry basic block. \\
+\hline
+\verb+EntryOrExitBasicBlock exit()+ & Unique exit basic block. \\
+\hline
+\verb+ListIterator.BasicBlock reversePostOrderIterator()+ & Iterator over all basic blocks in reverse post-order. \\
+\hline
+\verb+jq_Method getMethod()+ & Containing method of the CFG. \T \\
+\hline
+\end{mytable}
+
+The register factory contains one variable per argument of the method
+and one variable per temporary declared in the method body.
+Temporaries include those declared by programmers as well as those
+generated by Joeq.  The reason for the latter is that the quadcode
+representation is register-based and is constructed from Java
+bytecode, which is stack-based; moreover, Joeq does the Static Single
+Assignment (SSA) transformation by default, which introduces
+temporaries to ensure that there is at most one static assignment to
+any variable.
+
+The register factory is a unique object
+of class \javadoc{joeq.Compiler.Quad.RegisterFactory}{joeq/Compiler/Quad/RegisterFactory.html}.
+
+Each CFG has a set of basic blocks each of which contains a
+straight-line sequence of zero or more primitive statements called quads
+(Section \ref{sec:quads}).  Each CFG is guaranteed to contain at least
+two basic blocks: a unique entry basic block with no incoming edges
+and a unique exit block with no outgoing edges.  The entry and exit
+basic blocks do not contain any quads.
+
+Each basic block is a unique object of
+class \javadoc{joeq.Compiler.Quad.BasicBlock}{joeq/Compiler/Quad/BasicBlock.html}
+(the entry and exit basic blocks are instances of a
+subclass \javadoc{joeq.Compiler.Quad.EntryOrExitBasicBlock}{joeq/Compiler/Quad/EntryOrExitBsaicBlock.html}).
+Components of the basic block can be accessed using the following API
+provided by that class.
+
+\begin{mytable}{|l|l|}
+\hline
+\verb+int size()+ & Number of quads contained in the basic block. \\
+\hline
+\verb+Quad getQuad(int index)+ & Quad at the given 0-based index. \\
+\hline
+\verb+List.BasicBlock getPredecessors()+ & List of immediate predecessor basic blocks. \\
+\hline
+\verb+List.BasicBlock getSuccessors()+ & List of immediate successor basic blocks. \\
+\hline
+\verb+jq_Method getMethod()+ & Containing method of the basic block. \T \\
+\hline
+\end{mytable}
+
+\section{Quads}
+\label{sec:quads}
+
+Chord uses format \code{offset!mName:mDesc@cName}, described in
+class \javadoc{chord.program.MethodElem}{chord/program/MethodElem.html},
+to uniquely identify each bytecode instruction in the input Java
+program, where \code{offset} is the (0-based) bytecode offset of the
+instruction in its containing method, \code{mName} is the name of the
+method, \code{mDesc} is the descriptor of the method, and \code{cName}
+is the fully-qualified name of the class declaring the method.  For
+instance, ``\code{8!main:[Ljava/lang/String;@test.HelloWorld}''
+uniquely identifies the return instruction in the main method
+in the example above.
+
+The quadcode representation is register-based, as opposed to Java
+bytecode that is used to construct it, which is stack-based.  As a
+result, it uses {\it quads} to represent bytecode instructions.  A
+quad is a primitive statement that consists of an operator and upto
+four operands.  There is no one-to-one correspondence between bytecode
+instructions and quads: certain bytecode instructions generate a
+sequence of more than one quads while others do not generate any quad.
+The API of class \code{jq_Method} provides various \code{getQuad(...)}
+methods to access the quad(s)
+corresponding to a bytecode instruction (see Section \ref{sec:methods}).
+
+Each quad is a unique object of
+class \javadoc{joeq.Compiler.Quad.Quad}{joeq/Compiler/Quad/Quad.html}.
+Components of the quad can be accessed using the following API
+provided by that class.
+
+\begin{mytable}{|l|l|}
+\hline
+\verb+Operator getOperator()+ & Kind of the quad. \\
+\hline
+\verb+int getBCI()+ & Bytecode offset of the quad in its containing
+method. \\
+\hline
+\verb+String toByteLocStr()+ & Unique identifier of the quad in
+format \code{offset!mName:mDesc@cName}. \\
+\hline
+\verb+String toJavaLocStr()+ & Location of the quad in
+format \code{fileName:lineNum} in Java source code. \\
+\hline
+\verb+String toLocStr()+ & Location of the quad in both Java bytecode and source code. \\
+\hline
+\verb+String toVerboseStr()+ & Verbose description of the quad (its location plus contents). \\
+\hline
+\verb+jq_Method getMethod()+ & Containing method of the quad. \T \\
+\hline
+\end{mytable}
+
+The kind of each quad is determined by its operator which is a unique object of
+the appropriate subclass
+of \javadoc{joeq.Compiler.Quad.Operator}{joeq/Compiler/Quad/Operator.html}
+in the following hierarchy:
+
+\begin{framed}
+\begin{verbatim}
+Operator
+    |--- Move
+    |--- Phi
+    |--- Binary
+    |--- Unary
+    |--- New
+    |--- NewAray
+    |--- MultiNewArray
+    |--- CheckCast
+    |--- InstanceOf
+    |--- ALength
+    |--- ALoad
+    |--- AStore
+    |--- Getstatic
+    |--- Putstatic
+    |--- Getfield
+    |--- Putfield
+    |--- Return
+    |--- Branch
+    |       |--- IntIfCmp
+    |       |--- Goto
+    |       |--- Jsr
+    |       |--- Ret
+    |       |--- LookupSwitch
+    |       |--- TableSwitch
+    |--- Invoke
+    |       |--- InvokeVirtual
+    |       |--- InvokeStatic
+    |       |--- InvokeInterface
+    |--- Monitor
+            |--- MONITORENTER
+            |--- MONITOREXIT
+\end{verbatim}
+\end{framed}
+
+The number and kinds of operands of each quad depends upon the kind of
+the operator.  Each of the above subclasses of \code{Operator}
+provides an API to access the operands of the quad.  For instance,
+the components of a \code{Getfield} quad {\tt q} of the form ``l = b.f'' can be accessed as follows:
+
+\begin{framed}
+\begin{verbatim}
+Operand lo = Getfield.getSrc(q);
+Operand bo = Getfield.getBase(q);
+if (lo instanceof RegisterOperand && bo instanceof RegisterOperand) {
+    Register l = ((RegisterOperand) lo).getRegister();
+    Register b = ((RegisterOperand) bo).getRegister();
+    jq_Field f = Getfield.getField(q).getField();
+    ...
+}
+\end{verbatim}
+\end{framed}
+
+\section{Traversing Quadcode}
+\label{sec:traversing}
+
+A common way to traverse all quads in the input Java program is as follows:
+
+\texonly{\newpage}
+
+\begin{framed}
+\begin{verbatim}
+
+import chord.program.Program;
+import joeq.Compiler.Quad.QuadVisitor;
+import joeq.Class.jq_Method;
+import joeq.Compiler.Quad.ControlFlowGraph;
+import joeq.Util.Templates.ListIterator;
+import joeq.Compiler.Quad.BasicBlock;
+import joeq.Compiler.Quad.Quad;
+
+QuadVisitor qv = new QuadVisitor.EmptyVisitor() {
+    public void visitMove(Quad q) { ... }
+    public void visitPhi(Quad q) { ... }
+    public void visitBinary(Quad q) { ... }
+    ...
+};
+Program program = Program.g();
+for (jq_Method m : program.getMethods()) {
+    if (!m.isAbstract()) {
+        ControlFlowGraph cfg = m.getCFG();
+        ListIterator.BasicBlock it = cfg.reversePostOrderIterator();
+        while (it.hasNext()) {
+            BasicBlock b = it.nextBasicBlock();
+            for (int i = 0; i < b.size(); i++) {
+                Quad q = b.getQuad(i);
+                q.accept(qv);
+            }
+        }
+    }
+}
+\end{verbatim}
+\end{framed}

### 2011-03-30 main/doc/example.tex
documentation edits

@@
-\chapter{An Example Run of Chord}
+\xname{example}
+\chapter{An Example Run}
 \label{chap:example}
@@
 To run Chord on the above program, run the following command in
-Chord's {\tt main/} directory:
+Chord's main directory:
 
+\begin{framed}
 \begin{verbatim}
-    prompt> ant -Dchord.work.dir=<...>/example run
+ant -Dchord.work.dir=<...>/example run
 \end{verbatim}
+\end{framed}
 
@@
 \end{verbatim}
+\end{framed}
 \end{itemize}
 
-To run one or more such analyses, run the following command in Chord's \code{main/} directory:
+To run one or more such analyses, run the following command in Chord's main directory:
 
+\begin{framed}
 \begin{verbatim}
-    prompt> ant -Dchord.work.dir=<PROGRAM DIR> -Dchord.run.analyses=<ANALYSIS NAME> run
+ant -Dchord.work.dir=<PROGRAM DIR> -Dchord.run.analyses=<ANALYSIS NAME> run
 \end{verbatim}
+\end{framed}
 
@@
 For instance, to run a basic may-alias and call-graph analysis (called 0CFA),
-run the following command in Chord's \code{main/} directory:
+run the following command in Chord's main directory:
 
+\begin{framed}
 \begin{verbatim}
-    prompt> ant -Dchord.work.dir=<PROGRAM DIR> -Dchord.run.analyses=cipa-0cfa-dlog run
+ant -Dchord.work.dir=<PROGRAM DIR> -Dchord.run.analyses=cipa-0cfa-dlog run
 \end{verbatim}
+\end{framed}
 
@@
 
+\begin{framed}
+{\small
 \begin{verbatim}
@@
 \end{verbatim}
+}
+\end{framed}
 
@@
 \end{verbatim}
+}
+\end{framed}

### 2011-03-30 main/doc/user_guide.tex
documentation edits

@@
 	\newcommand\url[1]{\xlink{#1}{#1}}
-	\newenvironment{framed}{\xmlattributes*{table}{border=1 frame=box rules=none cellpadding=8}\begin{tabular}{l}}{\end{tabular}}
+	\newenvironment{framed}
+		{\xmlattributes*{table}{border=1 frame=box rules=none cellpadding=8 bgcolor=CCFFFF width=100%
+		}\begin{tabular}{l}}{\end{tabular}}
 }
@@
 
-\newenvironment{mytable}[1]
-	{\htmlattributes*{table}{border=1 cellpadding=10}\begin{tabular}{#1}}
-	{\end{tabular}}
+\newenvironment{mytable}[1]{\htmlattributes*{table}{border=1 cellpadding=10}\begin{tabular}{#1}}{\end{tabular}}
 

### 2011-04-04 main/doc/getting_started.tex
documentation

@@
+\xname{getting_started}
 \chapter{Getting Started}
@@
 This chapter describes how to download, install, and run Chord.
-Section \ref{sec:downloading-binaries} describes how to obtain pre-built binaries of Chord.
-Section \ref{sec:downloading-sources} describes how to obtain the source code of Chord
-and Section \ref{sec:downloading-sources} explains how to build it.
-Finally, Section \ref{sec:running-chord} describes how to run Chord.
+Section \ref{sec:downloading-binaries} describes how to obtain
+pre-built binaries of Chord.  Section \ref{sec:downloading-sources}
+describes how to obtain the source code of Chord and Section
+\ref{sec:downloading-sources} explains how to build it.  Finally,
+Section \ref{sec:running-chord} describes how to run Chord.
 
@@
 
-To obtain Chord's pre-built binaries, download file \xlink{chord-bin-2.0.tar.gz}{http://jchord.googlecode.com/files/chord-bin-2.0.tar.gz}.
-It contains the following files:
+To obtain Chord's pre-built binaries, download and uncompress file
+\xlink{chord-bin-2.0.tar.gz}{http://jchord.googlecode.com/files/chord-bin-2.0.tar.gz}.
+It includes the following files:
 
@@
 \item
-\code{chord.jar}, which contains the class files of Chord and of libraries used by Chord.
+\code{chord.jar}, which contains the class files of Chord and of
+libraries used by Chord.
 \item
-\code{libbuddy.so}, \code{buddy.dll}, and \code{libbuddy.dylib}: you can keep one of these files
-depending upon whether you intend to run Chord on Linux, Windows/Cygwin, or MacOS, respectively.
-These files are needed only if you want BDD library BuDDy to be used when 
-the BDD-based Datalog solver bddbddb in Chord runs analyses written in Datalog.
+\code{libbuddy.so}, \code{buddy.dll}, and \code{libbuddy.dylib}: you
+can keep one of these files depending upon whether you intend to run
+Chord on Linux, Windows/Cygwin, or MacOS, respectively.  These files
+are needed only if you want BDD library BuDDy to be used when the
+BDD-based Datalog solver bddbddb in Chord runs analyses written in
+Datalog.
 \item
-\code{libchord_instr_agent.so}: this file is needed only if you want the JVMTI-based bytecode
-instrumentation agent to be used when Chord runs dynamic analyses.
+\code{libchord_instr_agent.so}: this file is needed only if you want
+the JVMTI-based bytecode instrumentation agent to be used when Chord
+runs dynamic analyses.
 \end{enumerate}
 
-Novice users can ignore items (2) and (3) until they become more familiar with Chord.
-The binaries mentioned in items (2) and (3) might not be compatible with your machine, in which case you
-can either forgo using them (with hardly any noticeable difference in functionality),
-or you can download the sources (see Section \ref{sec:downloading-sources}) and build them yourself
-(see Section \ref{sec:compiling-sources}).
+Novice users can ignore items (2) and (3) until they become more
+familiar with Chord.  The binaries mentioned in items (2) and (3)
+might not be compatible with your machine, in which case you can
+either forgo using them (with hardly any noticeable difference in
+functionality), or you can download the sources (see Section
+\ref{sec:downloading-sources}) and build them yourself (see Section
+\ref{sec:compiling-sources}).
 
@@
 
-To obtain Chord's source code, download the following files:
+To obtain Chord's source code, download and uncompress the following
+files:
 
@@
 \item
-Mandatory: file \xlink{chord-src-2.0.tar.gz}{http://jchord.googlecode.com/files/chord-src-2.0.tar.gz}, which contains Chord's source code and the jars of libraries used by Chord.
+Mandatory: file
+\xlink{chord-src-2.0.tar.gz}{http://jchord.googlecode.com/files/chord-src-2.0.tar.gz},
+which contains Chord's source code and jars of libraries used by
+Chord.
 \item
-Optional: file \xlink{chord-libsrc-2.0.tar.gz}{http://jchord.googlecode.com/files/chord-libsrc-2.0.tar.gz}, which contains the source
-code of libraries used by Chord (e.g., joeq, javassist, bddbddb, etc.)
+Optional: file
+\xlink{chord-libsrc-2.0.tar.gz}{http://jchord.googlecode.com/files/chord-libsrc-2.0.tar.gz},
+which contains the source code of libraries used by Chord (e.g., joeq,
+javassist, bddbddb, etc.)
 \end{itemize}
 
-Alternatively, you can obtain the latest development snapshot from the SVN repository:
+Alternatively, you can obtain the latest development snapshot from the
+SVN repository by running the following command:
 
+\begin{framed}
 \begin{verbatim}
-    prompt> svn checkout http://jchord.googlecode.com/svn/trunk/ chord
+svn checkout http://jchord.googlecode.com/svn/trunk/ chord
 \end{verbatim}
+\end{framed}
 
-Instead of checking out the entire \code{trunk/}, which contains several sub-directories, you can check out
-specific sub-directories:
+Instead of checking out the entire \code{trunk/}, which contains
+several sub-directories, you can check out specific sub-directories:
 
@@
 \item
-\code{main/} contains Chord's source code and the jars of libraries used by Chord.
+\code{main/} contains Chord's source code and jars of libraries used by Chord.
 \item
-\code{libsrc/} contains the source code of libraries used by Chord (e.g., joeq, javassist, bddbddb, etc.).
+\code{libsrc/} contains the source code of libraries used by Chord
+(e.g., joeq, javassist, bddbddb, etc.).
 \item
@@
 \item
-many more; these might eventually move into \code{main/}.
+many more; these might eventually move under \code{main/}.
 \end{itemize}
 
-Files \code{chord-2.0-src.tar.gz} and \code{chord-2.0-libsrc.tar.gz} mentioned above are essentially stable releases of
-the \code{main/} and \code{libsrc/} directories, respectively.
+Files \code{chord-2.0-src.tar.gz} and \code{chord-2.0-libsrc.tar.gz}
+mentioned above are essentially stable releases of the \code{main/}
+and \code{libsrc/} directories, respectively.
 
@@
 \item
-A JVM with JDK 5 or higher, e.g. \xlink{IBM J9}{http://www.ibm.com/developerworks/java/jdk/} or
+JVM for Java 5 or higher, e.g.
+\xlink{IBM J9}{http://www.ibm.com/developerworks/java/jdk/} or
 \xlink{Oracle HotSpot}{http://www.oracle.com/technetwork/java/javase/downloads/index.html}.
@@
 
-Directory {\tt main/} contains a file named {\tt build.xml} which is
-interpreted by Apache Ant.  To see the various targets available, run
-the following command in that directory:
+Chord's main directory contains a file named {\tt build.xml} which is
+interpreted by Apache Ant.  To see the various possible targets,
+simply run command ``{\tt ant}" in that directory.
 
-\begin{verbatim}
-    prompt> ant
-\end{verbatim}
-
-To compile Chord, run the following command in the same directory:
-
-\begin{verbatim}
-    prompt> ant compile
-\end{verbatim}
-
-This will compile Chord's Java sources from \code{main/src/} to class files
-in \code{main/classes/}, as well as build a jar file \code{main/chord.jar} containing
-these class files as well as the class files in the jars of libraries that are used by Chord
-and are provided under \code{main/lib/} (e.g., \code{joeq.jar}, \code{javassist.jar},
-\code{bddbddb.jar}, etc.).
-Additionally:
+To compile Chord, run command ``{\tt ant compile}" in the same directory.
+This will compile Chord's Java sources from \code{src/} to class
+files in \code{classes/}, as well as build a jar file
+\code{chord.jar} that contains these class files as well as those
+in the jars of libraries that are used by Chord and are
+provided under \code{lib/} (e.g., \code{joeq.jar},
+\code{javassist.jar}, \code{bddbddb.jar}, etc.).  Additionally:
 
@@
 
-If system property \code{chord.use.buddy} is set to \code{true},
-then the C source code of BDD library \xlink{BuDDy}{http://buddy.sourceforge.net/} 
-from \code{main/bdd/} will be compiled to a shared library in \code{main/}
+If system property \code{chord.use.buddy} is set to \code{true}, then
+the C source code of BDD library
+\xlink{BuDDy}{http://buddy.sourceforge.net/} from directory \code{bdd/}
+will be compiled to a shared library named
 (\code{libbuddy.so} on Linux, \code{buddy.dll} on Windows, and
-\code{libbuddy.dylib} on MacOS); this library is used 
-by BDD-based Datalog solver \xlink{bddbddb}{http://bddbddb.sourceforge.net/} 
-in Chord for running analyses written in Datalog.
+\code{libbuddy.dylib} on MacOS; this library is used by BDD-based
+Datalog solver \xlink{bddbddb}{http://bddbddb.sourceforge.net/} in
+Chord for running analyses written in Datalog.
 
@@
 
-If system property \code{chord.use.jvmti} is set to \code{true},
-then the C++ source code of the JVMTI-based bytecode instrumentation agent from
-\code{main/agent/} will be compiled to a shared library in \code{main/}
-(\code{libchord_instr_agent.so} on all architectures); this agent is
-used in Chord for computing analysis scope dynamically and for running
-dynamic analyses.
+If system property \code{chord.use.jvmti} is set to \code{true}, then
+the C++ source code of the JVMTI-based bytecode instrumentation agent
+from directory \code{agent/} will be compiled to a shared library named
+\code{libchord_instr_agent.so} on all architectures;
+this agent is used in Chord for computing analysis scope dynamically
+and for running dynamic analyses.
 \end{itemize}
 
-Properties \code{chord.use.buddy} and \code{chord.use.jvmti} are defined in file \code{main/chord.properties}.  The
-default value of both these properties is \code{false}.
-If you set either of them to \code{true}, then you will also need a utility like GNU Make (to run the
-\code{Makefile}'s under \code{main/bdd/} and \code{main/agent/}) and a C++ compiler.
+Properties \code{chord.use.buddy} and \code{chord.use.jvmti} are
+defined in file \code{chord.properties} in Chord's main directory.
+The default value of both these properties is \code{false}.  If you set
+either of them to \code{true}, then you will also need a utility like
+GNU Make (to run the \code{Makefile}'s in directories \code{bdd/} and
+\code{agent/}) and a C++ compiler (to build the above shared libraries).
 
@@
 
-Running Chord requires a JVM with JDK 5 or higher.
+Running Chord requires a JVM supporting Java 5 or higher.
 There are two equivalent ways to run Chord.
-One way is to run the following command in the \code{main/} directory:
+One way, which is available only in the source installation of Chord, is to run the
+following command in Chord's main directory:
 
+\begin{framed}
 \begin{verbatim}
-    prompt> ant -D<key1>=<value1> ... -D<keyN>=<valueN> run
+ant -D<key1>=<val1> ... -D<keyN>=<valN> run
 \end{verbatim}
+\end{framed}
 
-The above requires Apache Ant (a Java build tool) to be installed on your machine.  The alternative, which
-does not require Apache Ant, is to run the following command:
+The above requires Apache Ant (a Java build tool) to be installed on
+your machine.  The alternative, which does not require Apache Ant and is
+available in both the source and binary installations of Chord, is
+to run the following command:
 
+\begin{framed}
 \begin{verbatim}
-    prompt> java -cp <...>/chord.jar -D<key1>=<value1> ... -D<keyN>=<valueN> chord.project.Boot
+java -cp <...>/chord.jar -D<key1>=<val1> ... -D<keyN>=<valN> chord.project.Boot
 \end{verbatim}
+\end{framed}
 
-where \code{<...>} denotes the absolute or relative path of the directory containing file \code{chord.jar};
-that directory is also expected to contain any other binaries in Chord's installation
-(e.g., \code{libbuddy.so} and \code{libchord_instr_agent.so}).
+where \code{<...>} denotes the absolute or relative path of the
+directory containing file \code{chord.jar}; that directory is also
+expected to contain any shared libraries in Chord's installation (e.g.,
+\code{libbuddy.so} and \code{libchord_instr_agent.so}).
 
-Each ``\code{-D<key>=<value>}" argument above sets the system property named \code{<key>} to
-the value denoted by \code{<value>}.  The only way to specify inputs to Chord is via system properties; there is no
-command-line argument processing.
-Chapter\ref{chap:properties} describes all system properties recognized by Chord.
+Each ``\code{-D<key>=<val>}" argument above sets the system property
+named \code{<key>} to the value denoted by \code{<val>}.  The only
+way to specify inputs to Chord is via system properties; there is no
+command-line argument processing.  Chapter \ref{chap:properties}
+describes all system properties recognized by Chord.

### 2011-04-04 main/doc/program_representation.tex
documentation

@@
 (CFG).  The representation of a CFG (Section \ref{sec:cfgs}) consists
-of a set of local variables and a set of basic blocks.  Each basic
+of a set of registers and a set of basic blocks.  Each basic
 block contains zero or more primitive statements called quads
@@
 The quadcode representation of each method is a unique object of
-class \javadoc{joeq.Class.jq_Method}{joeq/Class/jq_Method.html}.  Components of the method (e.g.,
-its control-flow graph) can be accessed using the following API provided
+class \javadoc{joeq.Class.jq_Method}{joeq/Class/jq_Method.html}.  Components of the method,
+most notably its control-flow graph,
+can be accessed using the following API provided
 by that class.
@@
 
-The control-flow graph (CFG) of each method consists of a set of local
-variables of the method, called a ``register factory'', and a graph
-whose nodes are basic blocks and whose edges denote flow of control
-between basic blocks.
+The control-flow graph (CFG) of each method consists of a set of registers,
+called the register factory, and a graph whose nodes are basic blocks and
+whose edges denote flow of control between basic blocks.
 
@@
 
-The register factory contains one variable per argument of the method
-and one variable per temporary declared in the method body.
+\subsection{Register Factory}
+
+The register factory contains one register per argument of the method
+(called {\it local variables}) and one register per temporary 
+in the method body (called {\it stack variables}).
 Temporaries include those declared by programmers as well as those
-generated by Joeq.  The reason for the latter is that the quadcode
-representation is register-based and is constructed from Java
+generated by Joeq.  The reason Joeq can generate temporaries is that the quadcode
+representation, which is register-based, is constructed from Java
 bytecode, which is stack-based; moreover, Joeq does the Static Single
@@
 temporaries to ensure that there is at most one static assignment to
-any variable.
+any variable.  Registers corresponding to local variables are named {\tt R0}, {\tt R1}, ..., {\tt Rn},
+while those corresponding to stack variables are named {\tt Tn+1}, {\tt Tn+2}, ..., {\tt Tm}.
 
-The register factory is a unique object
+For instance, the register factory of the main method in the example above 
+has 3 registers: {\tt R0} denoting the {\tt args} argument of the method
+and {\tt T1} and {\tt T2} denoting temporaries generated by Joeq.
+
+Each register factory is a unique object
 of class \javadoc{joeq.Compiler.Quad.RegisterFactory}{joeq/Compiler/Quad/RegisterFactory.html}.
+
+\subsection{Basic Blocks}
 
@@
 Operator
-    |--- Move
-    |--- Phi
-    |--- Binary
-    |--- Unary
-    |--- New
-    |--- NewAray
-    |--- MultiNewArray
-    |--- CheckCast
-    |--- InstanceOf
-    |--- ALength
-    |--- ALoad
-    |--- AStore
-    |--- Getstatic
-    |--- Putstatic
-    |--- Getfield
-    |--- Putfield
-    |--- Return
+    |    Move, Phi, Unary, Binary, New, NewArray, MultiNewArray, 
+    |--- Getstatic, Putstatic, Getfield, Putfield, ALoad, AStore,
+    |    CheckCast, InstanceOf, ALength, Return
+    |
     |--- Branch
-    |       |--- IntIfCmp
-    |       |--- Goto
-    |       |--- Jsr
-    |       |--- Ret
-    |       |--- LookupSwitch
-    |       |--- TableSwitch
+    |       |--- IntIfCmp, Goto, Jsr, Ret, LookupSwitch, TableSwitch
+    |
     |--- Invoke
-    |       |--- InvokeVirtual
-    |       |--- InvokeStatic
-    |       |--- InvokeInterface
+    |       |--- InvokeVirtual, InvokeStatic, InvokeInterface
+    |
     |--- Monitor
-            |--- MONITORENTER
-            |--- MONITOREXIT
+            |--- MONITORENTER, MONITOREXIT
 \end{verbatim}
@@
     public void visitPhi(Quad q) { ... }
-    public void visitBinary(Quad q) { ... }
+    public void visitUnary(Quad q) { ... }
     ...

### 2011-04-12 main/doc/architecture.tex
updated documentation

@@
 \begin{center}
-\texonly{\includegraphics[scale=0.7]{chord_arch.png}}
-\htmlonly{{\htmlimg{chord_arch.png}}}
+\includeimage{0.7}{chord_arch.png}
+%\texonly{\includegraphics[scale=0.7]{chord_arch.png}}
+%\htmlonly{{\htmlimg{chord_arch.png}}}
 \end{center}
@@
 The Java program to be analyzed is also specified via properties.
-These properties are described in Section \ref{sec:program-props}.
+%These properties are described in Section \ref{sec:program-props}.
+Chapter \ref{chap:program-setup} describes how to setup a Java program for analysis using Chord.
 Chord analyzes Java bytecode, not Java source code, and thus only requires
@@
 
-\subsubsection*{Java Program Representation}
+\subsubsection*{Program Representation}
 
@@
 
-\subsubsection*{Computing Analysis Scope}
+\subsubsection*{Analysis Scope Construction}
 
 A pre-requisite to analyzing a Java program using any program analysis framework, including Chord, is
-to compute the {\it analysis scope}: the set of methods of the program to be analyzed.  Chord implements
+to compute the {\it analysis scope}: which parts of the program to analyze.  Chord implements
 several standard scope construction algorithms from the literature that differ in aspects such as
@@
 
-Chord provides various {\it analysis templates}: classes containing boilerplate code that can be
-extended by users to rapidly prototype different kinds of analyses.  An example is class RHSAnalysis,
-named after [Reps, Horowitz, and Sagiv 1995], which can be extended by users to write a summary-based inter-procedural
-context-sensitive static analysis by merely specifying the abstract domain and intra-procedural transfer functions.
-Another exampls is DynamicAnalysis, which can be extended by
-users to write a dynamic analysis by merely specifying which of various provided events to instrument, and
-the transfer functions for those events.
-
 A distinctive aspect of Chord is that each analysis is written modularly, independent of other analyses, along with lightweight
 annotations specifying the inputs and outputs of the analysis.
-Users must provide a unique name to each analysis, which allows the desired analyses to be called from the
-command-line, via a property named \code{chord.run.analyses}.
 Chord's runtime automatically computes
-producer-consumer relationships between analyses (e.g., determines which analysis produces as output a
+dependencies between analyses (e.g., determines which analysis produces as output a
 result that is needed as input by another analysis).  Before running a desired analysis,
@@
 
-Chord can be invoked in one of two modes: classic or modern (depending upon whether the value of property
-\code{chord.classic} is \code{true} (default) or \code{false}, respectively).
+Chord can be invoked in one of two modes: classic or modern.
 These two modes defer in the semantics of dependencies between analyses.  In particular, the classic mode
@@
 
-Chapter \ref{chap:analyses} describes the various analysis templates and how to connect disparate analyses
-together via dependencies in either of the above two modes.
+Chapter \ref{chap:project} describes how to write and run analyses in Chord.
+
+\subsubsection*{Analysis Templates}
+
+Chord provides various {\it analysis templates}: classes containing boilerplate code that can be
+extended by users to rapidly prototype different kinds of analyses.  An example is class RHSAnalysis,
+named after [Reps, Horowitz, and Sagiv 1995], which can be extended by users to write a summary-based inter-procedural
+context-sensitive static analysis by merely specifying the abstract domain and intra-procedural transfer functions.
+Another example is DynamicAnalysis, which can be extended by
+users to write a dynamic analysis by merely specifying which of various provided events to instrument, and
+the transfer functions for those events.
+Chapter \ref{chap:analysis-templates} describes the various analysis templates.
 
@@
 %each item \code{<N>} in the value of property \code{chord.run.ids}).
+
+\subsubsection*{Predefined Analyses}
+
+Chapter \ref{chap:predefined-analyses} describes many standard analyses that are implemented in Chord.

### 2011-04-12 main/doc/datalog_analysis.tex
updated documentation

@@
 
-A common way to rapidly prototype program analyses in Chord is using a declarative
+A common way to rapidly prototype analyses in Chord is using a declarative
 logic-programming language called Datalog.  This chapter describes all aspects of
-Datalog analysis in Chord.
-Section \ref{sec:writing-datalog-analysis} explains how to write a Datalog analysis,
-Section \ref{sec:running-datalog-analysis} explains how to run it, and
+Datalog analyses in Chord.
+Section \ref{sec:writing-datalog-analysis} explains how to write and run a Datalog analysis.
 Section \ref{sec:tuning-datalog-analysis} explains how to tune its performance.
@@
 specifying how to compute the output relations from the input relations.
-An example of such an analysis is shown below:
+It can be defined in any file with extension {\tt .dlog}
+in any directory in the path specified by property \code{chord.dlog.analysis.path}.
+An example Datalog analysis is shown below:
 
@@
 
-Any line that begins with a {\tt \#} is regarded a comment, except a
-line of the form ``{\tt \#} {\tt name=<...>}", which specifies the name
-{\tt <...>} of the Datalog analysis.
+Any line that begins with a ``{\tt \#}" is regarded a comment, except a
+line of the form ``\code{# name=<ANALYSIS_NAME>}'', which specifies the name
+\code{<ANALYSIS_NAME>} of the Datalog analysis.
 Each Datalog analysis is expected to have exactly one such line.
 The above Datalog analysis is named {\tt datarace-dlog}.
-In Chord, all Datalog analysis names have suffix
-{\tt -dlog} and all Java analysis names have suffix {\tt -java}, but
-this is merely a convention that analysis writers are free to deviate from.
-The name of each analysis, written in Datalog or in Java, is expected to be
-unique across all analyses in scope (i.e., across all analyses in paths
-specified by properties \code{chord.dlog.analysis.path} and \code{chord.java.analysis.path}).
 
-The {\tt .include "<...>.dom"} lines specify each program domain named
-{\tt <...>} that is needed by the Datalog analysis, i.e., each domain over which
+The ``\code{.include "<DOM_NAME>.dom"}'' lines specify each program domain named
+\code{<DOM_NAME>} that is needed by the Datalog analysis, i.e., each domain over which
 any program relation that is input/output by the Datalog analysis is defined. 
@@
 in the above example, {\tt candidate} is a binary relation, both of whose
-attributes have domain E, and they are distinguished as E0 and E1.
+attributes have domain {\tt E}, and they are distinguished as {\tt E0} and {\tt E1}.
 
@@
 on the order of these BDD variables.
-The {\tt .bddvarorder <...>} line in the Datalog analysis enables the Datalog
+The ``\code{.bddvarorder <BDD_VAR_ORDER>}'' line in the Datalog analysis enables the Datalog
 analysis writer to specify this order.
 It must list all domains along with their numerical suffixes, separated
-by {\tt \_} or {\tt x}.
-Using a {\tt \_} between two domains, such as {\tt T0\_T1}, means that the BDD variables assigned
+by `{\tt \_}' or `{\tt x}'.
+Using a `{\tt \_}' between two domains, such as {\tt T0\_T1}, means that the BDD variables assigned
 to domain {\tt T0} precede those assigned to domain {\tt T1} in the BDD variable
 order for this Datalog analysis.
-Using a {\tt x} between two domains, such as {\tt E0xE1}, means that the
+Using a `{\tt x}' between two domains, such as {\tt E0xE1}, means that the
 BDD variables assigned to domains {\tt E0} and {\tt E1}
@@
 Each Datalog analysis rule is a Horn clause of the form
-{\tt R(t) :- R1(t1), ..., Rn(tn)}
+``{\tt R(t) :- R1(t1), ..., Rn(tn)}"
 meaning that if relations {\tt R1}, ..., {\tt Rn} contain tuples {\tt t1}, ..., {\tt tn}
@@
 
-\section{Running a Datalog Analysis}
-\label{sec:running-datalog-analysis}
+Running a Datalog analysis is no different from running any other kind of analysis
+in Chord.  See Chapter \ref{chap:running-analysis} for how to run an analysis.
 
-Chord interprets each file with extension {\tt .dlog} or {\tt .datalog} in the path specified by
-property \code{chord.dlog.analysis.path} as a program analysis expressed in Datalog.
-
-{\bf Under Construction}
-
-\section{Tuning a Datalog Analysis}
+\section{Tuning Performance}
 \label{sec:tuning-datalog-analysis}
@@
 \begin{verbatim}
-ant -Ddlog.file=<file> -Dwork.dir=<dir> solve
+ant -Ddlog.file=<FILE> -Dwork.dir=<DIR> solve
 \end{verbatim}
@@
 
-where {\tt <file>} denotes the file defining the Datalog analysis
-to be tuned, and {\tt <dir>} is the directory containing the
+where {\tt <FILE>} is the file defining the Datalog analysis
+to be tuned, and {\tt <DIR>} is the directory containing the
 program domains ({\tt *.dom} files) and program relations ({\tt *.bdd} files)
@@
 \begin{enumerate}
+
 \item
+
 Set properties \verb+noisy=yes+, \verb+tracesolve=yes+, and \verb+fulltracesolve=yes+
@@
 \item
+
 Once you have ensured that none of the rules is getting ``stuck",
@@
 \item
+
 You can add the \verb+.split+ keyword at the end of certain rules as a
@@
 \item
+
 You can try to decompose a single Datalog analysis file into two separate Datalog analysis
@@
 \item
+
 Observe the sizes of the BDDs representing the relations that are
@@
 \item
+
 The relative order of values *within* domains (e.g.,
@@
 \item
+
 Finally, it is worth emphasizing that BDDs are not magic.
@@
 
-If two domains are uninterleaved in the declared domain order in a Datalog program (i.e., {\tt \_} is used instead of {\tt x} between them),
+If two domains are uninterleaved in the declared domain order in a Datalog program (i.e., `{\tt \_}' is used instead of `{\tt x}' between them),
 then the BDD variables assigned to those domains are ordered in reverse order in the underlying BDD factory.
@@
 
-If two domains are interleaved in the declared domain order in a Datalog program (i.e., {\tt x} is used instead of {\tt \_} between them),
+If two domains are interleaved in the declared domain order in a Datalog program (i.e., `{\tt x}' is used instead of `{\tt \_}' between them),
 then the BDD variables assigned to those domains are still ordered in reverse order of IDs in the underlying BDD factory,

### 2011-04-12 main/doc/predefined_analyses.tex
updated documentation

@@
 
-To run this analysis, run the following command in Chord's main directory:
+To run this analysis, run the following command:
 
@@
 \begin{verbatim}
-ant -Dchord.work.dir=<...> -Dchord.run.analyses=cipa-0cfa-dlog run
+ant -Dchord.work.dir=<WORK_DIR> -Dchord.run.analyses=cipa-0cfa-dlog run
 \end{verbatim}
 \end{framed}
+
+where \code{<WORK_DIR>} is a directory
+containing a file named \code{chord.properties} that defines
+properties \code{chord.main.class} and \code{chord.class.path} specifying
+the main class and the application classpath, respectively, of the program
+to be analyzed.
 
@@
 \item rootM subset M contains the set of entry methods that may be reachable; this includes the program's main method
-as well as each static initializer method that may be reachable from the main method.
+as well as each static initializer method that may be reachable from the main
+method.
 \item reachableM subset M contains the set of methods that may be reachable from the program's main method.
@@
 
-In a context-sensitive analysis, there is no longer just one abstract memory location per object allocation site.
-Rather, the set of objects a reference can point to depends on the \textit{context} in which the method containing the
-reference is called.
-Whereas a context-insensitive analysis talks about the domain of methods (M) and the domain
-of allocation sites (H), a context-sensitive analysis talks about the domain of abstract contexts, labeled C.
-Elements of domain C contain both abstract calling contexts and abstract objects. (These are merged for reasons
-described below.)
+In a context-sensitive analysis, there is no longer just one abstract memory
+location per object allocation site.  Rather, the set of objects a reference can
+point to depends on the \textit{context} in which the method containing the
+reference is called.  Whereas a context-insensitive analysis talks about the
+domain of methods (M) and the domain of allocation sites (H), a
+context-sensitive analysis talks about the domain of abstract contexts, labeled
+C.  Elements of domain C contain both abstract calling contexts and abstract
+objects. (These are merged for reasons described below.)
 
-Chord has several context-sensitive analyses, but they all expose the same relations, which are
-described below:
+Chord has several context-sensitive analyses, but they all expose the same
+relations, which are described below:
 
@@
 \begin{itemize}
-\item C is the domain of all abstract calling contexts and abstract objects.  Each element in this domain is a sequence of zero or more sites,
-where each site can be a call site or an object allocation site.  A sequence may have mixed call sites and object allocation sites.
-The most significant site (i.e. the first site) in each sequence is called the head; the remaining sub-sequence is called the tail.  The
-below three relations relate a sequence with its head and tail.
+\item C is the domain of all abstract calling contexts and abstract objects.
+Each element in this domain is a sequence of zero or more sites, where each site
+can be a call site or an object allocation site.  A sequence may have mixed call
+sites and object allocation sites.  The most significant site (i.e. the first
+site) in each sequence is called the head; the remaining sub-sequence is called
+the tail.  The below three relations relate a sequence with its head and tail.
 \item CC subset (C $\times$ C) contains tuples \texttt{(c1,c2)} such that context \texttt{c2} is the tail of context \texttt{c1}.
@@
 \item CVC subset (C $\times$ V $\times$ C) contains tuples \texttt{(c,v,o)} such that local variable \texttt{v} may point to
-object \texttt{o} in context \texttt{c} of that variable's declaring method. Note that both \texttt{o} and \texttt{c} are elements of domain C.
+object \texttt{o} in context \texttt{c} of that variable's declaring
+method. Note that both \texttt{o} and \texttt{c} are elements of domain C.
 \item CFC subset (C $\times$ F $\times$ C) contains tuples \texttt{(o1,f,o2)} such that instance field \texttt{f} of object \texttt{o1} may point to object \texttt{o2}.
@@
 
-{\bf Under Construction: explain object-sensitive vs context-sensitive, plus how to invoke.}
+{\bf Under Construction: explain object-sensitive vs context-sensitive, plus how
+to invoke.}
 
@@
 
-To run Chord's static datarace analysis, run the following
-command in Chord's main directory:
+To run Chord's static datarace analysis, run the following command:
 
@@
 \begin{verbatim}
-ant -Dchord.work.dir=<...> -Dchord.run.analyses=datarace-java run
+ant -Dchord.work.dir=<WORK_DIR> -Dchord.run.analyses=datarace-java run
 \end{verbatim}
@@
 
-where {\tt <...>} denotes the absolute or relative path of a directory
-containing a file named \code{chord.properties} which defines
+where \code{<WORK_DIR>} is a directory
+containing a file named \code{chord.properties} that defines
 properties \code{chord.main.class}, \code{chord.class.path}, and
-\code{chord.src.path}.  See Section \ref{sec:program-props} for the
-meaning of these properties.
+\code{chord.src.path} specifying the main class, the application classpath, and
+the Java source path, respectively, of the program to be analyzed.
 
-Directory \code{examples/datarace_test/} in Chord's main directory provides a toy Java
-program on which one can run the datarace analysis.  First run {\tt
-  ant} in that directory (in order to compile the program's {\tt
-  .java} files to {\tt .class} files) and then run the above command
-in Chord's main directory with {\tt <...>} replaced by
+Directory \code{<CHORD_MAIN_DIR>/examples/datarace_test/} provides a toy Java
+program on which one can run the datarace analysis.  First run {\tt ant} in that
+directory (in order to compile the program's {\tt .java} files to {\tt .class}
+files) and then run the above command with \code{<WORK_DIR>} replaced by
 \code{examples/datarace_test/}.  Upon successful completion, the
@@
 \item
-File \code{dataraces_by_fld.html}, listing all dataraces grouped by
-the field on which they occur; all dataraces on the same instance
-field or the same static field are listed in the same group, and so
-are all dataraces on array elements.
+File \code{dataraces_by_fld.html}, listing all dataraces grouped by the field on
+which they occur; all dataraces on the same instance field or the same static
+field are listed in the same group, and so are all dataraces on array elements.
 \item
-File \code{dataraces_by_obj.html}, listing all dataraces grouped by
-the abstract object on whose field they occur; dataraces on all static
-fields are listed in the same group, and so are dataraces on different
-instance fields of the same abstract object.
+File \code{dataraces_by_obj.html}, listing all dataraces grouped by the abstract
+object on whose field they occur; dataraces on all static fields are listed in
+the same group, and so are dataraces on different instance fields of the same
+abstract object.
 \end{itemize}
@@
 
-To run Chord's static deadlock analysis,  run the following
-command in Chord's main directory:
+To run Chord's static deadlock analysis,  run the following command:
 
@@
 \begin{verbatim}
-ant -Dchord.work.dir=<...> -Dchord.run.analyses=deadlock-java run
+ant -Dchord.work.dir=<WORK_DIR> -Dchord.run.analyses=deadlock-java run
 \end{verbatim}
@@
 
-where {\tt <...>} denotes the absolute or relative path of a directory
-containing a file named \code{chord.properties} which defines
+where \code{<WORK_DIR>} is a directory
+containing a file named \code{chord.properties} that defines
 properties \code{chord.main.class}, \code{chord.class.path}, and
-\code{chord.src.path}.  See Section \ref{sec:program-props} for the
-meaning of these properties.
+\code{chord.src.path} specifying the main class, the application classpath,
+and the Java source path, respectively, of the program to be analyzed.
 
-Directory \code{examples/deadlock_test/} in Chord's main directory provides a toy Java
-program on which one can run the deadlock analysis.  First run {\tt
-  ant} in that directory (in order to compile the program's {\tt
-  .java} files to {\tt .class} files) and then run the above command
-in Chord's main directory with {\tt <...>} replaced by
+Directory \code{<CHORD_MAIN_DIR>/examples/deadlock_test/} provides a toy Java
+program on which one can run the deadlock analysis.  First run {\tt ant} in that
+directory (in order to compile the program's {\tt .java} files to {\tt .class}
+files) and then run the above command with \code{<WORK_DIR>} replaced by
 \code{examples/deadlock_test/}.  Upon successful completion, the

### 2011-04-12 main/doc/properties.tex
updated documentation

@@
+\xname{properties}
 \chapter{Chord Properties}
@@
 
-The only way to specify inputs to Chord is by means of system properties.
-There is no command-line argument processing in Chord and any command-line arguments are ignored.
-Section \ref{sec:properties-setting} explains how to set properties and
-Section \ref{sec:properties-meaning} explains the meaning of properties recognized by Chord.
+The only way to specify inputs to Chord is by means of system properties.  There
+is no command-line argument processing in Chord and any command-line arguments
+are ignored.  Section \ref{sec:properties-setting} explains how to set
+properties and Section \ref{sec:properties-meaning} explains the meaning of
+properties recognized by Chord.
+Notation {\tt [<...>]} is used in this chapter to denote the value of the property named {\tt <...>}.
 
@@
 
-A property can be passed to Chord in any of several ways.
-The reason for providing multiple ways is to provide users shorthand ways for defining properties
-once and for all for a particular input Java program, or even once and for all across all Chord runs. 
-The following are the different ways by which a property can be passed to Chord in decreasing order of precedence:
+A property can be passed to Chord in any of several ways.  The reason for
+providing multiple ways is to provide users with shorthand ways for defining
+properties once and for all for a particular Java program under analysis, or
+even once and for all across all Chord runs.  The following are the different
+ways by which a property can be passed to Chord in decreasing order of
+precedence:
 
@@
 
-On the command-line via the \code{-D<key>=<value>} format.
+{\bf How:} On the command-line via the ``\code{-D<key>=<val>}" format.
 
-Use this option to specify properties specific to the current run of Chord.
+{\bf When:} Use this option to specify properties specific to the {\it current run} of Chord.
 
-Typical usage of this option is by running the following command in Chord's \code{main/} directory:
+Typical usage of this option is by running the following command:
+
+\begin{framed}
 \begin{verbatim}
-    prompt> ant -D<key1>=<value1> ... -D<keyN>=<valueN> run
+ant -D<key1>=<val1> ... -D<keyN>=<valN> run
 \end{verbatim}
+\end{framed}
 
@@
 
-Via the file specified by property \code{chord.props.file}.
+{\bf How:} Via a user-defined {\it properties file} whose location is specified by property \code{chord.props.file}.
 
-Use this option to specify once and for all properties of the program to
-be analyzed (e.g., property \code{chord.main.class} specifying the name of that program's main class, property \code{chord.class.path}
-specifying the application classpath of that program, etc.).
+{\bf When:} Use this option to specify once and for all properties of the
+Java program to be analyzed (e.g., property \code{chord.main.class} specifying the
+name of that program's main class).
+Chapter \ref{chap:program-setup} presents an example properties file 
+that defines the program properties that are most commonly used.  Section \ref{sec:program-props}
+presents all program properties that are recognized by Chord.
 
-The default value of this property is \code{[chord.work.dir]/chord.properties} and typically
-does not need to be changed by users.
-Property \code{chord.work.dir} specifies the directory in which Chord will run.
-The default value of this property is the current directory but users must typically set
-it on the command line, as follows:
+The default value of property \code{chord.props.file} is \code{[chord.work.dir]/chord.properties}.
+Property \code{chord.work.dir} specifies the directory in which Chord must run; its default value
+is the current directory.
 
+There are three ways to use this option of setting properties:
+
+The first way is to override the default value of property \code{chord.work.dir} on the command-line.
+This requires naming the above properties file as \code{chord.properties}, placing it in the
+directory in which Chord will run, denoted \code{<WORK_DIR>}
+(e.g., this could be the top-level directory of the program to be analyzed),
+and running the following command:
+
+\begin{framed}
 \begin{verbatim}
-    prompt> ant -Dchord.work.dir=<...> run
+ant -Dchord.work.dir=<WORK_DIR> run
 \end{verbatim}
-where \code{<...>} denotes the absolute or relative path of the top-level directory of
-the program to be analyzed, which contains a file named \code{chord.properties} that users
-must create themselves and populate with properties that are recognized by chord and
-are specific to that program.
+\end{framed}
+
+The second way is to override the default value of property \code{chord.props.file} on the command-line.
+In this case, the properties file can be in any user-desired location, denoted \code{<PROPS_FILE>}, and Chord
+will run in the current directory:
+
+\begin{framed}
+\begin{verbatim}
+ant -Dchord.props.file=<PROPS_FILE> run
+\end{verbatim}
+\end{framed}
+
+The third (and most flexible) way is to override the default values of both properties \code{chord.work.dir}
+and \code{chord.props.file} on the command-line.
 
@@
 
-Via file \code{main/chord.properties}.
+{\bf How:} Via the properties file named \code{chord.properties} that is already provided in Chord's main directory.
 
-Use this option to specify once and for all properties you would like to hold in every run of Chord
-(e.g., property \code{chord.max.heap} specifying the maximum heap memory size to be used by the
-JVM running Chord).
+{\bf When:} Use this option to specify once and for all properties you would
+like to hold in {\it every run} of Chord (e.g., property \code{chord.max.heap}
+specifying the maximum heap memory size to be used by the JVM running Chord).
 \end{enumerate}
 
-\section{Meaning of Properties}
+\section{Recognized Properties}
 \label{sec:properties-meaning}
@@
 The separator for list-valued properties can be either a blank space, a comma, a colon, or a semi-colon.
-Notation {\tt [<...>]} is used in this section to denote the value of the property named {\tt <...>}.
 
@@
 be defined in this file to avoid defining them on the command line
-(using the \code{-D<key>=<value>} format) every time Chord is run.
+(using the \code{-D<key>=<val>} format) every time Chord is run.
 Each relative file/directory name in the value of any property defined
@@
 {\bf Note:} Chord analyzes only Java bytecode, not Java source code.
-This property is used only by the task of converting Java source files into HTML files by analyses that need to present their analysis results at the Java source code level (by calling method \code{chord.program.Program.g().HTMLizeJavaSrcFiles())}.
+This property is used only by the task of converting Java source files into HTML files by analyses that need to present their results at the Java source code level (by calling method \code{chord.program.Program.g().HTMLizeJavaSrcFiles())}.
 \end{quote}
@@
 \begin{quote}
-{\bf Type:} {\tt [none|static|dynamic|static\_cast]} \\
-{\bf Description:} TODO \\
+{\bf Type:} {\tt [none|dynamic|static|static\_cast]} \\
+{\bf Description:} Algorithm to resolve reflection.  The choices are {\tt none} (do not resolve any reflection),
+{\tt dynamic} (run the program and observe how reflection is resolved),
+{\tt static} (resolve reflection statically but without analyzing casts), and
+{\tt static\_cast} (resolve reflection statically by analyzing casts). \\
 {\bf Default value:} {\tt none}
@@
 {\bf Default value:} {\tt ""} \\
-{\bf Note:} If the analysis is written in Java, its name is specified via statement {\tt name=<...>} in its {\tt @Chord} annotation.
-If the analysis is written in Datalog, its name is specified via a line of the form ``{\tt \# name=<...>}".
+{\bf Note:} If the analysis is written in Java, its name is specified via statement {\tt name="..."} in its {\tt @Chord} annotation.
+If the analysis is written in Datalog, its name is specified via a line of the form ``{\tt \# name=...}".
 \end{quote}
@@
 {\bf Default value:} {\tt ""} \\
-{\bf Note:} Specify each method in format \code{<mname>:<mdesc>@<cname>} where {\tt <mname>} is the method's name, {\tt <mdesc>} is the method's descriptor, and {\tt <cname>} is the name of the method's declaring class.
-In {\tt <cname>}, use `.' instead of `/', and use {\tt \#} instead of the dollar character.
+{\bf Note:} Specify each method in format \code{mname:mdesc@cname} where {\tt mname} is the method's name, {\tt mdesc} is the method's descriptor, and {\tt cname} is the name of the method's declaring class.
+In {\tt cname}, use `.' instead of `/', and use `{\tt \#}' instead of `{\tt \$}'.
 \end{quote}
@@
 {\bf Type:} string list  \\
-{\bf Description:} List of names of program relations whose contents must be printed to files \code{[chord.out.dir]/<...>.txt} where {\tt <...>} denotes the relation name. \\
+{\bf Description:} List of names of program relations whose contents must be printed to files \code{[chord.out.dir]/<RELNAME>.txt} where
+{\tt <RELNAME>} denotes the relation name. \\
 {\bf Default value:} {\tt ""} \\
-{\bf Note:} This functionality must be used with caution as certain program relations, albeit represented compactly as BDDs, may contain a large number (e.g., millions) of tuples, resulting in voluminous output when printed in explicit form to a text file.
-See Section \ref{sec:tuning-datalog-analysis} for a more efficient way to query the contents of program relations (namely, by using the {\tt debug} target provided in file {\tt build.xml} in Chord's \code{main/} directory).
+{\bf Note:} This functionality must be used with caution as certain program relations, albeit represented compactly as BDDs, may contain
+ a large number (e.g., millions) of tuples, resulting in voluminous output when printed in explicit form to a text file.  See Section
+ \ref{sec:tuning-datalog-analysis} for a more efficient way to query the contents of program relations (namely, by using the {\tt debug}
+ target provided in file {\tt build.xml} in Chord's main directory).
 \end{quote}
@@
 {\bf Type:} bool \\
-{\bf Description:} Create files \code{targets_sortby_name.html}, \code{targets_sortby_kind.html}, and \code{targets_sortby_producers.html} in directory \code{[chord.out.dir]}, publishing all tasks and targets defined by analyses in paths \code{[chord.java.analysis.path]} and \code{[chord.dlog.analysis.path]}.
-\\
+{\bf Description:} Create files \code{targets_sortby_name.html}, \code{targets_sortby_kind.html}, and \code{targets_sortby_producers.html}
+ in directory \code{[chord.out.dir]}, publishing all tasks and targets defined by analyses in paths \code{[chord.java.analysis.path]} and
+ \code{[chord.dlog.analysis.path]}.  \\
 {\bf Default value:} {\tt false}
@@
 {\bf Type:} bool \\
-{\bf Description:} TODO \\
-{\bf Default value:} TODO
+{\bf Description:} Whether to use the classic project (as opposed to the modern project).  See Chapter\ref{chap:analyses} for the difference
+ between the two kinds of projects. \\
+{\bf Default value:} \code{true}
 \end{quote}
@@
 {\bf Description:} Partial classpath of analyses written in Java (i.e., {\tt @Chord}-annotated classes).
-Conventionally, it includes all such analyses that are predefined in Chord.  \\
-{\bf Default value:} The absolute path of file \code{chord.jar}
+Conventionally, it includes all Java analyses that are predefined in Chord.  \\
+{\bf Default value:} The absolute path of file \code{chord.jar}.
 \end{quote}
@@
 {\bf Description:} Partial classpath of analyses written in Java (i.e., {\tt @Chord}-annotated classes).
-Conventionally, it includes all such external analyses that users want to run. \\
+Conventionally, it includes all user-defined analyses. \\
 {\bf Default value:} {\tt ""}
@@
 {\bf Type:} path \\
-{\bf Description:} Partial classpath of analyses written in Datalog (i.e., files with suffix {\tt .dlog} or {\tt .datalog}).
-Conventionally, it includes all such analyses that are predefined in Chord.  \\
-{\bf Default value:} The absolute path of file \code{chord.jar}
+{\bf Description:} Partial path of analyses written in Datalog (i.e., files with suffix {\tt .dlog}).
+Conventionally, it includes all Datalog analyses that are predefined in Chord. \\
+{\bf Default value:} The absolute path of file \code{chord.jar}.
 \end{quote}
@@
 {\bf Type:} path \\
-{\bf Description:} Partial classpath of analyses written in Datalog (i.e., files with suffix {\tt .dlog} or {\tt .datalog}).
-Conventionally, it includes all such external analyses that users want to run. \\
+{\bf Description:} Partial path of analyses written in Datalog (i.e., files with suffix {\tt .dlog}).
+Conventionally, it includes all user-defined Datalog analyses. \\
 {\bf Default value:} {\tt ""}
@@
 {\bf Type:} path  \\
-{\bf Description:} Complete classpath of analyses written in Datalog (i.e., files with suffix {\tt .dlog} or {\tt .datalog}). \\
+{\bf Description:} Complete path of analyses written in Datalog (i.e., files with suffix {\tt .dlog}). \\
 {\bf Default value:} \code{[chord.std.dlog.analysis.path]:[chord.ext.dlog.analysis.path]}
@@
 %{\tt true}
-%{\bf Note:} This functionality is useful for debugging Datalog programs using the {\tt debug} target provided in file {\tt build.xml} in Chord's {\tt main/} directory (see Section \ref{sec:tuning-datalog-analysis}).
+%{\bf Note:} This functionality is useful for debugging Datalog programs using the {\tt debug} target provided in file {\tt build.xml} in Chord's main directory (see Section \ref{sec:tuning-datalog-analysis}).

### 2011-04-12 main/doc/writing_analysis.tex
updated documentation

--- 2010-04-01 main/doc/writing_analysis.tex
@@
-\section{Writing a Program Analysis}
-\label{sec:writing-analysis}
+\xname{writing_analysis}
+\chapter{Writing an Analysis}
+\label{chap:writing-analysis}
 
-TODO
+Chord provides several {\it analysis templates}: classes containing boilerplate
+code that can be extended by users to rapidly prototype different kinds of
+analyses.  These templates are organized in the following hierarchy in
+package \javadoc{chord.project.analyses}{chord/project/analyses/package-summary.html}:
+
+\begin{verbatim}
+JavaAnalysis
+    |
+    |--- ProgramDom
+    |
+    |--- ProgramRel
+    |
+    |--- DlogAnalysis
+    |
+    |--- RHSAnalysis
+    |        |
+    |        |--- ForwardRHSAnalysis
+    |        |
+    |        |--- BackwardRHSAnalysis
+    |
+    |--- BasicDynamicAnalysis
+             |
+             |--- DynamicAnalysis
+\end{verbatim}
+
+All these classes implement
+interface \javadoc{chord.project.ITask}{chord/project/ITask.html} and can
+therefore serve as tasks.  We next describe each of them in more detail.
+
+\section{JavaAnalysis}
+\label{sec:java-analysis}
+
+Class \code{chord.project.analyses.JavaAnalysis} provides the most general
+template for writing an analysis imperatively in Java.  An analysis can be
+created using this template by extending this class as follows:
+
+\begin{verbatim}
+@Chord(
+    name = "example-java",
+    consumes = { "C1", ..., "Cn" },
+    produces = { "P1", ..., "Pm" },
+    types = { A1.class, ..., An.class, B1.class, ..., Bm.class },
+    namesOfTypes = { "C1", ..., "Cn", "P1", ..., "Pm" },
+)
+public class ExampleAnalysis extends JavaAnalysis {
+    @Override
+    public void run() {
+        ClassicProject project = ClassicProject.g();
+
+        A1 c1 = (A1) project.getTrgt("C1");
+        ...
+        An cn = (An) project.getTrgt("Cn");
+
+        B1 p1 = (B1) project.getTrgt("P1");
+        ...
+        Bm pm = (Bm) project.getTrgt("Pm");
+
+        ... // analysis code computing p1, ..., pm from c1, ..., cn
+    }
+}
+\end{verbatim}
+
+The class must be compiled to a {\tt .class} file that occurs in
+some element (directory or jar/zip file) of the path
+specified by property \code{chord.java.analysis.path}.  Chord represents the
+corresponding analysis using a unique object of the class.
+
+The class is annotated with an {\tt @Chord} annotation defined in
+class \javadoc{chord.project.Chord}{chord/project/Chord.html}.  This annotation
+allows specifying the name of the analysis ({\tt example-java}), the names of
+the targets that are directly consumed ({\tt C1}, ..., {\tt Cn}) and directly
+produced ({\tt P1}, ..., {\tt Pm}) by the analysis, and the types of targets.
+There is a one-to-one correspondence between the arrays denoted by fields {\tt
+types} and {\tt namesOfTypes}, e.g., the type of the target named {\tt C1} is
+class {\tt A1}, and so on.  In principle, the type of {\it any} target in the
+project can be specified here.  In practice, however, the types of only the
+targets declared as directly consumed/produced by this analysis are specified.
+Moreover, although the type of each target that is directly consumed/produced by
+the above analysis is specified in the above annotation, in practice the types
+of hardly any targets need to be explicitly specified, because they can be
+automatically inferred by Chord from analyses created using more specialized
+templates discussed below that also consume/produce those targets.
+
+The code of the analysis must be supplied in the {\tt run()} method.  This
+method typically retrieves the representation of each directly consumed/produced
+target from the project, reads (the state of) the consumed targets, performs
+some computation over it, and writes to (the state of) the produced targets.
+
+The analysis templates presented in the following sections are more specialized
+forms of the JavaAnalysis template in that they restrict what the directly
+consumed/produced targets and the analysis code in the {\tt run()} method can
+be.
+
+%called JavaAnalysis (Section \ref{sec:java-analysis}), and the other for writing
+%analyses declaratively in Datalog, called DlogAnalysis
+%(Section \ref{sec:dlog-analysis}).  JavaAnalysis is the most general template:
+%it least constrains the kind of analysis that can be written but also offers the
+%least boilerplate code for writing an analysis.  Chord provides several
+%additional templates for writing analyses imperatively in Java that are
+%specialized forms of JavaAnalysis, namely, ProgramDom
+%(Section \ref{sec:program-dom}), ProgramRel (Section \ref{sec:program-rel}),
+%RHSAnalysis (Section \ref{sec:rhs-analysis}), and DynamicAnalysis
+%(Section \ref{sec:dynamic-analysis}).  We next explain each of these templates.
+
+\section{DlogAnalysis}
+\label{sec:dlog-analysis}
+
+A common way to rapidly prototype analyses in Chord is using a declarative
+logic-programming language called Datalog.  A Datalog analysis can be defined in
+any {\tt .dlog} file that occurs in some element (directory or jar/zip file)
+of the path specified by property \code{chord.dlog.analysis.path}.  Chord represents each such analysis
+using a unique object of class \javadoc{chord.project.analyses.DlogAnalysis}{chord/project/analyses/DlogAnalysis.html}.
+
+A Datalog analysis declares a bunch of input/output {\it program relations},
+each over one or more {\it program domains}, and provides a bunch of rules
+(constraints) specifying how to compute the output relations from the input
+relations.
+
+TODO: explain inputs, outputs, code
+
+Briefly, each Datalog analysis specifies the following:
+\begin{itemize}
+\item
+a set of program domains D1, ..., Dk; a program domain is a collection of program artifacts of a particular kind.
+\item
+a set of input program relations I1, .., In, each over one of more program domains D1, ..., Dk;
+a program relation
+\item
+I1, ..., In, each over one or more program domains D1, ..., Dk
+, of which , a set of output program
+relations O1, ..., Om, a set of rules specifying how to compute the output relations from the input
+relations, and a set of program domains D1, ..., Dk.  
+\end{itemize}
+
+Chapter \ref{chap:datalog-analysis} explains how to write a Datalog analysis in
+more detail.
+
+\section{ProgramDom}
+\label{sec:program-dom}
+
+\section{ProgramRel}
+\label{sec:program-rel}
+
+\section{DynamicAnalysis}
+\label{sec:dynamic-analysis}
+
+\section{RHSAnalysis}
+\label{sec:rhs-analysis}
+
+%\begin{itemize}
+%\item
+%B is the domain of basic blocks.
+%\item
+%C is the domain of abstract calling contexts.
+%\item
+%E is the domain of program points that read or write a field.
+%\item
+%F is the domain of fields inside a class.
+%\item
+%H is the domain of allocation sites.
+%\item
+%I is the domain of invocation statements (method calls)
+%\item
+%L is the domain of lock acquisition points.
+%\item
+%M is the domain of program methods.
+%\item
+%P is the domain of simple statements. %SAY MORE HERE?
+%\item
+%R is the domain of lock release points.
+%\item
+%T is the domain of types (classes and primitives).
+%\item
+%V is the domain of reference-typed variables.
+%\item
+%W is the domain of loop head statements. %SAY MORE HERE?
+%\item
+%Z is a domain of integers corresponding to argument positions.The size of the domain is chosen to 
+%accommodate the method with the greatest number of arguments.
+%\end{itemize}
+
+%\section{Core relations}
+
+%Chord includes code to build a very large number of relations.
+These standard relations are very useful in writing your own analyses.
+For example, relation \texttt{sub} includes all pairs of types \texttt{(t1, t2)} where \texttt{t2} is a subclass of \texttt{t1}.
+Most of these standard relations are in the package \texttt{chord.rels}.
+
+%A Java analysis should extend JavaAnalysis or ProgramRel.
+It's usually easiest if an analysis only produces one relation, in which case extending ProgramRel is the best approach.
+If you extend ProgramRel, all the work of creating and saving the relation is taken care of for you; all you need to do is add tuples to the relation.Typically you do this by calling \texttt{super.add(...)} with the tuple you wish to add to the relation.
+For forther convenience, you can implement one of the Visitor interfaces defined in \texttt{chord.program.visitors}.
+This will force you to implement some \texttt{visit} methods that iterate over all elements in some domain, such as all methods or all invocation points.
+
+%Sometimes, you need to create two different relations at the same time. In this case, the right approach is to extend JavaAnalysis, and explicitly create and save the relations on your own.
+
+%Regardless of which of these approaches you take, every Java analysis must have an \texttt{@Chord} annotation.
+This annotation is a set of named fields.
+At a minimum, this annotation must specify the name of the analysis.
+If you extend ProgramRel, the name of the analysis will be treated as the same as that of the generated relation.
+For example, \texttt{@Chord(name = "F")} is a valid annotation.
+The name of the Java class that generates the relation is ignored by Chord -- only the annotation matters.
+
+%\textbf{MAYUR, IS SIGN MANDATORY?   IF NOT, WHEN CAN IT BE OMITTED?}
+
+%There are also other annotation fields available.  You may specify the dependencies of your analysis via \texttt{consumes = \{ "xyz"\}}.
+
+%If your analysis does not extend ProgramRel, you should also specify the relations that your analysis produces. You do this via the \texttt{produces} relation.
+
+%Should mention the signs fields.

### 2011-04-15 main/doc/arch.tex
more documentation

--- 2011-04-12 main/doc/architecture.tex
@@
-\xname{architecture}
+\xname{arch}
 \chapter{Architecture of Chord}
-\label{chap:architecture}
+\label{chap:arch}
 
@@
 \includeimage{0.7}{chord_arch.png}
-%\texonly{\includegraphics[scale=0.7]{chord_arch.png}}
-%\htmlonly{{\htmlimg{chord_arch.png}}}
 \end{center}
@@
 
-All inputs to Chord are specified by means of system properties. 
-Chapter \ref{chap:properties} describes how to set properties and the meaning of 
-each property that is recognized by Chord.
+All inputs to Chord are specified by means of system properties.  Chapter
+\ref{chap:properties} describes how to set properties and the meaning of each
+property that is recognized by Chord.
 
-The Java program to be analyzed is also specified via properties.
-%These properties are described in Section \ref{sec:program-props}.
-Chapter \ref{chap:program-setup} describes how to setup a Java program for analysis using Chord.
-Chord analyzes Java bytecode, not Java source code, and thus only requires
-the program's class files.  Certain analyses, however, present
-their results at the Java source code level, and thus require the program's
-Java source files as well.
+The Java program to be analyzed is also specified via properties.  Chapter
+\ref{chap:setup} describes how to setup a Java program for analysis using Chord.
+Chord analyzes Java bytecode, not Java source code, and thus only requires the
+program's class files.  Certain analyses, however, present their results at the
+Java source code level, and thus require the program's Java source files as
+well.
 
-\subsubsection*{Program Representation}
+\subsubsection*{Java Program Representation}
 
-Chord uses the \xlink{Joeq}{http://joeq.sourceforce.net/} Java compiler framework
-to convert the Java bytecode of the input Java program, one class file at a time,
-into a three-address-like intermediate program representation called {\it quadcode}
-that is more suitable for analysis. Chapter \ref{chap:program-representation}
-describes the quadcode representation in detail.
+Chord uses the \xlink{Joeq}{http://joeq.sourceforce.net/} Java compiler
+framework to convert the Java bytecode of the input Java program, one class file
+at a time, into a three-address-like intermediate program representation called
+{\it quadcode} that is more suitable for analysis. Chapter
+\ref{chap:program} describes the quadcode representation in detail.
 
@@
 
-A pre-requisite to analyzing a Java program using any program analysis framework, including Chord, is
-to compute the {\it analysis scope}: which parts of the program to analyze.  Chord implements
-several standard scope construction algorithms from the literature that differ in aspects such as
-scalability, precision, and usability for the problem at hand.  Chapter \ref{chap:computing-scope}
+A pre-requisite to analyzing a Java program using any program analysis
+framework, including Chord, is to compute the {\it analysis scope}: which parts
+of the program to analyze.  Chord implements several standard scope construction
+algorithms from the literature that differ in aspects such as scalability,
+precision, and usability for the problem at hand.  Chapter \ref{chap:scope}
 describes these algorithms in detail.
@@
 
-A distinctive aspect of Chord is that each analysis is written modularly, independent of other analyses, along with lightweight
-annotations specifying the inputs and outputs of the analysis.
-Chord's runtime automatically computes
-dependencies between analyses (e.g., determines which analysis produces as output a
-result that is needed as input by another analysis).  Before running a desired analysis,
-Chord recursively runs other analyses until the inputs to the desired analysis have been computed; it
-finally runs the desired analysis to produce the outputs of that analysis.
+Chord provides many standard analyses.  Chapter \ref{chap:predefined} describes
+these analyses and how to run them.  Moreover, Chord allows users to define
+their own analyses, possibly atop the provided analyses.
 
-Chord can be invoked in one of two modes: classic or modern.
-These two modes defer in the semantics of dependencies between analyses.  In particular, the classic mode
-is simpler to understand for novice users (the dependencies are only data dependencies) but has a sequential
-runtime, whereas the modern mode is harder to understand (there are both data and control dependencies)
-but has a parallel runtime that is capable of running analyses without dependencies between
-them in parallel.  The parallel runtime is based on \xlink{Habanero-Java}{http://habanero.rice.edu/hj.html},
-and the semantics of the dependencies between analyses is based on the
-\xlink{Habanero Concurrent Collections (CnC)}{http://habanero.rice.edu/cnc.html} declarative parallel
-programming model.
+A distinctive aspect of Chord is that each analysis is written modularly,
+independent of other analyses, along with lightweight annotations specifying the
+inputs and outputs of the analysis.  Chord's runtime automatically computes
+dependencies between analyses (e.g., determines which analysis produces as
+output a result that is needed as input by another analysis).  Before running a
+desired analysis, Chord recursively runs other analyses until the inputs to the
+desired analysis have been computed; it finally runs the desired analysis to
+produce the outputs of that analysis.
 
-Chapter \ref{chap:project} describes how to write and run analyses in Chord.
+Chord can be invoked in one of two modes: {\it classic} or {\it modern}.  These
+two modes defer in the semantics of dependencies between analyses.  In
+particular, the classic mode is simpler to understand for novice users (the
+dependencies are only data dependencies) but has a sequential runtime, whereas
+the modern mode is harder to understand (there are both data and control
+dependencies) but has a parallel runtime that is capable of running analyses
+without dependencies between them in parallel.  The parallel runtime is based on
+\xlink{Habanero-Java}{http://habanero.rice.edu/hj.html}, and the semantics of
+the dependencies between analyses is based on the
+\xlink{Habanero Concurrent Collections (CnC)}{http://habanero.rice.edu/cnc.html}
+declarative parallel programming model.
 
-\subsubsection*{Analysis Templates}
+Chapter \ref{chap:project} expands upon the modular architecture of analyses in
+Chord that is described above.
 
-Chord provides various {\it analysis templates}: classes containing boilerplate code that can be
-extended by users to rapidly prototype different kinds of analyses.  An example is class RHSAnalysis,
-named after [Reps, Horowitz, and Sagiv 1995], which can be extended by users to write a summary-based inter-procedural
-context-sensitive static analysis by merely specifying the abstract domain and intra-procedural transfer functions.
-Another example is DynamicAnalysis, which can be extended by
-users to write a dynamic analysis by merely specifying which of various provided events to instrument, and
-the transfer functions for those events.
-Chapter \ref{chap:analysis-templates} describes the various analysis templates.
+Chord provides various {\it analysis templates}: classes containing boilerplate
+code that can be extended by users to rapidly prototype different kinds of
+analyses.  An example is class RHSAnalysis, named after [Reps, Horowitz, and
+Sagiv 1995], which can be extended by users to write a summary-based
+inter-procedural context-sensitive static analysis by merely specifying the
+abstract domain and intra-procedural transfer functions.  Another example is
+DynamicAnalysis, which can be extended by users to write a dynamic analysis by
+merely specifying which of various provided events to instrument, and the
+transfer functions for those events.
+
+Chapters \ref{chap:writing} and \ref{chap:running} describe how to write and run
+your own analyses in Chord using the provided analysis templates.
 
@@
 
-Chord uses the \xlink{Javassist}{http://www.javassist.org/} Java bytecode manipulation
-framework for instrumenting bytecode and doing dynamic analysis.
-Chord offers the most versatile capabilities of any existing dynamic analysis framework
-for Java, particularly the ability to instrument the entire JDK (including classes in
-package java.lang).
-Specifically, it includes support for:
+Chord uses the \xlink{Javassist}{http://www.javassist.org/} Java bytecode
+manipulation framework for instrumenting bytecode and doing dynamic analysis.
+Chord offers the most versatile capabilities of any existing dynamic analysis
+framework for Java, particularly the ability to instrument the entire JDK
+(including classes in package java.lang).  Specifically, it includes support
+for:
+
 \begin{itemize}
@@
 \item
-processing of dynamic analysis events online in the same JVM or offline in a different JVM with an uninstrumented JDK (the latter circumvents performance and correctness problems that can arise if a single JVM with an instrumented JDK is used to generate and handle events); and
+processing of dynamic analysis events online in the same JVM or offline in a
+different JVM with an uninstrumented JDK (the latter circumvents performance and
+correctness problems that can arise if a single JVM with an instrumented JDK is
+used to generate and handle events); and
 \item
-allowing the event-generating and event-handling JVMs to run either serially (by storing the entire trace of events to a regular file) or in parallel (by streaming the trace of events in a piped file).
+allowing the event-generating and event-handling JVMs to run either serially
+(by storing the entire trace of events to a regular file) or in parallel (by
+streaming the trace of events in a piped file).
 \end{itemize}
-Chapter \ref{chap:dynamic-analysis} describes all aspects of dynamic analysis in Chord.
+
+Chapter \ref{chap:dynamic} describes all aspects of dynamic analysis in Chord.
 
@@
 
-An analysis in Chord can be written either imperatively in Java or declaratively in Datalog.
-Chord uses the BDD-based Datalog solver \xlink{bddbddb}{http://bddbddb.sourceforge.net/} to
-run analyes written in Datalog.
-Chapter \ref{chap:datalog-analysis} describes all aspects of Datalog analysis.
-
-%If the analyses are purely static, it suffices to specify the main class of the program
-%(via property \code{chord.main.class}) and the application classpath of the program (via property \code{chord.class.path}).
-%If the analyses need to present their results at the Java source level, then
-%the path of the program's Java source files is also required (via property \code{chord.src.path}).
-%Finally, if any analysis is dynamic, then the command-line arguments to be used for running the
-%program must also be specified (via property \code{chord.run.ids} and properties \code{chord.args.<N>} for
-%each item \code{<N>} in the value of property \code{chord.run.ids}).
-
-\subsubsection*{Predefined Analyses}
-
-Chapter \ref{chap:predefined-analyses} describes many standard analyses that are implemented in Chord.
+A common way to rapidly prototype an analysis in Chord is using a declarative
+logic-programming language called Datalog.  Chord uses the BDD-based Datalog
+solver \xlink{bddbddb}{http://bddbddb.sourceforge.net/} to run analyes written
+in Datalog.  Chapter \ref{chap:datalog} describes all aspects of such analyses.

### 2011-04-15 main/doc/datalog.tex
more documentation

--- 2011-04-12 main/doc/datalog_analysis.tex
@@
-\xname{datalog_analysis}
+\xname{datalog}
 \chapter{Datalog Analysis}
-\label{chap:datalog-analysis}
+\label{chap:datalog}
 
-A common way to rapidly prototype analyses in Chord is using a declarative
-logic-programming language called Datalog.  This chapter describes all aspects of
-Datalog analyses in Chord.
-Section \ref{sec:writing-datalog-analysis} explains how to write and run a Datalog analysis.
-Section \ref{sec:tuning-datalog-analysis} explains how to tune its performance.
-Finally, Section \ref{sec:bdd} explains the representation of BDDs (Binary Decision Diagrams)
-which are a data structure central to executing Datalog analyses.
+A common way to rapidly prototype an analysis in Chord is using a declarative
+logic-programming language called Datalog.  This chapter describes all aspects
+of Datalog analyses in Chord.  Section \ref{sec:writing-datalog} explains how to
+write and run a Datalog analysis.  Section \ref{sec:tuning-datalog} explains how
+to tune its performance.  Finally, Section \ref{sec:bdd} explains the
+representation of BDDs (Binary Decision Diagrams) which are a data structure
+central to executing Datalog analyses.
 
 \section{Writing a Datalog Analysis}
-\label{sec:writing-datalog-analysis}
+\label{sec:writing-datalog}
 
-A Datalog analysis declares a bunch of input/output program relations,
-each over one or more program domains, and provides a bunch of rules (constraints)
-specifying how to compute the output relations from the input relations.
-It can be defined in any file with extension {\tt .dlog}
-in any directory in the path specified by property \code{chord.dlog.analysis.path}.
-An example Datalog analysis is shown below:
+A Datalog analysis declares a bunch of input/output program relations, each over
+one or more program domains, and provides a bunch of rules (constraints)
+specifying how to compute the output relations from the input relations.  It can
+be defined in any file with suffix {\tt .dlog} in any directory in the path
+specified by property \code{chord.dlog.analysis.path}.  An example Datalog
+analysis is shown below:
 
@@
 
-The ``\code{.include "<DOM_NAME>.dom"}'' lines specify each program domain named
+The ``\code{.include "<DOM_NAME>.dom"}'' lines specify each domain named
 \code{<DOM_NAME>} that is needed by the Datalog analysis, i.e., each domain over which
-any program relation that is input/output by the Datalog analysis is defined. 
-The declaration of each such relation specifies the domain of each of the relation's attributes.
-If the same domain appears in multiple attributes of a relation then
+any relation that is input/output by the Datalog analysis is defined. 
+The declaration of each such relation specifies the relation's schema:
+all the domains over which the relation is defined.
+If the same domain appears multiple times in a relation's schema then
 contiguous integers starting from 0 must be used to distinguish them; for instance,
@@
 long rules into multiple shorter rules communicating via intermediate relations.
-See Section \ref{sec:tuning-datalog-analysis} for hints on tuning the performance
+See Section \ref{sec:tuning-datalog} for hints on tuning the performance
 of Datalog analyses.
@@
 Running a Datalog analysis is no different from running any other kind of analysis
-in Chord.  See Chapter \ref{chap:running-analysis} for how to run an analysis.
+in Chord.  See Chapter \ref{chap:running} for how to run an analysis.
 
 \section{Tuning Performance}
-\label{sec:tuning-datalog-analysis}
+\label{sec:tuning-datalog}
 

### 2011-04-15 main/doc/program.tex
more documentation

--- 2011-04-04 main/doc/program_representation.tex
@@
-\xname{program_representation}
-\chapter{Program Representation}
-\label{chap:program-representation}
+\xname{program}
+\chapter{Java Program Representation}
+\label{chap:program}
 
@@
 consists of a set of classes that may be loaded
-(Section \ref{sec:program}).  Each class has a set of fields and
-methods, collectively called members of the class
-(Section \ref{sec:class-members}).  The representation of a concrete
+(Section \ref{sec:program}).  The representation of each class consists
+of a set of members (Section \ref{sec:class-members}) which are the fields and methods
+of the class.  The representation of a concrete
 method (Section \ref{sec:methods}) consists of a control-flow graph
 (CFG).  The representation of a CFG (Section \ref{sec:cfgs}) consists
-of a set of registers and a set of basic blocks.  Each basic
-block contains zero or more primitive statements called quads
+of a set of registers and a set of basic blocks linked by directed edges
+denoting flow of control between basic blocks.
+Each basic block contains zero or more primitive statements called quads
 (Section \ref{sec:quads}).  Finally, the most common way to traverse
@@
 \begin{verbatim}
-javap -classpath <...> -private -verbose <CLASS NAME>
+javap -classpath <CLASS_PATH> -bootclasspath <BOOT_CLASS_PATH> -private -verbose <CLASS_NAME>
 \end{verbatim}
@@
 \item
-{\tt <CLASS NAME>} is the fully-qualified name of the class whose
+\verb+<CLASS_NAME>+ is the fully-qualified name of the class whose
 bytecode is to be printed (\code{test.HelloWorld} in our example).
 \item
-\code{<...>} is the classpath of that class (\code{examples/hello_world/} in our example).
+\verb+<CLASS_PATH>+ is the classpath of that class (\code{examples/hello_world/} in our example).
+\item
+\verb+<BOOT_CLASS_PATH>+ is the boot classpath; it is optional and must be supplied if
+\verb+<CLASS_NAME>+ is a class from the JDK standard library (e.g., \code{java.util.ArrayList}) that has
+been modified and written to a user-defined location (e.g., it has been instrumented by Chord and
+written to \code{chord_output/boot_classes/}).
 \end{itemize}
@@
 \begin{verbatim}
-ant -Dchord.work.dir=<...> -Dchord.print.classes=<CLASS NAME> \
-    -Dchord.verbose=0 -Dchord.out.file=<FILE NAME> run
+ant -Dchord.work.dir=<WORK_DIR> -Dchord.print.classes=<CLASS_NAME> \
+    -Dchord.verbose=0 -Dchord.out.file=<OUT_FILE> run
 \end{verbatim}
@@
 \item
-\code{<...>} is the path of a directory (\code{examples/hello_world/} in our example) that is expected to
-contain a file named \code{chord.properties} which defines
+\code{<WORK_DIR>} is a directory (\code{examples/hello_world/} in our example) 
+that contains a file named \code{chord.properties} which defines
 properties \code{chord.main.class} and \code{chord.class.path}
-specifying the main class and the classpath of the input Java program
-(alternatively, these two properties can be defined directly on the
-above command line).
+specifying the main class and the classpath of the input Java program.
+Alternatively, these two properties can be defined directly on the
+above command-line.
 \item
-{\tt <CLASS NAME>} is the name of the class whose quadcode is to be
-printed.
+\code{<CLASS_NAME>} is the fully-qualified name of the class whose quadcode is to be printed
+({\tt test.HelloWorld} in our example).  Each occurrence of a `{\tt \$}' in the class name
+must be replaced by a `{\tt \#}'.
 \item
-{\tt <FILE NAME>} is the name of a file to which the quadcode is to be
-dumped.
+\code{<OUT_FILE>} is the file to which the quadcode must be written; if left unspecified,
+the quadcode is written to the standard output.
 \end{itemize}
@@
 This API is contained in
-packages \javadoc{chord.program}{chord/program/package-tree.html}, \javadoc{joeq.Class}{joeq/Class/package-tree.html},
-and \javadoc{joeq.Compiler.Quad}{joeq/Compiler/Quad/package-tree.html}.
+packages \javadoc{chord.program}{chord/program/package-summary.html}, \javadoc{joeq.Class}{joeq/Class/package-summary.html},
+and \javadoc{joeq.Compiler.Quad}{joeq/Compiler/Quad/package-summary.html}.
 
@@
 
-See Chapter \ref{chap:computing-scope} for how Chord determines which
+See Chapter \ref{chap:scope} for how Chord determines which
 classes may be loaded and which methods may be called.
@@
 
-\begin{framed}
 \begin{verbatim}
-joeq.Class.jq_Type
+                    jq_Type
                        |
-  |--- joeq.Class.jq_Primitive
+      ------------------------------------
+      |                                  |
+jq_Primitive                       jq_Reference
                                          |
-  |--- joeq.Class.jq_Reference
-            |
-            |--- joeq.Class.jq_Class
-            |
-            |--- joeq.Class.jq_Array
+                          -------------------------------
+                          |                             |
+                      jq_Class                       jq_Array
 \end{verbatim}
-\end{framed}
 
@@
 
-\begin{framed}
 \begin{verbatim}
@@
 \end{verbatim}
-\end{framed}
 
-The characters of \code{BaseType}, the {\tt L} and {\tt ;}
-of \code{ObjectType}, and the {\tt [} of
+The characters of \code{BaseType}, the `{\tt L}' and `{\tt ;}'
+of \code{ObjectType}, and the `{\tt [}' of
 \code{ArrayType} are all ASCII characters.
@@
 \end{verbatim}
-\end{framed}
 
@@
 A return descriptor represents the type of the return result of a
-method.  The character \code{V} indicates that the method returns no
+method.  The character `{\tt V}' indicates that the method returns no
 value (its return type is void).
@@
 The control-flow graph (CFG) of each method consists of a set of registers,
-called the register factory, and a graph whose nodes are basic blocks and
+called the register factory, and a directed graph whose nodes are basic blocks and
 whose edges denote flow of control between basic blocks.
@@
 
-\subsection{Register Factory}
+%\subsection{Register Factory}
 
@@
 
-\subsection{Basic Blocks}
-
-Each CFG has a set of basic blocks each of which contains a
-straight-line sequence of zero or more primitive statements called quads
+Besides the register factory, a CFG has a directed graph whose nodes are basic blocks and whose edges
+denote flow of control between basic blocks.
+Each basic block contains a straight-line sequence of zero or more primitive statements called quads
 (Section \ref{sec:quads}).  Each CFG is guaranteed to contain at least
@@
 
-\begin{framed}
 \begin{verbatim}
 Operator
-    |    Move, Phi, Unary, Binary, New, NewArray, MultiNewArray, 
-    |--- Getstatic, Putstatic, Getfield, Putfield, ALoad, AStore,
-    |    CheckCast, InstanceOf, ALength, Return
+    |
+    |--- Move
+    |--- Phi
+    |--- Unary
+    |--- Binary
+    |--- New
+    |--- NewArray
+    |--- MultiNewArray
+    |--- Getstatic
+    |--- Putstatic
+    |--- ALoad
+    |--- AStore
+    |--- Getfield
+    |--- Putfield
+    |--- CheckCast
+    |--- InstanceOf
+    |--- ALength
+    |--- Return
     |
     |--- Branch
-    |       |--- IntIfCmp, Goto, Jsr, Ret, LookupSwitch, TableSwitch
+    |       |
+    |       |--- IntIfCmp
+    |       |--- Goto
+    |       |--- Jsr
+    |       |--- Ret
+    |       |--- LookupSwitch
+    |       |--- TableSwitch
     |
     |--- Invoke
-    |       |--- InvokeVirtual, InvokeStatic, InvokeInterface
+    |       |
+    |       |--- InvokeVirtual
+    |       |--- InvokeStatic
+    |       |--- InvokeInterface
     |
     |--- Monitor
-            |--- MONITORENTER, MONITOREXIT
+            |
+            |--- MONITORENTER
+            |--- MONITOREXIT
 \end{verbatim}
-\end{framed}
 

### 2011-04-15 main/doc/properties.tex
more documentation

@@
 name of that program's main class).
-Chapter \ref{chap:program-setup} presents an example properties file 
+Chapter \ref{chap:setup} presents an example properties file 
 that defines the program properties that are most commonly used.  Section \ref{sec:program-props}
@@
 This section describes properties that specify how the analysis scope of the input Java program is computed.
-See Chapter \ref{chap:computing-scope} for more details.
+See Chapter \ref{chap:scope} for more details.
 
@@
  a large number (e.g., millions) of tuples, resulting in voluminous output when printed in explicit form to a text file.  See Section
- \ref{sec:tuning-datalog-analysis} for a more efficient way to query the contents of program relations (namely, by using the {\tt debug}
+ \ref{sec:tuning-datalog} for a more efficient way to query the contents of program relations (namely, by using the {\tt debug}
  target provided in file {\tt build.xml} in Chord's main directory).
@@
 {\bf Type:} bool \\
-{\bf Description:} Whether to use the classic project (as opposed to the modern project).  See Chapter\ref{chap:analyses} for the difference
+{\bf Description:} Whether to use the classic project (as opposed to the modern
+project).  See Chapter\ref{chap:project} for the difference
  between the two kinds of projects. \\
@@
 %{\tt true}
-%{\bf Note:} This functionality is useful for debugging Datalog programs using the {\tt debug} target provided in file {\tt build.xml} in Chord's main directory (see Section \ref{sec:tuning-datalog-analysis}).
+%{\bf Note:} This functionality is useful for debugging Datalog programs using
+%the {\tt debug} target provided in file {\tt build.xml} in Chord's main
+%directory (see Section \ref{sec:tuning-datalog}).

### 2011-04-15 main/doc/scope.tex
more documentation

--- 2011-03-27 main/doc/computing_scope.tex
@@
-\xname{computing_scope}
-\chapter{Computing Analysis Scope}
-\label{chap:computing-scope}
+\xname{scope}
+\chapter{Analysis Scope Construction}
+\label{chap:scope}
 
 A pre-requisite to analyzing a Java program using any program analysis framework,
-including Chord, is to compute the {\it analysis scope}: the set of methods
-of the program to be analyzed.  Several scope construction algorithms (so-called
-call-graph algorithms) exist in the literature that differ in scalability (i.e., how large
-a program they can handle with the available resources) and precision (i.e., how small a
-set of methods they deem is reachable).
+including Chord, is to compute the {\it analysis scope}: which parts of the
+program to analyze.  Several scope construction algorithms (so-called
+call-graph algorithms) exist in the literature that differ in scalability (i.e.,
+how large a program they can handle with the available resources) and precision (i.e., 
+how much of the program they deem is reachable).
 
@@
 an additional metric of these algorithms in Chord that can be controlled by users is usability,
-which concerns aspects such as excluding certain methods from being analyzed even if they
-are reachable, and modeling Java features such as reflection,
-dynamic class loading, and native methods.
-These features affect which methods are
+which concerns aspects such as excluding certain code from being analyzed even if it
+is reachable, and modeling Java features such as reflection, dynamic class loading, and native methods.
+These features affect which code are
 reachable but, in general, they cannot be modeled soundly by any program analysis framework.  The best
@@
 The possible values of property \code{chord.scope.kind} are
-[\code{dynamic}$|$\code{rta}$|$\code{cha}] (the default value is {\tt rta}).
+[\code{rta}$|$\code{cha}$|$\code{dynamic}] (the default value is {\tt rta}).
 The following subsections describe the scope construction algorithm
@@
 
-\subsection{Dynamic Scope Construction}
-
-If property \code{chord.scope.kind} has value {\tt dynamic}, then Chord
-computes analysis scope
-dynamically, by running the program and observing the
-classes that are loaded at run-time.  The number of times the program
-is run and the command-line arguments to be supplied to the program in
-each run is specified by properties \code{chord.run.ids} and
-\code{chord.args.<id>} for each run ID {\tt <id>}.  By default, the
-program is run only once, using run ID {\tt 0}, and without any
-command-line arguments.  Only classes loaded in some run are regarded
-as reachable but {\it all} methods of each loaded class are regarded
-as reachable regardless of whether they were invoked in the run.  The
-rationale behind this decision is to both reduce the run-time instrumentation
-overhead and increase the predictive power of program analyses
-performed using the computed analysis scope.
-
-\subsection{RTA Scope Construction}
+\subsection{Rapid Type Analysis}
 
@@
 
-\subsection{CHA Scope Construction}
+\subsection{Class Hierarchy Analysis}
 
@@
 
+\subsection{Dynamic Analysis}
+
+If property \code{chord.scope.kind} has value {\tt dynamic}, then Chord
+computes analysis scope
+dynamically, by running the program and observing the
+classes that are loaded at run-time.  The number of times the program
+is run and the command-line arguments to be supplied to the program in
+each run is specified by properties \code{chord.run.ids} and
+\code{chord.args.<id>} for each run ID {\tt <id>}.  By default, the
+program is run only once, using run ID {\tt 0}, and without any
+command-line arguments.  Only classes loaded in some run are regarded
+as reachable but {\it all} methods of each loaded class are regarded
+as reachable regardless of whether they were invoked in the run.  The
+rationale behind this decision is to both reduce the run-time instrumentation
+overhead and increase the predictive power of program analyses
+performed using the computed analysis scope.
+
 \section{Scope Exclusion}
@@
 {\bf Note:} The value of each of the above properties is a list of {\it prefixes},
-not {\it regular expressions}.  A valid value is \code{java.,com.sun.},
-but not \code{java.*,com.sun.*}.
+not {\it regular expressions}.  A valid value is ``\code{java.,com.sun.}",
+but not ``\code{java.*,com.sun.*}".
 

### 2011-04-15 main/doc/writing.tex
more documentation

\xname{writing}
\chapter{Writing an Analysis}
\label{chap:writing}

Chord provides several {\it analysis templates}: classes containing boilerplate
code that can be extended by users to rapidly prototype different kinds of
analyses.  These classes are organized in the following hierarchy in package
\javadoc{chord.project.analyses}{chord/project/analyses/package-summary.html}:

\begin{verbatim}
JavaAnalysis
    |
    |--- ProgramDom
    |
    |--- ProgramRel
    |
    |--- DlogAnalysis
    |
    |--- RHSAnalysis
    |        |
    |        |--- ForwardRHSAnalysis
    |        |
    |        |--- BackwardRHSAnalysis
    |
    |--- BasicDynamicAnalysis
             |
             |--- DynamicAnalysis
\end{verbatim}

The following sections describe each of these analysis templates in more detail.

\section{JavaAnalysis}
\label{sec:java}

Class \javadoc{chord.project.analyses.JavaAnalysis}{chord/project/analyses/JavaAnalysis.html}
is the most general template for writing an analysis.  An analysis can be
created using this template by extending this class as follows:

\begin{framed}
\begin{verbatim}
import chord.project.Chord;
import chord.project.ClassicProject;
import chord.project.analyses.JavaAnalysis;
import chord.program.Program;

@Chord(
    name = "<ANALYSIS_NAME>",
    consumes = { "C1", ..., "Cn" },
    produces = { "P1", ..., "Pm" },
    namesOfTypes = { "C1", ..., "Cn", "P1", ..., "Pm" },
    types = { A1.class, ..., An.class, B1.class, ..., Bm.class }
)
public class ExampleAnalysis extends JavaAnalysis {
    @Override public void run() {
        Program program = Program.g();
        ClassicProject project = ClassicProject.g();
        A1 c1 = (A1) project.getTrgt("C1");
        ...
        An cn = (An) project.getTrgt("Cn");
        B1 p1 = (B1) project.getTrgt("P1");
        ...
        Bm pm = (Bm) project.getTrgt("Pm");
        // compute produced targets p1, ..., pm from program and
        // consumed targets c1, ..., cn
        ...
    }
}
\end{verbatim}
\end{framed}

To run the analysis, class \code{ExampleAnalysis} must be compiled to a
{\tt .class} file that occurs in some element (directory or jar/zip file) of
the path specified by property \code{chord.java.analysis.path}.  This causes
the analysis to be included in a Chord project as a task that is represented
as a separate object of class \code{ExampleAnalysis}.

The {\tt @Chord} annotation, defined in class
\javadoc{chord.project.Chord}{chord/project/Chord.html}, specifies via fields
the following aspects of the analysis:
\begin{itemize}
\item
Field {\tt name} specifies the name of the analysis (\code{<ANALYSIS_NAME>}).
\item
Field {\tt consumes} specifies the names of targets that are consumed by the
analysis ({\tt C1}, ..., {\tt Cn}).
\item
Field {\tt produces} specifies the names of targets that are produced by the
analysis ({\tt P1}, ..., {\tt Pm}).
\item
Fields {\tt namesOfTypes} and {\tt types} specify the types of targets.  There
is a 1-to-1 correspondence between the arrays denoted by these two fields,
e.g., the type of the target named {\tt C1} is class {\tt A1}, and so on.  In
principle, the type of {\it any} target in the project can be specified here.
In practice, however, the types of only the targets declared as
consumed/produced by this analysis are specified.  Moreover, although the type
of each target that is consumed/produced by the above analysis is specified in
the above annotation, in practice the types of hardly any targets need to be
explicitly specified, because they can be automatically inferred by Chord from
analyses created using more specialized templates discussed below that also
consume/produce those targets.
\end{itemize}

The code of the analysis must be supplied in the {\tt run()} method.  This
method typically does the following in order:
(1) retrieves the program being analyzed and the representation of each
consumed/produced target from the project;
(2) performs some computation that uses the program and the consumed targets
as inputs; and
(3) writes the outputs of the computation to the produced targets.

The analysis templates presented in the following sections are more specialized
forms of the {\tt JavaAnalysis} template: they constrain the number and kinds of
consumed/produced targets and/or the analysis code in the {\tt run()} method.

%called JavaAnalysis (Section \ref{sec:java}), and the other for writing
%analyses declaratively in Datalog, called DlogAnalysis
%(Section \ref{sec:dlog}).  JavaAnalysis is the most general template:
%it least constrains the kind of analysis that can be written but also offers the
%least boilerplate code for writing an analysis.  Chord provides several
%additional templates for writing analyses imperatively in Java that are
%specialized forms of JavaAnalysis, namely, ProgramDom
%(Section \ref{sec:program-dom}), ProgramRel (Section \ref{sec:program-rel}),
%RHSAnalysis (Section \ref{sec:rhs}), and DynamicAnalysis
%(Section \ref{sec:dynamic}).  We next explain each of these templates.

\section{ProgramDom}
\label{sec:program-dom}

Class \javadoc{chord.project.analyses.ProgramDom}{chord/project/analyses/ProgramDom.html}
is a template for writing a {\it program domain analysis}.
A {\it program domain} represents an indexed set of values of a fixed
kind, typically from the program being analyzed, such as the set of all methods
in the program, the set of all fields in the program, etc.  Indices are 
assigned starting from 0 and in the order in which values are added to the set.
A program domain primarily serves as an input to Datalog analyses
(see Section \ref{sec:dlog}).  Thus, it is a kind of target (i.e.,
analysis result) in a Chord project.  A common way to define a program domain
is to create a program domain analysis by extending class {\tt ProgramDom} as
follows:

\begin{framed}
\begin{verbatim}
import chord.project.Chord;
import chord.project.ClassicProject;
import chord.project.analyses.ProgramDom;
import chord.program.Program;

@Chord(
    name = "<DOM_NAME>",
    consumes = { "C1", ..., "Cn" }
)
public class ExampleDom extends ProgramDom<DOM_TYPE> {
    @Override public void fill() {
        Program p = Program.g();
        ClassicProject project = ClassicProject.g();
        A1 c1 = (A1) project.getTrgt("C1");
        ...
        An cn = (An) project.getTrgt("Cn");
        // populate domain using program and consumed targets c1, ..., cn
        for (...) {
            DOM_TYPE e = ...;
            add(e); 
        }
    }
}
\end{verbatim}
\end{framed}

To run the analysis, class \code{ExampleDom} must be compiled to a {\tt .class}
file that occurs in some element (directory or jar/zip file) of the path
specified by property \code{chord.java.analysis.path}.  This causes the
analysis to be included in a Chord project as a task that is represented as a
separate object of class \code{ExampleDom}.  Moreover, that object also denotes
a target in the Chord project.  Both the task and target have the same name
\code{<DOM_NAME>}.

The \code{ProgramDom} template can be viewed as providing the following
specialized form of the general \code{JavaAnalysis} template:

\begin{itemize}
\item
It consumes any number and kinds of targets ({\tt C1}, ..., {\tt Cn}).
\item
It produces a single target, namely, the defined program domain itself
(named \code{<DOM_NAME>} and of type \code{ExampleDom}).
\item
Its \code{run()} method adds values to the defined program domain.  Typically,
it suffices to override the {\tt fill()} method (which is called by the
{\tt run()} method) and call from it the {\tt add(e)} method for each value
{\tt e} to be added to the domain in order.
\end{itemize}

It is a runtime error to explicitly specify any produced targets in the
{\tt @Chord} annotation of a class extending {\tt ProgramDom}.  If you wish to
define an analysis that produces additional targets besides a program domain,
then you can still define the program domain in a class such as
{\tt ExampleDom} that extends {\tt ProgramDom}, but you must not annotate it
with the {\tt @Chord} annotation (since this annotation causes the class to be 
egarded as defining an analysis).  Instead, define the analysis in a separate
class that extends {\tt JavaAnalysis}, as follows:

\begin{framed}
\begin{verbatim}
import chord.project.Chord;
import chord.project.ClassicProject;
import chord.project.analyses.JavaAnalysis;

@Chord(
    name = "<ANALYSIS_NAME>",
    consumes = { "C1", ..., "Cn" },
    produces = { "<DOM_NAME>", ... }
)
public class ExampleAnalysis extends JavaAnalysis {
    @Override public void run() {
        ExampleDom d = (ExampleDom) ClassicProject.g().getTrgt("<DOM_NAME>");
        d.run();  // produce domain named <DOM_NAME>
        ...
    }
}
\end{verbatim}
\end{framed}

Note that targets {\tt C1}, ..., {\tt Cn} that were declared as consumed in
the {\tt @Chord} annotation of class {\tt ExampleDom} (and any other fields
such as {\tt namesOfTypes} and {\tt types}) must now be provided in the
{\tt @Chord} annotation of class \code{ExampleAnalysis}.

\section{ProgramRel}
\label{sec:program-rel}

Class \javadoc{chord.project.analyses.ProgramRel}{chord/project/analyses/ProgramRel.html}
is a template for writing a {\it program relation analysis}.
A {\it program relation} represents a set of tuples over one or more fixed 
program domains.  A program relation primarily serves as an input or output of
Datalog analyses (see Section \ref{sec:dlog}).  Thus, it is a kind of
target (i.e., analysis result) in a Chord project.  A common way to define a
program relation is to create a program domain analysis by extending class
{\tt ProgramRel} as follows:

\begin{framed}
\begin{verbatim}
import chord.project.Chord;
import chord.project.ClassicProject;
import chord.project.analyses.ProgramDom;
import chord.project.analyses.ProgramRel;
import chord.program.Program;

@Chord(
    name = "<REL_NAME>",
    consumes = { "C1", ..., "Cn" },
    sign = "<DOM_NAMES>:<DOM_ORDER>"
)
public class ExampleRel extends ProgramRel {
    @Override public void fill() {
        Program p = Program.g();
        ProgramDom<T1> d1 = doms[0];
        ...
        ProgramDom<Tm> dm = doms[m-1];
        ClassicProject project = ClassicProject.g();
        A1 c1 = (A1) project.getTrgt("C1");
        ...
        An cn = (An) project.getTrgt("Cn");
        // populate relation using program, its domains d1, ..., dm, and
        // consumed targets c1, ..., cn
        for (...) {
            T1 o1 = ...;
            Tm om = ...;
            add(o1, ..., om); 
        }
    }
}
\end{verbatim}
\end{framed}

To run the analysis, class \code{ExampleRel} must be compiled to a {\tt .class}
file that occurs in some element (directory or jar/zip file) of the path
specified by property \code{chord.java.analysis.path}.  This causes the
analysis to be included in a Chord project as a task that is represented as a
separate object of class \code{ExampleRel}.  Moreover, that object also denotes
a target in the Chord project.  Both the task and target have the same name
\code{<REL_NAME>}.

The \code{ProgramRel} template can be viewed as providing the following
specialized form of the general \code{JavaAnalysis} template:

\begin{itemize}
\item
It consumes any number and kinds of targets ({\tt C1}, ..., {\tt Cn}).
\item
It produces a single target, namely, the defined program relation itself
(named \code{<REL_NAME>} and of type \code{ExampleRel}).
\item
Its \code{run()} method adds tuples to the defined program relation.  Typically,
it suffices to override the {\tt fill()} method (which is called by the
{\tt run()} method) and call from it the {\tt add(o1, ..., om)} method
for each tuple ({\tt o1}, ..., {\tt om}) to be added to the relation.
\end{itemize}

Unlike for program domains, the order in which tuples are added to a program
relation is irrelevant.  But the relative ordering of the program domains over
which the program relation is declared matters heavily for performance.
This is because each program relation in Chord is represented symbolically (as
oppoosed to explicitly) using a data structure called a Binary Decision Diagram
(BDD for short).  This in turn is because in pratice, a program relation (e.g.,
one representing context-sensitive points-to information)
can contain millions or billions of tuples even for a moderately-sized
input Java program; representing such a large number of tuples explicitly is
prohibitively and needlessly expensive.  The size of a BDD, on the other hand,
does not depend at all upon the number of tuples in the program relation that
the BDD represents.  Instead, it depends heavily upon the relative ordering
of the program domains over which the program relation is declared.
Hence, the {\tt @Chord} annotation on a class such as {\tt ExampleRel} that
extends \code{ProgramRel} is required to have a {\tt sign} field whose value is
the {\it sign} of the program relation.
A sign is a string of the form \code{<DOM_NAMES>:<DOM_ORDER>} where:

\begin{itemize}

\item

\code{<DOM_NAMES>} is mandatory and specifies the relation's {\it schema}: a comma-separated list of names of
the domains over which the relation is defined, with each domain name suffixed with a
non-negative integer that is typically 0 and must be unique across multiple
occurrences of the same domain name.
The order of domain names in the schema is {\it irrelevant}: users must
pick this order purely based on what order they find most convenient to remember.

For example, suppose the program relation represents the result of Class Hierarchy
Analysis (CHA), i.e., it contains each tuple of the form ($m_1$,$t$,$m_2$) such that
method $m_2$ is the resolved method of an invokevirtual or invokeinterface call
with resolved method $m_1$ on an object of class $t$.  Let {\tt M} and {\tt T}
denote the names of the program domains representing the set of all methods
and the set of all classes, respectively, in the Java program being analyzed.
Then, \code{<DOM_NAMES>} for this program relation could be any of ``{\tt M0,T0,M1}'',
``{\tt M1,T0,M0}'', ``{\tt M0,T1,M1}'', and so on.

\item

\code{<DOM_ORDER>} is optional and determines the relation's representation as a
BDD. It is a permutation of the names in \code{DOM_NAMES>} with a separator
`\_' or `x' between consecutive names.  The order of names in this
list, and the kind of separators used between them, are what determines both
the size of the BDD and the performance of operations on it (such as join,
selection, projection, etc.).

An example value of \code{<DOM_ORDER>} for the
CHA program relation above with \code{<DOM_LIST>} as ``\code{M0,T0,M1}''  is
``\code{M0xM1_T0}''; another example value is ``\code{M1_T0_M0}''.
\end{itemize}

Chapter \ref{chap:datalog} describes how BDDs are represented
(Section \ref{sec:bdd}) and how you can tune their size and the
performance of operations on them (Section \ref{sec:tuning-datalog}).

It is a runtime error to explicitly specify any produced targets in the
{\tt @Chord} annotation of a class extending {\tt ProgramRel}.  If you wish to
define an analysis that produces additional targets besides a program relation,
then you can still define the program relation in a class such as
{\tt ExampleRel} that extends {\tt ProgramRel}, but you must not annotate it
with the {\tt @Chord} annotation (since this annotation causes the class to be 
egarded as defining an analysis).  Instead, define the analysis in a separate
class that extends {\tt JavaAnalysis}, as follows:

\begin{framed}
\begin{verbatim}
import chord.project.Chord;
import chord.project.ClassicProject;
import chord.project.analyses.JavaAnalysis;

@Chord(
    name = "<ANALYSIS_NAME>",
    consumes = { "C1", ..., "Cn" },
    produces = { "<REL_NAME>", ... },
    namesOfSigns = { "<REL_NAME>", ... },
    signs = { "<DOM_ORDER>:<DOM_NAME>", ... }
)
public class ExampleAnalysis extends JavaAnalysis {
    @Override public void run() {
        ExampleRel r = (ExampleRel) ClassicProject.g().getTrgt("<REL_NAME>");
        r.run();  // produce program relation named <REL_NAME>
    }
}
\end{verbatim}
\end{framed}

Note that targets {\tt C1}, ..., {\tt Cn} that were declared as consumed in
the {\tt @Chord} annotation of class {\tt ExampleRel} (and any other fields
such as {\tt namesOfTypes} and {\tt types}) must now be provided in the
{\tt @Chord} annotation of class \code{ExampleAnalysis}.  Just like the
1-to-1 correspondence between the values of fields {\tt namesOfTypes} and
{\tt types}, there is a 1-to-1 correspondence between the values of fields
{\tt namesOfSigns} and {\tt signs}, which allow relating the name of any program
relation in the project with its sign.

\section{DlogAnalysis}
\label{sec:dlog}

A common way to rapidly prototype analyses in Chord is using a declarative
logic-programming language called Datalog.  A Datalog analysis is defined in a
{\tt .dlog} file that primarily specifies the following:
\begin{itemize}
\item
A set of {\it program domains} {\tt D1}, ..., {\tt Dk}.  A program domain is a
set of values of a fixed kind.  Each program domain in a Chord project is
a target that is represented as a separate object of class \code{ProgramDom}
(see Section \ref{sec:program-dom}).
\item
A set of {\it program relations}, including input relations {\tt I1}, ..., {\tt In}
and output relations {\tt O1}, ..., {\tt Om}.  A program relation is a set of
tuples over one or more fixed program domains {\tt D1}, ..., {\tt Dk}.  Each program
relation in a Chord project is a target that is represented as a separate object
of class \code{ProgramRel} (see Section \ref{sec:program-rel}).
\item
A set of rules (constraints) {\tt R} specifying how to compute the output
relations from the input relations.
\end{itemize}
An example such file is shown in Chapter \ref{chap:datalog} which also
explains all aspects of Datalog analyses.

To run the analysis, the {\tt .dlog} file must occur in some element (directory
or jar/zip file) of the path specified by property \code{chord.dlog.analysis.path}.
 This causes the analysis to be included in a Chord project as a task that is
represented as a separate object of class
\javadoc{chord.project.analyses.DlogAnalysis}{chord/project/analyses/DlogAnalysis.html}.
Note that, unlike class {\tt JavaAnalysis}, users must not extend class
{\tt DlogAnalysis} but must instead define the analysis in a {\tt .dlog} file.

The \code{DlogAnalysis} template can be viewed as providing the following
specialized form of the general \code{JavaAnalysis} template:
\begin{itemize}
\item
It consumes targets {\tt D1}, ..., {\tt Dk} of type {\tt ProgramDom} and
{\tt I1}, ..., {\tt In} of type {\tt ProgramRel}.
\item
It produces targets {\tt O1}, ..., {\tt Om} of type {\tt ProgramRel}.
\item
Its \code{run()} method executes Datalog solver \code{bddbddb} to compute output
relations {\tt O1}, ..., {\tt Om} from input relations {\tt I1}, ..., {\tt In}
by solving constraints {\tt R}.
\end{itemize}

\section{DynamicAnalysis}
\label{sec:dynamic}

\section{RHSAnalysis}
\label{sec:rhs}

%\begin{itemize}
%\item
%B is the domain of basic blocks.
%\item
%C is the domain of abstract calling contexts.
%\item
%E is the domain of program points that read or write a field.
%\item
%F is the domain of fields inside a class.
%\item
%H is the domain of allocation sites.
%\item
%I is the domain of invocation statements (method calls)
%\item
%L is the domain of lock acquisition points.
%\item
%M is the domain of program methods.
%\item
%P is the domain of simple statements. %SAY MORE HERE?
%\item
%R is the domain of lock release points.
%\item
%T is the domain of types (classes and primitives).
%\item
%V is the domain of reference-typed variables.
%\item
%W is the domain of loop head statements. %SAY MORE HERE?
%\item
%Z is a domain of integers corresponding to argument positions.The size of the domain is chosen to 
%accommodate the method with the greatest number of arguments.
%\end{itemize}

%\section{Core relations}

%Chord includes code to build a very large number of relations. These standard relations are very useful in writing your own analyses.  For example, relation \texttt{sub} includes all pairs of types \texttt{(t1, t2)} where \texttt{t2} is a subclass of \texttt{t1}.  Most of these standard relations are in the package \texttt{chord.rels}.

%A Java analysis should extend JavaAnalysis or ProgramRel.  It's usually easiest if an analysis only produces one relation, in which case extending ProgramRel is the best approach.  If you extend ProgramRel, all the work of creating and saving the relation is taken care of for you; all you need to do is add tuples to the relation.Typically you do this by calling \texttt{super.add(...)} with the tuple you wish to add to the relation.  For forther convenience, you can implement one of the Visitor interfaces defined in \texttt{chord.program.visitors}.  This will force you to implement some \texttt{visit} methods that iterate over all elements in some domain, such as all methods or all invocation points.

%Sometimes, you need to create two different relations at the same time. In this case, the right approach is to extend JavaAnalysis, and explicitly create and save the relations on your own.

%Regardless of which of these approaches you take, every Java analysis must have an \texttt{@Chord} annotation.  This annotation is a set of named fields.  At a minimum, this annotation must specify the name of the analysis.  If you extend ProgramRel, the name of the analysis will be treated as the same as that of the generated relation.  For example, \texttt{@Chord(name = "F")} is a valid annotation.  The name of the Java class that generates the relation is ignored by Chord -- only the annotation matters.

%\textbf{MAYUR, IS SIGN MANDATORY?   IF NOT, WHEN CAN IT BE OMITTED?}

%There are also other annotation fields available.  You may specify the dependencies of your analysis via \texttt{consumes = \{ "xyz"\}}.

%If your analysis does not extend ProgramRel, you should also specify the relations that your analysis produces. You do this via the \texttt{produces} relation.

%Should mention the signs fields.

### 2011-04-15 main/doc/arch.tex
more documentation

@@
 
-\subsubsection*{Chord Properties}
+{\bf Chord Properties}
 
@@
 
-\subsubsection*{Java Program Representation}
+{\bf Java Program Representation}
 
@@
 
-\subsubsection*{Analysis Scope Construction}
+{\bf Analysis Scope Construction}
 
@@
 
-\subsubsection*{Writing and Running Analyses}
+{\bf Writing and Running Analyses}
 
@@
 \xlink{Habanero Concurrent Collections (CnC)}{http://habanero.rice.edu/cnc.html}
-declarative parallel programming model.
-
-Chapter \ref{chap:project} expands upon the modular architecture of analyses in
-Chord that is described above.
+declarative parallel programming model.  Chapter \ref{chap:project} expands upon
+the modular architecture of analyses in Chord.
 
@@
 merely specifying which of various provided events to instrument, and the
-transfer functions for those events.
+transfer functions for those events.  Chapters \ref{chap:writing} and
+\ref{chap:running} describe how to write and run your own analyses in Chord
+using the provided analysis templates.
 
-Chapters \ref{chap:writing} and \ref{chap:running} describe how to write and run
-your own analyses in Chord using the provided analysis templates.
-
-\subsubsection*{Dynamic Analysis}
+{\bf Dynamic Analysis}
 
@@
 
-\subsubsection*{Datalog Analysis}
+{\bf Datalog Analysis}
 

### 2011-04-15 main/doc/organization.tex
more documentation

\xname{organization}
\chapter{Organization of this Guide}
\label{chap:organization}

Chord is a program analysis platform that is both {\it stand-alone}, in that it
provides many standard analyses for users to run, and {\it extensible}, in that
it allows users to write their own analyses, possibly atop the provided
analyses.
As a result, Chord has two kinds of users: {\it end-users}, who only wish to run predefined
analyses, and {\it developers}, who additionally wish to write and run their
own analyses.

For convenience, this user guide consists of two parts: a
\xlink{guide for end-users}{enduser\_guide.html} and a
\xlink{guide for developers}{developer\_guide.html}.  Unlike end-users,
developers need to understand Chord's source code and API, and code written by
them is executed as part of a Chord run.  Hence, the guide for end-users
concerns how to run Chord, and the guide for developers concerns how to extend
Chord.

### 2011-04-15 main/doc/predefined.tex
more documentation

--- 2011-04-12 main/doc/predefined_analyses.tex
@@
-\xname{predefined_analyses}
+\xname{predefined}
 \chapter{Predefined Analyses}
-\label{chap:predefined-analyses}
+\label{chap:predefined}
 
-This chapter describes various predefined analyses in Chord.
+Chord provides many standard analyses.
+This chapter first explains how to run any such analysis and then provides
+descriptions of various predefined analyses.
+
+\section{Running an Analysis}
+\label{sec:running-predefined}
+
+Each predefined analysis in Chord has a unique name that can be used to run the
+analysis from the command-line.  The following command runs the analysis named
+\code{<ANALYSIS_NAME>} on the program specified by directory \code{<WORK_DIR>}
+(see Chapter \ref{chap:setup} for how to setup a program):
+
+\begin{framed}
+\begin{verbatim}
+ant -Dchord.work.dir=<WORK_DIR> -Dchord.run.analyses=<ANALYSIS_NAME> run
+\end{verbatim}
+\end{framed}
+
+For instance, the following command runs a basic may-alias and call-graph
+analysis (called 0CFA) provided in Chord:
+
+\begin{framed}
+\begin{verbatim}
+ant -Dchord.work.dir=<WORK_DIR> -Dchord.run.analyses=cipa-0cfa-dlog run
+\end{verbatim}
+\end{framed}
+
+This instructs Chord to run the analysis named \code{cipa-0cfa-dlog},
+which is defined in file \code{main/src/chord/analyses/alias/cipa_0cfa.dlog}.
+
+The output of the above command is of the form:
+
+\begin{framed}
+{\small
+\begin{verbatim}
+Buildfile: build.xml
+
+run:
+     [java] Chord run initiated at: Mar 13, 2011 10:31:08 PM
+     [java] ENTER: cipa-0cfa-dlog
+     [java] ENTER: T
+     [java] ENTER: RTA
+     [java] Iteration: 0
+     [java] Iteration: 1
+     [java] Iteration: 2
+     [java] LEAVE: RTA
+     [java] SAVING dom T size: 1386
+     [java] LEAVE: T
+     [java] ENTER: F
+     [java] SAVING dom F size: 4120
+     [java] LEAVE: F
+     ...
+     [java] ENTER: MputStatFldInst
+     [java] SAVING rel MputStatFldInst size: 739
+     [java] LEAVE: MputStatFldInst
+     [java] ENTER: statIM
+     [java] SAVING rel statIM size: 3359
+     [java] LEAVE: statIM
+     [java] Starting command: 'java ... chord_analyses_alias_cipa_0cfa.dlog'
+     [java] Relation VH: 541 nodes, 449.0 elements (V0,H0)
+     [java] Relation FH: 137 nodes, 8.0 elements (H0,F0)
+     [java] Relation HFH: 199 nodes, 35.0 elements (H0,F0,H1)
+     [java] Relation IM: 590 nodes, 69.0 elements (I0,M0)
+     [java] Finished command: 'java ... chord_analyses_alias_cipa_0cfa.dlog'
+     [java] LEAVE: cipa-0cfa-dlog
+     [java] Chord run completed at: Mar 13, 2011 10:31:36 PM
+     [java] Total time: 00:00:27:671 hh:mm:ss:ms
+
+BUILD SUCCESSFUL
+Total time: 28 seconds
+\end{verbatim}
+}
+\end{framed}
+
+Each analysis in Chord is written modularly, independent of other analyses,
+along with lightweight annotations specifying the inputs and outputs of the
+analysis.  Chord's runtime automatically computes producer-consumer
+relationships between analyses (e.g., determines which analysis produces as
+output a result that is needed as input by another analysis).  Before running a
+desired analysis (such as 0CFA in the above example), Chord recursively runs
+other analyses until the inputs to the desired analysis have been computed; it
+finally runs the desired analysis to produce the outputs of that analysis.
+
+The 0CFA analysis consumes and produces multiple {\it program relations}.  The
+consumed program relations include \code{MputStatFldInst} and \code{statIM},
+each of which is produced by a separate imperative analysis with the
+corresponding name, and the produced program relations include \code{VH},
+\code{FH}, \code{HFH}, and \code{IM}.  We next briefly discuss these relations.
+
+The program relations consumed by the 0CFA analysis contain basic program facts.
+For instance, \code{MputStatFldInst} is a relation containing each tuple
+($m$,$f$,$v$) such that method $m$ in the input Java program contains a
+putstatic instruction of the form ``$f$ = $v$", while \code{statIM} is a
+relation containing each tuple ($i$,$m$) such that $m$ is the target method of
+invokestatic instruction $i$.
+
+The program relations produced by the 0CFA analysis represent points-to
+information and the call graph of the input Java program as computed by the
+analysis.  Specifically, relations \code{VH}, \code{FH}, and \code{HFH}
+represent points-to information for local variables, static fields, and instance
+fields, respectively, while relation \code{IM} represents the call graph,
+namely, it contain each tuple ($i$,$m$) such that $m$ is a possible target
+method of call site $i$.
+
+Metavariables $m$, $f$, $i$, and $v$ above range over entities in so-called
+{\it program domains} \code{M}, \code{F}, \code{I}, and \code{V}, respectively.
+A program domain is a set of entities of a certain kind in the input Java
+program.  For instance, \code{M} is the domain representing the set of all
+methods in the input Java program, \code{F} is the domain representing the set
+of all fields, \code{I} is the domain representing the set of all call sites in
+methods in \code{M}, and \code{V} is the domain representing the set of all
+local variables of reference type in methods in \code{M}.  Each of these domains
+is produced by a separate Java analysis with the corresponding name.  Notice
+that the analyses producing these domains run upfront because these domains are
+consumed by the analyses that produce relations such as \code{MputStatFldInst}
+and \code{statIM}, which in turn are consumed by the desired 0CFA analysis.
+
+During execution, Chord dumps intermediate and final results to files in the
+directory specified by property \code{chord.out.dir}, whose default value is
+\code{[chord.work.dir]/chord_output/} and typically does not need to be changed
+by users.  For the above example, this directory is
+\code{<WORK_DIR>/chord_output/}.
+
+The verbosity of Chord's output above is controlled by property
+\code{chord.verbose}, whose default value is 1.  At verbosity level 0, the above
+command produces less voluminous output of the form:
+
+\begin{framed}
+{\small
+\begin{verbatim}
+Buildfile: build.xml
+
+run:
+     [java] Chord run initiated at: Mar 13, 2011 10:35:01 PM
+     [java] Chord run completed at: Mar 13, 2011 10:35:28 PM
+     [java] Total time: 00:00:26:297 hh:mm:ss:ms
+
+BUILD SUCCESSFUL
+Total time: 26 seconds
+\end{verbatim}
+}
+\end{framed}
 
@@
 
-Chord offers several choices for computing points-to and call-graph information of Java programs.
-In each of these choices, points-to and call-graph information is computed simultaneously
-(called ``on-the-fly call-graph construction" in the literature in contrast to ``ahead-of-time call-graph
-construction" in which the call graph is computed first followed by points-to information).
-On-the-fly approaches are more precise because, in a dynamically dispatching language like Java,
-as more points-to facts are discovered, more (dynamically dispatched) methods are deemed reachable,
-thereby growing the call graph; the code in these newly added methods in turn results in more
-points-to facts.
+Chord offers several choices for computing points-to and call-graph information
+of Java programs.  In each of these choices, the points-to and call-graph
+information is computed simultaneously (called ``on-the-fly call-graph
+construction" in the literature in contrast to ``ahead-of-time call-graph
+construction" in which the call graph is computed first followed by points-to
+information).  On-the-fly approaches are more precise because, in a dynamically
+dispatching language like Java, as more points-to facts are discovered, more
+(dynamically dispatched) methods are deemed reachable, thereby growing the call
+graph; the code in these newly added methods in turn results in more points-to
+facts.
 
-Flow-insensitive analysis computes a single abstract heap whereas flow-sensitive analysis computes
-per-program-point abstract heaps.  Context-insensitive analysis analyzes each method at most once (i.e. in a single abstract context),
-whereas context-sensitive analyses potentially analyze each method multiple times, in different abstract contexts.
-Thus, flow- and context-sensitive analyses are more precise but less scalable than flow- and context-insensitive
+Flow-insensitive analysis computes a single abstract heap whereas flow-sensitive
+analysis computes per-program-point abstract heaps.  Context-insensitive
+analysis analyzes each method at most once (i.e. in a single abstract context),
+whereas context-sensitive analyses potentially analyze each method multiple
+times, in different abstract contexts.  Thus, flow- and context-sensitive
+analyses are more precise but less scalable than flow- and context-insensitive
 analyses, respectively.
 
-Flow-sensitive analysis does not offer much precision over flow-insensitive analysis in
-practice, especially in the absence of strong updates and in the presence of SSA (Static Single Assignment form),
-a program representation that renders a flow-insensitive analysis almost as precise as a flow-sensitive analysis.
-Since analyses in Chord currently perform only weak updates, and since they all operate on an SSA form of the
-input Java program by default, the rest of this section focuses only on flow-insensitive analysis, which is the predominant
-kind of points-to/call-graph analysis in Chord.
+Flow-sensitive analysis does not offer much precision over flow-insensitive
+analysis in practice, especially in the absence of strong updates and in the
+presence of SSA (Static Single Assignment form), a program representation that
+renders a flow-insensitive analysis almost as precise as a flow-sensitive
+analysis.  Since analyses in Chord currently perform only weak updates, and
+since they all operate on an SSA form of the input Java program by default, the
+rest of this section focuses only on flow-insensitive analysis, which is the
+predominant kind of points-to/call-graph analysis in Chord.
 
-We describe context-insensitive analysis first because understanding the concepts behind it will help understand
-the more sophisticated context-sensitive analyses.
-We first recall some relevant program domains:
+We describe context-insensitive analysis first because understanding the
+concepts behind it will help understand the more sophisticated context-sensitive
+analyses.  We first recall some relevant program domains:
 
@@
 
-The context-insensitive points-to/call-graph analysis 
-treats each object allocation site as a separate abstract memory location; in other words, it
-can distinguish objects created at different sites but not those created at the same site.
-Additionally, it is field-sensitive, in that it distinguishes between different instance fields of the same object,
-but array-insensitive, in that it cannot distinguish between different elements of the same array;
-all array elements are modeled using a distinguished hypothetical instance field (that has index 0 in domain F).
+The context-insensitive points-to/call-graph analysis treats each object
+allocation site as a separate abstract memory location; in other words, it can
+distinguish objects created at different sites but not those created at the same
+site.  Additionally, it is field-sensitive, in that it distinguishes between
+different instance fields of the same object, but array-insensitive, in that it
+cannot distinguish between different elements of the same array; all array
+elements are modeled using a distinguished hypothetical instance field (that has
+index 0 in domain F).
 
@@
 \begin{itemize}
+\item {\it Call-graph information:}
+\begin{itemize}
 \item
-{\it Call-graph information:}
-\begin{itemize}
-\item rootM subset M contains the set of entry methods that may be reachable; this includes the program's main method
-as well as each static initializer method that may be reachable from the main
-method.
-\item reachableM subset M contains the set of methods that may be reachable from the program's main method.
-\item IM subset (I $\times$ M) contains tuples \texttt{(i,m)} such that call site \texttt{i} may call method \texttt{m}.
-%\item MM subset (M $\times$ M) contains tuples \texttt{(m1,m2)} such that method \texttt{m1} may call method \texttt{m2}.
-%\item reachableI subset I contains the set of call sites that may be reachable.
-%\item reachableT subset T contains the set of classes that may be reachable.
+rootM subset M contains the set of entry methods that may be reachable; this
+includes the program's main method as well as each static initializer method
+that may be reachable from the main method.
+\item
+reachableM subset M contains the set of methods that may be reachable from the
+program's main method.
+\item
+IM subset (I $\times$ M) contains tuples \texttt{(i,m)} such that call site
+\texttt{i} may call method \texttt{m}.
+%\item
+% MM subset (M $\times$ M) contains tuples \texttt{(m1,m2)} such that
+%method \texttt{m1} may call method \texttt{m2}.
+%\item
+%reachableI subset I contains the set of call sites that may be reachable.
+%\item
+%reachableT subset T contains the set of classes that may be reachable.
 \end{itemize}
@@
 \begin{itemize}
-\item FH subset (F $\times$ H) contains tuples \texttt{(f,h)} such that static field (i.e. global variable) \texttt{f} may point to an object allocated at site \texttt{h}.
-\item VH subset (V $\times$ H) contains tuples \texttt{(v,h)} such that local variable \texttt{v} may point to an object allocated at site \texttt{h}.
-\item HFH subset (H $\times$ F $\times$ H) contains tuples \texttt{(h1,f,h2)} such that instance field \texttt{f} of some object allocated at
-site \texttt{h1} may point to some object allocated at site \texttt{h2}.
+\item
+FH subset (F $\times$ H) contains tuples \texttt{(f,h)} such that static field
+(i.e.  global variable) \texttt{f} may point to an object allocated at site
+\texttt{h}.
+\item
+VH subset (V $\times$ H) contains tuples \texttt{(v,h)} such that local variable
+\texttt{v} may point to an object allocated at site \texttt{h}.
+\item
+HFH subset (H $\times$ F $\times$ H) contains tuples \texttt{(h1,f,h2)} such
+that instance field \texttt{f} of some object allocated at site \texttt{h1} may
+point to some object allocated at site \texttt{h2}.
 \end{itemize}
@@
 \begin{itemize}
-\item C is the domain of all abstract calling contexts and abstract objects.
-Each element in this domain is a sequence of zero or more sites, where each site
-can be a call site or an object allocation site.  A sequence may have mixed call
+\item
+C is the domain of all abstract calling contexts and abstract objects.  Each
+element in this domain is a sequence of zero or more sites, where each site can
+be a call site or an object allocation site.  A sequence may have mixed call
 sites and object allocation sites.  The most significant site (i.e. the first
@@
 the tail.  The below three relations relate a sequence with its head and tail.
-\item CC subset (C $\times$ C) contains tuples \texttt{(c1,c2)} such that context \texttt{c2} is the tail of context \texttt{c1}.
-\item CH subset (C $\times$ H) contains tuples \texttt{(c,h)} such that object allocation site \texttt{h} is the head of context \texttt{c}.
-\item CI subset (C $\times$ I) contains tuples \texttt{(c,i)} such that call site \texttt{i} is the head of context \texttt{c}.
+\item
+CC subset (C $\times$ C) contains tuples \texttt{(c1,c2)} such that context
+\texttt{c2} is the tail of context \texttt{c1}.
+\item
+CH subset (C $\times$ H) contains tuples \texttt{(c,h)} such that object
+allocation site \texttt{h} is the head of context \texttt{c}.
+\item
+CI subset (C $\times$ I) contains tuples \texttt{(c,i)} such that call site
+\texttt{i} is the head of context \texttt{c}.
 \end{itemize}
@@
 \begin{itemize}
-\item rootCM subset (C $\times$ M) contains tuples \texttt{(c,m)} such that method \texttt{m} is an entry method in context \texttt{c}.
-\item reachableCM subset (C $\times$ M) contains tuples \texttt{(c,m)} such that method \texttt{m} may be reachable in context \texttt{c}.
-\item CICM subset (C $\times$ I $\times$ C $\times$ M) contains tuples \texttt{(c1,i,c2,m)} such that call site \texttt{i} in
-context \texttt{c1} may call method \texttt{m2} in context \texttt{c2}.
+\item
+rootCM subset (C $\times$ M) contains tuples \texttt{(c,m)} such that method
+\texttt{m} is an entry method in context \texttt{c}.
+\item
+reachableCM subset (C $\times$ M) contains tuples \texttt{(c,m)} such that
+method \texttt{m} may be reachable in context \texttt{c}.
+\item
+CICM subset (C $\times$ I $\times$ C $\times$ M) contains tuples
+\texttt{(c1,i,c2,m)} such that call site \texttt{i} in context \texttt{c1} may
+call method \texttt{m2} in context \texttt{c2}.
 \end{itemize}
@@
 \begin{itemize}
-\item FC subset (F $\times$ C) contains tuples \texttt{(f,o)} such that static field (i.e. global variable) \texttt{f} may point to object \texttt{o}.
-\item CVC subset (C $\times$ V $\times$ C) contains tuples \texttt{(c,v,o)} such that local variable \texttt{v} may point to
-object \texttt{o} in context \texttt{c} of that variable's declaring
-method. Note that both \texttt{o} and \texttt{c} are elements of domain C.
-\item CFC subset (C $\times$ F $\times$ C) contains tuples \texttt{(o1,f,o2)} such that instance field \texttt{f} of object \texttt{o1} may point to object \texttt{o2}.
-\end{itemize}
+\item
+FC subset (F $\times$ C) contains tuples \texttt{(f,o)} such that static field
+(i.e. global variable) \texttt{f} may point to object \texttt{o}.
+\item
+CVC subset (C $\times$ V $\times$ C) contains tuples \texttt{(c,v,o)} such that
+local variable \texttt{v} may point to object \texttt{o} in context \texttt{c}
+of that variable's declaring method.  Note that both \texttt{o} and \texttt{c}
+are elements of domain C.
+\item
+CFC subset (C $\times$ F $\times$ C) contains tuples \texttt{(o1,f,o2)} such
+that instance field \texttt{f} of object \texttt{o1} may point to object
+\texttt{o2}.  \end{itemize}
 \end{itemize}
@@
 
-where \code{<WORK_DIR>} is a directory
-containing a file named \code{chord.properties} that defines
-properties \code{chord.main.class}, \code{chord.class.path}, and
-\code{chord.src.path} specifying the main class, the application classpath, and
-the Java source path, respectively, of the program to be analyzed.
+where \code{<WORK_DIR>} is a directory containing a file named
+\code{chord.properties} that defines properties \code{chord.main.class},
+\code{chord.class.path}, and \code{chord.src.path} specifying the main class,
+the application classpath, and the Java source path, respectively, of the
+program to be analyzed.
 
@@
 files) and then run the above command with \code{<WORK_DIR>} replaced by
-\code{examples/datarace_test/}.  Upon successful completion, the
-following files should be produced in directory
-\code{examples/datarace_test/chord_output/}:
+\code{examples/datarace_test/}.  Upon successful completion, the following files
+should be produced in directory \code{examples/datarace_test/chord_output/}:
 
@@
 
-where \code{<WORK_DIR>} is a directory
-containing a file named \code{chord.properties} that defines
-properties \code{chord.main.class}, \code{chord.class.path}, and
-\code{chord.src.path} specifying the main class, the application classpath,
-and the Java source path, respectively, of the program to be analyzed.
+where \code{<WORK_DIR>} is a directory containing a file named
+\code{chord.properties} that defines properties \code{chord.main.class},
+\code{chord.class.path}, and \code{chord.src.path} specifying the main class,
+the application classpath, and the Java source path, respectively, of the
+program to be analyzed.
 
@@
 files) and then run the above command with \code{<WORK_DIR>} replaced by
-\code{examples/deadlock_test/}.  Upon successful completion, the
-file \code{deadlocks.html} should be produced in directory
+\code{examples/deadlock_test/}.  Upon successful completion, the file
+\code{deadlocks.html} should be produced in directory
 \code{examples/deadlock_test/chord_output/}.

### 2011-04-17 main/doc/predefined.tex
more documentation

@@
 
-Each predefined analysis in Chord has a unique name that can be used to run the
-analysis from the command-line.  The following command runs the analysis named
-\code{<ANALYSIS_NAME>} on the program specified by directory \code{<WORK_DIR>}
+Each analysis in Chord is written modularly, independent of other analyses,
+along with lightweight annotations specifying the inputs and outputs of the
+analysis.  Chord's runtime automatically computes dependencies between analyses
+(e.g., determines which analysis produces as output a result that is needed as
+input by another analysis).  Before running a desired analysis, Chord
+recursively runs other analyses until the inputs to the desired analysis have
+been computed; it finally runs the desired analysis to produce the outputs of
+that analysis.
+
+Each analysis in Chord has a unique name that can be used to run the
+analysis from the command-line.  You can obtain the names of all analyses
+predefined in Chord by running the following command:
+
+\begin{framed}
+\begin{verbatim}
+ant -Dchord.out.dir=<OUT_DIR> -Dchord.print.project=true run
+\end{verbatim}
+\end{framed}
+
+This command produces files \code{targets_sortby_name.html}, \code{targets_sortby_kind.html},
+and \code{targets_sortby_producers.html} in directory \code{<OUT_DIR>}.  These
+files publish the names and descriptions of all analyses that are predefined in
+Chord, along with the inputs and outputs of each analysis.
+
+The following command runs the analysis
+named \code{<ANALYSIS_NAME>} on the program specified by directory \code{<WORK_DIR>}
 (see Chapter \ref{chap:setup} for how to setup a program):
@@
 
-Each analysis in Chord is written modularly, independent of other analyses,
-along with lightweight annotations specifying the inputs and outputs of the
-analysis.  Chord's runtime automatically computes producer-consumer
-relationships between analyses (e.g., determines which analysis produces as
-output a result that is needed as input by another analysis).  Before running a
-desired analysis (such as 0CFA in the above example), Chord recursively runs
-other analyses until the inputs to the desired analysis have been computed; it
-finally runs the desired analysis to produce the outputs of that analysis.
-
 The 0CFA analysis consumes and produces multiple {\it program relations}.  The
 consumed program relations include \code{MputStatFldInst} and \code{statIM},
-each of which is produced by a separate imperative analysis with the
-corresponding name, and the produced program relations include \code{VH},
-\code{FH}, \code{HFH}, and \code{IM}.  We next briefly discuss these relations.
+each of which is produced by a separate analysis with the corresponding name.
+The produced program relations include \code{VH}, \code{FH}, \code{HFH}, and
+\code{IM}.  We next briefly discuss these relations.
 
@@
 
-Metavariables $m$, $f$, $i$, and $v$ above range over entities in so-called
+Metavariables $m$, $f$, $i$, and $v$ above range over values in so-called
 {\it program domains} \code{M}, \code{F}, \code{I}, and \code{V}, respectively.
-A program domain is a set of entities of a certain kind in the input Java
+A program domain is a set of values of a fixed kind in the input Java
 program.  For instance, \code{M} is the domain representing the set of all
@@
 local variables of reference type in methods in \code{M}.  Each of these domains
-is produced by a separate Java analysis with the corresponding name.  Notice
+is produced by a separate analysis with the corresponding name.  Notice
 that the analyses producing these domains run upfront because these domains are
@@
 
-The verbosity of Chord's output above is controlled by property
+The verbosity of Chord's output is controlled by property
 \code{chord.verbose}, whose default value is 1.  At verbosity level 0, the above
@@
 
-To run this analysis, run the following command:
+The following command can be used to run this analysis on the program specified
+by directory \code{<WORK_DIR>} (see Chapter \ref{chap:setup} for how to setup a
+program):
 
@@
 
-where \code{<WORK_DIR>} is a directory
-containing a file named \code{chord.properties} that defines
-properties \code{chord.main.class} and \code{chord.class.path} specifying
-the main class and the application classpath, respectively, of the program
-to be analyzed.
-
 This analysis outputs the following relations:
-
-%is implemented in \code{chord/analyses/alias/cipa-0cfa.dlog}.
 
@@
 
-To run Chord's static datarace analysis, run the following command:
+The following command can be used to run Chord's static datarace analysis on the
+program specified by directory \code{<WORK_DIR>} (see Chapter \ref{chap:setup}
+for how to setup a program):
 
@@
 
-where \code{<WORK_DIR>} is a directory containing a file named
-\code{chord.properties} that defines properties \code{chord.main.class},
-\code{chord.class.path}, and \code{chord.src.path} specifying the main class,
-the application classpath, and the Java source path, respectively, of the
-program to be analyzed.
-
-Directory \code{<CHORD_MAIN_DIR>/examples/datarace_test/} provides a toy Java
-program on which one can run the datarace analysis.  First run {\tt ant} in that
-directory (in order to compile the program's {\tt .java} files to {\tt .class}
-files) and then run the above command with \code{<WORK_DIR>} replaced by
-\code{examples/datarace_test/}.  Upon successful completion, the following files
-should be produced in directory \code{examples/datarace_test/chord_output/}:
+Directory \code{examples/datarace_test/} provides a toy Java program on which to
+run the datarace analysis.  First run {\tt ant} in that directory (in order to
+compile the program's {\tt .java} files to {\tt .class} files) and then run the
+above command with \code{<WORK_DIR>} replaced by \code{examples/datarace_test/}.
+Upon successful completion, the following files should be produced in directory
+\code{examples/datarace_test/chord_output/}:
 
@@
 
-To run Chord's static deadlock analysis, run the following command:
+The following command can be used to run Chord's static deadlock analysis on the
+program specified by directory \code{<WORK_DIR>} (see Chapter \ref{chap:setup}
+for how to setup a program):
 
@@
 
-where \code{<WORK_DIR>} is a directory containing a file named
-\code{chord.properties} that defines properties \code{chord.main.class},
-\code{chord.class.path}, and \code{chord.src.path} specifying the main class,
-the application classpath, and the Java source path, respectively, of the
-program to be analyzed.
-
-Directory \code{<CHORD_MAIN_DIR>/examples/deadlock_test/} provides a toy Java
-program on which one can run the deadlock analysis.  First run {\tt ant} in that
-directory (in order to compile the program's {\tt .java} files to {\tt .class}
-files) and then run the above command with \code{<WORK_DIR>} replaced by
-\code{examples/deadlock_test/}.  Upon successful completion, the file
-\code{deadlocks.html} should be produced in directory
-\code{examples/deadlock_test/chord_output/}.
+Directory \code{examples/deadlock_test/} provides a toy Java program on which to
+run the deadlock analysis.  First run {\tt ant} in that directory (in order to
+compile the program's {\tt .java} files to {\tt .class} files) and then run the
+above command with \code{<WORK_DIR>} replaced by \code{examples/deadlock_test/}.
+Upon successful completion, the file \code{deadlocks.html} should be produced in
+directory \code{examples/deadlock_test/chord_output/}.

### 2011-05-19 main/doc/predefined.tex
updated documentation

@@
 
-For instance, the following command runs a basic may-alias and call-graph
+For instance, the following command runs a basic points-to and call-graph
 analysis (called 0CFA) provided in Chord:
@@
 
-Chord offers several choices for computing points-to and call-graph information
-of Java programs.  In each of these choices, the points-to and call-graph
-information is computed simultaneously (called ``on-the-fly call-graph
-construction" in the literature in contrast to ``ahead-of-time call-graph
-construction" in which the call graph is computed first followed by points-to
-information).  On-the-fly approaches are more precise because, in a dynamically
-dispatching language like Java, as more points-to facts are discovered, more
-(dynamically dispatched) methods are deemed reachable, thereby growing the call
-graph; the code in these newly added methods in turn results in more points-to
-facts.
+Chord provides several analyses with different precision/scalability tradeoff for
+computing points-to and call-graph information of Java programs.  All these
+analyses are defined under directory \code{src/chord/analyses/alias/}.
+
+In all these analyses, the points-to and call-graph information is computed
+simultaneously (called ``on-the-fly call-graph construction" in the literature
+in contrast to ``ahead-of-time call-graph construction" in which the call graph
+is computed first followed by points-to information).  On-the-fly approaches are
+more precise because, in a dynamically dispatching language like Java, as more
+points-to facts are discovered, more (dynamically dispatched) methods are deemed
+reachable, thereby growing the call graph; the code in these newly added methods
+in turn results in more points-to facts.
+
+Broadly, the various analyses are characterized by two aspects (1) flow
+sensitive vs. insensitive and (2) context sensitive vs. insensitive.
 
@@
 
-The context-insensitive points-to/call-graph analysis treats each object
-allocation site as a separate abstract memory location; in other words, it can
-distinguish objects created at different sites but not those created at the same
-site.  Additionally, it is field-sensitive, in that it distinguishes between
-different instance fields of the same object, but array-insensitive, in that it
-cannot distinguish between different elements of the same array; all array
-elements are modeled using a distinguished hypothetical instance field (that has
-index 0 in domain F).
+The context-insensitive points-to/call-graph analysis derives its name from the
+fact that it analyzes each method in a single abstract context.
+Additionally, the analysis is:
+\begin{itemize}
+\item
+{\it object-insensitive}: it cannot distinguish between
+different objects created at the same object allocation site; the reason is that
+it treats each object allocation site as a separate abstract memory location.
+\item
+{\it field-sensitive}: it distinguishes between
+different instance fields of the same object.
+\item
+{\it array-insensitive}: it cannot distinguish between
+different elements of the same array; all array elements are modeled using a
+distinguished hypothetical instance field (that has index 0 in domain F).
+\end{itemize}
 
@@
 rootM subset M contains the set of entry methods that may be reachable; this
-includes the program's main method as well as each static initializer method
+includes the program's main method as well as each class initializer method
 that may be reachable from the main method.
@@
 
-In a context-sensitive analysis, there is no longer just one abstract memory
-location per object allocation site.  Rather, the set of objects a reference can
-point to depends on the \textit{context} in which the method containing the
-reference is called.  Whereas a context-insensitive analysis talks about the
-domain of methods (M) and the domain of allocation sites (H), a
-context-sensitive analysis talks about the domain of abstract contexts, labeled
-C.  Elements of domain C contain both abstract calling contexts and abstract
-objects. (These are merged for reasons described below.)
+In a context-sensitive analysis, each method is analyzed in potentially multiple
+contexts, in contrast to a context-insensitive analysis, in which each
+method is always analyzed in a single context.  Additionally, a
+context-sensitive analysis can be {\it object-sensitive} in that it can distinguish
+between different objects created at the same object allocation site, whereas
+a context-insensitive analysis is always object-insensitive.
 
-Chord has several context-sensitive analyses, but they all expose the same
-relations, which are described below:
+Different kinds of context- and object-sensitivity have been proposed in the
+literature.  Chord's context-sensitive points-to/call-graph analysis is highly
+parametric in the kind of context- and object-sensitivity.  It allows:
+\begin{itemize}
+\item
+each method to be analyzed using a different kind of context sensitivity,
+namely, one of context insensitivity, k-CFA, k-object-sensitivity, and
+copy-context-sensitivity;
+\item
+each local variable to be analyzed context sensitively or insensitively; and
+\item
+a different 'k' value to be used for each object allocation site and method
+call site.
+\end{itemize}
+
+The analysis can be called multiple times and in each invocation it can
+incorporate feedback from a client to adjust the precision of the points-to
+information and call graph computed subsequently by the points-to/call-graph
+analyses.  Clients can indicate in each invocation:
+
+\begin{itemize}
+\item
+Which methods must be analyzed context sensitively (in addition to those already being
+analyzed context sensitively in the previous invocation of this analysis) and using
+what kind of context sensitivity; the remaining methods will be analyzed context
+insensitively (that is, in the lone 'epsilon' context).
+\item
+Which local variables of reference type must be analyzed context sensitively (in
+addition to those already being analyzed context sensitively in the previous
+invocation of this analysis); the remaining ones will be analyzed context insensitively
+(that is, their points-to information will be tracked in the lone 'epsilon' context).
+\item
+The object allocation sites and method call sites whose 'k' values must be incremented
+(over those used in the previous invocation of this analysis).
+\end{itemize}
+
+The following command can be used to run this analysis on the program specified
+by directory \code{<WORK_DIR>} (see Chapter \ref{chap:setup} for how to setup a
+program):
+
+\begin{framed}
+\begin{verbatim}
+ant -Dchord.work.dir=<WORK_DIR> -Dchord.run.analyses=cspa-hybrid-dlog run
+\end{verbatim}
+\end{framed}
+
+System properties recognized by this analysis are:
+\begin{itemize}
+\item
+\code{chord.inst.ctxt.kind}: the kind of context sensitivity to use for each
+instance method (and all its locals).  It may be 'ci' (context insensitive), 'cs'
+(k-CFA), or 'co' (k-object-sensitive).  It is 'ci' by default.
+\item
+\code{chord.stat.ctxt.kind}: the kind of context sensitivity to use for each
+static method (and all its locals).  It may be 'ci' (context insensitive), 'cs'
+(k-CFA), or 'co' (copy-context-sensitive).  It is 'ci' by default.
+\item
+\code{chord.ctxt.kind}: the kind of context sensitivity to use for each method
+(and all its locals).  It may be one of 'ci', 'cs', or 'co', and serves as
+shorthand for properties \code{chord.inst.ctxt.kind} and \code{chord.stat.ctxt.kind}.
+\item
+\code{chord.kobj.k} and \code{chord.kcfa.k}: the 'k' value to use for each object
+allocation site and each method call site, respectively.
+It is 1 by default.
+\end{itemize}
+
+The above command runs the most general form of context-sensitive
+points-to/call-graph analysis,
+allowing different kinds of context- and object-sensitivity to be mixed in the analysis of
+the same program, but this flexibility comes at the cost of efficiency.
+Hence, Chord provides three specialized forms of this analysis that more
+efficiently handle certain common instantiations of this analysis that do not
+arbitrarily mix different kinds of context- and object-sensitivity.  We next describe these
+three specialized analyses.
+
+If \code{chord.ctxt.kind} is 'ci' (the default value),
+the above command essentially performs a purely context-insensitive
+analysis; in this case, it is preferable to run the following more efficient
+implementation of that analysis:
+
+\begin{framed}
+\begin{verbatim}
+ant -Dchord.work.dir=<WORK_DIR> -Dchord.run.analyses=cspa-0cfa-dlog run
+\end{verbatim}
+\end{framed}
+
+If \code{chord.ctxt.kind} is 'cs' and \code{chord.kcfa.k} is K,
+the above command essentially performs a purely K-CFA analysis; in this case, it
+is preferable to run the following more efficient implementation of that
+analysis:
+
+\begin{framed}
+\begin{verbatim}
+ant -Dchord.work.dir=<WORK_DIR> -Dchord.ctxt.kind=cs -Dchord.kcfa.k=K -Dchord.run.analyses=cspa-kcfa-dlog run
+\end{verbatim}
+\end{framed}
+
+If \code{chord.ctxt.kind} is 'co' and \code{chord.kobj.k} is K,
+the above command essentially performs a purely K-object-sensitive analysis; in
+this case, it is preferable to run the following more efficient implementation
+of that analysis:
+
+\begin{framed}
+\begin{verbatim}
+ant -Dchord.work.dir=<WORK_DIR> -Dchord.ctxt.kind=co -Dchord.kobj.k=K -Dchord.run.analyses=cspa-kobj-dlog run
+\end{verbatim}
+\end{framed}
+
+Each of the above four commands outputs the following relations:
 
@@
 \item
-C is the domain of all abstract calling contexts and abstract objects.  Each
+C is the domain of all abstract contexts and abstract objects.  Each
 element in this domain is a sequence of zero or more sites, where each site can
@@
 
-{\bf Under Construction: explain object-sensitive vs context-sensitive, plus how
-to invoke.}
+\subsection{Call Graph Variants}
+
+Chord offers two variants of each of the context-insensitive and
+context-sensitive baseline call graphs presented above, called {\it abbreviated
+thread-insensitive} and {\it abbreviated thread-sensitive} call graphs.  We
+explain these two variants for the context-insensitive case; they are analogous
+for the context-sensitive case.  Consider the following example Java program:
+
+\begin{framed}
+\begin{verbatim}
+public class T extends java.lang.Thread {
+    static {
+i1:     m1();
+    }
+    public static void m1() { }
+    public static void m2() { }
+    public static void m3() { }
+    public void run() {
+i2:     m2();
+    }
+    public static void main() {
+i3:     m3();
+        T t = ...;
+i4:     t.start();
+    }
+}
+\end{verbatim}
+\end{framed}
+
+The baseline call graph for this program is 
+rootM = \{ main, \code{<clinit>} \}, 
+reachableM = \{ main, \code{<clinit>}, m1, m2, m3, start, run \}, and
+IM = \{ (i1, m1), (i2, m2), (i3, m3), (i4, start), (i5, run) \}, where methods
+main, m1, m2, m3, and run are those declared in class C, method \code{<clinit>}
+is the class initializer method of class C, and the
+\code{java.lang.Thread.start()} method contains a call site labeled i5 that calls
+\code{this.run()}.
+
+An {\it abbreviated thread-insensitive} call graph excludes from the baseline call
+graph all methods that are unreachable from the main method.  In other words, it
+excludes all class initializer methods as well as methods that are reachable
+only from such methods.
+For the above example, this call graph is
+rootM = \{ main \},
+reachableM = \{ main, m2, m3, start, run \}, and
+IM = \{ (i2, m2), (i3, m3), (i4, start), (i5, run) \}.  Notice that methods
+\code{<clinit>} and m1 are omitted from reachableM, and call edge (i1, m1) is
+omitted from IM, since m1 is unreachable from the main method.
+
+An {\it abbreviated thread-sensitive} call graph excludes from the baseline call
+graph all methods that are unreachable from the main method, like the
+abbreviated thread-insensitive call graph, and additionally treats the
+\code{java.lang.Thread.start()} method as a root method and does not include
+call edges that target that method.
+For the above example, this call graph is
+rootM = \{ main, start \},
+reachableM = \{ main, m2, m3, start, run \}, and
+IM = \{ (i2, m2), (i3, m3), (i5, run) \}.  Notice that the start method is
+added to rootM and call edge (i4, start) is omitted from IM.
+
+The above two call graph variants are provided to simplify expressing
+certain forms of reachability:
+\begin{itemize}
+\item
+An {\it abbreviated} call graph allows to focus just on reachability
+from the main method, as opposed to from class initializer methods as well;
+this is used, for instance, in Chord's static datarace analysis to
+suppress reporting races on statements reachable only from class initializer
+methods, which are typically false races in practice.
+\item
+A {\it thread-sensitive} call graph additionally allows to focus just on
+``intra-thread'' reachability; this is used, for instance, in Chord's static datarace
+analysis to prevent reporting false races (t1,e1,t2,e2)
+where statement e1 (or e2 resp.) is unreachable from abstract thread
+t1 (or t2 resp.).
+\end{itemize}
 

### 2011-05-19 main/doc/relation.tex
updated documentation

--- 2011-04-15 main/doc/datalog.tex
@@
-\xname{datalog}
-\chapter{Datalog Analysis}
-\label{chap:datalog}
+\xname{relation}
+\chapter{Program Relation}
+\label{chap:relation}
 
-A common way to rapidly prototype an analysis in Chord is using a declarative
-logic-programming language called Datalog.  This chapter describes all aspects
-of Datalog analyses in Chord.  Section \ref{sec:writing-datalog} explains how to
-write and run a Datalog analysis.  Section \ref{sec:tuning-datalog} explains how
-to tune its performance.  Finally, Section \ref{sec:bdd} explains the
+Finally, Section \ref{sec:bdd} explains the
 representation of BDDs (Binary Decision Diagrams) which are a data structure
@@
 
-\section{Writing a Datalog Analysis}
-\label{sec:writing-datalog}
+\section{Constructing a Relation}
 
-A Datalog analysis declares a bunch of input/output program relations, each over
-one or more program domains, and provides a bunch of rules (constraints)
-specifying how to compute the output relations from the input relations.  It can
-be defined in any file with suffix {\tt .dlog} in any directory in the path
-specified by property \code{chord.dlog.analysis.path}.  An example Datalog
-analysis is shown below:
+1. Using ProgramRel
+2. Produced by analysis
+3. Manual
 
-\texonly{\newpage}
+\section{Operations on a Relation}
 
-\begin{framed}
-{\small
-\begin{verbatim}
-    # name=datarace-dlog
-
-    # Program domains
-    .include "E.dom"
-    .include "F.dom"
-    .include "T.dom"
-
-    # BDD variable order
-    .bddvarorder E0xE1_T0_T1_F0
-
-    # Input/intermediate/output program relations
-    field(e:E0,f:F0) input
-    write(e:E0) input
-    reach(t:T0,e:E0) input
-    alias(e1:E0,e2:E1) input
-    escape(e:E0) input
-    unguarded(t1:T0,e1:E0,t2:T1,e2:E1) input
-    hasWrite(e1:E0,e2:E1)
-    candidate(e1:E0,e2:E1) 
-    datarace(t1:T0,e1:E0, t2:T1,e2:E1) output
-
-    # Analysis constraints
-    hasWrite(e1,e2) :- write(e1).
-    hasWrite(e1,e2) :- write(e2).
-    candidate(e1,e2) :- field(e1,f), field(e2,f), hasWrite(e1,e2), e1 <= e2.
-    datarace(t1,e1,t2,e2) :- candidate(e1,e2), reach(t1,e1), reach(t2,e2), \
-        alias(e1,e2), escape(e1), escape(e2), unguarded(t1,e1,t2,e2).
-\end{verbatim}
-}
-\end{framed}
-
-Any line that begins with a ``{\tt \#}" is regarded a comment, except a
-line of the form ``\code{# name=<ANALYSIS_NAME>}'', which specifies the name
-\code{<ANALYSIS_NAME>} of the Datalog analysis.
-Each Datalog analysis is expected to have exactly one such line.
-The above Datalog analysis is named {\tt datarace-dlog}.
-
-The ``\code{.include "<DOM_NAME>.dom"}'' lines specify each domain named
-\code{<DOM_NAME>} that is needed by the Datalog analysis, i.e., each domain over which
-any relation that is input/output by the Datalog analysis is defined. 
-The declaration of each such relation specifies the relation's schema:
-all the domains over which the relation is defined.
-If the same domain appears multiple times in a relation's schema then
-contiguous integers starting from 0 must be used to distinguish them; for instance,
-in the above example, {\tt candidate} is a binary relation, both of whose
-attributes have domain {\tt E}, and they are distinguished as {\tt E0} and {\tt E1}.
-
-Each relation is represented symbolically (as opposed to explicitly)
-using a graph-based data structure called a Binary Decision Diagram (BDD for short).
-Each domain containing N elements is assigned log2(N) BDD variables.
-The size of a BDD and the efficiency of operations on it depends heavily
-on the order of these BDD variables.
-The ``\code{.bddvarorder <BDD_VAR_ORDER>}'' line in the Datalog analysis enables the Datalog
-analysis writer to specify this order.
-It must list all domains along with their numerical suffixes, separated
-by `{\tt \_}' or `{\tt x}'.
-Using a `{\tt \_}' between two domains, such as {\tt T0\_T1}, means that the BDD variables assigned
-to domain {\tt T0} precede those assigned to domain {\tt T1} in the BDD variable
-order for this Datalog analysis.
-Using a `{\tt x}' between two domains, such as {\tt E0xE1}, means that the
-BDD variables assigned to domains {\tt E0} and {\tt E1}
-will be interleaved in the BDD variable order for this Datalog analysis.
-See Section \ref{sec:bdd} for more details on BDD representations.
-
-Each Datalog analysis rule is a Horn clause of the form
-``{\tt R(t) :- R1(t1), ..., Rn(tn)}"
-meaning that if relations {\tt R1}, ..., {\tt Rn} contain tuples {\tt t1}, ..., {\tt tn}
-respectively, then relation {\tt R} contains tuple {\tt t}.
-A backslash may be used at the end of a line to break long rules for readability.
-The Datalog analysis solver bddbddb used in Chord does not apply any
-sophisticated optimizations to simplify the rules; besides the BDD variable order,
-the manner in which these rules are expressed heavily affects the performance of
-the solver.  For instance, an important manual optimization involves breaking down
-long rules into multiple shorter rules communicating via intermediate relations.
-See Section \ref{sec:tuning-datalog} for hints on tuning the performance
-of Datalog analyses.
-
-Running a Datalog analysis is no different from running any other kind of analysis
-in Chord.  See Chapter \ref{chap:running} for how to run an analysis.
-
-\section{Tuning Performance}
-\label{sec:tuning-datalog}
-
-There are several tricks analysis writers can try to improve the
-performance of bddbddb, the Datalog solver used by Chord, often
-by several orders of magnitude.
-Try these tricks by running the following command:
-
-\begin{framed}
-\begin{verbatim}
-ant -Ddlog.file=<FILE> -Dwork.dir=<DIR> solve
-\end{verbatim}
-\end{framed}
-
-where {\tt <FILE>} is the file defining the Datalog analysis
-to be tuned, and {\tt <DIR>} is the directory containing the
-program domains ({\tt *.dom} files) and program relations ({\tt *.bdd} files)
-consumed by the analysis (this is by default the
-\code{chord_output/bddbddb/} directory generated
-by a previous run of Chord.
-
-\begin{enumerate}
-
-\item
-
-Set properties \verb+noisy=yes+, \verb+tracesolve=yes+, and \verb+fulltracesolve=yes+
-on the above command line and observe which rule gets ``stuck" (i.e., takes several seconds to solve).
-\verb+fulltracesolve+ is seldom useful, but \verb+noisy+ and \verb+tracesolve+ are
-often very useful.  Once you identify the rule that is getting stuck, it
-will also tell you which relations and which domains used in that rule,
-and which operation on them, is taking a long time to solve.  Then try
-to fix the problem with that rule by doing either or both of the following:
-\begin{itemize}
-\item
-Break down the rule into multiple rules by creating intermediate relations (the more
-relations you have on the RHS of a rule the slower it generally takes to solve
-that rule).
-\item
-Change the relative order of the domains of those
-relations in the BDD variable order
-(note that you can use either `\_' or `x' between a pair of domains).
-\end{itemize}
-
-\item
-
-Once you have ensured that none of the rules is getting ``stuck",
-you will notice that some rules are applied too many times, and so
-although each application of the rule itself isn't taking too much
-time, the cumulative time for the rule is too much.  After finishing
-solving a Datalog analysis, bddbddb prints how long each rule took to
-solve (both in terms of the number of times it was applied and the
-cumulative time it took).  It sorts the rules in the order of the
-cumulative time.  You need to focus on the rules that took the most
-time to solve (they will be at the bottom of the list).  Assuming you
-removed the problem of rules getting ``stuck", the rules will roughly
-be in the order of the number of times they were applied.  Here is an
-example:
-
-\begin{framed}
-\begin{verbatim}
-OUT> Rule VH(u:V0,h:H0) :- VV(u:V0,v:V1), VH(v:V1,h:H0), VHfilter
-(u:V0,h:H0).
-OUT>    Updates: 2871
-OUT>    Time: 6798 ms
-OUT>    Longest Iteration: 0 (0 ms)
-OUT> Rule IM(i:I0,m:M0) :- reachableI(i:I0), specIMV(i:I0,m:M0,v:V0), VH(v:V0,_:H0).
-OUT>    Updates: 5031
-OUT>    Time: 6972 ms
-OUT>    Longest Iteration: 0 (0 ms)
-\end{verbatim}
-\end{framed}
-
-Notice that the second rule was applied 5031 times whereas the first
-was applied 2871 times.  More importantly, the second rule took 6972
-milliseconds in all, compared to 6798 for the first rule.  Hence, you
-should focus on the second rule first, and try to speed it up.  This
-means that you should focus only on relations IM, reachableI, specIMV,
-and VH, and the domains I0, M0, V0, and H0.  Any changes you make that
-do not affect these relations and domains are unlikely to make your
-solving faster.  In general, look at the last few rules, not just the
-last one, and try to identify the ``sub-analysis" of the Datalog analysis
-that seems problematic, and then focus on speeding up just that sub-
-analysis.
-
-\item
-
-You can add the \verb+.split+ keyword at the end of certain rules as a
-hint to bddbddb to decompose those rules into simpler ones that can be
-solved faster.  You can also set property \verb+split_all_rules=yes+ as shorthand
-for splitting all rules without adding the \verb+.split+ keyword to any of
-them, though I seldom find splitting all rules helpful.
-
-\item
-
-You can try to decompose a single Datalog analysis file into two separate Datalog analysis
-files.  Of course, you cannot separate mutually-recursive rules into two
-different analyses, but if you unnecessarily club
-together rules that could have gone into different analyses, then they
-can put conflicting demands on bddbddb (e.g., on the BDD variable order).
-So if rule 2 uses the result of rule 1 and rule 1 does not use the result of
-rule 2, then put rule 1 and rule 2 in separate Datalog analyses.
-
-\item
-
-Observe the sizes of the BDDs representing the relations that are
-input and output.  bddbddb prints both the number of tuples in each
-relation and the number of nodes in the BDD.  Try changing the
-BDD variable order for the domains of the relation, and observe how the
-number of nodes in the BDD for that relation change.  You will
-notice that some orders perform remarkably better than others.  Then
-note down these orders as invariants that you will not violate as
-you tweak other things.
-
-\item
-
-The relative order of values *within* domains (e.g.,
-in domains named \verb+M+, \verb+H+, \verb+C+, etc. in Chord) affects the
-performance of bddbddb, but
-I've never tried changing this and studying its effect.  It might be
-worth trying.  For instance, John Whaley's PLDI'04 paper describes a
-specific way in which he numbers contexts (in domain \verb+C+) and that it was
-fundamental to the speedup of his ``infinity"-CFA points-to analysis.
-
-\item
-
-Finally, it is worth emphasizing that BDDs are not magic.
-If your algorithm itself is fundamentally hard to scale, then BDDs are
-unlikely to help you a whole lot.  Secondly, many things are awkward to
-encode as integers (e.g., the abstract contexts in domain \verb+C+ 
-in Chord) or as Datalog rules.
-For instance, I've noticed that summary-based context-sensitive program
-analyses are hard to express in Datalog.  The may-happen-in-parallel
-analysis provided in Chord shows a relatively simple kind of summary-based
-analysis that uses the Reps-Horwitz-Sagiv tabulation algorithm.  But this
-is as far as I could get---more complicated summary-based algorithms are
-best written in Java itself instead of Datalog.
-\end{enumerate}
-
-\section{BDD Representation}
+\section{BDD Representation Internals}
 \label{sec:bdd}
 
-Each domain containing N elements is assigned log2(N) BDD variables in the underlying BDD factory with contiguous IDs.
+Each domain containing N elements is assigned log2(N) BDD variables in the
+underlying BDD factory with contiguous IDs.
 For instance,
-domain {\tt F0} containing [128..256) elements will be assigned 8 variables with IDs (say) 63,64,65,66,67,68,69,70 and
-domain {\tt Z0} containing [8..16) elements will be assigned 4 variables with IDs (say) 105,106,107,108.
+domain {\tt F0} containing [128..256) elements will be assigned 8 variables with
+IDs (say) 63,64,65,66,67,68,69,70 and
+domain {\tt Z0} containing [8..16) elements will be assigned 4 variables with
+IDs (say) 105,106,107,108.
 
-If two domains are uninterleaved in the declared domain order in a Datalog program (i.e., `{\tt \_}' is used instead of `{\tt x}' between them),
-then the BDD variables assigned to those domains are ordered in reverse order in the underlying BDD factory.
-For instance, the BDD variable order corresponding to the declared domain order {\tt F0\_Z0} is (in level2var format)
+If two domains are uninterleaved in the declared domain order in a Datalog
+program (i.e., `{\tt \_}' is used instead of `{\tt x}' between them),
+then the BDD variables assigned to those domains are ordered in reverse order in
+the underlying BDD factory.
+For instance, the BDD variable order corresponding to the declared domain order
+{\tt F0\_Z0} is (in level2var format)
 ``70,69,68,67,66,65,64,63,108,107,106,105".
 
-If two domains are interleaved in the declared domain order in a Datalog program (i.e., `{\tt x}' is used instead of `{\tt \_}' between them),
-then the BDD variables assigned to those domains are still ordered in reverse order of IDs in the underlying BDD factory,
+If two domains are interleaved in the declared domain order in a Datalog program
+(i.e., `{\tt x}' is used instead of `{\tt \_}' between them),
+then the BDD variables assigned to those domains are still ordered in reverse
+order of IDs in the underlying BDD factory,
 but they are also interleaved.
-For instance, the BDD variable order corresponding to the declared domain order {\tt F0xZ0} is (in level2var format)
+For instance, the BDD variable order corresponding to the declared domain order
+{\tt F0xZ0} is (in level2var format)
 ``70,108,69,107,68,106,67,105,66,65,64,63".
 
-Each BDD variable is at a unique ``level" which is its 0-based position in the BDD variable order in the underlying BDD factory.
+Each BDD variable is at a unique ``level" which is its 0-based position in the
+BDD variable order in the underlying BDD factory.
 
-We will next illustrate the format of a BDD stored on disk (in a .bdd file) using the following example:
+We will next illustrate the format of a BDD stored on disk (in a .bdd file)
+using the following example:
 
@@
 28489 134
-39 36 33 30 27 24 21 18 15 12 9 6 3 0 81 79 77 75 73 71 69 67 65 63 61 59 57 55  \
-   53 51 82 80 78 76 74 72 70 68 66 64 62 60 58 56 54 52 37 34 31 28 25 22 19 16 \
-   13 10 7 4 1 117 116 115 114 113 112 111 110 109 108 107 106 50 49 48 47 46 45 \
-   44 43 42 41 40 105 104 103 102 101 100 99 98 97 96 95 94 93 92 91 90 89 88 87 \
-   86 85 84 83 133 132 131 130 129 128 127 126 125 124 123 122 121 120 119 118   \
+39 36 33 30 27 24 21 18 15 12 9 6 3 0 81 79 77 75 73 71 69 67 65 63 61 59 57 55
+\
+   53 51 82 80 78 76 74 72 70 68 66 64 62 60 58 56 54 52 37 34 31 28 25 22 19 16
+\
+   13 10 7 4 1 117 116 115 114 113 112 111 110 109 108 107 106 50 49 48 47 46 45
+\
+   44 43 42 41 40 105 104 103 102 101 100 99 98 97 96 95 94 93 92 91 90 89 88 87
+\
+   86 85 84 83 133 132 131 130 129 128 127 126 125 124 123 122 121 120 119 118
+\
    38 35 32 29 26 23 20 17 14 11 8 5 2
@@
 
-The first comment line indicates the domains in the relation (in the above case, {\tt V0} and {\tt H0},
+The first comment line indicates the domains in the relation (in the above case,
+{\tt V0} and {\tt H0},
 represented using 16 and 12 unique BDD variables, respectively).
 
-If there are $N$ domains, there are $N$ following comment lines, each specifying the
+If there are $N$ domains, there are $N$ following comment lines, each specifying
+the
 BDD variables assigned to the corresponding domain.
 
-The following line has two numbers: the number of nodes in the represented BDD (28489 in this case) and the number of variables
-in the BDD factory from which the BDD was dumped to disk.  Note that the number of variables (134 in this case) is not
-necessarily the number of variables in the represented BDD (16+12=28 in this case) though it is guaranteed to be
+The following line has two numbers: the number of nodes in the represented BDD
+(28489 in this case) and the number of variables
+in the BDD factory from which the BDD was dumped to disk.  Note that the number
+of variables (134 in this case) is not
+necessarily the number of variables in the represented BDD (16+12=28 in this
+case) though it is guaranteed to be
 greater than or equal to it.
 
-The following line specifies the BDD variable order in var2level format.  In this case, the specified order subsumes
-{\tt V0\_H0} (notice that levels ``81 79 77 75 73 71 69 67 65 63 61 59 57 55 53 51", which are at positions ``14 15 ... 28 29" 
-in the specified order are lower than levels ``105 104 103 102 101 100 99 98 97 96 95 94" which are at positions
+The following line specifies the BDD variable order in var2level format.  In
+this case, the specified order subsumes
+{\tt V0\_H0} (notice that levels ``81 79 77 75 73 71 69 67 65 63 61 59 57 55 53
+51", which are at positions ``14 15 ... 28 29"
+in the specified order are lower than levels ``105 104 103 102 101 100 99 98 97
+96 95 94" which are at positions
 ``82 83 .. 92 93").
 
-Each of the following lines specifies a unique node in the represented BDD; it has format ``X V L H" where:
+Each of the following lines specifies a unique node in the represented BDD; it
+has format ``X V L H" where:
 \begin{itemize}
 \item X is the ID of the BDD node
-\item V is the ID of the bdd variable labeling that node (unless it is 0 or 1, in which case it represents a leaf node)
+\item V is the ID of the bdd variable labeling that node (unless it is 0 or 1,
+in which case it represents a leaf node)
 \item L is the ID of the BDD node's low child
@@
 
-The order of these lines specifying BDD nodes is such that the lines specifying the BDD nodes in the L and H entries
-of each BDD node are guaranteed to occur before the line specifying that BDD node (for instance, the L entry 287 on
-the second line and the R entry 349123 on the third line are IDs of BDD nodes specified on the first and second lines,
+The order of these lines specifying BDD nodes is such that the lines specifying
+the BDD nodes in the L and H entries
+of each BDD node are guaranteed to occur before the line specifying that BDD
+node (for instance, the L entry 287 on
+the second line and the R entry 349123 on the third line are IDs of BDD nodes
+specified on the first and second lines,
 respectively).
@@
 that represents all the variables used in {\tt b}.
-The support BDD {\tt r} is a linear tree each of whose nodes contains a separate variable,
+The support BDD {\tt r} is a linear tree each of whose nodes contains a separate
+variable,
 the low branch is 0, and high branch is the node representing the next variable.

### 2011-05-19 main/doc/user_guide.tex
updated documentation

@@
 
+\newcommand\includeimage[2]{\texonly{\includegraphics[scale=#1]{#2}}\htmlonly{{\htmlimg{#2}}}}
+
 \newcommand\bh{{\bf h}}
@@
 
-\setcounter{htmldepth}{2}
-\setcounter{htmlautomenu}{3}
+\newcommand\bsdlicense{\xlink{New BSD License}{http://www.opensource.org/licenses/bsd-license.php}}
+\newcommand\chordweb{\url{http://jchord.googlecode.com/}}
+\newcommand\buddy{\xlink{BuDDy}{http://buddy.sourceforge.net/}}
+\newcommand\bddbddb{\xlink{bddbddb}{http://bddbddb.sourceforge.net/}}
+\newcommand\ibmjvm{\xlink{IBM J9}{http://www.ibm.com/developerworks/java/jdk/}}
+\newcommand\sunjvm{\xlink{Oracle HotSpot}{http://www.oracle.com/technetwork/java/javase/downloads/index.html}}
+\newcommand\chordbinfile{\xlink{chord-bin-2.0.tar.gz}{http://jchord.googlecode.com/files/chord-bin-2.0.tar.gz}}
+\newcommand\chordsrcfile{\xlink{chord-src-2.0.tar.gz}{http://jchord.googlecode.com/files/chord-src-2.0.tar.gz}}
+\newcommand\chordlibfile{\xlink{chord-libsrc-2.0.tar.gz}{http://jchord.googlecode.com/files/chord-libsrc-2.0.tar.gz}}
+\newcommand\ant{\xlink{Apache Ant}{http://ant.apache.org/}}
+\newcommand\javassist{\xlink{Javassist}{http://www.javassist.org/}}
+\newcommand\joeq{\xlink{Joeq}{http://joeq.sourceforge.net/}}
+
+\setcounter{htmldepth}{3}
+\setcounter{htmlautomenu}{4}
 
@@
 
-\input{whatis_chord}
-\input{getting_started}
-\input{example}
-\input{architecture}
+%\texonly{\tableofcontents}
+
+\xname{preface}
+\part{Preface}
+
+\input{organization}
+\input{acks}
+
+\xname{enduser_guide}
+\part{Guide for End-Users}
+
+\input{whatis}
+\input{start}
 \input{properties}
-\input{predefined_analyses}
-\input{analyses}
-\input{dynamic_analysis}
-\input{datalog_analysis}
-\input{computing_scope}
-\input{program_representation}
-\input{acknowledgments}
+\input{setup}
+\input{scope}
+\input{predefined}
+
+\xname{developer_guide}
+\part{Guide for Developers}
+
+\input{arch}
+\input{program}
+\input{project}
+\input{writing}
+\input{running}
+\input{dynamic}
+\input{datalog}
 
